### 2020-10-09 report/report.tex
Scaffold LaTeX report

\documentclass[]{article}

%opening
\title{Detecting Causal Structure in Time Series Data}
\author{Lawrence Tray \\ Ioannis Kontoyiannis}

%packages
\usepackage[margin=0.5in]{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}

%package setup
\graphicspath{{./img/}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

%custom commands
\newcommand{\dft}{\mathcal{F}}
\newcommand{\idft}{\mathcal{F}^{-1}}
\newcommand{\Xcal}{\mathcal{X}}
\newcommand{\cmplx}{\mathbb{C}}
\newcommand{\figwidth}{0.55\linewidth}

\begin{document}

\maketitle

\begin{abstract}
\end{abstract}

\section{Introduction}

\end{document}

### 2020-10-22 report/report.tex
Continue mean proof

@@
 \usepackage{subcaption}
+\usepackage{mathtools}
 
@@
 \newcommand{\cmplx}{\mathbb{C}}
+\newcommand{\E}{\mathbb{E}}
+\newcommand{\Gaussian}{\mathcal{N}}
+\newcommand{\lik}{\mathcal{L}}
+\newcommand{\kl}{\mathcal{D}}
 \newcommand{\figwidth}{0.55\linewidth}
@@
 \section{Introduction}
+\section{Hypothesis Testing}
+\subsection{Single Sample against known mean}
+We start with the simple case of determining whether or not a coin is fair.
+Each coin flip can be represented by a random variable with a Bernoulli distribution $X_i \sim Bern(p)$.
+Each coin can result in either a Tails or a Heads denoted by $X_i \in \{0, 1\}$.
+We toss the coin $n$ times leading us to a set $\left\{ X_i \right\}_{i=1}^{n}$.
+We wish to test the null hypothesis $p=p_0$ against the alternative.
+We shall keep the $p_0$ notation for generality, though for a fair coin we require $p_0=1/2$.
+%
+\begin{align*}
+H_0:& \quad p = p_0 \\
+H_1:& \quad p \neq p_0
+\end{align*}
+%
+We denote the number of heads with the random variable $K \coloneqq \sum_{i=1}^{n} X_i \sim Bern(p, n)$.
+For a particular experiment we observe $K=k$.
+We employ a standard likelihood ratio test.
+The test statistic $t_n$ is calculated as the log-likelihood ratio of observing $K=k$ under $H_1$ and $H_0$.
+%
+\begin{equation}
+t_n \coloneqq \log \frac{\lik(H_1)}{\lik(H_0)}
+= \log \frac{\max_p P(K=k|p)}{P(K=k|p=p_0)}
+\end{equation} 
+%
+We note that since $K$ is distributed as a binomial, $P(K=k|p)=\binom{n}{k}p^k(1-p)^{n-k}$.
+If we can vary $p$, this probability is maximised for $p=\hat{p} \coloneqq k/n$.
+Therefore, the test statistic is given by.
+%
+\begin{equation}
+t_n = \log \frac{\binom{n}{k} \hat{p}^k (1 - \hat{p})^{n-k}}{\binom{n}{k} p_0^k (1 - p_0)^{n-k}}
+=  \log \frac{ \hat{p}^k (1 - \hat{p})^{n-k}}{ p_0^k (1 - p_0)^{n-k}}
+\end{equation}
+%
+The combinatoric term $\binom{n}{k}$ cancels out.
+This implies that the order in which the heads land does not matter when determining the fairness of the coin.
+We can work the above expression into a more usable form:
+%
+\begin{align}
+t_n =& k \log \frac{\hat{p}}{p_0} + (n-k) \log \frac{1 - \hat{p}}{1 - p_0} \nonumber \\
+=& n \left( \hat{p} \log \frac{\hat{p}}{p_0} + (1-\hat{p}) \log \frac{1 - \hat{p}}{1-p_0} \right) \nonumber \\
+=& n \kl \left(Bern(\hat{p}) || Bern(p_0) \right)
+\end{align}
+%
+Where $\kl(f | g) \coloneqq \sum_{x \in \Xcal} f(x) \log \frac{f(x)}{g(x)}$ is the Kullback-Leibler divergence between two arbitrary probability mass functions $f$ and $g$.
+This is also called the relative entropy.
+The KL divergence has the property that:
+%
+\begin{equation}
+\kl (f || g) \geq 0 \quad \text{with equality iff} \quad f(x) = g(x) \quad \forall x \in \Xcal
+\end{equation}
+%
+We can exploit this result for the case that $f \approx g$ to obtain a simplified expression for the KL divergence.
+We begin by defining $\delta (x) \coloneqq f(x) - g(x)$.
+We are interested in the region where $\delta$ is small.
+We start by substituting for $f=\delta + g$ and then taking the Taylor expansion of $\log 1+x$.
+%
+\begin{align*}
+\kl(f||g) &= \sum_{x \in \Xcal} (\delta + g) \log \left(1 + \frac{\delta}{g} \right) \\
+&= \sum_{x \in \Xcal} (\delta + g) \left( \frac{\delta}{g} - \frac{\delta^2}{2g^2} + O(\delta^3) \right) \\
+&= \sum_{x \in \Xcal} \delta + \frac{1}{2} \sum_{x \in \Xcal} \frac{\delta^2}{g} + O(\delta^3) \\
+&= \frac{1}{2} \sum_{x \in \Xcal} \frac{\delta^2}{g} + O(\delta^3) \\
+&= \frac{1}{2} \chi^2(f||g) + O(\delta^3)
+\end{align*}
+%
+Where the summation over $\delta$ evaluates to 0 because $\delta$ is the difference of two valid p.m.f's which each sum to 1 over $x \in \Xcal$.
+We are able to neglect the $O(\delta^3)$ terms for $f$ very close to $g$ we shall see what this means later.
+$\chi^2(f||g)$ is known as the chi-squared distance between two distributions and is defined simply as $\chi^2(f||g) \coloneqq \sum_{x \in \Xcal} (f-g)^2/g$.
+We now investigate the chi-squared divergence for $f = Bern(p)$ and $g = Bern(q)$.
+%
+\begin{align*}
+\chi^2(Bern(p)||Bern(q)) &= \frac{(p-q)^2}{q} + \frac{((1-p)-(1-q))^2}{1-q} \\
+&= \frac{(p-q)^2}{q(1-q)} \\
+&= \left(\frac{p-q}{\sqrt{q(1-q)}}\right)^2
+\end{align*}
+%
+Now we can exploit these results to get a workable expression for the test statistic $t_n$.
+%
+\begin{align*}
+t_n &= n \kl \left(Bern(\hat{p}) || Bern(p_0)\right) \\
+	&= \frac{n}{2} \chi^2 \left(Bern(\hat{p}) || Bern(p_0) \right) + nO(\delta^3) \\
+	&= \frac{1}{2} \left( \frac{\hat{p} - p_0}{\sqrt{p_0(1-p_0)/n}} \right)^2 + \epsilon
+\end{align*}
+%
+So far we have been treating $t_n$ as deterministic but it is merely an observation a random variable.
+To make this distinction clear we shall use upper-case to refer to random variables.
+Therefore, we have that the random variable $T_n$ is a function of $\hat{P} \coloneqq K/n$.
+For large n, we can find the distribution of $\hat{P}$ by the Central Limit Theorem.
+%
+\begin{align*}
+\hat{P} &= \frac{K}{n} = \frac{1}{n} \sum_{i=1}^{n} X_i \\
+H_0: \E[X_i] &= p_0, Var(X_i) = p_0(1-p_0) \\
+\text{CLT, as }n \rightarrow \infty: \hat{P} &\sim \Gaussian\left(\mu=p_0, \sigma^2=p_0(1-p_0)/n\right) \\
+\therefore \frac{\hat{P}-p_0}{\sqrt{p_0(1-p_0)/n}} = Z &\sim \Gaussian(\mu=0,\sigma^2=1)
+\end{align*}
+%
+Therefore neglecting the error term $\epsilon$\footnote{See the Appendix for proof of negligibility}, we have that under $H_0$, for sufficiently large $n$.
+%
+\begin{align}
+T_n = \frac{1}{2} Z^2 \sim \frac{1}{2} \chi_1^2
+\end{align}
+%
+By the definition of the chi-squared distribution with one degree of freedom.
+To reject the null hypothesis $H_0$ at the $100(1-\alpha)\%$ confidence level, we require that $P(T_n = t_n | H_0) < \alpha$.
+In other words, a low probability of observing this result under the null hypothesis.
+
+\subsection{Two samples equality of means}
+
+We now complicate things by introducing a second coin, with throws denoted by $\{Y_i\}_{i=1}^{m}$ where we assume each throw is i.i.d Bernoulli with parameter $q$ ($Y_i \sim Bern(q)$).
+Note that the population sizes $n$ and $m$ may be different.
+We define $L \coloneqq \sum_{i=1}^{m} Y_i$ which is the analogue of $K$.
+We now set up our hypotheses to be:
+%
+\begin{align*}
+H_0:& \quad p = q \\
+H_1:& \quad p \neq q
+\end{align*}
+%
+Proceeding as before we can derive a formula for the test statistic.
+%
+\begin{align*}
+t_n &\coloneqq \log \frac{\lik(H_1)}{\lik(H_0)}
+= \log \frac{\max_{p,q} P(K=k|p) P(L=l|q)}{\max_p P(K=k|p) P(L=l|q=p)} \\
+&= \log \frac
+{\binom{n}{k} \hat{p}^k (1 - \hat{p})^{n-k} \binom{m}{l} \hat{q}^k (1 - \hat{q})^{m-l}}
+{\binom{n}{k} \hat{r}^k (1 - \hat{r})^{n-k} \binom{m}{l} \hat{r}^l (1 - \hat{r})^{m-l}} \\
+&= \log \frac
+{ \hat{p}^k (1 - \hat{p})^{n-k}}
+{ \hat{r}^k (1 - \hat{r})^{n-k}}
++ \log \frac{\hat{q}^k (1 - \hat{q})^{m-l}}{ \hat{r}^l (1 - \hat{r})^{m-l}}
+\end{align*} 
+%
+Where, $\hat{p} \coloneqq k/n, \hat{q} \coloneqq l/m, \hat{r} \coloneqq (k+l)/(n+m) = (n\hat{p} + m\hat{q})/(n+m)$.
+Using the same tricks as before, we can express this in terms of the chi-squared distance between the various parameters:
+%
+\begin{align*}
+t_n &= \frac{n}{2} \chi^2\left(Bern(\hat{p}||\hat{r})\right)
++ \frac{m}{2} \chi^2\left(Bern(\hat{q}||\hat{r})\right)
++ \epsilon \\
+&\approx \frac{1}{2\hat{r}(1-\hat{r})} \left(
+n(\hat{p} - \hat{r})^2 + m(\hat{q} - \hat{r})^2
+\right) \\
+&= \frac{1}{2\hat{r}(1-\hat{r})} \left(
+n\left( \frac{m(\hat{p} - \hat{q})}{n+m}\right)^2 + 
+m\left( \frac{n(\hat{q} - \hat{p})}{n+m}\right)^2
+\right) \\
+&= \frac{nm(\hat{p} - \hat{q})^2}{2\hat{r}(1-\hat{r})(n+m)}
+\end{align*}
+%
+Under the null hypothesis $H_0$, we require $p=q(=\mu)$; we introduce this third variable $\mu$ to refer to the true mean to avoid ambiguity.
+Applying the central limit theorem and combining Gaussians in the standard way, we have that:
+%
+\begin{align*}
+\hat{P} &\sim \Gaussian \left( \mu, \frac{\mu(1-\mu)}{n} \right) \\
+\hat{Q} &\sim \Gaussian \left(\mu, \frac{\mu(1-\mu)}{m} \right) \\
+\hat{R} &\sim \Gaussian \left(\mu, \frac{\mu(1-\mu)}{n+m} \right) \\
+\therefore \hat{P} - \hat{Q} &\sim \Gaussian \left(0, \frac{\mu(1-\mu)(n+m)}{nm} \right) \\
+\delta \hat{R} \coloneqq \hat{R} - \mu &\sim \Gaussian \left( 0, \frac{\mu(1-\mu)}{n+m} \right)
+\end{align*}
+%
+\section{Appendix}
+
+\subsection{Proving H.O.T can be neglected}
+We have shown that under the null hypothesis, that for sufficiently large $n$, $(\hat{P}-p_0) \sim \Gaussian(0, \beta/n)$ for some positive finite constant $\beta$ (in this case $\beta = p_0(1-p_0)$ but we just require finiteness for this proof).
+Therefore, $Q \coloneqq \sqrt{n} (\hat{P} - p_0) \sim \Gaussian(0,\beta)$.
+Manipulating our expression for the error term $\epsilon$ we can show that it can be expressed as a sum of
+%
+\begin{equation}
+\begin{gathered}
+\epsilon = n O(\delta^3) \\
+\therefore |\epsilon| \leq n \sum_{i=0}^{\infty} \alpha_i|\hat{P} - p_0|^{3+i} 
+\quad \text{for some finite constants } \alpha_i \geq 0 \\
+|\epsilon| \leq \sum_{i=0}^{\infty} \alpha_i n^{-\frac{1+i}{2}}( \sqrt{n}|\hat{P} - p_0|)^{3+i} \\
+|\epsilon| \leq \sum_{i=0}^{\infty} \alpha_i n^{-\frac{1+i}{2}}|Q|^{3+i}
+\end{gathered}
+\end{equation}
+%
+We know that $Q$ is a Gaussian of zero mean and finite variance, therefore $|Q|$ will be a finite value.
+However, we see that it is scaled by a negative power of $n$, therefore for sufficiently large $n$ we have that the error term asymptotes to 0.
+To be precise:
+%
+\begin{equation}
+\begin{gathered}
+\lim_{n \rightarrow \infty} P(|\epsilon| < \eta) = 1 \text{ for arbitrarily small } \eta \ge 0
+\end{gathered}
+\end{equation}
 

### 2020-10-22 report/report.tex
Finish proof

@@
 %
-Proceeding as before we can derive a formula for the test statistic.
+Proceeding as before we can derive a formula for the test statistic. This time we denote the test statistic by $t_N$ where $N \coloneqq n + m$.
 %
 \begin{align*}
-t_n &\coloneqq \log \frac{\lik(H_1)}{\lik(H_0)}
+t_N &\coloneqq \log \frac{\lik(H_1)}{\lik(H_0)}
 = \log \frac{\max_{p,q} P(K=k|p) P(L=l|q)}{\max_p P(K=k|p) P(L=l|q=p)} \\
@@
 \begin{align*}
-t_n &= \frac{n}{2} \chi^2\left(Bern(\hat{p}||\hat{r})\right)
+t_N &= \frac{n}{2} \chi^2\left(Bern(\hat{p}||\hat{r})\right)
 + \frac{m}{2} \chi^2\left(Bern(\hat{q}||\hat{r})\right)
@@
 \right) \\
-&= \frac{nm(\hat{p} - \hat{q})^2}{2\hat{r}(1-\hat{r})(n+m)}
+&= \frac{nm(\hat{p} - \hat{q})^2}{2\hat{r}(1-\hat{r})(n+m)} \\
+&= \frac{1}{2} \left( \sqrt{\frac{nm}{\hat{r}(1-\hat{r})(n+m)}} \left(\hat{p} - \hat{q}\right) \right)^2
 \end{align*}
@@
 Under the null hypothesis $H_0$, we require $p=q(=\mu)$; we introduce this third variable $\mu$ to refer to the true mean to avoid ambiguity.
-Applying the central limit theorem and combining Gaussians in the standard way, we have that:
+Applying the central limit theorem (for sufficiently large $n$ and $m$) and combining Gaussians in the standard way, we have that:
+%
+\begin{equation*}
+\begin{gathered}
+\hat{P} \sim \Gaussian \left( \mu, \frac{\mu(1-\mu)}{n} \right) \\
+\hat{Q} \sim \Gaussian \left(\mu, \frac{\mu(1-\mu)}{m} \right) \\
+\hat{R} \sim \Gaussian \left(\mu, \frac{\mu(1-\mu)}{n+m} \right) \\
+\therefore  \sqrt{\frac{nm}{\mu(1-\mu)(n+m)}} (\hat{P} - \hat{Q}) = Z \sim \Gaussian \left(0, 1 \right) \\
+\delta \hat{R} \coloneqq \hat{R} - \mu \sim \Gaussian \left( 0, \frac{\mu(1-\mu)}{n+m} \right)
+\end{gathered}
+\end{equation*}
+%
+We almost have $T_n$ we just need to demonstrate that $\hat{R}(1-\hat{R})$ is sufficiently close to $\mu(1-\mu)$ for our purposes.
 %
 \begin{align*}
-\hat{P} &\sim \Gaussian \left( \mu, \frac{\mu(1-\mu)}{n} \right) \\
-\hat{Q} &\sim \Gaussian \left(\mu, \frac{\mu(1-\mu)}{m} \right) \\
-\hat{R} &\sim \Gaussian \left(\mu, \frac{\mu(1-\mu)}{n+m} \right) \\
-\therefore \hat{P} - \hat{Q} &\sim \Gaussian \left(0, \frac{\mu(1-\mu)(n+m)}{nm} \right) \\
-\delta \hat{R} \coloneqq \hat{R} - \mu &\sim \Gaussian \left( 0, \frac{\mu(1-\mu)}{n+m} \right)
+\hat{R}(1-\hat{R}) &= (\mu + \delta\hat{R})(1-(\mu + \delta \hat{R})) \\
+&= \mu(1-\mu) + O(\delta\hat{R}) \\
+\therefore \frac{1}{\hat{R}(1-\hat{R})} &= \frac{1}{\mu(1-\mu)} \left(\frac{1}{1+O(\delta\hat{R})}\right) \\
+&= \frac{1}{\mu(1-\mu)} (1 + O(\delta\hat{R})) \\
+&\approx \frac{1}{\mu(1-\mu)}
 \end{align*}
 %
+We can neglect the terms of order $\delta\hat{R}$ and higher powers, as it is zero mean and for sufficiently large $n+m$ the variance approaches 0. Therefore, we have the desired expression for $T_N$.
+%
+\begin{equation}
+T_N \approx \frac{1}{2} \left(
+\sqrt{\frac{nm}{\mu(1-\mu)(n+m)}}
+(\hat{P} - \hat{Q})
+\right)^2 = \frac{1}{2} Z^2 \sim \frac{1}{2} \chi^2_1
+\end{equation}
+%
+For this test we can also use the z-statistic instead of the t-statistic. Since the former is distributed like a Gaussian it may be easier to deal with.
+\begin{align}
+z_N = \frac{\hat{p} - \hat{q}}{\sqrt{\hat{r}(1-\hat{r})\left(1/n + 1/m\right)}} \sim \Gaussian(0, 1)
+\end{align}
 \section{Appendix}

### 2020-11-16 reading/reading.tex
Scaffold bibliography
More sources

--- 2020-10-09 report/report.tex
@@
 %opening
-\title{Detecting Causal Structure in Time Series Data}
+\title{Jormungandr - Reading Notes}
 \author{Lawrence Tray \\ Ioannis Kontoyiannis}
@@
 \usepackage{subcaption}
+\usepackage{mathtools}
+\usepackage{parskip}
 
-%package setup
-\graphicspath{{./img/}}
-\DeclareMathOperator*{\argmax}{arg\,max}
-\DeclareMathOperator*{\argmin}{arg\,min}
-
-%custom commands
-\newcommand{\dft}{\mathcal{F}}
-\newcommand{\idft}{\mathcal{F}^{-1}}
-\newcommand{\Xcal}{\mathcal{X}}
-\newcommand{\cmplx}{\mathbb{C}}
-\newcommand{\figwidth}{0.55\linewidth}
+\usepackage[backend=bibtex]{biblatex}
+\addbibresource{sources.bib}
 
@@
 
-\begin{abstract}
-\end{abstract}
+\section{Initial Reading}
 
-\section{Introduction}
+Summary of \cite{Abbe}.
+
+Types of recovery:
+\begin{enumerate}
+	\item Exact
+	\item Almost Exact
+	\item Partial
+	\item Weak
+\end{enumerate}
+
+For 2 balanced communities in a SSBM (Symmetric SBM) once can apply linear algebra pproaches to solve efficiently. But for more communities, the state of the art is belief propagation.
+
+Many properties are given as functions of the number of edges $n$. Therefore can maybe sub-partition graph to see how properties scale with $n$ to fit the most appropriate model.
+
+Best algorithm for Weak recovery is Belief Propagation (linear version in \cite{Linear-ABP}).
+
+\nocite{*}
+\printbibliography
 

### 2020-11-23 report/report.tex
Fix incorrect formula for null hypothesis rejection
Beautify ABP

@@
 By the definition of the chi-squared distribution with one degree of freedom.
-To reject the null hypothesis $H_0$ at the $100(1-\alpha)\%$ confidence level, we require that $P(T_n = t_n | H_0) < \alpha$.
+To reject the null hypothesis $H_0$ at the $100(1-\alpha)\%$ confidence level, we require that $P(T_n \geq t_n | H_0) < \alpha$.
 In other words, a low probability of observing this result under the null hypothesis.
@@
 \begin{align*}
-t_N &= \frac{n}{2} \chi^2\left(Bern(\hat{p}||\hat{r})\right)
-+ \frac{m}{2} \chi^2\left(Bern(\hat{q}||\hat{r})\right)
-+ \epsilon \\
-&\approx \frac{1}{2\hat{r}(1-\hat{r})} \left(
+t_N &= n \kl\left(Bern(\hat{p}) || Bern(\hat{r})\right)
++ m \kl\left(Bern(\hat{q}) || Bern(\hat{r})\right) \\ 
+&\approx \frac{n}{2} \chi^2\left(Bern(\hat{p})||Bern(\hat{r}))\right)
++ \frac{m}{2} \chi^2\left(Bern(\hat{q})||Bern(\hat{r}))\right) \\
+&= \frac{1}{2\hat{r}(1-\hat{r})} \left(
 n(\hat{p} - \hat{r})^2 + m(\hat{q} - \hat{r})^2
@@
 &= \frac{nm(\hat{p} - \hat{q})^2}{2\hat{r}(1-\hat{r})(n+m)} \\
-&= \frac{1}{2} \left( \sqrt{\frac{nm}{\hat{r}(1-\hat{r})(n+m)}} \left(\hat{p} - \hat{q}\right) \right)^2
+&= \frac{1}{2} \left( \frac{\left(\hat{p} - \hat{q}\right)}{\sqrt{\hat{r}(1-\hat{r})(1/n+1/m)}}  \right)^2
 \end{align*}
@@
 \end{align}
+
+\section{ABP}
+
+ABP(r, T) on a graph $G$ with vertex set $V = V(G)$ and edge set $E = E(G)$
+\begin{enumerate}
+	\item Initialise messages for $(v, v')$:
+		\subitem $y^{(0)}_{v' \rightarrow v} \leftarrow \Gaussian(0, 1)$
+	\item Iterate for $1 \leq t \leq T$ and for $(v, v') \in E$:
+		\subitem compute average $s^{(t-1)} \leftarrow \frac{1}{2|E|} \sum_{(v, v') \in E} y^{(t-1)}_{v' \rightarrow v}$
+		\subitem recentre messages $z^{(t-1)}_{v' \rightarrow v} \leftarrow y^{(t-1)}_{v' \rightarrow v} - s^{(t-1)}$ 
+		\subitem sum incoming $y^{(t)}_{v' \rightarrow v} \leftarrow \sum_{(v', v'') \in E \setminus \{v\}}z^{(t-1)}_{v'' \rightarrow v'}$
+		\subitem if $(v''' \rightarrow v \rightarrow v')$ on cycle of length $r' \leq r$ then correct:
+			\subsubitem $y^{(t)}_{v' \rightarrow v} \leftarrow y^{(t)}_{v' \rightarrow v} - \sum_{(v, v'') \in E \setminus \{v', v'''\}}z^{(t-r')}_{v'' \rightarrow v'}$
+	\item Assignment, for all $v \in V$:
+		\subitem Sum incoming $y_v^{(T)} = \sum_{(v, v') \in E} y^{(t)}_{v' \rightarrow v}$
+		\subitem Assign labels $\sigma_v = 1$ if $y_v^{(T)} > 0$ and $0$ otherwise
+\end{enumerate}
 \section{Appendix}

### 2021-01-12 interim-report/interim-report.tex
Work on interim

--- 2020-11-23 report/report.tex
@@
 %opening
-\title{Detecting Causal Structure in Time Series Data}
+\title{Detecting Structure in Graphical Data}
 \author{Lawrence Tray \\ Ioannis Kontoyiannis}
@@
 \usepackage{mathtools}
+\usepackage{parskip}
+
+%bibliography
+\usepackage[backend=bibtex]{biblatex}
+\addbibresource{interim-sources.bib}
 
@@
 \newcommand{\Xcal}{\mathcal{X}}
+\newcommand{\Dcal}{\mathcal{D}}
 \newcommand{\cmplx}{\mathbb{C}}
@@
 \newcommand{\Gaussian}{\mathcal{N}}
+\newcommand{\Gcal}{\mathcal{G}}
+\newcommand{\Vcal}{\mathcal{V}}
+\newcommand{\Ecal}{\mathcal{E}}
+\newcommand{\pbold}{\boldsymbol{p}}
+\newcommand{\sigmabold}{\boldsymbol{\sigma}}
 \newcommand{\lik}{\mathcal{L}}
 \newcommand{\kl}{\mathcal{D}}
+\newcommand{\Integers}{\mathbb{Z}}
+\newcommand{\SBM}{\textrm{SBM}}
+\newcommand{\iid}{\stackrel{iid}{\sim}}
 \newcommand{\figwidth}{0.55\linewidth}
+
+% definition
+\newtheorem{definition}{Definition}[section]
 
@@
 \begin{abstract}
+So the agenda for today.
+I will take you through my exploratory work around the field in rough chronological order.
+We’ll start off with the fundamentals of hypothesis testing.
+I will then introduce you to the most commonly used graphical model and talk about the ways of verifying structure in a graph with labelled nodes.
+However, that is not sufficient as we may want to detect structure in an unlabelled graph.
+Finally I’ll talk about the future direction of the project.
 \end{abstract}
 
+\tableofcontents
+
 \section{Introduction}
-\section{Hypothesis Testing}
+
+The title started off as ``detecting causal structure in time series data'' but the emphasis has pivoted to the more generic question: detecting structure in data and specifically graphical data.
+
+There is a wealth of graphical data in the world and more is being produced each second; social networks, website hyperlinks and academic collaborations are just some examples of this data.
+There is a wealth of algorithms developed to analyse graphical data.
+Nevertheless, that same principled framework we have for querying classical data (hypothesis testing) is less developed for graphical data.
+Do my friends vote the same way I do or do researchers collaborate with those of the same gender?
+We want to answer these questions and not only that, we wish to report our confidence in the answers.
+To that end there is space to expand the hypothesis testing framework to graphs.
+
+\section{The Stochastic Block Model}
+
+The most popular graphical model in industry and indeed academia is called the Stochastic Block Model (SBM). We use a definition adapted from Abbe \cite{Abbe}.
+
+\begin{definition}
+	\label{defn:sbm}
+	Let $n \in \Integers^+$ be the number of vertices and $k \in \Integers^+$ be the number of communities in an SBM graph.
+We define a probability vector $\pi = [\pi_1, \pi_2 \dots \pi_K]^T$ to be the prior on the K-communities.
+Each vertex $v \in \Vcal = \{1, 2 \dots N\}$ has a community label $X_v \in \{1, 2 \dots n\}$.
+Let $W$ be a symmetric $k \times k$ matrix called the connectivity matrix.
+We say that the pair $(X, \Gcal) \sim \textrm{SBM}(n, \pi, W)$ if X is an $N$-dimensional vector with each component independently distributed as the community prior $X_v \sim \pi$ and $\Gcal$ is an $N$-vertex graph where each pair of vertices $(i, j)$ is connected with probability $p(i \leftrightarrow j) = W_{X_i, X_j}$ independently of other pairs of vertices.
+Lastly, we define the community sets as $\Omega_i = \Omega_i(X) \coloneqq \{v \in \Vcal : X_v = i\}$ which groups our partitions the vertex-set.
+\end{definition}
+
+Obviously this definition imposes further constraints on the values that the connectivity matrix $W$ namely: $0 \leq W_{ij} \leq 1$.
+Though the definition of the SBM is simple, it allows for very deep and rich analysis of graphical datasets.
+
+\section{Verifying Structure}
+
+Armed with this definition we tackle the simplest problem in structure verification.
+Given a graph $\Gcal$ and vertex-labels $X$, we wish to determine whether the two communities $a$ and $b$ connect differently.
+Put formally this a hypothesis test on the parameters of $W$.
+There are three parameters we would wish to test: $W_{aa}, W_{ab}$ and $W_{bb}$ (note that the symmetry constraint requires $W_{ab} = W_{ba}$).
+To do this we can perform three-pairwise hypothesis tests.
+Here we test $W_{\alpha}$ against $W_{\beta}$ where $\alpha$ and $\beta$ are unique indices in $\{(a,a), (a, b), (b,b)\}$:
+%
+\begin{equation}
+\begin{aligned}
+	H_0:& \quad W_{\alpha} = W_{\beta} \\
+	H_1:& \quad W_{\alpha} \neq W_{\beta}
+\end{aligned}
+\end{equation}
+
+We formulate this as a likelihood ratio test.
+Letting $\lik(\Dcal | H)$ denote the likelihood of observing the data $\Dcal = (X, G)$ under hypothesis $H$.
+For ease of notation.
+Therefore, the test statistic is given by:
+%
+\begin{equation}
+	t_n \coloneqq \log \frac{\lik(\Dcal | H_1)}{\lik(\Dcal | H_0)}
+	\label{eqn:test-statistic-start}
+\end{equation}
+
+At this point it helps to introduce some more notation.
+We define $n_i \coloneqq |\Omega_i(X)|$ leading to the result $n = \sum_i n_i$.
+Furthermore, we define $E_{ij} = E_{ij}(X, \Gcal)$ to denote the number of realised edges between communities $i$ and $j$ (note that $i$ may be equal to $j$).
+Furthermore, define $M_{ij} = M_{ij}(X)$ as the maximum number of possible edges between the communities.
+The likelihood function can now be rewritten in more simple terms:
+%
+\begin{align}
+\lik(\Dcal | H) &= p(X| \pi) \cdot p(\Gcal | W, X) \nonumber \\
+&= p(X | \pi) \cdot \prod_{i=1}^{k} \prod_{j=i}^{k} p(E_{ij} | W, X) \nonumber \\
+&= p(X | \pi) \cdot \prod_{i=1}^{k} \prod_{j=i}^{k} W_{ij} ^ {E_{ij}} \cdot \left( 1 - W_{ij} \right) ^ {(M_{ij} - E_{ij})}
+\label{eqn:likelihood-verbose}
+\end{align}
+
+The form of $p(\Gcal | W, X)$ is simply a sequence of Bernoulli trials for each distinct community pair $(i, j)$ (edge present with probability $W_{ij}$ or edge absent with probability $1 - W_{ij}$ for every pair of vertices across those communities).
+A sequence of Bernoullis is the same as a Binomial distribution without the combinatoric term.
+By inspecting equation \ref{eqn:likelihood-verbose} we see that only terms involving $W_{\alpha}$ and $W_{\beta}$ are going to differ under the two hypotheses; the rest of the terms will cancel in our calculation.
+Therefore, we can reqrite the likelihood as follows:
+%
+\begin{equation}
+	\lik (\Dcal | H) \propto f (W_\alpha, E_\alpha, M_\alpha) \cdot f (W_\beta, E_\beta, M_\beta)
+\end{equation} 
+\begin{equation}
+	\textrm{where } f (w, e, m) \coloneqq w^e \cdot (1-w)^{(m - e)}
+\end{equation}
+
+As such we can simplify equation \ref{eqn:test-statistic-start} greatly to give:
+%
+\begin{align}
+	t_n &= \log \frac
+	{
+		\max_{W_{\alpha} \neq W_{\beta}}(f (W_\alpha, E_\alpha, M_\alpha) \cdot f (W_\beta, E_\beta, M_\beta))
+	}
+	{
+		\max_{W_\alpha = W_\beta} (f (W_\alpha, E_\alpha, M_\alpha) \cdot f (W_\beta, E_\beta, M_\beta))
+	} \nonumber \\
+	&= \log \frac{f(E_\alpha / M_\alpha, E_\alpha, M_\alpha) \cdot f(E_\beta / M_\beta, E_\beta, M_\beta)}
+	{f \left( (E_\alpha + E_\beta)/(M_\alpha + M_\beta), E_\alpha + E_\beta, M_\alpha + M_\beta \right)}
+\end{align}
+
+\section{Verifying Structure}
+
+So hypothesis testing.
+I’ll start with an example that may seem trivial now but will be extremely important later for graphical analysis.
+This is the test for equality of means.
+Say we have two coins, X and Y.
+We assume each flip is an independent Bernoulli trial.
+Coin X has probability p of coming up heads, Y has parameter q.
+We flip X, n times and record k heads.
+Y is flipped m times and l heads.
+
+Given this observed data we want to test the null hypothesis that p=q against the alternative that p != q.
+
+There is a very general theorem for hypothesis tests which states that the log-likelihood ratio test statistic (this is the likelihood of the observed data under H1 divided by the likelihood under H0 all logged) The theorem states that the log-likelihood ratio under the null is distributed like ½ times a chi-squared distribution with p degrees of freedom.
+Where p is the number of additional degrees of freedom introduced by H1.
+This allows us to compute the test statistic and reject the null hypothesis if the statistic is large enough.
+
+The first thing I did was formally prove that this theorem holds for the coin flipping example we just spoke about.
+For brevity I’ll omit the details of the proof but it essentially manipulates the expression into such a form that we can apply the Central Limit Theorem to show that the term in the brackets is distributed as a standard gaussian.
+So we now have a robust tool for testing populations for equality of means.
+Put it in your toolbox for now because we will come back to it later.
+
+So now on how to verify evidence of structure in a given graph.
+Suppose we have a graph G with known vertex labels.
+We assume it is drawn from a symmetric SBM with probability p of vertices within the same community being connected and probability q across communities.
+We say it is symmetric because these p and q are the same for all communities.
+We wish to test the hypothesis that these parameters are different against the null that they are equal.
+If the null is rejected then we can say that there is sufficient evidence to show the communities interact with one another differently.
+
+And this is where we reap the dividends from the earlier hypothesis test.
+Two nodes within a community are connected with probability p (this is the equivalent of coming up heads) and they are disconnected with probability 1-p.
+We can reduce this hypothesis test to exactly the one we had before by assigning the parameters as follows.
+
+n is now the number of possible edges within communities
+m is those possible across
+k is the number that are realised within communities
+And l are those realised across
+
+So we now try to apply this test to a real world dataset.
+The Stanford Network Analysis Project (SNAP) has a wealth of graphical datasets.
+In particular I analysed a couple of facebook ego-nets.
+So that is a graph of all my friends and the connections between them.
+Each vertex is also labelled by a feature vector which is entirely anonymised.
+I cannot even tell someone’s gender only whether it is the same as somebody else’s.
+The graph on the right is one such egonet with all gender-77 vertices coloured orange.
+
+So we apply three hypothesis tests on this graph to determine whether gender impacts how people connect on facebook.
+In every case, we reject the null with a very high degree of confidence (in one case we even surpass the precision of the chi-squared cdf).
+We can be certain that gender impacts how people connect but this approach has several limitations.
+
+First of all, we cannot say with certainty how large the impact is.
+Only that gender does impact graphical structure.
+Secondly we must be careful to avoid confounding variables.
+These are variables that we did not account for that could better explain the structure.
+To give an example, we might find evidence that students in Cambridge are more likely to be friends with those of the same gender.
+The confounding variable here could be college affiliation especially given the existence of all-female colleges such as Newnham.
+And lastly these approaches only work if the graph is labelled in advance.
+
+To overcome some of these limitations we now look at the related issue of detecting structure in an unlabeled graph.
+The question is how well we can recover the communities from the graph.
+This is dependent on the parameters of the SBM.
+Each regime here is less strict than the one before.
+We will focus on weak recovery as that is the regime that most real-life graphs occupy and the same algorithms to solve weak recovery can be applied to more neat graphs.
+
+Formally, weak recovery means that we can assign vertex labels with more accuracy than simply sampling from the community prior.
+A different way of putting this is that we simply need to partition the graph’s vertices into two sets S and the complement of S such that this partition cuts across communities.
+Formally, the fraction of nodes from a certain community in our set S is different for different communities.
+
+The most tractable weak recovery algorithm is something called Acyclic Belief Propagation.
+This is a form of message-passing that runs in O(n log n) on graphs with constant degree.
+Since we’re pressed on time I’ll skim over the details.
+We initialise the messages by drawing from a standard Gaussian.
+We propagate these messages along the edges of the graph correcting for cycles of length less than r and then take the output to be the sum of all incoming messages.
+If the output of a vertex is positive we assign it into set S.
+
+So we run ABP on our same ego-net and find that weak recovery is satisfied.
+We partition the graph into two sets and find that the fraction of orange gender in set S is different to the fraction ofpurple gender in set S.
+0.41 is not equal to 0.47.
+But these numbers are still quite close to one another so perhaps gender is not the most important feature by which we can partition the graph.
+And this is where I want to go to next.
+
+So we can run ABP on an unlabelled graph to partition into the two most distinctive sets.
+But what if we also have the vertex labels.
+We could then train a binary classifier to map from feature vectors to classification under ABP.
+This way we will be able to include all feature variables automatically and rank order them by importance.
+
+We would then start to ask more sophisticated questions of the dataset.
+Are friends made independent of politics given age?
+Is the gender of a research partner independent given the research topic?
+We can start to build a much more powerful tool.
+
+Looking slightly longer term, I would like to formalise this approach and link the ABP partition to a robust hypothesis testing framework.
+I may look to bring in causality and seeing if we cannot start to predict graph evolution.
+Remember that almost no graph is static, new vertices and edges are being added all the time.
+I may also extend this approach to more sophisticated graphical models.
+Indeed, the standard SBM is not well suited to tightly clustered graphs.
+The geometric SBM introduces an additional parameter to account for the mutual friends phenomenon.
+You are much more likely to be friends with a friend of your friend than with a random vertex on the graph.
+
+\section{Verifying Structure}
 \subsection{Single Sample against known mean}
@@
 
-\section{ABP}
+\section{Detecting Structure}
+\subsection{ABP}
 
@@
 \end{enumerate}
+
+\section{Composite Approaches}
+
 \section{Appendix}
@@
 
+\nocite{*}
+\printbibliography
+
 \end{document}

### 2021-01-16 interim-report/interim-report.tex
More work on interim

--- 2020-11-23 report/report.tex
@@
 \newcommand{\Gaussian}{\mathcal{N}}
+\newcommand{\Gcal}{\mathcal{G}}
+\newcommand{\Vcal}{\mathcal{V}}
+\newcommand{\Ecal}{\mathcal{E}}
+\newcommand{\pbold}{\boldsymbol{p}}
+\newcommand{\sigmabold}{\boldsymbol{\sigma}}
 \newcommand{\lik}{\mathcal{L}}
 \newcommand{\kl}{\mathcal{D}}
+\newcommand{\Integers}{\mathbb{Z}}
+\newcommand{\SBM}{\textrm{SBM}}
+\newcommand{\iid}{\stackrel{iid}{\sim}}
 \newcommand{\figwidth}{0.55\linewidth}
+
+% definition
+\newtheorem{definition}{Definition}[section]
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{corollary}{Corollary}[theorem]
+\newtheorem{lemma}[theorem]{Lemma}
 
@@
 \begin{abstract}
+So the agenda for today.
+I will take you through my exploratory work around the field in rough chronological order.
+We’ll start off with the fundamentals of hypothesis testing.
+I will then introduce you to the most commonly used graphical model and talk about the ways of verifying structure in a graph with labelled nodes.
+However, that is not sufficient as we may want to detect structure in an unlabelled graph.
+Finally I’ll talk about the future direction of the project.
 \end{abstract}
 
+\tableofcontents
+
 \section{Introduction}
-\section{Hypothesis Testing}
+
+The title started off as ``detecting causal structure in time series data'' but the emphasis has pivoted to the more generic question: detecting structure in data and specifically graphical data.
+
+There is a wealth of graphical data in the world and more is being produced each second; social networks, website hyperlinks and academic collaborations are just some examples of this data.
+There is a wealth of algorithms developed to analyse graphical data.
+Nevertheless, that same principled framework we have for querying classical data (hypothesis testing) is less developed for graphical data.
+Do my friends vote the same way I do or do researchers collaborate with those of the same gender?
+We want to answer these questions and not only that, we wish to report our confidence in the answers.
+To that end there is space to expand the hypothesis testing framework to graphs.
+
+\section{The Stochastic Block Model}
+
+The most popular graphical model in industry and indeed academia is called the Stochastic Block Model (SBM). We use a definition adapted from Abbe \cite{Abbe}.
+
+\begin{definition}
+	\label{defn:sbm}
+	Let $n \in \Integers^+$ be the number of vertices and $k \in \Integers^+$ be the number of communities in an SBM graph.
+We define a probability vector $\pi = [\pi_1, \pi_2 \dots \pi_K]^T$ to be the prior on the K-communities.
+Each vertex $v \in \Vcal = \{1, 2 \dots N\}$ has a community label $X_v \in \{1, 2 \dots n\}$.
+Let $W$ be a symmetric $k \times k$ matrix called the connectivity matrix.
+We say that the pair $(X, \Gcal) \sim \textrm{SBM}(n, \pi, W)$ if X is an $N$-dimensional vector with each component independently distributed as the community prior $X_v \sim \pi$ and $\Gcal$ is an $N$-vertex graph where each pair of vertices $(i, j)$ is connected with probability $p(i \leftrightarrow j) = W_{X_i, X_j}$ independently of other pairs of vertices.
+Lastly, we define the community sets as $\Omega_i = \Omega_i(X) \coloneqq \{v \in \Vcal : X_v = i\}$ which groups our partitions the vertex-set.
+\end{definition}
+
+Obviously this definition imposes further constraints on the values that the connectivity matrix $W$ namely: $0 \leq W_{ij} \leq 1$.
+Though the definition of the SBM is simple, it allows for very deep and rich analysis of graphical datasets.
+
+\section{Verifying Structure}
+
+Armed with this definition we tackle the simplest problem in structure verification.
+Given a graph $\Gcal$ and vertex-labels $X$, we wish to determine whether the two communities $a$ and $b$ connect differently.
+Put formally this a hypothesis test on the parameters of $W$.
+There are three parameters we would wish to test: $W_{aa}, W_{ab}$ and $W_{bb}$ (note that the symmetry constraint requires $W_{ab} = W_{ba}$).
+To do this we can perform three-pairwise hypothesis tests.
+Here we test $W_{\alpha}$ against $W_{\beta}$ where $\alpha$ and $\beta$ are unique indices in $\{(a,a), (a, b), (b,b)\}$:
+%
+\begin{equation}
+\begin{aligned}
+	H_0:& \quad W_{\alpha} = W_{\beta} \\
+	H_1:& \quad W_{\alpha} \neq W_{\beta}
+\end{aligned}
+\end{equation}
+
+We formulate this as a likelihood ratio test.
+Letting $\lik(\Dcal | H)$ denote the likelihood of observing the data $\Dcal = (X, G)$ under hypothesis $H$.
+For ease of notation.
+Therefore, the test statistic is given by:
+%
+\begin{equation}
+	t_n \coloneqq \log \frac{\lik(\Dcal | H_1)}{\lik(\Dcal | H_0)}
+	\label{eqn:test-statistic-start}
+\end{equation}
+
+At this point it helps to introduce some more notation.
+We define $n_i \coloneqq |\Omega_i(X)|$ leading to the result $n = \sum_i n_i$.
+Furthermore, we define $E_{ij} = E_{ij}(X, \Gcal)$ to denote the number of realised edges between communities $i$ and $j$ (in generality $i$ may be equal to $j$).
+Furthermore, define $M_{ij} = M_{ij}(X)$ as the maximum number of possible edges between the communities.
+For an undirected graph this can be computed simply as follows:
+%
+\begin{equation}
+	M_{ij} = M_{ij} (X) = \begin{cases}
+		n_i n_j &\text{for } i \neq j \\
+		\frac{1}{2}n_i (n_i - 1) &\text{for } i = j
+	\end{cases}
+\end{equation}
+
+With this new notation, likelihood function can be written explicitly:
+%
+\begin{align}
+\lik(\Dcal | H) &= p(X| \pi) \cdot p(\Gcal | W, X) \nonumber \\
+&= p(X | \pi) \cdot \prod_{i=1}^{k} \prod_{j=i}^{k} p(E_{ij} | W, X) \nonumber \\
+&= p(X | \pi) \cdot \prod_{i=1}^{k} \prod_{j=i}^{k} W_{ij} ^ {E_{ij}} \cdot \left( 1 - W_{ij} \right) ^ {(M_{ij} - E_{ij})}
+\label{eqn:likelihood-verbose}
+\end{align}
+
+The form of $p(\Gcal | W, X)$ is simply a sequence of Bernoulli trials for each distinct community pair $(i, j)$ (edge present with probability $W_{ij}$ or edge absent with probability $1 - W_{ij}$ for every pair of vertices across those communities).
+A sequence of Bernoullis is the same as a Binomial distribution without the combinatoric term.
+By inspecting equation \ref{eqn:likelihood-verbose} we see that only terms involving $W_{\alpha}$ and $W_{\beta}$ are going to differ under the two hypotheses; the rest of the terms will cancel in our calculation.
+Therefore, we can rewrite the likelihood as follows:
+%
+\begin{equation}
+	\lik (\Dcal | H) \propto f (W_\alpha, E_\alpha, M_\alpha) \cdot f (W_\beta, E_\beta, M_\beta)
+\end{equation} 
+\begin{equation}
+	\textrm{where } f (w, e, m) \coloneqq w^e \cdot (1-w)^{(m - e)}
+	\label{eqn:f-defn}
+\end{equation}
+
+We note that $f(w, e, m)$ is simply the probability of observing a specific sequence of $e$ successes in $m$ independent Bernoulli trials with parameter $w$.
+Its maximiser with respect to the first argument is easily computed through partial differentiation giving:
+%
+\begin{equation}
+	\argmax_w f(w, e, m) = \hat{w} = e / m
+	\label{eqn:f-maximiser}
+\end{equation}
+
+Furthermore, we spot the following property $f(w, e_1, m_1) \cdot f(w, e_2, m_2) = f(w, e_1 + e_2, m_1 + m_2)$ or in other words, the function $f$ is linear in its second and third arguments given the same first argument.
+As such we can manipulate equation \ref{eqn:test-statistic-start} greatly to give:
+%
+\begin{align}
+	t_n &= \log \frac
+	{
+		\max_{W_{\alpha} \neq W_{\beta}}(f (W_\alpha, E_\alpha, M_\alpha) \cdot f (W_\beta, E_\beta, M_\beta))
+	}
+	{
+		\max_{W_\alpha = W_\beta} (f (W_\alpha, E_\alpha, M_\alpha) \cdot f (W_\beta, E_\beta, M_\beta))
+	} \nonumber \\
+	&= \log \frac{
+		\max_p f(p, E_\alpha, M_\alpha) \cdot \max_q f(q, E_\alpha, M_\alpha)
+	}{
+		\max_r f(r, E_\alpha + E_\beta, M_\alpha + M_\beta)
+	} \nonumber \\
+	&= \log \frac{
+		f(\hat{p}, E_\alpha, M_\alpha) \cdot f(\hat{q}, E_\alpha, M_\alpha)
+	}{
+		f(\hat{r}, E_\alpha + E_\beta, M_\alpha + M_\beta)
+	} \nonumber \\
+	&= \log \frac{f(\hat{p}, E_\alpha, M_\alpha)}{f(\hat{r}, E_\alpha, M_\alpha)} + \log \frac{f(\hat{q}, E_\beta, M_\beta)}{f(\hat{r}, E_\beta, M_\beta)}
+\end{align}
+
+Where $\hat{p} \coloneqq E_\alpha / M_\alpha$, $\hat{q} \coloneqq E_\beta / M_\beta$ and $\hat{r} \coloneqq (E_\alpha + E_\beta) / (M_\alpha + M_\beta)$.
+These symbols are introduced to make the notation more succinct.
+
+\begin{lemma}
+	With $f$ defined as in equation \ref{eqn:f-defn}, $0 \leq e \leq m$ and $r \in [0, 1]$ then it holds that $\log \frac{f(e/m, e, m)}{f(r, e, m)} = m \cdot \kl \left( \textrm{Bern}(e/m) || \textrm{Bern}(r) \right)$ where $\kl(g || h)$ is the Kullback-Leibler divergence between two probability mass functions $g, h: \Xcal \mapsto [0, 1]$ defined in discrete space as $\kl(g || h) \coloneqq \sum_{x \in \Xcal} g(x) \log \frac{g(x)}{h(x)}$ and $\textrm{Bern(p)}$ denotes the Bernoulli p.m.f with parameter $p$.
+	\label{lem:kl-div}
+\end{lemma}
+
+Proving lemma \ref{lem:kl-div} is simply a case of algebraic manipulation:
+%
+\begin{align}
+	\log \frac{f(e/m, e, m)}{f(r, e, m)} &= e \cdot \log \frac{e/m}{r} + (m-e) \cdot \log \frac{1- e/m}{1 - r} \nonumber \\
+	&= m \cdot \left( (e/m) \cdot \log \frac{e/m}{r} + (1 - e/m) \cdot \log \frac{1- e/m}{1 - r} \right) \nonumber \\
+	&= m \sum_{x \in \{0, 1\}} \textrm{Bern}(x; e/m) \cdot \log \frac{\textrm{Bern}(x; e/m)}{\textrm{Bern}(x; r)} \nonumber \\
+	&= m \kl \left( \textrm{Bern}(e/m) || \textrm{Bern}(r) \right) \quad \therefore \textrm{QED}
+\end{align}
+
+Thereby proving lemma \ref{lem:kl-div}. This allows us to simplify the test-statistic into a form that is more numerically stable:
+%
+\begin{equation}
+	t_n = M_\alpha \cdot \kl\left( \textrm{Bern}(\hat{p}) || \textrm{Bern}(\hat{r})\right) + 
+	M_\beta \cdot \kl\left( \textrm{Bern}(\hat{q}) || \textrm{Bern}(\hat{r})\right)
+	\label{eqn:test-stat-compute-form}
+\end{equation}
+
+\begin{lemma}
+	For $p \approx r$ then 
+\end{lemma}
+
+\begin{theorem}
+	We posit that $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$.
+Given the realised graph and class labels $(X, \Gcal)$ we can perform a hypothesis test on parameters $W_\alpha$ and $W_\beta$ of the connectivity matrix $W$.
+	%
+	\begin{align*}
+	H_0:& \quad W_{\alpha} = W_{\beta} \\
+	H_1:& \quad W_{\alpha} \neq W_{\beta}
+	\end{align*}
+	If the log-likelihood ratio test statistic $t_n$ is computed as in equation \ref{eqn:test-stat-compute-form}, repeated here:
+	%
+	\begin{equation*}
+		t_n \coloneqq  M_\alpha \cdot \kl\left( \textrm{Bern}(\hat{p}) || \textrm{Bern}(\hat{r})\right) + 
+		M_\beta \cdot \kl\left( \textrm{Bern}(\hat{q}) || \textrm{Bern}(\hat{r})\right)
+	\end{equation*}
+
+	Where $\hat{p} \coloneqq E_\alpha / M_\alpha$, $\hat{q} \coloneqq E_\beta / M_\beta$ and $\hat{r} \coloneqq (E_\alpha + E_\beta) / (M_\alpha + M_\beta)$.
+Then as the number of vertices $n \rightarrow \infty, t_n \sim \frac{1}{2} \chi^2_1$ under the null $H_0$.
+Therefore, we reject $H_0$ at the $100(\zeta)\%$ confidence level if and only if $2t_n \geq \psi^{-1}(\zeta)$, where $\psi^{-1}$ is the $\chi^2_1$ inverse cdf satisfying $Pr(Y \leq \psi^{-1}(\zeta)) = \zeta$ given $Y \sim \chi^2_1$.
+
+	\label{theorem:hyp-test-sbm-chi}
+\end{theorem}
+
+\begin{corollary}
+	We can also use a slightly simpler test-statistic $z_n$ albeit with some loss of generality. If we define $z_n$ to be:
+	%
+	\begin{equation*}
+		z_n \coloneqq \sqrt{\frac{M_\alpha M_\beta}{\hat{r}(1-\hat{r})(M_\alpha + M_\beta)}} \left( \hat{p} - \hat{q} \right)
+	\end{equation*}
+
+	With the symbols retaining their previous meanings.
+Then we show that under the null $H_0$, as $n \rightarrow \infty$ then $z_n \sim \Gaussian(0, 1)$.
+Meaning that, we can construct a similar test to reject $H_0$ at the $100(\zeta)\%$ confidence level if and only if $|z_n| \geq \phi^{-1}(\zeta)$, where $\phi^{-1}$ is the standard inverse Gaussian cdf on magnitude satisfying $Pr(|Y| \leq \phi^{-1}(\zeta)) = \zeta$ given $Y \sim \Gaussian(0, 1)$
+\end{corollary}
+
+\section{Early results}
+
+We seek to apply theorem \ref{theorem:hyp-test-sbm-chi} to various real-world graphical datasets.
+We start by analysing social network graphs.
+The Stanford Network Analysis Project (SNAP for short) \cite{snapnets} offers a wealth of Facebook egonets.
+An egonet is simply a graph where all vertices (in this case each representing a Facebook user) are guaranteed to be connected to one central node (the ego-node).
+The data consists of the undirected set of edges $\Gcal$ indicating whether any two vertices (Facebook users) are connected (friends on Facebook).
+We also have a set of binary labels for each vertex $X$.
+However, for the sake of privacy these features are anonymised.
+This is best explained through illustration:
+%
+\begin{verbatim}
+75 first_name;anonymized feature 75
+76 first_name;anonymized feature 76
+77 gender;anonymized feature 77
+78 gender;anonymized feature 78
+79 hometown;id;anonymized feature 79
+\end{verbatim}
+
+If we have a total of $f$ features and $n$ vertices, then the feature matrix $X$ would be an $n \times f$ where each row is the feature vector for the corresponding vertex.
+The features are binary such that $X_{ij} \in \{0, 1\}$ indicating the feature of and on respectively.
+However, each feature is anonymised to avoid giving away personally identifiable information.
+I cannot tell whether vertex $v$ is male or female but I can tell whether they are the same gender as another vertex $w$.
+
+\section{Verifying Structure}
 \subsection{Single Sample against known mean}

### 2021-01-18 interim-report/interim-report.tex
More work

@@
 \usepackage{parskip}
+\usepackage{fancyvrb}
+\usepackage{bbm}
 
@@
 \newcommand{\iid}{\stackrel{iid}{\sim}}
+\newcommand{\one}{\mathbbm{1}}
 \newcommand{\figwidth}{0.55\linewidth}
@@
 
-The title started off as ``detecting causal structure in time series data'' but the emphasis has pivoted to the more generic question: detecting structure in data and specifically graphical data.
-
 There is a wealth of graphical data in the world and more is being produced each second; social networks, website hyperlinks and academic collaborations are just some examples of this data.
@@
 	Let $n \in \Integers^+$ be the number of vertices and $k \in \Integers^+$ be the number of communities in an SBM graph.
-We define a probability vector $\pi = [\pi_1, \pi_2 \dots \pi_K]^T$ to be the prior on the K-communities.
+We define a probability vector $\pi = [\pi_1, \pi_2 \dots \pi_k]^T$ to be the prior on the k-communities.
 Each vertex $v \in \Vcal = \{1, 2 \dots N\}$ has a community label $X_v \in \{1, 2 \dots n\}$.
-Let $W$ be a symmetric $k \times k$ matrix called the connectivity matrix.
+Let $W$ be a symmetric $k \times k$ matrix with entries in $[0,1]$ called the connectivity matrix.
 We say that the pair $(X, \Gcal) \sim \textrm{SBM}(n, \pi, W)$ if X is an $N$-dimensional vector with each component independently distributed as the community prior $X_v \sim \pi$ and $\Gcal$ is an $N$-vertex graph where each pair of vertices $(i, j)$ is connected with probability $p(i \leftrightarrow j) = W_{X_i, X_j}$ independently of other pairs of vertices.
-Lastly, we define the community sets as $\Omega_i = \Omega_i(X) \coloneqq \{v \in \Vcal : X_v = i\}$ which groups our partitions the vertex-set.
+Lastly, we define the community sets as $\Omega_i = \Omega_i(X) \coloneqq \{v \in \Vcal : X_v = i\}$ which contains all vertices belonging to community $i$.
 \end{definition}
 
-Obviously this definition imposes further constraints on the values that the connectivity matrix $W$ namely: $0 \leq W_{ij} \leq 1$.
-Though the definition of the SBM is simple, it allows for very deep and rich analysis of graphical datasets.
+Though the definition of the SBM is simple, it allows for very deep and rich analysis of graphical datasets. For certain problems it helps to define the symmetric SBM.
+
+\begin{definition}
+	\label{defn:sym-sbm}
+	The symmetric SBM is a special case denoted by $\textrm{SSBM}(n, k, q_{in}, q_{out}) \equiv \textrm{SBM}(n, p, W)$ if the community prior $p$ is uniform ($p_i = 1/k$ for $i \in \{1, 2, \dots k\}$) and $W_{ij}$ takes only two values, one on diagonal and another off diagonal such that $W_{ij} = q_{in}$ for $i=j$ and $W_{ij} = q_{out}$ for $i \neq j$.
+\end{definition}
 
@@
 Armed with this definition we tackle the simplest problem in structure verification.
-Given a graph $\Gcal$ and vertex-labels $X$, we wish to determine whether the two communities $a$ and $b$ connect differently.
-Put formally this a hypothesis test on the parameters of $W$.
-There are three parameters we would wish to test: $W_{aa}, W_{ab}$ and $W_{bb}$ (note that the symmetry constraint requires $W_{ab} = W_{ba}$).
+Given an undirected graph $\Gcal$ and vertex-labels $X$, we wish to determine whether the two communities $a$ and $b$ connect differently.
+Put formally, this is a hypothesis test on the parameters of $W$.
+There are three parameters we would wish to test: $W_{aa}, W_{ab}$ and $W_{bb}$ (note that for an undirected graph $W = W^T$ necessarily so $W_{ab} = W_{ba}$).
 To do this we can perform three-pairwise hypothesis tests.
@@
 
-We formulate this as a likelihood ratio test.
-Letting $\lik(\Dcal | H)$ denote the likelihood of observing the data $\Dcal = (X, G)$ under hypothesis $H$.
-For ease of notation.
-Therefore, the test statistic is given by:
+We formulate this as a likelihood ratio test. Letting $\lik(\Dcal | H)$ denote the likelihood of observing the data $\Dcal = (X, G)$ under hypothesis $H$. Therefore, the test statistic is given by:
 %
@@
 At this point it helps to introduce some more notation.
-We define $n_i \coloneqq |\Omega_i(X)|$ leading to the result $n = \sum_i n_i$.
-Furthermore, we define $E_{ij} = E_{ij}(X, \Gcal)$ to denote the number of realised edges between communities $i$ and $j$ (in generality $i$ may be equal to $j$).
-Furthermore, define $M_{ij} = M_{ij}(X)$ as the maximum number of possible edges between the communities.
+We define the number of vertices in community $i$ by $n_i \coloneqq |\Omega_i(X)|$ leading to the result $n = \sum_i n_i$.
+Furthermore, we use $E_{ij} = E_{ij}(X, \Gcal)$ to denote the number of realised edges between communities $i$ and $j$ (in generality $i$ may be equal to $j$) and similarly define $M_{ij} = M_{ij}(X)$ as the maximum number of possible edges between the communities.
 For an undirected graph this can be computed simply as follows:
@@
 
-With this new notation, likelihood function can be written explicitly:
+With this new notation, the likelihood function can be written explicitly:
 %
@@
 A sequence of Bernoullis is the same as a Binomial distribution without the combinatoric term.
-By inspecting equation \ref{eqn:likelihood-verbose} we see that only terms involving $W_{\alpha}$ and $W_{\beta}$ are going to differ under the two hypotheses; the rest of the terms will cancel in our calculation.
+By inspecting equation \ref{eqn:likelihood-verbose} we see that only terms involving $W_{\alpha}$ and $W_{\beta}$ are going to differ under the two hypotheses; the rest of the terms will cancel in our calculation of the test-statistic $t_n$.
 Therefore, we can rewrite the likelihood as follows:
@@
 \begin{lemma}
-	With $f$ defined as in equation \ref{eqn:f-defn}, $0 \leq e \leq m$ and $r \in [0, 1]$ then it holds that $\log \frac{f(e/m, e, m)}{f(r, e, m)} = m \cdot \kl \left( \textrm{Bern}(e/m) || \textrm{Bern}(r) \right)$ where $\kl(g || h)$ is the Kullback-Leibler divergence between two probability mass functions $g, h: \Xcal \mapsto [0, 1]$ defined in discrete space as $\kl(g || h) \coloneqq \sum_{x \in \Xcal} g(x) \log \frac{g(x)}{h(x)}$ and $\textrm{Bern(p)}$ denotes the Bernoulli p.m.f with parameter $p$.
+	With $f$ defined as in equation \ref{eqn:f-defn}, $0 \leq e \leq m$ and $r \in [0, 1]$ it holds that $\log \frac{f(e/m, e, m)}{f(r, e, m)} = m \cdot \kl \left( \textrm{Bern}(e/m) || \textrm{Bern}(r) \right)$ where $\kl(g || h)$ is the Kullback-Leibler divergence between two probability mass functions $g, h: \Xcal \mapsto [0, 1]$ defined in discrete space as $\kl(g || h) \coloneqq \sum_{x \in \Xcal} g(x) \log \frac{g(x)}{h(x)}$ and $\textrm{Bern(p)}$ denotes the Bernoulli p.m.f with parameter $p$.
 	\label{lem:kl-div}
@@
 \section{Early results}
+\subsection{Social Networks}
 
@@
 However, for the sake of privacy these features are anonymised.
-This is best explained through illustration:
-%
-\begin{verbatim}
+This is best explained through the example below: \\
+
+\begin{center}
+\begin{minipage}{8cm}
+\begin{Verbatim}[fontsize=\small, frame=single, label={\fbox{Example anonymised feature flags}}]
 75 first_name;anonymized feature 75
@@
 79 hometown;id;anonymized feature 79
-\end{verbatim}
+\end{Verbatim}
+\end{minipage}
+\end{center}
 
+Each feature is anonymised to avoid disclosing personally identifiable information.
+I cannot tell whether vertex $v$ is male or female but I can tell whether they are the same gender as another vertex $w$ which suffices for our analysis.
 If we have a total of $f$ features and $n$ vertices, then the feature matrix $X$ would be an $n \times f$ where each row is the feature vector for the corresponding vertex.
 The features are binary such that $X_{ij} \in \{0, 1\}$ indicating the feature of and on respectively.
-However, each feature is anonymised to avoid giving away personally identifiable information.
-I cannot tell whether vertex $v$ is male or female but I can tell whether they are the same gender as another vertex $w$.
+
+We perform a hypothesis test in the manner described by theorem \ref{theorem:hyp-test-sbm-chi} to determine whether gender influences how friends connect on Facebook.
+We choose to analyse SNAP egonet with id 0 (the id of the egonet is the id of the single egonode) though any choice is possible.
+The egonet is plotted on figure \ref{fig:ego-0-by-gender}.
+We use the Python package NetworkX \cite{networkx} for its visualisation tools.
+Indeed, we will discuss layout algorithms later as for now we focus on the simple hypothesis test.
+%
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=\figwidth]{ego-0-by-gender.png}
+	\caption{Egonet 0 with nodes coloured by gender}
+	\label{fig:ego-0-by-gender}
+\end{figure}
+
+We use an SBM model with $k=2$ communities (1: gender-77 on and 2: gender-77 off) to model the egonet and perform a three-way hypothesis test on the parameters of the connectivity Matrix $W$.
+%
+\begin{equation}
+	W = \begin{bmatrix}
+		p_1 & q \\
+		q & p_2
+	\end{bmatrix}
+\end{equation}
+
+Therefore, $p_1$ is the probability that two vertices of gender-77 are connected, $q$ is the crossover probability and $p_2$ is the connection probability within the gender-78 community.
+The results of the hypothesis tests are given in table \ref{tab:egonet-0-hyp-tests} alongside p-values (we choose a 95\% significance level to reject the null).
+However, in some cases the test statistics were so extreme that the p-value saturated to 0.
+%
+\begin{table}[!h]
+	\centering
+	\begin{tabular}{c c c c c}
+		$n_1$                & $n_2$                & $\hat{p_1}$ & $\hat{q}$ & $\hat{p_2}$       \\ \hline
+		130 & 218 & $8.13 \times 10^{-2}$ & $7.66 \times 10^{-2}$ & $10.6 \times 10^{-2}$
+	\end{tabular}
+	\caption{Egonet-0 properties and parameter estimates}
+	\label{tab:egonet-0-props}
+\end{table}
+
+\begin{table}[!h]
+	\centering
+	\begin{tabular}{cccccc}
+		Test & $H_0$                & $H_1$                & p-value & $H_0$ rejected       \\ \hline
+		1    & $p_1 = q$            & $p_1 \neq q$         & 0.158 &  No                \\
+		2    & $p_1 = p_2$ & $p_1 \neq p_2$ & 0.000 & Yes \\
+		3    & $q = p_2$ & $q \neq p_2$ &  0.000 & Yes
+	\end{tabular}
+	\caption{Egonet-0 hypothesis tests that gender influences friendship formation}
+	\label{tab:egonet-0-hyp-tests}
+\end{table}
+
+Test 2 being rejected gives evidence that one gender has on average more friends from their own gender.
+We see that gender-78 people (community 2) have on average more friends ($\hat{p_2} > \hat{p_1}$ in table \ref{tab:egonet-0-props}).
+The rejection of the null in test 3 gives evidence that gender-78 treat gender-77 people differently to their own gender.
+Indeed, a gender-78 person is more likely to be friends with a fellow 78-er than with a 77-er.
+However, we do not reject the null in test 1 so there is insufficient evidence to claim the reverse (that 77-ers tend to stay away from 78-ers).
+Nevertheless, these are results on a single egonet so can hardly be generalised to society as a whole.
+However, the exercise has highlighted some interesting points:
+
+\begin{enumerate}
+	\item The hypothesis test framework does not quantify the magnitude of the difference just whether or not a difference exists.
+	\item As the degree $n$ increases, the p-values will necessarily become more extreme.
+	\item The method relies on the analyst to specify the features of interest so results may be misleading if third variables are missing from the analysis.
+\end{enumerate}
+
+To address these shortcomings we instead approach the problem from a different angle.
+For now we have been given a graph with fully labelled vertices and ask if a given feature affect the structure.
+Instead, we could take a graph and partition into communities first (without using feature information) and then ask which features most reliably explain the partition.
+For that we need a way of detecting structure.
+
+\section{Detecting Structure}
+
+This is the problem of recovering the communities from the graph $\Gcal$.
+We use the notation $\hat{X}$ to denote the produced estimate of the community labels for the $n$ vertices.
+Obviously, if the SBM is symmetric then we must allow for an arbitrary relabelling of the communities $r: \{1, 2 \dots k\} \mapsto \{1, 2 \dots k\}$ before we compare the agreement of the two vectors $\hat{X}$ and $X$.
+The agreement $A$ and normalised agreement $\tilde{A}$ between two vectors $x, y$ are computed as below:
+%
+\begin{align}
+	A(x, y) &= \max_r \frac{1}{n} \sum_{i=1}^{n} \one (x_i, r(y_i)) \\
+	\tilde{A}(x, y) &= \max_r \frac{1}{k} \sum_{i=1}^{k} \frac{1}{n_i} \sum_{v \in \Omega_i(x)} \one(x_v, r(y_v))
+\end{align}
+
+The normalised agreement is important for asymmetric SBMs as it takes the agreement averaged over communities rather than vertices.
+We define 4 recovery regimes (plus the trivial no recovery), such that the following holds (adapted from Abbe \cite{Abbe}):
+
+\begin{definition}
+	For $(X, \Gcal) \sim \textrm{SBM}(m, p, W)$ the following recovery requirements are solved if there exists an algorithm which takes $\Gcal$ and input and outputs a vector of classifications $\hat{X} = \hat{X}(\Gcal)$ such that as $n \rightarrow \infty$:
+	\begin{itemize}
+		\item \textbf{Exact recovery:} $Pr\{A(X, \hat{X}) = 1\} = 1 - o(1)$ correct classification recovered almost surely
+		\item \textbf{Almost exact recovery:} $Pr\{A(X, \hat{X}) = 1 - o(1)\} = 1 - o(1)$ vanishing fraction misclassified almost surely
+		\item \textbf{Partial recovery:} $Pr\{\tilde{A}(X, \hat{X}) \geq \alpha\} = 1 - o(1), 1 > \alpha > 1/k$ better than choosing from uniform prior
+		\item \textbf{Weak recovery:} $Pr\{\tilde{A}(X, \hat{X}) \geq 1/k + \epsilon \} = 1 - o(1), \epsilon > 0$ marginally better than choosing from uniform prior
+		\item \textbf{No recovery:} graph provides no information as to class labels
+	\end{itemize}
+	Where $o(1)$ denotes a family of functions that tend to 0 as $n \rightarrow \infty$.
+The regime $\Gcal$ occupies depends on the parameters of the SBM.
+Each recovery regime is weaker than the one above it so Exact recovery implies weak recovery is possible..
+\end{definition}
+
+We choose to focus on weak recovery as it is the least strict of all regimes and in algorithms for weak recovery can solve stricter recovery regimes too.
+One of the most promising weak recovery algorithms is called Acyclic Belief Propagataion (ABP).
+We skim over the details here and merely represent our form of the algorithm, adapted from \cite{Linear-ABP}.
+ABP(r, T) on a graph $\Gcal$ with vertex set $\Vcal = \Vcal(\Gcal)$ and edge set $\Ecal = \Ecal(\Gcal)$:
+
+\begin{enumerate}
+	\item Initialise messages for $(v, v')$:
+	\subitem $y^{(0)}_{v' \rightarrow v} \leftarrow \Gaussian(0, 1)$
+	\item Iterate for $1 \leq t \leq T$ and for $(v, v') \in \Ecal$:
+	\subitem compute average $s^{(t-1)} \leftarrow \frac{1}{2|E|} \sum_{(v, v') \in \Ecal} y^{(t-1)}_{v' \rightarrow v}$
+	\subitem recentre messages $z^{(t-1)}_{v' \rightarrow v} \leftarrow y^{(t-1)}_{v' \rightarrow v} - s^{(t-1)}$ 
+	\subitem sum incoming $y^{(t)}_{v' \rightarrow v} \leftarrow \sum_{(v', v'') \in \Ecal \setminus \{v\}}z^{(t-1)}_{v'' \rightarrow v'}$
+	\subitem if $(v''' \rightarrow v \rightarrow v')$ on cycle of length $r' \leq r$ then correct:
+	\subsubitem $y^{(t)}_{v' \rightarrow v} \leftarrow y^{(t)}_{v' \rightarrow v} - \sum_{(v, v'') \in \Ecal \setminus \{v', v'''\}}z^{(t-r')}_{v'' \rightarrow v'}$
+	\item Assignment, for all $v \in V$:
+	\subitem Sum incoming $y_v^{(T)} = \sum_{(v, v') \in \Ecal} y^{(t)}_{v' \rightarrow v}$
+	\subitem Assign labels $\sigma_v = 1$ if $y_v^{(T)} > 0$ and $0$ otherwise
+\end{enumerate}
+
+We implement this algorithm in Python and apply it to the Facebook Egonets described earlier.
 

### 2021-01-18 interim-report/interim-report.tex
More work

@@
 %opening
-\title{Detecting Structure in Graphical Data}
+\title{TMR: Detecting Structure in Graphical Data}
 \author{Lawrence Tray \\ Ioannis Kontoyiannis}
@@
 \usepackage{bbm}
+\usepackage{algorithm}% http://ctan.org/pkg/algorithm
+\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
 
@@
 \newcommand{\one}{\mathbbm{1}}
-\newcommand{\figwidth}{0.55\linewidth}
+\newcommand{\figwidth}{0.4\linewidth}
 
@@
 \begin{abstract}
-So the agenda for today.
-I will take you through my exploratory work around the field in rough chronological order.
-We’ll start off with the fundamentals of hypothesis testing.
-I will then introduce you to the most commonly used graphical model and talk about the ways of verifying structure in a graph with labelled nodes.
-However, that is not sufficient as we may want to detect structure in an unlabelled graph.
-Finally I’ll talk about the future direction of the project.
+This is the Technical Milestone Report for the II-B project: detecting structure in graphical data.
+We produce vast quantities of graphical data every day and yet the techniques used to analyse their structure are still in their infancy.
+We start by defining the Stochastic Block Model (SBM) widely used in academia and develop a theorem to verify structure in a given labelled graph.
 \end{abstract}
@@
 \section{Verifying Structure}
+\subsection{Theory}
 
@@
 	} \nonumber \\
-	&= \log \frac{
-		f(\hat{p}, E_\alpha, M_\alpha) \cdot f(\hat{q}, E_\alpha, M_\alpha)
-	}{
-		f(\hat{r}, E_\alpha + E_\beta, M_\alpha + M_\beta)
-	} \nonumber \\
 	&= \log \frac{f(\hat{p}, E_\alpha, M_\alpha)}{f(\hat{r}, E_\alpha, M_\alpha)} + \log \frac{f(\hat{q}, E_\beta, M_\beta)}{f(\hat{r}, E_\beta, M_\beta)}
@@
 
-\begin{lemma}
-	With $f$ defined as in equation \ref{eqn:f-defn}, $0 \leq e \leq m$ and $r \in [0, 1]$ it holds that $\log \frac{f(e/m, e, m)}{f(r, e, m)} = m \cdot \kl \left( \textrm{Bern}(e/m) || \textrm{Bern}(r) \right)$ where $\kl(g || h)$ is the Kullback-Leibler divergence between two probability mass functions $g, h: \Xcal \mapsto [0, 1]$ defined in discrete space as $\kl(g || h) \coloneqq \sum_{x \in \Xcal} g(x) \log \frac{g(x)}{h(x)}$ and $\textrm{Bern(p)}$ denotes the Bernoulli p.m.f with parameter $p$.
+\begin{lemma}[KL divergence]
+	With $f$ defined as in equation \ref{eqn:f-defn}, $0 \leq e \leq m$ and $r \in [0, 1]$ it holds that:
+	%
+	\begin{equation*}
+		\log \frac{f(e/m, e, m)}{f(r, e, m)} = m \cdot \kl \left( \textrm{Bern}(e/m) || \textrm{Bern}(r) \right)
+	\end{equation*}
+
+	where $\kl(g || h)$ is the Kullback-Leibler divergence between two probability mass functions $g, h: \Xcal \mapsto [0, 1]$.
+$\kl$ is defined in discrete space as $\kl(g || h) \coloneqq \sum_{x \in \Xcal} g(x) \log \frac{g(x)}{h(x)}$ and $\textrm{Bern(p)}$ denotes the Bernoulli p.m.f with parameter $p$.
 	\label{lem:kl-div}
@@
 	&= m \cdot \left( (e/m) \cdot \log \frac{e/m}{r} + (1 - e/m) \cdot \log \frac{1- e/m}{1 - r} \right) \nonumber \\
-	&= m \sum_{x \in \{0, 1\}} \textrm{Bern}(x; e/m) \cdot \log \frac{\textrm{Bern}(x; e/m)}{\textrm{Bern}(x; r)} \nonumber \\
 	&= m \kl \left( \textrm{Bern}(e/m) || \textrm{Bern}(r) \right) \quad \therefore \textrm{QED}
@@
 
-\begin{lemma}
-	For $p \approx r$ then 
+However, we must simplify further still.
+%
+\begin{lemma}[$\chi^2$ approximation]
+	\label{lem:kl-approx-chi}
+	For $g(x) \approx h(x)$ then $\kl(g || h) \approx \frac{1}{2} \chi^2(g || h)$ where $\chi^2$ is the chi-squared distance between two distributions defined as $\chi^2(g || h) \coloneqq \sum_{x \in \Xcal} \frac{(g(x)-h(x))^2}{h(x)}$
 \end{lemma}
 
-\begin{theorem}
-	We posit that $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$.
-Given the realised graph and class labels $(X, \Gcal)$ we can perform a hypothesis test on parameters $W_\alpha$ and $W_\beta$ of the connectivity matrix $W$.
+The proof requires defining $\delta(x) \coloneqq g(x) - h(x)$ and taking a Taylor expansion of $\log 1 + x$:
+%
+\begin{align*}
+	\kl(g||h) &= \sum_{x \in \Xcal} (\delta + h) \log \left(1 + \frac{\delta}{g} \right) 
+	= \sum_{x \in \Xcal} (\delta + h) \left( \frac{\delta}{h} - \frac{\delta^2}{2h^2} + O(\delta^3) \right)  \\
+	&= \sum_{x \in \Xcal} \delta + \frac{1}{2} \sum_{x \in \Xcal} \frac{\delta^2}{h} + O(\delta^3) 
+	= \frac{1}{2} \sum_{x \in \Xcal} \frac{\delta^2}{h} + O(\delta^3) \\
+	&= \frac{1}{2} \chi^2(g||h) + O(\delta^3) \approx \frac{1}{2} \chi^2(h||g) \quad \therefore \textrm{QED}
+\end{align*}
+%
+Where the summation over $\delta$ evaluates to 0 because $\delta$ is the difference of two valid p.m.f's which each sum to 1 over $x \in \Xcal$.
+We are able to neglect the $O(\delta^3)$ terms for $g$ very close to $h$.
+The chi-squared distance between two Bernoullis for $g = Bern(p)$ and $h = Bern(q)$ is simply:
+%
+\begin{equation}
+\chi^2(\textrm{Bern}(p) || \textrm{Bern}(q)) = \left( \frac{p-q}{\sqrt{q(1-q)}} \right)^2
+\end{equation}
+
+We can apply these results to equation \ref{eqn:test-stat-compute-form} to obtain an approximate expression for $t_n$ under the null for large $n$ such that $\hat{p}$, $\hat{q}$ and $\hat{r}$ are all close meaning Lemma \ref{lem:kl-approx-chi} is valid, giving:
+%
+\begin{align}
+	t_n &\approx \frac{1}{2 \hat{r}(1 - \hat{r})} \left( 
+	M_\alpha \left(\hat{p} - \hat{r}\right)^2 +
+	M_\beta \left(\hat{q} - \hat{r} \right)^2 
+	\right) \nonumber \\
+	&= \frac{1}{2 \hat{r}(1-\hat{r})} \left( 
+	M_\alpha \left(\frac{M_\alpha (\hat{p} - \hat{q})}{M_\alpha + M_\beta} \right)^2 +
+	M_\beta \left(\frac{M_\beta (\hat{q} - \hat{p})}{M_\alpha + M_\beta} \right)^2
+	\right) \nonumber \\
+	&= \frac{1}{2} \left( \frac{\hat{p} - \hat{q}}{\sqrt{\hat{r}(1- \hat{r})(1/M_\alpha + 1/M_\beta)}} \right)^2
+	= \frac{1}{2} z_n^2
+\end{align}
+
+Through application of the Central Limit Theorem (CLT) and Slutsky's Theorem it can be shown that as $n \rightarrow \infty$ under the null $H_0$ the term in brackets on the final line is distributed like a standard Gaussian $z_n \sim \Gaussian(0,1)$.
+This is because under the same assumptions, assuming that $W_\alpha = W_\beta = \mu$, by the CLT $\hat{p} \sim (\mu, \mu(1 - \mu)/ M_\alpha)$, $\hat{q} \sim (\mu, \mu(1 - \mu)/ M_\beta)$ and $\hat{p} \sim (\mu, \mu(1 - \mu)/ (M_\alpha + M_\beta))$.
+Therefore, the whole test-statistic is distributed as $t_n \sim \frac{1}{2} \chi^2_1$ by the definition of the $\chi^2_p$ distribution: $\chi^2_p \sim \sum_{i=1}{p} Z_i^2$ for $Z_i \sim \Gaussian(0,1)$.
+This leads to the following theorem:
+
+\begin{theorem}[$\chi^2$ test]
+	For $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$, given the realised graph and class labels $(X, \Gcal)$ we can perform a hypothesis test on parameters $W_\alpha$ and $W_\beta$ of the connectivity matrix $W$.
 	%
@@
 
-\begin{corollary}
+\begin{corollary}[z-test]
 	We can also use a slightly simpler test-statistic $z_n$ albeit with some loss of generality. If we define $z_n$ to be:
@@
 	\begin{equation*}
-		z_n \coloneqq \sqrt{\frac{M_\alpha M_\beta}{\hat{r}(1-\hat{r})(M_\alpha + M_\beta)}} \left( \hat{p} - \hat{q} \right)
+		z_n \coloneqq \frac{\hat{p} - \hat{q}}{\sqrt{\hat{r}(1- \hat{r})(1/M_\alpha + 1/M_\beta)}}
 	\end{equation*}
@@
 	With the symbols retaining their previous meanings.
-Then we show that under the null $H_0$, as $n \rightarrow \infty$ then $z_n \sim \Gaussian(0, 1)$.
+Under the null $H_0$, as $n \rightarrow \infty$ then $z_n \sim \Gaussian(0, 1)$.
 Meaning that, we can construct a similar test to reject $H_0$ at the $100(\zeta)\%$ confidence level if and only if $|z_n| \geq \phi^{-1}(\zeta)$, where $\phi^{-1}$ is the standard inverse Gaussian cdf on magnitude satisfying $Pr(|Y| \leq \phi^{-1}(\zeta)) = \zeta$ given $Y \sim \Gaussian(0, 1)$
@@
 
-\section{Early results}
-\subsection{Social Networks}
+\subsection{Early results}
 
@@
 \section{Detecting Structure}
+\subsection{Weak recovery}
 
@@
 
-\begin{definition}
-	For $(X, \Gcal) \sim \textrm{SBM}(m, p, W)$ the following recovery requirements are solved if there exists an algorithm which takes $\Gcal$ and input and outputs a vector of classifications $\hat{X} = \hat{X}(\Gcal)$ such that as $n \rightarrow \infty$:
+\begin{definition}[Recovery regimes]
+	for $(X, \Gcal) \sim \textrm{SBM}(m, p, W)$ the following recovery requirements are solved if there exists an algorithm which takes $\Gcal$ and input and outputs a vector of classifications $\hat{X} = \hat{X}(\Gcal)$ such that as $n \rightarrow \infty$:
 	\begin{itemize}
@@
 
-We choose to focus on weak recovery as it is the least strict of all regimes and in algorithms for weak recovery can solve stricter recovery regimes too.
+We choose to focus on weak recovery as it is the least strict of all regimes. An alternative statement of weak recovery is given below:
+
+\begin{definition}[Weak recovery]
+	\label{defn:weak-recovery}
+	for  $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$ weak recovery is solved if there exists $\epsilon > 0$ and a certain choice of indices $i, j \in \{1, 2 \dots k\}$ and an algorithm that takes as input $\Gcal$ and outputs a partition of the vertex set $\Vcal$ into two distinct sets $S$ and $S^C$ (detected communities 1 and 2 respectively) such that:
+	\begin{equation*}
+		Pr\left\{ \frac{|\Omega_i \cap S|}{|\Omega_i|} - \frac{|\Omega_j \cap S|}{|\Omega_j|} \geq \epsilon \right\} = 1 - o(1)
+	\end{equation*}
+
+	Bearing in mind the definition of $\Omega_i = \Omega_i(X) \coloneqq \{v \in \Vcal : X_v = i\}$.
+In other words, we require the partition $S \subset \Vcal$ to cut across communities such that the fraction of a particular community in partition $S$ with respect to the whole community is different for a choice of two communities.
+\end{definition}
+
 One of the most promising weak recovery algorithms is called Acyclic Belief Propagataion (ABP).
-We skim over the details here and merely represent our form of the algorithm, adapted from \cite{Linear-ABP}.
-ABP(r, T) on a graph $\Gcal$ with vertex set $\Vcal = \Vcal(\Gcal)$ and edge set $\Ecal = \Ecal(\Gcal)$:
+This is a message-passing algorithm that solves weak recovery in the sense of definition \ref{defn:weak-recovery}.
+Algorithm \ref{alg:lin-abp} is heavily adapated from Abbe and Sandon \cite{Linear-ABP}.
+The original paper goes into much more detail but it works on the premise that one can infer the community of a particular vertex by taking a majority vote among the  of all its neighbours.
+The algorithm is non-determinsitic and requires the vertex set $\Vcal = \Vcal(\Gcal)$ and edge set $\Ecal = \Ecal(\Gcal)$ as well as two hyper-parameters $r$ (the maximum cyclic length to correct for) and $T$ (the number of iterations):
 
-\begin{enumerate}
-	\item Initialise messages for $(v, v')$:
-	\subitem $y^{(0)}_{v' \rightarrow v} \leftarrow \Gaussian(0, 1)$
-	\item Iterate for $1 \leq t \leq T$ and for $(v, v') \in \Ecal$:
-	\subitem compute average $s^{(t-1)} \leftarrow \frac{1}{2|E|} \sum_{(v, v') \in \Ecal} y^{(t-1)}_{v' \rightarrow v}$
-	\subitem recentre messages $z^{(t-1)}_{v' \rightarrow v} \leftarrow y^{(t-1)}_{v' \rightarrow v} - s^{(t-1)}$ 
-	\subitem sum incoming $y^{(t)}_{v' \rightarrow v} \leftarrow \sum_{(v', v'') \in \Ecal \setminus \{v\}}z^{(t-1)}_{v'' \rightarrow v'}$
-	\subitem if $(v''' \rightarrow v \rightarrow v')$ on cycle of length $r' \leq r$ then correct:
-	\subsubitem $y^{(t)}_{v' \rightarrow v} \leftarrow y^{(t)}_{v' \rightarrow v} - \sum_{(v, v'') \in \Ecal \setminus \{v', v'''\}}z^{(t-r')}_{v'' \rightarrow v'}$
-	\item Assignment, for all $v \in V$:
-	\subitem Sum incoming $y_v^{(T)} = \sum_{(v, v') \in \Ecal} y^{(t)}_{v' \rightarrow v}$
-	\subitem Assign labels $\sigma_v = 1$ if $y_v^{(T)} > 0$ and $0$ otherwise
-\end{enumerate}
+%
+\begin{algorithm}
+	\caption{Linearised Acyclic Belief Propagation (ABP)}
+	\label{alg:lin-abp}
+	\begin{algorithmic}
+		\Function{ABP}{$\Ecal$, $\Vcal$, r, T}
+		\State $y^{(0)}_{v' \rightarrow v} \gets \Gaussian(0, 1)$ \Comment{Initialise messages randomly}
+		\For{$t \in \{1, 2 \cdots T\}$}
+		\For{$(v, v') \in \Ecal$}
+		\State $s^{(t-1)} \gets \frac{1}{2|E|} \sum_{(v, v') \in \Ecal} y^{(t-1)}_{v' \rightarrow v}$ \Comment{Compute average}
+		\State $z^{(t-1)}_{v' \rightarrow v} \gets y^{(t-1)}_{v' \rightarrow v} - s^{(t-1)}$ \Comment{Recentre messages}
+		\State $y^{(t)}_{v' \rightarrow v} \gets \sum_{(v', v'') \in \Ecal \setminus \{v\}}z^{(t-1)}_{v'' \rightarrow v'}$ \Comment{Sum incoming messages} \\
+		\If {$\exists v''' \in \Vcal$ such that $(v''' \rightarrow v \rightarrow v')$ is on a cycle of length $r' \leq r$}
+		\State $y^{(t)}_{v' \rightarrow v} \leftarrow y^{(t)}_{v' \rightarrow v} - \sum_{(v, v'') \in \Ecal \setminus \{v', v'''\}}z^{(t-r')}_{v'' \rightarrow v'}$ \Comment{Cyclic correction}
+		\EndIf
+		\EndFor
+		\EndFor
+		\State $\sigma_v = \sum_{(v, v') \in \Ecal} y^{(T)}_{v' \rightarrow v}$ \Comment{Sum incoming messages}
+		\State $\hat{X}_v \gets 1$ if $\sigma_v > 0$ else $2$ \Comment{Hard assignment}
+		\State \textbf{return} $\hat{X}_v$
+		\EndFunction
+	\end{algorithmic}
+\end{algorithm}
 
 We implement this algorithm in Python and apply it to the Facebook Egonets described earlier.
-
-\section{Verifying Structure}
-\subsection{Single Sample against known mean}
-We start with the simple case of determining whether or not a coin is fair.
-Each coin flip can be represented by a random variable with a Bernoulli distribution $X_i \sim Bern(p)$.
-Each coin can result in either a Tails or a Heads denoted by $X_i \in \{0, 1\}$.
-We toss the coin $n$ times leading us to a set $\left\{ X_i \right\}_{i=1}^{n}$.
-We wish to test the null hypothesis $p=p_0$ against the alternative.
-We shall keep the $p_0$ notation for generality, though for a fair coin we require $p_0=1/2$.
-%
-\begin{align*}
-H_0:& \quad p = p_0 \\
-H_1:& \quad p \neq p_0
-\end{align*}
-%
-We denote the number of heads with the random variable $K \coloneqq \sum_{i=1}^{n} X_i \sim Bern(p, n)$.
-For a particular experiment we observe $K=k$.
-We employ a standard likelihood ratio test.
-The test statistic $t_n$ is calculated as the log-likelihood ratio of observing $K=k$ under $H_1$ and $H_0$.
+We found that the hyperparameter settings $r=3$ and $T=5$ yielded good results.
+We define the normalised ABP output below:
 %
 \begin{equation}
-t_n \coloneqq \log \frac{\lik(H_1)}{\lik(H_0)}
-= \log \frac{\max_p P(K=k|p)}{P(K=k|p=p_0)}
+	\tilde{\sigma}_v \coloneqq \begin{cases}
+		\frac{\sigma_v}{|\sigma_{max}|} \textrm{ for } \sigma_j > 0 \\
+		\frac{\sigma_v}{|\sigma_{min}|} \textrm{ for } \sigma_j \leq 0
+	\end{cases}
 \end{equation}
+
+Where $\sigma_{max}$ is the largest positive value of $\sigma_v$ and $\sigma_{min}$ is the most negative.
+The normalised quantity $\tilde{\sigma}_v \in [-1, +1]$ with $0$ still being the midpoint that determines which detected community vertex $v$ is assigned to.
+We colour each vertex according to this quantity $\tilde{\sigma}_v$ and plot the graph on figure \ref{fig:abp-output-norm}.
+We see intuitively that ABP has worked.
+The blue cluster in the lower right is clearly part of one community and a great many vertices do not seem to belong to any cluster at all so their classification is uncertain (the colour is close to green).
 %
-We note that since $K$ is distributed as a binomial, $P(K=k|p)=\binom{n}{k}p^k(1-p)^{n-k}$.
-If we can vary $p$, this probability is maximised for $p=\hat{p} \coloneqq k/n$.
-Therefore, the test statistic is given by.
+\begin{figure}[!h]
+	\centering
+	\begin{subfigure}{0.4\linewidth}
+		\includegraphics[width=\linewidth]{ego-0-abp.png}
+		\caption{Normalised ABP output $\tilde{\sigma}_v$ colouring vertices}
+		\label{fig:abp-output-norm}
+	\end{subfigure}
+	\begin{subfigure}{0.4\linewidth}
+		\includegraphics[width=\linewidth]{ego-0-abp-by-gender.png}
+		\caption{Vertices grouped by hard ABP assignments and coloured by gender.
+In assigned community 1 ($S$ - left) we have \{58, 122\} gender-77 and gender-78 respectively and in detected community 2 ($S^C$ - right) we have \{72, 96\}}
+		\label{fig:abp-output-by-gender}
+	\end{subfigure}
+	\caption{Egonet-0 weak recovery results}
+\end{figure}
+
+We can evaluate formally whether weak recovery has succeeded by inspecting figure \ref{fig:abp-output-by-gender} and treating gender-77 and gender-78 as the two communities of interest:
 %
 \begin{equation}
-t_n = \log \frac{\binom{n}{k} \hat{p}^k (1 - \hat{p})^{n-k}}{\binom{n}{k} p_0^k (1 - p_0)^{n-k}}
-=  \log \frac{ \hat{p}^k (1 - \hat{p})^{n-k}}{ p_0^k (1 - p_0)^{n-k}}
+	\frac{|\Omega_{77} \cap S|}{|\Omega_{77}|} = \frac{58}{58+72} = 0.446
+	\quad \textrm{and} \quad
+	 \frac{|\Omega_{78} \cap S|}{|\Omega_{78}|} = \frac{122}{122+96} = 0.560
 \end{equation}
+
+Clearly the fraction of each community in the partition $S$ differ substantially so we say that weak recovery is solved.
+The set $S$ contains a higher proportion of the overall 78-population than of the 77-population.
+
+\subsection{From detection to inference}
+
+We now have an algorithm (ABP) that partitions our graph into two sets that are most readily explained by an SBM with $k=2$ communities.
+We can now ask the question of which vertex labels most readily explain the separation.
+Therefore, we can determine which features have the largest impact on the graphical structure; this circumvents the binary resolution problem with the hypothesis testing approach.
+
+This part of the project is still in its early stages so these are just preliminary results. For reference we plot the community fractions in the detected set $S$
+\footnote{ABP is non-deterministic so each subplot is the result of a different run of ABP. Subfigures use the same partition $S$ but no exact comparison is possible between figures.}
+on figure \ref{fig:abp-output-community-fractions}.
+It is interesting to note that if two bars on this plot differ substantially then we have satisfied the requirment for weak recovery as defined in \ref{defn:weak-recovery}.
+We necessarily remove the constraint on the community sets $\Omega_i$ to be disjoint as one vertex can have multiple features turned on.
+This can be modelled more precisely through a Mixed Membership Model but for now an SBM with $k=2$ communities suffices.
+
+We perform a linear regression on the normalised ABP output $\tilde{\sigma}_v$
 %
-The combinatoric term $\binom{n}{k}$ cancels out.
-This implies that the order in which the heads land does not matter when determining the fairness of the coin.
-We can work the above expression into a more usable form:
-%
-\begin{align}
-t_n =& k \log \frac{\hat{p}}{p_0} + (n-k) \log \frac{1 - \hat{p}}{1 - p_0} \nonumber \\
-=& n \left( \hat{p} \log \frac{\hat{p}}{p_0} + (1-\hat{p}) \log \frac{1 - \hat{p}}{1-p_0} \right) \nonumber \\
-=& n \kl \left(Bern(\hat{p}) || Bern(p_0) \right)
-\end{align}
-%
-Where $\kl(f | g) \coloneqq \sum_{x \in \Xcal} f(x) \log \frac{f(x)}{g(x)}$ is the Kullback-Leibler divergence between two arbitrary probability mass functions $f$ and $g$.
-This is also called the relative entropy.
-The KL divergence has the property that:
-%
-\begin{equation}
-\kl (f || g) \geq 0 \quad \text{with equality iff} \quad f(x) = g(x) \quad \forall x \in \Xcal
-\end{equation}
-%
-We can exploit this result for the case that $f \approx g$ to obtain a simplified expression for the KL divergence.
-We begin by defining $\delta (x) \coloneqq f(x) - g(x)$.
-We are interested in the region where $\delta$ is small.
-We start by substituting for $f=\delta + g$ and then taking the Taylor expansion of $\log 1+x$.
-%
-\begin{align*}
-\kl(f||g) &= \sum_{x \in \Xcal} (\delta + g) \log \left(1 + \frac{\delta}{g} \right) \\
-&= \sum_{x \in \Xcal} (\delta + g) \left( \frac{\delta}{g} - \frac{\delta^2}{2g^2} + O(\delta^3) \right) \\
-&= \sum_{x \in \Xcal} \delta + \frac{1}{2} \sum_{x \in \Xcal} \frac{\delta^2}{g} + O(\delta^3) \\
-&= \frac{1}{2} \sum_{x \in \Xcal} \frac{\delta^2}{g} + O(\delta^3) \\
-&= \frac{1}{2} \chi^2(f||g) + O(\delta^3)
-\end{align*}
-%
-Where the summation over $\delta$ evaluates to 0 because $\delta$ is the difference of two valid p.m.f's which each sum to 1 over $x \in \Xcal$.
-We are able to neglect the $O(\delta^3)$ terms for $f$ very close to $g$ we shall see what this means later.
-$\chi^2(f||g)$ is known as the chi-squared distance between two distributions and is defined simply as $\chi^2(f||g) \coloneqq \sum_{x \in \Xcal} (f-g)^2/g$.
-We now investigate the chi-squared divergence for $f = Bern(p)$ and $g = Bern(q)$.
-%
-\begin{align*}
-\chi^2(Bern(p)||Bern(q)) &= \frac{(p-q)^2}{q} + \frac{((1-p)-(1-q))^2}{1-q} \\
-&= \frac{(p-q)^2}{q(1-q)} \\
-&= \left(\frac{p-q}{\sqrt{q(1-q)}}\right)^2
-\end{align*}
-%
-Now we can exploit these results to get a workable expression for the test statistic $t_n$.
-%
-\begin{align*}
-t_n &= n \kl \left(Bern(\hat{p}) || Bern(p_0)\right) \\
-	&= \frac{n}{2} \chi^2 \left(Bern(\hat{p}) || Bern(p_0) \right) + nO(\delta^3) \\
-	&= \frac{1}{2} \left( \frac{\hat{p} - p_0}{\sqrt{p_0(1-p_0)/n}} \right)^2 + \epsilon
-\end{align*}
-%
-So far we have been treating $t_n$ as deterministic but it is merely an observation a random variable.
-To make this distinction clear we shall use upper-case to refer to random variables.
-Therefore, we have that the random variable $T_n$ is a function of $\hat{P} \coloneqq K/n$.
-For large n, we can find the distribution of $\hat{P}$ by the Central Limit Theorem.
-%
-\begin{align*}
-\hat{P} &= \frac{K}{n} = \frac{1}{n} \sum_{i=1}^{n} X_i \\
-H_0: \E[X_i] &= p_0, Var(X_i) = p_0(1-p_0) \\
-\text{CLT, as }n \rightarrow \infty: \hat{P} &\sim \Gaussian\left(\mu=p_0, \sigma^2=p_0(1-p_0)/n\right) \\
-\therefore \frac{\hat{P}-p_0}{\sqrt{p_0(1-p_0)/n}} = Z &\sim \Gaussian(\mu=0,\sigma^2=1)
-\end{align*}
-%
-Therefore neglecting the error term $\epsilon$\footnote{See the Appendix for proof of negligibility}, we have that under $H_0$, for sufficiently large $n$.
-%
-\begin{align}
-T_n = \frac{1}{2} Z^2 \sim \frac{1}{2} \chi_1^2
-\end{align}
-%
-By the definition of the chi-squared distribution with one degree of freedom.
-To reject the null hypothesis $H_0$ at the $100(1-\alpha)\%$ confidence level, we require that $P(T_n \geq t_n | H_0) < \alpha$.
-In other words, a low probability of observing this result under the null hypothesis.
+\begin{figure}[!h]
+	\centering
+	\begin{subfigure}{0.6\linewidth}
+		\includegraphics[width=\linewidth]{ego-0-abp-community-fractions.png}
+		\caption{Community fractions $|\Omega_i \cap S| / |\Omega_i|$}
+		\label{fig:abp-output-community-fractions}
+	\end{subfigure}
+	\begin{subfigure}{0.6\linewidth}
+		\includegraphics[width=\linewidth]{ego-0-abp-lin-regression.png}
+		\caption{Linear regression of $\tilde{\sigma}_v$ on select feature ($R^2 = 0.113$)}
+		\label{fig:abp-output-lin-regression}
+	\end{subfigure}
+	\caption{Egonet-0 linear classifier results}
+\end{figure}
 
-\subsection{Two samples equality of means}
+\section{Future Direction}
 
-We now complicate things by introducing a second coin, with throws denoted by $\{Y_i\}_{i=1}^{m}$ where we assume each throw is i.i.d Bernoulli with parameter $q$ ($Y_i \sim Bern(q)$).
-Note that the population sizes $n$ and $m$ may be different.
-We define $L \coloneqq \sum_{i=1}^{m} Y_i$ which is the analogue of $K$.
-We now set up our hypotheses to be:
-%
-\begin{align*}
-H_0:& \quad p = q \\
-H_1:& \quad p \neq q
-\end{align*}
-%
-Proceeding as before we can derive a formula for the test statistic. This time we denote the test statistic by $t_N$ where $N \coloneqq n + m$.
-%
-\begin{align*}
-t_N &\coloneqq \log \frac{\lik(H_1)}{\lik(H_0)}
-= \log \frac{\max_{p,q} P(K=k|p) P(L=l|q)}{\max_p P(K=k|p) P(L=l|q=p)} \\
-&= \log \frac
-{\binom{n}{k} \hat{p}^k (1 - \hat{p})^{n-k} \binom{m}{l} \hat{q}^k (1 - \hat{q})^{m-l}}
-{\binom{n}{k} \hat{r}^k (1 - \hat{r})^{n-k} \binom{m}{l} \hat{r}^l (1 - \hat{r})^{m-l}} \\
-&= \log \frac
-{ \hat{p}^k (1 - \hat{p})^{n-k}}
-{ \hat{r}^k (1 - \hat{r})^{n-k}}
-+ \log \frac{\hat{q}^k (1 - \hat{q})^{m-l}}{ \hat{r}^l (1 - \hat{r})^{m-l}}
-\end{align*} 
-%
-Where, $\hat{p} \coloneqq k/n, \hat{q} \coloneqq l/m, \hat{r} \coloneqq (k+l)/(n+m) = (n\hat{p} + m\hat{q})/(n+m)$.
-Using the same tricks as before, we can express this in terms of the chi-squared distance between the various parameters:
-%
-\begin{align*}
-t_N &= n \kl\left(Bern(\hat{p}) || Bern(\hat{r})\right)
-+ m \kl\left(Bern(\hat{q}) || Bern(\hat{r})\right) \\ 
-&\approx \frac{n}{2} \chi^2\left(Bern(\hat{p})||Bern(\hat{r}))\right)
-+ \frac{m}{2} \chi^2\left(Bern(\hat{q})||Bern(\hat{r}))\right) \\
-&= \frac{1}{2\hat{r}(1-\hat{r})} \left(
-n(\hat{p} - \hat{r})^2 + m(\hat{q} - \hat{r})^2
-\right) \\
-&= \frac{1}{2\hat{r}(1-\hat{r})} \left(
-n\left( \frac{m(\hat{p} - \hat{q})}{n+m}\right)^2 + 
-m\left( \frac{n(\hat{q} - \hat{p})}{n+m}\right)^2
-\right) \\
-&= \frac{nm(\hat{p} - \hat{q})^2}{2\hat{r}(1-\hat{r})(n+m)} \\
-&= \frac{1}{2} \left( \frac{\left(\hat{p} - \hat{q}\right)}{\sqrt{\hat{r}(1-\hat{r})(1/n+1/m)}}  \right)^2
-\end{align*}
-%
-Under the null hypothesis $H_0$, we require $p=q(=\mu)$; we introduce this third variable $\mu$ to refer to the true mean to avoid ambiguity.
-Applying the central limit theorem (for sufficiently large $n$ and $m$) and combining Gaussians in the standard way, we have that:
-%
-\begin{equation*}
-\begin{gathered}
-\hat{P} \sim \Gaussian \left( \mu, \frac{\mu(1-\mu)}{n} \right) \\
-\hat{Q} \sim \Gaussian \left(\mu, \frac{\mu(1-\mu)}{m} \right) \\
-\hat{R} \sim \Gaussian \left(\mu, \frac{\mu(1-\mu)}{n+m} \right) \\
-\therefore  \sqrt{\frac{nm}{\mu(1-\mu)(n+m)}} (\hat{P} - \hat{Q}) = Z \sim \Gaussian \left(0, 1 \right) \\
-\delta \hat{R} \coloneqq \hat{R} - \mu \sim \Gaussian \left( 0, \frac{\mu(1-\mu)}{n+m} \right)
-\end{gathered}
-\end{equation*}
-%
-We almost have $T_n$ we just need to demonstrate that $\hat{R}(1-\hat{R})$ is sufficiently close to $\mu(1-\mu)$ for our purposes.
-%
-\begin{align*}
-\hat{R}(1-\hat{R}) &= (\mu + \delta\hat{R})(1-(\mu + \delta \hat{R})) \\
-&= \mu(1-\mu) + O(\delta\hat{R}) \\
-\therefore \frac{1}{\hat{R}(1-\hat{R})} &= \frac{1}{\mu(1-\mu)} \left(\frac{1}{1+O(\delta\hat{R})}\right) \\
-&= \frac{1}{\mu(1-\mu)} (1 + O(\delta\hat{R})) \\
-&\approx \frac{1}{\mu(1-\mu)}
-\end{align*}
-%
-We can neglect the terms of order $\delta\hat{R}$ and higher powers, as it is zero mean and for sufficiently large $n+m$ the variance approaches 0. Therefore, we have the desired expression for $T_N$.
-%
-\begin{equation}
-T_N \approx \frac{1}{2} \left(
-\sqrt{\frac{nm}{\mu(1-\mu)(n+m)}}
-(\hat{P} - \hat{Q})
-\right)^2 = \frac{1}{2} Z^2 \sim \frac{1}{2} \chi^2_1
-\end{equation}
-%
-For this test we can also use the z-statistic instead of the t-statistic. Since the former is distributed like a Gaussian it may be easier to deal with.
-\begin{align}
-z_N = \frac{\hat{p} - \hat{q}}{\sqrt{\hat{r}(1-\hat{r})\left(1/n + 1/m\right)}} \sim \Gaussian(0, 1)
-\end{align}
-
-\section{Detecting Structure}
-\subsection{ABP}
-
-ABP(r, T) on a graph $G$ with vertex set $V = V(G)$ and edge set $E = E(G)$
-\begin{enumerate}
-	\item Initialise messages for $(v, v')$:
-		\subitem $y^{(0)}_{v' \rightarrow v} \leftarrow \Gaussian(0, 1)$
-	\item Iterate for $1 \leq t \leq T$ and for $(v, v') \in E$:
-		\subitem compute average $s^{(t-1)} \leftarrow \frac{1}{2|E|} \sum_{(v, v') \in E} y^{(t-1)}_{v' \rightarrow v}$
-		\subitem recentre messages $z^{(t-1)}_{v' \rightarrow v} \leftarrow y^{(t-1)}_{v' \rightarrow v} - s^{(t-1)}$ 
-		\subitem sum incoming $y^{(t)}_{v' \rightarrow v} \leftarrow \sum_{(v', v'') \in E \setminus \{v\}}z^{(t-1)}_{v'' \rightarrow v'}$
-		\subitem if $(v''' \rightarrow v \rightarrow v')$ on cycle of length $r' \leq r$ then correct:
-			\subsubitem $y^{(t)}_{v' \rightarrow v} \leftarrow y^{(t)}_{v' \rightarrow v} - \sum_{(v, v'') \in E \setminus \{v', v'''\}}z^{(t-r')}_{v'' \rightarrow v'}$
-	\item Assignment, for all $v \in V$:
-		\subitem Sum incoming $y_v^{(T)} = \sum_{(v, v') \in E} y^{(t)}_{v' \rightarrow v}$
-		\subitem Assign labels $\sigma_v = 1$ if $y_v^{(T)} > 0$ and $0$ otherwise
-\end{enumerate}
-
-\section{Composite Approaches}
-
-\section{Appendix}
-
-\subsection{Proving H.O.T can be neglected}
-We have shown that under the null hypothesis, that for sufficiently large $n$, $(\hat{P}-p_0) \sim \Gaussian(0, \beta/n)$ for some positive finite constant $\beta$ (in this case $\beta = p_0(1-p_0)$ but we just require finiteness for this proof).
-Therefore, $Q \coloneqq \sqrt{n} (\hat{P} - p_0) \sim \Gaussian(0,\beta)$.
-Manipulating our expression for the error term $\epsilon$ we can show that it can be expressed as a sum of
-%
-\begin{equation}
-\begin{gathered}
-\epsilon = n O(\delta^3) \\
-\therefore |\epsilon| \leq n \sum_{i=0}^{\infty} \alpha_i|\hat{P} - p_0|^{3+i} 
-\quad \text{for some finite constants } \alpha_i \geq 0 \\
-|\epsilon| \leq \sum_{i=0}^{\infty} \alpha_i n^{-\frac{1+i}{2}}( \sqrt{n}|\hat{P} - p_0|)^{3+i} \\
-|\epsilon| \leq \sum_{i=0}^{\infty} \alpha_i n^{-\frac{1+i}{2}}|Q|^{3+i}
-\end{gathered}
-\end{equation}
-%
-We know that $Q$ is a Gaussian of zero mean and finite variance, therefore $|Q|$ will be a finite value.
-However, we see that it is scaled by a negative power of $n$, therefore for sufficiently large $n$ we have that the error term asymptotes to 0.
-To be precise:
-%
-\begin{equation}
-\begin{gathered}
-\lim_{n \rightarrow \infty} P(|\epsilon| < \eta) = 1 \text{ for arbitrarily small } \eta \ge 0
-\end{gathered}
-\end{equation}
+\section{Conclusion}
 

### 2021-01-18 interim-report/interim-report.tex
First draft

@@
 \title{TMR: Detecting Structure in Graphical Data}
-\author{Lawrence Tray \\ Ioannis Kontoyiannis}
+\author{Lawrence Tray \\ Sup: Ioannis Kontoyiannis}
 
@@
 \begin{abstract}
-This is the Technical Milestone Report for the II-B project: detecting structure in graphical data.
 We produce vast quantities of graphical data every day and yet the techniques used to analyse their structure are still in their infancy.
-We start by defining the Stochastic Block Model (SBM) widely used in academia and develop a theorem to verify structure in a given labelled graph.
+We start by defining the Stochastic Block Model (SBM) widely used in academia and develop a theorem to verify structure in a labelled graph through a hypothesis test formulation.
+We explore the results of applying this theorem to Facebook friendship egonets and find that gender does indeed influence how friendships form on Facebook but this binary yes-no result is unsatisfactory.
+
+To rank order input features by importance we instead approach the problem from the opposite direction: given a graph we detect the partitions that are in some sense typical and then develop a crude classifier to map from vertex features to partition.
+The classifier can then rank order the input features by importance.
+Nevertheless, work must still be done to refine this classifier and tie it back to the rigorous hypothesis testing framework we have already developed.
 \end{abstract}
@@
 		\EndFor
-		\State $\sigma_v = \sum_{(v, v') \in \Ecal} y^{(T)}_{v' \rightarrow v}$ \Comment{Sum incoming messages}
-		\State $\hat{X}_v \gets 1$ if $\sigma_v > 0$ else $2$ \Comment{Hard assignment}
-		\State \textbf{return} $\hat{X}_v$
+		\State $\sigma_v \gets \sum_{(v, v') \in \Ecal} y^{(T)}_{v' \rightarrow v}$ \Comment{Sum incoming messages}
+		\State $S \gets \{v \in \Vcal : \sigma_v > 0\}$ \Comment{Hard assignment}
+		\State $S^C \gets \{v \in \Vcal : \sigma_v \leq 0\}$
+		\State \textbf{return} $S, S^C$
 		\EndFunction
@@
 \subsection{From detection to inference}
+\label{sect:detection-to-inference}
 
@@
 
-We perform a linear regression on the normalised ABP output $\tilde{\sigma}_v$
+We perform a linear regression on the normalised ABP output vector $\tilde{\sigma}$ with the feature matrix $n \times f$ feature matrix $X$ as the explanatory variables.
+$X_{ij} = 1$ if vertex $i$ has feature $j$ turned on and $0$ if feature $j$ is off.
+The equation we try to fit is given below:
+%
+\begin{equation}
+	\tilde{\sigma} = a + X b + \epsilon
+	\label{eqn:lin-regression}
+\end{equation}
+
+Linear regression finds the constant bias $a$ and $f \times 1$ weight vector $b$ that minimises the mean squared prediction error $\epsilon^T \epsilon / n$.
+The theory for linear regression is well known so omitted here (see \cite{yale-linear-regression} for a reference).
+We perform a linear regression using only gender and language features available in the dataset and plot the parameter coefficient in order of decreasing magnitude (figure \ref{fig:abp-output-lin-regression}).
 %
@@
 
+The results are rather interesting.
+Certain languages are a far stronger predictor of community assignment than language.
+After all, one would hope that a language barrier poses a more tangible obstacle to friendship than gender.
+The bias term has rather low magnitude which just implies that the partition $S$ and $S^C$ are of similar size.
+A keen reader would notice that the bias term is included in figure \ref{fig:abp-output-community-fractions}; this is because we implement the bias by augmenting $X$ with an additional feature flag always to $1$ such that the constant term $a$ in equation \ref{eqn:lin-regression} is subsumed into the $b$ vector.
+Indeed, in figure \ref{fig:abp-output-community-fractions} the fraction of vertices with bias turned on in partition $S$ is very close to 0.5.
+Since all vertices have the bias feature turned on $\Omega_{bias} = \Vcal$, this is equivalent to the number of vertices in $S$ as a fraction of the total: $|S| / |\Vcal|$.
+
+This analysis is of course rather crude as we have failed to take into account multi-collinearity or even determined if a linear regression model is appropriate; indeed, an $R^2$ value of just 0.113 would suggest that it is not.
+Nevertheless, this serves as the platform for subsequent more rigorous analysis.
+
 \section{Future Direction}
 
+The immediate next steps for the project are to refine the classification approach outlined in section \ref{sect:detection-to-inference}.
+Firstly, it should be tied back to the hypothesis testing framework to make sure that the results are proven to be meaningful.
+Secondly, the exact form of the classifier can be explored; a linear regression model is a good starting point but we can expand to a softmax neural network as that would allow us to generalise to more than just two partitions.
+
+Indeed, ABP is severely limited by the fact that it only outputs 2 partitions $S$ and $S^C$.
+Therefore, though weak recovery is solved the results from applying the classification approach may not be natural or easy to interpret.
+We instead would prefer to investigate various different algorithms.
+Force-directed approaches are the standard in industry.
+One of the most common is the Fruchterman-Reingold algorithm which treats vertices as repelling objects and edges as attractive forces (springs) and then performs iterative updates on vertex positions for a finite number of iterations.
+This algorithm is what is used by the NetworkX package \cite{networkx} to determine graph layouts as seen in figures \ref{fig:ego-0-by-gender} and \ref{fig:abp-output-norm}.
+Song and Bressan \cite{force-directed} have had great success using the Fruchterman-Reingold algorithm followed by k-means clustering for community detection.
+Nevertheless, though they show the method produces accurate results there is no rigorous proof that it solves weak recovery.
+
+Lastly, the analysis must of course be expanded to more datasets to benchmark efficacy against other methods.
+Work has already begun analysing the influence of gender academic collaboration graphs from Aminer \cite{aminer}.
+Nevertheless, this dataset does not contain gender explicitly, only first-name.
+However, this is cross-referenced with a list of popular baby-names to get a good replacement (as suggested by West et al.
+\cite{gender-scholarship}).
+Furthermore, the seminal dataset of political blogs from the 2004 US election \cite{polblogs} which started much of the analysis of the SBM will of course be analysed.
+
 \section{Conclusion}
+
+This concludes the Technical Milestone Report for the project thus far.
+We start by fully fleshing out the theoretical foundation of the Stochastic Block Model and hypothesis tests on connectivity parameters.
+Nevertheless, this approach is too narrow in scope to be a useful tool on its own.
+We therefore set about the problem from the opposite direction; rather than verify structure we are already given, we detect structure in the graph without labels and seek to identify which labels best recreate the detected partition.
+Early results are promising but this approach needs to be given the same rigorous treatment as the hypothesis testing framework.
+That will be the emphasis of future work.
 

### 2021-01-18 interim-report/interim-report.tex
Remove table of contents

@@
 
-\tableofcontents
-
 \section{Introduction}
 
-There is a wealth of graphical data in the world and more is being produced each second; social networks, website hyperlinks and academic collaborations are just some examples of this data.
-There is a wealth of algorithms developed to analyse graphical data.
-Nevertheless, that same principled framework we have for querying classical data (hypothesis testing) is less developed for graphical data.
+There is a wealth of graphical data in the world and more is being produced each second; social networks, website hyperlinks and academic collaborations are just some examples.
+There are many algorithms developed to analyse graphical data.
+Nevertheless, that same principled hypothesis testing framework we have for querying classical data is less well-developed for graphical data.
 Do my friends vote the same way I do or do researchers collaborate with those of the same gender?

### 2021-01-18 interim-report/interim-report.tex
rename to hyp-test

@@
 We produce vast quantities of graphical data every day and yet the techniques used to analyse their structure are still in their infancy.
-We start by defining the Stochastic Block Model (SBM) widely used in academia and develop a theorem to verify structure in a labelled graph through a hypothesis test formulation.
-We explore the results of applying this theorem to Facebook friendship egonets and find that gender does indeed influence how friendships form on Facebook but this binary yes-no result is unsatisfactory.
+We start with the definition of the Stochastic Block Model (SBM) widely used in academia and develop a theorem to verify structure in a labelled graph through a hypothesis test.
+We explore the results of applying this theorem to Facebook friendship egonets and find that gender does indeed influence how friendships form on Facebook.
+Nevertheless, this binary yes-no result is unsatisfactory.
 
-To rank order input features by importance we instead approach the problem from the opposite direction: given a graph we detect the partitions that are in some sense typical and then develop a crude classifier to map from vertex features to partition.
-The classifier can then rank order the input features by importance.
-Nevertheless, work must still be done to refine this classifier and tie it back to the rigorous hypothesis testing framework we have already developed.
+To rank order vertex labels by importance we instead approach the problem from the opposite direction: given a graph we detect the partitions that are in some sense typical and then develop a classifier to map from vertex features to partition.
+The classifier then implicitly rank orders the input features by importance.
+Nevertheless, work must still be done to refine this classifier and tie it back to the rigorous hypothesis testing framework already developed.
 \end{abstract}

### 2021-01-19 interim-report/interim-report.tex
Proof-read

@@
 
-Armed with this definition we tackle the simplest problem in structure verification.
-Given an undirected graph $\Gcal$ and vertex-labels $X$, we wish to determine whether the two communities $a$ and $b$ connect differently.
+Armed with definition \ref{defn:sbm} we tackle the simplest problem in structure verification.
+Given an undirected graph $\Gcal$ and vertex-labels $X$ such that $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$, we wish to determine whether the two communities $a$ and $b$ connect differently.
 Put formally, this is a hypothesis test on the parameters of $W$.
@@
 To do this we can perform three-pairwise hypothesis tests.
-Here we test $W_{\alpha}$ against $W_{\beta}$ where $\alpha$ and $\beta$ are unique indices in $\{(a,a), (a, b), (b,b)\}$:
+Here we test $W_{\alpha}$ against $W_{\beta}$ where $\alpha$ and $\beta$ are unique indices in $\{(a,a), (a, b), (b,b)\}$.
+Therefore, each hypothesis test can be formulated as below:
 %
@@
 
-We formulate this as a likelihood ratio test. Letting $\lik(\Dcal | H)$ denote the likelihood of observing the data $\Dcal = (X, G)$ under hypothesis $H$. Therefore, the test statistic is given by:
+We formulate this as a likelihood ratio test. Letting $\lik(D | H)$ denote the likelihood of observing the data $D = (X, \Gcal)$ under hypothesis $H$. Therefore, the test statistic is given by:
 %
 \begin{equation}
-	t_n \coloneqq \log \frac{\lik(\Dcal | H_1)}{\lik(\Dcal | H_0)}
+	t_n \coloneqq \log \frac{\lik(D | H_1)}{\lik(D | H_0)}
 	\label{eqn:test-statistic-start}
@@
 At this point it helps to introduce some more notation.
-We define the number of vertices in community $i$ by $n_i \coloneqq |\Omega_i(X)|$ leading to the result $n = \sum_i n_i$.
-Furthermore, we use $E_{ij} = E_{ij}(X, \Gcal)$ to denote the number of realised edges between communities $i$ and $j$ (in generality $i$ may be equal to $j$) and similarly define $M_{ij} = M_{ij}(X)$ as the maximum number of possible edges between the communities.
-For an undirected graph this can be computed simply as follows:
+We define the number of vertices in community $i$ by $n_i \coloneqq |\Omega_i(X)|$.
+Furthermore, we use $E_{ij} = E_{ij}(X, \Gcal)$ to denote the number of realised edges between communities $i$ and $j$ (in generality $i$ may be equal to $j$) and similarly define $M_{ij} = M_{ij}(X)$ as the maximum number of possible edges between communities $i$ and $j$.
+This is computed simply as follows:
 %
@@
 \begin{align}
-\lik(\Dcal | H) &= p(X| \pi) \cdot p(\Gcal | W, X) \nonumber \\
-&= p(X | \pi) \cdot \prod_{i=1}^{k} \prod_{j=i}^{k} p(E_{ij} | W, X) \nonumber \\
-&= p(X | \pi) \cdot \prod_{i=1}^{k} \prod_{j=i}^{k} W_{ij} ^ {E_{ij}} \cdot \left( 1 - W_{ij} \right) ^ {(M_{ij} - E_{ij})}
+\lik(D | H) &= Pr(X| p) \cdot Pr(\Gcal | W, X) \nonumber \\
+&= Pr(X | p) \cdot \prod_{i=1}^{k} \prod_{j=i}^{k} p(E_{ij} | W, X) \nonumber \\
+&= Pr(X | p) \cdot \prod_{i=1}^{k} \prod_{j=i}^{k} W_{ij} ^ {E_{ij}} \cdot \left( 1 - W_{ij} \right) ^ {(M_{ij} - E_{ij})}
 \label{eqn:likelihood-verbose}
@@
 
-The form of $p(\Gcal | W, X)$ is simply a sequence of Bernoulli trials for each distinct community pair $(i, j)$ (edge present with probability $W_{ij}$ or edge absent with probability $1 - W_{ij}$ for every pair of vertices across those communities).
+The form of $Pr(\Gcal | W, X)$ is simply a sequence of Bernoulli trials for each distinct community pair $(i, j)$; there are $E_{ij}$ edges present each with probability $W_{ij}$ and $(M_{ij} - E_{ij})$ absent with probability $(1 - W_{ij})$.
 A sequence of Bernoullis is the same as a Binomial distribution without the combinatoric term.
-By inspecting equation \ref{eqn:likelihood-verbose} we see that only terms involving $W_{\alpha}$ and $W_{\beta}$ are going to differ under the two hypotheses; the rest of the terms will cancel in our calculation of the test-statistic $t_n$.
+By inspecting equation \ref{eqn:likelihood-verbose} we see that only terms involving $W_{\alpha}$ and $W_{\beta}$ are going to differ under the two hypotheses $H_0$ and $H_1$; the rest of the terms will cancel in our calculation of the test-statistic $t_n$.
 Therefore, we can rewrite the likelihood as follows:
@@
 \begin{equation}
-	\lik (\Dcal | H) \propto f (W_\alpha, E_\alpha, M_\alpha) \cdot f (W_\beta, E_\beta, M_\beta)
+	\lik (D | H) \propto f (W_\alpha, E_\alpha, M_\alpha) \cdot f (W_\beta, E_\beta, M_\beta)
 \end{equation} 
@@
 Where $\hat{p} \coloneqq E_\alpha / M_\alpha$, $\hat{q} \coloneqq E_\beta / M_\beta$ and $\hat{r} \coloneqq (E_\alpha + E_\beta) / (M_\alpha + M_\beta)$.
-These symbols are introduced to make the notation more succinct.
+These are essentially maximum-likelihood estimates of parameters of $W$ and the symbols are introduced to make the notation more succinct.
+Note that the max-likelihood estimate of $W_alpha$ under $H_1$ is denoted by the symbol $\hat{p}$ and should not be confused with the community prior $p$ though context will make the distinction apparent.
+We now state lemma \ref{lem:kl-div}.
 
@@
 
-Thereby proving lemma \ref{lem:kl-div}. This allows us to simplify the test-statistic into a form that is more numerically stable:
+Thereby proving lemma \ref{lem:kl-div}.
+This allows us to simplify the test-statistic into a form that is more numerically stable (equation \ref{eqn:test-stat-compute-form}).
+However, we must simplify further still using lemma \ref{lem:kl-approx-chi}.
 %
@@
 
-However, we must simplify further still.
 %
@@
 
-We can apply these results to equation \ref{eqn:test-stat-compute-form} to obtain an approximate expression for $t_n$ under the null for large $n$ such that $\hat{p}$, $\hat{q}$ and $\hat{r}$ are all close meaning Lemma \ref{lem:kl-approx-chi} is valid, giving:
+We can apply these results to equation \ref{eqn:test-stat-compute-form} to obtain an approximate expression for $t_n$ under the null.
+Under $H_0$ and for large $n$, the estimates $\hat{p}$, $\hat{q}$ and $\hat{r}$ are all close meaning Lemma \ref{lem:kl-approx-chi} is applicable.
+This gives:
 %
@@
 Through application of the Central Limit Theorem (CLT) and Slutsky's Theorem it can be shown that as $n \rightarrow \infty$ under the null $H_0$ the term in brackets on the final line is distributed like a standard Gaussian $z_n \sim \Gaussian(0,1)$.
-This is because under the same assumptions, assuming that $W_\alpha = W_\beta = \mu$, by the CLT $\hat{p} \sim (\mu, \mu(1 - \mu)/ M_\alpha)$, $\hat{q} \sim (\mu, \mu(1 - \mu)/ M_\beta)$ and $\hat{p} \sim (\mu, \mu(1 - \mu)/ (M_\alpha + M_\beta))$.
+This is because under the same assumptions, assuming that $W_\alpha = W_\beta = \mu$, by the CLT $\hat{p} \sim (\mu, \mu(1 - \mu)/ M_\alpha)$, $\hat{q} \sim (\mu, \mu(1 - \mu)/ M_\beta)$ and $\hat{r} \sim (\mu, \mu(1 - \mu)/ (M_\alpha + M_\beta))$.
+The proof is long-winded and therefore omitted from this short report.
 Therefore, the whole test-statistic is distributed as $t_n \sim \frac{1}{2} \chi^2_1$ by the definition of the $\chi^2_p$ distribution: $\chi^2_p \sim \sum_{i=1}{p} Z_i^2$ for $Z_i \sim \Gaussian(0,1)$.
-This leads to the following theorem:
+This leads to theorem \ref{theorem:hyp-test-sbm-chi} and corollary \ref{corollary:z-test}.
 
@@
 Meaning that, we can construct a similar test to reject $H_0$ at the $100(\zeta)\%$ confidence level if and only if $|z_n| \geq \phi^{-1}(\zeta)$, where $\phi^{-1}$ is the standard inverse Gaussian cdf on magnitude satisfying $Pr(|Y| \leq \phi^{-1}(\zeta)) = \zeta$ given $Y \sim \Gaussian(0, 1)$
+	\label{corollary:z-test}
 \end{corollary}
@@
 
-We seek to apply theorem \ref{theorem:hyp-test-sbm-chi} to various real-world graphical datasets.
+We seek to apply theorem \ref{theorem:hyp-test-sbm-chi} to real-world graphical datasets.
 We start by analysing social network graphs.
@@
 The data consists of the undirected set of edges $\Gcal$ indicating whether any two vertices (Facebook users) are connected (friends on Facebook).
-We also have a set of binary labels for each vertex $X$.
+We also have a set of binary feature flags for each vertex $X$.
 However, for the sake of privacy these features are anonymised.
@@
 
-Each feature is anonymised to avoid disclosing personally identifiable information.
-I cannot tell whether vertex $v$ is male or female but I can tell whether they are the same gender as another vertex $w$ which suffices for our analysis.
+Each feature is given an anonymised signature to avoid disclosing personally identifiable information.
+We cannot tell whether vertex $v$ is male or female but we can tell whether they are the same gender as another vertex $w$ which suffices for our analysis.
 If we have a total of $f$ features and $n$ vertices, then the feature matrix $X$ would be an $n \times f$ where each row is the feature vector for the corresponding vertex.
-The features are binary such that $X_{ij} \in \{0, 1\}$ indicating the feature of and on respectively.
+The features are binary such that $X_{ij} \in \{0, 1\}$ indicating the feature being turned off and on respectively.
 
@@
 We choose to analyse SNAP egonet with id 0 (the id of the egonet is the id of the single egonode) though any choice is possible.
-The egonet is plotted on figure \ref{fig:ego-0-by-gender}.
+The egonet is plotted on figure \ref{fig:ego-0-by-gender} with nodes coloured by gender.
 We use the Python package NetworkX \cite{networkx} for its visualisation tools.
@@
 
-Therefore, $p_1$ is the probability that two vertices of gender-77 are connected, $q$ is the crossover probability and $p_2$ is the connection probability within the gender-78 community.
+Therefore, $p_1$ is the probability that two vertices of gender-77 are connected, $q$ is the crossover probability and $p_2$ is the connection probability within the gender-78 community\footnote{Strictly speaking this is the not gender-77 community as there are a handful of vertices that do not disclose their gender to Facebook or do not fall into this binary}.
 The results of the hypothesis tests are given in table \ref{tab:egonet-0-hyp-tests} alongside p-values (we choose a 95\% significance level to reject the null).
@@
 
-Test 2 being rejected gives evidence that one gender has on average more friends from their own gender.
-We see that gender-78 people (community 2) have on average more friends ($\hat{p_2} > \hat{p_1}$ in table \ref{tab:egonet-0-props}).
+Test 2 rejecting the null gives evidence that gender-78 people (community 2) have on average more same-gender friends than 77-ers have ($\hat{p_2} > \hat{p_1}$ in table \ref{tab:egonet-0-props}).
 The rejection of the null in test 3 gives evidence that gender-78 treat gender-77 people differently to their own gender.
@@
 
-\begin{enumerate}
+\begin{itemize}
 	\item The hypothesis test framework does not quantify the magnitude of the difference just whether or not a difference exists.
@@
 	\item The method relies on the analyst to specify the features of interest so results may be misleading if third variables are missing from the analysis.
-\end{enumerate}
+\end{itemize}
 
 To address these shortcomings we instead approach the problem from a different angle.
-For now we have been given a graph with fully labelled vertices and ask if a given feature affect the structure.
-Instead, we could take a graph and partition into communities first (without using feature information) and then ask which features most reliably explain the partition.
+For now we have been given a graph with fully labelled vertices and ask if a given feature affects the structure.
+Instead, we could take a graph and partition first (without using feature information) and then ask which features explain the partition.
 For that we need a way of detecting structure.
@@
 
-We choose to focus on weak recovery as it is the least strict of all regimes. An alternative statement of weak recovery is given below:
+We choose to focus on weak recovery as it is the least strict of all regimes. An alternative statement of weak recovery that implies the one above is given below \cite{Abbe}:
 
@@
 Algorithm \ref{alg:lin-abp} is heavily adapated from Abbe and Sandon \cite{Linear-ABP}.
-The original paper goes into much more detail but it works on the premise that one can infer the community of a particular vertex by taking a majority vote among the  of all its neighbours.
-The algorithm is non-determinsitic and requires the vertex set $\Vcal = \Vcal(\Gcal)$ and edge set $\Ecal = \Ecal(\Gcal)$ as well as two hyper-parameters $r$ (the maximum cyclic length to correct for) and $T$ (the number of iterations):
+The original paper goes into much more detail but it works on the premise that one can infer the community of a particular vertex simply from the labels of all its neighbours.
+The algorithm is non-deterministic and requires the vertex set $\Vcal = \Vcal(\Gcal)$ and edge set $\Ecal = \Ecal(\Gcal)$ as well as two hyper-parameters $r$ (the maximum cyclic length to correct for) and $T$ (the number of iterations).
 
@@
 
-We implement this algorithm in Python and apply it to the Facebook Egonets described earlier.
+We implement this algorithm in Python and apply it to the Facebook egonets described earlier.
 We found that the hyperparameter settings $r=3$ and $T=5$ yielded good results.
-We define the normalised ABP output below:
+As well as the hard partition of $\Vcal$ into $S$ and $S^C$ we can use the normalised ABP output $\tilde{\sigma}_v$ defined below:
 %
@@
 Where $\sigma_{max}$ is the largest positive value of $\sigma_v$ and $\sigma_{min}$ is the most negative.
-The normalised quantity $\tilde{\sigma}_v \in [-1, +1]$ with $0$ still being the midpoint that determines which detected community vertex $v$ is assigned to.
+The normalised quantity has restricted range $\tilde{\sigma}_v \in [-1, +1]$ with $0$ still being the decision boundary for partitioning.
 We colour each vertex according to this quantity $\tilde{\sigma}_v$ and plot the graph on figure \ref{fig:abp-output-norm}.
 We see intuitively that ABP has worked.
-The blue cluster in the lower right is clearly part of one community and a great many vertices do not seem to belong to any cluster at all so their classification is uncertain (the colour is close to green).
+The blue cluster in the lower right is clearly part of the same community.
+Furthermore, many vertices that do not seem to belong to any cluster have uncertain classifications (the colour is close to green).
 %
@@
 We now have an algorithm (ABP) that partitions our graph into two sets that are most readily explained by an SBM with $k=2$ communities.
-We can now ask the question of which vertex labels most readily explain the separation.
-Therefore, we can determine which features have the largest impact on the graphical structure; this circumvents the binary resolution problem with the hypothesis testing approach.
+We can now ask the question of which vertex labels most readily explain the separation and therefore determine which features have the largest impact on the graphical structure; this circumvents the binary resolution problem with the hypothesis testing approach.
 
@@
 on figure \ref{fig:abp-output-community-fractions}.
-It is interesting to note that if two bars on this plot differ substantially then we have satisfied the requirment for weak recovery as defined in \ref{defn:weak-recovery}.
-We necessarily remove the constraint on the community sets $\Omega_i$ to be disjoint as one vertex can have multiple features turned on.
+It is interesting to note that if two bars on this plot differ substantially then we have strong evidence that we have solved weak recovery as in definition \ref{defn:weak-recovery}.
+In a slight abuse of notation, we necessarily remove the constraint on the community sets $\Omega_i$ to be disjoint as one vertex can have multiple features turned on.
+These should no longer be thought of as community sets but rather feature sets.
 This can be modelled more precisely through a Mixed Membership Model but for now an SBM with $k=2$ communities suffices.
 
-We perform a linear regression on the normalised ABP output vector $\tilde{\sigma}$ with the feature matrix $n \times f$ feature matrix $X$ as the explanatory variables.
+We perform a linear regression on the $n \times 1$ normalised ABP output vector $\tilde{\sigma}$ with the $n \times f$ feature matrix $X$ as the explanatory variable.
 $X_{ij} = 1$ if vertex $i$ has feature $j$ turned on and $0$ if feature $j$ is off.
@@
 The results are rather interesting.
-Certain languages are a far stronger predictor of community assignment than language.
+Certain languages are a far stronger predictor of community assignment than gender.
 After all, one would hope that a language barrier poses a more tangible obstacle to friendship than gender.
+Furthermore, the gender coefficients are of the same sign further decreasing the causal impact that gender has given language.
 The bias term has rather low magnitude which just implies that the partition $S$ and $S^C$ are of similar size.
-A keen reader would notice that the bias term is included in figure \ref{fig:abp-output-community-fractions}; this is because we implement the bias by augmenting $X$ with an additional feature flag always to $1$ such that the constant term $a$ in equation \ref{eqn:lin-regression} is subsumed into the $b$ vector.
-Indeed, in figure \ref{fig:abp-output-community-fractions} the fraction of vertices with bias turned on in partition $S$ is very close to 0.5.
-Since all vertices have the bias feature turned on $\Omega_{bias} = \Vcal$, this is equivalent to the number of vertices in $S$ as a fraction of the total: $|S| / |\Vcal|$.
+A keen reader would notice that the bias term is included in figure \ref{fig:abp-output-community-fractions}.
+This is because we implement the bias by augmenting $X$ with an additional feature flag always set to $1$ such that the constant term $a$ in equation \ref{eqn:lin-regression} is subsumed into the $b$ vector.
+Since all vertices have the bias feature turned on $\Omega_{bias} = \Vcal$, the fraction of vertices with bias turned on in $S$ is equivalent to the total number of vertices in $S$ as a fraction of the total in $\Vcal$, or put formally: $|S| / |\Vcal|$.
+Indeed, in figure \ref{fig:abp-output-community-fractions} we see that this quantity is very close to 0.5.
 
@@
 Nevertheless, this serves as the platform for subsequent more rigorous analysis.
+Furthermore, we have not proven the conjecture that the regression coefficients are valid measures of causal impact on graphical structure.
 
@@
 We instead would prefer to investigate various different algorithms.
-Force-directed approaches are the standard in industry.
-One of the most common is the Fruchterman-Reingold algorithm which treats vertices as repelling objects and edges as attractive forces (springs) and then performs iterative updates on vertex positions for a finite number of iterations.
+Force-directed approaches are the standard in industry for their speed of execution.
+One of the most common is the Fruchterman-Reingold algorithm which treats vertices as repelling objects and edges as attractive forces (springs).
+It can then caluclate forces and perform iterative updates on vertex positions for a finite number of iterations.
 This algorithm is what is used by the NetworkX package \cite{networkx} to determine graph layouts as seen in figures \ref{fig:ego-0-by-gender} and \ref{fig:abp-output-norm}.
 Song and Bressan \cite{force-directed} have had great success using the Fruchterman-Reingold algorithm followed by k-means clustering for community detection.
-Nevertheless, though they show the method produces accurate results there is no rigorous proof that it solves weak recovery.
+Nevertheless, though they show the method produces accurate results there is no rigorous proof yet that it solves weak recovery.
 
-Lastly, the analysis must of course be expanded to more datasets to benchmark efficacy against other methods.
+Lastly, the analysis must of course be expanded to more datasets to benchmark efficacy against other methods and ensure the methods developed are generally applicable.
 Work has already begun analysing the influence of gender academic collaboration graphs from Aminer \cite{aminer}.

### 2021-01-19 interim-report/interim-report.tex
Change to 2 columns

@@
 \usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
+\usepackage{multicol}
 
@@
 \newcommand{\one}{\mathbbm{1}}
-\newcommand{\figwidth}{0.4\linewidth}
+\newcommand{\figwidth}{\linewidth}
 
@@
 \end{abstract}
+
+\begin{multicols*}{2}
 
@@
 
-Proving lemma \ref{lem:kl-div} is simply a case of algebraic manipulation:
-%
-\begin{align}
-	\log \frac{f(e/m, e, m)}{f(r, e, m)} &= e \cdot \log \frac{e/m}{r} + (m-e) \cdot \log \frac{1- e/m}{1 - r} \nonumber \\
-	&= m \cdot \left( (e/m) \cdot \log \frac{e/m}{r} + (1 - e/m) \cdot \log \frac{1- e/m}{1 - r} \right) \nonumber \\
-	&= m \kl \left( \textrm{Bern}(e/m) || \textrm{Bern}(r) \right) \quad \therefore \textrm{QED}
-\end{align}
-
-Thereby proving lemma \ref{lem:kl-div}.
-This allows us to simplify the test-statistic into a form that is more numerically stable (equation \ref{eqn:test-stat-compute-form}).
+Proving lemma \ref{lem:kl-div} is simply a case of algebraic manipulation and is omitted here.
+However, this allows us to simplify the test-statistic into a form that is more numerically stable (equation \ref{eqn:test-stat-compute-form}).
 However, we must simplify further still using lemma \ref{lem:kl-approx-chi}.
@@
 \begin{equation}
-	t_n = M_\alpha \cdot \kl\left( \textrm{Bern}(\hat{p}) || \textrm{Bern}(\hat{r})\right) + 
-	M_\beta \cdot \kl\left( \textrm{Bern}(\hat{q}) || \textrm{Bern}(\hat{r})\right)
+	t_n = M_\alpha \kl\left( \textrm{Bern}(\hat{p}) || \textrm{Bern}(\hat{r})\right) + 
+	M_\beta \kl\left( \textrm{Bern}(\hat{q}) || \textrm{Bern}(\hat{r})\right)
 	\label{eqn:test-stat-compute-form}
@@
 
-The proof requires defining $\delta(x) \coloneqq g(x) - h(x)$ and taking a Taylor expansion of $\log 1 + x$:
+The proof requires defining $\delta(x) \coloneqq g(x) - h(x)$ and taking a Taylor expansion of $\log 1 + \epsilon$:
 %
 \begin{align*}
-	\kl(g||h) &= \sum_{x \in \Xcal} (\delta + h) \log \left(1 + \frac{\delta}{g} \right) 
-	= \sum_{x \in \Xcal} (\delta + h) \left( \frac{\delta}{h} - \frac{\delta^2}{2h^2} + O(\delta^3) \right)  \\
-	&= \sum_{x \in \Xcal} \delta + \frac{1}{2} \sum_{x \in \Xcal} \frac{\delta^2}{h} + O(\delta^3) 
-	= \frac{1}{2} \sum_{x \in \Xcal} \frac{\delta^2}{h} + O(\delta^3) \\
+	\kl(g||h) &= \sum_{x \in \Xcal} (\delta + h) \log \left(1 + \frac{\delta}{g} \right) \\
+	&= \sum_{x \in \Xcal} (\delta + h) \left( \frac{\delta}{h} - \frac{\delta^2}{2h^2} + O(\delta^3) \right)  \\
+	&= \sum_{x \in \Xcal} \delta + \frac{1}{2} \sum_{x \in \Xcal} \frac{\delta^2}{h} + O(\delta^3) \\
+	&= \frac{1}{2} \sum_{x \in \Xcal} \frac{\delta^2}{h} + O(\delta^3) \\
 	&= \frac{1}{2} \chi^2(g||h) + O(\delta^3) \approx \frac{1}{2} \chi^2(h||g) \quad \therefore \textrm{QED}
@@
 	\begin{equation*}
-		t_n \coloneqq  M_\alpha \cdot \kl\left( \textrm{Bern}(\hat{p}) || \textrm{Bern}(\hat{r})\right) + 
-		M_\beta \cdot \kl\left( \textrm{Bern}(\hat{q}) || \textrm{Bern}(\hat{r})\right)
+		t_n \coloneqq  M_\alpha \kl\left( \textrm{Bern}(\hat{p}) || \textrm{Bern}(\hat{r})\right) + 
+		M_\beta \kl\left( \textrm{Bern}(\hat{q}) || \textrm{Bern}(\hat{r})\right)
 	\end{equation*}
@@
 %
-\begin{figure}[!h]
+\begin{figure}[H]
 	\centering
@@
 %
-\begin{table}[!h]
+\begin{table}[H]
 	\centering
@@
 
-\begin{table}[!h]
+\begin{table}[H]
 	\centering
@@
 %
-\begin{algorithm}
+\begin{algorithm}[H]
 	\caption{Linearised Acyclic Belief Propagation (ABP)}
@@
 		\Function{ABP}{$\Ecal$, $\Vcal$, r, T}
-		\State $y^{(0)}_{v' \rightarrow v} \gets \Gaussian(0, 1)$ \Comment{Initialise messages randomly}
-		\For{$t \in \{1, 2 \cdots T\}$}
-		\For{$(v, v') \in \Ecal$}
-		\State $s^{(t-1)} \gets \frac{1}{2|E|} \sum_{(v, v') \in \Ecal} y^{(t-1)}_{v' \rightarrow v}$ \Comment{Compute average}
-		\State $z^{(t-1)}_{v' \rightarrow v} \gets y^{(t-1)}_{v' \rightarrow v} - s^{(t-1)}$ \Comment{Recentre messages}
-		\State $y^{(t)}_{v' \rightarrow v} \gets \sum_{(v', v'') \in \Ecal \setminus \{v\}}z^{(t-1)}_{v'' \rightarrow v'}$ \Comment{Sum incoming messages} \\
-		\If {$\exists v''' \in \Vcal$ such that $(v''' \rightarrow v \rightarrow v')$ is on a cycle of length $r' \leq r$}
-		\State $y^{(t)}_{v' \rightarrow v} \leftarrow y^{(t)}_{v' \rightarrow v} - \sum_{(v, v'') \in \Ecal \setminus \{v', v'''\}}z^{(t-r')}_{v'' \rightarrow v'}$ \Comment{Cyclic correction}
+		\State $y^{(0)}_{v' \rightarrow v} \gets \Gaussian(0, 1)$
+		\For{$t \in \{1, 2 \cdots T\}$ and $(v, v') \in \Ecal$}
+		\State $s^{(t-1)} \gets \frac{1}{2|E|} \sum_{(v, v') \in \Ecal} y^{(t-1)}_{v' \rightarrow v}$ \Comment{average}
+		\State $z^{(t-1)}_{v' \rightarrow v} \gets y^{(t-1)}_{v' \rightarrow v} - s^{(t-1)}$ \Comment{recentre}
+		\State $y^{(t)}_{v' \rightarrow v} \gets \sum_{(v', v'') \in \Ecal \setminus \{v\}}z^{(t-1)}_{v'' \rightarrow v'}$ \Comment{sum} \\
+		\If {$(v''' \rightarrow v \rightarrow v')$ on a cycle of length $r' \leq r$}
+		\State $y^{(t)}_{v' \rightarrow v} \leftarrow y^{(t)}_{v' \rightarrow v} - \sum_{(v, v'') \in \Ecal \setminus \{v', v'''\}}z^{(t-r')}_{v'' \rightarrow v'}$
 		\EndIf
 		\EndFor
-		\EndFor
-		\State $\sigma_v \gets \sum_{(v, v') \in \Ecal} y^{(T)}_{v' \rightarrow v}$ \Comment{Sum incoming messages}
-		\State $S \gets \{v \in \Vcal : \sigma_v > 0\}$ \Comment{Hard assignment}
+		\State $\sigma_v \gets \sum_{(v, v') \in \Ecal} y^{(T)}_{v' \rightarrow v}$ \Comment{sum final messages}
+		\State $S \gets \{v \in \Vcal : \sigma_v > 0\}$ \Comment{hard assignment}
 		\State $S^C \gets \{v \in \Vcal : \sigma_v \leq 0\}$
@@
 %
-\begin{figure}[!h]
+\begin{figure}[H]
 	\centering
-	\begin{subfigure}{0.4\linewidth}
+	\begin{subfigure}{\figwidth}
 		\includegraphics[width=\linewidth]{ego-0-abp.png}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.4\linewidth}
+	\begin{subfigure}{\figwidth}
 		\includegraphics[width=\linewidth]{ego-0-abp-by-gender.png}
@@
 \begin{equation}
-	\frac{|\Omega_{77} \cap S|}{|\Omega_{77}|} = \frac{58}{58+72} = 0.446
-	\quad \textrm{and} \quad
-	 \frac{|\Omega_{78} \cap S|}{|\Omega_{78}|} = \frac{122}{122+96} = 0.560
+	\begin{aligned}
+	\frac{|\Omega_{77} \cap S|}{|\Omega_{77}|} &= \frac{58}{58+72} = 0.446 \\
+	\frac{|\Omega_{78} \cap S|}{|\Omega_{78}|} &= \frac{122}{122+96} = 0.560
+	\end{aligned}
 \end{equation}
@@
 %
-\begin{figure}[!h]
+\begin{figure}[H]
 	\centering
-	\begin{subfigure}{0.6\linewidth}
+	\begin{subfigure}{\figwidth}
 		\includegraphics[width=\linewidth]{ego-0-abp-community-fractions.png}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.6\linewidth}
+	\begin{subfigure}{\figwidth}
 		\includegraphics[width=\linewidth]{ego-0-abp-lin-regression.png}
@@
 
-\nocite{*}
+%\nocite{*}
 \printbibliography
 
+\end{multicols*}
+
 \end{document}

### 2021-01-19 interim-report/interim-report.tex
Cut down to 6

@@
 
-Though the definition of the SBM is simple, it allows for very deep and rich analysis of graphical datasets. For certain problems it helps to define the symmetric SBM.
-
-\begin{definition}
-	\label{defn:sym-sbm}
-	The symmetric SBM is a special case denoted by $\textrm{SSBM}(n, k, q_{in}, q_{out}) \equiv \textrm{SBM}(n, p, W)$ if the community prior $p$ is uniform ($p_i = 1/k$ for $i \in \{1, 2, \dots k\}$) and $W_{ij}$ takes only two values, one on diagonal and another off diagonal such that $W_{ij} = q_{in}$ for $i=j$ and $W_{ij} = q_{out}$ for $i \neq j$.
-\end{definition}
+Though the definition of the SBM is simple, it allows for very deep and rich analysis of graphical datasets.
 
@@
 The proof is long-winded and therefore omitted from this short report.
-Therefore, the whole test-statistic is distributed as $t_n \sim \frac{1}{2} \chi^2_1$ by the definition of the $\chi^2_p$ distribution: $\chi^2_p \sim \sum_{i=1}{p} Z_i^2$ for $Z_i \sim \Gaussian(0,1)$.
-This leads to theorem \ref{theorem:hyp-test-sbm-chi} and corollary \ref{corollary:z-test}.
+The result leads to theorem \ref{theorem:hyp-test-sbm-chi}.
 
@@
 
-\begin{corollary}[z-test]
-	We can also use a slightly simpler test-statistic $z_n$ albeit with some loss of generality. If we define $z_n$ to be:
-	%
-	\begin{equation*}
-		z_n \coloneqq \frac{\hat{p} - \hat{q}}{\sqrt{\hat{r}(1- \hat{r})(1/M_\alpha + 1/M_\beta)}}
-	\end{equation*}
-
-	With the symbols retaining their previous meanings.
-Under the null $H_0$, as $n \rightarrow \infty$ then $z_n \sim \Gaussian(0, 1)$.
-Meaning that, we can construct a similar test to reject $H_0$ at the $100(\zeta)\%$ confidence level if and only if $|z_n| \geq \phi^{-1}(\zeta)$, where $\phi^{-1}$ is the standard inverse Gaussian cdf on magnitude satisfying $Pr(|Y| \leq \phi^{-1}(\zeta)) = \zeta$ given $Y \sim \Gaussian(0, 1)$
-	\label{corollary:z-test}
-\end{corollary}
+Note that an alternative test is possible by comparing $z_n$ to a Gaussian.
 
@@
 
-Test 2 rejecting the null gives evidence that gender-78 people (community 2) have on average more same-gender friends than 77-ers have ($\hat{p_2} > \hat{p_1}$ in table \ref{tab:egonet-0-props}).
+Test 2 rejecting the null gives evidence that gender-78 people (community 2) have on average more same-gender friends than 77-ers have ($\hat{p}_2 > \hat{p}_1$ in table \ref{tab:egonet-0-props}).
 The rejection of the null in test 3 gives evidence that gender-78 treat gender-77 people differently to their own gender.
@@
 
-This is the problem of recovering the communities from the graph $\Gcal$.
-We use the notation $\hat{X}$ to denote the produced estimate of the community labels for the $n$ vertices.
+Community detection is the problem of recovering the communities from the graph $\Gcal$.
+We use the notation $\hat{X}$ to denote the estimate of the community labels for the $n$ vertices.
 Obviously, if the SBM is symmetric then we must allow for an arbitrary relabelling of the communities $r: \{1, 2 \dots k\} \mapsto \{1, 2 \dots k\}$ before we compare the agreement of the two vectors $\hat{X}$ and $X$.
-The agreement $A$ and normalised agreement $\tilde{A}$ between two vectors $x, y$ are computed as below:
-%
-\begin{align}
-	A(x, y) &= \max_r \frac{1}{n} \sum_{i=1}^{n} \one (x_i, r(y_i)) \\
-	\tilde{A}(x, y) &= \max_r \frac{1}{k} \sum_{i=1}^{k} \frac{1}{n_i} \sum_{v \in \Omega_i(x)} \one(x_v, r(y_v))
-\end{align}
-
-The normalised agreement is important for asymmetric SBMs as it takes the agreement averaged over communities rather than vertices.
-We define 4 recovery regimes (plus the trivial no recovery), such that the following holds (adapted from Abbe \cite{Abbe}):
-
-\begin{definition}[Recovery regimes]
-	for $(X, \Gcal) \sim \textrm{SBM}(m, p, W)$ the following recovery requirements are solved if there exists an algorithm which takes $\Gcal$ and input and outputs a vector of classifications $\hat{X} = \hat{X}(\Gcal)$ such that as $n \rightarrow \infty$:
-	\begin{itemize}
-		\item \textbf{Exact recovery:} $Pr\{A(X, \hat{X}) = 1\} = 1 - o(1)$ correct classification recovered almost surely
-		\item \textbf{Almost exact recovery:} $Pr\{A(X, \hat{X}) = 1 - o(1)\} = 1 - o(1)$ vanishing fraction misclassified almost surely
-		\item \textbf{Partial recovery:} $Pr\{\tilde{A}(X, \hat{X}) \geq \alpha\} = 1 - o(1), 1 > \alpha > 1/k$ better than choosing from uniform prior
-		\item \textbf{Weak recovery:} $Pr\{\tilde{A}(X, \hat{X}) \geq 1/k + \epsilon \} = 1 - o(1), \epsilon > 0$ marginally better than choosing from uniform prior
-		\item \textbf{No recovery:} graph provides no information as to class labels
-	\end{itemize}
-	Where $o(1)$ denotes a family of functions that tend to 0 as $n \rightarrow \infty$.
-The regime $\Gcal$ occupies depends on the parameters of the SBM.
-Each recovery regime is weaker than the one above it so Exact recovery implies weak recovery is possible..
-\end{definition}
-
-We choose to focus on weak recovery as it is the least strict of all regimes. An alternative statement of weak recovery that implies the one above is given below \cite{Abbe}:
+There are 4 recovery regimes (plus the trivial no recovery), as defined by Abbe \cite{Abbe}.
+We choose to focus on weak recovery as it is the least strict of all.
+Therefore, there are very few constraints on the SBM parameters and the theory can be applied to a wide range of graphs.
+A statement of weak recovery is given in definition \ref{defn:weak-recovery} \cite{Abbe}.
 
@@
 	\label{defn:weak-recovery}
-	for  $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$ weak recovery is solved if there exists $\epsilon > 0$ and a certain choice of indices $i, j \in \{1, 2 \dots k\}$ and an algorithm that takes as input $\Gcal$ and outputs a partition of the vertex set $\Vcal$ into two distinct sets $S$ and $S^C$ (detected communities 1 and 2 respectively) such that:
+	weak recovery is solved for a graph  $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$ if there exists $\epsilon > 0$ and a certain choice of indices $i, j \in \{1, 2 \dots k\}$ and an algorithm that takes as input $\Gcal$ and outputs a partition of the vertex set $\Vcal$ into two distinct sets $S$ and $S^C$ (detected communities 1 and 2 respectively) such that:
 	\begin{equation*}
@@
 
-%\nocite{*}
 \printbibliography

### 2021-01-19 interim-report/interim-report.tex
Heavy cut but not much reduction

@@
-\documentclass[]{article}
+\documentclass[11pt]{article}
 
@@
 \newcommand{\one}{\mathbbm{1}}
-\newcommand{\figwidth}{\linewidth}
+\newcommand{\figwidth}{0.7\linewidth}
 
@@
 Nevertheless, these are results on a single egonet so can hardly be generalised to society as a whole.
-However, the exercise has highlighted some interesting points:
 
-\begin{itemize}
-	\item The hypothesis test framework does not quantify the magnitude of the difference just whether or not a difference exists.
-	\item As the degree $n$ increases, the p-values will necessarily become more extreme.
-	\item The method relies on the analyst to specify the features of interest so results may be misleading if third variables are missing from the analysis.
-\end{itemize}
+However, the exercise has highlighted some interesting points.
+The hypothesis test framework does not quantify the magnitude of the difference just whether or not a difference exists.
+Furthermore, The method relies on the analyst to specify the features of interest so results may be misleading if third variables are missing from the analysis.
 
@@
 One of the most promising weak recovery algorithms is called Acyclic Belief Propagataion (ABP).
+We implement a linearised version adapted from Abbe and Sandon \cite{Linear-ABP}.
 This is a message-passing algorithm that solves weak recovery in the sense of definition \ref{defn:weak-recovery}.
-Algorithm \ref{alg:lin-abp} is heavily adapated from Abbe and Sandon \cite{Linear-ABP}.
-The original paper goes into much more detail but it works on the premise that one can infer the community of a particular vertex simply from the labels of all its neighbours.
+The original paper goes into much more detail than we have space to discuss here.
+Nevertheless, the premise is that one can infer the community of a particular vertex simply from the labels of all its neighbours.
 The algorithm is non-deterministic and requires the vertex set $\Vcal = \Vcal(\Gcal)$ and edge set $\Ecal = \Ecal(\Gcal)$ as well as two hyper-parameters $r$ (the maximum cyclic length to correct for) and $T$ (the number of iterations).
 
-%
-\begin{algorithm}[H]
-	\caption{Linearised Acyclic Belief Propagation (ABP)}
-	\label{alg:lin-abp}
-	\begin{algorithmic}
-		\Function{ABP}{$\Ecal$, $\Vcal$, r, T}
-		\State $y^{(0)}_{v' \rightarrow v} \gets \Gaussian(0, 1)$
-		\For{$t \in \{1, 2 \cdots T\}$ and $(v, v') \in \Ecal$}
-		\State $s^{(t-1)} \gets \frac{1}{2|E|} \sum_{(v, v') \in \Ecal} y^{(t-1)}_{v' \rightarrow v}$ \Comment{average}
-		\State $z^{(t-1)}_{v' \rightarrow v} \gets y^{(t-1)}_{v' \rightarrow v} - s^{(t-1)}$ \Comment{recentre}
-		\State $y^{(t)}_{v' \rightarrow v} \gets \sum_{(v', v'') \in \Ecal \setminus \{v\}}z^{(t-1)}_{v'' \rightarrow v'}$ \Comment{sum} \\
-		\If {$(v''' \rightarrow v \rightarrow v')$ on a cycle of length $r' \leq r$}
-		\State $y^{(t)}_{v' \rightarrow v} \leftarrow y^{(t)}_{v' \rightarrow v} - \sum_{(v, v'') \in \Ecal \setminus \{v', v'''\}}z^{(t-r')}_{v'' \rightarrow v'}$
-		\EndIf
-		\EndFor
-		\State $\sigma_v \gets \sum_{(v, v') \in \Ecal} y^{(T)}_{v' \rightarrow v}$ \Comment{sum final messages}
-		\State $S \gets \{v \in \Vcal : \sigma_v > 0\}$ \Comment{hard assignment}
-		\State $S^C \gets \{v \in \Vcal : \sigma_v \leq 0\}$
-		\State \textbf{return} $S, S^C$
-		\EndFunction
-	\end{algorithmic}
-\end{algorithm}
+We implement linearised ABP in Python and apply it to the Facebook egonets described earlier.
+We found that the hyperparameter settings $r=3$ and $T=5$ yielded good results.
+The algorithm returns the belief $\sigma_v$ associated with each vertex $v$.
+If this value is positive $\sigma_v > 0$ then we assign $v \in S$; $S^C$ contains all vertices with $\sigma_v \leq 0$.
+As well as the hard partition of $\Vcal$ into $S$ and $S^C$, the magnitude of $\sigma_v$ encodes how strongly we believe that this vertex belongs to its assigned partition.
+We can rescale the positive and negative components independently to obtain the normalised belief $\tilde{\sigma}_v$ such that $\tilde{\sigma}_v \in [-1, +1]$ and no signs are flipped by the rescaling.
 
-We implement this algorithm in Python and apply it to the Facebook egonets described earlier.
-We found that the hyperparameter settings $r=3$ and $T=5$ yielded good results.
-As well as the hard partition of $\Vcal$ into $S$ and $S^C$ we can use the normalised ABP output $\tilde{\sigma}_v$ defined below:
-%
-\begin{equation}
-	\tilde{\sigma}_v \coloneqq \begin{cases}
-		\frac{\sigma_v}{|\sigma_{max}|} \textrm{ for } \sigma_j > 0 \\
-		\frac{\sigma_v}{|\sigma_{min}|} \textrm{ for } \sigma_j \leq 0
-	\end{cases}
-\end{equation}
-
-Where $\sigma_{max}$ is the largest positive value of $\sigma_v$ and $\sigma_{min}$ is the most negative.
-The normalised quantity has restricted range $\tilde{\sigma}_v \in [-1, +1]$ with $0$ still being the decision boundary for partitioning.
 We colour each vertex according to this quantity $\tilde{\sigma}_v$ and plot the graph on figure \ref{fig:abp-output-norm}.
@@
 		\includegraphics[width=\linewidth]{ego-0-abp.png}
-		\caption{Normalised ABP output $\tilde{\sigma}_v$ colouring vertices}
+		\caption{Normalised ABP output $\tilde{\sigma}_v$ colouring vertices, spring layout}
 		\label{fig:abp-output-norm}
@@
 		\includegraphics[width=\linewidth]{ego-0-abp-by-gender.png}
-		\caption{Vertices grouped by hard ABP assignments and coloured by gender.
-In assigned community 1 ($S$ - left) we have \{58, 122\} gender-77 and gender-78 respectively and in detected community 2 ($S^C$ - right) we have \{72, 96\}}
+		\caption{Vertices grouped into set $S$ (left cluster) and set $S^C$ (right cluster)}
 		\label{fig:abp-output-by-gender}
@@
 
-We can evaluate formally whether weak recovery has succeeded by inspecting figure \ref{fig:abp-output-by-gender} and treating gender-77 and gender-78 as the two communities of interest:
+We can evaluate formally whether weak recovery has succeeded by inspecting figure \ref{fig:abp-output-by-gender} and treating gender-77 and gender-78 as the two communities of interest.
+Within set $S$ we have $\{58, 122\}$ 77-ers and 78-ers respectively and for $S^C$ the breakdown is $\{72, 96\}$.
+Therefore, the fraction of each gender cohort in set S can be calculated:
 %
@@
 
-Clearly the fraction of each community in the partition $S$ differ substantially so we say that weak recovery is solved.
+Clearly the fractions of each community in the partition $S$ differ substantially so we say that weak recovery is solved.
 The set $S$ contains a higher proportion of the overall 78-population than of the 77-population.
@@
 
-We perform a linear regression on the $n \times 1$ normalised ABP output vector $\tilde{\sigma}$ with the $n \times f$ feature matrix $X$ as the explanatory variable.
+We perform a linear regression on the $n \times 1$ normalised belief vector $\tilde{\sigma}$ with the $n \times f$ feature matrix $X$ as the explanatory variable.
 $X_{ij} = 1$ if vertex $i$ has feature $j$ turned on and $0$ if feature $j$ is off.
@@
 	\centering
-	\begin{subfigure}{\figwidth}
+	\begin{subfigure}{\linewidth}
 		\includegraphics[width=\linewidth]{ego-0-abp-community-fractions.png}
@@
 	\end{subfigure}
-	\begin{subfigure}{\figwidth}
+	\begin{subfigure}{\linewidth}
 		\includegraphics[width=\linewidth]{ego-0-abp-lin-regression.png}
-		\caption{Linear regression of $\tilde{\sigma}_v$ on select feature ($R^2 = 0.113$)}
+		\caption{Linear regression of $\tilde{\sigma}_v$ on select features ($R^2 = 0.113$)}
 		\label{fig:abp-output-lin-regression}

### 2021-01-19 interim-report/interim-report.tex
5.5 pages

@@
-\documentclass[11pt]{article}
+\documentclass[10pt]{article}
 
@@
 \end{equation}
-
+%
 We formulate this as a likelihood ratio test. Letting $\lik(D | H)$ denote the likelihood of observing the data $D = (X, \Gcal)$ under hypothesis $H$. Therefore, the test statistic is given by:
@@
 \end{equation}
-
+%
 At this point it helps to introduce some more notation.
@@
 \end{equation}
-
+%
 With this new notation, the likelihood function can be written explicitly:
@@
 \lik(D | H) &= Pr(X| p) \cdot Pr(\Gcal | W, X) \nonumber \\
-&= Pr(X | p) \cdot \prod_{i=1}^{k} \prod_{j=i}^{k} p(E_{ij} | W, X) \nonumber \\
-&= Pr(X | p) \cdot \prod_{i=1}^{k} \prod_{j=i}^{k} W_{ij} ^ {E_{ij}} \cdot \left( 1 - W_{ij} \right) ^ {(M_{ij} - E_{ij})}
+&= Pr(X | p) \prod_{i=1}^{k} \prod_{j=i}^{k} p(E_{ij} | W, X) \nonumber \\
+&= Pr(X | p) \prod_{i=1}^{k} \prod_{j=i}^{k} W_{ij} ^ {E_{ij}} \left( 1 - W_{ij} \right) ^ {(M_{ij} - E_{ij})}
 \label{eqn:likelihood-verbose}
 \end{align}
-
+%
 The form of $Pr(\Gcal | W, X)$ is simply a sequence of Bernoulli trials for each distinct community pair $(i, j)$; there are $E_{ij}$ edges present each with probability $W_{ij}$ and $(M_{ij} - E_{ij})$ absent with probability $(1 - W_{ij})$.
@@
 \end{equation}
-
+%
 We note that $f(w, e, m)$ is simply the probability of observing a specific sequence of $e$ successes in $m$ independent Bernoulli trials with parameter $w$.
@@
 \end{equation}
-
+%
 Furthermore, we spot the following property $f(w, e_1, m_1) \cdot f(w, e_2, m_2) = f(w, e_1 + e_2, m_1 + m_2)$ or in other words, the function $f$ is linear in its second and third arguments given the same first argument.
@@
 \end{align}
-
+%
 Where $\hat{p} \coloneqq E_\alpha / M_\alpha$, $\hat{q} \coloneqq E_\beta / M_\beta$ and $\hat{r} \coloneqq (E_\alpha + E_\beta) / (M_\alpha + M_\beta)$.
@@
 \begin{equation}
-	t_n = M_\alpha \kl\left( \textrm{Bern}(\hat{p}) || \textrm{Bern}(\hat{r})\right) + 
-	M_\beta \kl\left( \textrm{Bern}(\hat{q}) || \textrm{Bern}(\hat{r})\right)
+	\begin{aligned}
+	t_n = &M_\alpha \kl\left( \textrm{Bern}(\hat{p}) || \textrm{Bern}(\hat{r})\right) \\
+	&+ M_\beta \kl\left( \textrm{Bern}(\hat{q}) || \textrm{Bern}(\hat{r})\right)
+	\end{aligned}
 	\label{eqn:test-stat-compute-form}
 \end{equation}
-
 %
@@
 \begin{align*}
-	\kl(g||h) &= \sum_{x \in \Xcal} (\delta + h) \log \left(1 + \frac{\delta}{g} \right) \\
+	\kl(g||h) &= \sum_{x \in \Xcal} (\delta + h) \log \left(1 + \frac{\delta}{h} \right) \\
 	&= \sum_{x \in \Xcal} (\delta + h) \left( \frac{\delta}{h} - \frac{\delta^2}{2h^2} + O(\delta^3) \right)  \\
 	&= \sum_{x \in \Xcal} \delta + \frac{1}{2} \sum_{x \in \Xcal} \frac{\delta^2}{h} + O(\delta^3) \\
-	&= \frac{1}{2} \sum_{x \in \Xcal} \frac{\delta^2}{h} + O(\delta^3) \\
 	&= \frac{1}{2} \chi^2(g||h) + O(\delta^3) \approx \frac{1}{2} \chi^2(h||g) \quad \therefore \textrm{QED}
@@
 \end{equation}
-
+%
 We can apply these results to equation \ref{eqn:test-stat-compute-form} to obtain an approximate expression for $t_n$ under the null.
@@
 	\right) \nonumber \\
-	&= \frac{1}{2 \hat{r}(1-\hat{r})} \left( 
-	M_\alpha \left(\frac{M_\alpha (\hat{p} - \hat{q})}{M_\alpha + M_\beta} \right)^2 +
-	M_\beta \left(\frac{M_\beta (\hat{q} - \hat{p})}{M_\alpha + M_\beta} \right)^2
-	\right) \nonumber \\
 	&= \frac{1}{2} \left( \frac{\hat{p} - \hat{q}}{\sqrt{\hat{r}(1- \hat{r})(1/M_\alpha + 1/M_\beta)}} \right)^2
@@
 \end{align}
-
+%
 Through application of the Central Limit Theorem (CLT) and Slutsky's Theorem it can be shown that as $n \rightarrow \infty$ under the null $H_0$ the term in brackets on the final line is distributed like a standard Gaussian $z_n \sim \Gaussian(0,1)$.
@@
 	\end{align*}
-	If the log-likelihood ratio test statistic $t_n$ is computed as in equation \ref{eqn:test-stat-compute-form}, repeated here:
-	%
-	\begin{equation*}
-		t_n \coloneqq  M_\alpha \kl\left( \textrm{Bern}(\hat{p}) || \textrm{Bern}(\hat{r})\right) + 
-		M_\beta \kl\left( \textrm{Bern}(\hat{q}) || \textrm{Bern}(\hat{r})\right)
-	\end{equation*}
-
-	Where $\hat{p} \coloneqq E_\alpha / M_\alpha$, $\hat{q} \coloneqq E_\beta / M_\beta$ and $\hat{r} \coloneqq (E_\alpha + E_\beta) / (M_\alpha + M_\beta)$.
-Then as the number of vertices $n \rightarrow \infty, t_n \sim \frac{1}{2} \chi^2_1$ under the null $H_0$.
+	If the log-likelihood ratio test statistic $t_n$ is computed as in equation \ref{eqn:test-stat-compute-form}, then as the number of vertices $n \rightarrow \infty, t_n \sim \frac{1}{2} \chi^2_1$ under the null $H_0$.
 Therefore, we reject $H_0$ at the $100(\zeta)\%$ confidence level if and only if $2t_n \geq \psi^{-1}(\zeta)$, where $\psi^{-1}$ is the $\chi^2_1$ inverse cdf satisfying $Pr(Y \leq \psi^{-1}(\zeta)) = \zeta$ given $Y \sim \chi^2_1$.
@@
 
-Note that an alternative test is possible by comparing $z_n$ to a Gaussian.
+Indeed, a related test is possible by comparing $z_n$ to a standard Gaussian.
 
@@
 Each feature is given an anonymised signature to avoid disclosing personally identifiable information.
-We cannot tell whether vertex $v$ is male or female but we can tell whether they are the same gender as another vertex $w$ which suffices for our analysis.
-If we have a total of $f$ features and $n$ vertices, then the feature matrix $X$ would be an $n \times f$ where each row is the feature vector for the corresponding vertex.
-The features are binary such that $X_{ij} \in \{0, 1\}$ indicating the feature being turned off and on respectively.
+If we have a total of $f$ features and $n$ vertices, then the feature matrix $X$ would have shape $n \times f$ where each row is the feature vector for the corresponding vertex.
+Each feature is encoded with a binary flag such that $X_{ij} \in \{0, 1\}$ indicating feature turned off and on respectively.
 
@@
 \end{equation}
-
+%
 Therefore, $p_1$ is the probability that two vertices of gender-77 are connected, $q$ is the crossover probability and $p_2$ is the connection probability within the gender-78 community\footnote{Strictly speaking this is the not gender-77 community as there are a handful of vertices that do not disclose their gender to Facebook or do not fall into this binary}.
@@
 	\centering
-	\begin{subfigure}{\figwidth}
-		\includegraphics[width=\linewidth]{ego-0-abp.png}
+	\includegraphics[width=\figwidth]{ego-0-abp.png}
 	\caption{Normalised ABP output $\tilde{\sigma}_v$ colouring vertices, spring layout}
 	\label{fig:abp-output-norm}
-	\end{subfigure}
-	\begin{subfigure}{\figwidth}
-		\includegraphics[width=\linewidth]{ego-0-abp-by-gender.png}
-		\caption{Vertices grouped into set $S$ (left cluster) and set $S^C$ (right cluster)}
+\end{figure}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\figwidth]{ego-0-abp-by-gender.png}
+	\caption{Vertices grouped by ABP partition: set $S$ (left cluster) and set $S^C$ (right cluster)}
 	\label{fig:abp-output-by-gender}
-	\end{subfigure}
-	\caption{Egonet-0 weak recovery results}
 \end{figure}
@@
 This part of the project is still in its early stages so these are just preliminary results. For reference we plot the community fractions in the detected set $S$
-\footnote{ABP is non-deterministic so each subplot is the result of a different run of ABP. Subfigures use the same partition $S$ but no exact comparison is possible between figures.}
+\footnote{ABP is non-deterministic so each figure is the result of a different run of ABP. No exact comparison is possible between figures but the results are in some sense typical.}
 on figure \ref{fig:abp-output-community-fractions}.
@@
 \begin{figure}[H]
-	\centering
-	\begin{subfigure}{\linewidth}
-		\includegraphics[width=\linewidth]{ego-0-abp-community-fractions.png}
-		\caption{Community fractions $|\Omega_i \cap S| / |\Omega_i|$}
-		\label{fig:abp-output-community-fractions}
-	\end{subfigure}
-	\begin{subfigure}{\linewidth}
 	\includegraphics[width=\linewidth]{ego-0-abp-lin-regression.png}
-		\caption{Linear regression of $\tilde{\sigma}_v$ on select features ($R^2 = 0.113$)}
+	\caption{Egonet-0 linear regression of $\tilde{\sigma}_v$ on select features ($R^2 = 0.113$)}
 	\label{fig:abp-output-lin-regression}
-	\end{subfigure}
-	\caption{Egonet-0 linear classifier results}
 \end{figure}
@@
 The bias term has rather low magnitude which just implies that the partition $S$ and $S^C$ are of similar size.
-A keen reader would notice that the bias term is included in figure \ref{fig:abp-output-community-fractions}.
-This is because we implement the bias by augmenting $X$ with an additional feature flag always set to $1$ such that the constant term $a$ in equation \ref{eqn:lin-regression} is subsumed into the $b$ vector.
-Since all vertices have the bias feature turned on $\Omega_{bias} = \Vcal$, the fraction of vertices with bias turned on in $S$ is equivalent to the total number of vertices in $S$ as a fraction of the total in $\Vcal$, or put formally: $|S| / |\Vcal|$.
-Indeed, in figure \ref{fig:abp-output-community-fractions} we see that this quantity is very close to 0.5.
 

### 2021-01-20 interim-report/interim-report.tex
6-page

@@
-\documentclass[10pt]{article}
+\documentclass[11pt]{article}
 
@@
 \newcommand{\one}{\mathbbm{1}}
+\newcommand{\prob}{\mathbbm{P}}
 \newcommand{\figwidth}{0.7\linewidth}
@@
 We produce vast quantities of graphical data every day and yet the techniques used to analyse their structure are still in their infancy.
-We start with the definition of the Stochastic Block Model (SBM) widely used in academia and develop a theorem to verify structure in a labelled graph through a hypothesis test.
+We begin with the definition of the Stochastic Block Model (SBM) widely used in academia and develop a theorem to verify structure in a labelled graph through a hypothesis test.
 We explore the results of applying this theorem to Facebook friendship egonets and find that gender does indeed influence how friendships form on Facebook.
@@
 To rank order vertex labels by importance we instead approach the problem from the opposite direction: given a graph we detect the partitions that are in some sense typical and then develop a classifier to map from vertex features to partition.
-The classifier then implicitly rank orders the input features by importance.
+The classifier then implicitly rank orders the input features automatically.
 Nevertheless, work must still be done to refine this classifier and tie it back to the rigorous hypothesis testing framework already developed.
@@
 	Let $n \in \Integers^+$ be the number of vertices and $k \in \Integers^+$ be the number of communities in an SBM graph.
-We define a probability vector $\pi = [\pi_1, \pi_2 \dots \pi_k]^T$ to be the prior on the k-communities.
-Each vertex $v \in \Vcal = \{1, 2 \dots N\}$ has a community label $X_v \in \{1, 2 \dots n\}$.
+We define a probability vector $p = [p_1, p_2 \dots p_k]^T$ to be the prior on the k-communities.
+Each vertex $v \in \Vcal = \{1, 2 \dots n\}$ has a community label $X_v \in \{1, 2 \dots n\}$.
 Let $W$ be a symmetric $k \times k$ matrix with entries in $[0,1]$ called the connectivity matrix.
-We say that the pair $(X, \Gcal) \sim \textrm{SBM}(n, \pi, W)$ if X is an $N$-dimensional vector with each component independently distributed as the community prior $X_v \sim \pi$ and $\Gcal$ is an $N$-vertex graph where each pair of vertices $(i, j)$ is connected with probability $p(i \leftrightarrow j) = W_{X_i, X_j}$ independently of other pairs of vertices.
+We say that the pair $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$ if X is an $n$-dimensional vector with each component independently distributed as the community prior $X_v \sim p$ and $\Gcal$ is an $n$-vertex graph where each pair of vertices $(i, j)$ is connected with probability $\prob(i \leftrightarrow j) = W_{X_i, X_j}$ independently of other pairs of vertices.
 Lastly, we define the community sets as $\Omega_i = \Omega_i(X) \coloneqq \{v \in \Vcal : X_v = i\}$ which contains all vertices belonging to community $i$.
@@
 Armed with definition \ref{defn:sbm} we tackle the simplest problem in structure verification.
-Given an undirected graph $\Gcal$ and vertex-labels $X$ such that $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$, we wish to determine whether the two communities $a$ and $b$ connect differently.
+Given an undirected graph $\Gcal$ and vertex-labels $X$ such that $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$, we wish to determine whether two communities $a$ and $b$ connect differently.
 Put formally, this is a hypothesis test on the parameters of $W$.
@@
 To do this we can perform three-pairwise hypothesis tests.
-Here we test $W_{\alpha}$ against $W_{\beta}$ where $\alpha$ and $\beta$ are unique indices in $\{(a,a), (a, b), (b,b)\}$.
+Here we test $W_{\alpha}$ against $W_{\beta}$ where $\alpha$ and $\beta$ are unique index pairs in $\{(a,a), (a, b), (b,b)\}$.
 Therefore, each hypothesis test can be formulated as below:
@@
 \begin{align}
-\lik(D | H) &= Pr(X| p) \cdot Pr(\Gcal | W, X) \nonumber \\
-&= Pr(X | p) \prod_{i=1}^{k} \prod_{j=i}^{k} p(E_{ij} | W, X) \nonumber \\
-&= Pr(X | p) \prod_{i=1}^{k} \prod_{j=i}^{k} W_{ij} ^ {E_{ij}} \left( 1 - W_{ij} \right) ^ {(M_{ij} - E_{ij})}
+\lik(D | H) &= \prob(X| p) \cdot \prob(\Gcal | W, X) \nonumber \\
+&= \prob(X | p) \prod_{i=1}^{k} \prod_{j=i}^{k} \prob(E_{ij} | W, X) \nonumber \\
+&= \prob(X | p) \prod_{i=1}^{k} \prod_{j=i}^{k} W_{ij} ^ {E_{ij}} \left( 1 - W_{ij} \right) ^ {(M_{ij} - E_{ij})}
 \label{eqn:likelihood-verbose}
@@
 %
-The form of $Pr(\Gcal | W, X)$ is simply a sequence of Bernoulli trials for each distinct community pair $(i, j)$; there are $E_{ij}$ edges present each with probability $W_{ij}$ and $(M_{ij} - E_{ij})$ absent with probability $(1 - W_{ij})$.
-A sequence of Bernoullis is the same as a Binomial distribution without the combinatoric term.
 By inspecting equation \ref{eqn:likelihood-verbose} we see that only terms involving $W_{\alpha}$ and $W_{\beta}$ are going to differ under the two hypotheses $H_0$ and $H_1$; the rest of the terms will cancel in our calculation of the test-statistic $t_n$.
@@
 Where $\hat{p} \coloneqq E_\alpha / M_\alpha$, $\hat{q} \coloneqq E_\beta / M_\beta$ and $\hat{r} \coloneqq (E_\alpha + E_\beta) / (M_\alpha + M_\beta)$.
-These are essentially maximum-likelihood estimates of parameters of $W$ and the symbols are introduced to make the notation more succinct.
-Note that the max-likelihood estimate of $W_alpha$ under $H_1$ is denoted by the symbol $\hat{p}$ and should not be confused with the community prior $p$ though context will make the distinction apparent.
+These are essentially maximum-likelihood estimates of parameters of $W$ constrained by the corresponding hypothesis.
+Note that the max-likelihood estimate of $W_\alpha$ under $H_1$ is denoted by the symbol $\hat{p}$ and should not be confused with the community prior $p$ though context will make the distinction apparent.
 We now state lemma \ref{lem:kl-div}.
@@
 Proving lemma \ref{lem:kl-div} is simply a case of algebraic manipulation and is omitted here.
-However, this allows us to simplify the test-statistic into a form that is more numerically stable (equation \ref{eqn:test-stat-compute-form}).
+However, this allows us to simplify the test-statistic into a form that is easier to compute (equation \ref{eqn:test-stat-compute-form}).
 However, we must simplify further still using lemma \ref{lem:kl-approx-chi}.
@@
 
-The proof requires defining $\delta(x) \coloneqq g(x) - h(x)$ and taking a Taylor expansion of $\log 1 + \epsilon$:
-%
-\begin{align*}
-	\kl(g||h) &= \sum_{x \in \Xcal} (\delta + h) \log \left(1 + \frac{\delta}{h} \right) \\
-	&= \sum_{x \in \Xcal} (\delta + h) \left( \frac{\delta}{h} - \frac{\delta^2}{2h^2} + O(\delta^3) \right)  \\
-	&= \sum_{x \in \Xcal} \delta + \frac{1}{2} \sum_{x \in \Xcal} \frac{\delta^2}{h} + O(\delta^3) \\
-	&= \frac{1}{2} \chi^2(g||h) + O(\delta^3) \approx \frac{1}{2} \chi^2(h||g) \quad \therefore \textrm{QED}
-\end{align*}
-%
-Where the summation over $\delta$ evaluates to 0 because $\delta$ is the difference of two valid p.m.f's which each sum to 1 over $x \in \Xcal$.
-We are able to neglect the $O(\delta^3)$ terms for $g$ very close to $h$.
-The chi-squared distance between two Bernoullis for $g = Bern(p)$ and $h = Bern(q)$ is simply:
+The proof requires defining $\delta(x) \coloneqq g(x) - h(x)$ and taking a Taylor expansion of $\log (1 + \delta / h)$ - omitted for bervity.
+The chi-squared distance between two Bernoullis $g = Bern(p)$ and $h = Bern(q)$ has a very simple form:
 %
@@
 %
-Through application of the Central Limit Theorem (CLT) and Slutsky's Theorem it can be shown that as $n \rightarrow \infty$ under the null $H_0$ the term in brackets on the final line is distributed like a standard Gaussian $z_n \sim \Gaussian(0,1)$.
-This is because under the same assumptions, assuming that $W_\alpha = W_\beta = \mu$, by the CLT $\hat{p} \sim (\mu, \mu(1 - \mu)/ M_\alpha)$, $\hat{q} \sim (\mu, \mu(1 - \mu)/ M_\beta)$ and $\hat{r} \sim (\mu, \mu(1 - \mu)/ (M_\alpha + M_\beta))$.
+Through application of the Central Limit Theorem (CLT) and Slutsky's Theorem it can be shown that as $n \rightarrow \infty$ under the null, the term in brackets on the final line is distributed like a standard Gaussian $z_n \sim \Gaussian(0,1)$.
 The proof is long-winded and therefore omitted from this short report.
@@
 \begin{theorem}[$\chi^2$ test]
-	For $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$, given the realised graph and class labels $(X, \Gcal)$ we can perform a hypothesis test on parameters $W_\alpha$ and $W_\beta$ of the connectivity matrix $W$.
+	For $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$, given the realised graph and class labels $(X, \Gcal)$ we can perform a hypothesis test on entries $W_\alpha$ and $W_\beta$ of the connectivity matrix $W$.
 	%
@@
 	If the log-likelihood ratio test statistic $t_n$ is computed as in equation \ref{eqn:test-stat-compute-form}, then as the number of vertices $n \rightarrow \infty, t_n \sim \frac{1}{2} \chi^2_1$ under the null $H_0$.
-Therefore, we reject $H_0$ at the $100(\zeta)\%$ confidence level if and only if $2t_n \geq \psi^{-1}(\zeta)$, where $\psi^{-1}$ is the $\chi^2_1$ inverse cdf satisfying $Pr(Y \leq \psi^{-1}(\zeta)) = \zeta$ given $Y \sim \chi^2_1$.
+Therefore, we reject $H_0$ at the $100(\zeta)\%$ confidence level if and only if $2t_n \geq \psi^{-1}(\zeta)$, where $\psi^{-1}$ is the $\chi^2_1$ inverse cdf satisfying $\prob(Y \leq \psi^{-1}(\zeta)) = \zeta$ given $Y \sim \chi^2_1$.
 
@@
 
-Indeed, a related test is possible by comparing $z_n$ to a standard Gaussian.
+Indeed, a corollary is the hypothesis test possible by comparing $z_n$ to a standard Gaussian.
 
@@
 We start by analysing social network graphs.
-The Stanford Network Analysis Project (SNAP for short) \cite{snapnets} offers a wealth of Facebook egonets.
-An egonet is simply a graph where all vertices (in this case each representing a Facebook user) are guaranteed to be connected to one central node (the ego-node).
-The data consists of the undirected set of edges $\Gcal$ indicating whether any two vertices (Facebook users) are connected (friends on Facebook).
-We also have a set of binary feature flags for each vertex $X$.
-However, for the sake of privacy these features are anonymised.
-This is best explained through the example below: \\
+The Stanford Network Analysis Project (SNAP) \cite{snapnets} offers a wealth of Facebook egonets.
+An egonet is simply a graph where all vertices (in this case each representing a Facebook user) are connected to one central node (the ego-node).
+The data consists of the undirected set of edges $\Gcal$ indicating whether any two vertices (Facebook users) are connected (friends on Facebook) and matrix $F$ containing binary feature flags for each vertex.
+Example feature flags are given below: \\
 
@@
 Each feature is given an anonymised signature to avoid disclosing personally identifiable information.
-If we have a total of $f$ features and $n$ vertices, then the feature matrix $X$ would have shape $n \times f$ where each row is the feature vector for the corresponding vertex.
-Each feature is encoded with a binary flag such that $X_{ij} \in \{0, 1\}$ indicating feature turned off and on respectively.
+If we have a total of $f$ features and $n$ vertices, then the feature matrix $F$ would have shape $n \times f$ where each row is the feature vector for the corresponding vertex.
+Each feature is encoded with a binary flag such that $F_{ij} \in \{0, 1\}$ indicating feature turned off and on respectively.
 
 We perform a hypothesis test in the manner described by theorem \ref{theorem:hyp-test-sbm-chi} to determine whether gender influences how friends connect on Facebook.
-We choose to analyse SNAP egonet with id 0 (the id of the egonet is the id of the single egonode) though any choice is possible.
+We choose to analyse SNAP egonet with id 0 (the id of the egonet is the id of the single egonode).
 The egonet is plotted on figure \ref{fig:ego-0-by-gender} with nodes coloured by gender.
-We use the Python package NetworkX \cite{networkx} for its visualisation tools.
-Indeed, we will discuss layout algorithms later as for now we focus on the simple hypothesis test.
+We use the Python package NetworkX \cite{networkx} for its graph visualisation tools.
+Indeed, we will discuss layout algorithms later as for now we focus on the hypothesis test.
 %
@@
 %
-Therefore, $p_1$ is the probability that two vertices of gender-77 are connected, $q$ is the crossover probability and $p_2$ is the connection probability within the gender-78 community\footnote{Strictly speaking this is the not gender-77 community as there are a handful of vertices that do not disclose their gender to Facebook or do not fall into this binary}.
+Therefore, $p_1$ is the probability that two vertices of gender-77 are connected, $q$ is the crossover probability and $p_2$ is the connection probability within the gender-78 community
+\footnote{Strictly speaking this is the not gender-77 community as there are a handful of vertices that do not disclose their gender to Facebook or do not fall into this binary}
+.
 The results of the hypothesis tests are given in table \ref{tab:egonet-0-hyp-tests} alongside p-values (we choose a 95\% significance level to reject the null).
@@
 The rejection of the null in test 3 gives evidence that gender-78 treat gender-77 people differently to their own gender.
-Indeed, a gender-78 person is more likely to be friends with a fellow 78-er than with a 77-er.
-However, we do not reject the null in test 1 so there is insufficient evidence to claim the reverse (that 77-ers tend to stay away from 78-ers).
+Indeed, a gender-78 person is more likely to be friends with a fellow 78-er than with a 77-er ($\hat{p}_2 > \hat{q}$).
+
+However, we do not reject the null in test 1 as there is insufficient evidence to claim the reverse (that 77-ers tend to stay away from 78-ers).
 Nevertheless, these are results on a single egonet so can hardly be generalised to society as a whole.
@@
 Community detection is the problem of recovering the communities from the graph $\Gcal$.
-We use the notation $\hat{X}$ to denote the estimate of the community labels for the $n$ vertices.
-Obviously, if the SBM is symmetric then we must allow for an arbitrary relabelling of the communities $r: \{1, 2 \dots k\} \mapsto \{1, 2 \dots k\}$ before we compare the agreement of the two vectors $\hat{X}$ and $X$.
 There are 4 recovery regimes (plus the trivial no recovery), as defined by Abbe \cite{Abbe}.
@@
 Therefore, there are very few constraints on the SBM parameters and the theory can be applied to a wide range of graphs.
-A statement of weak recovery is given in definition \ref{defn:weak-recovery} \cite{Abbe}.
+A statement of weak recovery is given in definition \ref{defn:weak-recovery} (adapated from \cite{Abbe}).
 
@@
 	\label{defn:weak-recovery}
-	weak recovery is solved for a graph  $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$ if there exists $\epsilon > 0$ and a certain choice of indices $i, j \in \{1, 2 \dots k\}$ and an algorithm that takes as input $\Gcal$ and outputs a partition of the vertex set $\Vcal$ into two distinct sets $S$ and $S^C$ (detected communities 1 and 2 respectively) such that:
+	weak recovery is solved for a graph  $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$ if there exists $\epsilon > 0$, a certain choice of indices $i, j \in \{1, 2 \dots k\}$ and an algorithm that takes as input $\Gcal$ and outputs a partition of the vertex set $\Vcal$ into two distinct sets $S$ and $S^C$ (detected communities 1 and 2 respectively) such that:
 	\begin{equation*}
-		Pr\left\{ \frac{|\Omega_i \cap S|}{|\Omega_i|} - \frac{|\Omega_j \cap S|}{|\Omega_j|} \geq \epsilon \right\} = 1 - o(1)
+		\prob \left\{ \frac{|\Omega_i \cap S|}{|\Omega_i|} - \frac{|\Omega_j \cap S|}{|\Omega_j|} \geq \epsilon \right\} = 1 - o(1)
 	\end{equation*}
 
-	Bearing in mind the definition of $\Omega_i = \Omega_i(X) \coloneqq \{v \in \Vcal : X_v = i\}$.
+	Where $o(1)$ denotes a family of functions that tend to $0$ as $n$ gets large.
+We recall the definition of $\Omega_i = \Omega_i(X) \coloneqq \{v \in \Vcal : X_v = i\}$.
 In other words, we require the partition $S \subset \Vcal$ to cut across communities such that the fraction of a particular community in partition $S$ with respect to the whole community is different for a choice of two communities.
@@
 Nevertheless, the premise is that one can infer the community of a particular vertex simply from the labels of all its neighbours.
-The algorithm is non-deterministic and requires the vertex set $\Vcal = \Vcal(\Gcal)$ and edge set $\Ecal = \Ecal(\Gcal)$ as well as two hyper-parameters $r$ (the maximum cyclic length to correct for) and $T$ (the number of iterations).
+The algorithm is non-deterministic and requires the graph $\Gcal$ as well as two hyper-parameters $r$ (the maximum cyclic length to correct for) and $T$ (the number of iterations).
 
@@
 
-We colour each vertex according to this quantity $\tilde{\sigma}_v$ and plot the graph on figure \ref{fig:abp-output-norm}.
+We colour each vertex according to this quantity $\tilde{\sigma}_v$ and plot the graph
+\footnote{ABP is non-deterministic so each figure is the result of a different run of ABP. No exact comparison is possible between figures but the results are in some sense typical.}
+on figure \ref{fig:abp-output-norm}.
 We see intuitively that ABP has worked.
@@
 Furthermore, many vertices that do not seem to belong to any cluster have uncertain classifications (the colour is close to green).
+Inspecting \ref{fig:ego-0-by-gender} again shows us that the sets $S$ and $S^C$ to separate nicely with high intra-group connectivity and lower inter-group connection density.
 %
@@
 
-We can evaluate formally whether weak recovery has succeeded by inspecting figure \ref{fig:abp-output-by-gender} and treating gender-77 and gender-78 as the two communities of interest.
-Within set $S$ we have $\{58, 122\}$ 77-ers and 78-ers respectively and for $S^C$ the breakdown is $\{72, 96\}$.
-Therefore, the fraction of each gender cohort in set S can be calculated:
+Within set $S$ we have $\{58, 122\}$ 77-ers and 78-ers respectively and for $S^C$ the breakdown is $\{72, 96\}$. Therefore, the fraction of each gender cohort in set S can be calculated:
 %
@@
 
-We now have an algorithm (ABP) that partitions our graph into two sets that are most readily explained by an SBM with $k=2$ communities.
+We now have an algorithm (ABP) that partitions our graph into two sets that are maximally distinct in a graphical sense.
 We can now ask the question of which vertex labels most readily explain the separation and therefore determine which features have the largest impact on the graphical structure; this circumvents the binary resolution problem with the hypothesis testing approach.
 
-This part of the project is still in its early stages so these are just preliminary results. For reference we plot the community fractions in the detected set $S$
-\footnote{ABP is non-deterministic so each figure is the result of a different run of ABP. No exact comparison is possible between figures but the results are in some sense typical.}
-on figure \ref{fig:abp-output-community-fractions}.
-It is interesting to note that if two bars on this plot differ substantially then we have strong evidence that we have solved weak recovery as in definition \ref{defn:weak-recovery}.
-In a slight abuse of notation, we necessarily remove the constraint on the community sets $\Omega_i$ to be disjoint as one vertex can have multiple features turned on.
+This part of the project is still in its early stages so these are just preliminary results.
+We necessarily remove the constraint on the community sets $\Omega_i$ to be disjoint as one vertex can have multiple features turned on.
 These should no longer be thought of as community sets but rather feature sets.
-This can be modelled more precisely through a Mixed Membership Model but for now an SBM with $k=2$ communities suffices.
 
-We perform a linear regression on the $n \times 1$ normalised belief vector $\tilde{\sigma}$ with the $n \times f$ feature matrix $X$ as the explanatory variable.
+We perform a linear regression on the $n \times 1$ normalised belief vector $\tilde{\sigma}$ with the $n \times f$ feature matrix $F$ as the explanatory variable.
 $X_{ij} = 1$ if vertex $i$ has feature $j$ turned on and $0$ if feature $j$ is off.

### 2021-01-20 interim-report/interim-report.tex
2nd-draft

@@
 \title{TMR: Detecting Structure in Graphical Data}
-\author{Lawrence Tray \\ Sup: Ioannis Kontoyiannis}
+\author{Lawrence Tray \and Ioannis Kontoyiannis}
 
 %packages
-\usepackage[margin=0.5in]{geometry}
+\usepackage[margin=0.6in]{geometry}
 \usepackage{graphicx}
@@
 \usepackage{multicol}
+\usepackage{titling}
 
@@
 \addbibresource{interim-sources.bib}
+
+%title
+\setlength{\droptitle}{-4em}   % This is your set screw
 
@@
 
-There is a wealth of graphical data in the world and more is being produced each second; social networks, website hyperlinks and academic collaborations are just some examples.
-There are many algorithms developed to analyse graphical data.
-Nevertheless, that same principled hypothesis testing framework we have for querying classical data is less well-developed for graphical data.
+There is a wealth of graphical data in the world with more produced every second; social networks, website hyperlinks and academic collaboration are just some examples.
+Many algorithms are used to analyse this data.
+Nevertheless, the same principled hypothesis testing framework we have for querying classical data is less well-developed for graphical data.
 Do my friends vote the same way I do or do researchers collaborate with those of the same gender?
 We want to answer these questions and not only that, we wish to report our confidence in the answers.
-To that end there is space to expand the hypothesis testing framework to graphs.
+To that end we seek to expand the hypothesis testing framework to graphs.
 
@@
 
-The most popular graphical model in industry and indeed academia is called the Stochastic Block Model (SBM). We use a definition adapted from Abbe \cite{Abbe}.
+The most popular graphical model in academia is called the Stochastic Block Model (SBM). We use a definition adapted from Abbe \cite{Abbe}.
 
@@
 	\label{defn:sbm}
-	Let $n \in \Integers^+$ be the number of vertices and $k \in \Integers^+$ be the number of communities in an SBM graph.
-We define a probability vector $p = [p_1, p_2 \dots p_k]^T$ to be the prior on the k-communities.
+	Let $n \in \Integers^+$ be the number of vertices and $k \in \Integers^+$ be the number of communities.
+We define the vector $p = [p_1, p_2 \dots p_k]^T$ to be the prior on the k-communities.
 Each vertex $v \in \Vcal = \{1, 2 \dots n\}$ has a community label $X_v \in \{1, 2 \dots n\}$.
 Let $W$ be a symmetric $k \times k$ matrix with entries in $[0,1]$ called the connectivity matrix.
-We say that the pair $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$ if X is an $n$-dimensional vector with each component independently distributed as the community prior $X_v \sim p$ and $\Gcal$ is an $n$-vertex graph where each pair of vertices $(i, j)$ is connected with probability $\prob(i \leftrightarrow j) = W_{X_i, X_j}$ independently of other pairs of vertices.
+The pair $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$ if X is an $n \times 1$ vector with each component independently drawn from the community prior $X_v \sim p$ and $\Gcal$ is an $n$-vertex graph where each vertex-pair $(i, j)$ is connected with probability $\prob(i \leftrightarrow j) = W_{X_i, X_j}$ independently of other edges.
 Lastly, we define the community sets as $\Omega_i = \Omega_i(X) \coloneqq \{v \in \Vcal : X_v = i\}$ which contains all vertices belonging to community $i$.
 \end{definition}
-
-Though the definition of the SBM is simple, it allows for very deep and rich analysis of graphical datasets.
 
@@
 The Stanford Network Analysis Project (SNAP) \cite{snapnets} offers a wealth of Facebook egonets.
-An egonet is simply a graph where all vertices (in this case each representing a Facebook user) are connected to one central node (the ego-node).
-The data consists of the undirected set of edges $\Gcal$ indicating whether any two vertices (Facebook users) are connected (friends on Facebook) and matrix $F$ containing binary feature flags for each vertex.
+An egonet is simply a graph where all vertices (in this case Facebook users) are connected to one central node (the ego-node).
+The data consists of the undirected set of edges $\Gcal$ indicating whether any two vertices are connected (friends on Facebook) and matrix $F$ containing binary feature flags for each vertex.
 Example feature flags are given below: \\
@@
 \end{table}
-
+%
 \begin{table}[H]
@@
 	\end{tabular}
-	\caption{Egonet-0 hypothesis tests that gender influences friendship formation}
+	\caption{Egonet-0 hypothesis test results}
 	\label{tab:egonet-0-hyp-tests}
@@
 We can rescale the positive and negative components independently to obtain the normalised belief $\tilde{\sigma}_v$ such that $\tilde{\sigma}_v \in [-1, +1]$ and no signs are flipped by the rescaling.
+%
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\figwidth]{ego-0-abp.png}
+	\caption{Normalised ABP output $\tilde{\sigma}_v$ colouring vertices, spring layout}
+	\label{fig:abp-output-norm}
+\end{figure}
 
@@
 Furthermore, many vertices that do not seem to belong to any cluster have uncertain classifications (the colour is close to green).
-Inspecting \ref{fig:ego-0-by-gender} again shows us that the sets $S$ and $S^C$ to separate nicely with high intra-group connectivity and lower inter-group connection density.
+Inspecting figure \ref{fig:ego-0-by-gender} shows us that the sets $S$ and $S^C$ do separate nicely with high intra-group connectivity and lower inter-group connection density.
+After all, we seek to find a roughly balanced partition that minimises the number of edges between the two sets.
 %
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=\figwidth]{ego-0-abp.png}
-	\caption{Normalised ABP output $\tilde{\sigma}_v$ colouring vertices, spring layout}
-	\label{fig:abp-output-norm}
-\end{figure}
 \begin{figure}[H]
@@
 \end{equation}
-
+%
 Clearly the fractions of each community in the partition $S$ differ substantially so we say that weak recovery is solved.
@@
 
-We now have an algorithm (ABP) that partitions our graph into two sets that are maximally distinct in a graphical sense.
-We can now ask the question of which vertex labels most readily explain the separation and therefore determine which features have the largest impact on the graphical structure; this circumvents the binary resolution problem with the hypothesis testing approach.
+We now have an algorithm (ABP) that partitions our graph into two sets that are maximally distinct from a clustering sense.
+We can now ask the question of which vertex labels most readily explain the separation.
+Therefore, we can determine which features have the largest impact on the graphical structure; this circumvents the binary resolution problem with the hypothesis testing approach.
 
@@
 We necessarily remove the constraint on the community sets $\Omega_i$ to be disjoint as one vertex can have multiple features turned on.
-These should no longer be thought of as community sets but rather feature sets.
-
 We perform a linear regression on the $n \times 1$ normalised belief vector $\tilde{\sigma}$ with the $n \times f$ feature matrix $F$ as the explanatory variable.
-$X_{ij} = 1$ if vertex $i$ has feature $j$ turned on and $0$ if feature $j$ is off.
+$F_{ij} = 1$ if vertex $i$ has feature $j$ turned on and $0$ if feature $j$ is off.
 The equation we try to fit is given below:
@@
 \begin{equation}
-	\tilde{\sigma} = a + X b + \epsilon
+	\tilde{\sigma} = a + F b + \epsilon
 	\label{eqn:lin-regression}
 \end{equation}
-
+%
 Linear regression finds the constant bias $a$ and $f \times 1$ weight vector $b$ that minimises the mean squared prediction error $\epsilon^T \epsilon / n$.
 The theory for linear regression is well known so omitted here (see \cite{yale-linear-regression} for a reference).
-We perform a linear regression using only gender and language features available in the dataset and plot the parameter coefficient in order of decreasing magnitude (figure \ref{fig:abp-output-lin-regression}).
+We restrict $F$ to only include gender and language features and plot the regression coefficient for each feature in order of decreasing magnitude (figure \ref{fig:abp-output-lin-regression}).
 %
@@
 After all, one would hope that a language barrier poses a more tangible obstacle to friendship than gender.
-Furthermore, the gender coefficients are of the same sign further decreasing the causal impact that gender has given language.
 The bias term has rather low magnitude which just implies that the partition $S$ and $S^C$ are of similar size.
@@
 Nevertheless, this serves as the platform for subsequent more rigorous analysis.
-Furthermore, we have not proven the conjecture that the regression coefficients are valid measures of causal impact on graphical structure.
+Furthermore, we have not yet proven the conjecture that the regression coefficients are valid measures of causal impact on graphical structure.
 
@@
 
-Indeed, ABP is severely limited by the fact that it only outputs 2 partitions $S$ and $S^C$.
-Therefore, though weak recovery is solved the results from applying the classification approach may not be natural or easy to interpret.
-We instead would prefer to investigate various different algorithms.
+Indeed, ABP is severely limited by the fact that it only outputs 2 sets.
+Though weak recovery is solved, the partition is often hard to interpret as various clusters get subsumed into one.
+We instead would prefer to investigate algorithms that can identify arbitrarily many clusters.
 Force-directed approaches are the standard in industry for their speed of execution.
-One of the most common is the Fruchterman-Reingold algorithm which treats vertices as repelling objects and edges as attractive forces (springs).
-It can then caluclate forces and perform iterative updates on vertex positions for a finite number of iterations.
-This algorithm is what is used by the NetworkX package \cite{networkx} to determine graph layouts as seen in figures \ref{fig:ego-0-by-gender} and \ref{fig:abp-output-norm}.
+One of the most common is the Fruchterman-Reingold algorithm which treats vertices as repelling objects and edges as restoring forces.
+It can then perform iterative updates on vertex positions until convergence or termination.
+This algorithm is what is used by the NetworkX package \cite{networkx} to compute graph layouts as seen in figures \ref{fig:ego-0-by-gender} and \ref{fig:abp-output-norm}.
 Song and Bressan \cite{force-directed} have had great success using the Fruchterman-Reingold algorithm followed by k-means clustering for community detection.
-Nevertheless, though they show the method produces accurate results there is no rigorous proof yet that it solves weak recovery.
 
-Lastly, the analysis must of course be expanded to more datasets to benchmark efficacy against other methods and ensure the methods developed are generally applicable.
-Work has already begun analysing the influence of gender academic collaboration graphs from Aminer \cite{aminer}.
+Lastly, the analysis must of course be expanded to more datasets to benchmark efficacy against existing methods and ensure the approach is generally applicable.
+Work has already begun analysing the influence of gender on academic collaboration graphs provided by Aminer \cite{aminer}.
 Nevertheless, this dataset does not contain gender explicitly, only first-name.
@@
 Nevertheless, this approach is too narrow in scope to be a useful tool on its own.
-We therefore set about the problem from the opposite direction; rather than verify structure we are already given, we detect structure in the graph without labels and seek to identify which labels best recreate the detected partition.
+We therefore set about the problem from the opposite direction; rather than verify structure we are already given, we detect structure in the graph without using labels and seek to identify which labels best recreate the detected partition.
 Early results are promising but this approach needs to be given the same rigorous treatment as the hypothesis testing framework.

### 2021-01-20 interim-report/interim-report.tex
Final draft

@@
 \newcommand{\prob}{\mathbbm{P}}
-\newcommand{\figwidth}{0.7\linewidth}
+\newcommand{\figwidth}{0.8\linewidth}
 
@@
 We produce vast quantities of graphical data every day and yet the techniques used to analyse their structure are still in their infancy.
-We begin with the definition of the Stochastic Block Model (SBM) widely used in academia and develop a theorem to verify structure in a labelled graph through a hypothesis test.
-We explore the results of applying this theorem to Facebook friendship egonets and find that gender does indeed influence how friendships form on Facebook.
+We begin with the Stochastic Block Model (SBM) widely used in academia and derive a theorem to verify structure in a labelled graph through a hypothesis test.
+We apply this theorem to Facebook egonets and find that gender does indeed influence how friendships form on Facebook.
 Nevertheless, this binary yes-no result is unsatisfactory.
 
-To rank order vertex labels by importance we instead approach the problem from the opposite direction: given a graph we detect the partitions that are in some sense typical and then develop a classifier to map from vertex features to partition.
-The classifier then implicitly rank orders the input features automatically.
+We instead approach the problem from the opposite direction: given a graph we detect the partitions that are in some sense typical and then develop a classifier to map from vertex features to partition.
+The classifier then implicitly rank orders the input features by importance.
 Nevertheless, work must still be done to refine this classifier and tie it back to the rigorous hypothesis testing framework already developed.
@@
 Furthermore, we spot the following property $f(w, e_1, m_1) \cdot f(w, e_2, m_2) = f(w, e_1 + e_2, m_1 + m_2)$ or in other words, the function $f$ is linear in its second and third arguments given the same first argument.
-As such we can manipulate equation \ref{eqn:test-statistic-start} greatly to give:
+As such, we can manipulate equation \ref{eqn:test-statistic-start} greatly to give:
 %
@@
 Where $\hat{p} \coloneqq E_\alpha / M_\alpha$, $\hat{q} \coloneqq E_\beta / M_\beta$ and $\hat{r} \coloneqq (E_\alpha + E_\beta) / (M_\alpha + M_\beta)$.
-These are essentially maximum-likelihood estimates of parameters of $W$ constrained by the corresponding hypothesis.
+These are essentially max-likelihood estimates of the parameters of $W$ constrained by the corresponding hypothesis.
 Note that the max-likelihood estimate of $W_\alpha$ under $H_1$ is denoted by the symbol $\hat{p}$ and should not be confused with the community prior $p$ though context will make the distinction apparent.
@@
 
-	where $\kl(g || h)$ is the Kullback-Leibler divergence between two probability mass functions $g, h: \Xcal \mapsto [0, 1]$.
-$\kl$ is defined in discrete space as $\kl(g || h) \coloneqq \sum_{x \in \Xcal} g(x) \log \frac{g(x)}{h(x)}$ and $\textrm{Bern(p)}$ denotes the Bernoulli p.m.f with parameter $p$.
+	where $\kl(g || h)$ is the Kullback-Leibler divergence between two probability mass functions (pmf's) $g, h: \Xcal \mapsto [0, 1]$.
+$\kl$ is defined in discrete space as $\kl(g || h) \coloneqq \sum_{x \in \Xcal} g(x) \log \frac{g(x)}{h(x)}$ and $\textrm{Bern(p)}$ denotes the Bernoulli pmf with parameter $p$.
 	\label{lem:kl-div}
@@
 Proving lemma \ref{lem:kl-div} is simply a case of algebraic manipulation and is omitted here.
-However, this allows us to simplify the test-statistic into a form that is easier to compute (equation \ref{eqn:test-stat-compute-form}).
-However, we must simplify further still using lemma \ref{lem:kl-approx-chi}.
+The lemma allows us to simplify the test-statistic into a form that is easier to compute (equation \ref{eqn:test-stat-compute-form}).
 %
@@
 %
+However, we must simplify further using lemma \ref{lem:kl-approx-chi}.
+%
 \begin{lemma}[$\chi^2$ approximation]
@@
 
-The proof requires defining $\delta(x) \coloneqq g(x) - h(x)$ and taking a Taylor expansion of $\log (1 + \delta / h)$ - omitted for bervity.
+The proof requires defining $\delta(x) \coloneqq g(x) - h(x)$ and taking a Taylor expansion of $\log (1 + \delta / h)$ - omitted for brevity.
 The chi-squared distance between two Bernoullis $g = Bern(p)$ and $h = Bern(q)$ has a very simple form:
@@
 %
-We can apply these results to equation \ref{eqn:test-stat-compute-form} to obtain an approximate expression for $t_n$ under the null.
-Under $H_0$ and for large $n$, the estimates $\hat{p}$, $\hat{q}$ and $\hat{r}$ are all close meaning Lemma \ref{lem:kl-approx-chi} is applicable.
+We can apply these results to equation \ref{eqn:test-stat-compute-form} to obtain an approximate expression for $t_n$ valid for large $n$ under the null hypothesis $H_0$.
+Under these conditions, the estimates $\hat{p}$, $\hat{q}$ and $\hat{r}$ are all close meaning Lemma \ref{lem:kl-approx-chi} is applicable.
 This gives:
@@
 %
-Through application of the Central Limit Theorem (CLT) and Slutsky's Theorem it can be shown that as $n \rightarrow \infty$ under the null, the term in brackets on the final line is distributed like a standard Gaussian $z_n \sim \Gaussian(0,1)$.
+Through application of the Central Limit Theorem (CLT) and Slutsky's Theorem it can be shown that as $n \rightarrow \infty$ under the null $H_0$, the term in brackets on the final line is distributed like a standard Gaussian $z_n \sim \Gaussian(0,1)$.
 The proof is long-winded and therefore omitted from this short report.
@@
 \begin{theorem}[$\chi^2$ test]
-	For $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$, given the realised graph and class labels $(X, \Gcal)$ we can perform a hypothesis test on entries $W_\alpha$ and $W_\beta$ of the connectivity matrix $W$.
+	For $(X, \Gcal) \sim \textrm{SBM}(n, p, W)$, given the realised graph and class labels $(X, \Gcal)$ we can perform a hypothesis test on entries $W_\alpha$ and $W_\beta$ of the connectivity matrix $W$:
 	%
@@
 The Stanford Network Analysis Project (SNAP) \cite{snapnets} offers a wealth of Facebook egonets.
-An egonet is simply a graph where all vertices (in this case Facebook users) are connected to one central node (the ego-node).
-The data consists of the undirected set of edges $\Gcal$ indicating whether any two vertices are connected (friends on Facebook) and matrix $F$ containing binary feature flags for each vertex.
+An egonet is simply a graph where all vertices (in this case Facebook users) are connected to one central node (the egonode).
+The data consists of the undirected set of edges $\Gcal$, indicating whether any two vertices are connected (friends on Facebook) and matrix $F$, containing binary feature flags for each vertex.
 Example feature flags are given below: \\
@@
 
-We use an SBM model with $k=2$ communities (1: gender-77 on and 2: gender-77 off) to model the egonet and perform a three-way hypothesis test on the parameters of the connectivity Matrix $W$.
+We use an SBM with $k=2$ communities (1: gender-77 on and 2: gender-77 off) to model the egonet and perform a three-way hypothesis test on the parameters of the connectivity Matrix $W$.
 %
@@
 
-However, the exercise has highlighted some interesting points.
+That said, the exercise has highlighted some interesting points.
 The hypothesis test framework does not quantify the magnitude of the difference just whether or not a difference exists.
@@
 To address these shortcomings we instead approach the problem from a different angle.
-For now we have been given a graph with fully labelled vertices and ask if a given feature affects the structure.
+For now we have been given a graph with fully labelled vertices and ask if a given label affects the structure.
 Instead, we could take a graph and partition first (without using feature information) and then ask which features explain the partition.
@@
 Community detection is the problem of recovering the communities from the graph $\Gcal$.
-There are 4 recovery regimes (plus the trivial no recovery), as defined by Abbe \cite{Abbe}.
+There are 4 recovery regimes, as defined by Abbe \cite{Abbe}.
 We choose to focus on weak recovery as it is the least strict of all.
 Therefore, there are very few constraints on the SBM parameters and the theory can be applied to a wide range of graphs.
-A statement of weak recovery is given in definition \ref{defn:weak-recovery} (adapated from \cite{Abbe}).
+A statement of weak recovery is given in definition \ref{defn:weak-recovery} (adapted from \cite{Abbe}).
 
@@
 
-	Where $o(1)$ denotes a family of functions that tend to $0$ as $n$ gets large.
+	Where $o(1)$ denotes a family of functions that tend to $0$ as $n$ increases.
 We recall the definition of $\Omega_i = \Omega_i(X) \coloneqq \{v \in \Vcal : X_v = i\}$.
@@
 Nevertheless, the premise is that one can infer the community of a particular vertex simply from the labels of all its neighbours.
-The algorithm is non-deterministic and requires the graph $\Gcal$ as well as two hyper-parameters $r$ (the maximum cyclic length to correct for) and $T$ (the number of iterations).
+The algorithm is non-deterministic
+\footnote{As ABP is non-deterministic, each figure is the result of a different run of ABP. No exact comparison is possible between figures but the results are in some sense typical.}
+and requires the graph $\Gcal$ as well as two hyper-parameters $r$ (the maximum cyclic length to correct for) and $T$ (the number of iterations).
 
@@
 
-We colour each vertex according to this quantity $\tilde{\sigma}_v$ and plot the graph
-\footnote{ABP is non-deterministic so each figure is the result of a different run of ABP. No exact comparison is possible between figures but the results are in some sense typical.}
-on figure \ref{fig:abp-output-norm}.
+We colour each vertex according to this quantity $\tilde{\sigma}_v$ and plot the graph on figure \ref{fig:abp-output-norm}.
 We see intuitively that ABP has worked.
@@
 Inspecting figure \ref{fig:ego-0-by-gender} shows us that the sets $S$ and $S^C$ do separate nicely with high intra-group connectivity and lower inter-group connection density.
-After all, we seek to find a roughly balanced partition that minimises the number of edges between the two sets.
+Indeed, an alternative interpretation of weak recovery is finding the partition that minimises the number of edges between the two sets.
 %
@@
 %
-Clearly the fractions of each community in the partition $S$ differ substantially so we say that weak recovery is solved.
+Clearly the fractions of each community in the partition $S$ differ substantially so we say that weak recovery is solved in the sense of definition \ref{defn:weak-recovery}.
 The set $S$ contains a higher proportion of the overall 78-population than of the 77-population.
@@
 
-This part of the project is still in its early stages so these are just preliminary results.
+This part of the project is still in its early stages so we present preliminary results.
 We necessarily remove the constraint on the community sets $\Omega_i$ to be disjoint as one vertex can have multiple features turned on.
@@
 %
-Linear regression finds the constant bias $a$ and $f \times 1$ weight vector $b$ that minimises the mean squared prediction error $\epsilon^T \epsilon / n$.
+Linear regression finds the constant bias $a$ and weight vector $b$ that minimises the mean squared prediction error $\epsilon^T \epsilon / n$.
 The theory for linear regression is well known so omitted here (see \cite{yale-linear-regression} for a reference).
-We restrict $F$ to only include gender and language features and plot the regression coefficient for each feature in order of decreasing magnitude (figure \ref{fig:abp-output-lin-regression}).
+We restrict $F$ to only include gender and language features and plot the corresponding element of $b$ for each feature in order of decreasing magnitude (figure \ref{fig:abp-output-lin-regression}).
 %
@@
 This analysis is of course rather crude as we have failed to take into account multi-collinearity or even determined if a linear regression model is appropriate; indeed, an $R^2$ value of just 0.113 would suggest that it is not.
-Nevertheless, this serves as the platform for subsequent more rigorous analysis.
+Nevertheless, this serves as the starting point for subsequent more rigorous analysis.
 Furthermore, we have not yet proven the conjecture that the regression coefficients are valid measures of causal impact on graphical structure.
 
+\pagebreak
 \section{Future Direction}
@@
 The immediate next steps for the project are to refine the classification approach outlined in section \ref{sect:detection-to-inference}.
-Firstly, it should be tied back to the hypothesis testing framework to make sure that the results are proven to be meaningful.
-Secondly, the exact form of the classifier can be explored; a linear regression model is a good starting point but we can expand to a softmax neural network as that would allow us to generalise to more than just two partitions.
+Firstly, it should be tied back to the hypothesis testing framework to prove the results are be meaningful.
+Secondly, the exact form of the classifier can be explored; a linear regression model is a good starting point but we can expand to a simple softmax neural network as that would allow us to generalise to more than just two partitions.
 
 Indeed, ABP is severely limited by the fact that it only outputs 2 sets.
-Though weak recovery is solved, the partition is often hard to interpret as various clusters get subsumed into one.
+Though weak recovery is solved, the partition is often hard to interpret as multiple clusters can get subsumed into one.
 We instead would prefer to investigate algorithms that can identify arbitrarily many clusters.
@@
 Work has already begun analysing the influence of gender on academic collaboration graphs provided by Aminer \cite{aminer}.
-Nevertheless, this dataset does not contain gender explicitly, only first-name.
-However, this is cross-referenced with a list of popular baby-names to get a good replacement (as suggested by West et al.
-\cite{gender-scholarship}).
-Furthermore, the seminal dataset of political blogs from the 2004 US election \cite{polblogs} which started much of the analysis of the SBM will of course be analysed.
+Furthermore, the seminal dataset of political blogs from the 2004 US election \cite{polblogs} which started much of the analysis of the SBM is in the works.
 
@@
 
-This concludes the Technical Milestone Report for the project thus far.
-We start by fully fleshing out the theoretical foundation of the Stochastic Block Model and hypothesis tests on connectivity parameters.
+We began by fully fleshing out the theoretical foundation of the Stochastic Block Model and hypothesis tests on connectivity parameters.
 Nevertheless, this approach is too narrow in scope to be a useful tool on its own.

### 2021-04-01 final-report/final-report.tex
Scaffold final report
Slight report work
Change plotting function and generalise graph init

--- 2021-01-20 interim-report/interim-report.tex
@@
 \usepackage[backend=bibtex]{biblatex}
-\addbibresource{interim-sources.bib}
+\addbibresource{final-sources.bib}
 
@@
 \begin{align}
-\lik(D | H) &= \prob(X| p) \cdot \prob(\Gcal | W, X) \nonumber \\
+\lik(D | &H) = \prob(X| p) \cdot \prob(\Gcal | W, X) \nonumber \\
 &= \prob(X | p) \prod_{i=1}^{k} \prod_{j=i}^{k} \prob(E_{ij} | W, X) \nonumber \\
@@
 
+\pagebreak
 \begin{theorem}[$\chi^2$ test]
@@
 
+\section{Bayesian Exploration}
+
+For our inference goals we wish to sample block memberships $b$ according to the posterior $p(b | \Gcal)$.
+Peixoto \cite{Peixoto-MCMC} proposes a Monte-Carlo method for generating these samples.
+We simply
+
+At each iteration of the Markov chain we can use the most recent block assignments to train the classifier, mapping from vertex features to detected membership.
+The classifier has parameters $\theta$.
+We wish to apply Bayesian techniques such that we retain a measurement of the uncertainty in the classification.
+We propose to use Stochastic Gradient Langevin Dynamics (SGLD) as described by Nemeth and Fearnhead \cite{Nemeth-SGMCMC}.
+This allows us to draw samples from an arbitrary distribution $\pi$ on $\theta$.
+We can write $\pi$ in the following form.
+%
+\begin{equation}
+	\pi(\theta) \propto \exp \left( - U(\theta) \right)
+\end{equation}
+
+We wish to draw samples from, the posterior.
+%
+\begin{equation}
+	p(\theta | b) = \frac{p(\theta) p(b | \theta)}{p(b)}
+\end{equation}
+
+We can generate these samples and then use them to approximate the true posterior:
+%
+\begin{equation}
+	p(b_v = j | \Gcal) \approx \frac{1}{N} \sum_{n=1}^{N} \one \left\{ b_v^{(n)} = j \right\}
+\end{equation}
+
 \section{Conclusion}

### 2021-04-08 final-report/final-report.tex
Implement uniform prior on b

@@
-\documentclass[11pt]{article}
+\documentclass[]{article}
 
@@
 \usepackage{titling}
+\usepackage{tikz}
 
@@
 
-\begin{definition}
+\begin{definition}[Canonical SBM]
 	\label{defn:sbm}
@@
 Peixoto \cite{Peixoto-MCMC} proposes a Monte-Carlo method for generating these samples.
-We simply
+However, he proposes a slight modification of the Stochastic Block Model.
+The so called microcanonical formulation.
+In contrast to the canonical version where edge parameters are obeyed only on expectation, the microcanonical model has these constraints obeyed with certainty.
+The definition follows from \cite{Peixoto-Bayesian-Microcanonical}.
+
+\begin{definition}[Microcanonical SBM]
+	\label{defn:microcan-sbm}
+	Let $N \in \Integers^{+}$ denote the number of vertices in our graph.
+The block memberships are denoted by a vector $b$ of length $N$ where each entry $b_i \in \{1, 2 \dots B\}$ where $B$ is the number of nonempty blocks.
+Let $e$ be a $B \times B$ matrix of edge counts between blocks ($e_{rs}$ is number of edges from block $r$ onto block $s$ - or twice that number if $r=s$).
+We restrict our analysis to undirected graphs so $e$ is symmetric.
+For a non-degree-corrected stochastic block model (NDC-SBM), we say that the graph $A$ is generated as follows:
+	%
+	\begin{equation}
+		A \sim \textrm{NDC-SBM}_{\textrm{MC}} (b, e)
+	\end{equation}
+	%
+	Where edges are placed at random but respecting the constraints imposed by $e$ and $b$.
+The additional parameters $N$ and $B$ are omitted as they are inferred from the shape ofs of $b$ and $e$.
+If we interpret $A$ as an adjacency matrix, then this constraint can be written formally as: $e_{rs} = \sum_{i=1}^{N} \sum_{j=1}^{N} A_{ij} \one \{b_i = r\} \one \{b_j = s\}$.
+However, this formulation does not tolerate high degree variation within blocks as is typical of real-world data.
+We therefore introduce the degree-corrected SBM (DC-SBM) which has an additional parameter $k$ which is a vector of length $N$ encoding the degree sequence ($k_i$ is the degree of vertex $i$).
+Therefore, we write:
+	%
+	\begin{equation}
+		A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, e, k)
+	\end{equation}
+	%
+	This imposes the additional constraint that $k_i = \sum_{j=1}^{N} A_{ij}$. In what follows, we will always assume the degree-corrected model unless otherwise specified.
+
+\end{definition}
+
+\begin{figure}[!h]
+	\begin{tikzpicture}
+		\draw (0, 0) circle (2);
+		\draw (2, 0) -> (4, 0);
+		\draw (6, 0) circle (2);
+		\draw (8, 0) -> (10, 0);
+		\draw (12, 0) circle (2);
+	\end{tikzpicture}
+\end{figure}
 

### 2021-04-15 nips/submission.tex
nips work

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2021

% ready for submission
\usepackage[final]{neurips_2021}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2021}

% to compile a camera-ready version, add the [final] option, e.g.:
%\usepackage[final]{neurips_2021}

% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_2021}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{bbm}

%bibliography
\bibliographystyle{ieeetr}

\title{Automatic feature classification in stochastic block models}
% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{
  Lawrence Tray \\
  Department of Engineering \\
  University of Cambridge \\
  \texttt{lpt30@cam.ac.uk} \\
  % examples of more authors
  \And
  Ioannis Kontoyiannis \\
  Department of Mathematics \\
  University of Cambridge \\
  \texttt{ik355@cam.ac.uk}
}

%custom commands
\newcommand{\Xcal}{\mathcal{X}}
\newcommand{\Integers}{\mathbb{Z}}
\newcommand{\one}{\mathbbm{1}}
\newcommand{\Gaussian}{\mathcal{N}}
\newcommand{\indep}{\perp \!\!\! \perp}
\newcommand{\specialchoose}{\genfrac{\{}{\}}{0pt}{}}

%envs
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}

\begin{document}

\maketitle

\begin{abstract}
  Paper abstract.
\end{abstract}

\section{Introduction}

\section{Preliminaries}

We will be using the microcanonical stochastic block model, proposed by \cite{Peixoto-Bayesian-Microcanonical}. A paraphrased definition is given below.

\begin{definition}[Microcanonical SBM]
	\label{defn:microcan-sbm}
	Let $N \in \Integers^{+}$ denote the number of vertices in our graph. The block memberships are denoted by a vector $b$ of length $N$ where each entry $b_i \in \{1, 2 \dots B\}$ where $B$ is the number of nonempty blocks. Let $e$ be a $B \times B$ matrix of edge counts between blocks ($e_{rs}$ is number of edges from block $r$ onto block $s$ - or twice that number if $r=s$). We restrict our analysis to undirected graphs so $e$ is symmetric. For a non-degree-corrected stochastic block model (NDC-SBM), we say that the graph $A$ is generated as follows:
	%
	\begin{equation}
		A \sim \textrm{NDC-SBM}_{\textrm{MC}} (b, e)
	\end{equation}
	%
	Where edges are placed at random but respecting the constraints imposed by $e$ and $b$. The additional parameters $N$ and $B$ are omitted as they are inferred from the shape ofs of $b$ and $e$. If we interpret $A$ as an adjacency matrix, then this constraint can be written formally as: $e_{rs} = \sum_{i=1}^{N} \sum_{j=1}^{N} A_{ij} \one \{b_i = r\} \one \{b_j = s\}$. However, this formulation does not tolerate high degree variation within blocks as is typical of real-world data. We therefore introduce the degree-corrected SBM (DC-SBM) which has an additional parameter $k$ which is a vector of length $N$ encoding the degree sequence ($k_i$ is the degree of vertex $i$). Therefore, we write:
	%
	\begin{equation}
		A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, e, k)
	\end{equation}
	%
	This imposes the additional constraint that $k_i = \sum_{j=1}^{N} A_{ij}$. In what follows, we will always assume the degree-corrected model unless otherwise specified.

\end{definition}

\section{Latent block generative model}

We restrict our analysis to labelled, undirected graphs with $N$ nodes. We define the vector $x_i \in \Xcal^D$ as the feature vector for the $i$'th vertex. Each vertex has $D$ total features and we assume all entries take values from the same set $\Xcal$. For the majority of datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$. The feature vectors $\{x_i\}_{i=1}^{N}$ are subsumed into the $N \times D$ matrix $X$.

The proposed generative model is given in figure \ref{fig:lbgm}. We start, with the feature matrix X and generate a vector of block memberships $b$. The parameters of this generator are encapsulated by $\theta$. Each feature vector $x_i$ is treated independently and used to generate the block membership $b_i$. We choose a single softmax layer to model $p(b_i | x_i, \theta)$. More complex models are possible but then deriving meaning from inferred parameter distributions is complicated. Summarising, we can write $p(b | X, \theta)$ as follows:
%
\begin{equation}
	p(b| X, \theta) = \prod_{i=1}^{N} p(b_i | x_i, \theta) = \prod_{i=1}^{N} \phi_{b_i} (x_i; \theta)
	= \prod_{i=1}^{N} \frac{\exp\left(w_{b_i}^T \tilde{x}_i\right)}{\sum_{k=1}^{B} \exp \left( w_k^T \tilde{x}_i\right)}
\end{equation}
%
Where $\tilde{x} \coloneqq \left[x_1, x_2, \dots x_D, 1\right]^T$ is an augmented version of $x$ that allows for a bias term. The parameters $\theta$ just contain the $B \times (D+1)$ matrix of weight values or alternatively, $\theta = \{w_k\}_{k=1}^{B}$.

\begin{figure}[!h]
	\centering
	\begin{tikzpicture}[
		roundnode/.style={circle, draw=black, minimum size=7mm},
		squarednode/.style={rectangle, draw=black, minimum size=7mm}
		]
		% nodes
		\node[roundnode] (X) at (0, 0) {X};
		\node[squarednode] (b) at (3, 0) {b};
		\node[roundnode] (A) at (6, 0) {A};

		% arrows
		\draw[->] (X.east) -- node[above] {$\theta$} (b.west);
		\draw[->] (b.east) -- node[above] {$\psi$}(A.west);
	\end{tikzpicture}
	\caption{Latent block generative model}
	\label{fig:lbgm}
\end{figure}

Once we have generated the block memberships $b$, we proceed to draw the graph $A$ from the microcanical SBM (\ref{defn:microcan-sbm}) with additional parameters $\psi = \{e, k\}$. In a slight abuse of notation we denote the inter-block edge count matrix with $e = \psi_e$ and the degree sequence $k = \psi_k$ to make explicit that these are contained in $\psi$.
%
\begin{equation}
	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k)
\end{equation}

\subsection{Prior selection}

Before performing any inference, we must specify priors on $\theta$ and $\psi$. For $\theta$ it seems sensible to choose a Gaussian prior, with zero mean and variance matrix $\sigma^2_\theta I$ such that each element of $\theta$ is independent and distributed like $\sim \Gaussian(0, \sigma)$. In vector form, the prior for $\theta$ is therefore:
%
\begin{equation}
	p(\theta) = \Gaussian \left( \theta ; 0, \sigma_\theta^2 I \right)
\end{equation}
%
We will see that this form of prior is equivalent to a regularisation term in neural network training that penalises extreme weight magnitudes. As $\sigma_\theta^2 \rightarrow \infty$ this becomes an uninformative uniform prior.

In our model $b$ is now an intermediate variable and so we cannot choose a prior. The closest thing we can get to a prior is $p(b | X)$. As far as inference on the right-hand-side of figure \ref{fig:lbgm}, we regard $p(b | X)$ as a pseudo-prior on $b$. We can show that our choice of prior for $p(\theta)$ leads to the following form for $p(b | X)$.
%
\begin{equation}
	p(b | X) = \int p(b | X, \theta) p(\theta) d\theta = B^{-N}
\end{equation}
%
In his paper, \cite{Peixoto-Bayesian-Microcanonical} proposes careful choices for the SBM parameters. The proposal is to write the joint prior on $(b, e, k)$ as a product of conditionals $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$. For our purposes we must insert a conditioning on $X$, to form our pseudo-prior for $b$ and $\psi$.
%
\begin{equation}
	p(b, \psi | X) = p(b | X) p(\psi | b, X) = p(b | X) p(\psi | b)
\end{equation}
%
Where it is apparent by regarding figure \ref{fig:lbgm} as a Markov model that $(\psi \indep X) | b$. We then borrow the priors proposed by \cite{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference.
%
\begin{equation}
	p(\psi | b) = p(e | b) p(k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right]
\end{equation}
%
Where $\specialchoose{n}{m}$ is shorthand for ${{n+m-1} \choose {m}} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of histograms (non-negative bin values) with $n$ bins that are constrained to sum to $m$. $E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges. Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform with respect to $e$. The variable $\eta_j^r$ is introduced to denote the number of vertices in block $r$ that have degree $j$. Formally, $\eta_j^r \coloneqq \sum_{i} \one\left\{b_i = r \right\} \one \left\{k_i = j \right\}$. Furthermore, $q(m, n)$ is the number of different histograms with at most $n$ non-zero bins that sum to $m$. Lastly, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one{b_i = r}$ is the number of vertices assigned to block $r$.

Importantly, we have computable forms for $p(\theta)$ and $p(b, \psi | X)$ which will be useful for performing inference.

\section{Inference}

We are presented with a vertex-labelled graph $(A, X)$. The goal is to draw samples from equation \ref{eqn:theta-target}. However, this is not easily done in practice.

\begin{equation}
	\label{eqn:theta-target}
	\theta^{(i)} \sim p(\theta | A, X)
\end{equation}
%
We instead propose an iterative approach:
%
\begin{align}
	b^{(i)} &\sim p(b | A, X) \\
	\theta^{(i)} &\sim p(\theta | X, b^{(i)})
\end{align}
%

\begin{ack}
Use unnumbered first level headings for the acknowledgments. All acknowledgments
go at the end of the paper before the list of references. Moreover, you are required to declare
funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2021/PaperInformation/FundingDisclosure}.

Do {\bf not} include this section in the anonymized submission, only in the final paper. You can use the \texttt{ack} environment provided in the style file to automatically hide this section in the anonymized submission.
\end{ack}

%references
\bibliography{sources.bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Checklist}

%%% BEGIN INSTRUCTIONS %%%
The checklist follows the references.  Please
read the checklist guidelines carefully for information on how to answer these
questions.  For each question, change the default \answerTODO{} to \answerYes{},
\answerNo{}, or \answerNA{}.  You are strongly encouraged to include a {\bf
justification to your answer}, either by referencing the appropriate section of
your paper or providing a brief inline description.  For example:
\begin{itemize}
  \item Did you include the license to the code and datasets? \answerYes{See Section~\ref{gen_inst}.}
  \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
  \item Did you include the license to the code and datasets? \answerNA{}
\end{itemize}
Please do not modify the questions and only use the provided macros for your
answers.  Note that the Checklist section does not count towards the page
limit.  In your paper, please delete this instructions block and only keep the
Checklist section heading above along with the questions/answers below.
%%% END INSTRUCTIONS %%%

\begin{enumerate}

\item For all authors...
\begin{enumerate}
  \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
    \answerTODO{}
  \item Did you describe the limitations of your work?
    \answerTODO{}
  \item Did you discuss any potential negative societal impacts of your work?
    \answerTODO{}
  \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
    \answerTODO{}
\end{enumerate}

\item If you are including theoretical results...
\begin{enumerate}
  \item Did you state the full set of assumptions of all theoretical results?
    \answerTODO{}
	\item Did you include complete proofs of all theoretical results?
    \answerTODO{}
\end{enumerate}

\item If you ran experiments...
\begin{enumerate}
  \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
    \answerTODO{}
  \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
    \answerTODO{}
	\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
    \answerTODO{}
	\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
    \answerTODO{}
\end{enumerate}

\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
\begin{enumerate}
  \item If your work uses existing assets, did you cite the creators?
    \answerTODO{}
  \item Did you mention the license of the assets?
    \answerTODO{}
  \item Did you include any new assets either in the supplemental material or as a URL?
    \answerTODO{}
  \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
    \answerTODO{}
  \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
    \answerTODO{}
\end{enumerate}

\item If you used crowdsourcing or conducted research with human subjects...
\begin{enumerate}
  \item Did you include the full text of instructions given to participants and screenshots, if applicable?
    \answerTODO{}
  \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
    \answerTODO{}
  \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
    \answerTODO{}
\end{enumerate}

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Appendix}

\subsection{Derivation of conditional block distribution given feature matrix}
\label{appdx:b|x}

We wish to determine the form of $p(b| X)$. This can be done by integrating over the joint probability with respect to $\theta$.
%
\begin{align*}
	p(b | X) &= \int p(b , \theta| X, \theta) d\theta = \int p(b | X, \theta) p(\theta | X) d\theta \\
	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i=1}^{N} \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
	&= \prod_{i=1}^{N} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j=1}^{B} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k=1}^{B} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}
\end{align*}
%
We note that $b_i \in {1, 2, \dots B}$ and so the integral's value is unchanged with respect to $b_i$. The integrand has the same form no matter which value $b_i$ takes as the prior is the same for each $w_j$. As such the integral can only be a function of at most $\tilde{x}_i$ and $\sigma_\theta^2$ as it is symmetric with respect to $b_i$ and all the various $w_j$ are integrated out as they are dummy variables. Therefore, denoting the integral by the (unknown) function $f(\tilde{x}_i, \sigma_\theta^2)$, we write $p(b| X)$ as follows:
%
\begin{align*}
	p(b | X) &= \prod_{i=1}^{N} f(\tilde{x}_i, \sigma_\theta^2) = \textrm{const w.r.t } b = c
\end{align*}
%
As this is a constant with respect to $b$ we conclude that $p(b | X)$ must be a uniform distribution. $\nicefrac{1}{c}$ is simply the size of the set of values that $b$ can take. We know $b_i \in \mathcal{B} = \{1, 2, \dots B\}$. Therefore, $b \in \mathcal{B}^N$ and $|\mathcal{B}^N| = |\mathcal{B}|^N = B^N= \nicefrac{1}{c}$. Putting this all together we show that:
%
\begin{equation}
	p(b | X) = B^{-N}
\end{equation}

\end{document}

### 2021-04-15 nips/rules/neurips_2021.tex
Compartmentalise

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2021

% ready for submission
\usepackage{neurips_2021}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2021}

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2021}

% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_2021}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors

\title{Formatting Instructions For NeurIPS 2021}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{Lawrence Tray}

\author{%
  David S.~Hippocampus\thanks{Use footnote for providing further information
    about author (webpage, alternative address)---\emph{not} for acknowledging
    funding agencies.} \\
  Department of Computer Science\\
  Cranberry-Lemon University\\
  Pittsburgh, PA 15213 \\
  \texttt{hippo@cs.cranberry-lemon.edu} \\
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

\begin{document}

\maketitle

\begin{abstract}
  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
  both the left- and right-hand margins. Use 10~point type, with a vertical
  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
  bold, and in point size 12. Two line spaces precede the abstract. The abstract
  must be limited to one paragraph.
\end{abstract}

\section{Submission of papers to NeurIPS 2021}

Please read the instructions below carefully and follow them faithfully.

\subsection{Style}

Papers to be submitted to NeurIPS 2021 must be prepared according to the
instructions presented here. Papers may only be up to {\bf nine} pages long,
including figures. Additional pages \emph{containing only acknowledgments and
references} are allowed. Papers that exceed the page limit will not be
reviewed, or in any other way considered for presentation at the conference.

The margins in 2021 are the same as those in 2007, which allow for $\sim$$15\%$
more words in the paper compared to earlier years.

Authors are required to use the NeurIPS \LaTeX{} style files obtainable at the
NeurIPS website as indicated below. Please make sure you use the current files
and not previous versions. Tweaking the style files may be grounds for
rejection.

\subsection{Retrieval of style files}

The style files for NeurIPS and other conference information are available on
the World Wide Web at
\begin{center}
  \url{http://www.neurips.cc/}
\end{center}
The file \verb+neurips_2021.pdf+ contains these instructions and illustrates the
various formatting requirements your NeurIPS paper must satisfy.

The only supported style file for NeurIPS 2021 is \verb+neurips_2021.sty+,
rewritten for \LaTeXe{}.  \textbf{Previous style files for \LaTeX{} 2.09,
  Microsoft Word, and RTF are no longer supported!}

The \LaTeX{} style file contains three optional arguments: \verb+final+, which
creates a camera-ready copy, \verb+preprint+, which creates a preprint for
submission to, e.g., arXiv, and \verb+nonatbib+, which will not load the
\verb+natbib+ package for you in case of package clash.

\paragraph{Preprint option}
If you wish to post a preprint of your work online, e.g., on arXiv, using the
NeurIPS style, please use the \verb+preprint+ option. This will create a
nonanonymized version of your work with the text ``Preprint. Work in progress.''
in the footer. This version may be distributed as you see fit. Please \textbf{do
  not} use the \verb+final+ option, which should \textbf{only} be used for
papers accepted to NeurIPS.

At submission time, please omit the \verb+final+ and \verb+preprint+
options. This will anonymize your submission and add line numbers to aid
review. Please do \emph{not} refer to these line numbers in your paper as they
will be removed during generation of camera-ready copies.

The file \verb+neurips_2021.tex+ may be used as a ``shell'' for writing your
paper. All you have to do is replace the author, title, abstract, and text of
the paper with your own.

The formatting instructions contained in these style files are summarized in
Sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.

\section{General formatting instructions}
\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas) wide and
9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).  Use 10~point
type with a vertical spacing (leading) of 11~points.  Times New Roman is the
preferred typeface throughout, and will be selected for you by default.
Paragraphs are separated by \nicefrac{1}{2}~line space (5.5 points), with no
indentation.

The paper title should be 17~point, initial caps/lower case, bold, centered
between two horizontal rules. The top rule should be 4~points thick and the
bottom rule should be 1~point thick. Allow \nicefrac{1}{4}~inch space above and
below the title to rules. All pages should start at 1~inch (6~picas) from the
top of the page.

For the final version, authors' names are set in boldface, and each name is
centered above the corresponding address. The lead author's name is to be listed
first (left-most), and the co-authors' names (if different address) are set to
follow. If there is only one co-author, list both author and co-author side by
side.

Please pay special attention to the instructions in Section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Headings: first level}
\label{headings}

All headings should be lower case (except for first word and proper nouns),
flush left, and bold.

First-level headings should be in 12-point type.

\subsection{Headings: second level}

Second-level headings should be in 10-point type.

\subsubsection{Headings: third level}

Third-level headings should be in 10-point type.

\paragraph{Paragraphs}

There is also a \verb+\paragraph+ command available, which sets the heading in
bold, flush left, and inline with the text, with the heading followed by 1\,em
of space.

\section{Citations, figures, tables, references}
\label{others}

These instructions apply to everyone.

\subsection{Citations within the text}

The \verb+natbib+ package will be loaded for you by default.  Citations may be
author/year or numeric, as long as you maintain internal consistency.  As to the
format of the references themselves, any style is acceptable as long as it is
used consistently.

The documentation for \verb+natbib+ may be found at
\begin{center}
  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
\end{center}
Of note is the command \verb+\citet+, which produces citations appropriate for
use in inline text.  For example,
\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}
produces
\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

If you wish to load the \verb+natbib+ package with options, you may add the
following before loading the \verb+neurips_2021+ package:
\begin{verbatim}
   \PassOptionsToPackage{options}{natbib}
\end{verbatim}

If \verb+natbib+ clashes with another package you load, you can add the optional
argument \verb+nonatbib+ when loading the style file:
\begin{verbatim}
   \usepackage[nonatbib]{neurips_2021}
\end{verbatim}

As submission is double blind, refer to your own published work in the third
person. That is, use ``In the previous work of Jones et al.\ [4],'' not ``In our
previous work [4].'' If you cite your other papers that are not widely available
(e.g., a journal paper under review), use anonymous author names in the
citation, e.g., an author of the form ``A.\ Anonymous.''

\subsection{Footnotes}

Footnotes should be used sparingly.  If you do require a footnote, indicate
footnotes with a number\footnote{Sample of the first footnote.} in the
text. Place the footnotes at the bottom of the page on which they appear.
Precede the footnote with a horizontal rule of 2~inches (12~picas).

Note that footnotes are properly typeset \emph{after} punctuation
marks.\footnote{As in this example.}

\subsection{Figures}

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
\end{figure}

All artwork must be neat, clean, and legible. Lines should be dark enough for
purposes of reproduction. The figure number and caption always appear after the
figure. Place one line space before the figure caption and one line space after
the figure. The figure caption should be lower case (except for first word and
proper nouns); figures are numbered consecutively.

You may use color figures.  However, it is best for the figure captions and the
paper body to be legible if the paper is printed in either black/white or in
color.

\subsection{Tables}

All tables must be centered, neat, clean and legible.  The table number and
title always appear before the table.  See Table~\ref{sample-table}.

Place one line space before the table title, one line space after the
table title, and one line space after the table. The table title must
be lower case (except for first word and proper nouns); tables are
numbered consecutively.

Note that publication-quality tables \emph{do not contain vertical rules.} We
strongly suggest the use of the \verb+booktabs+ package, which allows for
typesetting high-quality, professional tables:
\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}
This package was used to typeset Table~\ref{sample-table}.

\begin{table}
  \caption{Sample table title}
  \label{sample-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
\end{table}

\section{Final instructions}

Do not change any aspects of the formatting parameters in the style files.  In
particular, do not modify the width or length of the rectangle the text should
fit into, and do not change font sizes (except perhaps in the
\textbf{References} section; see below). Please note that pages should be
numbered.

\section{Preparing PDF files}

Please prepare submission files with paper size ``US Letter,'' and not, for
example, ``A4.''

Fonts were the main cause of problems in the past years. Your PDF file must only
contain Type 1 or Embedded TrueType fonts. Here are a few instructions to
achieve this.

\begin{itemize}

\item You should directly generate PDF files using \verb+pdflatex+.

\item You can check which fonts a PDF files uses.  In Acrobat Reader, select the
  menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
  also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
  available out-of-the-box on most Linux machines.

\item The IEEE has recommendations for generating PDF files whose fonts are also
  acceptable for NeurIPS. Please see
  \url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}

\item \verb+xfig+ "patterned" shapes are implemented with bitmap fonts.  Use
  "solid" shapes instead.

\item The \verb+\bbold+ package almost always uses bitmap fonts.  You should use
  the equivalent AMS Fonts:
\begin{verbatim}
   \usepackage{amsfonts}
\end{verbatim}
followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or \verb+\mathbb{C}+
for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You can also use the following
workaround for reals, natural and complex:
\begin{verbatim}
   \newcommand{\RR}{I\!\!R} %real numbers
   \newcommand{\Nat}{I\!\!N} %natural numbers
   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
\end{verbatim}
Note that \verb+amsfonts+ is automatically loaded by the \verb+amssymb+ package.

\end{itemize}

If your file contains type 3 fonts or non embedded TrueType fonts, we will ask
you to fix it.

\subsection{Margins in \LaTeX{}}

Most of the margin problems come from figures positioned by hand using
\verb+\special+ or other commands. We suggest using the command
\verb+\includegraphics+ from the \verb+graphicx+ package. Always specify the
figure width as a multiple of the line width as in the example below:
\begin{verbatim}
   \usepackage[pdftex]{graphicx} ...
   \includegraphics[width=0.8\linewidth]{myfile.pdf}
\end{verbatim}
See Section 4.4 in the graphics bundle documentation
(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})

A number of width problems arise when \LaTeX{} cannot properly hyphenate a
line. Please give LaTeX hyphenation hints using the \verb+\-+ command when
necessary.

\begin{ack}
Use unnumbered first level headings for the acknowledgments. All acknowledgments
go at the end of the paper before the list of references. Moreover, you are required to declare
funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2021/PaperInformation/FundingDisclosure}.

Do {\bf not} include this section in the anonymized submission, only in the final paper. You can use the \texttt{ack} environment provided in the style file to autmoatically hide this section in the anonymized submission.
\end{ack}

\section*{References}

References follow the acknowledgments. Use unnumbered first-level heading for
the references. Any choice of citation style is acceptable as long as you are
consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
when listing the references.
Note that the Reference section does not count towards the page limit.
\medskip

{
\small

[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
(eds.), {\it Advances in Neural Information Processing Systems 7},
pp.\ 609--616. Cambridge, MA: MIT Press.

[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
TELOS/Springer--Verlag.

[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
recall at excitatory recurrent synapses and cholinergic modulation in rat
hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Checklist}

%%% BEGIN INSTRUCTIONS %%%
The checklist follows the references.  Please
read the checklist guidelines carefully for information on how to answer these
questions.  For each question, change the default \answerTODO{} to \answerYes{},
\answerNo{}, or \answerNA{}.  You are strongly encouraged to include a {\bf
justification to your answer}, either by referencing the appropriate section of
your paper or providing a brief inline description.  For example:
\begin{itemize}
  \item Did you include the license to the code and datasets? \answerYes{See Section~\ref{gen_inst}.}
  \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
  \item Did you include the license to the code and datasets? \answerNA{}
\end{itemize}
Please do not modify the questions and only use the provided macros for your
answers.  Note that the Checklist section does not count towards the page
limit.  In your paper, please delete this instructions block and only keep the
Checklist section heading above along with the questions/answers below.
%%% END INSTRUCTIONS %%%

\begin{enumerate}

\item For all authors...
\begin{enumerate}
  \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
    \answerTODO{}
  \item Did you describe the limitations of your work?
    \answerTODO{}
  \item Did you discuss any potential negative societal impacts of your work?
    \answerTODO{}
  \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
    \answerTODO{}
\end{enumerate}

\item If you are including theoretical results...
\begin{enumerate}
  \item Did you state the full set of assumptions of all theoretical results?
    \answerTODO{}
	\item Did you include complete proofs of all theoretical results?
    \answerTODO{}
\end{enumerate}

\item If you ran experiments...
\begin{enumerate}
  \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
    \answerTODO{}
  \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
    \answerTODO{}
	\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
    \answerTODO{}
	\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
    \answerTODO{}
\end{enumerate}

\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
\begin{enumerate}
  \item If your work uses existing assets, did you cite the creators?
    \answerTODO{}
  \item Did you mention the license of the assets?
    \answerTODO{}
  \item Did you include any new assets either in the supplemental material or as a URL?
    \answerTODO{}
  \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
    \answerTODO{}
  \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
    \answerTODO{}
\end{enumerate}

\item If you used crowdsourcing or conducted research with human subjects...
\begin{enumerate}
  \item Did you include the full text of instructions given to participants and screenshots, if applicable?
    \answerTODO{}
  \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
    \answerTODO{}
  \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
    \answerTODO{}
\end{enumerate}

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Appendix}

Optionally include extra information (complete proofs, additional experiments and plots) in the appendix.
This section will often be part of the supplemental material.

\end{document}

### 2021-04-15 nips/submission.tex
Compartmentalise

@@
 %bibliography
-\bibliographystyle{ieeetr}
+\bibliographystyle{plainnat}
 
@@
 %
-In his paper, \cite{Peixoto-Bayesian-Microcanonical} proposes careful choices for the SBM parameters.
+In his paper, Peixoto \cite{Peixoto-Bayesian-Microcanonical} proposes careful choices for the SBM parameters.
 The proposal is to write the joint prior on $(b, e, k)$ as a product of conditionals $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$.
@@
 Where it is apparent by regarding figure \ref{fig:lbgm} as a Markov model that $(\psi \indep X) | b$.
-We then borrow the priors proposed by \cite{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference.
+We then borrow the priors proposed by Peixoto \cite{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference.
 %
@@
 %
-Where $\specialchoose{n}{m}$ is shorthand for ${{n+m-1} \choose {m}} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of histograms (non-negative bin values) with $n$ bins that are constrained to sum to $m$.
+Where $\specialchoose{n}{m}$ is shorthand for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of histograms (non-negative bin values) with $n$ bins that are constrained to sum to $m$.
 $E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges.
@@
 %
-We instead propose an iterative approach:
+We instead propose an iterative approach.
+First drawing samples from the block membership posterior $p(b | A, X)$ and then using these samples to draw samples from $\theta$ conditioned on the block membership sample $b^{(i)}$ and feature matrix.
 %
 \begin{align}
-	b^{(i)} &\sim p(b | A, X) \\
-	\theta^{(i)} &\sim p(\theta | X, b^{(i)})
+	b^{(i)} &\sim p \Big( b | A, X \Big) \\
+	\theta^{(i)} &\sim p\Big(\theta | X, b^{(i)} \Big)
 \end{align}
 %
+Both of these can be implemented using a variation of the Metropolis-Hastings algorithm \cite{hastings-alg}.
+We just need to define a proposal distribution $q(\cdot, \cdot)$ for proposing moves and be able to evaluate the target distribution point-wise.
+
+\subsection{Inferring block memeberships from graph $p (b | A, X)$}
+
+Peixoto \cite{Peixoto-MCMC} proposes a Monte Carlo method which we will base our approach on. It relies on writing the posterior in the following form.
+
+\begin{equation}
+	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) \coloneqq \pi_b(b)
+\end{equation}
+
+\section{Experiments}
+
+\section{Conclusion}
 
@@
 \begin{itemize}
-  \item Did you include the license to the code and datasets? \answerYes{See Section~\ref{gen_inst}.}
+  \item Did you include the license to the code and datasets? \answerYes{See Section gen-inst.}
   \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}

### 2021-04-15 nips/submission.tex
Sort out references

@@
 % ready for submission
+\PassOptionsToPackage{numbers, compress}{natbib}
 \usepackage[final]{neurips_2021}
@@
 %
-In his paper, Peixoto \cite{Peixoto-Bayesian-Microcanonical} proposes careful choices for the SBM parameters.
+In his paper, \citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for the SBM parameters.
 The proposal is to write the joint prior on $(b, e, k)$ as a product of conditionals $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$.
@@
 Where it is apparent by regarding figure \ref{fig:lbgm} as a Markov model that $(\psi \indep X) | b$.
-We then borrow the priors proposed by Peixoto \cite{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference.
+We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference.
 %
@@
 
-Peixoto \cite{Peixoto-MCMC} proposes a Monte Carlo method which we will base our approach on. It relies on writing the posterior in the following form.
+\citet{Peixoto-MCMC} proposes a Monte Carlo method which we will base our approach on. It relies on writing the posterior in the following form.
 

### 2021-04-15 nips/submission.tex
More inference fleshing out

@@
 \newcommand{\Xcal}{\mathcal{X}}
+\newcommand{\Bcal}{\mathcal{B}}
 \newcommand{\Integers}{\mathbb{Z}}
@@
 \newcommand{\specialchoose}{\genfrac{\{}{\}}{0pt}{}}
+\newcommand{\Expect}{\mathbb{E}}
 
@@
 We are presented with a vertex-labelled graph $(A, X)$. The goal is to draw samples from equation \ref{eqn:theta-target}. However, this is not easily done in practice.
-
+%
 \begin{equation}
@@
 We instead propose an iterative approach.
-First drawing samples from the block membership posterior $p(b | A, X)$ and then using these samples to draw samples from $\theta$ conditioned on the block membership sample $b^{(i)}$ and feature matrix.
+First drawing samples $b^{(i)}$ from the block membership posterior (equation \ref{eqn:b-samples}).
+We then use each $b^{(i)}$ to draw samples for $\theta$ as in equation \ref{eqn:theta-samples}.
 %
 \begin{align}
-	b^{(i)} &\sim p \Big( b | A, X \Big) \\
-	\theta^{(i)} &\sim p\Big(\theta | X, b^{(i)} \Big)
+	b^{(i)} &\sim p \Big( b | A, X \Big)  \label{eqn:b-samples}\\
+	\theta^{(i)} &\sim p\Big(\theta | X, b^{(i)} \Big) \label{eqn:theta-samples}
 \end{align}
 %
-Both of these can be implemented using a variation of the Metropolis-Hastings algorithm \cite{hastings-alg}.
-We just need to define a proposal distribution $q(\cdot, \cdot)$ for proposing moves and be able to evaluate the target distribution point-wise.
+Both of these can be implemented with a Markov Chain through the Metropolis-Hastings algorithm \cite{hastings-alg}.
+We just need to define a proposal distribution $q(x, y)$ for proposing a move $x \rightarrow y$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(\cdot)$, point-wise.
+The proposed move is then accepted with acceptance probability $\alpha$ else it is rejected and we stay at $x$.
+%
+\begin{equation}
+	\alpha = \min \left( \frac{\pi(y) q(y, x)}{\pi(x) q(x, y)} , 1 \right)
+\end{equation}
+%
+This accept-reject step ensures the resulting Markov Chain is in detailed balance with the target distribution $\pi(\cdot)$.
+What we propose in equations \ref{eqn:b-samples} and \ref{eqn:theta-samples} is therefore implemented through a 2-level Markov chain.
+The resulting samples for $\theta^{(i)}$ are unbiased in the sense that the expectation of their distribution is the posterior we are targeting in equation \ref{eqn:theta-target}.
+%
+\begin{align*}
+	\Expect_{b^{(i)}} \left[p \left( \theta | X, b^{(i)} \right) \right] &= \sum_{b \in \mathcal{B}^N} p(\theta | X, b) p(b | A, X) \\
+	&= \sum_{b \in \mathcal{B}^N} p(\theta, b | A, X) \\
+	&= p(\theta | A, X) \\
+\end{align*}
+%
+Which is indeed the distribution we are targeting from equation \ref{eqn:theta-target}.
+The reason we split the Markov chain into two stages is because the summation over all latent states $b \in \mathcal{B}^N$ required to compute the likelihood $p(A| X, \theta) = \sum_{b \in \Bcal^N}$ is computationally intensive
 
-\subsection{Inferring block memeberships from graph $p (b | A, X)$}
+\subsection{Inferring block memberships}
 
 \citet{Peixoto-MCMC} proposes a Monte Carlo method which we will base our approach on. It relies on writing the posterior in the following form.
+%
+\begin{equation}
+	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) = \pi_b(b)
+\end{equation}
+%
+Now $\pi_b(\cdot)$ is the un-normalised density we wish to sample from. We can break $\pi_b$ down as follows:
+%
+\begin{align*}
+	\pi_b(b) &= p(b|X) \sum_{\psi} \nolimits p(A , \psi | b, X) \\
+	&= p(b|X) p(A, \psi^* | b, X) \\
+	&= p(A | b, \psi^*) \cdot p(\psi^* | b) \cdot p(b | X)
+\end{align*}
+%
+Since we are using a microcanonical formulation, there is only one value of $\psi$ which we denote $\psi^*$ that is compatible with the given $A$ and $b$.
+Namely, $k^*_i = \sum_j A_{ij} $ and $e^*_{rs} = \sum_{i, j} A_ij \one \{b_i=r\} \one\{b_j=s\}$.
+Therefore, the summation over all $\psi$ reduces to just one term that has the value $\psi^*$ compatible with $b$ and $\theta$.
+We also define the description length or entropy of the configuration as.
+%
+\begin{equation}
+	S(b) = - \log \pi_b(b) = - \left[ \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \right]
+\end{equation}
+%
+Now this entropy can be thought of as the description length of the graph $A$ because it is the sum of the information required to represent the graph given the parameters and the amount of information required to store the parameters given $X$.
 
-\begin{equation}
-	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) \coloneqq \pi_b(b)
-\end{equation}
+\subsection{Inferring feature to block generator}
 

### 2021-04-18 nips/submission.tex
Work on feat-gen

@@
 \usepackage{bbm}
+\usepackage{amssymb}
 
@@
 Which is indeed the distribution we are targeting from equation \ref{eqn:theta-target}.
-The reason we split the Markov chain into two stages is because the summation over all latent states $b \in \mathcal{B}^N$ required to compute the likelihood $p(A| X, \theta) = \sum_{b \in \Bcal^N}$ is computationally intensive
+The reason we split the Markov chain into two stages is because the summation over all latent states $b \in \mathcal{B}^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in \Bcal^N} p(A | b) P(b | X, \theta)$ is computationally intensive.
 
-\subsection{Inferring block memberships}
+\subsection{Sampling block memberships}
 
@@
 %
-Now $\pi_b(\cdot)$ is the un-normalised density we wish to sample from. We can break $\pi_b$ down as follows:
+Now $\pi_b(\cdot)$ is the un-normalised density we wish to sample from.
+In other words, we wish to construct a Markov chain that has $\pi_b(\cdot)$ as its invariant distribution.
+We can break $\pi_b$ down as follows:
 %
@@
 %
-Since we are using a microcanonical formulation, there is only one value of $\psi$ which we denote $\psi^*$ that is compatible with the given $A$ and $b$.
-Namely, $k^*_i = \sum_j A_{ij} $ and $e^*_{rs} = \sum_{i, j} A_ij \one \{b_i=r\} \one\{b_j=s\}$.
-Therefore, the summation over all $\psi$ reduces to just one term that has the value $\psi^*$ compatible with $b$ and $\theta$.
-We also define the description length or entropy of the configuration as.
+Since we are using a microcanonical formulation, there is only one value of $\psi$ which we denote $\psi^*$ that is compatible with the given $(A, b)$ pair.
+Specifically, $k^*_i = \sum_j A_{ij} $ and $e^*_{rs} = \sum_{i, j} A_ij \one \{b_i=r\} \one\{b_j=s\}$.
+Therefore, the summation over all $\psi$ reduces to just the single $\psi^*$ term.
+We also define the microcanonical entropy of the configuration as.
 %
@@
 %
-Now this entropy can be thought of as the description length of the graph $A$ because it is the sum of the information required to represent the graph given the parameters and the amount of information required to store the parameters given $X$.
+This entropy can be thought of as the description length of the graph because it is the sum of the information required to represent the graph given the parameters and the amount of information required to store the parameters (given the feature matrix $X$).
 
 \subsection{Inferring feature to block generator}
+
+Now the invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. We write this as follows:
+%
+\begin{align}
+	p(\theta | X, b) &\propto p(b | X, \theta) p(\theta) = \pi_\theta (\theta) =  \exp \left( - U(\theta) \right) \\
+	\therefore U(\theta) &= - \left( \log p(b | X, \theta) + \log (\theta) \right)
+\end{align}
+%
+Where we have introduced $U(\theta)$ equal to the negative log posterior, because it simplifies analysis.
+Each of these terms is easily computed, by first defining $t_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $y_{ij} = \phi_j(x_i; \theta)$.
+%
+\begin{equation}
+	\log p(b | X, \theta) = \sum_{i=1}^{N} \sum_{j=1}^{B} t_{ij} \log y_{ij}  \quad \textrm{and} \quad
+	\log p(\theta) = -\frac{d}{2} \log 2\pi - \frac{1}{2} || \theta || ^2
+\end{equation}
+%
+The proposal can then be written as:
+%
+\begin{equation}
+	\theta' = \theta - h_t \nabla U(\theta) + \sqrt{2h_t} \cdot \xi_t
+\end{equation}
+%
+Where $\xi_t \sim \Gaussian(0, I)$. As such the proposal distribution, is easy to compute
+%
+\begin{equation}
+	q(\theta, \theta') = \Gaussian \left( \theta' ; \theta - h_t \nabla U(\theta), 2h_t I \right)
+\end{equation}
+%
+The term $\nabla U$ has an easy to compute analytic form. By noting that $\theta = \{w_k\}_{k=1}^{B}$
+%
+\begin{equation}
+	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (t_{ik} - y_{ik}) \Big\} - w_k \right)
+\end{equation}
+%
+. We would then accept these samples with probability
+%
+\begin{equation}
+	\alpha = \min \left( \exp \left(-\Delta U \right) \frac{ q(\theta', \theta)}{q(\theta, \theta')} , 1 \right)
+\end{equation}
 
@@
 
+\subsection{Derivation of gradient with respect to feature parameters}
+
 \end{document}

### 2021-04-18 nips/submission.tex
More work

@@
 \usepackage{tikz}
+\usetikzlibrary{arrows}
 \usepackage{amsmath}
@@
 \begin{abstract}
-  Paper abstract.
+We present a novel generative model for describing vertex-labelled undirected graphs.
+This is a two-layered model.
+First a probabilistic mapping from vertex labels to latent block memberships and from these block memberships we generate a graph according to the micro-canonical stochastic block model.
+
+With this framework we are able to efficiently sample the parameters of the vertex label to block membership mapping.
+This allows us to automatically determine which features have the largest impact on graphical structure.
 \end{abstract}
@@
 \section{Introduction}
+
+There is a wealth of graphical data in the world with more produced every second; social networks, website hyperlinks and academic collaboration are just some examples.
+Many algorithms are used to analyse this data.
+Nevertheless, many of the approaches have been developed in industry to yield empirically good results.
+There is space to develop a principled framework for Bayesian inference on graphical data.
 
@@
 		% nodes
-		\node[roundnode] (X) at (0, 0) {X};
-		\node[squarednode] (b) at (3, 0) {b};
-		\node[roundnode] (A) at (6, 0) {A};
+		\node[roundnode] (X) at (0, 0) {$X$};
+		\node[squarednode] (b) at (3, 0) {$b$};
+		\node[roundnode] (A) at (6, 0) {$A$};
 
@@
 Lastly, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one{b_i = r}$ is the number of vertices assigned to block $r$.
-
 Importantly, we have computable forms for $p(\theta)$ and $p(b, \psi | X)$ which will be useful for performing inference.
@@
 Both of these can be implemented with a Markov Chain through the Metropolis-Hastings algorithm \cite{hastings-alg}.
-We just need to define a proposal distribution $q(x, y)$ for proposing a move $x \rightarrow y$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(\cdot)$, point-wise.
-The proposed move is then accepted with acceptance probability $\alpha$ else it is rejected and we stay at $x$.
+We just need to define a proposal distribution $q(x, x')$ for proposing a move $x \rightarrow x'$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(\cdot)$, point-wise.
+The proposed move is then accepted with probability $\alpha$ (equation \ref{eqn:mh-accept}) else it is rejected and we stay at $x$.
 %
 \begin{equation}
-	\alpha = \min \left( \frac{\pi(y) q(y, x)}{\pi(x) q(x, y)} , 1 \right)
+	\alpha = \min \left( \frac{\pi(x') q(x', x)}{\pi(x) q(x, x')} , 1 \right)
+	\label{eqn:mh-accept}
 \end{equation}
@@
 %
-Which is indeed the distribution we are targeting from equation \ref{eqn:theta-target}.
 The reason we split the Markov chain into two stages is because the summation over all latent states $b \in \mathcal{B}^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in \Bcal^N} p(A | b) P(b | X, \theta)$ is computationally intensive.
+%
+\begin{figure}[!h]
+	\centering
+	\begin{tikzpicture}[
+		roundnode/.style={circle, draw=black, minimum size=12mm},
+		squarednode/.style={rectangle, draw=black, minimum size=12mm}
+		]
+		% nodes
+		\node[roundnode] (b0) at (0, 2) {$b^{(t)}$};
+		\node[roundnode] (b1) at (4, 2) {$b'$};
+		\node[roundnode] (b2) at (8, 2) {$b^{(t+1)}$};
+		\node[roundnode] (t0) at (0, 0) {$\theta^{(t)}$};
+		\node[roundnode] (t1) at (4, 0) {$\theta'$};
+		\node[roundnode] (t2) at (8, 0) {$\theta^{(t+1)}$};
+
+		% arrows
+		\draw[->] (b0) to node[above] {$\sim q_b(b^{(t)}, b' | A)$} (b1);
+		\draw[->] (b1) to node[above] {$\alpha_b (b, b' | A)$} (b2);
+		\draw[->] (b0) [out=30, in=150] to node[above] {$1-\alpha_b$} (b2);
+
+		\draw[->] (t0) to node[above] {$\sim q_\theta(\theta, \theta' | X, b^{(t)})$} (t1);
+		\draw[->] (t1) to node[above] {$\alpha_\theta (\theta, \theta' | X, b^{(t)})$} (t2);
+		\draw[->] (t0) [out=-30, in=-150] to node[below] {$1-\alpha_\theta$} (t2);
+
+		\draw[dashed, ->] (b0) to (2, 0.8);
+		\draw[dashed, ->] (b0) to (5.5, 0.8);
+
+	\end{tikzpicture}
+	\caption{Sampling sequence}
+	\label{fig:samp-sequence}
+\end{figure}
+%
+Figure \ref{fig:samp-sequence} shows an overview of the proposed method.
+We have introduced subscripts and conditionings to make explicit what parameters each step utilises.
+In an important simplification, $p(b| X) = B^{-N}$ which is unchanged no matter the value of X.
+Therefore, we do not need to know the value of $X$ to perform the sampling on $b$.
 
@@
 \begin{equation}
-	S(b) = - \log \pi_b(b) = - \left[ \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \right]
+	S(b) = - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big)
 \end{equation}
@@
 This entropy can be thought of as the description length of the graph because it is the sum of the information required to represent the graph given the parameters and the amount of information required to store the parameters (given the feature matrix $X$).
+The exact, from of the proposal distribution and accept-reject step is explored throughly by \citet{Peixoto-MCMC}.
+There is a widely available library \verb*|graph-tool| \cite{peixoto_graph-tool_2014} implementing this algorithm, which we use.
+The only modification is in the block membership prior $p(b)$ which we replace with $p(b|X)=B^{-N}$ which is a uniform distribution and so does not impact the MH accept-reject step.
 
-\subsection{Inferring feature to block generator}
+\subsection{Sampling feature classifier parameters}
 
@@
 Where we have introduced $U(\theta)$ equal to the negative log posterior, because it simplifies analysis.
-Each of these terms is easily computed, by first defining $t_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $y_{ij} = \phi_j(x_i; \theta)$.
+Each of the constituent terms of$U(\theta)$ is easily computed.
+To simplify notation, we define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} = \phi_j(x_i; \theta)$.
 %
 \begin{equation}
-	\log p(b | X, \theta) = \sum_{i=1}^{N} \sum_{j=1}^{B} t_{ij} \log y_{ij}  \quad \textrm{and} \quad
-	\log p(\theta) = -\frac{d}{2} \log 2\pi - \frac{1}{2} || \theta || ^2
+	\log p(b | X, \theta) = \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
+	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta || ^2
 \end{equation}
 %
-The proposal can then be written as:
+This means that $U(\theta) = H_c(t, y) + \nicefrac{1}{2\sigma_\theta^2} || \theta ||^2$ plus a constant, where $H_c(\cdot || \star)$ is the cross entropy between two distributions and the prior introduces a regularisation term.
+This is a typical objective function used in neural network training.
+In typical applications we only seek the value of $\theta$ that minimises the objective function, which in our case would yield the maximum a posteriori (MAP) estimate.
+This is often done through some kind of gradient descent.
+However, our goal is not to find the MAP estimate but to draw samples from the posterior.
+We therefore adopt the Metropolis Adjusted Langevin Algorithm (first proposed by \citet{mala-tweedie}).
+Given the current sample $\theta$, we propose a new sample $\theta'$ according to equation \ref{eqn:theta-update}
 %
 \begin{equation}
-	\theta' = \theta - h_t \nabla U(\theta) + \sqrt{2h_t} \cdot \xi_t
+	\theta' = \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi
+	\label{eqn:theta-update}
 \end{equation}
 %
-Where $\xi_t \sim \Gaussian(0, I)$. As such the proposal distribution, is easy to compute
+Where $\xi \sim \Gaussian(0, I)$.
+Indeed without the injected noise term, this is equivalent to gradient descent.
+We require the noise term to fully explore the parameter space.
+As such the proposal distribution, is a simple multivariate Gaussian which can be evaluated easily.
 %
 \begin{equation}
-	q(\theta, \theta') = \Gaussian \left( \theta' ; \theta - h_t \nabla U(\theta), 2h_t I \right)
+	q_\theta(\theta, \theta') = \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right)
 \end{equation}
@@
 \begin{equation}
-	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (t_{ik} - y_{ik}) \Big\} - w_k \right)
+	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - w_k \right)
 \end{equation}
 %
-. We would then accept these samples with probability
+After a proposed move is generated, in typical Metropolis-Hastings fashion we accept the move with probability $\alpha$.
 %
 \begin{equation}
-	\alpha = \min \left( \exp \left(-\Delta U \right) \frac{ q(\theta', \theta)}{q(\theta, \theta')} , 1 \right)
+	\alpha(\theta, \theta') = \min \left( \exp \left(-\Delta U \right) \frac{ q(\theta', \theta)}{q(\theta, \theta')} , 1 \right)
 \end{equation}
+
+\subsection{Sampling sequence}
+
+So far each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample.
+This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance.
+$U(\theta)$ is fed in the information $y_{ij}^{(t)} = \one\{b_i^{t} = j\}$.
+We would rather deal with the expectation of this value.
+%
+\begin{equation}
+	\Expect \left[ y_{ij}^{(t)} \right] = \Expect \left[ \one(b_{i}^{(t)} = j) \right]
+	= p(b_i = j)
+\end{equation}
+%
+We obtain an unbiased estimate for this quantity as simply the empirical distribution of the block membership samples $\left\{ b^{(t)} \right\}_{t=1}^T$.
+%
+\begin{equation}
+	\hat{y}_{ij} \coloneqq \frac{1}{T} \sum_{t=1}^{T} y_{ij}^T
+\end{equation}
+%
+This means we no longer need to run the $b$ and $\theta$ Markov chains concurrently.
+Instead, we run the $b$-chain to completion and use it to generate $\hat{y}_{ij}$ for $i \in \{1 \dots N\}$ and $j \in \{1 \dots B\}$.
+This is an estimate of $p(b | A, X)$ that we use for every iteration of the $\theta$ Markov chain.
+This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$ as well as reducing the burn-in time for the $\theta$-chain.
 
@@
 
+We apply the developed methods to a variety of datasets.
+
+\begin{itemize}
+	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier} - This is a network of the author's Facebook friends list.
+Each vertex has been manually labelled with a variety of features describing their relationship to the author.
+	\item \textbf{College football teams} \cite{Evans_2010_football} - network of American college football teams and their interactions. Vertex labels are the division a team belongs to.
+	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.
+\end{itemize}
 \section{Conclusion}
@@
 \subsection{Derivation of gradient with respect to feature parameters}
+\label{appdx:gradu}
 

### 2021-04-19 nips/submission.tex
Proof read 3 and 4

@@
 	\label{defn:microcan-sbm}
-	Let $N \in \Integers^{+}$ denote the number of vertices in our graph.
-The block memberships are denoted by a vector $b$ of length $N$ where each entry $b_i \in \{1, 2 \dots B\}$ where $B$ is the number of nonempty blocks.
+	Let $N \in \Integers^{+}$ denote the number of vertices in a graph.
+The block memberships are encoded by a vector $b$ of length $N$ where each entry $b_i \in \{1, 2 \dots B\}$.
+$B$ is the number of non-empty blocks.
 Let $e$ be a $B \times B$ matrix of edge counts between blocks ($e_{rs}$ is number of edges from block $r$ onto block $s$ - or twice that number if $r=s$).
-We restrict our analysis to undirected graphs so $e$ is symmetric.
+For undirected graphs $e$ is symmetric.
 For a non-degree-corrected stochastic block model (NDC-SBM), we say that the graph $A$ is generated as follows:
@@
 	%
-	Where edges are placed at random but respecting the constraints imposed by $e$ and $b$.
-The additional parameters $N$ and $B$ are omitted as they are inferred from the shape ofs of $b$ and $e$.
-If we interpret $A$ as an adjacency matrix, then this constraint can be written formally as: $e_{rs} = \sum_{i=1}^{N} \sum_{j=1}^{N} A_{ij} \one \{b_i = r\} \one \{b_j = s\}$.
-However, this formulation does not tolerate high degree variation within blocks as is typical of real-world data.
-We therefore introduce the degree-corrected SBM (DC-SBM) which has an additional parameter $k$ which is a vector of length $N$ encoding the degree sequence ($k_i$ is the degree of vertex $i$).
+	Where edges are placed uniformly at random but respecting the constraints imposed by $e$ and $b$.
+The additional parameters $N$ and $B$ are omitted as they are inferred from the shapes of $b$ and $e$.
+If we interpret $A$ as an adjacency matrix, then this constraint can be written formally as: $e_{rs} = \sum_{i,j} A_{ij} \one \{b_i = r\} \one \{b_j = s\}$.
+Nevertheless, this formulation does not accept high degree variability within blocks as is typical of real-world data.
+We therefore introduce the degree-corrected SBM (DC-SBM).
+This has an additional parameter $k$ which is an $N$-length vector encoding the degree sequence ($k_i$ is the degree of vertex $i$).
 Therefore, we write:
@@
 	%
-	This imposes the additional constraint that $k_i = \sum_{j=1}^{N} A_{ij}$. In what follows, we will always assume the degree-corrected model unless otherwise specified.
+	Once again, edges are placed uniformly at random but respecting the constraints imposed by the parameters.
+The DC-SBM has the additional constraint that $k_i = \sum_{j} A_{ij}$.
+In what follows, we will always assume the degree-corrected model unless otherwise specified.
 
@@
 
-The proposed generative model is given in figure \ref{fig:lbgm}.
+The proposed generative model (which we call the latent block generative model - LBGM) is given in figure \ref{fig:lbgm}.
 We start, with the feature matrix X and generate a vector of block memberships $b$.
@@
 We choose a single softmax layer to model $p(b_i | x_i, \theta)$.
-More complex models are possible but then deriving meaning from inferred parameter distributions is complicated.
-Summarising, we can write $p(b | X, \theta)$ as follows:
+More complex models are possible but then deriving meaning from inferred parameter distributions is more difficult.
+Summarising, we write $p(b | X, \theta)$ as follows:
 %
@@
 Where $\tilde{x} \coloneqq \left[x_1, x_2, \dots x_D, 1\right]^T$ is an augmented version of $x$ that allows for a bias term.
-The parameters $\theta$ just contain the $B \times (D+1)$ matrix of weight values or alternatively, $\theta = \{w_k\}_{k=1}^{B}$.
+The parameters for this stage are denoted $\theta$ and consist of the all weight $B$ weight vectors $\theta = \{w_k\}_{k=1}^{B}$.
+Each $w_k$ has dimension $D+1$.
 
@@
 
-Once we have generated the block memberships $b$, we proceed to draw the graph $A$ from the microcanical SBM (\ref{defn:microcan-sbm}) with additional parameters $\psi = \{e, k\}$.
-In a slight abuse of notation we denote the inter-block edge count matrix with $e = \psi_e$ and the degree sequence $k = \psi_k$ to make explicit that these are contained in $\psi$.
+Once we have generated the block memberships $b$, we proceed to draw the graph $A$ from the microcanical DC-SBM (\ref{defn:microcan-sbm}) with additional parameters $\psi = \{e, k\}$.
+In a slight abuse of notation we denote the inter-block edge count matrix with $e = \psi_e$ and the degree sequence $k = \psi_k$ to make explicit that these parameters are contained in $\psi$.
 %
@@
 	p(\theta) = \Gaussian \left( \theta ; 0, \sigma_\theta^2 I \right)
+	\label{eqn:theta-prior}
 \end{equation}
@@
 We will see that this form of prior is equivalent to a regularisation term in neural network training that penalises extreme weight magnitudes.
-As $\sigma_\theta^2 \rightarrow \infty$ this becomes an uninformative uniform prior.
+For $\sigma_\theta^2 \rightarrow \infty$ this becomes an uninformative uniform prior.
 
-In our model $b$ is now an intermediate variable and so we cannot choose a prior.
+In our model, the block memberships vector $b$ is an intermediate latent variable and so we are not free to choose a prior for it.
 The closest thing we can get to a prior is $p(b | X)$.
 As far as inference on the right-hand-side of figure \ref{fig:lbgm}, we regard $p(b | X)$ as a pseudo-prior on $b$.
-We can show that our choice of prior for $p(\theta)$ leads to the following form for $p(b | X)$.
+We can show (appendix \ref{appdx:b|x}) that our choice of prior for $p(\theta)$ (equation \ref{eqn:theta-prior}) leads to the following form for $p(b | X)$.
 %
@@
 %
-In his paper, \citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for the SBM parameters.
-The proposal is to write the joint prior on $(b, e, k)$ as a product of conditionals $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$.
+This is a uniform distribution that only depends on the number of blocks $B$.
+In a nice simplification we find that evaluating $p(b | X)$ does not require knowing the exact value $X$ takes.
+In his paper, \citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for the microcanonical SBM parameters $\psi$.
+In the absence of the feature matrix layer $X$, the idea is to write the joint prior on $(b, e, k)$ as a product of conditionals $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$.
 For our purposes we must insert a conditioning on $X$, to form our pseudo-prior for $b$ and $\psi$.
@@
 %
-Where it is apparent by regarding figure \ref{fig:lbgm} as a Markov model that $(\psi \indep X) | b$.
+Where it is apparent by regarding figure \ref{fig:lbgm} as a Bayesian model that $(\psi \indep X) | b$.
 We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference.
@@
 %
-Where $\specialchoose{n}{m}$ is shorthand for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of histograms (non-negative bin values) with $n$ bins that are constrained to sum to $m$.
-$E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges.
+Where $\specialchoose{n}{m}$ is shorthand for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of distinct histograms that have $n$ with constrained sum $m$.
+$E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges in the graph.
 Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform with respect to $e$.
@@
 Furthermore, $q(m, n)$ is the number of different histograms with at most $n$ non-zero bins that sum to $m$.
-Lastly, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one{b_i = r}$ is the number of vertices assigned to block $r$.
+Lastly, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ is the number of vertices assigned to block $r$.
 Importantly, we have computable forms for $p(\theta)$ and $p(b, \psi | X)$ which will be useful for performing inference.
@@
 
-We are presented with a vertex-labelled graph $(A, X)$. The goal is to draw samples from equation \ref{eqn:theta-target}. However, this is not easily done in practice.
+Suppose we are presented with a vertex-labelled graph $(A, X)$. The goal is to draw samples for $\theta$ according to equation \ref{eqn:theta-target}. However, this is not easily done in practice.
 %
@@
 %
-Both of these can be implemented with a Markov Chain through the Metropolis-Hastings algorithm \cite{hastings-alg}.
+Both of these sampling steps implemented with a Markov Chain through the Metropolis-Hastings algorithm \cite{hastings-alg}.
 We just need to define a proposal distribution $q(x, x')$ for proposing a move $x \rightarrow x'$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(\cdot)$, point-wise.
@@
 We have introduced subscripts and conditionings to make explicit what parameters each step utilises.
-In an important simplification, $p(b| X) = B^{-N}$ which is unchanged no matter the value of X.
+In an important simplification, we note that $p(b| X) = B^{-N}$ which does not depend on the exact value of X.
 Therefore, we do not need to know the value of $X$ to perform the sampling on $b$.
+Conversely, for the $\theta^{(t)}$ samples, we use $b^{(t)}$ but not $A$ as $(\theta \indep A) | b$
 
@@
 
-\citet{Peixoto-MCMC} proposes a Monte Carlo method which we will base our approach on. It relies on writing the posterior in the following form.
+\citet{Peixoto-MCMC} proposes a Monte Carlo method which we will base our approach on. It relies on writing the posterior in the following form:
 %
@@
 %
-Since we are using a microcanonical formulation, there is only one value of $\psi$ which we denote $\psi^*$ that is compatible with the given $(A, b)$ pair.
+Since we are using the microcanonical SBM formulation, there is only one value of $\psi$ that is compatible with the given $(A, b)$ pair.
+We denote this value $\psi^*$.
 Specifically, $k^*_i = \sum_j A_{ij} $ and $e^*_{rs} = \sum_{i, j} A_ij \one \{b_i=r\} \one\{b_j=s\}$.
@@
 The exact, from of the proposal distribution and accept-reject step is explored throughly by \citet{Peixoto-MCMC}.
-There is a widely available library \verb*|graph-tool| \cite{peixoto_graph-tool_2014} implementing this algorithm, which we use.
-The only modification is in the block membership prior $p(b)$ which we replace with $p(b|X)=B^{-N}$ which is a uniform distribution and so does not impact the MH accept-reject step.
+There is a widely available library, \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, implementing this algorithm.
+The only modification we make is in the block membership prior $p(b)$ which we replace with $p(b|X)=B^{-N}$ which is a uniform distribution and so cancels out in the MH accept-reject step.
 
@@
 
-Now the invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. We write this as follows:
+The invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. We write this as follows:
 %
@@
 Where we have introduced $U(\theta)$ equal to the negative log posterior, because it simplifies analysis.
-Each of the constituent terms of$U(\theta)$ is easily computed.
+Each of the constituent terms of $U(\theta)$ is easily computed (equation \ref{eqn:u-form}).
 To simplify notation, we define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} = \phi_j(x_i; \theta)$.
@@
 	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta || ^2
+	\label{eqn:u-form}
 \end{equation}
 %
-This means that $U(\theta) = H_c(t, y) + \nicefrac{1}{2\sigma_\theta^2} || \theta ||^2$ plus a constant, where $H_c(\cdot || \star)$ is the cross entropy between two distributions and the prior introduces a regularisation term.
+This means that $U(\theta) = H_c(t, y) + \nicefrac{1}{2\sigma_\theta^2} || \theta ||^2$ plus a constant, where $H_c(\cdot || \star)$ is the cross entropy between two distributions.
+We see that the prior is equivalent to a regularisation term.
 This is a typical objective function used in neural network training.
-In typical applications we only seek the value of $\theta$ that minimises the objective function, which in our case would yield the maximum a posteriori (MAP) estimate.
+In traditional applications we only seek the value of $\theta$ that minimises the objective function, which in our case would yield the maximum a posteriori (MAP) estimate.
 This is often done through some kind of gradient descent.
@@
 %
-Where $\xi \sim \Gaussian(0, I)$.
+Where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter - which may vary with the sample index.
 Indeed without the injected noise term, this is equivalent to gradient descent.
@@
 %
-The term $\nabla U$ has an easy to compute analytic form. By noting that $\theta = \{w_k\}_{k=1}^{B}$
+The term $\nabla U$ has an easy to compute analytic form. By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as:
 %
@@
 \begin{equation}
-	\hat{y}_{ij} \coloneqq \frac{1}{T} \sum_{t=1}^{T} y_{ij}^T
+	\hat{y}_{ij} \coloneqq \frac{1}{T} \sum_{t=1}^{T} y_{ij}^{(t)}
 \end{equation}
@@
 This is an estimate of $p(b | A, X)$ that we use for every iteration of the $\theta$ Markov chain.
-This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$ as well as reducing the burn-in time for the $\theta$-chain.
+This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$ as well as reducing the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
 

### 2021-05-13 nips/submission.tex
Implement MALA
Improve method section

@@
 
-\title{Automatic feature classification in stochastic block models}
+\title{Inferring community characteristics in labelled networks}
 % The \author macro works with any number of authors. There are two commands
@@
 
-There is a wealth of graphical data in the world with more produced every second; social networks, website hyperlinks and academic collaboration are just some examples.
-Many algorithms are used to analyse this data.
-Nevertheless, many of the approaches have been developed in industry to yield empirically good results.
-There is space to develop a principled framework for Bayesian inference on graphical data.
+There is a wealth of networks in the world.
+Some examples of this graphical data are social networks, website hyperlinks and academic collaboration with more being produced each second.
+It is clear we need tools to analyse this increasingly omni-present form of data.
+
+A somewhat surprising property of real-world graphs is that they exhibit strong community structure.
+In other words, each node will often belong to a cluster of densely connected nodes.
+This property is often exploited by graph compression algorithms and there is high interest in recovering the communities from the observed graph.
+
+A common subset of graphical data is the labelled network.
+This is a graph where we have information about the properties of each node.
+We shall refer to these node properties as features.
+One of the most common questions we can ask of this dataset is what features have the largest impact on the structure of the graph.
+For example, when analysing an academic collaboration graph, one may wish to ask what impact gender has on the structure of the network.
+Nevertheless, this analysis is often vulnerable to confounding variables.
+While gender very well may impact the structure of the graph, there is often a better explanatory variable for the structure.
+
+There is space to bridge the gap between these two approaches.
+Rather than determining whether a feature impacts the graphical structure directly, we can use the community concept as a stepping stone in our analysis.
+Therefore, we extract which features have the largest impact on overall graphical structure.
+This can be thought of as a form of dimensionality-reduction where we only keep the node features that are useful in predicting community memberships.
 
@@
 
-We will be using the microcanonical stochastic block model, proposed by \cite{Peixoto-Bayesian-Microcanonical}. A paraphrased definition is given below.
+This section defines some preliminary concepts required for the subsequent analysis.
+We first need a model for community-like structure in a graphical network.
+For this we adopt the stochastic block model (SBM) - widely used across academia.
+The premise is that each node in the graph belongs to a unique community called a block.
+The probability that two nodes are connected depends only on the block memberships of each.
+Graphs drawn from this ensemble exhibit community structure.
+Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}.
+A paraphrased definition is given below.
 
-\begin{definition}[Microcanonical SBM]
-	\label{defn:microcan-sbm}
-	Let $N \in \Integers^{+}$ denote the number of vertices in a graph.
+\begin{definition}[Microcanonical NDC-SBM]
+	\label{defn:microcan-ndc-sbm}
+	Let $N \in \Integers^{+}$ denote the number of vertices in an undirected graph.
 The block memberships are encoded by a vector $b$ of length $N$ where each entry $b_i \in \{1, 2 \dots B\}$.
-$B$ is the number of non-empty blocks.
+$B \in \Integers^{+}$ is the number of non-empty blocks.
 Let $e$ be a $B \times B$ matrix of edge counts between blocks ($e_{rs}$ is number of edges from block $r$ onto block $s$ - or twice that number if $r=s$).
@@
 	%
-	Where edges are placed uniformly at random but respecting the constraints imposed by $e$ and $b$.
+	Where edges are placed uniformly at random but respecting the constraint imposed by $e$ and $b$.
 The additional parameters $N$ and $B$ are omitted as they are inferred from the shapes of $b$ and $e$.
 If we interpret $A$ as an adjacency matrix, then this constraint can be written formally as: $e_{rs} = \sum_{i,j} A_{ij} \one \{b_i = r\} \one \{b_j = s\}$.
+\end{definition}
+
 Nevertheless, this formulation does not accept high degree variability within blocks as is typical of real-world data.
-We therefore introduce the degree-corrected SBM (DC-SBM).
-This has an additional parameter $k$ which is an $N$-length vector encoding the degree sequence ($k_i$ is the degree of vertex $i$).
-Therefore, we write:
+Indeed, the NDC-SBM favours a partition into high-degree and low-degree nodes rather than clusters of inter-connected nodes.
+We therefore introduce the degree-corrected SBM (DC-SBM) to circumvent these issues.
+
+\begin{definition}[Microcanonical DC-SBM]
+	\label{defn:microcan-dc-sbm}
+	This is much like the NDC-SBM but has an additional parameter $k$ which is an $N$-length vector encoding the degree sequence ($k_i$ is the degree of vertex $i$). Therefore, we write:
 	%
@@
 
+In this section we propose a novel generative model for modelling labelled networks.
 We restrict our analysis to labelled, undirected graphs with $N$ nodes.
@@
 We start, with the feature matrix X and generate a vector of block memberships $b$.
-The parameters of this generator are encapsulated by $\theta$.
+The parameters of this step are encapsulated by $\theta$.
 Each feature vector $x_i$ is treated independently and used to generate the block membership $b_i$.
+Each entry $b_i \in \{1, 2 \dots B\}$ where $B$ is the number of blocks and a constant in our model.
 We choose a single softmax layer to model $p(b_i | x_i, \theta)$.
@@
 Where $\tilde{x} \coloneqq \left[x_1, x_2, \dots x_D, 1\right]^T$ is an augmented version of $x$ that allows for a bias term.
-The parameters for this stage are denoted $\theta$ and consist of the all weight $B$ weight vectors $\theta = \{w_k\}_{k=1}^{B}$.
+The parameters for this stage are denoted $\theta$ and consist of the all $B$ weight vectors $\theta = \{w_k\}_{k=1}^{B}$.
 Each $w_k$ has dimension $D+1$.
+We could instead write the parameters $\theta$ as a $B \times (D+1)$ matrix of weights $W$; this form has use computationally as then $Z_i = W \tilde{x}_i$ is the input to the softmax activation function.
 
@@
 
-Once we have generated the block memberships $b$, we proceed to draw the graph $A$ from the microcanical DC-SBM (\ref{defn:microcan-sbm}) with additional parameters $\psi = \{e, k\}$.
+With the block memberships $b$ generated, we then draw the graph $A$ from the microcanical DC-SBM (\ref{defn:microcan-dc-sbm}) with additional parameters $\psi = \{e, k\}$.
 In a slight abuse of notation we denote the inter-block edge count matrix with $e = \psi_e$ and the degree sequence $k = \psi_k$ to make explicit that these parameters are contained in $\psi$.
@@
 Before performing any inference, we must specify priors on $\theta$ and $\psi$.
-For $\theta$ it seems sensible to choose a Gaussian prior, with zero mean and variance matrix $\sigma^2_\theta I$ such that each element of $\theta$ is independent and distributed like $\sim \Gaussian(0, \sigma)$.
+For $\theta$ it seems sensible to choose a Gaussian prior, with zero mean and variance matrix $\sigma^2_\theta I$ such that each element of $\theta$ is independent and distributed like $\sim \Gaussian(0, \sigma_\theta^2)$.
 In vector form, the prior for $\theta$ is therefore:
@@
 In our model, the block memberships vector $b$ is an intermediate latent variable and so we are not free to choose a prior for it.
-The closest thing we can get to a prior is $p(b | X)$.
-As far as inference on the right-hand-side of figure \ref{fig:lbgm}, we regard $p(b | X)$ as a pseudo-prior on $b$.
-We can show (appendix \ref{appdx:b|x}) that our choice of prior for $p(\theta)$ (equation \ref{eqn:theta-prior}) leads to the following form for $p(b | X)$.
+Nevertheless, as far as inference on the right-hand-side of figure \ref{fig:lbgm}, we regard $p(b | X)$ as a pseudo-prior on $b$.
+We can show (appendix \ref{appdx:b|x}) that our choice of prior for $p(\theta)$ in equation \ref{eqn:theta-prior} leads to a uniform $p(b | X)$ in equation \ref{eqn:b-pseudo-prior}.
 %
@@
 	p(b | X) = \int p(b | X, \theta) p(\theta) d\theta = B^{-N}
+	\label{eqn:b-pseudo-prior}
 \end{equation}
@@
 This is a uniform distribution that only depends on the number of blocks $B$.
-In a nice simplification we find that evaluating $p(b | X)$ does not require knowing the exact value $X$ takes.
-In his paper, \citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for the microcanonical SBM parameters $\psi$.
-In the absence of the feature matrix layer $X$, the idea is to write the joint prior on $(b, e, k)$ as a product of conditionals $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$.
-For our purposes we must insert a conditioning on $X$, to form our pseudo-prior for $b$ and $\psi$.
+This simplifies the problem nicely as evaluating $p(b | X)$ does not require knowing the exact value $X$ takes.
+\citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for the additional microcanonical SBM parameters $\psi$ which we adopt.
+Peixoto's idea is to write the joint prior on $(b, e, k)$ as a product of conditionals $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$.
+For our purposes we must insert a conditioning on $X$, to form our pseudo-prior for $b$ and $\psi$, to give equation \ref{eqn:joint-pseudo-prior}.
 %
@@
 	p(b, \psi | X) = p(b | X) p(\psi | b, X) = p(b | X) p(\psi | b)
+	\label{eqn:joint-pseudo-prior}
 \end{equation}
 %
-Where it is apparent by regarding figure \ref{fig:lbgm} as a Bayesian model that $(\psi \indep X) | b$.
-We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference.
+Where the simplification is made apparent by noting $(\psi \indep X) | b$. We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference.
 %
@@
 %
-Where $\specialchoose{n}{m}$ is shorthand for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of distinct histograms that have $n$ with constrained sum $m$.
+Where $\specialchoose{n}{m}$ is shorthand for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of distinct histograms with $n$ bins under the constraint they sum to $m$.
 $E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges in the graph.
@@
 Furthermore, $q(m, n)$ is the number of different histograms with at most $n$ non-zero bins that sum to $m$.
+$q(m, n)$ is related to but different from $\specialchoose{n}{m}$.
 Lastly, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ is the number of vertices assigned to block $r$.
@@
 
-Suppose we are presented with a vertex-labelled graph $(A, X)$. The goal is to draw samples for $\theta$ according to equation \ref{eqn:theta-target}. However, this is not easily done in practice.
+Now that we have defined the LBGM, we wish to leverage it to perform inference.
+Suppose we are presented with a vertex-labelled graph $(A, X)$; the goal is to draw samples for $\theta$ according to the posterior given the observed graph (equation \ref{eqn:theta-target}).
 %
@@
 	\label{eqn:theta-target}
-	\theta^{(i)} \sim p(\theta | A, X)
+	\theta^{(t)} \sim p(\theta | A, X)
 \end{equation}
 %
+These samples allow us to approximate the posterior distribution for $\theta$ as well as compute a predictive distribution $p(b^* | x^*, A, X) = \int p(b^* | x^*, \theta) p(\theta | X, A) d\theta \approx \frac{1}{T} \sum_{t=1}^{T} p(b^* | x^*, \theta^{(t)})$.
+However, generating these samples is not easily done in practice.
+
 We instead propose an iterative approach.
-First drawing samples $b^{(i)}$ from the block membership posterior (equation \ref{eqn:b-samples}).
+First drawing samples $b^{(t)}$ from the block membership posterior (equation \ref{eqn:b-samples}).
 We then use each $b^{(i)}$ to draw samples for $\theta$ as in equation \ref{eqn:theta-samples}.
@@
 \begin{align*}
-	\Expect_{b^{(i)}} \left[p \left( \theta | X, b^{(i)} \right) \right] &= \sum_{b \in \mathcal{B}^N} p(\theta | X, b) p(b | A, X) \\
+	\Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] &= \sum_{b \in \mathcal{B}^N} p(\theta | X, b) p(b | A, X) \\
 	&= \sum_{b \in \mathcal{B}^N} p(\theta, b | A, X) \\
@@
 %
-The reason we split the Markov chain into two stages is because the summation over all latent states $b \in \mathcal{B}^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in \Bcal^N} p(A | b) P(b | X, \theta)$ is computationally intensive.
+This is an example of a pseudo-marginal approach \cite{pseudo-marginal}.
+Indeed, the unbiased result is sufficient to prove that for sufficient samples, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{t})\right] = p(\theta
+\ A, X)$ which is exactly the distribution we were targetting (equation \ref{eqn:theta-target}).
+
+The reason we split the Markov chain into two stages is because the summation over all latent states $b \in \mathcal{B}^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in \Bcal^N} p(A | b) P(b | X, \theta)$ is intractable $O(B^N)$.
 %
@@
 Therefore, we do not need to know the value of $X$ to perform the sampling on $b$.
-Conversely, for the $\theta^{(t)}$ samples, we use $b^{(t)}$ but not $A$ as $(\theta \indep A) | b$
+Conversely, for the $\theta^{(t)}$ samples, we use $b^{(t)}$ but not $A$ as $(\theta \indep A) | b$.
 
@@
 Since we are using the microcanonical SBM formulation, there is only one value of $\psi$ that is compatible with the given $(A, b)$ pair.
-We denote this value $\psi^*$.
-Specifically, $k^*_i = \sum_j A_{ij} $ and $e^*_{rs} = \sum_{i, j} A_ij \one \{b_i=r\} \one\{b_j=s\}$.
+We denote this value $\psi^* = \{k^*, e^*\}$.
+Specifically, $k^*_i = \sum_j A_{ij} $ and $e^*_{rs} = \sum_{i, j} A_{ij} \one \{b_i=r\} \one\{b_j=s\}$.
 Therefore, the summation over all $\psi$ reduces to just the single $\psi^*$ term.
@@
 
-\subsection{Sampling feature classifier parameters}
+The end result of this is that we can generate a set of block membership samples $\{b^{(t)}\}_{t=1}^{T}$ with each $b^{(t)} \sim p(b | A, X)$.
+Each of these samples can then be used for the $\theta$-chain.
+
+\subsection{Sampling feature-to-block classifier parameters}
 
@@
 \begin{align}
-	p(\theta | X, b) &\propto p(b | X, \theta) p(\theta) = \pi_\theta (\theta) =  \exp \left( - U(\theta) \right) \\
-	\therefore U(\theta) &= - \left( \log p(b | X, \theta) + \log (\theta) \right)
+	p(\theta | X, b) &\propto p(b | X, \theta) p(\theta) = \pi_\theta (\theta) \propto  \exp \left( - U(\theta) \right) \\
+	\therefore U(\theta) &= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}
 \end{align}
 %
-Where we have introduced $U(\theta)$ equal to the negative log posterior, because it simplifies analysis.
+Where we have introduced $U(\theta)$ equal to the negative log posterior.
+This $U(\theta)$ term makes subsequent analysis more concise.
 Each of the constituent terms of $U(\theta)$ is easily computed (equation \ref{eqn:u-form}).
-To simplify notation, we define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} = \phi_j(x_i; \theta)$.
+To simplify notation, we define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
 %
@@
 	\log p(b | X, \theta) = \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
-	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta || ^2
+	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta ||^2
 	\label{eqn:u-form}
@@
 %
-This means that $U(\theta) = H_c(t, y) + \nicefrac{1}{2\sigma_\theta^2} || \theta ||^2$ plus a constant, where $H_c(\cdot || \star)$ is the cross entropy between two distributions.
-We see that the prior is equivalent to a regularisation term.
-This is a typical objective function used in neural network training.
-In traditional applications we only seek the value of $\theta$ that minimises the objective function, which in our case would yield the maximum a posteriori (MAP) estimate.
-This is often done through some kind of gradient descent.
-However, our goal is not to find the MAP estimate but to draw samples from the posterior.
-We therefore adopt the Metropolis Adjusted Langevin Algorithm (first proposed by \citet{mala-tweedie}).
+This means that if we ignore constant terms we can write $U(\theta)$ as in equation \ref{eqn:U-form}.
+We see that the prior effectively introduces a regularisation term.
+Note that $||\theta||^2 = \sum_{i} \theta_{i}^2 = \sum_{j=1}^{B} ||w_j||^2$ is the Euclidean norm of the vector of parameters $\theta$.
+%
+\begin{equation}
+	U(\theta) = \left( \sum_{i=N}^{T} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
+	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
+	\label{eqn:U-form}
+\end{equation}
+%
+$U(\theta)$ in equation \ref{eqn:U-form} appears a typical objective function to be minimised for neural network training.
+The first term is the cross-entropy between actual and predicted labels.
+The second term - introduced by the prior - brings a form of regularisation, preventing over-fitting.
+In traditional applications we only seek the value of $\theta$ that minimises the objective function $U(\theta)$, which in our case would yield the maximum a posteriori (MAP) estimate.
+This is often done through some kind of gradient descent as $\nabla U$ is easily computable (equation \ref{eqn:U-derivative}).
+
+However, our goal is not to find the MAP estimate but to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
+As discussed earlier, given the invariant $\pi_\theta(\cdot)$, it is sufficient to specify a proposal distribution and then apply a MH accept-reject step to ensure detailed balance of the Markov Chain.
+Nevertheless, we can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
+We therefore adopt the Metropolis Adjusted Langevin Algorithm (MALA) - first proposed by \citet{mala-tweedie} - which leverages just that.
 Given the current sample $\theta$, we propose a new sample $\theta'$ according to equation \ref{eqn:theta-update}
@@
 We require the noise term to fully explore the parameter space.
-As such the proposal distribution, is a simple multivariate Gaussian which can be evaluated easily.
+As such the proposal distribution, is a simple multivariate Gaussian which can be easily evaluated.
 %
@@
 \begin{equation}
-	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - w_k \right)
+	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
+	\label{eqn:U-derivative}
 \end{equation}
 %
-After a proposed move is generated, in typical Metropolis-Hastings fashion we accept the move with probability $\alpha$.
+After a proposed move is generated, in typical Metropolis-Hastings fashion we accept the move with probability $\alpha_\theta$.
 %
 \begin{equation}
-	\alpha(\theta, \theta') = \min \left( \exp \left(-\Delta U \right) \frac{ q(\theta', \theta)}{q(\theta, \theta')} , 1 \right)
+	\alpha_\theta(\theta, \theta') = \min \left( 
+	\exp \left( U(\theta) - U(\theta')\right)
+	\frac{ 
+		q_\theta(\theta', \theta)
+	}{
+		q_\theta(\theta, \theta')
+	} 
+	, 1 \right)
 \end{equation}
+%
+This fully specifies, the sampling procedure to generate $\theta^{(t)} \sim p(\theta | X, b^{(t)})$.
+This is a pseudo-marginal approach, meaning that $\theta^{(t)} \sim p(\theta | X, A)$ as $b^{(t)} \sim p(b | A, X)$.
 
@@
 \end{itemize}
+
+\section{Extensions}
+
 \section{Conclusion}

### 2021-05-13 nips/submission.tex
Write grad U appendix

@@
 	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta ||^2
-	\label{eqn:u-form}
+	\label{eqn:U-constituent-terms}
 \end{equation}
@@
 \begin{equation}
-	U(\theta) = \left( \sum_{i=N}^{T} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
+	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
 	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
@@
 %
-The term $\nabla U$ has an easy to compute analytic form. By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as:
+The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}). By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as:
 %
@@
 %
-This fully specifies, the sampling procedure to generate $\theta^{(t)} \sim p(\theta | X, b^{(t)})$.
-This is a pseudo-marginal approach, meaning that $\theta^{(t)} \sim p(\theta | X, A)$ as $b^{(t)} \sim p(b | A, X)$.
+This fully specifies, the sampling procedure to generate $\{\theta^{(t)}\}_{t=1}^T$. Bear in mind that each $\theta^{(t)}$ update step uses its corresponding $b^{(t)}$ block membership sample.
 
@@
 
-So far each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample.
+So far, each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample.
 This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance.
-$U(\theta)$ is fed in the information $y_{ij}^{(t)} = \one\{b_i^{t} = j\}$.
-We would rather deal with the expectation of this value.
+This may lead to longer burn-in and autocorrelation times of the resulting Markov Chain.
+The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends on $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
+We would rather deal with the expectation of each $y_{ij}^{(t)}$:
 %
 \begin{equation}
-	\Expect \left[ y_{ij}^{(t)} \right] = \Expect \left[ \one(b_{i}^{(t)} = j) \right]
-	= p(b_i = j)
+	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one(b_{i}^{(t)} = j) \right]
+	= p(b_i = j | A, X)
 \end{equation}
@@
 \begin{equation}
-	\hat{y}_{ij} \coloneqq \frac{1}{T} \sum_{t=1}^{T} y_{ij}^{(t)}
+	\hat{y}_{ij} \coloneqq \frac{1}{T} \sum_{t=1}^{T} y_{ij}^{(t)} = \frac{1}{T} \sum_{t=1}^{T} \one\{b_i^{(t)} = j\}
 \end{equation}
 %
+We therefore, choose to feed each $\theta^{(t)}$ update step the same $\hat{y}_{ij}$ for all $t$ rather than the corresponding $y^{(t)}_{ij}$.
 This means we no longer need to run the $b$ and $\theta$ Markov chains concurrently.
@@
 This is an estimate of $p(b | A, X)$ that we use for every iteration of the $\theta$ Markov chain.
-This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$ as well as reducing the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
+This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$; wed refer to these as $T_b$ and $T_\theta$ henceforth.
+Furthermore, this changeover reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
 
@@
 
-\subsection{Derivation of gradient with respect to feature parameters}
+\subsection{Derivation of U gradient with respect to feature parameters}
 \label{appdx:gradu}
+The goal is to determine $\nabla U(\theta)$, the gradient of the negative log posterior with respect to the parameters. We repeat the form of $U(\theta)$ in equation \ref{eqn:U-form-appdx}.
+%
+\begin{equation}
+	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
+	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
+	\label{eqn:U-form-appdx}
+\end{equation}
+%
+Where $y_{ij}$ is independent of $\theta$ and $a_{ij}$ is the output from the softmax layer, with form as given in equation \ref{eqn:a-ij}.
+%
+\begin{equation}
+	a_{ij} \coloneqq \phi_{j} (x_i; \theta) = \frac{\exp(w_j^T \tilde{x}_i)}{\sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i)}
+	\label{eqn:a-ij} 
+\end{equation}
+%
+We note that $\theta = \{w_k\}_{k=1}^B$, and as such we can write this in vector form $\theta = \left[w_1^T, w_2^T \dots w_B^T  \right]^T$.
+Therefore, $\nabla U(\theta) = \left[\nicefrac{\partial U}{\partial w_1}^T,\nicefrac{\partial U}{\partial w_2}^T \dots \nicefrac{\partial U}{\partial w_B}^T  \right]^T$; to compute $\nabla U(\theta)$ it suffices to find the form of $\nicefrac{\partial U}{\partial w_k}$ with respect to a general $k$.
+
+To this end, we must first find partial derivatives of $a_{ij}$ and $||\theta||$ with respect to $w_k$. Starting with $a_{ij}$:
+%
+\begin{align}
+	\frac{\partial a_{ij}}{\partial w_k} &= \frac
+	{\tilde{x}_i \exp(w_j^T \tilde{x}_i) \delta_{jk} \cdot \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) 
+		- 
+	\exp(w_j^T \tilde{x}_i) \cdot \tilde{x}_i \exp(w_k^T \tilde{x}_i)}
+	{\left( \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) \right)^2} \nonumber \\
+	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right) 
+\end{align}
+%
+Where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$. Now moving onto the derivative of $||\theta||^2$:
+%
+\begin{equation}
+	\frac{ \partial}{\partial w_k} ||\theta||^2 = \frac{\partial}{\partial w_k} \left( \sum_{b=1}^B ||w_b||^2 \right) = 2w_k
+\end{equation}
+%
+We are ready to put this all together, to find the partial derivative of $U(\theta)$ with respect to each $w_k$:
+\begin{align}
+	\frac{\partial U}{\partial w_k} &= 
+	\sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} 
+	\left( \frac{-\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
+	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
+	&=  - \left( \sum_{i=1}^{N} \tilde{x}_i \left( y_{ik} - a_{ik} \sum_{j=1}^{B} y_{ij} \right)
+	- \frac{w_k}{\sigma_\theta^2} \right) \nonumber \\
+	&= - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
+\end{align}
+%
+This is the required result.
+This form can be computed efficiently through matrix operations.
+The only property of $y_{ij}$ we have used in the derivation is the sum-to-one constraint $\sum_{j=1}^{B} y_{ij} = 1$ for all $i$.
 

### 2021-05-13 nips/submission.tex
Scaffold algorithms

@@
 \usepackage{amssymb}
+\usepackage{algorithmic}
+\usepackage{algorithm}
 
@@
 
+\begin{algorithm} % enter the algorithm environment
+	\caption{Block membership sample generation} % give the algorithm a caption
+	\label{alg:b-samples} % and a label for \ref{} commands later in the document
+	\begin{algorithmic} % enter the algorithmic environment
+
+		\FOR{$t \in \{0, 1 \dots T_b - 1\}$}
+			\STATE $b' \gets \sim q_b(b^{(t)}, b' | A)$
+			\STATE $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
+			\STATE $u \gets \sim \textrm{Unif}(0,1)$
+			\IF{$\log u < \log \alpha_b$}
+				\STATE $b^{(t+1)} \gets b'$
+			\ELSE
+				\STATE $b^{(t+1)} \gets b^{(t)}$
+			\ENDIF
+		\ENDFOR
+
+		\STATE \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$
+
+	\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm} % enter the algorithm environment
+	\caption{LBGM parameter pseudo-marginal inference} % give the algorithm a caption
+	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
+	\begin{algorithmic} % enter the algorithmic environment
+
+		\STATE $\mathcal{S} \gets \textrm{range}(\textrm{burn-in}, T_b, \textrm{thinning})$
+		\STATE $\hat{Y}_{ij} \gets \frac{1}{\mathcal{|S|}} \sum_{t \in \mathcal{S}} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
+
+		\item[]
+
+		\FOR{$t \in \{0, 1 \dots T_\theta - 1\}$}
+		\STATE $\theta' \gets \sim q_\theta(\theta^{(t)}, \theta' | X, \hat{Y})$
+		\STATE $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
+		\STATE $u \gets \sim \textrm{Unif}(0,1)$
+		\IF{$\log u < \log \alpha_\theta$}
+		\STATE $\theta^{(t+1)} \gets \theta'$
+		\ELSE
+		\STATE $\theta^{(t+1)} \gets \theta^{(t)}$
+		\ENDIF
+		\ENDFOR
+
+		\STATE \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$
+
+	\end{algorithmic}
+\end{algorithm}
+
 \section{Experiments}

### 2021-05-14 nips/submission.tex
More work on twitch data

@@
 
+\subsection{Complexity Analysis}
+
 \section{Experiments}
@@
 \begin{itemize}
-	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier} - This is a network of the author's Facebook friends list.
-Each vertex has been manually labelled with a variety of features describing their relationship to the author.
 	\item \textbf{College football teams} \cite{Evans_2010_football} - network of American college football teams and their interactions. Vertex labels are the division a team belongs to.
 	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.
+	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier} - network of the author's Facebook friends list.
+Each vertex has been manually labelled with a variety of features describing their relationship to the author.
+	\item \textbf{Twitch users} \cite{twitch} - a network of user-user friendships on the streaming service Twitch.
+Vertex labels are extracted according to video-games played, location and streaming habits.
+This dataset is also broken down into disjoint networks according to language.
+We only consider the English users with is a subnet with $N=7126$ vertices and $E=35324$ edges).
 \end{itemize}

### 2021-05-14 nips/submission.tex
More twitch work

@@
 			\STATE $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
-			\STATE $u \gets \sim \textrm{Unif}(0,1)$
-			\IF{$\log u < \log \alpha_b$}
+			\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
+			\IF{$\log \eta < \log \alpha_b$}
 				\STATE $b^{(t+1)} \gets b'$
@@
 		\FOR{$t \in \{0, 1 \dots T_\theta - 1\}$}
-		\STATE $\theta' \gets \sim q_\theta(\theta^{(t)}, \theta' | X, \hat{Y})$
+		\STATE $\xi \gets \sim \Gaussian(0, I)$
+		\STATE $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
 		\STATE $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
-		\STATE $u \gets \sim \textrm{Unif}(0,1)$
-		\IF{$\log u < \log \alpha_\theta$}
+		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
+		\IF{$\log \eta < \log \alpha_\theta$}
 		\STATE $\theta^{(t+1)} \gets \theta'$
@@
 
+\subsection{Choosing the MALA step-size}
+\label{appdx:step-size}
+
+For sampling from the $\theta$-chain of the block membership generator parameters, we employed the Metropolis Adjusted Langevin Algorithm (MALA). At iteration $t$, the proposed sample is generated by:
+%
+\begin{equation}
+	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi
+\end{equation}
+%
+There are two competing objectives when choosing the step-size $h_t$.
+On the one hand, we want the step-size to be large so that we arrive at a high density region quickly.
+However, too large a step-size will lead to a lower acceptance ratio and thus inefficient sampling.
+A solution to this problem would be to slowly decrease the step-size with $t$ - often called simulated annealing.
+Therefore, we still have a short burn-in time but will not bounce around the mode for large $t$.
+As well as the trivial constraint for $h_t$ to be strictly positive and we introduce two further constraints as outlined by \citet{Bayesian-SGLD}:
+%
+\begin{equation}
+	\sum_{t=1}^{\infty} h_t = \infty \qquad \textrm{and} \qquad
+	\sum_{t=1}^{\infty} h_t^2 < \infty
+	\label{eqn:h-constraints}
+\end{equation}
+%
+The first constraint ensures that we have cover sufficient distance to arrive at any arbitrary point in our domain, no matter the starting point.
+The second constraint ensures that once we converge to the mode rather than simply bouncing around it.
+\citet{Bayesian-SGLD} propose the following form for a polynomially decaying step-size which we adopt:
+%
+\begin{equation}
+	h_t = a(b + t)^{-\gamma}
+\end{equation}
+%
+Where $a, b, \gamma$ are hyper-parameters to be chosen.
+We require $a,b > 0$ and $\gamma \in (0.5, 1]$ to satisfy equation \ref{eqn:h-constraints}.
+We find empirically that $a=250, b=1000, \gamma=0.8$ yield good burn-in times for our MALA implementations.
+
 \end{document}

### 2021-05-15 nips/submission.tex
school work

@@
 \usepackage{algorithm}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{graphicx}
+
+%images
+\graphicspath{{./img/}}
 
@@
 
-We apply the developed methods to a variety of datasets.
+We apply the developed methods to a variety of datasets, to show the power of the method across different scenarios.
+We choose a range of node counts $N$, edge counts $E$ and feature space dimension $D$.
 
 \begin{itemize}
+	\item \textbf{Primary school dynamic contacts} \cite{schools} ($N=238, E=5539, D=13$) - network of face-to-face contacts amongst students and teachers at a primary school in Lyon, France.
+Two nodes are connected if the two parties shared a face-to-face interaction over the course of the day.
+Vertex features include class membership, gender and whether or not the student is a teacher or pupil.
+These data were collected on consecutive days in October 2009.
+We choose to analyse just the second day.
 	\item \textbf{College football teams} \cite{Evans_2010_football} - network of American college football teams and their interactions. Vertex labels are the division a team belongs to.
@@
 \end{itemize}
+
+\subsection{Primary school dynamic contacts}
+
+These data were originally collected by \citet{schools} to quantify the transmission opportunities for respiratory infections within a primary school context (ages 6-11 in France).
+However, we seek to ask a simpler question.
+What features best describe how people interact with one another in a primary school context.
+The only vertex features we have available are school-class (one of 10 values - 2 per year group), gender and a distinction between teachers and pupils.
+
+We must first choose the number of blocks to consider $B$ to define the coarseness of our analysis.
+A total of 10 classes would suggest that $B=10$ is a natural starting point.
+With this value of $B$ we first sample from the block membership posterior $b^{(t)} \sim p(b | A, X)$.
+We visualise the inferred partition in figure \ref{fig:school-graph}
+
+\begin{figure}[!h]
+	\centering
+	\begin{subfigure}{0.4\textwidth}
+		\centering
+		\includegraphics{width=\textwidth}{school-graph.png}
+		\caption{Graph with nodes coloured to show soft inferred block memberships $\hat{Y}_ij \approx p(b_i=j|A, X)$}
+	\end{subfigure}
+\end{figure}
 

### 2021-05-15 nips/submission.tex
Custom legend

@@
 This $U(\theta)$ term makes subsequent analysis more concise.
-Each of the constituent terms of $U(\theta)$ is easily computed (equation \ref{eqn:u-form}).
+Each of the constituent terms of $U(\theta)$ is easily computed (equation \ref{eqn:U-constituent-terms}).
 To simplify notation, we define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
@@
 
+\subsection{Choosing the number of blocks}
+
+For the purposes of our model (the LBGM), the number of blocks $B$ is a constant which must be specified by the data scientist.
+We could however, allow our choice of $B$ to be influenced by the observed data.
+This places us in the domain of empirical Bayes, which must be negotiated carefully.
+Prior beliefs must be determined a priori else they are not prior.
+However, as the number of blocks only specifies the coarseness of the analysis, it is fine to allow it to vary.
+Indeed, \citet{peixoto-determine-B} shows that for a fixed average degree the maximum number of detectable blocks scales as $O(\sqrt{N})$ where $N$ is the number of vertices.
+
+If we allow $B$ to vary in the $b$-chain (i.e.
+new blocks can be created and we permit empty blocks) then it can be run  until a minimum description length (MDL) solution is reached.
+We take the number of non-empty blocks at the MDL to be our fixed block number $B$ for subsequent analysis.
+Indeed, it is prudent to start our $b$-chain at this MDL solution as then the burn-in time is greatly reduced.
+
 \section{Experiments}
@@
 
+A table summarising the results is given in table \ref{tab:results}.
+
+\begin{table}[]
+	\centering
+	\begin{tabular}{c|ccc|c|cc}
+		Dataset & $N$ & $E$ & $D$ & $B$ & $\bar{S}_b /N$ & $\bar{U}_\theta /N$ \\ \hline
+		Primary school & 238 & 5539 & 13 &  18 & 43.0  & 2.75 \\
+		Maier FB egonet &     &     &     &   &           &             \\
+		&     &     &     &              &     &       
+	\end{tabular}
+\end{table}
+
 \subsection{Primary school dynamic contacts}
@@
 	\centering
-	\begin{subfigure}{0.4\textwidth}
+	\begin{subfigure}{0.45\linewidth}
 		\centering
-		\includegraphics{width=\textwidth}{school-graph.png}
-		\caption{Graph with nodes coloured to show soft inferred block memberships $\hat{Y}_ij \approx p(b_i=j|A, X)$}
+		\includegraphics[width=\linewidth]{school-graph.png}
+		\caption{Softmax weighting samples for each block index}
+		\label{fig:school-graph}
 	\end{subfigure}
+	\hfill
+	\begin{subfigure}{0.45\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{school-null.png}
+		\caption{}
+		\label{fig:school-null}
+	\end{subfigure}
+	\caption{Primary school dynamic contacts network \cite{schools}}
 \end{figure}

### 2021-05-16 nips/ack.tex
Split into multiple tex files

--- 2021-05-14 nips/submission.tex
@@
-\documentclass{article}
-
-% if you need to pass options to natbib, use, e.g.:
-%     \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading neurips_2021
-
-% ready for submission
-\PassOptionsToPackage{numbers, compress}{natbib}
-\usepackage[final]{neurips_2021}
-
-% to compile a preprint version, e.g., for submission to arXiv, add add the
-% [preprint] option:
-%     \usepackage[preprint]{neurips_2021}
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-%\usepackage[final]{neurips_2021}
-
-% to avoid loading the natbib package, add option nonatbib:
-%    \usepackage[nonatbib]{neurips_2021}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{xcolor}         % colors
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-\usepackage{amsmath}
-\usepackage{mathtools}
-\usepackage{bbm}
-\usepackage{amssymb}
-\usepackage{algorithmic}
-\usepackage{algorithm}
-
-%bibliography
-\bibliographystyle{plainnat}
-
-\title{Inferring community characteristics in labelled networks}
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to break the
-% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
-% authors names on the first line, and the last on the second line, try using
-% \AND instead of \And before the third author name.
-
-\author{
-  Lawrence Tray \\
-  Department of Engineering \\
-  University of Cambridge \\
-  \texttt{lpt30@cam.ac.uk} \\
-  % examples of more authors
-  \And
-  Ioannis Kontoyiannis \\
-  Department of Mathematics \\
-  University of Cambridge \\
-  \texttt{ik355@cam.ac.uk}
-}
-
-%custom commands
-\newcommand{\Xcal}{\mathcal{X}}
-\newcommand{\Bcal}{\mathcal{B}}
-\newcommand{\Integers}{\mathbb{Z}}
-\newcommand{\one}{\mathbbm{1}}
-\newcommand{\Gaussian}{\mathcal{N}}
-\newcommand{\indep}{\perp \!\!\! \perp}
-\newcommand{\specialchoose}{\genfrac{\{}{\}}{0pt}{}}
-\newcommand{\Expect}{\mathbb{E}}
-
-%envs
-\newtheorem{definition}{Definition}[section]
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[theorem]
-\newtheorem{lemma}[theorem]{Lemma}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-We present a novel generative model for describing vertex-labelled undirected graphs.
-This is a two-layered model.
-First a probabilistic mapping from vertex labels to latent block memberships and from these block memberships we generate a graph according to the micro-canonical stochastic block model.
-
-With this framework we are able to efficiently sample the parameters of the vertex label to block membership mapping.
-This allows us to automatically determine which features have the largest impact on graphical structure.
-\end{abstract}
-
-\section{Introduction}
-
-There is a wealth of networks in the world.
-Some examples of this graphical data are social networks, website hyperlinks and academic collaboration with more being produced each second.
-It is clear we need tools to analyse this increasingly omni-present form of data.
-
-A somewhat surprising property of real-world graphs is that they exhibit strong community structure.
-In other words, each node will often belong to a cluster of densely connected nodes.
-This property is often exploited by graph compression algorithms and there is high interest in recovering the communities from the observed graph.
-
-A common subset of graphical data is the labelled network.
-This is a graph where we have information about the properties of each node.
-We shall refer to these node properties as features.
-One of the most common questions we can ask of this dataset is what features have the largest impact on the structure of the graph.
-For example, when analysing an academic collaboration graph, one may wish to ask what impact gender has on the structure of the network.
-Nevertheless, this analysis is often vulnerable to confounding variables.
-While gender very well may impact the structure of the graph, there is often a better explanatory variable for the structure.
-
-There is space to bridge the gap between these two approaches.
-Rather than determining whether a feature impacts the graphical structure directly, we can use the community concept as a stepping stone in our analysis.
-Therefore, we extract which features have the largest impact on overall graphical structure.
-This can be thought of as a form of dimensionality-reduction where we only keep the node features that are useful in predicting community memberships.
-
-\section{Preliminaries}
-
-This section defines some preliminary concepts required for the subsequent analysis.
-We first need a model for community-like structure in a graphical network.
-For this we adopt the stochastic block model (SBM) - widely used across academia.
-The premise is that each node in the graph belongs to a unique community called a block.
-The probability that two nodes are connected depends only on the block memberships of each.
-Graphs drawn from this ensemble exhibit community structure.
-Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}.
-A paraphrased definition is given below.
-
-\begin{definition}[Microcanonical NDC-SBM]
-	\label{defn:microcan-ndc-sbm}
-	Let $N \in \Integers^{+}$ denote the number of vertices in an undirected graph.
-The block memberships are encoded by a vector $b$ of length $N$ where each entry $b_i \in \{1, 2 \dots B\}$.
-$B \in \Integers^{+}$ is the number of non-empty blocks.
-Let $e$ be a $B \times B$ matrix of edge counts between blocks ($e_{rs}$ is number of edges from block $r$ onto block $s$ - or twice that number if $r=s$).
-For undirected graphs $e$ is symmetric.
-For a non-degree-corrected stochastic block model (NDC-SBM), we say that the graph $A$ is generated as follows:
-	%
-	\begin{equation}
-		A \sim \textrm{NDC-SBM}_{\textrm{MC}} (b, e)
-	\end{equation}
-	%
-	Where edges are placed uniformly at random but respecting the constraint imposed by $e$ and $b$.
-The additional parameters $N$ and $B$ are omitted as they are inferred from the shapes of $b$ and $e$.
-If we interpret $A$ as an adjacency matrix, then this constraint can be written formally as: $e_{rs} = \sum_{i,j} A_{ij} \one \{b_i = r\} \one \{b_j = s\}$.
-\end{definition}
-
-Nevertheless, this formulation does not accept high degree variability within blocks as is typical of real-world data.
-Indeed, the NDC-SBM favours a partition into high-degree and low-degree nodes rather than clusters of inter-connected nodes.
-We therefore introduce the degree-corrected SBM (DC-SBM) to circumvent these issues.
-
-\begin{definition}[Microcanonical DC-SBM]
-	\label{defn:microcan-dc-sbm}
-	This is much like the NDC-SBM but has an additional parameter $k$ which is an $N$-length vector encoding the degree sequence ($k_i$ is the degree of vertex $i$). Therefore, we write:
-	%
-	\begin{equation}
-		A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, e, k)
-	\end{equation}
-	%
-	Once again, edges are placed uniformly at random but respecting the constraints imposed by the parameters.
-The DC-SBM has the additional constraint that $k_i = \sum_{j} A_{ij}$.
-In what follows, we will always assume the degree-corrected model unless otherwise specified.
-
-\end{definition}
-
-\section{Latent block generative model}
-
-In this section we propose a novel generative model for modelling labelled networks.
-We restrict our analysis to labelled, undirected graphs with $N$ nodes.
-We define the vector $x_i \in \Xcal^D$ as the feature vector for the $i$'th vertex.
-Each vertex has $D$ total features and we assume all entries take values from the same set $\Xcal$.
-For the majority of datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$.
-The feature vectors $\{x_i\}_{i=1}^{N}$ are subsumed into the $N \times D$ matrix $X$.
-
-The proposed generative model (which we call the latent block generative model - LBGM) is given in figure \ref{fig:lbgm}.
-We start, with the feature matrix X and generate a vector of block memberships $b$.
-The parameters of this step are encapsulated by $\theta$.
-Each feature vector $x_i$ is treated independently and used to generate the block membership $b_i$.
-Each entry $b_i \in \{1, 2 \dots B\}$ where $B$ is the number of blocks and a constant in our model.
-We choose a single softmax layer to model $p(b_i | x_i, \theta)$.
-More complex models are possible but then deriving meaning from inferred parameter distributions is more difficult.
-Summarising, we write $p(b | X, \theta)$ as follows:
-%
-\begin{equation}
-	p(b| X, \theta) = \prod_{i=1}^{N} p(b_i | x_i, \theta) = \prod_{i=1}^{N} \phi_{b_i} (x_i; \theta)
-	= \prod_{i=1}^{N} \frac{\exp\left(w_{b_i}^T \tilde{x}_i\right)}{\sum_{k=1}^{B} \exp \left( w_k^T \tilde{x}_i\right)}
-\end{equation}
-%
-Where $\tilde{x} \coloneqq \left[x_1, x_2, \dots x_D, 1\right]^T$ is an augmented version of $x$ that allows for a bias term.
-The parameters for this stage are denoted $\theta$ and consist of the all $B$ weight vectors $\theta = \{w_k\}_{k=1}^{B}$.
-Each $w_k$ has dimension $D+1$.
-We could instead write the parameters $\theta$ as a $B \times (D+1)$ matrix of weights $W$; this form has use computationally as then $Z_i = W \tilde{x}_i$ is the input to the softmax activation function.
-
-\begin{figure}[!h]
-	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=7mm},
-		squarednode/.style={rectangle, draw=black, minimum size=7mm}
-		]
-		% nodes
-		\node[roundnode] (X) at (0, 0) {$X$};
-		\node[squarednode] (b) at (3, 0) {$b$};
-		\node[roundnode] (A) at (6, 0) {$A$};
-
-		% arrows
-		\draw[->] (X.east) -- node[above] {$\theta$} (b.west);
-		\draw[->] (b.east) -- node[above] {$\psi$}(A.west);
-	\end{tikzpicture}
-	\caption{Latent block generative model}
-	\label{fig:lbgm}
-\end{figure}
-
-With the block memberships $b$ generated, we then draw the graph $A$ from the microcanical DC-SBM (\ref{defn:microcan-dc-sbm}) with additional parameters $\psi = \{e, k\}$.
-In a slight abuse of notation we denote the inter-block edge count matrix with $e = \psi_e$ and the degree sequence $k = \psi_k$ to make explicit that these parameters are contained in $\psi$.
-%
-\begin{equation}
-	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k)
-\end{equation}
-
-\subsection{Prior selection}
-
-Before performing any inference, we must specify priors on $\theta$ and $\psi$.
-For $\theta$ it seems sensible to choose a Gaussian prior, with zero mean and variance matrix $\sigma^2_\theta I$ such that each element of $\theta$ is independent and distributed like $\sim \Gaussian(0, \sigma_\theta^2)$.
-In vector form, the prior for $\theta$ is therefore:
-%
-\begin{equation}
-	p(\theta) = \Gaussian \left( \theta ; 0, \sigma_\theta^2 I \right)
-	\label{eqn:theta-prior}
-\end{equation}
-%
-We will see that this form of prior is equivalent to a regularisation term in neural network training that penalises extreme weight magnitudes.
-For $\sigma_\theta^2 \rightarrow \infty$ this becomes an uninformative uniform prior.
-
-In our model, the block memberships vector $b$ is an intermediate latent variable and so we are not free to choose a prior for it.
-Nevertheless, as far as inference on the right-hand-side of figure \ref{fig:lbgm}, we regard $p(b | X)$ as a pseudo-prior on $b$.
-We can show (appendix \ref{appdx:b|x}) that our choice of prior for $p(\theta)$ in equation \ref{eqn:theta-prior} leads to a uniform $p(b | X)$ in equation \ref{eqn:b-pseudo-prior}.
-%
-\begin{equation}
-	p(b | X) = \int p(b | X, \theta) p(\theta) d\theta = B^{-N}
-	\label{eqn:b-pseudo-prior}
-\end{equation}
-%
-This is a uniform distribution that only depends on the number of blocks $B$.
-This simplifies the problem nicely as evaluating $p(b | X)$ does not require knowing the exact value $X$ takes.
-\citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for the additional microcanonical SBM parameters $\psi$ which we adopt.
-Peixoto's idea is to write the joint prior on $(b, e, k)$ as a product of conditionals $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$.
-For our purposes we must insert a conditioning on $X$, to form our pseudo-prior for $b$ and $\psi$, to give equation \ref{eqn:joint-pseudo-prior}.
-%
-\begin{equation}
-	p(b, \psi | X) = p(b | X) p(\psi | b, X) = p(b | X) p(\psi | b)
-	\label{eqn:joint-pseudo-prior}
-\end{equation}
-%
-Where the simplification is made apparent by noting $(\psi \indep X) | b$. We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference.
-%
-\begin{equation}
-	p(\psi | b) = p(e | b) p(k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
-	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right]
-\end{equation}
-%
-Where $\specialchoose{n}{m}$ is shorthand for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of distinct histograms with $n$ bins under the constraint they sum to $m$.
-$E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges in the graph.
-Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform with respect to $e$.
-The variable $\eta_j^r$ is introduced to denote the number of vertices in block $r$ that have degree $j$.
-Formally, $\eta_j^r \coloneqq \sum_{i} \one\left\{b_i = r \right\} \one \left\{k_i = j \right\}$.
-Furthermore, $q(m, n)$ is the number of different histograms with at most $n$ non-zero bins that sum to $m$.
-$q(m, n)$ is related to but different from $\specialchoose{n}{m}$.
-Lastly, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ is the number of vertices assigned to block $r$.
-Importantly, we have computable forms for $p(\theta)$ and $p(b, \psi | X)$ which will be useful for performing inference.
-
-\section{Inference}
-
-Now that we have defined the LBGM, we wish to leverage it to perform inference.
-Suppose we are presented with a vertex-labelled graph $(A, X)$; the goal is to draw samples for $\theta$ according to the posterior given the observed graph (equation \ref{eqn:theta-target}).
-%
-\begin{equation}
-	\label{eqn:theta-target}
-	\theta^{(t)} \sim p(\theta | A, X)
-\end{equation}
-%
-These samples allow us to approximate the posterior distribution for $\theta$ as well as compute a predictive distribution $p(b^* | x^*, A, X) = \int p(b^* | x^*, \theta) p(\theta | X, A) d\theta \approx \frac{1}{T} \sum_{t=1}^{T} p(b^* | x^*, \theta^{(t)})$.
-However, generating these samples is not easily done in practice.
-
-We instead propose an iterative approach.
-First drawing samples $b^{(t)}$ from the block membership posterior (equation \ref{eqn:b-samples}).
-We then use each $b^{(i)}$ to draw samples for $\theta$ as in equation \ref{eqn:theta-samples}.
-%
-\begin{align}
-	b^{(i)} &\sim p \Big( b | A, X \Big)  \label{eqn:b-samples}\\
-	\theta^{(i)} &\sim p\Big(\theta | X, b^{(i)} \Big) \label{eqn:theta-samples}
-\end{align}
-%
-Both of these sampling steps implemented with a Markov Chain through the Metropolis-Hastings algorithm \cite{hastings-alg}.
-We just need to define a proposal distribution $q(x, x')$ for proposing a move $x \rightarrow x'$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(\cdot)$, point-wise.
-The proposed move is then accepted with probability $\alpha$ (equation \ref{eqn:mh-accept}) else it is rejected and we stay at $x$.
-%
-\begin{equation}
-	\alpha = \min \left( \frac{\pi(x') q(x', x)}{\pi(x) q(x, x')} , 1 \right)
-	\label{eqn:mh-accept}
-\end{equation}
-%
-This accept-reject step ensures the resulting Markov Chain is in detailed balance with the target distribution $\pi(\cdot)$.
-What we propose in equations \ref{eqn:b-samples} and \ref{eqn:theta-samples} is therefore implemented through a 2-level Markov chain.
-The resulting samples for $\theta^{(i)}$ are unbiased in the sense that the expectation of their distribution is the posterior we are targeting in equation \ref{eqn:theta-target}.
-%
-\begin{align*}
-	\Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] &= \sum_{b \in \mathcal{B}^N} p(\theta | X, b) p(b | A, X) \\
-	&= \sum_{b \in \mathcal{B}^N} p(\theta, b | A, X) \\
-	&= p(\theta | A, X) \\
-\end{align*}
-%
-This is an example of a pseudo-marginal approach \cite{pseudo-marginal}.
-Indeed, the unbiased result is sufficient to prove that for sufficient samples, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{t})\right] = p(\theta
-\ A, X)$ which is exactly the distribution we were targetting (equation \ref{eqn:theta-target}).
-
-The reason we split the Markov chain into two stages is because the summation over all latent states $b \in \mathcal{B}^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in \Bcal^N} p(A | b) P(b | X, \theta)$ is intractable $O(B^N)$.
-%
-\begin{figure}[!h]
-	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=12mm},
-		squarednode/.style={rectangle, draw=black, minimum size=12mm}
-		]
-		% nodes
-		\node[roundnode] (b0) at (0, 2) {$b^{(t)}$};
-		\node[roundnode] (b1) at (4, 2) {$b'$};
-		\node[roundnode] (b2) at (8, 2) {$b^{(t+1)}$};
-		\node[roundnode] (t0) at (0, 0) {$\theta^{(t)}$};
-		\node[roundnode] (t1) at (4, 0) {$\theta'$};
-		\node[roundnode] (t2) at (8, 0) {$\theta^{(t+1)}$};
-
-		% arrows
-		\draw[->] (b0) to node[above] {$\sim q_b(b^{(t)}, b' | A)$} (b1);
-		\draw[->] (b1) to node[above] {$\alpha_b (b, b' | A)$} (b2);
-		\draw[->] (b0) [out=30, in=150] to node[above] {$1-\alpha_b$} (b2);
-
-		\draw[->] (t0) to node[above] {$\sim q_\theta(\theta, \theta' | X, b^{(t)})$} (t1);
-		\draw[->] (t1) to node[above] {$\alpha_\theta (\theta, \theta' | X, b^{(t)})$} (t2);
-		\draw[->] (t0) [out=-30, in=-150] to node[below] {$1-\alpha_\theta$} (t2);
-
-		\draw[dashed, ->] (b0) to (2, 0.8);
-		\draw[dashed, ->] (b0) to (5.5, 0.8);
-
-	\end{tikzpicture}
-	\caption{Sampling sequence}
-	\label{fig:samp-sequence}
-\end{figure}
-%
-Figure \ref{fig:samp-sequence} shows an overview of the proposed method.
-We have introduced subscripts and conditionings to make explicit what parameters each step utilises.
-In an important simplification, we note that $p(b| X) = B^{-N}$ which does not depend on the exact value of X.
-Therefore, we do not need to know the value of $X$ to perform the sampling on $b$.
-Conversely, for the $\theta^{(t)}$ samples, we use $b^{(t)}$ but not $A$ as $(\theta \indep A) | b$.
-
-\subsection{Sampling block memberships}
-
-\citet{Peixoto-MCMC} proposes a Monte Carlo method which we will base our approach on. It relies on writing the posterior in the following form:
-%
-\begin{equation}
-	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) = \pi_b(b)
-\end{equation}
-%
-Now $\pi_b(\cdot)$ is the un-normalised density we wish to sample from.
-In other words, we wish to construct a Markov chain that has $\pi_b(\cdot)$ as its invariant distribution.
-We can break $\pi_b$ down as follows:
-%
-\begin{align*}
-	\pi_b(b) &= p(b|X) \sum_{\psi} \nolimits p(A , \psi | b, X) \\
-	&= p(b|X) p(A, \psi^* | b, X) \\
-	&= p(A | b, \psi^*) \cdot p(\psi^* | b) \cdot p(b | X)
-\end{align*}
-%
-Since we are using the microcanonical SBM formulation, there is only one value of $\psi$ that is compatible with the given $(A, b)$ pair.
-We denote this value $\psi^* = \{k^*, e^*\}$.
-Specifically, $k^*_i = \sum_j A_{ij} $ and $e^*_{rs} = \sum_{i, j} A_{ij} \one \{b_i=r\} \one\{b_j=s\}$.
-Therefore, the summation over all $\psi$ reduces to just the single $\psi^*$ term.
-We also define the microcanonical entropy of the configuration as.
-%
-\begin{equation}
-	S(b) = - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big)
-\end{equation}
-%
-This entropy can be thought of as the description length of the graph because it is the sum of the information required to represent the graph given the parameters and the amount of information required to store the parameters (given the feature matrix $X$).
-The exact, from of the proposal distribution and accept-reject step is explored throughly by \citet{Peixoto-MCMC}.
-There is a widely available library, \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, implementing this algorithm.
-The only modification we make is in the block membership prior $p(b)$ which we replace with $p(b|X)=B^{-N}$ which is a uniform distribution and so cancels out in the MH accept-reject step.
-
-The end result of this is that we can generate a set of block membership samples $\{b^{(t)}\}_{t=1}^{T}$ with each $b^{(t)} \sim p(b | A, X)$.
-Each of these samples can then be used for the $\theta$-chain.
-
-\subsection{Sampling feature-to-block classifier parameters}
-
-The invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. We write this as follows:
-%
-\begin{align}
-	p(\theta | X, b) &\propto p(b | X, \theta) p(\theta) = \pi_\theta (\theta) \propto  \exp \left( - U(\theta) \right) \\
-	\therefore U(\theta) &= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}
-\end{align}
-%
-Where we have introduced $U(\theta)$ equal to the negative log posterior.
-This $U(\theta)$ term makes subsequent analysis more concise.
-Each of the constituent terms of $U(\theta)$ is easily computed (equation \ref{eqn:u-form}).
-To simplify notation, we define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
-%
-\begin{equation}
-	\log p(b | X, \theta) = \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
-	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta ||^2
-	\label{eqn:U-constituent-terms}
-\end{equation}
-%
-This means that if we ignore constant terms we can write $U(\theta)$ as in equation \ref{eqn:U-form}.
-We see that the prior effectively introduces a regularisation term.
-Note that $||\theta||^2 = \sum_{i} \theta_{i}^2 = \sum_{j=1}^{B} ||w_j||^2$ is the Euclidean norm of the vector of parameters $\theta$.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form}
-\end{equation}
-%
-$U(\theta)$ in equation \ref{eqn:U-form} appears a typical objective function to be minimised for neural network training.
-The first term is the cross-entropy between actual and predicted labels.
-The second term - introduced by the prior - brings a form of regularisation, preventing over-fitting.
-In traditional applications we only seek the value of $\theta$ that minimises the objective function $U(\theta)$, which in our case would yield the maximum a posteriori (MAP) estimate.
-This is often done through some kind of gradient descent as $\nabla U$ is easily computable (equation \ref{eqn:U-derivative}).
-
-However, our goal is not to find the MAP estimate but to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
-As discussed earlier, given the invariant $\pi_\theta(\cdot)$, it is sufficient to specify a proposal distribution and then apply a MH accept-reject step to ensure detailed balance of the Markov Chain.
-Nevertheless, we can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
-We therefore adopt the Metropolis Adjusted Langevin Algorithm (MALA) - first proposed by \citet{mala-tweedie} - which leverages just that.
-Given the current sample $\theta$, we propose a new sample $\theta'$ according to equation \ref{eqn:theta-update}
-%
-\begin{equation}
-	\theta' = \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi
-	\label{eqn:theta-update}
-\end{equation}
-%
-Where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter - which may vary with the sample index.
-Indeed without the injected noise term, this is equivalent to gradient descent.
-We require the noise term to fully explore the parameter space.
-As such the proposal distribution, is a simple multivariate Gaussian which can be easily evaluated.
-%
-\begin{equation}
-	q_\theta(\theta, \theta') = \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right)
-\end{equation}
-%
-The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}). By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as:
-%
-\begin{equation}
-	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
-	\label{eqn:U-derivative}
-\end{equation}
-%
-After a proposed move is generated, in typical Metropolis-Hastings fashion we accept the move with probability $\alpha_\theta$.
-%
-\begin{equation}
-	\alpha_\theta(\theta, \theta') = \min \left( 
-	\exp \left( U(\theta) - U(\theta')\right)
-	\frac{ 
-		q_\theta(\theta', \theta)
-	}{
-		q_\theta(\theta, \theta')
-	} 
-	, 1 \right)
-\end{equation}
-%
-This fully specifies, the sampling procedure to generate $\{\theta^{(t)}\}_{t=1}^T$. Bear in mind that each $\theta^{(t)}$ update step uses its corresponding $b^{(t)}$ block membership sample.
-
-\subsection{Sampling sequence}
-
-So far, each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample.
-This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance.
-This may lead to longer burn-in and autocorrelation times of the resulting Markov Chain.
-The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends on $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
-We would rather deal with the expectation of each $y_{ij}^{(t)}$:
-%
-\begin{equation}
-	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one(b_{i}^{(t)} = j) \right]
-	= p(b_i = j | A, X)
-\end{equation}
-%
-We obtain an unbiased estimate for this quantity as simply the empirical distribution of the block membership samples $\left\{ b^{(t)} \right\}_{t=1}^T$.
-%
-\begin{equation}
-	\hat{y}_{ij} \coloneqq \frac{1}{T} \sum_{t=1}^{T} y_{ij}^{(t)} = \frac{1}{T} \sum_{t=1}^{T} \one\{b_i^{(t)} = j\}
-\end{equation}
-%
-We therefore, choose to feed each $\theta^{(t)}$ update step the same $\hat{y}_{ij}$ for all $t$ rather than the corresponding $y^{(t)}_{ij}$.
-This means we no longer need to run the $b$ and $\theta$ Markov chains concurrently.
-Instead, we run the $b$-chain to completion and use it to generate $\hat{y}_{ij}$ for $i \in \{1 \dots N\}$ and $j \in \{1 \dots B\}$.
-This is an estimate of $p(b | A, X)$ that we use for every iteration of the $\theta$ Markov chain.
-This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$; wed refer to these as $T_b$ and $T_\theta$ henceforth.
-Furthermore, this changeover reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{Block membership sample generation} % give the algorithm a caption
-	\label{alg:b-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\FOR{$t \in \{0, 1 \dots T_b - 1\}$}
-			\STATE $b' \gets \sim q_b(b^{(t)}, b' | A)$
-			\STATE $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
-			\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-			\IF{$\log \eta < \log \alpha_b$}
-				\STATE $b^{(t+1)} \gets b'$
-			\ELSE
-				\STATE $b^{(t+1)} \gets b^{(t)}$
-			\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{LBGM parameter pseudo-marginal inference} % give the algorithm a caption
-	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\STATE $\mathcal{S} \gets \textrm{range}(\textrm{burn-in}, T_b, \textrm{thinning})$
-		\STATE $\hat{Y}_{ij} \gets \frac{1}{\mathcal{|S|}} \sum_{t \in \mathcal{S}} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
-
-		\item[]
-
-		\FOR{$t \in \{0, 1 \dots T_\theta - 1\}$}
-		\STATE $\xi \gets \sim \Gaussian(0, I)$
-		\STATE $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
-		\STATE $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
-		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-		\IF{$\log \eta < \log \alpha_\theta$}
-		\STATE $\theta^{(t+1)} \gets \theta'$
-		\ELSE
-		\STATE $\theta^{(t+1)} \gets \theta^{(t)}$
-		\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\subsection{Complexity Analysis}
-
-\section{Experiments}
-
-We apply the developed methods to a variety of datasets.
-
-\begin{itemize}
-	\item \textbf{College football teams} \cite{Evans_2010_football} - network of American college football teams and their interactions. Vertex labels are the division a team belongs to.
-	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.
-	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier} - network of the author's Facebook friends list.
-Each vertex has been manually labelled with a variety of features describing their relationship to the author.
-	\item \textbf{Twitch users} \cite{twitch} - a network of user-user friendships on the streaming service Twitch.
-Vertex labels are extracted according to video-games played, location and streaming habits.
-This dataset is also broken down into disjoint networks according to language.
-We only consider the English users with is a subnet with $N=7126$ vertices and $E=35324$ edges).
-\end{itemize}
-
-\section{Extensions}
-
-\section{Conclusion}
-
 \begin{ack}
@@
 \end{ack}
-
-%references
-\bibliography{sources.bib}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Checklist}
-
-%%% BEGIN INSTRUCTIONS %%%
-The checklist follows the references.  Please
-read the checklist guidelines carefully for information on how to answer these
-questions.  For each question, change the default \answerTODO{} to \answerYes{},
-\answerNo{}, or \answerNA{}.  You are strongly encouraged to include a {\bf
-justification to your answer}, either by referencing the appropriate section of
-your paper or providing a brief inline description.  For example:
-\begin{itemize}
-  \item Did you include the license to the code and datasets? \answerYes{See Section gen-inst.}
-  \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
-  \item Did you include the license to the code and datasets? \answerNA{}
-\end{itemize}
-Please do not modify the questions and only use the provided macros for your
-answers.  Note that the Checklist section does not count towards the page
-limit.  In your paper, please delete this instructions block and only keep the
-Checklist section heading above along with the questions/answers below.
-%%% END INSTRUCTIONS %%%
-
-\begin{enumerate}
-
-\item For all authors...
-\begin{enumerate}
-  \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
-    \answerTODO{}
-  \item Did you describe the limitations of your work?
-    \answerTODO{}
-  \item Did you discuss any potential negative societal impacts of your work?
-    \answerTODO{}
-  \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you are including theoretical results...
-\begin{enumerate}
-  \item Did you state the full set of assumptions of all theoretical results?
-    \answerTODO{}
-	\item Did you include complete proofs of all theoretical results?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you ran experiments...
-\begin{enumerate}
-  \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
-    \answerTODO{}
-  \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
-    \answerTODO{}
-	\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
-    \answerTODO{}
-	\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
-\begin{enumerate}
-  \item If your work uses existing assets, did you cite the creators?
-    \answerTODO{}
-  \item Did you mention the license of the assets?
-    \answerTODO{}
-  \item Did you include any new assets either in the supplemental material or as a URL?
-    \answerTODO{}
-  \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
-    \answerTODO{}
-  \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you used crowdsourcing or conducted research with human subjects...
-\begin{enumerate}
-  \item Did you include the full text of instructions given to participants and screenshots, if applicable?
-    \answerTODO{}
-  \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
-    \answerTODO{}
-  \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
-    \answerTODO{}
-\end{enumerate}
-
-\end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\appendix
-
-\section{Appendix}
-
-\subsection{Derivation of conditional block distribution given feature matrix}
-\label{appdx:b|x}
-
-We wish to determine the form of $p(b| X)$. This can be done by integrating over the joint probability with respect to $\theta$.
-%
-\begin{align*}
-	p(b | X) &= \int p(b , \theta| X, \theta) d\theta = \int p(b | X, \theta) p(\theta | X) d\theta \\
-	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i=1}^{N} \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
-	&= \prod_{i=1}^{N} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j=1}^{B} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k=1}^{B} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}
-\end{align*}
-%
-We note that $b_i \in {1, 2, \dots B}$ and so the integral's value is unchanged with respect to $b_i$.
-The integrand has the same form no matter which value $b_i$ takes as the prior is the same for each $w_j$.
-As such the integral can only be a function of at most $\tilde{x}_i$ and $\sigma_\theta^2$ as it is symmetric with respect to $b_i$ and all the various $w_j$ are integrated out as they are dummy variables.
-Therefore, denoting the integral by the (unknown) function $f(\tilde{x}_i, \sigma_\theta^2)$, we write $p(b| X)$ as follows:
-%
-\begin{align*}
-	p(b | X) &= \prod_{i=1}^{N} f(\tilde{x}_i, \sigma_\theta^2) = \textrm{const w.r.t } b = c
-\end{align*}
-%
-As this is a constant with respect to $b$ we conclude that $p(b | X)$ must be a uniform distribution.
-$\nicefrac{1}{c}$ is simply the size of the set of values that $b$ can take.
-We know $b_i \in \mathcal{B} = \{1, 2, \dots B\}$.
-Therefore, $b \in \mathcal{B}^N$ and $|\mathcal{B}^N| = |\mathcal{B}|^N = B^N= \nicefrac{1}{c}$.
-Putting this all together we show that:
-%
-\begin{equation}
-	p(b | X) = B^{-N}
-\end{equation}
-
-\subsection{Derivation of U gradient with respect to feature parameters}
-\label{appdx:gradu}
-The goal is to determine $\nabla U(\theta)$, the gradient of the negative log posterior with respect to the parameters. We repeat the form of $U(\theta)$ in equation \ref{eqn:U-form-appdx}.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form-appdx}
-\end{equation}
-%
-Where $y_{ij}$ is independent of $\theta$ and $a_{ij}$ is the output from the softmax layer, with form as given in equation \ref{eqn:a-ij}.
-%
-\begin{equation}
-	a_{ij} \coloneqq \phi_{j} (x_i; \theta) = \frac{\exp(w_j^T \tilde{x}_i)}{\sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i)}
-	\label{eqn:a-ij} 
-\end{equation}
-%
-We note that $\theta = \{w_k\}_{k=1}^B$, and as such we can write this in vector form $\theta = \left[w_1^T, w_2^T \dots w_B^T  \right]^T$.
-Therefore, $\nabla U(\theta) = \left[\nicefrac{\partial U}{\partial w_1}^T,\nicefrac{\partial U}{\partial w_2}^T \dots \nicefrac{\partial U}{\partial w_B}^T  \right]^T$; to compute $\nabla U(\theta)$ it suffices to find the form of $\nicefrac{\partial U}{\partial w_k}$ with respect to a general $k$.
-
-To this end, we must first find partial derivatives of $a_{ij}$ and $||\theta||$ with respect to $w_k$. Starting with $a_{ij}$:
-%
-\begin{align}
-	\frac{\partial a_{ij}}{\partial w_k} &= \frac
-	{\tilde{x}_i \exp(w_j^T \tilde{x}_i) \delta_{jk} \cdot \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) 
-		- 
-	\exp(w_j^T \tilde{x}_i) \cdot \tilde{x}_i \exp(w_k^T \tilde{x}_i)}
-	{\left( \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) \right)^2} \nonumber \\
-	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right) 
-\end{align}
-%
-Where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$. Now moving onto the derivative of $||\theta||^2$:
-%
-\begin{equation}
-	\frac{ \partial}{\partial w_k} ||\theta||^2 = \frac{\partial}{\partial w_k} \left( \sum_{b=1}^B ||w_b||^2 \right) = 2w_k
-\end{equation}
-%
-We are ready to put this all together, to find the partial derivative of $U(\theta)$ with respect to each $w_k$:
-\begin{align}
-	\frac{\partial U}{\partial w_k} &= 
-	\sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} 
-	\left( \frac{-\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
-	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
-	&=  - \left( \sum_{i=1}^{N} \tilde{x}_i \left( y_{ik} - a_{ik} \sum_{j=1}^{B} y_{ij} \right)
-	- \frac{w_k}{\sigma_\theta^2} \right) \nonumber \\
-	&= - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
-\end{align}
-%
-This is the required result.
-This form can be computed efficiently through matrix operations.
-The only property of $y_{ij}$ we have used in the derivation is the sum-to-one constraint $\sum_{j=1}^{B} y_{ij} = 1$ for all $i$.
-
-\subsection{Choosing the MALA step-size}
-\label{appdx:step-size}
-
-For sampling from the $\theta$-chain of the block membership generator parameters, we employed the Metropolis Adjusted Langevin Algorithm (MALA). At iteration $t$, the proposed sample is generated by:
-%
-\begin{equation}
-	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi
-\end{equation}
-%
-There are two competing objectives when choosing the step-size $h_t$.
-On the one hand, we want the step-size to be large so that we arrive at a high density region quickly.
-However, too large a step-size will lead to a lower acceptance ratio and thus inefficient sampling.
-A solution to this problem would be to slowly decrease the step-size with $t$ - often called simulated annealing.
-Therefore, we still have a short burn-in time but will not bounce around the mode for large $t$.
-As well as the trivial constraint for $h_t$ to be strictly positive and we introduce two further constraints as outlined by \citet{Bayesian-SGLD}:
-%
-\begin{equation}
-	\sum_{t=1}^{\infty} h_t = \infty \qquad \textrm{and} \qquad
-	\sum_{t=1}^{\infty} h_t^2 < \infty
-	\label{eqn:h-constraints}
-\end{equation}
-%
-The first constraint ensures that we have cover sufficient distance to arrive at any arbitrary point in our domain, no matter the starting point.
-The second constraint ensures that once we converge to the mode rather than simply bouncing around it.
-\citet{Bayesian-SGLD} propose the following form for a polynomially decaying step-size which we adopt:
-%
-\begin{equation}
-	h_t = a(b + t)^{-\gamma}
-\end{equation}
-%
-Where $a, b, \gamma$ are hyper-parameters to be chosen.
-We require $a,b > 0$ and $\gamma \in (0.5, 1]$ to satisfy equation \ref{eqn:h-constraints}.
-We find empirically that $a=250, b=1000, \gamma=0.8$ yield good burn-in times for our MALA implementations.
-
-\end{document}

### 2021-05-16 nips/appendices.tex
Split into multiple tex files

--- 2021-05-14 nips/submission.tex
@@
-\documentclass{article}
-
-% if you need to pass options to natbib, use, e.g.:
-%     \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading neurips_2021
-
-% ready for submission
-\PassOptionsToPackage{numbers, compress}{natbib}
-\usepackage[final]{neurips_2021}
-
-% to compile a preprint version, e.g., for submission to arXiv, add add the
-% [preprint] option:
-%     \usepackage[preprint]{neurips_2021}
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-%\usepackage[final]{neurips_2021}
-
-% to avoid loading the natbib package, add option nonatbib:
-%    \usepackage[nonatbib]{neurips_2021}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{xcolor}         % colors
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-\usepackage{amsmath}
-\usepackage{mathtools}
-\usepackage{bbm}
-\usepackage{amssymb}
-\usepackage{algorithmic}
-\usepackage{algorithm}
-
-%bibliography
-\bibliographystyle{plainnat}
-
-\title{Inferring community characteristics in labelled networks}
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to break the
-% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
-% authors names on the first line, and the last on the second line, try using
-% \AND instead of \And before the third author name.
-
-\author{
-  Lawrence Tray \\
-  Department of Engineering \\
-  University of Cambridge \\
-  \texttt{lpt30@cam.ac.uk} \\
-  % examples of more authors
-  \And
-  Ioannis Kontoyiannis \\
-  Department of Mathematics \\
-  University of Cambridge \\
-  \texttt{ik355@cam.ac.uk}
-}
-
-%custom commands
-\newcommand{\Xcal}{\mathcal{X}}
-\newcommand{\Bcal}{\mathcal{B}}
-\newcommand{\Integers}{\mathbb{Z}}
-\newcommand{\one}{\mathbbm{1}}
-\newcommand{\Gaussian}{\mathcal{N}}
-\newcommand{\indep}{\perp \!\!\! \perp}
-\newcommand{\specialchoose}{\genfrac{\{}{\}}{0pt}{}}
-\newcommand{\Expect}{\mathbb{E}}
-
-%envs
-\newtheorem{definition}{Definition}[section]
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[theorem]
-\newtheorem{lemma}[theorem]{Lemma}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-We present a novel generative model for describing vertex-labelled undirected graphs.
-This is a two-layered model.
-First a probabilistic mapping from vertex labels to latent block memberships and from these block memberships we generate a graph according to the micro-canonical stochastic block model.
-
-With this framework we are able to efficiently sample the parameters of the vertex label to block membership mapping.
-This allows us to automatically determine which features have the largest impact on graphical structure.
-\end{abstract}
-
-\section{Introduction}
-
-There is a wealth of networks in the world.
-Some examples of this graphical data are social networks, website hyperlinks and academic collaboration with more being produced each second.
-It is clear we need tools to analyse this increasingly omni-present form of data.
-
-A somewhat surprising property of real-world graphs is that they exhibit strong community structure.
-In other words, each node will often belong to a cluster of densely connected nodes.
-This property is often exploited by graph compression algorithms and there is high interest in recovering the communities from the observed graph.
-
-A common subset of graphical data is the labelled network.
-This is a graph where we have information about the properties of each node.
-We shall refer to these node properties as features.
-One of the most common questions we can ask of this dataset is what features have the largest impact on the structure of the graph.
-For example, when analysing an academic collaboration graph, one may wish to ask what impact gender has on the structure of the network.
-Nevertheless, this analysis is often vulnerable to confounding variables.
-While gender very well may impact the structure of the graph, there is often a better explanatory variable for the structure.
-
-There is space to bridge the gap between these two approaches.
-Rather than determining whether a feature impacts the graphical structure directly, we can use the community concept as a stepping stone in our analysis.
-Therefore, we extract which features have the largest impact on overall graphical structure.
-This can be thought of as a form of dimensionality-reduction where we only keep the node features that are useful in predicting community memberships.
-
-\section{Preliminaries}
-
-This section defines some preliminary concepts required for the subsequent analysis.
-We first need a model for community-like structure in a graphical network.
-For this we adopt the stochastic block model (SBM) - widely used across academia.
-The premise is that each node in the graph belongs to a unique community called a block.
-The probability that two nodes are connected depends only on the block memberships of each.
-Graphs drawn from this ensemble exhibit community structure.
-Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}.
-A paraphrased definition is given below.
-
-\begin{definition}[Microcanonical NDC-SBM]
-	\label{defn:microcan-ndc-sbm}
-	Let $N \in \Integers^{+}$ denote the number of vertices in an undirected graph.
-The block memberships are encoded by a vector $b$ of length $N$ where each entry $b_i \in \{1, 2 \dots B\}$.
-$B \in \Integers^{+}$ is the number of non-empty blocks.
-Let $e$ be a $B \times B$ matrix of edge counts between blocks ($e_{rs}$ is number of edges from block $r$ onto block $s$ - or twice that number if $r=s$).
-For undirected graphs $e$ is symmetric.
-For a non-degree-corrected stochastic block model (NDC-SBM), we say that the graph $A$ is generated as follows:
-	%
-	\begin{equation}
-		A \sim \textrm{NDC-SBM}_{\textrm{MC}} (b, e)
-	\end{equation}
-	%
-	Where edges are placed uniformly at random but respecting the constraint imposed by $e$ and $b$.
-The additional parameters $N$ and $B$ are omitted as they are inferred from the shapes of $b$ and $e$.
-If we interpret $A$ as an adjacency matrix, then this constraint can be written formally as: $e_{rs} = \sum_{i,j} A_{ij} \one \{b_i = r\} \one \{b_j = s\}$.
-\end{definition}
-
-Nevertheless, this formulation does not accept high degree variability within blocks as is typical of real-world data.
-Indeed, the NDC-SBM favours a partition into high-degree and low-degree nodes rather than clusters of inter-connected nodes.
-We therefore introduce the degree-corrected SBM (DC-SBM) to circumvent these issues.
-
-\begin{definition}[Microcanonical DC-SBM]
-	\label{defn:microcan-dc-sbm}
-	This is much like the NDC-SBM but has an additional parameter $k$ which is an $N$-length vector encoding the degree sequence ($k_i$ is the degree of vertex $i$). Therefore, we write:
-	%
-	\begin{equation}
-		A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, e, k)
-	\end{equation}
-	%
-	Once again, edges are placed uniformly at random but respecting the constraints imposed by the parameters.
-The DC-SBM has the additional constraint that $k_i = \sum_{j} A_{ij}$.
-In what follows, we will always assume the degree-corrected model unless otherwise specified.
-
-\end{definition}
-
-\section{Latent block generative model}
-
-In this section we propose a novel generative model for modelling labelled networks.
-We restrict our analysis to labelled, undirected graphs with $N$ nodes.
-We define the vector $x_i \in \Xcal^D$ as the feature vector for the $i$'th vertex.
-Each vertex has $D$ total features and we assume all entries take values from the same set $\Xcal$.
-For the majority of datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$.
-The feature vectors $\{x_i\}_{i=1}^{N}$ are subsumed into the $N \times D$ matrix $X$.
-
-The proposed generative model (which we call the latent block generative model - LBGM) is given in figure \ref{fig:lbgm}.
-We start, with the feature matrix X and generate a vector of block memberships $b$.
-The parameters of this step are encapsulated by $\theta$.
-Each feature vector $x_i$ is treated independently and used to generate the block membership $b_i$.
-Each entry $b_i \in \{1, 2 \dots B\}$ where $B$ is the number of blocks and a constant in our model.
-We choose a single softmax layer to model $p(b_i | x_i, \theta)$.
-More complex models are possible but then deriving meaning from inferred parameter distributions is more difficult.
-Summarising, we write $p(b | X, \theta)$ as follows:
-%
-\begin{equation}
-	p(b| X, \theta) = \prod_{i=1}^{N} p(b_i | x_i, \theta) = \prod_{i=1}^{N} \phi_{b_i} (x_i; \theta)
-	= \prod_{i=1}^{N} \frac{\exp\left(w_{b_i}^T \tilde{x}_i\right)}{\sum_{k=1}^{B} \exp \left( w_k^T \tilde{x}_i\right)}
-\end{equation}
-%
-Where $\tilde{x} \coloneqq \left[x_1, x_2, \dots x_D, 1\right]^T$ is an augmented version of $x$ that allows for a bias term.
-The parameters for this stage are denoted $\theta$ and consist of the all $B$ weight vectors $\theta = \{w_k\}_{k=1}^{B}$.
-Each $w_k$ has dimension $D+1$.
-We could instead write the parameters $\theta$ as a $B \times (D+1)$ matrix of weights $W$; this form has use computationally as then $Z_i = W \tilde{x}_i$ is the input to the softmax activation function.
-
-\begin{figure}[!h]
-	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=7mm},
-		squarednode/.style={rectangle, draw=black, minimum size=7mm}
-		]
-		% nodes
-		\node[roundnode] (X) at (0, 0) {$X$};
-		\node[squarednode] (b) at (3, 0) {$b$};
-		\node[roundnode] (A) at (6, 0) {$A$};
-
-		% arrows
-		\draw[->] (X.east) -- node[above] {$\theta$} (b.west);
-		\draw[->] (b.east) -- node[above] {$\psi$}(A.west);
-	\end{tikzpicture}
-	\caption{Latent block generative model}
-	\label{fig:lbgm}
-\end{figure}
-
-With the block memberships $b$ generated, we then draw the graph $A$ from the microcanical DC-SBM (\ref{defn:microcan-dc-sbm}) with additional parameters $\psi = \{e, k\}$.
-In a slight abuse of notation we denote the inter-block edge count matrix with $e = \psi_e$ and the degree sequence $k = \psi_k$ to make explicit that these parameters are contained in $\psi$.
-%
-\begin{equation}
-	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k)
-\end{equation}
-
-\subsection{Prior selection}
-
-Before performing any inference, we must specify priors on $\theta$ and $\psi$.
-For $\theta$ it seems sensible to choose a Gaussian prior, with zero mean and variance matrix $\sigma^2_\theta I$ such that each element of $\theta$ is independent and distributed like $\sim \Gaussian(0, \sigma_\theta^2)$.
-In vector form, the prior for $\theta$ is therefore:
-%
-\begin{equation}
-	p(\theta) = \Gaussian \left( \theta ; 0, \sigma_\theta^2 I \right)
-	\label{eqn:theta-prior}
-\end{equation}
-%
-We will see that this form of prior is equivalent to a regularisation term in neural network training that penalises extreme weight magnitudes.
-For $\sigma_\theta^2 \rightarrow \infty$ this becomes an uninformative uniform prior.
-
-In our model, the block memberships vector $b$ is an intermediate latent variable and so we are not free to choose a prior for it.
-Nevertheless, as far as inference on the right-hand-side of figure \ref{fig:lbgm}, we regard $p(b | X)$ as a pseudo-prior on $b$.
-We can show (appendix \ref{appdx:b|x}) that our choice of prior for $p(\theta)$ in equation \ref{eqn:theta-prior} leads to a uniform $p(b | X)$ in equation \ref{eqn:b-pseudo-prior}.
-%
-\begin{equation}
-	p(b | X) = \int p(b | X, \theta) p(\theta) d\theta = B^{-N}
-	\label{eqn:b-pseudo-prior}
-\end{equation}
-%
-This is a uniform distribution that only depends on the number of blocks $B$.
-This simplifies the problem nicely as evaluating $p(b | X)$ does not require knowing the exact value $X$ takes.
-\citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for the additional microcanonical SBM parameters $\psi$ which we adopt.
-Peixoto's idea is to write the joint prior on $(b, e, k)$ as a product of conditionals $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$.
-For our purposes we must insert a conditioning on $X$, to form our pseudo-prior for $b$ and $\psi$, to give equation \ref{eqn:joint-pseudo-prior}.
-%
-\begin{equation}
-	p(b, \psi | X) = p(b | X) p(\psi | b, X) = p(b | X) p(\psi | b)
-	\label{eqn:joint-pseudo-prior}
-\end{equation}
-%
-Where the simplification is made apparent by noting $(\psi \indep X) | b$. We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference.
-%
-\begin{equation}
-	p(\psi | b) = p(e | b) p(k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
-	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right]
-\end{equation}
-%
-Where $\specialchoose{n}{m}$ is shorthand for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of distinct histograms with $n$ bins under the constraint they sum to $m$.
-$E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges in the graph.
-Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform with respect to $e$.
-The variable $\eta_j^r$ is introduced to denote the number of vertices in block $r$ that have degree $j$.
-Formally, $\eta_j^r \coloneqq \sum_{i} \one\left\{b_i = r \right\} \one \left\{k_i = j \right\}$.
-Furthermore, $q(m, n)$ is the number of different histograms with at most $n$ non-zero bins that sum to $m$.
-$q(m, n)$ is related to but different from $\specialchoose{n}{m}$.
-Lastly, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ is the number of vertices assigned to block $r$.
-Importantly, we have computable forms for $p(\theta)$ and $p(b, \psi | X)$ which will be useful for performing inference.
-
-\section{Inference}
-
-Now that we have defined the LBGM, we wish to leverage it to perform inference.
-Suppose we are presented with a vertex-labelled graph $(A, X)$; the goal is to draw samples for $\theta$ according to the posterior given the observed graph (equation \ref{eqn:theta-target}).
-%
-\begin{equation}
-	\label{eqn:theta-target}
-	\theta^{(t)} \sim p(\theta | A, X)
-\end{equation}
-%
-These samples allow us to approximate the posterior distribution for $\theta$ as well as compute a predictive distribution $p(b^* | x^*, A, X) = \int p(b^* | x^*, \theta) p(\theta | X, A) d\theta \approx \frac{1}{T} \sum_{t=1}^{T} p(b^* | x^*, \theta^{(t)})$.
-However, generating these samples is not easily done in practice.
-
-We instead propose an iterative approach.
-First drawing samples $b^{(t)}$ from the block membership posterior (equation \ref{eqn:b-samples}).
-We then use each $b^{(i)}$ to draw samples for $\theta$ as in equation \ref{eqn:theta-samples}.
-%
-\begin{align}
-	b^{(i)} &\sim p \Big( b | A, X \Big)  \label{eqn:b-samples}\\
-	\theta^{(i)} &\sim p\Big(\theta | X, b^{(i)} \Big) \label{eqn:theta-samples}
-\end{align}
-%
-Both of these sampling steps implemented with a Markov Chain through the Metropolis-Hastings algorithm \cite{hastings-alg}.
-We just need to define a proposal distribution $q(x, x')$ for proposing a move $x \rightarrow x'$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(\cdot)$, point-wise.
-The proposed move is then accepted with probability $\alpha$ (equation \ref{eqn:mh-accept}) else it is rejected and we stay at $x$.
-%
-\begin{equation}
-	\alpha = \min \left( \frac{\pi(x') q(x', x)}{\pi(x) q(x, x')} , 1 \right)
-	\label{eqn:mh-accept}
-\end{equation}
-%
-This accept-reject step ensures the resulting Markov Chain is in detailed balance with the target distribution $\pi(\cdot)$.
-What we propose in equations \ref{eqn:b-samples} and \ref{eqn:theta-samples} is therefore implemented through a 2-level Markov chain.
-The resulting samples for $\theta^{(i)}$ are unbiased in the sense that the expectation of their distribution is the posterior we are targeting in equation \ref{eqn:theta-target}.
-%
-\begin{align*}
-	\Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] &= \sum_{b \in \mathcal{B}^N} p(\theta | X, b) p(b | A, X) \\
-	&= \sum_{b \in \mathcal{B}^N} p(\theta, b | A, X) \\
-	&= p(\theta | A, X) \\
-\end{align*}
-%
-This is an example of a pseudo-marginal approach \cite{pseudo-marginal}.
-Indeed, the unbiased result is sufficient to prove that for sufficient samples, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{t})\right] = p(\theta
-\ A, X)$ which is exactly the distribution we were targetting (equation \ref{eqn:theta-target}).
-
-The reason we split the Markov chain into two stages is because the summation over all latent states $b \in \mathcal{B}^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in \Bcal^N} p(A | b) P(b | X, \theta)$ is intractable $O(B^N)$.
-%
-\begin{figure}[!h]
-	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=12mm},
-		squarednode/.style={rectangle, draw=black, minimum size=12mm}
-		]
-		% nodes
-		\node[roundnode] (b0) at (0, 2) {$b^{(t)}$};
-		\node[roundnode] (b1) at (4, 2) {$b'$};
-		\node[roundnode] (b2) at (8, 2) {$b^{(t+1)}$};
-		\node[roundnode] (t0) at (0, 0) {$\theta^{(t)}$};
-		\node[roundnode] (t1) at (4, 0) {$\theta'$};
-		\node[roundnode] (t2) at (8, 0) {$\theta^{(t+1)}$};
-
-		% arrows
-		\draw[->] (b0) to node[above] {$\sim q_b(b^{(t)}, b' | A)$} (b1);
-		\draw[->] (b1) to node[above] {$\alpha_b (b, b' | A)$} (b2);
-		\draw[->] (b0) [out=30, in=150] to node[above] {$1-\alpha_b$} (b2);
-
-		\draw[->] (t0) to node[above] {$\sim q_\theta(\theta, \theta' | X, b^{(t)})$} (t1);
-		\draw[->] (t1) to node[above] {$\alpha_\theta (\theta, \theta' | X, b^{(t)})$} (t2);
-		\draw[->] (t0) [out=-30, in=-150] to node[below] {$1-\alpha_\theta$} (t2);
-
-		\draw[dashed, ->] (b0) to (2, 0.8);
-		\draw[dashed, ->] (b0) to (5.5, 0.8);
-
-	\end{tikzpicture}
-	\caption{Sampling sequence}
-	\label{fig:samp-sequence}
-\end{figure}
-%
-Figure \ref{fig:samp-sequence} shows an overview of the proposed method.
-We have introduced subscripts and conditionings to make explicit what parameters each step utilises.
-In an important simplification, we note that $p(b| X) = B^{-N}$ which does not depend on the exact value of X.
-Therefore, we do not need to know the value of $X$ to perform the sampling on $b$.
-Conversely, for the $\theta^{(t)}$ samples, we use $b^{(t)}$ but not $A$ as $(\theta \indep A) | b$.
-
-\subsection{Sampling block memberships}
-
-\citet{Peixoto-MCMC} proposes a Monte Carlo method which we will base our approach on. It relies on writing the posterior in the following form:
-%
-\begin{equation}
-	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) = \pi_b(b)
-\end{equation}
-%
-Now $\pi_b(\cdot)$ is the un-normalised density we wish to sample from.
-In other words, we wish to construct a Markov chain that has $\pi_b(\cdot)$ as its invariant distribution.
-We can break $\pi_b$ down as follows:
-%
-\begin{align*}
-	\pi_b(b) &= p(b|X) \sum_{\psi} \nolimits p(A , \psi | b, X) \\
-	&= p(b|X) p(A, \psi^* | b, X) \\
-	&= p(A | b, \psi^*) \cdot p(\psi^* | b) \cdot p(b | X)
-\end{align*}
-%
-Since we are using the microcanonical SBM formulation, there is only one value of $\psi$ that is compatible with the given $(A, b)$ pair.
-We denote this value $\psi^* = \{k^*, e^*\}$.
-Specifically, $k^*_i = \sum_j A_{ij} $ and $e^*_{rs} = \sum_{i, j} A_{ij} \one \{b_i=r\} \one\{b_j=s\}$.
-Therefore, the summation over all $\psi$ reduces to just the single $\psi^*$ term.
-We also define the microcanonical entropy of the configuration as.
-%
-\begin{equation}
-	S(b) = - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big)
-\end{equation}
-%
-This entropy can be thought of as the description length of the graph because it is the sum of the information required to represent the graph given the parameters and the amount of information required to store the parameters (given the feature matrix $X$).
-The exact, from of the proposal distribution and accept-reject step is explored throughly by \citet{Peixoto-MCMC}.
-There is a widely available library, \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, implementing this algorithm.
-The only modification we make is in the block membership prior $p(b)$ which we replace with $p(b|X)=B^{-N}$ which is a uniform distribution and so cancels out in the MH accept-reject step.
-
-The end result of this is that we can generate a set of block membership samples $\{b^{(t)}\}_{t=1}^{T}$ with each $b^{(t)} \sim p(b | A, X)$.
-Each of these samples can then be used for the $\theta$-chain.
-
-\subsection{Sampling feature-to-block classifier parameters}
-
-The invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. We write this as follows:
-%
-\begin{align}
-	p(\theta | X, b) &\propto p(b | X, \theta) p(\theta) = \pi_\theta (\theta) \propto  \exp \left( - U(\theta) \right) \\
-	\therefore U(\theta) &= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}
-\end{align}
-%
-Where we have introduced $U(\theta)$ equal to the negative log posterior.
-This $U(\theta)$ term makes subsequent analysis more concise.
-Each of the constituent terms of $U(\theta)$ is easily computed (equation \ref{eqn:u-form}).
-To simplify notation, we define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
-%
-\begin{equation}
-	\log p(b | X, \theta) = \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
-	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta ||^2
-	\label{eqn:U-constituent-terms}
-\end{equation}
-%
-This means that if we ignore constant terms we can write $U(\theta)$ as in equation \ref{eqn:U-form}.
-We see that the prior effectively introduces a regularisation term.
-Note that $||\theta||^2 = \sum_{i} \theta_{i}^2 = \sum_{j=1}^{B} ||w_j||^2$ is the Euclidean norm of the vector of parameters $\theta$.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form}
-\end{equation}
-%
-$U(\theta)$ in equation \ref{eqn:U-form} appears a typical objective function to be minimised for neural network training.
-The first term is the cross-entropy between actual and predicted labels.
-The second term - introduced by the prior - brings a form of regularisation, preventing over-fitting.
-In traditional applications we only seek the value of $\theta$ that minimises the objective function $U(\theta)$, which in our case would yield the maximum a posteriori (MAP) estimate.
-This is often done through some kind of gradient descent as $\nabla U$ is easily computable (equation \ref{eqn:U-derivative}).
-
-However, our goal is not to find the MAP estimate but to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
-As discussed earlier, given the invariant $\pi_\theta(\cdot)$, it is sufficient to specify a proposal distribution and then apply a MH accept-reject step to ensure detailed balance of the Markov Chain.
-Nevertheless, we can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
-We therefore adopt the Metropolis Adjusted Langevin Algorithm (MALA) - first proposed by \citet{mala-tweedie} - which leverages just that.
-Given the current sample $\theta$, we propose a new sample $\theta'$ according to equation \ref{eqn:theta-update}
-%
-\begin{equation}
-	\theta' = \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi
-	\label{eqn:theta-update}
-\end{equation}
-%
-Where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter - which may vary with the sample index.
-Indeed without the injected noise term, this is equivalent to gradient descent.
-We require the noise term to fully explore the parameter space.
-As such the proposal distribution, is a simple multivariate Gaussian which can be easily evaluated.
-%
-\begin{equation}
-	q_\theta(\theta, \theta') = \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right)
-\end{equation}
-%
-The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}). By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as:
-%
-\begin{equation}
-	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
-	\label{eqn:U-derivative}
-\end{equation}
-%
-After a proposed move is generated, in typical Metropolis-Hastings fashion we accept the move with probability $\alpha_\theta$.
-%
-\begin{equation}
-	\alpha_\theta(\theta, \theta') = \min \left( 
-	\exp \left( U(\theta) - U(\theta')\right)
-	\frac{ 
-		q_\theta(\theta', \theta)
-	}{
-		q_\theta(\theta, \theta')
-	} 
-	, 1 \right)
-\end{equation}
-%
-This fully specifies, the sampling procedure to generate $\{\theta^{(t)}\}_{t=1}^T$. Bear in mind that each $\theta^{(t)}$ update step uses its corresponding $b^{(t)}$ block membership sample.
-
-\subsection{Sampling sequence}
-
-So far, each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample.
-This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance.
-This may lead to longer burn-in and autocorrelation times of the resulting Markov Chain.
-The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends on $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
-We would rather deal with the expectation of each $y_{ij}^{(t)}$:
-%
-\begin{equation}
-	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one(b_{i}^{(t)} = j) \right]
-	= p(b_i = j | A, X)
-\end{equation}
-%
-We obtain an unbiased estimate for this quantity as simply the empirical distribution of the block membership samples $\left\{ b^{(t)} \right\}_{t=1}^T$.
-%
-\begin{equation}
-	\hat{y}_{ij} \coloneqq \frac{1}{T} \sum_{t=1}^{T} y_{ij}^{(t)} = \frac{1}{T} \sum_{t=1}^{T} \one\{b_i^{(t)} = j\}
-\end{equation}
-%
-We therefore, choose to feed each $\theta^{(t)}$ update step the same $\hat{y}_{ij}$ for all $t$ rather than the corresponding $y^{(t)}_{ij}$.
-This means we no longer need to run the $b$ and $\theta$ Markov chains concurrently.
-Instead, we run the $b$-chain to completion and use it to generate $\hat{y}_{ij}$ for $i \in \{1 \dots N\}$ and $j \in \{1 \dots B\}$.
-This is an estimate of $p(b | A, X)$ that we use for every iteration of the $\theta$ Markov chain.
-This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$; wed refer to these as $T_b$ and $T_\theta$ henceforth.
-Furthermore, this changeover reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{Block membership sample generation} % give the algorithm a caption
-	\label{alg:b-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\FOR{$t \in \{0, 1 \dots T_b - 1\}$}
-			\STATE $b' \gets \sim q_b(b^{(t)}, b' | A)$
-			\STATE $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
-			\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-			\IF{$\log \eta < \log \alpha_b$}
-				\STATE $b^{(t+1)} \gets b'$
-			\ELSE
-				\STATE $b^{(t+1)} \gets b^{(t)}$
-			\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{LBGM parameter pseudo-marginal inference} % give the algorithm a caption
-	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\STATE $\mathcal{S} \gets \textrm{range}(\textrm{burn-in}, T_b, \textrm{thinning})$
-		\STATE $\hat{Y}_{ij} \gets \frac{1}{\mathcal{|S|}} \sum_{t \in \mathcal{S}} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
-
-		\item[]
-
-		\FOR{$t \in \{0, 1 \dots T_\theta - 1\}$}
-		\STATE $\xi \gets \sim \Gaussian(0, I)$
-		\STATE $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
-		\STATE $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
-		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-		\IF{$\log \eta < \log \alpha_\theta$}
-		\STATE $\theta^{(t+1)} \gets \theta'$
-		\ELSE
-		\STATE $\theta^{(t+1)} \gets \theta^{(t)}$
-		\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\subsection{Complexity Analysis}
-
-\section{Experiments}
-
-We apply the developed methods to a variety of datasets.
-
-\begin{itemize}
-	\item \textbf{College football teams} \cite{Evans_2010_football} - network of American college football teams and their interactions. Vertex labels are the division a team belongs to.
-	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.
-	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier} - network of the author's Facebook friends list.
-Each vertex has been manually labelled with a variety of features describing their relationship to the author.
-	\item \textbf{Twitch users} \cite{twitch} - a network of user-user friendships on the streaming service Twitch.
-Vertex labels are extracted according to video-games played, location and streaming habits.
-This dataset is also broken down into disjoint networks according to language.
-We only consider the English users with is a subnet with $N=7126$ vertices and $E=35324$ edges).
-\end{itemize}
-
-\section{Extensions}
-
-\section{Conclusion}
-
-\begin{ack}
-Use unnumbered first level headings for the acknowledgments. All acknowledgments
-go at the end of the paper before the list of references. Moreover, you are required to declare
-funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
-More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2021/PaperInformation/FundingDisclosure}.
-
-Do {\bf not} include this section in the anonymized submission, only in the final paper.
-You can use the \texttt{ack} environment provided in the style file to automatically hide this section in the anonymized submission.
-\end{ack}
-
-%references
-\bibliography{sources.bib}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Checklist}
-
-%%% BEGIN INSTRUCTIONS %%%
-The checklist follows the references.  Please
-read the checklist guidelines carefully for information on how to answer these
-questions.  For each question, change the default \answerTODO{} to \answerYes{},
-\answerNo{}, or \answerNA{}.  You are strongly encouraged to include a {\bf
-justification to your answer}, either by referencing the appropriate section of
-your paper or providing a brief inline description.  For example:
-\begin{itemize}
-  \item Did you include the license to the code and datasets? \answerYes{See Section gen-inst.}
-  \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
-  \item Did you include the license to the code and datasets? \answerNA{}
-\end{itemize}
-Please do not modify the questions and only use the provided macros for your
-answers.  Note that the Checklist section does not count towards the page
-limit.  In your paper, please delete this instructions block and only keep the
-Checklist section heading above along with the questions/answers below.
-%%% END INSTRUCTIONS %%%
-
-\begin{enumerate}
-
-\item For all authors...
-\begin{enumerate}
-  \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
-    \answerTODO{}
-  \item Did you describe the limitations of your work?
-    \answerTODO{}
-  \item Did you discuss any potential negative societal impacts of your work?
-    \answerTODO{}
-  \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you are including theoretical results...
-\begin{enumerate}
-  \item Did you state the full set of assumptions of all theoretical results?
-    \answerTODO{}
-	\item Did you include complete proofs of all theoretical results?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you ran experiments...
-\begin{enumerate}
-  \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
-    \answerTODO{}
-  \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
-    \answerTODO{}
-	\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
-    \answerTODO{}
-	\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
-\begin{enumerate}
-  \item If your work uses existing assets, did you cite the creators?
-    \answerTODO{}
-  \item Did you mention the license of the assets?
-    \answerTODO{}
-  \item Did you include any new assets either in the supplemental material or as a URL?
-    \answerTODO{}
-  \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
-    \answerTODO{}
-  \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you used crowdsourcing or conducted research with human subjects...
-\begin{enumerate}
-  \item Did you include the full text of instructions given to participants and screenshots, if applicable?
-    \answerTODO{}
-  \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
-    \answerTODO{}
-  \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
-    \answerTODO{}
-\end{enumerate}
-
-\end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\appendix
-
 \section{Appendix}
@@
 We find empirically that $a=250, b=1000, \gamma=0.8$ yield good burn-in times for our MALA implementations.
-
-\end{document}

### 2021-05-16 nips/checklist.tex
Split into multiple tex files

--- 2021-05-14 nips/submission.tex
@@
-\documentclass{article}
-
-% if you need to pass options to natbib, use, e.g.:
-%     \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading neurips_2021
-
-% ready for submission
-\PassOptionsToPackage{numbers, compress}{natbib}
-\usepackage[final]{neurips_2021}
-
-% to compile a preprint version, e.g., for submission to arXiv, add add the
-% [preprint] option:
-%     \usepackage[preprint]{neurips_2021}
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-%\usepackage[final]{neurips_2021}
-
-% to avoid loading the natbib package, add option nonatbib:
-%    \usepackage[nonatbib]{neurips_2021}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{xcolor}         % colors
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-\usepackage{amsmath}
-\usepackage{mathtools}
-\usepackage{bbm}
-\usepackage{amssymb}
-\usepackage{algorithmic}
-\usepackage{algorithm}
-
-%bibliography
-\bibliographystyle{plainnat}
-
-\title{Inferring community characteristics in labelled networks}
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to break the
-% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
-% authors names on the first line, and the last on the second line, try using
-% \AND instead of \And before the third author name.
-
-\author{
-  Lawrence Tray \\
-  Department of Engineering \\
-  University of Cambridge \\
-  \texttt{lpt30@cam.ac.uk} \\
-  % examples of more authors
-  \And
-  Ioannis Kontoyiannis \\
-  Department of Mathematics \\
-  University of Cambridge \\
-  \texttt{ik355@cam.ac.uk}
-}
-
-%custom commands
-\newcommand{\Xcal}{\mathcal{X}}
-\newcommand{\Bcal}{\mathcal{B}}
-\newcommand{\Integers}{\mathbb{Z}}
-\newcommand{\one}{\mathbbm{1}}
-\newcommand{\Gaussian}{\mathcal{N}}
-\newcommand{\indep}{\perp \!\!\! \perp}
-\newcommand{\specialchoose}{\genfrac{\{}{\}}{0pt}{}}
-\newcommand{\Expect}{\mathbb{E}}
-
-%envs
-\newtheorem{definition}{Definition}[section]
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[theorem]
-\newtheorem{lemma}[theorem]{Lemma}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-We present a novel generative model for describing vertex-labelled undirected graphs.
-This is a two-layered model.
-First a probabilistic mapping from vertex labels to latent block memberships and from these block memberships we generate a graph according to the micro-canonical stochastic block model.
-
-With this framework we are able to efficiently sample the parameters of the vertex label to block membership mapping.
-This allows us to automatically determine which features have the largest impact on graphical structure.
-\end{abstract}
-
-\section{Introduction}
-
-There is a wealth of networks in the world.
-Some examples of this graphical data are social networks, website hyperlinks and academic collaboration with more being produced each second.
-It is clear we need tools to analyse this increasingly omni-present form of data.
-
-A somewhat surprising property of real-world graphs is that they exhibit strong community structure.
-In other words, each node will often belong to a cluster of densely connected nodes.
-This property is often exploited by graph compression algorithms and there is high interest in recovering the communities from the observed graph.
-
-A common subset of graphical data is the labelled network.
-This is a graph where we have information about the properties of each node.
-We shall refer to these node properties as features.
-One of the most common questions we can ask of this dataset is what features have the largest impact on the structure of the graph.
-For example, when analysing an academic collaboration graph, one may wish to ask what impact gender has on the structure of the network.
-Nevertheless, this analysis is often vulnerable to confounding variables.
-While gender very well may impact the structure of the graph, there is often a better explanatory variable for the structure.
-
-There is space to bridge the gap between these two approaches.
-Rather than determining whether a feature impacts the graphical structure directly, we can use the community concept as a stepping stone in our analysis.
-Therefore, we extract which features have the largest impact on overall graphical structure.
-This can be thought of as a form of dimensionality-reduction where we only keep the node features that are useful in predicting community memberships.
-
-\section{Preliminaries}
-
-This section defines some preliminary concepts required for the subsequent analysis.
-We first need a model for community-like structure in a graphical network.
-For this we adopt the stochastic block model (SBM) - widely used across academia.
-The premise is that each node in the graph belongs to a unique community called a block.
-The probability that two nodes are connected depends only on the block memberships of each.
-Graphs drawn from this ensemble exhibit community structure.
-Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}.
-A paraphrased definition is given below.
-
-\begin{definition}[Microcanonical NDC-SBM]
-	\label{defn:microcan-ndc-sbm}
-	Let $N \in \Integers^{+}$ denote the number of vertices in an undirected graph.
-The block memberships are encoded by a vector $b$ of length $N$ where each entry $b_i \in \{1, 2 \dots B\}$.
-$B \in \Integers^{+}$ is the number of non-empty blocks.
-Let $e$ be a $B \times B$ matrix of edge counts between blocks ($e_{rs}$ is number of edges from block $r$ onto block $s$ - or twice that number if $r=s$).
-For undirected graphs $e$ is symmetric.
-For a non-degree-corrected stochastic block model (NDC-SBM), we say that the graph $A$ is generated as follows:
-	%
-	\begin{equation}
-		A \sim \textrm{NDC-SBM}_{\textrm{MC}} (b, e)
-	\end{equation}
-	%
-	Where edges are placed uniformly at random but respecting the constraint imposed by $e$ and $b$.
-The additional parameters $N$ and $B$ are omitted as they are inferred from the shapes of $b$ and $e$.
-If we interpret $A$ as an adjacency matrix, then this constraint can be written formally as: $e_{rs} = \sum_{i,j} A_{ij} \one \{b_i = r\} \one \{b_j = s\}$.
-\end{definition}
-
-Nevertheless, this formulation does not accept high degree variability within blocks as is typical of real-world data.
-Indeed, the NDC-SBM favours a partition into high-degree and low-degree nodes rather than clusters of inter-connected nodes.
-We therefore introduce the degree-corrected SBM (DC-SBM) to circumvent these issues.
-
-\begin{definition}[Microcanonical DC-SBM]
-	\label{defn:microcan-dc-sbm}
-	This is much like the NDC-SBM but has an additional parameter $k$ which is an $N$-length vector encoding the degree sequence ($k_i$ is the degree of vertex $i$). Therefore, we write:
-	%
-	\begin{equation}
-		A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, e, k)
-	\end{equation}
-	%
-	Once again, edges are placed uniformly at random but respecting the constraints imposed by the parameters.
-The DC-SBM has the additional constraint that $k_i = \sum_{j} A_{ij}$.
-In what follows, we will always assume the degree-corrected model unless otherwise specified.
-
-\end{definition}
-
-\section{Latent block generative model}
-
-In this section we propose a novel generative model for modelling labelled networks.
-We restrict our analysis to labelled, undirected graphs with $N$ nodes.
-We define the vector $x_i \in \Xcal^D$ as the feature vector for the $i$'th vertex.
-Each vertex has $D$ total features and we assume all entries take values from the same set $\Xcal$.
-For the majority of datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$.
-The feature vectors $\{x_i\}_{i=1}^{N}$ are subsumed into the $N \times D$ matrix $X$.
-
-The proposed generative model (which we call the latent block generative model - LBGM) is given in figure \ref{fig:lbgm}.
-We start, with the feature matrix X and generate a vector of block memberships $b$.
-The parameters of this step are encapsulated by $\theta$.
-Each feature vector $x_i$ is treated independently and used to generate the block membership $b_i$.
-Each entry $b_i \in \{1, 2 \dots B\}$ where $B$ is the number of blocks and a constant in our model.
-We choose a single softmax layer to model $p(b_i | x_i, \theta)$.
-More complex models are possible but then deriving meaning from inferred parameter distributions is more difficult.
-Summarising, we write $p(b | X, \theta)$ as follows:
-%
-\begin{equation}
-	p(b| X, \theta) = \prod_{i=1}^{N} p(b_i | x_i, \theta) = \prod_{i=1}^{N} \phi_{b_i} (x_i; \theta)
-	= \prod_{i=1}^{N} \frac{\exp\left(w_{b_i}^T \tilde{x}_i\right)}{\sum_{k=1}^{B} \exp \left( w_k^T \tilde{x}_i\right)}
-\end{equation}
-%
-Where $\tilde{x} \coloneqq \left[x_1, x_2, \dots x_D, 1\right]^T$ is an augmented version of $x$ that allows for a bias term.
-The parameters for this stage are denoted $\theta$ and consist of the all $B$ weight vectors $\theta = \{w_k\}_{k=1}^{B}$.
-Each $w_k$ has dimension $D+1$.
-We could instead write the parameters $\theta$ as a $B \times (D+1)$ matrix of weights $W$; this form has use computationally as then $Z_i = W \tilde{x}_i$ is the input to the softmax activation function.
-
-\begin{figure}[!h]
-	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=7mm},
-		squarednode/.style={rectangle, draw=black, minimum size=7mm}
-		]
-		% nodes
-		\node[roundnode] (X) at (0, 0) {$X$};
-		\node[squarednode] (b) at (3, 0) {$b$};
-		\node[roundnode] (A) at (6, 0) {$A$};
-
-		% arrows
-		\draw[->] (X.east) -- node[above] {$\theta$} (b.west);
-		\draw[->] (b.east) -- node[above] {$\psi$}(A.west);
-	\end{tikzpicture}
-	\caption{Latent block generative model}
-	\label{fig:lbgm}
-\end{figure}
-
-With the block memberships $b$ generated, we then draw the graph $A$ from the microcanical DC-SBM (\ref{defn:microcan-dc-sbm}) with additional parameters $\psi = \{e, k\}$.
-In a slight abuse of notation we denote the inter-block edge count matrix with $e = \psi_e$ and the degree sequence $k = \psi_k$ to make explicit that these parameters are contained in $\psi$.
-%
-\begin{equation}
-	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k)
-\end{equation}
-
-\subsection{Prior selection}
-
-Before performing any inference, we must specify priors on $\theta$ and $\psi$.
-For $\theta$ it seems sensible to choose a Gaussian prior, with zero mean and variance matrix $\sigma^2_\theta I$ such that each element of $\theta$ is independent and distributed like $\sim \Gaussian(0, \sigma_\theta^2)$.
-In vector form, the prior for $\theta$ is therefore:
-%
-\begin{equation}
-	p(\theta) = \Gaussian \left( \theta ; 0, \sigma_\theta^2 I \right)
-	\label{eqn:theta-prior}
-\end{equation}
-%
-We will see that this form of prior is equivalent to a regularisation term in neural network training that penalises extreme weight magnitudes.
-For $\sigma_\theta^2 \rightarrow \infty$ this becomes an uninformative uniform prior.
-
-In our model, the block memberships vector $b$ is an intermediate latent variable and so we are not free to choose a prior for it.
-Nevertheless, as far as inference on the right-hand-side of figure \ref{fig:lbgm}, we regard $p(b | X)$ as a pseudo-prior on $b$.
-We can show (appendix \ref{appdx:b|x}) that our choice of prior for $p(\theta)$ in equation \ref{eqn:theta-prior} leads to a uniform $p(b | X)$ in equation \ref{eqn:b-pseudo-prior}.
-%
-\begin{equation}
-	p(b | X) = \int p(b | X, \theta) p(\theta) d\theta = B^{-N}
-	\label{eqn:b-pseudo-prior}
-\end{equation}
-%
-This is a uniform distribution that only depends on the number of blocks $B$.
-This simplifies the problem nicely as evaluating $p(b | X)$ does not require knowing the exact value $X$ takes.
-\citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for the additional microcanonical SBM parameters $\psi$ which we adopt.
-Peixoto's idea is to write the joint prior on $(b, e, k)$ as a product of conditionals $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$.
-For our purposes we must insert a conditioning on $X$, to form our pseudo-prior for $b$ and $\psi$, to give equation \ref{eqn:joint-pseudo-prior}.
-%
-\begin{equation}
-	p(b, \psi | X) = p(b | X) p(\psi | b, X) = p(b | X) p(\psi | b)
-	\label{eqn:joint-pseudo-prior}
-\end{equation}
-%
-Where the simplification is made apparent by noting $(\psi \indep X) | b$. We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference.
-%
-\begin{equation}
-	p(\psi | b) = p(e | b) p(k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
-	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right]
-\end{equation}
-%
-Where $\specialchoose{n}{m}$ is shorthand for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of distinct histograms with $n$ bins under the constraint they sum to $m$.
-$E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges in the graph.
-Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform with respect to $e$.
-The variable $\eta_j^r$ is introduced to denote the number of vertices in block $r$ that have degree $j$.
-Formally, $\eta_j^r \coloneqq \sum_{i} \one\left\{b_i = r \right\} \one \left\{k_i = j \right\}$.
-Furthermore, $q(m, n)$ is the number of different histograms with at most $n$ non-zero bins that sum to $m$.
-$q(m, n)$ is related to but different from $\specialchoose{n}{m}$.
-Lastly, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ is the number of vertices assigned to block $r$.
-Importantly, we have computable forms for $p(\theta)$ and $p(b, \psi | X)$ which will be useful for performing inference.
-
-\section{Inference}
-
-Now that we have defined the LBGM, we wish to leverage it to perform inference.
-Suppose we are presented with a vertex-labelled graph $(A, X)$; the goal is to draw samples for $\theta$ according to the posterior given the observed graph (equation \ref{eqn:theta-target}).
-%
-\begin{equation}
-	\label{eqn:theta-target}
-	\theta^{(t)} \sim p(\theta | A, X)
-\end{equation}
-%
-These samples allow us to approximate the posterior distribution for $\theta$ as well as compute a predictive distribution $p(b^* | x^*, A, X) = \int p(b^* | x^*, \theta) p(\theta | X, A) d\theta \approx \frac{1}{T} \sum_{t=1}^{T} p(b^* | x^*, \theta^{(t)})$.
-However, generating these samples is not easily done in practice.
-
-We instead propose an iterative approach.
-First drawing samples $b^{(t)}$ from the block membership posterior (equation \ref{eqn:b-samples}).
-We then use each $b^{(i)}$ to draw samples for $\theta$ as in equation \ref{eqn:theta-samples}.
-%
-\begin{align}
-	b^{(i)} &\sim p \Big( b | A, X \Big)  \label{eqn:b-samples}\\
-	\theta^{(i)} &\sim p\Big(\theta | X, b^{(i)} \Big) \label{eqn:theta-samples}
-\end{align}
-%
-Both of these sampling steps implemented with a Markov Chain through the Metropolis-Hastings algorithm \cite{hastings-alg}.
-We just need to define a proposal distribution $q(x, x')$ for proposing a move $x \rightarrow x'$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(\cdot)$, point-wise.
-The proposed move is then accepted with probability $\alpha$ (equation \ref{eqn:mh-accept}) else it is rejected and we stay at $x$.
-%
-\begin{equation}
-	\alpha = \min \left( \frac{\pi(x') q(x', x)}{\pi(x) q(x, x')} , 1 \right)
-	\label{eqn:mh-accept}
-\end{equation}
-%
-This accept-reject step ensures the resulting Markov Chain is in detailed balance with the target distribution $\pi(\cdot)$.
-What we propose in equations \ref{eqn:b-samples} and \ref{eqn:theta-samples} is therefore implemented through a 2-level Markov chain.
-The resulting samples for $\theta^{(i)}$ are unbiased in the sense that the expectation of their distribution is the posterior we are targeting in equation \ref{eqn:theta-target}.
-%
-\begin{align*}
-	\Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] &= \sum_{b \in \mathcal{B}^N} p(\theta | X, b) p(b | A, X) \\
-	&= \sum_{b \in \mathcal{B}^N} p(\theta, b | A, X) \\
-	&= p(\theta | A, X) \\
-\end{align*}
-%
-This is an example of a pseudo-marginal approach \cite{pseudo-marginal}.
-Indeed, the unbiased result is sufficient to prove that for sufficient samples, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{t})\right] = p(\theta
-\ A, X)$ which is exactly the distribution we were targetting (equation \ref{eqn:theta-target}).
-
-The reason we split the Markov chain into two stages is because the summation over all latent states $b \in \mathcal{B}^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in \Bcal^N} p(A | b) P(b | X, \theta)$ is intractable $O(B^N)$.
-%
-\begin{figure}[!h]
-	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=12mm},
-		squarednode/.style={rectangle, draw=black, minimum size=12mm}
-		]
-		% nodes
-		\node[roundnode] (b0) at (0, 2) {$b^{(t)}$};
-		\node[roundnode] (b1) at (4, 2) {$b'$};
-		\node[roundnode] (b2) at (8, 2) {$b^{(t+1)}$};
-		\node[roundnode] (t0) at (0, 0) {$\theta^{(t)}$};
-		\node[roundnode] (t1) at (4, 0) {$\theta'$};
-		\node[roundnode] (t2) at (8, 0) {$\theta^{(t+1)}$};
-
-		% arrows
-		\draw[->] (b0) to node[above] {$\sim q_b(b^{(t)}, b' | A)$} (b1);
-		\draw[->] (b1) to node[above] {$\alpha_b (b, b' | A)$} (b2);
-		\draw[->] (b0) [out=30, in=150] to node[above] {$1-\alpha_b$} (b2);
-
-		\draw[->] (t0) to node[above] {$\sim q_\theta(\theta, \theta' | X, b^{(t)})$} (t1);
-		\draw[->] (t1) to node[above] {$\alpha_\theta (\theta, \theta' | X, b^{(t)})$} (t2);
-		\draw[->] (t0) [out=-30, in=-150] to node[below] {$1-\alpha_\theta$} (t2);
-
-		\draw[dashed, ->] (b0) to (2, 0.8);
-		\draw[dashed, ->] (b0) to (5.5, 0.8);
-
-	\end{tikzpicture}
-	\caption{Sampling sequence}
-	\label{fig:samp-sequence}
-\end{figure}
-%
-Figure \ref{fig:samp-sequence} shows an overview of the proposed method.
-We have introduced subscripts and conditionings to make explicit what parameters each step utilises.
-In an important simplification, we note that $p(b| X) = B^{-N}$ which does not depend on the exact value of X.
-Therefore, we do not need to know the value of $X$ to perform the sampling on $b$.
-Conversely, for the $\theta^{(t)}$ samples, we use $b^{(t)}$ but not $A$ as $(\theta \indep A) | b$.
-
-\subsection{Sampling block memberships}
-
-\citet{Peixoto-MCMC} proposes a Monte Carlo method which we will base our approach on. It relies on writing the posterior in the following form:
-%
-\begin{equation}
-	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) = \pi_b(b)
-\end{equation}
-%
-Now $\pi_b(\cdot)$ is the un-normalised density we wish to sample from.
-In other words, we wish to construct a Markov chain that has $\pi_b(\cdot)$ as its invariant distribution.
-We can break $\pi_b$ down as follows:
-%
-\begin{align*}
-	\pi_b(b) &= p(b|X) \sum_{\psi} \nolimits p(A , \psi | b, X) \\
-	&= p(b|X) p(A, \psi^* | b, X) \\
-	&= p(A | b, \psi^*) \cdot p(\psi^* | b) \cdot p(b | X)
-\end{align*}
-%
-Since we are using the microcanonical SBM formulation, there is only one value of $\psi$ that is compatible with the given $(A, b)$ pair.
-We denote this value $\psi^* = \{k^*, e^*\}$.
-Specifically, $k^*_i = \sum_j A_{ij} $ and $e^*_{rs} = \sum_{i, j} A_{ij} \one \{b_i=r\} \one\{b_j=s\}$.
-Therefore, the summation over all $\psi$ reduces to just the single $\psi^*$ term.
-We also define the microcanonical entropy of the configuration as.
-%
-\begin{equation}
-	S(b) = - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big)
-\end{equation}
-%
-This entropy can be thought of as the description length of the graph because it is the sum of the information required to represent the graph given the parameters and the amount of information required to store the parameters (given the feature matrix $X$).
-The exact, from of the proposal distribution and accept-reject step is explored throughly by \citet{Peixoto-MCMC}.
-There is a widely available library, \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, implementing this algorithm.
-The only modification we make is in the block membership prior $p(b)$ which we replace with $p(b|X)=B^{-N}$ which is a uniform distribution and so cancels out in the MH accept-reject step.
-
-The end result of this is that we can generate a set of block membership samples $\{b^{(t)}\}_{t=1}^{T}$ with each $b^{(t)} \sim p(b | A, X)$.
-Each of these samples can then be used for the $\theta$-chain.
-
-\subsection{Sampling feature-to-block classifier parameters}
-
-The invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. We write this as follows:
-%
-\begin{align}
-	p(\theta | X, b) &\propto p(b | X, \theta) p(\theta) = \pi_\theta (\theta) \propto  \exp \left( - U(\theta) \right) \\
-	\therefore U(\theta) &= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}
-\end{align}
-%
-Where we have introduced $U(\theta)$ equal to the negative log posterior.
-This $U(\theta)$ term makes subsequent analysis more concise.
-Each of the constituent terms of $U(\theta)$ is easily computed (equation \ref{eqn:u-form}).
-To simplify notation, we define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
-%
-\begin{equation}
-	\log p(b | X, \theta) = \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
-	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta ||^2
-	\label{eqn:U-constituent-terms}
-\end{equation}
-%
-This means that if we ignore constant terms we can write $U(\theta)$ as in equation \ref{eqn:U-form}.
-We see that the prior effectively introduces a regularisation term.
-Note that $||\theta||^2 = \sum_{i} \theta_{i}^2 = \sum_{j=1}^{B} ||w_j||^2$ is the Euclidean norm of the vector of parameters $\theta$.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form}
-\end{equation}
-%
-$U(\theta)$ in equation \ref{eqn:U-form} appears a typical objective function to be minimised for neural network training.
-The first term is the cross-entropy between actual and predicted labels.
-The second term - introduced by the prior - brings a form of regularisation, preventing over-fitting.
-In traditional applications we only seek the value of $\theta$ that minimises the objective function $U(\theta)$, which in our case would yield the maximum a posteriori (MAP) estimate.
-This is often done through some kind of gradient descent as $\nabla U$ is easily computable (equation \ref{eqn:U-derivative}).
-
-However, our goal is not to find the MAP estimate but to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
-As discussed earlier, given the invariant $\pi_\theta(\cdot)$, it is sufficient to specify a proposal distribution and then apply a MH accept-reject step to ensure detailed balance of the Markov Chain.
-Nevertheless, we can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
-We therefore adopt the Metropolis Adjusted Langevin Algorithm (MALA) - first proposed by \citet{mala-tweedie} - which leverages just that.
-Given the current sample $\theta$, we propose a new sample $\theta'$ according to equation \ref{eqn:theta-update}
-%
-\begin{equation}
-	\theta' = \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi
-	\label{eqn:theta-update}
-\end{equation}
-%
-Where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter - which may vary with the sample index.
-Indeed without the injected noise term, this is equivalent to gradient descent.
-We require the noise term to fully explore the parameter space.
-As such the proposal distribution, is a simple multivariate Gaussian which can be easily evaluated.
-%
-\begin{equation}
-	q_\theta(\theta, \theta') = \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right)
-\end{equation}
-%
-The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}). By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as:
-%
-\begin{equation}
-	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
-	\label{eqn:U-derivative}
-\end{equation}
-%
-After a proposed move is generated, in typical Metropolis-Hastings fashion we accept the move with probability $\alpha_\theta$.
-%
-\begin{equation}
-	\alpha_\theta(\theta, \theta') = \min \left( 
-	\exp \left( U(\theta) - U(\theta')\right)
-	\frac{ 
-		q_\theta(\theta', \theta)
-	}{
-		q_\theta(\theta, \theta')
-	} 
-	, 1 \right)
-\end{equation}
-%
-This fully specifies, the sampling procedure to generate $\{\theta^{(t)}\}_{t=1}^T$. Bear in mind that each $\theta^{(t)}$ update step uses its corresponding $b^{(t)}$ block membership sample.
-
-\subsection{Sampling sequence}
-
-So far, each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample.
-This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance.
-This may lead to longer burn-in and autocorrelation times of the resulting Markov Chain.
-The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends on $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
-We would rather deal with the expectation of each $y_{ij}^{(t)}$:
-%
-\begin{equation}
-	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one(b_{i}^{(t)} = j) \right]
-	= p(b_i = j | A, X)
-\end{equation}
-%
-We obtain an unbiased estimate for this quantity as simply the empirical distribution of the block membership samples $\left\{ b^{(t)} \right\}_{t=1}^T$.
-%
-\begin{equation}
-	\hat{y}_{ij} \coloneqq \frac{1}{T} \sum_{t=1}^{T} y_{ij}^{(t)} = \frac{1}{T} \sum_{t=1}^{T} \one\{b_i^{(t)} = j\}
-\end{equation}
-%
-We therefore, choose to feed each $\theta^{(t)}$ update step the same $\hat{y}_{ij}$ for all $t$ rather than the corresponding $y^{(t)}_{ij}$.
-This means we no longer need to run the $b$ and $\theta$ Markov chains concurrently.
-Instead, we run the $b$-chain to completion and use it to generate $\hat{y}_{ij}$ for $i \in \{1 \dots N\}$ and $j \in \{1 \dots B\}$.
-This is an estimate of $p(b | A, X)$ that we use for every iteration of the $\theta$ Markov chain.
-This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$; wed refer to these as $T_b$ and $T_\theta$ henceforth.
-Furthermore, this changeover reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{Block membership sample generation} % give the algorithm a caption
-	\label{alg:b-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\FOR{$t \in \{0, 1 \dots T_b - 1\}$}
-			\STATE $b' \gets \sim q_b(b^{(t)}, b' | A)$
-			\STATE $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
-			\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-			\IF{$\log \eta < \log \alpha_b$}
-				\STATE $b^{(t+1)} \gets b'$
-			\ELSE
-				\STATE $b^{(t+1)} \gets b^{(t)}$
-			\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{LBGM parameter pseudo-marginal inference} % give the algorithm a caption
-	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\STATE $\mathcal{S} \gets \textrm{range}(\textrm{burn-in}, T_b, \textrm{thinning})$
-		\STATE $\hat{Y}_{ij} \gets \frac{1}{\mathcal{|S|}} \sum_{t \in \mathcal{S}} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
-
-		\item[]
-
-		\FOR{$t \in \{0, 1 \dots T_\theta - 1\}$}
-		\STATE $\xi \gets \sim \Gaussian(0, I)$
-		\STATE $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
-		\STATE $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
-		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-		\IF{$\log \eta < \log \alpha_\theta$}
-		\STATE $\theta^{(t+1)} \gets \theta'$
-		\ELSE
-		\STATE $\theta^{(t+1)} \gets \theta^{(t)}$
-		\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\subsection{Complexity Analysis}
-
-\section{Experiments}
-
-We apply the developed methods to a variety of datasets.
-
-\begin{itemize}
-	\item \textbf{College football teams} \cite{Evans_2010_football} - network of American college football teams and their interactions. Vertex labels are the division a team belongs to.
-	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.
-	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier} - network of the author's Facebook friends list.
-Each vertex has been manually labelled with a variety of features describing their relationship to the author.
-	\item \textbf{Twitch users} \cite{twitch} - a network of user-user friendships on the streaming service Twitch.
-Vertex labels are extracted according to video-games played, location and streaming habits.
-This dataset is also broken down into disjoint networks according to language.
-We only consider the English users with is a subnet with $N=7126$ vertices and $E=35324$ edges).
-\end{itemize}
-
-\section{Extensions}
-
-\section{Conclusion}
-
-\begin{ack}
-Use unnumbered first level headings for the acknowledgments. All acknowledgments
-go at the end of the paper before the list of references. Moreover, you are required to declare
-funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
-More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2021/PaperInformation/FundingDisclosure}.
-
-Do {\bf not} include this section in the anonymized submission, only in the final paper.
-You can use the \texttt{ack} environment provided in the style file to automatically hide this section in the anonymized submission.
-\end{ack}
-
-%references
-\bibliography{sources.bib}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Checklist}
@@
 \end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\appendix
-
-\section{Appendix}
-
-\subsection{Derivation of conditional block distribution given feature matrix}
-\label{appdx:b|x}
-
-We wish to determine the form of $p(b| X)$. This can be done by integrating over the joint probability with respect to $\theta$.
-%
-\begin{align*}
-	p(b | X) &= \int p(b , \theta| X, \theta) d\theta = \int p(b | X, \theta) p(\theta | X) d\theta \\
-	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i=1}^{N} \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
-	&= \prod_{i=1}^{N} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j=1}^{B} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k=1}^{B} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}
-\end{align*}
-%
-We note that $b_i \in {1, 2, \dots B}$ and so the integral's value is unchanged with respect to $b_i$.
-The integrand has the same form no matter which value $b_i$ takes as the prior is the same for each $w_j$.
-As such the integral can only be a function of at most $\tilde{x}_i$ and $\sigma_\theta^2$ as it is symmetric with respect to $b_i$ and all the various $w_j$ are integrated out as they are dummy variables.
-Therefore, denoting the integral by the (unknown) function $f(\tilde{x}_i, \sigma_\theta^2)$, we write $p(b| X)$ as follows:
-%
-\begin{align*}
-	p(b | X) &= \prod_{i=1}^{N} f(\tilde{x}_i, \sigma_\theta^2) = \textrm{const w.r.t } b = c
-\end{align*}
-%
-As this is a constant with respect to $b$ we conclude that $p(b | X)$ must be a uniform distribution.
-$\nicefrac{1}{c}$ is simply the size of the set of values that $b$ can take.
-We know $b_i \in \mathcal{B} = \{1, 2, \dots B\}$.
-Therefore, $b \in \mathcal{B}^N$ and $|\mathcal{B}^N| = |\mathcal{B}|^N = B^N= \nicefrac{1}{c}$.
-Putting this all together we show that:
-%
-\begin{equation}
-	p(b | X) = B^{-N}
-\end{equation}
-
-\subsection{Derivation of U gradient with respect to feature parameters}
-\label{appdx:gradu}
-The goal is to determine $\nabla U(\theta)$, the gradient of the negative log posterior with respect to the parameters. We repeat the form of $U(\theta)$ in equation \ref{eqn:U-form-appdx}.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form-appdx}
-\end{equation}
-%
-Where $y_{ij}$ is independent of $\theta$ and $a_{ij}$ is the output from the softmax layer, with form as given in equation \ref{eqn:a-ij}.
-%
-\begin{equation}
-	a_{ij} \coloneqq \phi_{j} (x_i; \theta) = \frac{\exp(w_j^T \tilde{x}_i)}{\sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i)}
-	\label{eqn:a-ij} 
-\end{equation}
-%
-We note that $\theta = \{w_k\}_{k=1}^B$, and as such we can write this in vector form $\theta = \left[w_1^T, w_2^T \dots w_B^T  \right]^T$.
-Therefore, $\nabla U(\theta) = \left[\nicefrac{\partial U}{\partial w_1}^T,\nicefrac{\partial U}{\partial w_2}^T \dots \nicefrac{\partial U}{\partial w_B}^T  \right]^T$; to compute $\nabla U(\theta)$ it suffices to find the form of $\nicefrac{\partial U}{\partial w_k}$ with respect to a general $k$.
-
-To this end, we must first find partial derivatives of $a_{ij}$ and $||\theta||$ with respect to $w_k$. Starting with $a_{ij}$:
-%
-\begin{align}
-	\frac{\partial a_{ij}}{\partial w_k} &= \frac
-	{\tilde{x}_i \exp(w_j^T \tilde{x}_i) \delta_{jk} \cdot \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) 
-		- 
-	\exp(w_j^T \tilde{x}_i) \cdot \tilde{x}_i \exp(w_k^T \tilde{x}_i)}
-	{\left( \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) \right)^2} \nonumber \\
-	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right) 
-\end{align}
-%
-Where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$. Now moving onto the derivative of $||\theta||^2$:
-%
-\begin{equation}
-	\frac{ \partial}{\partial w_k} ||\theta||^2 = \frac{\partial}{\partial w_k} \left( \sum_{b=1}^B ||w_b||^2 \right) = 2w_k
-\end{equation}
-%
-We are ready to put this all together, to find the partial derivative of $U(\theta)$ with respect to each $w_k$:
-\begin{align}
-	\frac{\partial U}{\partial w_k} &= 
-	\sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} 
-	\left( \frac{-\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
-	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
-	&=  - \left( \sum_{i=1}^{N} \tilde{x}_i \left( y_{ik} - a_{ik} \sum_{j=1}^{B} y_{ij} \right)
-	- \frac{w_k}{\sigma_\theta^2} \right) \nonumber \\
-	&= - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
-\end{align}
-%
-This is the required result.
-This form can be computed efficiently through matrix operations.
-The only property of $y_{ij}$ we have used in the derivation is the sum-to-one constraint $\sum_{j=1}^{B} y_{ij} = 1$ for all $i$.
-
-\subsection{Choosing the MALA step-size}
-\label{appdx:step-size}
-
-For sampling from the $\theta$-chain of the block membership generator parameters, we employed the Metropolis Adjusted Langevin Algorithm (MALA). At iteration $t$, the proposed sample is generated by:
-%
-\begin{equation}
-	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi
-\end{equation}
-%
-There are two competing objectives when choosing the step-size $h_t$.
-On the one hand, we want the step-size to be large so that we arrive at a high density region quickly.
-However, too large a step-size will lead to a lower acceptance ratio and thus inefficient sampling.
-A solution to this problem would be to slowly decrease the step-size with $t$ - often called simulated annealing.
-Therefore, we still have a short burn-in time but will not bounce around the mode for large $t$.
-As well as the trivial constraint for $h_t$ to be strictly positive and we introduce two further constraints as outlined by \citet{Bayesian-SGLD}:
-%
-\begin{equation}
-	\sum_{t=1}^{\infty} h_t = \infty \qquad \textrm{and} \qquad
-	\sum_{t=1}^{\infty} h_t^2 < \infty
-	\label{eqn:h-constraints}
-\end{equation}
-%
-The first constraint ensures that we have cover sufficient distance to arrive at any arbitrary point in our domain, no matter the starting point.
-The second constraint ensures that once we converge to the mode rather than simply bouncing around it.
-\citet{Bayesian-SGLD} propose the following form for a polynomially decaying step-size which we adopt:
-%
-\begin{equation}
-	h_t = a(b + t)^{-\gamma}
-\end{equation}
-%
-Where $a, b, \gamma$ are hyper-parameters to be chosen.
-We require $a,b > 0$ and $\gamma \in (0.5, 1]$ to satisfy equation \ref{eqn:h-constraints}.
-We find empirically that $a=250, b=1000, \gamma=0.8$ yield good burn-in times for our MALA implementations.
-
-\end{document}

### 2021-05-16 nips/experiments.tex
Split into multiple tex files

--- 2021-05-15 nips/submission.tex
@@
-\documentclass{article}
-
-% if you need to pass options to natbib, use, e.g.:
-%     \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading neurips_2021
-
-% ready for submission
-\PassOptionsToPackage{numbers, compress}{natbib}
-\usepackage[final]{neurips_2021}
-
-% to compile a preprint version, e.g., for submission to arXiv, add add the
-% [preprint] option:
-%     \usepackage[preprint]{neurips_2021}
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-%\usepackage[final]{neurips_2021}
-
-% to avoid loading the natbib package, add option nonatbib:
-%    \usepackage[nonatbib]{neurips_2021}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{xcolor}         % colors
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-\usepackage{amsmath}
-\usepackage{mathtools}
-\usepackage{bbm}
-\usepackage{amssymb}
-\usepackage{algorithmic}
-\usepackage{algorithm}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{graphicx}
-
-%images
-\graphicspath{{./img/}}
-
-%bibliography
-\bibliographystyle{plainnat}
-
-\title{Inferring community characteristics in labelled networks}
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to break the
-% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
-% authors names on the first line, and the last on the second line, try using
-% \AND instead of \And before the third author name.
-
-\author{
-  Lawrence Tray \\
-  Department of Engineering \\
-  University of Cambridge \\
-  \texttt{lpt30@cam.ac.uk} \\
-  % examples of more authors
-  \And
-  Ioannis Kontoyiannis \\
-  Department of Mathematics \\
-  University of Cambridge \\
-  \texttt{ik355@cam.ac.uk}
-}
-
-%custom commands
-\newcommand{\Xcal}{\mathcal{X}}
-\newcommand{\Bcal}{\mathcal{B}}
-\newcommand{\Integers}{\mathbb{Z}}
-\newcommand{\one}{\mathbbm{1}}
-\newcommand{\Gaussian}{\mathcal{N}}
-\newcommand{\indep}{\perp \!\!\! \perp}
-\newcommand{\specialchoose}{\genfrac{\{}{\}}{0pt}{}}
-\newcommand{\Expect}{\mathbb{E}}
-
-%envs
-\newtheorem{definition}{Definition}[section]
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[theorem]
-\newtheorem{lemma}[theorem]{Lemma}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-We present a novel generative model for describing vertex-labelled undirected graphs.
-This is a two-layered model.
-First a probabilistic mapping from vertex labels to latent block memberships and from these block memberships we generate a graph according to the micro-canonical stochastic block model.
-
-With this framework we are able to efficiently sample the parameters of the vertex label to block membership mapping.
-This allows us to automatically determine which features have the largest impact on graphical structure.
-\end{abstract}
-
-\section{Introduction}
-
-There is a wealth of networks in the world.
-Some examples of this graphical data are social networks, website hyperlinks and academic collaboration with more being produced each second.
-It is clear we need tools to analyse this increasingly omni-present form of data.
-
-A somewhat surprising property of real-world graphs is that they exhibit strong community structure.
-In other words, each node will often belong to a cluster of densely connected nodes.
-This property is often exploited by graph compression algorithms and there is high interest in recovering the communities from the observed graph.
-
-A common subset of graphical data is the labelled network.
-This is a graph where we have information about the properties of each node.
-We shall refer to these node properties as features.
-One of the most common questions we can ask of this dataset is what features have the largest impact on the structure of the graph.
-For example, when analysing an academic collaboration graph, one may wish to ask what impact gender has on the structure of the network.
-Nevertheless, this analysis is often vulnerable to confounding variables.
-While gender very well may impact the structure of the graph, there is often a better explanatory variable for the structure.
-
-There is space to bridge the gap between these two approaches.
-Rather than determining whether a feature impacts the graphical structure directly, we can use the community concept as a stepping stone in our analysis.
-Therefore, we extract which features have the largest impact on overall graphical structure.
-This can be thought of as a form of dimensionality-reduction where we only keep the node features that are useful in predicting community memberships.
-
-\section{Preliminaries}
-
-This section defines some preliminary concepts required for the subsequent analysis.
-We first need a model for community-like structure in a graphical network.
-For this we adopt the stochastic block model (SBM) - widely used across academia.
-The premise is that each node in the graph belongs to a unique community called a block.
-The probability that two nodes are connected depends only on the block memberships of each.
-Graphs drawn from this ensemble exhibit community structure.
-Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}.
-A paraphrased definition is given below.
-
-\begin{definition}[Microcanonical NDC-SBM]
-	\label{defn:microcan-ndc-sbm}
-	Let $N \in \Integers^{+}$ denote the number of vertices in an undirected graph.
-The block memberships are encoded by a vector $b$ of length $N$ where each entry $b_i \in \{1, 2 \dots B\}$.
-$B \in \Integers^{+}$ is the number of non-empty blocks.
-Let $e$ be a $B \times B$ matrix of edge counts between blocks ($e_{rs}$ is number of edges from block $r$ onto block $s$ - or twice that number if $r=s$).
-For undirected graphs $e$ is symmetric.
-For a non-degree-corrected stochastic block model (NDC-SBM), we say that the graph $A$ is generated as follows:
-	%
-	\begin{equation}
-		A \sim \textrm{NDC-SBM}_{\textrm{MC}} (b, e)
-	\end{equation}
-	%
-	Where edges are placed uniformly at random but respecting the constraint imposed by $e$ and $b$.
-The additional parameters $N$ and $B$ are omitted as they are inferred from the shapes of $b$ and $e$.
-If we interpret $A$ as an adjacency matrix, then this constraint can be written formally as: $e_{rs} = \sum_{i,j} A_{ij} \one \{b_i = r\} \one \{b_j = s\}$.
-\end{definition}
-
-Nevertheless, this formulation does not accept high degree variability within blocks as is typical of real-world data.
-Indeed, the NDC-SBM favours a partition into high-degree and low-degree nodes rather than clusters of inter-connected nodes.
-We therefore introduce the degree-corrected SBM (DC-SBM) to circumvent these issues.
-
-\begin{definition}[Microcanonical DC-SBM]
-	\label{defn:microcan-dc-sbm}
-	This is much like the NDC-SBM but has an additional parameter $k$ which is an $N$-length vector encoding the degree sequence ($k_i$ is the degree of vertex $i$). Therefore, we write:
-	%
-	\begin{equation}
-		A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, e, k)
-	\end{equation}
-	%
-	Once again, edges are placed uniformly at random but respecting the constraints imposed by the parameters.
-The DC-SBM has the additional constraint that $k_i = \sum_{j} A_{ij}$.
-In what follows, we will always assume the degree-corrected model unless otherwise specified.
-
-\end{definition}
-
-\section{Latent block generative model}
-
-In this section we propose a novel generative model for modelling labelled networks.
-We restrict our analysis to labelled, undirected graphs with $N$ nodes.
-We define the vector $x_i \in \Xcal^D$ as the feature vector for the $i$'th vertex.
-Each vertex has $D$ total features and we assume all entries take values from the same set $\Xcal$.
-For the majority of datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$.
-The feature vectors $\{x_i\}_{i=1}^{N}$ are subsumed into the $N \times D$ matrix $X$.
-
-The proposed generative model (which we call the latent block generative model - LBGM) is given in figure \ref{fig:lbgm}.
-We start, with the feature matrix X and generate a vector of block memberships $b$.
-The parameters of this step are encapsulated by $\theta$.
-Each feature vector $x_i$ is treated independently and used to generate the block membership $b_i$.
-Each entry $b_i \in \{1, 2 \dots B\}$ where $B$ is the number of blocks and a constant in our model.
-We choose a single softmax layer to model $p(b_i | x_i, \theta)$.
-More complex models are possible but then deriving meaning from inferred parameter distributions is more difficult.
-Summarising, we write $p(b | X, \theta)$ as follows:
-%
-\begin{equation}
-	p(b| X, \theta) = \prod_{i=1}^{N} p(b_i | x_i, \theta) = \prod_{i=1}^{N} \phi_{b_i} (x_i; \theta)
-	= \prod_{i=1}^{N} \frac{\exp\left(w_{b_i}^T \tilde{x}_i\right)}{\sum_{k=1}^{B} \exp \left( w_k^T \tilde{x}_i\right)}
-\end{equation}
-%
-Where $\tilde{x} \coloneqq \left[x_1, x_2, \dots x_D, 1\right]^T$ is an augmented version of $x$ that allows for a bias term.
-The parameters for this stage are denoted $\theta$ and consist of the all $B$ weight vectors $\theta = \{w_k\}_{k=1}^{B}$.
-Each $w_k$ has dimension $D+1$.
-We could instead write the parameters $\theta$ as a $B \times (D+1)$ matrix of weights $W$; this form has use computationally as then $Z_i = W \tilde{x}_i$ is the input to the softmax activation function.
-
-\begin{figure}[!h]
-	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=7mm},
-		squarednode/.style={rectangle, draw=black, minimum size=7mm}
-		]
-		% nodes
-		\node[roundnode] (X) at (0, 0) {$X$};
-		\node[squarednode] (b) at (3, 0) {$b$};
-		\node[roundnode] (A) at (6, 0) {$A$};
-
-		% arrows
-		\draw[->] (X.east) -- node[above] {$\theta$} (b.west);
-		\draw[->] (b.east) -- node[above] {$\psi$}(A.west);
-	\end{tikzpicture}
-	\caption{Latent block generative model}
-	\label{fig:lbgm}
-\end{figure}
-
-With the block memberships $b$ generated, we then draw the graph $A$ from the microcanical DC-SBM (\ref{defn:microcan-dc-sbm}) with additional parameters $\psi = \{e, k\}$.
-In a slight abuse of notation we denote the inter-block edge count matrix with $e = \psi_e$ and the degree sequence $k = \psi_k$ to make explicit that these parameters are contained in $\psi$.
-%
-\begin{equation}
-	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k)
-\end{equation}
-
-\subsection{Prior selection}
-
-Before performing any inference, we must specify priors on $\theta$ and $\psi$.
-For $\theta$ it seems sensible to choose a Gaussian prior, with zero mean and variance matrix $\sigma^2_\theta I$ such that each element of $\theta$ is independent and distributed like $\sim \Gaussian(0, \sigma_\theta^2)$.
-In vector form, the prior for $\theta$ is therefore:
-%
-\begin{equation}
-	p(\theta) = \Gaussian \left( \theta ; 0, \sigma_\theta^2 I \right)
-	\label{eqn:theta-prior}
-\end{equation}
-%
-We will see that this form of prior is equivalent to a regularisation term in neural network training that penalises extreme weight magnitudes.
-For $\sigma_\theta^2 \rightarrow \infty$ this becomes an uninformative uniform prior.
-
-In our model, the block memberships vector $b$ is an intermediate latent variable and so we are not free to choose a prior for it.
-Nevertheless, as far as inference on the right-hand-side of figure \ref{fig:lbgm}, we regard $p(b | X)$ as a pseudo-prior on $b$.
-We can show (appendix \ref{appdx:b|x}) that our choice of prior for $p(\theta)$ in equation \ref{eqn:theta-prior} leads to a uniform $p(b | X)$ in equation \ref{eqn:b-pseudo-prior}.
-%
-\begin{equation}
-	p(b | X) = \int p(b | X, \theta) p(\theta) d\theta = B^{-N}
-	\label{eqn:b-pseudo-prior}
-\end{equation}
-%
-This is a uniform distribution that only depends on the number of blocks $B$.
-This simplifies the problem nicely as evaluating $p(b | X)$ does not require knowing the exact value $X$ takes.
-\citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for the additional microcanonical SBM parameters $\psi$ which we adopt.
-Peixoto's idea is to write the joint prior on $(b, e, k)$ as a product of conditionals $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$.
-For our purposes we must insert a conditioning on $X$, to form our pseudo-prior for $b$ and $\psi$, to give equation \ref{eqn:joint-pseudo-prior}.
-%
-\begin{equation}
-	p(b, \psi | X) = p(b | X) p(\psi | b, X) = p(b | X) p(\psi | b)
-	\label{eqn:joint-pseudo-prior}
-\end{equation}
-%
-Where the simplification is made apparent by noting $(\psi \indep X) | b$. We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference.
-%
-\begin{equation}
-	p(\psi | b) = p(e | b) p(k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
-	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right]
-\end{equation}
-%
-Where $\specialchoose{n}{m}$ is shorthand for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of distinct histograms with $n$ bins under the constraint they sum to $m$.
-$E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges in the graph.
-Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform with respect to $e$.
-The variable $\eta_j^r$ is introduced to denote the number of vertices in block $r$ that have degree $j$.
-Formally, $\eta_j^r \coloneqq \sum_{i} \one\left\{b_i = r \right\} \one \left\{k_i = j \right\}$.
-Furthermore, $q(m, n)$ is the number of different histograms with at most $n$ non-zero bins that sum to $m$.
-$q(m, n)$ is related to but different from $\specialchoose{n}{m}$.
-Lastly, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ is the number of vertices assigned to block $r$.
-Importantly, we have computable forms for $p(\theta)$ and $p(b, \psi | X)$ which will be useful for performing inference.
-
-\section{Inference}
-
-Now that we have defined the LBGM, we wish to leverage it to perform inference.
-Suppose we are presented with a vertex-labelled graph $(A, X)$; the goal is to draw samples for $\theta$ according to the posterior given the observed graph (equation \ref{eqn:theta-target}).
-%
-\begin{equation}
-	\label{eqn:theta-target}
-	\theta^{(t)} \sim p(\theta | A, X)
-\end{equation}
-%
-These samples allow us to approximate the posterior distribution for $\theta$ as well as compute a predictive distribution $p(b^* | x^*, A, X) = \int p(b^* | x^*, \theta) p(\theta | X, A) d\theta \approx \frac{1}{T} \sum_{t=1}^{T} p(b^* | x^*, \theta^{(t)})$.
-However, generating these samples is not easily done in practice.
-
-We instead propose an iterative approach.
-First drawing samples $b^{(t)}$ from the block membership posterior (equation \ref{eqn:b-samples}).
-We then use each $b^{(i)}$ to draw samples for $\theta$ as in equation \ref{eqn:theta-samples}.
-%
-\begin{align}
-	b^{(i)} &\sim p \Big( b | A, X \Big)  \label{eqn:b-samples}\\
-	\theta^{(i)} &\sim p\Big(\theta | X, b^{(i)} \Big) \label{eqn:theta-samples}
-\end{align}
-%
-Both of these sampling steps implemented with a Markov Chain through the Metropolis-Hastings algorithm \cite{hastings-alg}.
-We just need to define a proposal distribution $q(x, x')$ for proposing a move $x \rightarrow x'$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(\cdot)$, point-wise.
-The proposed move is then accepted with probability $\alpha$ (equation \ref{eqn:mh-accept}) else it is rejected and we stay at $x$.
-%
-\begin{equation}
-	\alpha = \min \left( \frac{\pi(x') q(x', x)}{\pi(x) q(x, x')} , 1 \right)
-	\label{eqn:mh-accept}
-\end{equation}
-%
-This accept-reject step ensures the resulting Markov Chain is in detailed balance with the target distribution $\pi(\cdot)$.
-What we propose in equations \ref{eqn:b-samples} and \ref{eqn:theta-samples} is therefore implemented through a 2-level Markov chain.
-The resulting samples for $\theta^{(i)}$ are unbiased in the sense that the expectation of their distribution is the posterior we are targeting in equation \ref{eqn:theta-target}.
-%
-\begin{align*}
-	\Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] &= \sum_{b \in \mathcal{B}^N} p(\theta | X, b) p(b | A, X) \\
-	&= \sum_{b \in \mathcal{B}^N} p(\theta, b | A, X) \\
-	&= p(\theta | A, X) \\
-\end{align*}
-%
-This is an example of a pseudo-marginal approach \cite{pseudo-marginal}.
-Indeed, the unbiased result is sufficient to prove that for sufficient samples, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{t})\right] = p(\theta
-\ A, X)$ which is exactly the distribution we were targetting (equation \ref{eqn:theta-target}).
-
-The reason we split the Markov chain into two stages is because the summation over all latent states $b \in \mathcal{B}^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in \Bcal^N} p(A | b) P(b | X, \theta)$ is intractable $O(B^N)$.
-%
-\begin{figure}[!h]
-	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=12mm},
-		squarednode/.style={rectangle, draw=black, minimum size=12mm}
-		]
-		% nodes
-		\node[roundnode] (b0) at (0, 2) {$b^{(t)}$};
-		\node[roundnode] (b1) at (4, 2) {$b'$};
-		\node[roundnode] (b2) at (8, 2) {$b^{(t+1)}$};
-		\node[roundnode] (t0) at (0, 0) {$\theta^{(t)}$};
-		\node[roundnode] (t1) at (4, 0) {$\theta'$};
-		\node[roundnode] (t2) at (8, 0) {$\theta^{(t+1)}$};
-
-		% arrows
-		\draw[->] (b0) to node[above] {$\sim q_b(b^{(t)}, b' | A)$} (b1);
-		\draw[->] (b1) to node[above] {$\alpha_b (b, b' | A)$} (b2);
-		\draw[->] (b0) [out=30, in=150] to node[above] {$1-\alpha_b$} (b2);
-
-		\draw[->] (t0) to node[above] {$\sim q_\theta(\theta, \theta' | X, b^{(t)})$} (t1);
-		\draw[->] (t1) to node[above] {$\alpha_\theta (\theta, \theta' | X, b^{(t)})$} (t2);
-		\draw[->] (t0) [out=-30, in=-150] to node[below] {$1-\alpha_\theta$} (t2);
-
-		\draw[dashed, ->] (b0) to (2, 0.8);
-		\draw[dashed, ->] (b0) to (5.5, 0.8);
-
-	\end{tikzpicture}
-	\caption{Sampling sequence}
-	\label{fig:samp-sequence}
-\end{figure}
-%
-Figure \ref{fig:samp-sequence} shows an overview of the proposed method.
-We have introduced subscripts and conditionings to make explicit what parameters each step utilises.
-In an important simplification, we note that $p(b| X) = B^{-N}$ which does not depend on the exact value of X.
-Therefore, we do not need to know the value of $X$ to perform the sampling on $b$.
-Conversely, for the $\theta^{(t)}$ samples, we use $b^{(t)}$ but not $A$ as $(\theta \indep A) | b$.
-
-\subsection{Sampling block memberships}
-
-\citet{Peixoto-MCMC} proposes a Monte Carlo method which we will base our approach on. It relies on writing the posterior in the following form:
-%
-\begin{equation}
-	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) = \pi_b(b)
-\end{equation}
-%
-Now $\pi_b(\cdot)$ is the un-normalised density we wish to sample from.
-In other words, we wish to construct a Markov chain that has $\pi_b(\cdot)$ as its invariant distribution.
-We can break $\pi_b$ down as follows:
-%
-\begin{align*}
-	\pi_b(b) &= p(b|X) \sum_{\psi} \nolimits p(A , \psi | b, X) \\
-	&= p(b|X) p(A, \psi^* | b, X) \\
-	&= p(A | b, \psi^*) \cdot p(\psi^* | b) \cdot p(b | X)
-\end{align*}
-%
-Since we are using the microcanonical SBM formulation, there is only one value of $\psi$ that is compatible with the given $(A, b)$ pair.
-We denote this value $\psi^* = \{k^*, e^*\}$.
-Specifically, $k^*_i = \sum_j A_{ij} $ and $e^*_{rs} = \sum_{i, j} A_{ij} \one \{b_i=r\} \one\{b_j=s\}$.
-Therefore, the summation over all $\psi$ reduces to just the single $\psi^*$ term.
-We also define the microcanonical entropy of the configuration as.
-%
-\begin{equation}
-	S(b) = - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big)
-\end{equation}
-%
-This entropy can be thought of as the description length of the graph because it is the sum of the information required to represent the graph given the parameters and the amount of information required to store the parameters (given the feature matrix $X$).
-The exact, from of the proposal distribution and accept-reject step is explored throughly by \citet{Peixoto-MCMC}.
-There is a widely available library, \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, implementing this algorithm.
-The only modification we make is in the block membership prior $p(b)$ which we replace with $p(b|X)=B^{-N}$ which is a uniform distribution and so cancels out in the MH accept-reject step.
-
-The end result of this is that we can generate a set of block membership samples $\{b^{(t)}\}_{t=1}^{T}$ with each $b^{(t)} \sim p(b | A, X)$.
-Each of these samples can then be used for the $\theta$-chain.
-
-\subsection{Sampling feature-to-block classifier parameters}
-
-The invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. We write this as follows:
-%
-\begin{align}
-	p(\theta | X, b) &\propto p(b | X, \theta) p(\theta) = \pi_\theta (\theta) \propto  \exp \left( - U(\theta) \right) \\
-	\therefore U(\theta) &= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}
-\end{align}
-%
-Where we have introduced $U(\theta)$ equal to the negative log posterior.
-This $U(\theta)$ term makes subsequent analysis more concise.
-Each of the constituent terms of $U(\theta)$ is easily computed (equation \ref{eqn:u-form}).
-To simplify notation, we define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
-%
-\begin{equation}
-	\log p(b | X, \theta) = \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
-	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta ||^2
-	\label{eqn:U-constituent-terms}
-\end{equation}
-%
-This means that if we ignore constant terms we can write $U(\theta)$ as in equation \ref{eqn:U-form}.
-We see that the prior effectively introduces a regularisation term.
-Note that $||\theta||^2 = \sum_{i} \theta_{i}^2 = \sum_{j=1}^{B} ||w_j||^2$ is the Euclidean norm of the vector of parameters $\theta$.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form}
-\end{equation}
-%
-$U(\theta)$ in equation \ref{eqn:U-form} appears a typical objective function to be minimised for neural network training.
-The first term is the cross-entropy between actual and predicted labels.
-The second term - introduced by the prior - brings a form of regularisation, preventing over-fitting.
-In traditional applications we only seek the value of $\theta$ that minimises the objective function $U(\theta)$, which in our case would yield the maximum a posteriori (MAP) estimate.
-This is often done through some kind of gradient descent as $\nabla U$ is easily computable (equation \ref{eqn:U-derivative}).
-
-However, our goal is not to find the MAP estimate but to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
-As discussed earlier, given the invariant $\pi_\theta(\cdot)$, it is sufficient to specify a proposal distribution and then apply a MH accept-reject step to ensure detailed balance of the Markov Chain.
-Nevertheless, we can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
-We therefore adopt the Metropolis Adjusted Langevin Algorithm (MALA) - first proposed by \citet{mala-tweedie} - which leverages just that.
-Given the current sample $\theta$, we propose a new sample $\theta'$ according to equation \ref{eqn:theta-update}
-%
-\begin{equation}
-	\theta' = \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi
-	\label{eqn:theta-update}
-\end{equation}
-%
-Where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter - which may vary with the sample index.
-Indeed without the injected noise term, this is equivalent to gradient descent.
-We require the noise term to fully explore the parameter space.
-As such the proposal distribution, is a simple multivariate Gaussian which can be easily evaluated.
-%
-\begin{equation}
-	q_\theta(\theta, \theta') = \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right)
-\end{equation}
-%
-The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}). By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as:
-%
-\begin{equation}
-	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
-	\label{eqn:U-derivative}
-\end{equation}
-%
-After a proposed move is generated, in typical Metropolis-Hastings fashion we accept the move with probability $\alpha_\theta$.
-%
-\begin{equation}
-	\alpha_\theta(\theta, \theta') = \min \left( 
-	\exp \left( U(\theta) - U(\theta')\right)
-	\frac{ 
-		q_\theta(\theta', \theta)
-	}{
-		q_\theta(\theta, \theta')
-	} 
-	, 1 \right)
-\end{equation}
-%
-This fully specifies, the sampling procedure to generate $\{\theta^{(t)}\}_{t=1}^T$. Bear in mind that each $\theta^{(t)}$ update step uses its corresponding $b^{(t)}$ block membership sample.
-
-\subsection{Sampling sequence}
-
-So far, each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample.
-This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance.
-This may lead to longer burn-in and autocorrelation times of the resulting Markov Chain.
-The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends on $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
-We would rather deal with the expectation of each $y_{ij}^{(t)}$:
-%
-\begin{equation}
-	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one(b_{i}^{(t)} = j) \right]
-	= p(b_i = j | A, X)
-\end{equation}
-%
-We obtain an unbiased estimate for this quantity as simply the empirical distribution of the block membership samples $\left\{ b^{(t)} \right\}_{t=1}^T$.
-%
-\begin{equation}
-	\hat{y}_{ij} \coloneqq \frac{1}{T} \sum_{t=1}^{T} y_{ij}^{(t)} = \frac{1}{T} \sum_{t=1}^{T} \one\{b_i^{(t)} = j\}
-\end{equation}
-%
-We therefore, choose to feed each $\theta^{(t)}$ update step the same $\hat{y}_{ij}$ for all $t$ rather than the corresponding $y^{(t)}_{ij}$.
-This means we no longer need to run the $b$ and $\theta$ Markov chains concurrently.
-Instead, we run the $b$-chain to completion and use it to generate $\hat{y}_{ij}$ for $i \in \{1 \dots N\}$ and $j \in \{1 \dots B\}$.
-This is an estimate of $p(b | A, X)$ that we use for every iteration of the $\theta$ Markov chain.
-This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$; wed refer to these as $T_b$ and $T_\theta$ henceforth.
-Furthermore, this changeover reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{Block membership sample generation} % give the algorithm a caption
-	\label{alg:b-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\FOR{$t \in \{0, 1 \dots T_b - 1\}$}
-			\STATE $b' \gets \sim q_b(b^{(t)}, b' | A)$
-			\STATE $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
-			\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-			\IF{$\log \eta < \log \alpha_b$}
-				\STATE $b^{(t+1)} \gets b'$
-			\ELSE
-				\STATE $b^{(t+1)} \gets b^{(t)}$
-			\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{LBGM parameter pseudo-marginal inference} % give the algorithm a caption
-	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\STATE $\mathcal{S} \gets \textrm{range}(\textrm{burn-in}, T_b, \textrm{thinning})$
-		\STATE $\hat{Y}_{ij} \gets \frac{1}{\mathcal{|S|}} \sum_{t \in \mathcal{S}} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
-
-		\item[]
-
-		\FOR{$t \in \{0, 1 \dots T_\theta - 1\}$}
-		\STATE $\xi \gets \sim \Gaussian(0, I)$
-		\STATE $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
-		\STATE $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
-		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-		\IF{$\log \eta < \log \alpha_\theta$}
-		\STATE $\theta^{(t+1)} \gets \theta'$
-		\ELSE
-		\STATE $\theta^{(t+1)} \gets \theta^{(t)}$
-		\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\subsection{Complexity Analysis}
-
 \section{Experiments}
@@
 
+A table summarising the results is given in table \ref{tab:results}.
+
+\begin{table}[]
+	\centering
+	\begin{tabular}{c|ccc|c|cc}
+		Dataset & $N$ & $E$ & $D$ & $B$ & $\bar{S}_b /N$ & $\bar{U}_\theta /N$ \\ \hline
+		Primary school & 238 & 5539 & 13 &  18 & 43.0  & 2.75 \\
+		Maier FB egonet &  &     &     &   &           &             \\
+		&     &     &     &              &     &       
+	\end{tabular}
+\end{table}
+
 \subsection{Primary school dynamic contacts}
@@
 	\centering
-	\begin{subfigure}{0.4\textwidth}
+	\begin{subfigure}{0.45\linewidth}
 		\centering
-		\includegraphics{width=\textwidth}{school-graph.png}
-		\caption{Graph with nodes coloured to show soft inferred block memberships $\hat{Y}_ij \approx p(b_i=j|A, X)$}
+		\includegraphics[width=\linewidth]{school-graph.png}
+		\caption{Softmax weighting samples for each block index}
+		\label{fig:school-graph}
 	\end{subfigure}
+	\hfill
+	\begin{subfigure}{0.45\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{school-null.png}
+		\caption{}
+		\label{fig:school-null}
+	\end{subfigure}
+	\caption{Primary school dynamic contacts network \cite{schools}}
 \end{figure}
-
-\section{Extensions}
-
-\section{Conclusion}
-
-\begin{ack}
-Use unnumbered first level headings for the acknowledgments. All acknowledgments
-go at the end of the paper before the list of references. Moreover, you are required to declare
-funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
-More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2021/PaperInformation/FundingDisclosure}.
-
-Do {\bf not} include this section in the anonymized submission, only in the final paper.
-You can use the \texttt{ack} environment provided in the style file to automatically hide this section in the anonymized submission.
-\end{ack}
-
-%references
-\bibliography{sources.bib}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Checklist}
-
-%%% BEGIN INSTRUCTIONS %%%
-The checklist follows the references.  Please
-read the checklist guidelines carefully for information on how to answer these
-questions.  For each question, change the default \answerTODO{} to \answerYes{},
-\answerNo{}, or \answerNA{}.  You are strongly encouraged to include a {\bf
-justification to your answer}, either by referencing the appropriate section of
-your paper or providing a brief inline description.  For example:
-\begin{itemize}
-  \item Did you include the license to the code and datasets? \answerYes{See Section gen-inst.}
-  \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
-  \item Did you include the license to the code and datasets? \answerNA{}
-\end{itemize}
-Please do not modify the questions and only use the provided macros for your
-answers.  Note that the Checklist section does not count towards the page
-limit.  In your paper, please delete this instructions block and only keep the
-Checklist section heading above along with the questions/answers below.
-%%% END INSTRUCTIONS %%%
-
-\begin{enumerate}
-
-\item For all authors...
-\begin{enumerate}
-  \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
-    \answerTODO{}
-  \item Did you describe the limitations of your work?
-    \answerTODO{}
-  \item Did you discuss any potential negative societal impacts of your work?
-    \answerTODO{}
-  \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you are including theoretical results...
-\begin{enumerate}
-  \item Did you state the full set of assumptions of all theoretical results?
-    \answerTODO{}
-	\item Did you include complete proofs of all theoretical results?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you ran experiments...
-\begin{enumerate}
-  \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
-    \answerTODO{}
-  \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
-    \answerTODO{}
-	\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
-    \answerTODO{}
-	\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
-\begin{enumerate}
-  \item If your work uses existing assets, did you cite the creators?
-    \answerTODO{}
-  \item Did you mention the license of the assets?
-    \answerTODO{}
-  \item Did you include any new assets either in the supplemental material or as a URL?
-    \answerTODO{}
-  \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
-    \answerTODO{}
-  \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you used crowdsourcing or conducted research with human subjects...
-\begin{enumerate}
-  \item Did you include the full text of instructions given to participants and screenshots, if applicable?
-    \answerTODO{}
-  \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
-    \answerTODO{}
-  \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
-    \answerTODO{}
-\end{enumerate}
-
-\end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\appendix
-
-\section{Appendix}
-
-\subsection{Derivation of conditional block distribution given feature matrix}
-\label{appdx:b|x}
-
-We wish to determine the form of $p(b| X)$. This can be done by integrating over the joint probability with respect to $\theta$.
-%
-\begin{align*}
-	p(b | X) &= \int p(b , \theta| X, \theta) d\theta = \int p(b | X, \theta) p(\theta | X) d\theta \\
-	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i=1}^{N} \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
-	&= \prod_{i=1}^{N} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j=1}^{B} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k=1}^{B} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}
-\end{align*}
-%
-We note that $b_i \in {1, 2, \dots B}$ and so the integral's value is unchanged with respect to $b_i$.
-The integrand has the same form no matter which value $b_i$ takes as the prior is the same for each $w_j$.
-As such the integral can only be a function of at most $\tilde{x}_i$ and $\sigma_\theta^2$ as it is symmetric with respect to $b_i$ and all the various $w_j$ are integrated out as they are dummy variables.
-Therefore, denoting the integral by the (unknown) function $f(\tilde{x}_i, \sigma_\theta^2)$, we write $p(b| X)$ as follows:
-%
-\begin{align*}
-	p(b | X) &= \prod_{i=1}^{N} f(\tilde{x}_i, \sigma_\theta^2) = \textrm{const w.r.t } b = c
-\end{align*}
-%
-As this is a constant with respect to $b$ we conclude that $p(b | X)$ must be a uniform distribution.
-$\nicefrac{1}{c}$ is simply the size of the set of values that $b$ can take.
-We know $b_i \in \mathcal{B} = \{1, 2, \dots B\}$.
-Therefore, $b \in \mathcal{B}^N$ and $|\mathcal{B}^N| = |\mathcal{B}|^N = B^N= \nicefrac{1}{c}$.
-Putting this all together we show that:
-%
-\begin{equation}
-	p(b | X) = B^{-N}
-\end{equation}
-
-\subsection{Derivation of U gradient with respect to feature parameters}
-\label{appdx:gradu}
-The goal is to determine $\nabla U(\theta)$, the gradient of the negative log posterior with respect to the parameters. We repeat the form of $U(\theta)$ in equation \ref{eqn:U-form-appdx}.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form-appdx}
-\end{equation}
-%
-Where $y_{ij}$ is independent of $\theta$ and $a_{ij}$ is the output from the softmax layer, with form as given in equation \ref{eqn:a-ij}.
-%
-\begin{equation}
-	a_{ij} \coloneqq \phi_{j} (x_i; \theta) = \frac{\exp(w_j^T \tilde{x}_i)}{\sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i)}
-	\label{eqn:a-ij} 
-\end{equation}
-%
-We note that $\theta = \{w_k\}_{k=1}^B$, and as such we can write this in vector form $\theta = \left[w_1^T, w_2^T \dots w_B^T  \right]^T$.
-Therefore, $\nabla U(\theta) = \left[\nicefrac{\partial U}{\partial w_1}^T,\nicefrac{\partial U}{\partial w_2}^T \dots \nicefrac{\partial U}{\partial w_B}^T  \right]^T$; to compute $\nabla U(\theta)$ it suffices to find the form of $\nicefrac{\partial U}{\partial w_k}$ with respect to a general $k$.
-
-To this end, we must first find partial derivatives of $a_{ij}$ and $||\theta||$ with respect to $w_k$. Starting with $a_{ij}$:
-%
-\begin{align}
-	\frac{\partial a_{ij}}{\partial w_k} &= \frac
-	{\tilde{x}_i \exp(w_j^T \tilde{x}_i) \delta_{jk} \cdot \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) 
-		- 
-	\exp(w_j^T \tilde{x}_i) \cdot \tilde{x}_i \exp(w_k^T \tilde{x}_i)}
-	{\left( \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) \right)^2} \nonumber \\
-	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right) 
-\end{align}
-%
-Where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$. Now moving onto the derivative of $||\theta||^2$:
-%
-\begin{equation}
-	\frac{ \partial}{\partial w_k} ||\theta||^2 = \frac{\partial}{\partial w_k} \left( \sum_{b=1}^B ||w_b||^2 \right) = 2w_k
-\end{equation}
-%
-We are ready to put this all together, to find the partial derivative of $U(\theta)$ with respect to each $w_k$:
-\begin{align}
-	\frac{\partial U}{\partial w_k} &= 
-	\sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} 
-	\left( \frac{-\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
-	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
-	&=  - \left( \sum_{i=1}^{N} \tilde{x}_i \left( y_{ik} - a_{ik} \sum_{j=1}^{B} y_{ij} \right)
-	- \frac{w_k}{\sigma_\theta^2} \right) \nonumber \\
-	&= - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
-\end{align}
-%
-This is the required result.
-This form can be computed efficiently through matrix operations.
-The only property of $y_{ij}$ we have used in the derivation is the sum-to-one constraint $\sum_{j=1}^{B} y_{ij} = 1$ for all $i$.
-
-\subsection{Choosing the MALA step-size}
-\label{appdx:step-size}
-
-For sampling from the $\theta$-chain of the block membership generator parameters, we employed the Metropolis Adjusted Langevin Algorithm (MALA). At iteration $t$, the proposed sample is generated by:
-%
-\begin{equation}
-	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi
-\end{equation}
-%
-There are two competing objectives when choosing the step-size $h_t$.
-On the one hand, we want the step-size to be large so that we arrive at a high density region quickly.
-However, too large a step-size will lead to a lower acceptance ratio and thus inefficient sampling.
-A solution to this problem would be to slowly decrease the step-size with $t$ - often called simulated annealing.
-Therefore, we still have a short burn-in time but will not bounce around the mode for large $t$.
-As well as the trivial constraint for $h_t$ to be strictly positive and we introduce two further constraints as outlined by \citet{Bayesian-SGLD}:
-%
-\begin{equation}
-	\sum_{t=1}^{\infty} h_t = \infty \qquad \textrm{and} \qquad
-	\sum_{t=1}^{\infty} h_t^2 < \infty
-	\label{eqn:h-constraints}
-\end{equation}
-%
-The first constraint ensures that we have cover sufficient distance to arrive at any arbitrary point in our domain, no matter the starting point.
-The second constraint ensures that once we converge to the mode rather than simply bouncing around it.
-\citet{Bayesian-SGLD} propose the following form for a polynomially decaying step-size which we adopt:
-%
-\begin{equation}
-	h_t = a(b + t)^{-\gamma}
-\end{equation}
-%
-Where $a, b, \gamma$ are hyper-parameters to be chosen.
-We require $a,b > 0$ and $\gamma \in (0.5, 1]$ to satisfy equation \ref{eqn:h-constraints}.
-We find empirically that $a=250, b=1000, \gamma=0.8$ yield good burn-in times for our MALA implementations.
-
-\end{document}

### 2021-05-16 nips/inference.tex
Split into multiple tex files

--- 2021-05-14 nips/submission.tex
@@
-\documentclass{article}
-
-% if you need to pass options to natbib, use, e.g.:
-%     \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading neurips_2021
-
-% ready for submission
-\PassOptionsToPackage{numbers, compress}{natbib}
-\usepackage[final]{neurips_2021}
-
-% to compile a preprint version, e.g., for submission to arXiv, add add the
-% [preprint] option:
-%     \usepackage[preprint]{neurips_2021}
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-%\usepackage[final]{neurips_2021}
-
-% to avoid loading the natbib package, add option nonatbib:
-%    \usepackage[nonatbib]{neurips_2021}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{xcolor}         % colors
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-\usepackage{amsmath}
-\usepackage{mathtools}
-\usepackage{bbm}
-\usepackage{amssymb}
-\usepackage{algorithmic}
-\usepackage{algorithm}
-
-%bibliography
-\bibliographystyle{plainnat}
-
-\title{Inferring community characteristics in labelled networks}
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to break the
-% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
-% authors names on the first line, and the last on the second line, try using
-% \AND instead of \And before the third author name.
-
-\author{
-  Lawrence Tray \\
-  Department of Engineering \\
-  University of Cambridge \\
-  \texttt{lpt30@cam.ac.uk} \\
-  % examples of more authors
-  \And
-  Ioannis Kontoyiannis \\
-  Department of Mathematics \\
-  University of Cambridge \\
-  \texttt{ik355@cam.ac.uk}
-}
-
-%custom commands
-\newcommand{\Xcal}{\mathcal{X}}
-\newcommand{\Bcal}{\mathcal{B}}
-\newcommand{\Integers}{\mathbb{Z}}
-\newcommand{\one}{\mathbbm{1}}
-\newcommand{\Gaussian}{\mathcal{N}}
-\newcommand{\indep}{\perp \!\!\! \perp}
-\newcommand{\specialchoose}{\genfrac{\{}{\}}{0pt}{}}
-\newcommand{\Expect}{\mathbb{E}}
-
-%envs
-\newtheorem{definition}{Definition}[section]
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[theorem]
-\newtheorem{lemma}[theorem]{Lemma}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-We present a novel generative model for describing vertex-labelled undirected graphs.
-This is a two-layered model.
-First a probabilistic mapping from vertex labels to latent block memberships and from these block memberships we generate a graph according to the micro-canonical stochastic block model.
-
-With this framework we are able to efficiently sample the parameters of the vertex label to block membership mapping.
-This allows us to automatically determine which features have the largest impact on graphical structure.
-\end{abstract}
-
-\section{Introduction}
-
-There is a wealth of networks in the world.
-Some examples of this graphical data are social networks, website hyperlinks and academic collaboration with more being produced each second.
-It is clear we need tools to analyse this increasingly omni-present form of data.
-
-A somewhat surprising property of real-world graphs is that they exhibit strong community structure.
-In other words, each node will often belong to a cluster of densely connected nodes.
-This property is often exploited by graph compression algorithms and there is high interest in recovering the communities from the observed graph.
-
-A common subset of graphical data is the labelled network.
-This is a graph where we have information about the properties of each node.
-We shall refer to these node properties as features.
-One of the most common questions we can ask of this dataset is what features have the largest impact on the structure of the graph.
-For example, when analysing an academic collaboration graph, one may wish to ask what impact gender has on the structure of the network.
-Nevertheless, this analysis is often vulnerable to confounding variables.
-While gender very well may impact the structure of the graph, there is often a better explanatory variable for the structure.
-
-There is space to bridge the gap between these two approaches.
-Rather than determining whether a feature impacts the graphical structure directly, we can use the community concept as a stepping stone in our analysis.
-Therefore, we extract which features have the largest impact on overall graphical structure.
-This can be thought of as a form of dimensionality-reduction where we only keep the node features that are useful in predicting community memberships.
-
-\section{Preliminaries}
-
-This section defines some preliminary concepts required for the subsequent analysis.
-We first need a model for community-like structure in a graphical network.
-For this we adopt the stochastic block model (SBM) - widely used across academia.
-The premise is that each node in the graph belongs to a unique community called a block.
-The probability that two nodes are connected depends only on the block memberships of each.
-Graphs drawn from this ensemble exhibit community structure.
-Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}.
-A paraphrased definition is given below.
-
-\begin{definition}[Microcanonical NDC-SBM]
-	\label{defn:microcan-ndc-sbm}
-	Let $N \in \Integers^{+}$ denote the number of vertices in an undirected graph.
-The block memberships are encoded by a vector $b$ of length $N$ where each entry $b_i \in \{1, 2 \dots B\}$.
-$B \in \Integers^{+}$ is the number of non-empty blocks.
-Let $e$ be a $B \times B$ matrix of edge counts between blocks ($e_{rs}$ is number of edges from block $r$ onto block $s$ - or twice that number if $r=s$).
-For undirected graphs $e$ is symmetric.
-For a non-degree-corrected stochastic block model (NDC-SBM), we say that the graph $A$ is generated as follows:
-	%
-	\begin{equation}
-		A \sim \textrm{NDC-SBM}_{\textrm{MC}} (b, e)
-	\end{equation}
-	%
-	Where edges are placed uniformly at random but respecting the constraint imposed by $e$ and $b$.
-The additional parameters $N$ and $B$ are omitted as they are inferred from the shapes of $b$ and $e$.
-If we interpret $A$ as an adjacency matrix, then this constraint can be written formally as: $e_{rs} = \sum_{i,j} A_{ij} \one \{b_i = r\} \one \{b_j = s\}$.
-\end{definition}
-
-Nevertheless, this formulation does not accept high degree variability within blocks as is typical of real-world data.
-Indeed, the NDC-SBM favours a partition into high-degree and low-degree nodes rather than clusters of inter-connected nodes.
-We therefore introduce the degree-corrected SBM (DC-SBM) to circumvent these issues.
-
-\begin{definition}[Microcanonical DC-SBM]
-	\label{defn:microcan-dc-sbm}
-	This is much like the NDC-SBM but has an additional parameter $k$ which is an $N$-length vector encoding the degree sequence ($k_i$ is the degree of vertex $i$). Therefore, we write:
-	%
-	\begin{equation}
-		A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, e, k)
-	\end{equation}
-	%
-	Once again, edges are placed uniformly at random but respecting the constraints imposed by the parameters.
-The DC-SBM has the additional constraint that $k_i = \sum_{j} A_{ij}$.
-In what follows, we will always assume the degree-corrected model unless otherwise specified.
-
-\end{definition}
-
-\section{Latent block generative model}
-
-In this section we propose a novel generative model for modelling labelled networks.
-We restrict our analysis to labelled, undirected graphs with $N$ nodes.
-We define the vector $x_i \in \Xcal^D$ as the feature vector for the $i$'th vertex.
-Each vertex has $D$ total features and we assume all entries take values from the same set $\Xcal$.
-For the majority of datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$.
-The feature vectors $\{x_i\}_{i=1}^{N}$ are subsumed into the $N \times D$ matrix $X$.
-
-The proposed generative model (which we call the latent block generative model - LBGM) is given in figure \ref{fig:lbgm}.
-We start, with the feature matrix X and generate a vector of block memberships $b$.
-The parameters of this step are encapsulated by $\theta$.
-Each feature vector $x_i$ is treated independently and used to generate the block membership $b_i$.
-Each entry $b_i \in \{1, 2 \dots B\}$ where $B$ is the number of blocks and a constant in our model.
-We choose a single softmax layer to model $p(b_i | x_i, \theta)$.
-More complex models are possible but then deriving meaning from inferred parameter distributions is more difficult.
-Summarising, we write $p(b | X, \theta)$ as follows:
-%
-\begin{equation}
-	p(b| X, \theta) = \prod_{i=1}^{N} p(b_i | x_i, \theta) = \prod_{i=1}^{N} \phi_{b_i} (x_i; \theta)
-	= \prod_{i=1}^{N} \frac{\exp\left(w_{b_i}^T \tilde{x}_i\right)}{\sum_{k=1}^{B} \exp \left( w_k^T \tilde{x}_i\right)}
-\end{equation}
-%
-Where $\tilde{x} \coloneqq \left[x_1, x_2, \dots x_D, 1\right]^T$ is an augmented version of $x$ that allows for a bias term.
-The parameters for this stage are denoted $\theta$ and consist of the all $B$ weight vectors $\theta = \{w_k\}_{k=1}^{B}$.
-Each $w_k$ has dimension $D+1$.
-We could instead write the parameters $\theta$ as a $B \times (D+1)$ matrix of weights $W$; this form has use computationally as then $Z_i = W \tilde{x}_i$ is the input to the softmax activation function.
-
-\begin{figure}[!h]
-	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=7mm},
-		squarednode/.style={rectangle, draw=black, minimum size=7mm}
-		]
-		% nodes
-		\node[roundnode] (X) at (0, 0) {$X$};
-		\node[squarednode] (b) at (3, 0) {$b$};
-		\node[roundnode] (A) at (6, 0) {$A$};
-
-		% arrows
-		\draw[->] (X.east) -- node[above] {$\theta$} (b.west);
-		\draw[->] (b.east) -- node[above] {$\psi$}(A.west);
-	\end{tikzpicture}
-	\caption{Latent block generative model}
-	\label{fig:lbgm}
-\end{figure}
-
-With the block memberships $b$ generated, we then draw the graph $A$ from the microcanical DC-SBM (\ref{defn:microcan-dc-sbm}) with additional parameters $\psi = \{e, k\}$.
-In a slight abuse of notation we denote the inter-block edge count matrix with $e = \psi_e$ and the degree sequence $k = \psi_k$ to make explicit that these parameters are contained in $\psi$.
-%
-\begin{equation}
-	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k)
-\end{equation}
-
-\subsection{Prior selection}
-
-Before performing any inference, we must specify priors on $\theta$ and $\psi$.
-For $\theta$ it seems sensible to choose a Gaussian prior, with zero mean and variance matrix $\sigma^2_\theta I$ such that each element of $\theta$ is independent and distributed like $\sim \Gaussian(0, \sigma_\theta^2)$.
-In vector form, the prior for $\theta$ is therefore:
-%
-\begin{equation}
-	p(\theta) = \Gaussian \left( \theta ; 0, \sigma_\theta^2 I \right)
-	\label{eqn:theta-prior}
-\end{equation}
-%
-We will see that this form of prior is equivalent to a regularisation term in neural network training that penalises extreme weight magnitudes.
-For $\sigma_\theta^2 \rightarrow \infty$ this becomes an uninformative uniform prior.
-
-In our model, the block memberships vector $b$ is an intermediate latent variable and so we are not free to choose a prior for it.
-Nevertheless, as far as inference on the right-hand-side of figure \ref{fig:lbgm}, we regard $p(b | X)$ as a pseudo-prior on $b$.
-We can show (appendix \ref{appdx:b|x}) that our choice of prior for $p(\theta)$ in equation \ref{eqn:theta-prior} leads to a uniform $p(b | X)$ in equation \ref{eqn:b-pseudo-prior}.
-%
-\begin{equation}
-	p(b | X) = \int p(b | X, \theta) p(\theta) d\theta = B^{-N}
-	\label{eqn:b-pseudo-prior}
-\end{equation}
-%
-This is a uniform distribution that only depends on the number of blocks $B$.
-This simplifies the problem nicely as evaluating $p(b | X)$ does not require knowing the exact value $X$ takes.
-\citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for the additional microcanonical SBM parameters $\psi$ which we adopt.
-Peixoto's idea is to write the joint prior on $(b, e, k)$ as a product of conditionals $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$.
-For our purposes we must insert a conditioning on $X$, to form our pseudo-prior for $b$ and $\psi$, to give equation \ref{eqn:joint-pseudo-prior}.
-%
-\begin{equation}
-	p(b, \psi | X) = p(b | X) p(\psi | b, X) = p(b | X) p(\psi | b)
-	\label{eqn:joint-pseudo-prior}
-\end{equation}
-%
-Where the simplification is made apparent by noting $(\psi \indep X) | b$. We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference.
-%
-\begin{equation}
-	p(\psi | b) = p(e | b) p(k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
-	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right]
-\end{equation}
-%
-Where $\specialchoose{n}{m}$ is shorthand for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of distinct histograms with $n$ bins under the constraint they sum to $m$.
-$E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges in the graph.
-Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform with respect to $e$.
-The variable $\eta_j^r$ is introduced to denote the number of vertices in block $r$ that have degree $j$.
-Formally, $\eta_j^r \coloneqq \sum_{i} \one\left\{b_i = r \right\} \one \left\{k_i = j \right\}$.
-Furthermore, $q(m, n)$ is the number of different histograms with at most $n$ non-zero bins that sum to $m$.
-$q(m, n)$ is related to but different from $\specialchoose{n}{m}$.
-Lastly, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ is the number of vertices assigned to block $r$.
-Importantly, we have computable forms for $p(\theta)$ and $p(b, \psi | X)$ which will be useful for performing inference.
-
 \section{Inference}
@@
 The exact, from of the proposal distribution and accept-reject step is explored throughly by \citet{Peixoto-MCMC}.
-There is a widely available library, \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, implementing this algorithm.
+There is a widely used library for Python made available under LGPL called \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, which implements this algorithm.
 The only modification we make is in the block membership prior $p(b)$ which we replace with $p(b|X)=B^{-N}$ which is a uniform distribution and so cancels out in the MH accept-reject step.
@@
 
-\section{Experiments}
+\subsection{Choosing the number of blocks}
 
-We apply the developed methods to a variety of datasets.
+For the purposes of our model (the LBGM), the number of blocks $B$ is a constant which must be specified by the data scientist.
+We could however, allow our choice of $B$ to be influenced by the observed data.
+This places us in the domain of empirical Bayes, which must be negotiated carefully.
+Prior beliefs must be determined a priori else they are not prior.
+However, as the number of blocks only specifies the coarseness of the analysis, it is fine to allow it to vary.
+Indeed, \citet{peixoto-determine-B} shows that for a fixed average degree the maximum number of detectable blocks scales as $O(\sqrt{N})$ where $N$ is the number of vertices.
 
-\begin{itemize}
-	\item \textbf{College football teams} \cite{Evans_2010_football} - network of American college football teams and their interactions. Vertex labels are the division a team belongs to.
-	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.
-	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier} - network of the author's Facebook friends list.
-Each vertex has been manually labelled with a variety of features describing their relationship to the author.
-	\item \textbf{Twitch users} \cite{twitch} - a network of user-user friendships on the streaming service Twitch.
-Vertex labels are extracted according to video-games played, location and streaming habits.
-This dataset is also broken down into disjoint networks according to language.
-We only consider the English users with is a subnet with $N=7126$ vertices and $E=35324$ edges).
-\end{itemize}
-
-\section{Extensions}
-
-\section{Conclusion}
-
-\begin{ack}
-Use unnumbered first level headings for the acknowledgments. All acknowledgments
-go at the end of the paper before the list of references. Moreover, you are required to declare
-funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
-More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2021/PaperInformation/FundingDisclosure}.
-
-Do {\bf not} include this section in the anonymized submission, only in the final paper.
-You can use the \texttt{ack} environment provided in the style file to automatically hide this section in the anonymized submission.
-\end{ack}
-
-%references
-\bibliography{sources.bib}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Checklist}
-
-%%% BEGIN INSTRUCTIONS %%%
-The checklist follows the references.  Please
-read the checklist guidelines carefully for information on how to answer these
-questions.  For each question, change the default \answerTODO{} to \answerYes{},
-\answerNo{}, or \answerNA{}.  You are strongly encouraged to include a {\bf
-justification to your answer}, either by referencing the appropriate section of
-your paper or providing a brief inline description.  For example:
-\begin{itemize}
-  \item Did you include the license to the code and datasets? \answerYes{See Section gen-inst.}
-  \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
-  \item Did you include the license to the code and datasets? \answerNA{}
-\end{itemize}
-Please do not modify the questions and only use the provided macros for your
-answers.  Note that the Checklist section does not count towards the page
-limit.  In your paper, please delete this instructions block and only keep the
-Checklist section heading above along with the questions/answers below.
-%%% END INSTRUCTIONS %%%
-
-\begin{enumerate}
-
-\item For all authors...
-\begin{enumerate}
-  \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
-    \answerTODO{}
-  \item Did you describe the limitations of your work?
-    \answerTODO{}
-  \item Did you discuss any potential negative societal impacts of your work?
-    \answerTODO{}
-  \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you are including theoretical results...
-\begin{enumerate}
-  \item Did you state the full set of assumptions of all theoretical results?
-    \answerTODO{}
-	\item Did you include complete proofs of all theoretical results?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you ran experiments...
-\begin{enumerate}
-  \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
-    \answerTODO{}
-  \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
-    \answerTODO{}
-	\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
-    \answerTODO{}
-	\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
-\begin{enumerate}
-  \item If your work uses existing assets, did you cite the creators?
-    \answerTODO{}
-  \item Did you mention the license of the assets?
-    \answerTODO{}
-  \item Did you include any new assets either in the supplemental material or as a URL?
-    \answerTODO{}
-  \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
-    \answerTODO{}
-  \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you used crowdsourcing or conducted research with human subjects...
-\begin{enumerate}
-  \item Did you include the full text of instructions given to participants and screenshots, if applicable?
-    \answerTODO{}
-  \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
-    \answerTODO{}
-  \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
-    \answerTODO{}
-\end{enumerate}
-
-\end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\appendix
-
-\section{Appendix}
-
-\subsection{Derivation of conditional block distribution given feature matrix}
-\label{appdx:b|x}
-
-We wish to determine the form of $p(b| X)$. This can be done by integrating over the joint probability with respect to $\theta$.
-%
-\begin{align*}
-	p(b | X) &= \int p(b , \theta| X, \theta) d\theta = \int p(b | X, \theta) p(\theta | X) d\theta \\
-	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i=1}^{N} \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
-	&= \prod_{i=1}^{N} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j=1}^{B} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k=1}^{B} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}
-\end{align*}
-%
-We note that $b_i \in {1, 2, \dots B}$ and so the integral's value is unchanged with respect to $b_i$.
-The integrand has the same form no matter which value $b_i$ takes as the prior is the same for each $w_j$.
-As such the integral can only be a function of at most $\tilde{x}_i$ and $\sigma_\theta^2$ as it is symmetric with respect to $b_i$ and all the various $w_j$ are integrated out as they are dummy variables.
-Therefore, denoting the integral by the (unknown) function $f(\tilde{x}_i, \sigma_\theta^2)$, we write $p(b| X)$ as follows:
-%
-\begin{align*}
-	p(b | X) &= \prod_{i=1}^{N} f(\tilde{x}_i, \sigma_\theta^2) = \textrm{const w.r.t } b = c
-\end{align*}
-%
-As this is a constant with respect to $b$ we conclude that $p(b | X)$ must be a uniform distribution.
-$\nicefrac{1}{c}$ is simply the size of the set of values that $b$ can take.
-We know $b_i \in \mathcal{B} = \{1, 2, \dots B\}$.
-Therefore, $b \in \mathcal{B}^N$ and $|\mathcal{B}^N| = |\mathcal{B}|^N = B^N= \nicefrac{1}{c}$.
-Putting this all together we show that:
-%
-\begin{equation}
-	p(b | X) = B^{-N}
-\end{equation}
-
-\subsection{Derivation of U gradient with respect to feature parameters}
-\label{appdx:gradu}
-The goal is to determine $\nabla U(\theta)$, the gradient of the negative log posterior with respect to the parameters. We repeat the form of $U(\theta)$ in equation \ref{eqn:U-form-appdx}.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form-appdx}
-\end{equation}
-%
-Where $y_{ij}$ is independent of $\theta$ and $a_{ij}$ is the output from the softmax layer, with form as given in equation \ref{eqn:a-ij}.
-%
-\begin{equation}
-	a_{ij} \coloneqq \phi_{j} (x_i; \theta) = \frac{\exp(w_j^T \tilde{x}_i)}{\sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i)}
-	\label{eqn:a-ij} 
-\end{equation}
-%
-We note that $\theta = \{w_k\}_{k=1}^B$, and as such we can write this in vector form $\theta = \left[w_1^T, w_2^T \dots w_B^T  \right]^T$.
-Therefore, $\nabla U(\theta) = \left[\nicefrac{\partial U}{\partial w_1}^T,\nicefrac{\partial U}{\partial w_2}^T \dots \nicefrac{\partial U}{\partial w_B}^T  \right]^T$; to compute $\nabla U(\theta)$ it suffices to find the form of $\nicefrac{\partial U}{\partial w_k}$ with respect to a general $k$.
-
-To this end, we must first find partial derivatives of $a_{ij}$ and $||\theta||$ with respect to $w_k$. Starting with $a_{ij}$:
-%
-\begin{align}
-	\frac{\partial a_{ij}}{\partial w_k} &= \frac
-	{\tilde{x}_i \exp(w_j^T \tilde{x}_i) \delta_{jk} \cdot \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) 
-		- 
-	\exp(w_j^T \tilde{x}_i) \cdot \tilde{x}_i \exp(w_k^T \tilde{x}_i)}
-	{\left( \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) \right)^2} \nonumber \\
-	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right) 
-\end{align}
-%
-Where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$. Now moving onto the derivative of $||\theta||^2$:
-%
-\begin{equation}
-	\frac{ \partial}{\partial w_k} ||\theta||^2 = \frac{\partial}{\partial w_k} \left( \sum_{b=1}^B ||w_b||^2 \right) = 2w_k
-\end{equation}
-%
-We are ready to put this all together, to find the partial derivative of $U(\theta)$ with respect to each $w_k$:
-\begin{align}
-	\frac{\partial U}{\partial w_k} &= 
-	\sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} 
-	\left( \frac{-\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
-	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
-	&=  - \left( \sum_{i=1}^{N} \tilde{x}_i \left( y_{ik} - a_{ik} \sum_{j=1}^{B} y_{ij} \right)
-	- \frac{w_k}{\sigma_\theta^2} \right) \nonumber \\
-	&= - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
-\end{align}
-%
-This is the required result.
-This form can be computed efficiently through matrix operations.
-The only property of $y_{ij}$ we have used in the derivation is the sum-to-one constraint $\sum_{j=1}^{B} y_{ij} = 1$ for all $i$.
-
-\subsection{Choosing the MALA step-size}
-\label{appdx:step-size}
-
-For sampling from the $\theta$-chain of the block membership generator parameters, we employed the Metropolis Adjusted Langevin Algorithm (MALA). At iteration $t$, the proposed sample is generated by:
-%
-\begin{equation}
-	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi
-\end{equation}
-%
-There are two competing objectives when choosing the step-size $h_t$.
-On the one hand, we want the step-size to be large so that we arrive at a high density region quickly.
-However, too large a step-size will lead to a lower acceptance ratio and thus inefficient sampling.
-A solution to this problem would be to slowly decrease the step-size with $t$ - often called simulated annealing.
-Therefore, we still have a short burn-in time but will not bounce around the mode for large $t$.
-As well as the trivial constraint for $h_t$ to be strictly positive and we introduce two further constraints as outlined by \citet{Bayesian-SGLD}:
-%
-\begin{equation}
-	\sum_{t=1}^{\infty} h_t = \infty \qquad \textrm{and} \qquad
-	\sum_{t=1}^{\infty} h_t^2 < \infty
-	\label{eqn:h-constraints}
-\end{equation}
-%
-The first constraint ensures that we have cover sufficient distance to arrive at any arbitrary point in our domain, no matter the starting point.
-The second constraint ensures that once we converge to the mode rather than simply bouncing around it.
-\citet{Bayesian-SGLD} propose the following form for a polynomially decaying step-size which we adopt:
-%
-\begin{equation}
-	h_t = a(b + t)^{-\gamma}
-\end{equation}
-%
-Where $a, b, \gamma$ are hyper-parameters to be chosen.
-We require $a,b > 0$ and $\gamma \in (0.5, 1]$ to satisfy equation \ref{eqn:h-constraints}.
-We find empirically that $a=250, b=1000, \gamma=0.8$ yield good burn-in times for our MALA implementations.
-
-\end{document}
+If we allow $B$ to vary in the $b$-chain (i.e.
+new blocks can be created and we permit empty blocks) then it can be run  until a minimum description length (MDL) solution is reached.
+We take the number of non-empty blocks at the MDL to be our fixed block number $B$ for subsequent analysis.
+Indeed, it is prudent to start our $b$-chain at this MDL solution as then the burn-in time is greatly reduced.

### 2021-05-16 nips/introduction.tex
Split into multiple tex files

--- 2021-05-14 nips/submission.tex
@@
-\documentclass{article}
-
-% if you need to pass options to natbib, use, e.g.:
-%     \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading neurips_2021
-
-% ready for submission
-\PassOptionsToPackage{numbers, compress}{natbib}
-\usepackage[final]{neurips_2021}
-
-% to compile a preprint version, e.g., for submission to arXiv, add add the
-% [preprint] option:
-%     \usepackage[preprint]{neurips_2021}
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-%\usepackage[final]{neurips_2021}
-
-% to avoid loading the natbib package, add option nonatbib:
-%    \usepackage[nonatbib]{neurips_2021}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{xcolor}         % colors
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-\usepackage{amsmath}
-\usepackage{mathtools}
-\usepackage{bbm}
-\usepackage{amssymb}
-\usepackage{algorithmic}
-\usepackage{algorithm}
-
-%bibliography
-\bibliographystyle{plainnat}
-
-\title{Inferring community characteristics in labelled networks}
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to break the
-% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
-% authors names on the first line, and the last on the second line, try using
-% \AND instead of \And before the third author name.
-
-\author{
-  Lawrence Tray \\
-  Department of Engineering \\
-  University of Cambridge \\
-  \texttt{lpt30@cam.ac.uk} \\
-  % examples of more authors
-  \And
-  Ioannis Kontoyiannis \\
-  Department of Mathematics \\
-  University of Cambridge \\
-  \texttt{ik355@cam.ac.uk}
-}
-
-%custom commands
-\newcommand{\Xcal}{\mathcal{X}}
-\newcommand{\Bcal}{\mathcal{B}}
-\newcommand{\Integers}{\mathbb{Z}}
-\newcommand{\one}{\mathbbm{1}}
-\newcommand{\Gaussian}{\mathcal{N}}
-\newcommand{\indep}{\perp \!\!\! \perp}
-\newcommand{\specialchoose}{\genfrac{\{}{\}}{0pt}{}}
-\newcommand{\Expect}{\mathbb{E}}
-
-%envs
-\newtheorem{definition}{Definition}[section]
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[theorem]
-\newtheorem{lemma}[theorem]{Lemma}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-We present a novel generative model for describing vertex-labelled undirected graphs.
-This is a two-layered model.
-First a probabilistic mapping from vertex labels to latent block memberships and from these block memberships we generate a graph according to the micro-canonical stochastic block model.
-
-With this framework we are able to efficiently sample the parameters of the vertex label to block membership mapping.
-This allows us to automatically determine which features have the largest impact on graphical structure.
-\end{abstract}
-
 \section{Introduction}
@@
 This can be thought of as a form of dimensionality-reduction where we only keep the node features that are useful in predicting community memberships.
-
-\section{Preliminaries}
-
-This section defines some preliminary concepts required for the subsequent analysis.
-We first need a model for community-like structure in a graphical network.
-For this we adopt the stochastic block model (SBM) - widely used across academia.
-The premise is that each node in the graph belongs to a unique community called a block.
-The probability that two nodes are connected depends only on the block memberships of each.
-Graphs drawn from this ensemble exhibit community structure.
-Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}.
-A paraphrased definition is given below.
-
-\begin{definition}[Microcanonical NDC-SBM]
-	\label{defn:microcan-ndc-sbm}
-	Let $N \in \Integers^{+}$ denote the number of vertices in an undirected graph.
-The block memberships are encoded by a vector $b$ of length $N$ where each entry $b_i \in \{1, 2 \dots B\}$.
-$B \in \Integers^{+}$ is the number of non-empty blocks.
-Let $e$ be a $B \times B$ matrix of edge counts between blocks ($e_{rs}$ is number of edges from block $r$ onto block $s$ - or twice that number if $r=s$).
-For undirected graphs $e$ is symmetric.
-For a non-degree-corrected stochastic block model (NDC-SBM), we say that the graph $A$ is generated as follows:
-	%
-	\begin{equation}
-		A \sim \textrm{NDC-SBM}_{\textrm{MC}} (b, e)
-	\end{equation}
-	%
-	Where edges are placed uniformly at random but respecting the constraint imposed by $e$ and $b$.
-The additional parameters $N$ and $B$ are omitted as they are inferred from the shapes of $b$ and $e$.
-If we interpret $A$ as an adjacency matrix, then this constraint can be written formally as: $e_{rs} = \sum_{i,j} A_{ij} \one \{b_i = r\} \one \{b_j = s\}$.
-\end{definition}
-
-Nevertheless, this formulation does not accept high degree variability within blocks as is typical of real-world data.
-Indeed, the NDC-SBM favours a partition into high-degree and low-degree nodes rather than clusters of inter-connected nodes.
-We therefore introduce the degree-corrected SBM (DC-SBM) to circumvent these issues.
-
-\begin{definition}[Microcanonical DC-SBM]
-	\label{defn:microcan-dc-sbm}
-	This is much like the NDC-SBM but has an additional parameter $k$ which is an $N$-length vector encoding the degree sequence ($k_i$ is the degree of vertex $i$). Therefore, we write:
-	%
-	\begin{equation}
-		A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, e, k)
-	\end{equation}
-	%
-	Once again, edges are placed uniformly at random but respecting the constraints imposed by the parameters.
-The DC-SBM has the additional constraint that $k_i = \sum_{j} A_{ij}$.
-In what follows, we will always assume the degree-corrected model unless otherwise specified.
-
-\end{definition}
-
-\section{Latent block generative model}
-
-In this section we propose a novel generative model for modelling labelled networks.
-We restrict our analysis to labelled, undirected graphs with $N$ nodes.
-We define the vector $x_i \in \Xcal^D$ as the feature vector for the $i$'th vertex.
-Each vertex has $D$ total features and we assume all entries take values from the same set $\Xcal$.
-For the majority of datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$.
-The feature vectors $\{x_i\}_{i=1}^{N}$ are subsumed into the $N \times D$ matrix $X$.
-
-The proposed generative model (which we call the latent block generative model - LBGM) is given in figure \ref{fig:lbgm}.
-We start, with the feature matrix X and generate a vector of block memberships $b$.
-The parameters of this step are encapsulated by $\theta$.
-Each feature vector $x_i$ is treated independently and used to generate the block membership $b_i$.
-Each entry $b_i \in \{1, 2 \dots B\}$ where $B$ is the number of blocks and a constant in our model.
-We choose a single softmax layer to model $p(b_i | x_i, \theta)$.
-More complex models are possible but then deriving meaning from inferred parameter distributions is more difficult.
-Summarising, we write $p(b | X, \theta)$ as follows:
-%
-\begin{equation}
-	p(b| X, \theta) = \prod_{i=1}^{N} p(b_i | x_i, \theta) = \prod_{i=1}^{N} \phi_{b_i} (x_i; \theta)
-	= \prod_{i=1}^{N} \frac{\exp\left(w_{b_i}^T \tilde{x}_i\right)}{\sum_{k=1}^{B} \exp \left( w_k^T \tilde{x}_i\right)}
-\end{equation}
-%
-Where $\tilde{x} \coloneqq \left[x_1, x_2, \dots x_D, 1\right]^T$ is an augmented version of $x$ that allows for a bias term.
-The parameters for this stage are denoted $\theta$ and consist of the all $B$ weight vectors $\theta = \{w_k\}_{k=1}^{B}$.
-Each $w_k$ has dimension $D+1$.
-We could instead write the parameters $\theta$ as a $B \times (D+1)$ matrix of weights $W$; this form has use computationally as then $Z_i = W \tilde{x}_i$ is the input to the softmax activation function.
-
-\begin{figure}[!h]
-	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=7mm},
-		squarednode/.style={rectangle, draw=black, minimum size=7mm}
-		]
-		% nodes
-		\node[roundnode] (X) at (0, 0) {$X$};
-		\node[squarednode] (b) at (3, 0) {$b$};
-		\node[roundnode] (A) at (6, 0) {$A$};
-
-		% arrows
-		\draw[->] (X.east) -- node[above] {$\theta$} (b.west);
-		\draw[->] (b.east) -- node[above] {$\psi$}(A.west);
-	\end{tikzpicture}
-	\caption{Latent block generative model}
-	\label{fig:lbgm}
-\end{figure}
-
-With the block memberships $b$ generated, we then draw the graph $A$ from the microcanical DC-SBM (\ref{defn:microcan-dc-sbm}) with additional parameters $\psi = \{e, k\}$.
-In a slight abuse of notation we denote the inter-block edge count matrix with $e = \psi_e$ and the degree sequence $k = \psi_k$ to make explicit that these parameters are contained in $\psi$.
-%
-\begin{equation}
-	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k)
-\end{equation}
-
-\subsection{Prior selection}
-
-Before performing any inference, we must specify priors on $\theta$ and $\psi$.
-For $\theta$ it seems sensible to choose a Gaussian prior, with zero mean and variance matrix $\sigma^2_\theta I$ such that each element of $\theta$ is independent and distributed like $\sim \Gaussian(0, \sigma_\theta^2)$.
-In vector form, the prior for $\theta$ is therefore:
-%
-\begin{equation}
-	p(\theta) = \Gaussian \left( \theta ; 0, \sigma_\theta^2 I \right)
-	\label{eqn:theta-prior}
-\end{equation}
-%
-We will see that this form of prior is equivalent to a regularisation term in neural network training that penalises extreme weight magnitudes.
-For $\sigma_\theta^2 \rightarrow \infty$ this becomes an uninformative uniform prior.
-
-In our model, the block memberships vector $b$ is an intermediate latent variable and so we are not free to choose a prior for it.
-Nevertheless, as far as inference on the right-hand-side of figure \ref{fig:lbgm}, we regard $p(b | X)$ as a pseudo-prior on $b$.
-We can show (appendix \ref{appdx:b|x}) that our choice of prior for $p(\theta)$ in equation \ref{eqn:theta-prior} leads to a uniform $p(b | X)$ in equation \ref{eqn:b-pseudo-prior}.
-%
-\begin{equation}
-	p(b | X) = \int p(b | X, \theta) p(\theta) d\theta = B^{-N}
-	\label{eqn:b-pseudo-prior}
-\end{equation}
-%
-This is a uniform distribution that only depends on the number of blocks $B$.
-This simplifies the problem nicely as evaluating $p(b | X)$ does not require knowing the exact value $X$ takes.
-\citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for the additional microcanonical SBM parameters $\psi$ which we adopt.
-Peixoto's idea is to write the joint prior on $(b, e, k)$ as a product of conditionals $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$.
-For our purposes we must insert a conditioning on $X$, to form our pseudo-prior for $b$ and $\psi$, to give equation \ref{eqn:joint-pseudo-prior}.
-%
-\begin{equation}
-	p(b, \psi | X) = p(b | X) p(\psi | b, X) = p(b | X) p(\psi | b)
-	\label{eqn:joint-pseudo-prior}
-\end{equation}
-%
-Where the simplification is made apparent by noting $(\psi \indep X) | b$. We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference.
-%
-\begin{equation}
-	p(\psi | b) = p(e | b) p(k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
-	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right]
-\end{equation}
-%
-Where $\specialchoose{n}{m}$ is shorthand for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of distinct histograms with $n$ bins under the constraint they sum to $m$.
-$E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges in the graph.
-Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform with respect to $e$.
-The variable $\eta_j^r$ is introduced to denote the number of vertices in block $r$ that have degree $j$.
-Formally, $\eta_j^r \coloneqq \sum_{i} \one\left\{b_i = r \right\} \one \left\{k_i = j \right\}$.
-Furthermore, $q(m, n)$ is the number of different histograms with at most $n$ non-zero bins that sum to $m$.
-$q(m, n)$ is related to but different from $\specialchoose{n}{m}$.
-Lastly, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ is the number of vertices assigned to block $r$.
-Importantly, we have computable forms for $p(\theta)$ and $p(b, \psi | X)$ which will be useful for performing inference.
-
-\section{Inference}
-
-Now that we have defined the LBGM, we wish to leverage it to perform inference.
-Suppose we are presented with a vertex-labelled graph $(A, X)$; the goal is to draw samples for $\theta$ according to the posterior given the observed graph (equation \ref{eqn:theta-target}).
-%
-\begin{equation}
-	\label{eqn:theta-target}
-	\theta^{(t)} \sim p(\theta | A, X)
-\end{equation}
-%
-These samples allow us to approximate the posterior distribution for $\theta$ as well as compute a predictive distribution $p(b^* | x^*, A, X) = \int p(b^* | x^*, \theta) p(\theta | X, A) d\theta \approx \frac{1}{T} \sum_{t=1}^{T} p(b^* | x^*, \theta^{(t)})$.
-However, generating these samples is not easily done in practice.
-
-We instead propose an iterative approach.
-First drawing samples $b^{(t)}$ from the block membership posterior (equation \ref{eqn:b-samples}).
-We then use each $b^{(i)}$ to draw samples for $\theta$ as in equation \ref{eqn:theta-samples}.
-%
-\begin{align}
-	b^{(i)} &\sim p \Big( b | A, X \Big)  \label{eqn:b-samples}\\
-	\theta^{(i)} &\sim p\Big(\theta | X, b^{(i)} \Big) \label{eqn:theta-samples}
-\end{align}
-%
-Both of these sampling steps implemented with a Markov Chain through the Metropolis-Hastings algorithm \cite{hastings-alg}.
-We just need to define a proposal distribution $q(x, x')$ for proposing a move $x \rightarrow x'$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(\cdot)$, point-wise.
-The proposed move is then accepted with probability $\alpha$ (equation \ref{eqn:mh-accept}) else it is rejected and we stay at $x$.
-%
-\begin{equation}
-	\alpha = \min \left( \frac{\pi(x') q(x', x)}{\pi(x) q(x, x')} , 1 \right)
-	\label{eqn:mh-accept}
-\end{equation}
-%
-This accept-reject step ensures the resulting Markov Chain is in detailed balance with the target distribution $\pi(\cdot)$.
-What we propose in equations \ref{eqn:b-samples} and \ref{eqn:theta-samples} is therefore implemented through a 2-level Markov chain.
-The resulting samples for $\theta^{(i)}$ are unbiased in the sense that the expectation of their distribution is the posterior we are targeting in equation \ref{eqn:theta-target}.
-%
-\begin{align*}
-	\Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] &= \sum_{b \in \mathcal{B}^N} p(\theta | X, b) p(b | A, X) \\
-	&= \sum_{b \in \mathcal{B}^N} p(\theta, b | A, X) \\
-	&= p(\theta | A, X) \\
-\end{align*}
-%
-This is an example of a pseudo-marginal approach \cite{pseudo-marginal}.
-Indeed, the unbiased result is sufficient to prove that for sufficient samples, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{t})\right] = p(\theta
-\ A, X)$ which is exactly the distribution we were targetting (equation \ref{eqn:theta-target}).
-
-The reason we split the Markov chain into two stages is because the summation over all latent states $b \in \mathcal{B}^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in \Bcal^N} p(A | b) P(b | X, \theta)$ is intractable $O(B^N)$.
-%
-\begin{figure}[!h]
-	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=12mm},
-		squarednode/.style={rectangle, draw=black, minimum size=12mm}
-		]
-		% nodes
-		\node[roundnode] (b0) at (0, 2) {$b^{(t)}$};
-		\node[roundnode] (b1) at (4, 2) {$b'$};
-		\node[roundnode] (b2) at (8, 2) {$b^{(t+1)}$};
-		\node[roundnode] (t0) at (0, 0) {$\theta^{(t)}$};
-		\node[roundnode] (t1) at (4, 0) {$\theta'$};
-		\node[roundnode] (t2) at (8, 0) {$\theta^{(t+1)}$};
-
-		% arrows
-		\draw[->] (b0) to node[above] {$\sim q_b(b^{(t)}, b' | A)$} (b1);
-		\draw[->] (b1) to node[above] {$\alpha_b (b, b' | A)$} (b2);
-		\draw[->] (b0) [out=30, in=150] to node[above] {$1-\alpha_b$} (b2);
-
-		\draw[->] (t0) to node[above] {$\sim q_\theta(\theta, \theta' | X, b^{(t)})$} (t1);
-		\draw[->] (t1) to node[above] {$\alpha_\theta (\theta, \theta' | X, b^{(t)})$} (t2);
-		\draw[->] (t0) [out=-30, in=-150] to node[below] {$1-\alpha_\theta$} (t2);
-
-		\draw[dashed, ->] (b0) to (2, 0.8);
-		\draw[dashed, ->] (b0) to (5.5, 0.8);
-
-	\end{tikzpicture}
-	\caption{Sampling sequence}
-	\label{fig:samp-sequence}
-\end{figure}
-%
-Figure \ref{fig:samp-sequence} shows an overview of the proposed method.
-We have introduced subscripts and conditionings to make explicit what parameters each step utilises.
-In an important simplification, we note that $p(b| X) = B^{-N}$ which does not depend on the exact value of X.
-Therefore, we do not need to know the value of $X$ to perform the sampling on $b$.
-Conversely, for the $\theta^{(t)}$ samples, we use $b^{(t)}$ but not $A$ as $(\theta \indep A) | b$.
-
-\subsection{Sampling block memberships}
-
-\citet{Peixoto-MCMC} proposes a Monte Carlo method which we will base our approach on. It relies on writing the posterior in the following form:
-%
-\begin{equation}
-	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) = \pi_b(b)
-\end{equation}
-%
-Now $\pi_b(\cdot)$ is the un-normalised density we wish to sample from.
-In other words, we wish to construct a Markov chain that has $\pi_b(\cdot)$ as its invariant distribution.
-We can break $\pi_b$ down as follows:
-%
-\begin{align*}
-	\pi_b(b) &= p(b|X) \sum_{\psi} \nolimits p(A , \psi | b, X) \\
-	&= p(b|X) p(A, \psi^* | b, X) \\
-	&= p(A | b, \psi^*) \cdot p(\psi^* | b) \cdot p(b | X)
-\end{align*}
-%
-Since we are using the microcanonical SBM formulation, there is only one value of $\psi$ that is compatible with the given $(A, b)$ pair.
-We denote this value $\psi^* = \{k^*, e^*\}$.
-Specifically, $k^*_i = \sum_j A_{ij} $ and $e^*_{rs} = \sum_{i, j} A_{ij} \one \{b_i=r\} \one\{b_j=s\}$.
-Therefore, the summation over all $\psi$ reduces to just the single $\psi^*$ term.
-We also define the microcanonical entropy of the configuration as.
-%
-\begin{equation}
-	S(b) = - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big)
-\end{equation}
-%
-This entropy can be thought of as the description length of the graph because it is the sum of the information required to represent the graph given the parameters and the amount of information required to store the parameters (given the feature matrix $X$).
-The exact, from of the proposal distribution and accept-reject step is explored throughly by \citet{Peixoto-MCMC}.
-There is a widely available library, \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, implementing this algorithm.
-The only modification we make is in the block membership prior $p(b)$ which we replace with $p(b|X)=B^{-N}$ which is a uniform distribution and so cancels out in the MH accept-reject step.
-
-The end result of this is that we can generate a set of block membership samples $\{b^{(t)}\}_{t=1}^{T}$ with each $b^{(t)} \sim p(b | A, X)$.
-Each of these samples can then be used for the $\theta$-chain.
-
-\subsection{Sampling feature-to-block classifier parameters}
-
-The invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. We write this as follows:
-%
-\begin{align}
-	p(\theta | X, b) &\propto p(b | X, \theta) p(\theta) = \pi_\theta (\theta) \propto  \exp \left( - U(\theta) \right) \\
-	\therefore U(\theta) &= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}
-\end{align}
-%
-Where we have introduced $U(\theta)$ equal to the negative log posterior.
-This $U(\theta)$ term makes subsequent analysis more concise.
-Each of the constituent terms of $U(\theta)$ is easily computed (equation \ref{eqn:u-form}).
-To simplify notation, we define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
-%
-\begin{equation}
-	\log p(b | X, \theta) = \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
-	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta ||^2
-	\label{eqn:U-constituent-terms}
-\end{equation}
-%
-This means that if we ignore constant terms we can write $U(\theta)$ as in equation \ref{eqn:U-form}.
-We see that the prior effectively introduces a regularisation term.
-Note that $||\theta||^2 = \sum_{i} \theta_{i}^2 = \sum_{j=1}^{B} ||w_j||^2$ is the Euclidean norm of the vector of parameters $\theta$.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form}
-\end{equation}
-%
-$U(\theta)$ in equation \ref{eqn:U-form} appears a typical objective function to be minimised for neural network training.
-The first term is the cross-entropy between actual and predicted labels.
-The second term - introduced by the prior - brings a form of regularisation, preventing over-fitting.
-In traditional applications we only seek the value of $\theta$ that minimises the objective function $U(\theta)$, which in our case would yield the maximum a posteriori (MAP) estimate.
-This is often done through some kind of gradient descent as $\nabla U$ is easily computable (equation \ref{eqn:U-derivative}).
-
-However, our goal is not to find the MAP estimate but to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
-As discussed earlier, given the invariant $\pi_\theta(\cdot)$, it is sufficient to specify a proposal distribution and then apply a MH accept-reject step to ensure detailed balance of the Markov Chain.
-Nevertheless, we can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
-We therefore adopt the Metropolis Adjusted Langevin Algorithm (MALA) - first proposed by \citet{mala-tweedie} - which leverages just that.
-Given the current sample $\theta$, we propose a new sample $\theta'$ according to equation \ref{eqn:theta-update}
-%
-\begin{equation}
-	\theta' = \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi
-	\label{eqn:theta-update}
-\end{equation}
-%
-Where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter - which may vary with the sample index.
-Indeed without the injected noise term, this is equivalent to gradient descent.
-We require the noise term to fully explore the parameter space.
-As such the proposal distribution, is a simple multivariate Gaussian which can be easily evaluated.
-%
-\begin{equation}
-	q_\theta(\theta, \theta') = \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right)
-\end{equation}
-%
-The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}). By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as:
-%
-\begin{equation}
-	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
-	\label{eqn:U-derivative}
-\end{equation}
-%
-After a proposed move is generated, in typical Metropolis-Hastings fashion we accept the move with probability $\alpha_\theta$.
-%
-\begin{equation}
-	\alpha_\theta(\theta, \theta') = \min \left( 
-	\exp \left( U(\theta) - U(\theta')\right)
-	\frac{ 
-		q_\theta(\theta', \theta)
-	}{
-		q_\theta(\theta, \theta')
-	} 
-	, 1 \right)
-\end{equation}
-%
-This fully specifies, the sampling procedure to generate $\{\theta^{(t)}\}_{t=1}^T$. Bear in mind that each $\theta^{(t)}$ update step uses its corresponding $b^{(t)}$ block membership sample.
-
-\subsection{Sampling sequence}
-
-So far, each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample.
-This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance.
-This may lead to longer burn-in and autocorrelation times of the resulting Markov Chain.
-The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends on $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
-We would rather deal with the expectation of each $y_{ij}^{(t)}$:
-%
-\begin{equation}
-	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one(b_{i}^{(t)} = j) \right]
-	= p(b_i = j | A, X)
-\end{equation}
-%
-We obtain an unbiased estimate for this quantity as simply the empirical distribution of the block membership samples $\left\{ b^{(t)} \right\}_{t=1}^T$.
-%
-\begin{equation}
-	\hat{y}_{ij} \coloneqq \frac{1}{T} \sum_{t=1}^{T} y_{ij}^{(t)} = \frac{1}{T} \sum_{t=1}^{T} \one\{b_i^{(t)} = j\}
-\end{equation}
-%
-We therefore, choose to feed each $\theta^{(t)}$ update step the same $\hat{y}_{ij}$ for all $t$ rather than the corresponding $y^{(t)}_{ij}$.
-This means we no longer need to run the $b$ and $\theta$ Markov chains concurrently.
-Instead, we run the $b$-chain to completion and use it to generate $\hat{y}_{ij}$ for $i \in \{1 \dots N\}$ and $j \in \{1 \dots B\}$.
-This is an estimate of $p(b | A, X)$ that we use for every iteration of the $\theta$ Markov chain.
-This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$; wed refer to these as $T_b$ and $T_\theta$ henceforth.
-Furthermore, this changeover reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{Block membership sample generation} % give the algorithm a caption
-	\label{alg:b-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\FOR{$t \in \{0, 1 \dots T_b - 1\}$}
-			\STATE $b' \gets \sim q_b(b^{(t)}, b' | A)$
-			\STATE $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
-			\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-			\IF{$\log \eta < \log \alpha_b$}
-				\STATE $b^{(t+1)} \gets b'$
-			\ELSE
-				\STATE $b^{(t+1)} \gets b^{(t)}$
-			\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{LBGM parameter pseudo-marginal inference} % give the algorithm a caption
-	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\STATE $\mathcal{S} \gets \textrm{range}(\textrm{burn-in}, T_b, \textrm{thinning})$
-		\STATE $\hat{Y}_{ij} \gets \frac{1}{\mathcal{|S|}} \sum_{t \in \mathcal{S}} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
-
-		\item[]
-
-		\FOR{$t \in \{0, 1 \dots T_\theta - 1\}$}
-		\STATE $\xi \gets \sim \Gaussian(0, I)$
-		\STATE $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
-		\STATE $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
-		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-		\IF{$\log \eta < \log \alpha_\theta$}
-		\STATE $\theta^{(t+1)} \gets \theta'$
-		\ELSE
-		\STATE $\theta^{(t+1)} \gets \theta^{(t)}$
-		\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\subsection{Complexity Analysis}
-
-\section{Experiments}
-
-We apply the developed methods to a variety of datasets.
-
-\begin{itemize}
-	\item \textbf{College football teams} \cite{Evans_2010_football} - network of American college football teams and their interactions. Vertex labels are the division a team belongs to.
-	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.
-	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier} - network of the author's Facebook friends list.
-Each vertex has been manually labelled with a variety of features describing their relationship to the author.
-	\item \textbf{Twitch users} \cite{twitch} - a network of user-user friendships on the streaming service Twitch.
-Vertex labels are extracted according to video-games played, location and streaming habits.
-This dataset is also broken down into disjoint networks according to language.
-We only consider the English users with is a subnet with $N=7126$ vertices and $E=35324$ edges).
-\end{itemize}
-
-\section{Extensions}
-
-\section{Conclusion}
-
-\begin{ack}
-Use unnumbered first level headings for the acknowledgments. All acknowledgments
-go at the end of the paper before the list of references. Moreover, you are required to declare
-funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
-More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2021/PaperInformation/FundingDisclosure}.
-
-Do {\bf not} include this section in the anonymized submission, only in the final paper.
-You can use the \texttt{ack} environment provided in the style file to automatically hide this section in the anonymized submission.
-\end{ack}
-
-%references
-\bibliography{sources.bib}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Checklist}
-
-%%% BEGIN INSTRUCTIONS %%%
-The checklist follows the references.  Please
-read the checklist guidelines carefully for information on how to answer these
-questions.  For each question, change the default \answerTODO{} to \answerYes{},
-\answerNo{}, or \answerNA{}.  You are strongly encouraged to include a {\bf
-justification to your answer}, either by referencing the appropriate section of
-your paper or providing a brief inline description.  For example:
-\begin{itemize}
-  \item Did you include the license to the code and datasets? \answerYes{See Section gen-inst.}
-  \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
-  \item Did you include the license to the code and datasets? \answerNA{}
-\end{itemize}
-Please do not modify the questions and only use the provided macros for your
-answers.  Note that the Checklist section does not count towards the page
-limit.  In your paper, please delete this instructions block and only keep the
-Checklist section heading above along with the questions/answers below.
-%%% END INSTRUCTIONS %%%
-
-\begin{enumerate}
-
-\item For all authors...
-\begin{enumerate}
-  \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
-    \answerTODO{}
-  \item Did you describe the limitations of your work?
-    \answerTODO{}
-  \item Did you discuss any potential negative societal impacts of your work?
-    \answerTODO{}
-  \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you are including theoretical results...
-\begin{enumerate}
-  \item Did you state the full set of assumptions of all theoretical results?
-    \answerTODO{}
-	\item Did you include complete proofs of all theoretical results?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you ran experiments...
-\begin{enumerate}
-  \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
-    \answerTODO{}
-  \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
-    \answerTODO{}
-	\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
-    \answerTODO{}
-	\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
-\begin{enumerate}
-  \item If your work uses existing assets, did you cite the creators?
-    \answerTODO{}
-  \item Did you mention the license of the assets?
-    \answerTODO{}
-  \item Did you include any new assets either in the supplemental material or as a URL?
-    \answerTODO{}
-  \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
-    \answerTODO{}
-  \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you used crowdsourcing or conducted research with human subjects...
-\begin{enumerate}
-  \item Did you include the full text of instructions given to participants and screenshots, if applicable?
-    \answerTODO{}
-  \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
-    \answerTODO{}
-  \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
-    \answerTODO{}
-\end{enumerate}
-
-\end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\appendix
-
-\section{Appendix}
-
-\subsection{Derivation of conditional block distribution given feature matrix}
-\label{appdx:b|x}
-
-We wish to determine the form of $p(b| X)$. This can be done by integrating over the joint probability with respect to $\theta$.
-%
-\begin{align*}
-	p(b | X) &= \int p(b , \theta| X, \theta) d\theta = \int p(b | X, \theta) p(\theta | X) d\theta \\
-	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i=1}^{N} \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
-	&= \prod_{i=1}^{N} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j=1}^{B} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k=1}^{B} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}
-\end{align*}
-%
-We note that $b_i \in {1, 2, \dots B}$ and so the integral's value is unchanged with respect to $b_i$.
-The integrand has the same form no matter which value $b_i$ takes as the prior is the same for each $w_j$.
-As such the integral can only be a function of at most $\tilde{x}_i$ and $\sigma_\theta^2$ as it is symmetric with respect to $b_i$ and all the various $w_j$ are integrated out as they are dummy variables.
-Therefore, denoting the integral by the (unknown) function $f(\tilde{x}_i, \sigma_\theta^2)$, we write $p(b| X)$ as follows:
-%
-\begin{align*}
-	p(b | X) &= \prod_{i=1}^{N} f(\tilde{x}_i, \sigma_\theta^2) = \textrm{const w.r.t } b = c
-\end{align*}
-%
-As this is a constant with respect to $b$ we conclude that $p(b | X)$ must be a uniform distribution.
-$\nicefrac{1}{c}$ is simply the size of the set of values that $b$ can take.
-We know $b_i \in \mathcal{B} = \{1, 2, \dots B\}$.
-Therefore, $b \in \mathcal{B}^N$ and $|\mathcal{B}^N| = |\mathcal{B}|^N = B^N= \nicefrac{1}{c}$.
-Putting this all together we show that:
-%
-\begin{equation}
-	p(b | X) = B^{-N}
-\end{equation}
-
-\subsection{Derivation of U gradient with respect to feature parameters}
-\label{appdx:gradu}
-The goal is to determine $\nabla U(\theta)$, the gradient of the negative log posterior with respect to the parameters. We repeat the form of $U(\theta)$ in equation \ref{eqn:U-form-appdx}.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form-appdx}
-\end{equation}
-%
-Where $y_{ij}$ is independent of $\theta$ and $a_{ij}$ is the output from the softmax layer, with form as given in equation \ref{eqn:a-ij}.
-%
-\begin{equation}
-	a_{ij} \coloneqq \phi_{j} (x_i; \theta) = \frac{\exp(w_j^T \tilde{x}_i)}{\sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i)}
-	\label{eqn:a-ij} 
-\end{equation}
-%
-We note that $\theta = \{w_k\}_{k=1}^B$, and as such we can write this in vector form $\theta = \left[w_1^T, w_2^T \dots w_B^T  \right]^T$.
-Therefore, $\nabla U(\theta) = \left[\nicefrac{\partial U}{\partial w_1}^T,\nicefrac{\partial U}{\partial w_2}^T \dots \nicefrac{\partial U}{\partial w_B}^T  \right]^T$; to compute $\nabla U(\theta)$ it suffices to find the form of $\nicefrac{\partial U}{\partial w_k}$ with respect to a general $k$.
-
-To this end, we must first find partial derivatives of $a_{ij}$ and $||\theta||$ with respect to $w_k$. Starting with $a_{ij}$:
-%
-\begin{align}
-	\frac{\partial a_{ij}}{\partial w_k} &= \frac
-	{\tilde{x}_i \exp(w_j^T \tilde{x}_i) \delta_{jk} \cdot \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) 
-		- 
-	\exp(w_j^T \tilde{x}_i) \cdot \tilde{x}_i \exp(w_k^T \tilde{x}_i)}
-	{\left( \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) \right)^2} \nonumber \\
-	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right) 
-\end{align}
-%
-Where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$. Now moving onto the derivative of $||\theta||^2$:
-%
-\begin{equation}
-	\frac{ \partial}{\partial w_k} ||\theta||^2 = \frac{\partial}{\partial w_k} \left( \sum_{b=1}^B ||w_b||^2 \right) = 2w_k
-\end{equation}
-%
-We are ready to put this all together, to find the partial derivative of $U(\theta)$ with respect to each $w_k$:
-\begin{align}
-	\frac{\partial U}{\partial w_k} &= 
-	\sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} 
-	\left( \frac{-\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
-	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
-	&=  - \left( \sum_{i=1}^{N} \tilde{x}_i \left( y_{ik} - a_{ik} \sum_{j=1}^{B} y_{ij} \right)
-	- \frac{w_k}{\sigma_\theta^2} \right) \nonumber \\
-	&= - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
-\end{align}
-%
-This is the required result.
-This form can be computed efficiently through matrix operations.
-The only property of $y_{ij}$ we have used in the derivation is the sum-to-one constraint $\sum_{j=1}^{B} y_{ij} = 1$ for all $i$.
-
-\subsection{Choosing the MALA step-size}
-\label{appdx:step-size}
-
-For sampling from the $\theta$-chain of the block membership generator parameters, we employed the Metropolis Adjusted Langevin Algorithm (MALA). At iteration $t$, the proposed sample is generated by:
-%
-\begin{equation}
-	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi
-\end{equation}
-%
-There are two competing objectives when choosing the step-size $h_t$.
-On the one hand, we want the step-size to be large so that we arrive at a high density region quickly.
-However, too large a step-size will lead to a lower acceptance ratio and thus inefficient sampling.
-A solution to this problem would be to slowly decrease the step-size with $t$ - often called simulated annealing.
-Therefore, we still have a short burn-in time but will not bounce around the mode for large $t$.
-As well as the trivial constraint for $h_t$ to be strictly positive and we introduce two further constraints as outlined by \citet{Bayesian-SGLD}:
-%
-\begin{equation}
-	\sum_{t=1}^{\infty} h_t = \infty \qquad \textrm{and} \qquad
-	\sum_{t=1}^{\infty} h_t^2 < \infty
-	\label{eqn:h-constraints}
-\end{equation}
-%
-The first constraint ensures that we have cover sufficient distance to arrive at any arbitrary point in our domain, no matter the starting point.
-The second constraint ensures that once we converge to the mode rather than simply bouncing around it.
-\citet{Bayesian-SGLD} propose the following form for a polynomially decaying step-size which we adopt:
-%
-\begin{equation}
-	h_t = a(b + t)^{-\gamma}
-\end{equation}
-%
-Where $a, b, \gamma$ are hyper-parameters to be chosen.
-We require $a,b > 0$ and $\gamma \in (0.5, 1]$ to satisfy equation \ref{eqn:h-constraints}.
-We find empirically that $a=250, b=1000, \gamma=0.8$ yield good burn-in times for our MALA implementations.
-
-\end{document}

### 2021-05-16 nips/lbgm.tex
Split into multiple tex files

--- 2021-05-14 nips/submission.tex
@@
-\documentclass{article}
-
-% if you need to pass options to natbib, use, e.g.:
-%     \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading neurips_2021
-
-% ready for submission
-\PassOptionsToPackage{numbers, compress}{natbib}
-\usepackage[final]{neurips_2021}
-
-% to compile a preprint version, e.g., for submission to arXiv, add add the
-% [preprint] option:
-%     \usepackage[preprint]{neurips_2021}
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-%\usepackage[final]{neurips_2021}
-
-% to avoid loading the natbib package, add option nonatbib:
-%    \usepackage[nonatbib]{neurips_2021}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{xcolor}         % colors
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-\usepackage{amsmath}
-\usepackage{mathtools}
-\usepackage{bbm}
-\usepackage{amssymb}
-\usepackage{algorithmic}
-\usepackage{algorithm}
-
-%bibliography
-\bibliographystyle{plainnat}
-
-\title{Inferring community characteristics in labelled networks}
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to break the
-% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
-% authors names on the first line, and the last on the second line, try using
-% \AND instead of \And before the third author name.
-
-\author{
-  Lawrence Tray \\
-  Department of Engineering \\
-  University of Cambridge \\
-  \texttt{lpt30@cam.ac.uk} \\
-  % examples of more authors
-  \And
-  Ioannis Kontoyiannis \\
-  Department of Mathematics \\
-  University of Cambridge \\
-  \texttt{ik355@cam.ac.uk}
-}
-
-%custom commands
-\newcommand{\Xcal}{\mathcal{X}}
-\newcommand{\Bcal}{\mathcal{B}}
-\newcommand{\Integers}{\mathbb{Z}}
-\newcommand{\one}{\mathbbm{1}}
-\newcommand{\Gaussian}{\mathcal{N}}
-\newcommand{\indep}{\perp \!\!\! \perp}
-\newcommand{\specialchoose}{\genfrac{\{}{\}}{0pt}{}}
-\newcommand{\Expect}{\mathbb{E}}
-
-%envs
-\newtheorem{definition}{Definition}[section]
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[theorem]
-\newtheorem{lemma}[theorem]{Lemma}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-We present a novel generative model for describing vertex-labelled undirected graphs.
-This is a two-layered model.
-First a probabilistic mapping from vertex labels to latent block memberships and from these block memberships we generate a graph according to the micro-canonical stochastic block model.
-
-With this framework we are able to efficiently sample the parameters of the vertex label to block membership mapping.
-This allows us to automatically determine which features have the largest impact on graphical structure.
-\end{abstract}
-
-\section{Introduction}
-
-There is a wealth of networks in the world.
-Some examples of this graphical data are social networks, website hyperlinks and academic collaboration with more being produced each second.
-It is clear we need tools to analyse this increasingly omni-present form of data.
-
-A somewhat surprising property of real-world graphs is that they exhibit strong community structure.
-In other words, each node will often belong to a cluster of densely connected nodes.
-This property is often exploited by graph compression algorithms and there is high interest in recovering the communities from the observed graph.
-
-A common subset of graphical data is the labelled network.
-This is a graph where we have information about the properties of each node.
-We shall refer to these node properties as features.
-One of the most common questions we can ask of this dataset is what features have the largest impact on the structure of the graph.
-For example, when analysing an academic collaboration graph, one may wish to ask what impact gender has on the structure of the network.
-Nevertheless, this analysis is often vulnerable to confounding variables.
-While gender very well may impact the structure of the graph, there is often a better explanatory variable for the structure.
-
-There is space to bridge the gap between these two approaches.
-Rather than determining whether a feature impacts the graphical structure directly, we can use the community concept as a stepping stone in our analysis.
-Therefore, we extract which features have the largest impact on overall graphical structure.
-This can be thought of as a form of dimensionality-reduction where we only keep the node features that are useful in predicting community memberships.
-
-\section{Preliminaries}
-
-This section defines some preliminary concepts required for the subsequent analysis.
-We first need a model for community-like structure in a graphical network.
-For this we adopt the stochastic block model (SBM) - widely used across academia.
-The premise is that each node in the graph belongs to a unique community called a block.
-The probability that two nodes are connected depends only on the block memberships of each.
-Graphs drawn from this ensemble exhibit community structure.
-Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}.
-A paraphrased definition is given below.
-
-\begin{definition}[Microcanonical NDC-SBM]
-	\label{defn:microcan-ndc-sbm}
-	Let $N \in \Integers^{+}$ denote the number of vertices in an undirected graph.
-The block memberships are encoded by a vector $b$ of length $N$ where each entry $b_i \in \{1, 2 \dots B\}$.
-$B \in \Integers^{+}$ is the number of non-empty blocks.
-Let $e$ be a $B \times B$ matrix of edge counts between blocks ($e_{rs}$ is number of edges from block $r$ onto block $s$ - or twice that number if $r=s$).
-For undirected graphs $e$ is symmetric.
-For a non-degree-corrected stochastic block model (NDC-SBM), we say that the graph $A$ is generated as follows:
-	%
-	\begin{equation}
-		A \sim \textrm{NDC-SBM}_{\textrm{MC}} (b, e)
-	\end{equation}
-	%
-	Where edges are placed uniformly at random but respecting the constraint imposed by $e$ and $b$.
-The additional parameters $N$ and $B$ are omitted as they are inferred from the shapes of $b$ and $e$.
-If we interpret $A$ as an adjacency matrix, then this constraint can be written formally as: $e_{rs} = \sum_{i,j} A_{ij} \one \{b_i = r\} \one \{b_j = s\}$.
-\end{definition}
-
-Nevertheless, this formulation does not accept high degree variability within blocks as is typical of real-world data.
-Indeed, the NDC-SBM favours a partition into high-degree and low-degree nodes rather than clusters of inter-connected nodes.
-We therefore introduce the degree-corrected SBM (DC-SBM) to circumvent these issues.
-
-\begin{definition}[Microcanonical DC-SBM]
-	\label{defn:microcan-dc-sbm}
-	This is much like the NDC-SBM but has an additional parameter $k$ which is an $N$-length vector encoding the degree sequence ($k_i$ is the degree of vertex $i$). Therefore, we write:
-	%
-	\begin{equation}
-		A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, e, k)
-	\end{equation}
-	%
-	Once again, edges are placed uniformly at random but respecting the constraints imposed by the parameters.
-The DC-SBM has the additional constraint that $k_i = \sum_{j} A_{ij}$.
-In what follows, we will always assume the degree-corrected model unless otherwise specified.
-
-\end{definition}
-
 \section{Latent block generative model}
@@
 Importantly, we have computable forms for $p(\theta)$ and $p(b, \psi | X)$ which will be useful for performing inference.
-
-\section{Inference}
-
-Now that we have defined the LBGM, we wish to leverage it to perform inference.
-Suppose we are presented with a vertex-labelled graph $(A, X)$; the goal is to draw samples for $\theta$ according to the posterior given the observed graph (equation \ref{eqn:theta-target}).
-%
-\begin{equation}
-	\label{eqn:theta-target}
-	\theta^{(t)} \sim p(\theta | A, X)
-\end{equation}
-%
-These samples allow us to approximate the posterior distribution for $\theta$ as well as compute a predictive distribution $p(b^* | x^*, A, X) = \int p(b^* | x^*, \theta) p(\theta | X, A) d\theta \approx \frac{1}{T} \sum_{t=1}^{T} p(b^* | x^*, \theta^{(t)})$.
-However, generating these samples is not easily done in practice.
-
-We instead propose an iterative approach.
-First drawing samples $b^{(t)}$ from the block membership posterior (equation \ref{eqn:b-samples}).
-We then use each $b^{(i)}$ to draw samples for $\theta$ as in equation \ref{eqn:theta-samples}.
-%
-\begin{align}
-	b^{(i)} &\sim p \Big( b | A, X \Big)  \label{eqn:b-samples}\\
-	\theta^{(i)} &\sim p\Big(\theta | X, b^{(i)} \Big) \label{eqn:theta-samples}
-\end{align}
-%
-Both of these sampling steps implemented with a Markov Chain through the Metropolis-Hastings algorithm \cite{hastings-alg}.
-We just need to define a proposal distribution $q(x, x')$ for proposing a move $x \rightarrow x'$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(\cdot)$, point-wise.
-The proposed move is then accepted with probability $\alpha$ (equation \ref{eqn:mh-accept}) else it is rejected and we stay at $x$.
-%
-\begin{equation}
-	\alpha = \min \left( \frac{\pi(x') q(x', x)}{\pi(x) q(x, x')} , 1 \right)
-	\label{eqn:mh-accept}
-\end{equation}
-%
-This accept-reject step ensures the resulting Markov Chain is in detailed balance with the target distribution $\pi(\cdot)$.
-What we propose in equations \ref{eqn:b-samples} and \ref{eqn:theta-samples} is therefore implemented through a 2-level Markov chain.
-The resulting samples for $\theta^{(i)}$ are unbiased in the sense that the expectation of their distribution is the posterior we are targeting in equation \ref{eqn:theta-target}.
-%
-\begin{align*}
-	\Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] &= \sum_{b \in \mathcal{B}^N} p(\theta | X, b) p(b | A, X) \\
-	&= \sum_{b \in \mathcal{B}^N} p(\theta, b | A, X) \\
-	&= p(\theta | A, X) \\
-\end{align*}
-%
-This is an example of a pseudo-marginal approach \cite{pseudo-marginal}.
-Indeed, the unbiased result is sufficient to prove that for sufficient samples, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{t})\right] = p(\theta
-\ A, X)$ which is exactly the distribution we were targetting (equation \ref{eqn:theta-target}).
-
-The reason we split the Markov chain into two stages is because the summation over all latent states $b \in \mathcal{B}^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in \Bcal^N} p(A | b) P(b | X, \theta)$ is intractable $O(B^N)$.
-%
-\begin{figure}[!h]
-	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=12mm},
-		squarednode/.style={rectangle, draw=black, minimum size=12mm}
-		]
-		% nodes
-		\node[roundnode] (b0) at (0, 2) {$b^{(t)}$};
-		\node[roundnode] (b1) at (4, 2) {$b'$};
-		\node[roundnode] (b2) at (8, 2) {$b^{(t+1)}$};
-		\node[roundnode] (t0) at (0, 0) {$\theta^{(t)}$};
-		\node[roundnode] (t1) at (4, 0) {$\theta'$};
-		\node[roundnode] (t2) at (8, 0) {$\theta^{(t+1)}$};
-
-		% arrows
-		\draw[->] (b0) to node[above] {$\sim q_b(b^{(t)}, b' | A)$} (b1);
-		\draw[->] (b1) to node[above] {$\alpha_b (b, b' | A)$} (b2);
-		\draw[->] (b0) [out=30, in=150] to node[above] {$1-\alpha_b$} (b2);
-
-		\draw[->] (t0) to node[above] {$\sim q_\theta(\theta, \theta' | X, b^{(t)})$} (t1);
-		\draw[->] (t1) to node[above] {$\alpha_\theta (\theta, \theta' | X, b^{(t)})$} (t2);
-		\draw[->] (t0) [out=-30, in=-150] to node[below] {$1-\alpha_\theta$} (t2);
-
-		\draw[dashed, ->] (b0) to (2, 0.8);
-		\draw[dashed, ->] (b0) to (5.5, 0.8);
-
-	\end{tikzpicture}
-	\caption{Sampling sequence}
-	\label{fig:samp-sequence}
-\end{figure}
-%
-Figure \ref{fig:samp-sequence} shows an overview of the proposed method.
-We have introduced subscripts and conditionings to make explicit what parameters each step utilises.
-In an important simplification, we note that $p(b| X) = B^{-N}$ which does not depend on the exact value of X.
-Therefore, we do not need to know the value of $X$ to perform the sampling on $b$.
-Conversely, for the $\theta^{(t)}$ samples, we use $b^{(t)}$ but not $A$ as $(\theta \indep A) | b$.
-
-\subsection{Sampling block memberships}
-
-\citet{Peixoto-MCMC} proposes a Monte Carlo method which we will base our approach on. It relies on writing the posterior in the following form:
-%
-\begin{equation}
-	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) = \pi_b(b)
-\end{equation}
-%
-Now $\pi_b(\cdot)$ is the un-normalised density we wish to sample from.
-In other words, we wish to construct a Markov chain that has $\pi_b(\cdot)$ as its invariant distribution.
-We can break $\pi_b$ down as follows:
-%
-\begin{align*}
-	\pi_b(b) &= p(b|X) \sum_{\psi} \nolimits p(A , \psi | b, X) \\
-	&= p(b|X) p(A, \psi^* | b, X) \\
-	&= p(A | b, \psi^*) \cdot p(\psi^* | b) \cdot p(b | X)
-\end{align*}
-%
-Since we are using the microcanonical SBM formulation, there is only one value of $\psi$ that is compatible with the given $(A, b)$ pair.
-We denote this value $\psi^* = \{k^*, e^*\}$.
-Specifically, $k^*_i = \sum_j A_{ij} $ and $e^*_{rs} = \sum_{i, j} A_{ij} \one \{b_i=r\} \one\{b_j=s\}$.
-Therefore, the summation over all $\psi$ reduces to just the single $\psi^*$ term.
-We also define the microcanonical entropy of the configuration as.
-%
-\begin{equation}
-	S(b) = - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big)
-\end{equation}
-%
-This entropy can be thought of as the description length of the graph because it is the sum of the information required to represent the graph given the parameters and the amount of information required to store the parameters (given the feature matrix $X$).
-The exact, from of the proposal distribution and accept-reject step is explored throughly by \citet{Peixoto-MCMC}.
-There is a widely available library, \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, implementing this algorithm.
-The only modification we make is in the block membership prior $p(b)$ which we replace with $p(b|X)=B^{-N}$ which is a uniform distribution and so cancels out in the MH accept-reject step.
-
-The end result of this is that we can generate a set of block membership samples $\{b^{(t)}\}_{t=1}^{T}$ with each $b^{(t)} \sim p(b | A, X)$.
-Each of these samples can then be used for the $\theta$-chain.
-
-\subsection{Sampling feature-to-block classifier parameters}
-
-The invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. We write this as follows:
-%
-\begin{align}
-	p(\theta | X, b) &\propto p(b | X, \theta) p(\theta) = \pi_\theta (\theta) \propto  \exp \left( - U(\theta) \right) \\
-	\therefore U(\theta) &= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}
-\end{align}
-%
-Where we have introduced $U(\theta)$ equal to the negative log posterior.
-This $U(\theta)$ term makes subsequent analysis more concise.
-Each of the constituent terms of $U(\theta)$ is easily computed (equation \ref{eqn:u-form}).
-To simplify notation, we define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
-%
-\begin{equation}
-	\log p(b | X, \theta) = \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
-	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta ||^2
-	\label{eqn:U-constituent-terms}
-\end{equation}
-%
-This means that if we ignore constant terms we can write $U(\theta)$ as in equation \ref{eqn:U-form}.
-We see that the prior effectively introduces a regularisation term.
-Note that $||\theta||^2 = \sum_{i} \theta_{i}^2 = \sum_{j=1}^{B} ||w_j||^2$ is the Euclidean norm of the vector of parameters $\theta$.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form}
-\end{equation}
-%
-$U(\theta)$ in equation \ref{eqn:U-form} appears a typical objective function to be minimised for neural network training.
-The first term is the cross-entropy between actual and predicted labels.
-The second term - introduced by the prior - brings a form of regularisation, preventing over-fitting.
-In traditional applications we only seek the value of $\theta$ that minimises the objective function $U(\theta)$, which in our case would yield the maximum a posteriori (MAP) estimate.
-This is often done through some kind of gradient descent as $\nabla U$ is easily computable (equation \ref{eqn:U-derivative}).
-
-However, our goal is not to find the MAP estimate but to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
-As discussed earlier, given the invariant $\pi_\theta(\cdot)$, it is sufficient to specify a proposal distribution and then apply a MH accept-reject step to ensure detailed balance of the Markov Chain.
-Nevertheless, we can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
-We therefore adopt the Metropolis Adjusted Langevin Algorithm (MALA) - first proposed by \citet{mala-tweedie} - which leverages just that.
-Given the current sample $\theta$, we propose a new sample $\theta'$ according to equation \ref{eqn:theta-update}
-%
-\begin{equation}
-	\theta' = \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi
-	\label{eqn:theta-update}
-\end{equation}
-%
-Where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter - which may vary with the sample index.
-Indeed without the injected noise term, this is equivalent to gradient descent.
-We require the noise term to fully explore the parameter space.
-As such the proposal distribution, is a simple multivariate Gaussian which can be easily evaluated.
-%
-\begin{equation}
-	q_\theta(\theta, \theta') = \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right)
-\end{equation}
-%
-The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}). By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as:
-%
-\begin{equation}
-	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
-	\label{eqn:U-derivative}
-\end{equation}
-%
-After a proposed move is generated, in typical Metropolis-Hastings fashion we accept the move with probability $\alpha_\theta$.
-%
-\begin{equation}
-	\alpha_\theta(\theta, \theta') = \min \left( 
-	\exp \left( U(\theta) - U(\theta')\right)
-	\frac{ 
-		q_\theta(\theta', \theta)
-	}{
-		q_\theta(\theta, \theta')
-	} 
-	, 1 \right)
-\end{equation}
-%
-This fully specifies, the sampling procedure to generate $\{\theta^{(t)}\}_{t=1}^T$. Bear in mind that each $\theta^{(t)}$ update step uses its corresponding $b^{(t)}$ block membership sample.
-
-\subsection{Sampling sequence}
-
-So far, each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample.
-This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance.
-This may lead to longer burn-in and autocorrelation times of the resulting Markov Chain.
-The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends on $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
-We would rather deal with the expectation of each $y_{ij}^{(t)}$:
-%
-\begin{equation}
-	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one(b_{i}^{(t)} = j) \right]
-	= p(b_i = j | A, X)
-\end{equation}
-%
-We obtain an unbiased estimate for this quantity as simply the empirical distribution of the block membership samples $\left\{ b^{(t)} \right\}_{t=1}^T$.
-%
-\begin{equation}
-	\hat{y}_{ij} \coloneqq \frac{1}{T} \sum_{t=1}^{T} y_{ij}^{(t)} = \frac{1}{T} \sum_{t=1}^{T} \one\{b_i^{(t)} = j\}
-\end{equation}
-%
-We therefore, choose to feed each $\theta^{(t)}$ update step the same $\hat{y}_{ij}$ for all $t$ rather than the corresponding $y^{(t)}_{ij}$.
-This means we no longer need to run the $b$ and $\theta$ Markov chains concurrently.
-Instead, we run the $b$-chain to completion and use it to generate $\hat{y}_{ij}$ for $i \in \{1 \dots N\}$ and $j \in \{1 \dots B\}$.
-This is an estimate of $p(b | A, X)$ that we use for every iteration of the $\theta$ Markov chain.
-This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$; wed refer to these as $T_b$ and $T_\theta$ henceforth.
-Furthermore, this changeover reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{Block membership sample generation} % give the algorithm a caption
-	\label{alg:b-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\FOR{$t \in \{0, 1 \dots T_b - 1\}$}
-			\STATE $b' \gets \sim q_b(b^{(t)}, b' | A)$
-			\STATE $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
-			\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-			\IF{$\log \eta < \log \alpha_b$}
-				\STATE $b^{(t+1)} \gets b'$
-			\ELSE
-				\STATE $b^{(t+1)} \gets b^{(t)}$
-			\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{LBGM parameter pseudo-marginal inference} % give the algorithm a caption
-	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\STATE $\mathcal{S} \gets \textrm{range}(\textrm{burn-in}, T_b, \textrm{thinning})$
-		\STATE $\hat{Y}_{ij} \gets \frac{1}{\mathcal{|S|}} \sum_{t \in \mathcal{S}} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
-
-		\item[]
-
-		\FOR{$t \in \{0, 1 \dots T_\theta - 1\}$}
-		\STATE $\xi \gets \sim \Gaussian(0, I)$
-		\STATE $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
-		\STATE $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
-		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-		\IF{$\log \eta < \log \alpha_\theta$}
-		\STATE $\theta^{(t+1)} \gets \theta'$
-		\ELSE
-		\STATE $\theta^{(t+1)} \gets \theta^{(t)}$
-		\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\subsection{Complexity Analysis}
-
-\section{Experiments}
-
-We apply the developed methods to a variety of datasets.
-
-\begin{itemize}
-	\item \textbf{College football teams} \cite{Evans_2010_football} - network of American college football teams and their interactions. Vertex labels are the division a team belongs to.
-	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.
-	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier} - network of the author's Facebook friends list.
-Each vertex has been manually labelled with a variety of features describing their relationship to the author.
-	\item \textbf{Twitch users} \cite{twitch} - a network of user-user friendships on the streaming service Twitch.
-Vertex labels are extracted according to video-games played, location and streaming habits.
-This dataset is also broken down into disjoint networks according to language.
-We only consider the English users with is a subnet with $N=7126$ vertices and $E=35324$ edges).
-\end{itemize}
-
-\section{Extensions}
-
-\section{Conclusion}
-
-\begin{ack}
-Use unnumbered first level headings for the acknowledgments. All acknowledgments
-go at the end of the paper before the list of references. Moreover, you are required to declare
-funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
-More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2021/PaperInformation/FundingDisclosure}.
-
-Do {\bf not} include this section in the anonymized submission, only in the final paper.
-You can use the \texttt{ack} environment provided in the style file to automatically hide this section in the anonymized submission.
-\end{ack}
-
-%references
-\bibliography{sources.bib}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Checklist}
-
-%%% BEGIN INSTRUCTIONS %%%
-The checklist follows the references.  Please
-read the checklist guidelines carefully for information on how to answer these
-questions.  For each question, change the default \answerTODO{} to \answerYes{},
-\answerNo{}, or \answerNA{}.  You are strongly encouraged to include a {\bf
-justification to your answer}, either by referencing the appropriate section of
-your paper or providing a brief inline description.  For example:
-\begin{itemize}
-  \item Did you include the license to the code and datasets? \answerYes{See Section gen-inst.}
-  \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
-  \item Did you include the license to the code and datasets? \answerNA{}
-\end{itemize}
-Please do not modify the questions and only use the provided macros for your
-answers.  Note that the Checklist section does not count towards the page
-limit.  In your paper, please delete this instructions block and only keep the
-Checklist section heading above along with the questions/answers below.
-%%% END INSTRUCTIONS %%%
-
-\begin{enumerate}
-
-\item For all authors...
-\begin{enumerate}
-  \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
-    \answerTODO{}
-  \item Did you describe the limitations of your work?
-    \answerTODO{}
-  \item Did you discuss any potential negative societal impacts of your work?
-    \answerTODO{}
-  \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you are including theoretical results...
-\begin{enumerate}
-  \item Did you state the full set of assumptions of all theoretical results?
-    \answerTODO{}
-	\item Did you include complete proofs of all theoretical results?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you ran experiments...
-\begin{enumerate}
-  \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
-    \answerTODO{}
-  \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
-    \answerTODO{}
-	\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
-    \answerTODO{}
-	\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
-\begin{enumerate}
-  \item If your work uses existing assets, did you cite the creators?
-    \answerTODO{}
-  \item Did you mention the license of the assets?
-    \answerTODO{}
-  \item Did you include any new assets either in the supplemental material or as a URL?
-    \answerTODO{}
-  \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
-    \answerTODO{}
-  \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you used crowdsourcing or conducted research with human subjects...
-\begin{enumerate}
-  \item Did you include the full text of instructions given to participants and screenshots, if applicable?
-    \answerTODO{}
-  \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
-    \answerTODO{}
-  \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
-    \answerTODO{}
-\end{enumerate}
-
-\end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\appendix
-
-\section{Appendix}
-
-\subsection{Derivation of conditional block distribution given feature matrix}
-\label{appdx:b|x}
-
-We wish to determine the form of $p(b| X)$. This can be done by integrating over the joint probability with respect to $\theta$.
-%
-\begin{align*}
-	p(b | X) &= \int p(b , \theta| X, \theta) d\theta = \int p(b | X, \theta) p(\theta | X) d\theta \\
-	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i=1}^{N} \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
-	&= \prod_{i=1}^{N} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j=1}^{B} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k=1}^{B} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}
-\end{align*}
-%
-We note that $b_i \in {1, 2, \dots B}$ and so the integral's value is unchanged with respect to $b_i$.
-The integrand has the same form no matter which value $b_i$ takes as the prior is the same for each $w_j$.
-As such the integral can only be a function of at most $\tilde{x}_i$ and $\sigma_\theta^2$ as it is symmetric with respect to $b_i$ and all the various $w_j$ are integrated out as they are dummy variables.
-Therefore, denoting the integral by the (unknown) function $f(\tilde{x}_i, \sigma_\theta^2)$, we write $p(b| X)$ as follows:
-%
-\begin{align*}
-	p(b | X) &= \prod_{i=1}^{N} f(\tilde{x}_i, \sigma_\theta^2) = \textrm{const w.r.t } b = c
-\end{align*}
-%
-As this is a constant with respect to $b$ we conclude that $p(b | X)$ must be a uniform distribution.
-$\nicefrac{1}{c}$ is simply the size of the set of values that $b$ can take.
-We know $b_i \in \mathcal{B} = \{1, 2, \dots B\}$.
-Therefore, $b \in \mathcal{B}^N$ and $|\mathcal{B}^N| = |\mathcal{B}|^N = B^N= \nicefrac{1}{c}$.
-Putting this all together we show that:
-%
-\begin{equation}
-	p(b | X) = B^{-N}
-\end{equation}
-
-\subsection{Derivation of U gradient with respect to feature parameters}
-\label{appdx:gradu}
-The goal is to determine $\nabla U(\theta)$, the gradient of the negative log posterior with respect to the parameters. We repeat the form of $U(\theta)$ in equation \ref{eqn:U-form-appdx}.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form-appdx}
-\end{equation}
-%
-Where $y_{ij}$ is independent of $\theta$ and $a_{ij}$ is the output from the softmax layer, with form as given in equation \ref{eqn:a-ij}.
-%
-\begin{equation}
-	a_{ij} \coloneqq \phi_{j} (x_i; \theta) = \frac{\exp(w_j^T \tilde{x}_i)}{\sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i)}
-	\label{eqn:a-ij} 
-\end{equation}
-%
-We note that $\theta = \{w_k\}_{k=1}^B$, and as such we can write this in vector form $\theta = \left[w_1^T, w_2^T \dots w_B^T  \right]^T$.
-Therefore, $\nabla U(\theta) = \left[\nicefrac{\partial U}{\partial w_1}^T,\nicefrac{\partial U}{\partial w_2}^T \dots \nicefrac{\partial U}{\partial w_B}^T  \right]^T$; to compute $\nabla U(\theta)$ it suffices to find the form of $\nicefrac{\partial U}{\partial w_k}$ with respect to a general $k$.
-
-To this end, we must first find partial derivatives of $a_{ij}$ and $||\theta||$ with respect to $w_k$. Starting with $a_{ij}$:
-%
-\begin{align}
-	\frac{\partial a_{ij}}{\partial w_k} &= \frac
-	{\tilde{x}_i \exp(w_j^T \tilde{x}_i) \delta_{jk} \cdot \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) 
-		- 
-	\exp(w_j^T \tilde{x}_i) \cdot \tilde{x}_i \exp(w_k^T \tilde{x}_i)}
-	{\left( \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) \right)^2} \nonumber \\
-	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right) 
-\end{align}
-%
-Where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$. Now moving onto the derivative of $||\theta||^2$:
-%
-\begin{equation}
-	\frac{ \partial}{\partial w_k} ||\theta||^2 = \frac{\partial}{\partial w_k} \left( \sum_{b=1}^B ||w_b||^2 \right) = 2w_k
-\end{equation}
-%
-We are ready to put this all together, to find the partial derivative of $U(\theta)$ with respect to each $w_k$:
-\begin{align}
-	\frac{\partial U}{\partial w_k} &= 
-	\sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} 
-	\left( \frac{-\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
-	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
-	&=  - \left( \sum_{i=1}^{N} \tilde{x}_i \left( y_{ik} - a_{ik} \sum_{j=1}^{B} y_{ij} \right)
-	- \frac{w_k}{\sigma_\theta^2} \right) \nonumber \\
-	&= - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
-\end{align}
-%
-This is the required result.
-This form can be computed efficiently through matrix operations.
-The only property of $y_{ij}$ we have used in the derivation is the sum-to-one constraint $\sum_{j=1}^{B} y_{ij} = 1$ for all $i$.
-
-\subsection{Choosing the MALA step-size}
-\label{appdx:step-size}
-
-For sampling from the $\theta$-chain of the block membership generator parameters, we employed the Metropolis Adjusted Langevin Algorithm (MALA). At iteration $t$, the proposed sample is generated by:
-%
-\begin{equation}
-	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi
-\end{equation}
-%
-There are two competing objectives when choosing the step-size $h_t$.
-On the one hand, we want the step-size to be large so that we arrive at a high density region quickly.
-However, too large a step-size will lead to a lower acceptance ratio and thus inefficient sampling.
-A solution to this problem would be to slowly decrease the step-size with $t$ - often called simulated annealing.
-Therefore, we still have a short burn-in time but will not bounce around the mode for large $t$.
-As well as the trivial constraint for $h_t$ to be strictly positive and we introduce two further constraints as outlined by \citet{Bayesian-SGLD}:
-%
-\begin{equation}
-	\sum_{t=1}^{\infty} h_t = \infty \qquad \textrm{and} \qquad
-	\sum_{t=1}^{\infty} h_t^2 < \infty
-	\label{eqn:h-constraints}
-\end{equation}
-%
-The first constraint ensures that we have cover sufficient distance to arrive at any arbitrary point in our domain, no matter the starting point.
-The second constraint ensures that once we converge to the mode rather than simply bouncing around it.
-\citet{Bayesian-SGLD} propose the following form for a polynomially decaying step-size which we adopt:
-%
-\begin{equation}
-	h_t = a(b + t)^{-\gamma}
-\end{equation}
-%
-Where $a, b, \gamma$ are hyper-parameters to be chosen.
-We require $a,b > 0$ and $\gamma \in (0.5, 1]$ to satisfy equation \ref{eqn:h-constraints}.
-We find empirically that $a=250, b=1000, \gamma=0.8$ yield good burn-in times for our MALA implementations.
-
-\end{document}

### 2021-05-16 nips/prelim.tex
Split into multiple tex files

--- 2021-05-14 nips/submission.tex
@@
-\documentclass{article}
-
-% if you need to pass options to natbib, use, e.g.:
-%     \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading neurips_2021
-
-% ready for submission
-\PassOptionsToPackage{numbers, compress}{natbib}
-\usepackage[final]{neurips_2021}
-
-% to compile a preprint version, e.g., for submission to arXiv, add add the
-% [preprint] option:
-%     \usepackage[preprint]{neurips_2021}
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-%\usepackage[final]{neurips_2021}
-
-% to avoid loading the natbib package, add option nonatbib:
-%    \usepackage[nonatbib]{neurips_2021}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{xcolor}         % colors
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-\usepackage{amsmath}
-\usepackage{mathtools}
-\usepackage{bbm}
-\usepackage{amssymb}
-\usepackage{algorithmic}
-\usepackage{algorithm}
-
-%bibliography
-\bibliographystyle{plainnat}
-
-\title{Inferring community characteristics in labelled networks}
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to break the
-% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
-% authors names on the first line, and the last on the second line, try using
-% \AND instead of \And before the third author name.
-
-\author{
-  Lawrence Tray \\
-  Department of Engineering \\
-  University of Cambridge \\
-  \texttt{lpt30@cam.ac.uk} \\
-  % examples of more authors
-  \And
-  Ioannis Kontoyiannis \\
-  Department of Mathematics \\
-  University of Cambridge \\
-  \texttt{ik355@cam.ac.uk}
-}
-
-%custom commands
-\newcommand{\Xcal}{\mathcal{X}}
-\newcommand{\Bcal}{\mathcal{B}}
-\newcommand{\Integers}{\mathbb{Z}}
-\newcommand{\one}{\mathbbm{1}}
-\newcommand{\Gaussian}{\mathcal{N}}
-\newcommand{\indep}{\perp \!\!\! \perp}
-\newcommand{\specialchoose}{\genfrac{\{}{\}}{0pt}{}}
-\newcommand{\Expect}{\mathbb{E}}
-
-%envs
-\newtheorem{definition}{Definition}[section]
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[theorem]
-\newtheorem{lemma}[theorem]{Lemma}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-We present a novel generative model for describing vertex-labelled undirected graphs.
-This is a two-layered model.
-First a probabilistic mapping from vertex labels to latent block memberships and from these block memberships we generate a graph according to the micro-canonical stochastic block model.
-
-With this framework we are able to efficiently sample the parameters of the vertex label to block membership mapping.
-This allows us to automatically determine which features have the largest impact on graphical structure.
-\end{abstract}
-
-\section{Introduction}
-
-There is a wealth of networks in the world.
-Some examples of this graphical data are social networks, website hyperlinks and academic collaboration with more being produced each second.
-It is clear we need tools to analyse this increasingly omni-present form of data.
-
-A somewhat surprising property of real-world graphs is that they exhibit strong community structure.
-In other words, each node will often belong to a cluster of densely connected nodes.
-This property is often exploited by graph compression algorithms and there is high interest in recovering the communities from the observed graph.
-
-A common subset of graphical data is the labelled network.
-This is a graph where we have information about the properties of each node.
-We shall refer to these node properties as features.
-One of the most common questions we can ask of this dataset is what features have the largest impact on the structure of the graph.
-For example, when analysing an academic collaboration graph, one may wish to ask what impact gender has on the structure of the network.
-Nevertheless, this analysis is often vulnerable to confounding variables.
-While gender very well may impact the structure of the graph, there is often a better explanatory variable for the structure.
-
-There is space to bridge the gap between these two approaches.
-Rather than determining whether a feature impacts the graphical structure directly, we can use the community concept as a stepping stone in our analysis.
-Therefore, we extract which features have the largest impact on overall graphical structure.
-This can be thought of as a form of dimensionality-reduction where we only keep the node features that are useful in predicting community memberships.
-
 \section{Preliminaries}
@@
 \end{definition}
-
-\section{Latent block generative model}
-
-In this section we propose a novel generative model for modelling labelled networks.
-We restrict our analysis to labelled, undirected graphs with $N$ nodes.
-We define the vector $x_i \in \Xcal^D$ as the feature vector for the $i$'th vertex.
-Each vertex has $D$ total features and we assume all entries take values from the same set $\Xcal$.
-For the majority of datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$.
-The feature vectors $\{x_i\}_{i=1}^{N}$ are subsumed into the $N \times D$ matrix $X$.
-
-The proposed generative model (which we call the latent block generative model - LBGM) is given in figure \ref{fig:lbgm}.
-We start, with the feature matrix X and generate a vector of block memberships $b$.
-The parameters of this step are encapsulated by $\theta$.
-Each feature vector $x_i$ is treated independently and used to generate the block membership $b_i$.
-Each entry $b_i \in \{1, 2 \dots B\}$ where $B$ is the number of blocks and a constant in our model.
-We choose a single softmax layer to model $p(b_i | x_i, \theta)$.
-More complex models are possible but then deriving meaning from inferred parameter distributions is more difficult.
-Summarising, we write $p(b | X, \theta)$ as follows:
-%
-\begin{equation}
-	p(b| X, \theta) = \prod_{i=1}^{N} p(b_i | x_i, \theta) = \prod_{i=1}^{N} \phi_{b_i} (x_i; \theta)
-	= \prod_{i=1}^{N} \frac{\exp\left(w_{b_i}^T \tilde{x}_i\right)}{\sum_{k=1}^{B} \exp \left( w_k^T \tilde{x}_i\right)}
-\end{equation}
-%
-Where $\tilde{x} \coloneqq \left[x_1, x_2, \dots x_D, 1\right]^T$ is an augmented version of $x$ that allows for a bias term.
-The parameters for this stage are denoted $\theta$ and consist of the all $B$ weight vectors $\theta = \{w_k\}_{k=1}^{B}$.
-Each $w_k$ has dimension $D+1$.
-We could instead write the parameters $\theta$ as a $B \times (D+1)$ matrix of weights $W$; this form has use computationally as then $Z_i = W \tilde{x}_i$ is the input to the softmax activation function.
-
-\begin{figure}[!h]
-	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=7mm},
-		squarednode/.style={rectangle, draw=black, minimum size=7mm}
-		]
-		% nodes
-		\node[roundnode] (X) at (0, 0) {$X$};
-		\node[squarednode] (b) at (3, 0) {$b$};
-		\node[roundnode] (A) at (6, 0) {$A$};
-
-		% arrows
-		\draw[->] (X.east) -- node[above] {$\theta$} (b.west);
-		\draw[->] (b.east) -- node[above] {$\psi$}(A.west);
-	\end{tikzpicture}
-	\caption{Latent block generative model}
-	\label{fig:lbgm}
-\end{figure}
-
-With the block memberships $b$ generated, we then draw the graph $A$ from the microcanical DC-SBM (\ref{defn:microcan-dc-sbm}) with additional parameters $\psi = \{e, k\}$.
-In a slight abuse of notation we denote the inter-block edge count matrix with $e = \psi_e$ and the degree sequence $k = \psi_k$ to make explicit that these parameters are contained in $\psi$.
-%
-\begin{equation}
-	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k)
-\end{equation}
-
-\subsection{Prior selection}
-
-Before performing any inference, we must specify priors on $\theta$ and $\psi$.
-For $\theta$ it seems sensible to choose a Gaussian prior, with zero mean and variance matrix $\sigma^2_\theta I$ such that each element of $\theta$ is independent and distributed like $\sim \Gaussian(0, \sigma_\theta^2)$.
-In vector form, the prior for $\theta$ is therefore:
-%
-\begin{equation}
-	p(\theta) = \Gaussian \left( \theta ; 0, \sigma_\theta^2 I \right)
-	\label{eqn:theta-prior}
-\end{equation}
-%
-We will see that this form of prior is equivalent to a regularisation term in neural network training that penalises extreme weight magnitudes.
-For $\sigma_\theta^2 \rightarrow \infty$ this becomes an uninformative uniform prior.
-
-In our model, the block memberships vector $b$ is an intermediate latent variable and so we are not free to choose a prior for it.
-Nevertheless, as far as inference on the right-hand-side of figure \ref{fig:lbgm}, we regard $p(b | X)$ as a pseudo-prior on $b$.
-We can show (appendix \ref{appdx:b|x}) that our choice of prior for $p(\theta)$ in equation \ref{eqn:theta-prior} leads to a uniform $p(b | X)$ in equation \ref{eqn:b-pseudo-prior}.
-%
-\begin{equation}
-	p(b | X) = \int p(b | X, \theta) p(\theta) d\theta = B^{-N}
-	\label{eqn:b-pseudo-prior}
-\end{equation}
-%
-This is a uniform distribution that only depends on the number of blocks $B$.
-This simplifies the problem nicely as evaluating $p(b | X)$ does not require knowing the exact value $X$ takes.
-\citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for the additional microcanonical SBM parameters $\psi$ which we adopt.
-Peixoto's idea is to write the joint prior on $(b, e, k)$ as a product of conditionals $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$.
-For our purposes we must insert a conditioning on $X$, to form our pseudo-prior for $b$ and $\psi$, to give equation \ref{eqn:joint-pseudo-prior}.
-%
-\begin{equation}
-	p(b, \psi | X) = p(b | X) p(\psi | b, X) = p(b | X) p(\psi | b)
-	\label{eqn:joint-pseudo-prior}
-\end{equation}
-%
-Where the simplification is made apparent by noting $(\psi \indep X) | b$. We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference.
-%
-\begin{equation}
-	p(\psi | b) = p(e | b) p(k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
-	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right]
-\end{equation}
-%
-Where $\specialchoose{n}{m}$ is shorthand for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of distinct histograms with $n$ bins under the constraint they sum to $m$.
-$E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges in the graph.
-Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform with respect to $e$.
-The variable $\eta_j^r$ is introduced to denote the number of vertices in block $r$ that have degree $j$.
-Formally, $\eta_j^r \coloneqq \sum_{i} \one\left\{b_i = r \right\} \one \left\{k_i = j \right\}$.
-Furthermore, $q(m, n)$ is the number of different histograms with at most $n$ non-zero bins that sum to $m$.
-$q(m, n)$ is related to but different from $\specialchoose{n}{m}$.
-Lastly, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ is the number of vertices assigned to block $r$.
-Importantly, we have computable forms for $p(\theta)$ and $p(b, \psi | X)$ which will be useful for performing inference.
-
-\section{Inference}
-
-Now that we have defined the LBGM, we wish to leverage it to perform inference.
-Suppose we are presented with a vertex-labelled graph $(A, X)$; the goal is to draw samples for $\theta$ according to the posterior given the observed graph (equation \ref{eqn:theta-target}).
-%
-\begin{equation}
-	\label{eqn:theta-target}
-	\theta^{(t)} \sim p(\theta | A, X)
-\end{equation}
-%
-These samples allow us to approximate the posterior distribution for $\theta$ as well as compute a predictive distribution $p(b^* | x^*, A, X) = \int p(b^* | x^*, \theta) p(\theta | X, A) d\theta \approx \frac{1}{T} \sum_{t=1}^{T} p(b^* | x^*, \theta^{(t)})$.
-However, generating these samples is not easily done in practice.
-
-We instead propose an iterative approach.
-First drawing samples $b^{(t)}$ from the block membership posterior (equation \ref{eqn:b-samples}).
-We then use each $b^{(i)}$ to draw samples for $\theta$ as in equation \ref{eqn:theta-samples}.
-%
-\begin{align}
-	b^{(i)} &\sim p \Big( b | A, X \Big)  \label{eqn:b-samples}\\
-	\theta^{(i)} &\sim p\Big(\theta | X, b^{(i)} \Big) \label{eqn:theta-samples}
-\end{align}
-%
-Both of these sampling steps implemented with a Markov Chain through the Metropolis-Hastings algorithm \cite{hastings-alg}.
-We just need to define a proposal distribution $q(x, x')$ for proposing a move $x \rightarrow x'$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(\cdot)$, point-wise.
-The proposed move is then accepted with probability $\alpha$ (equation \ref{eqn:mh-accept}) else it is rejected and we stay at $x$.
-%
-\begin{equation}
-	\alpha = \min \left( \frac{\pi(x') q(x', x)}{\pi(x) q(x, x')} , 1 \right)
-	\label{eqn:mh-accept}
-\end{equation}
-%
-This accept-reject step ensures the resulting Markov Chain is in detailed balance with the target distribution $\pi(\cdot)$.
-What we propose in equations \ref{eqn:b-samples} and \ref{eqn:theta-samples} is therefore implemented through a 2-level Markov chain.
-The resulting samples for $\theta^{(i)}$ are unbiased in the sense that the expectation of their distribution is the posterior we are targeting in equation \ref{eqn:theta-target}.
-%
-\begin{align*}
-	\Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] &= \sum_{b \in \mathcal{B}^N} p(\theta | X, b) p(b | A, X) \\
-	&= \sum_{b \in \mathcal{B}^N} p(\theta, b | A, X) \\
-	&= p(\theta | A, X) \\
-\end{align*}
-%
-This is an example of a pseudo-marginal approach \cite{pseudo-marginal}.
-Indeed, the unbiased result is sufficient to prove that for sufficient samples, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{t})\right] = p(\theta
-\ A, X)$ which is exactly the distribution we were targetting (equation \ref{eqn:theta-target}).
-
-The reason we split the Markov chain into two stages is because the summation over all latent states $b \in \mathcal{B}^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in \Bcal^N} p(A | b) P(b | X, \theta)$ is intractable $O(B^N)$.
-%
-\begin{figure}[!h]
-	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=12mm},
-		squarednode/.style={rectangle, draw=black, minimum size=12mm}
-		]
-		% nodes
-		\node[roundnode] (b0) at (0, 2) {$b^{(t)}$};
-		\node[roundnode] (b1) at (4, 2) {$b'$};
-		\node[roundnode] (b2) at (8, 2) {$b^{(t+1)}$};
-		\node[roundnode] (t0) at (0, 0) {$\theta^{(t)}$};
-		\node[roundnode] (t1) at (4, 0) {$\theta'$};
-		\node[roundnode] (t2) at (8, 0) {$\theta^{(t+1)}$};
-
-		% arrows
-		\draw[->] (b0) to node[above] {$\sim q_b(b^{(t)}, b' | A)$} (b1);
-		\draw[->] (b1) to node[above] {$\alpha_b (b, b' | A)$} (b2);
-		\draw[->] (b0) [out=30, in=150] to node[above] {$1-\alpha_b$} (b2);
-
-		\draw[->] (t0) to node[above] {$\sim q_\theta(\theta, \theta' | X, b^{(t)})$} (t1);
-		\draw[->] (t1) to node[above] {$\alpha_\theta (\theta, \theta' | X, b^{(t)})$} (t2);
-		\draw[->] (t0) [out=-30, in=-150] to node[below] {$1-\alpha_\theta$} (t2);
-
-		\draw[dashed, ->] (b0) to (2, 0.8);
-		\draw[dashed, ->] (b0) to (5.5, 0.8);
-
-	\end{tikzpicture}
-	\caption{Sampling sequence}
-	\label{fig:samp-sequence}
-\end{figure}
-%
-Figure \ref{fig:samp-sequence} shows an overview of the proposed method.
-We have introduced subscripts and conditionings to make explicit what parameters each step utilises.
-In an important simplification, we note that $p(b| X) = B^{-N}$ which does not depend on the exact value of X.
-Therefore, we do not need to know the value of $X$ to perform the sampling on $b$.
-Conversely, for the $\theta^{(t)}$ samples, we use $b^{(t)}$ but not $A$ as $(\theta \indep A) | b$.
-
-\subsection{Sampling block memberships}
-
-\citet{Peixoto-MCMC} proposes a Monte Carlo method which we will base our approach on. It relies on writing the posterior in the following form:
-%
-\begin{equation}
-	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) = \pi_b(b)
-\end{equation}
-%
-Now $\pi_b(\cdot)$ is the un-normalised density we wish to sample from.
-In other words, we wish to construct a Markov chain that has $\pi_b(\cdot)$ as its invariant distribution.
-We can break $\pi_b$ down as follows:
-%
-\begin{align*}
-	\pi_b(b) &= p(b|X) \sum_{\psi} \nolimits p(A , \psi | b, X) \\
-	&= p(b|X) p(A, \psi^* | b, X) \\
-	&= p(A | b, \psi^*) \cdot p(\psi^* | b) \cdot p(b | X)
-\end{align*}
-%
-Since we are using the microcanonical SBM formulation, there is only one value of $\psi$ that is compatible with the given $(A, b)$ pair.
-We denote this value $\psi^* = \{k^*, e^*\}$.
-Specifically, $k^*_i = \sum_j A_{ij} $ and $e^*_{rs} = \sum_{i, j} A_{ij} \one \{b_i=r\} \one\{b_j=s\}$.
-Therefore, the summation over all $\psi$ reduces to just the single $\psi^*$ term.
-We also define the microcanonical entropy of the configuration as.
-%
-\begin{equation}
-	S(b) = - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big)
-\end{equation}
-%
-This entropy can be thought of as the description length of the graph because it is the sum of the information required to represent the graph given the parameters and the amount of information required to store the parameters (given the feature matrix $X$).
-The exact, from of the proposal distribution and accept-reject step is explored throughly by \citet{Peixoto-MCMC}.
-There is a widely available library, \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, implementing this algorithm.
-The only modification we make is in the block membership prior $p(b)$ which we replace with $p(b|X)=B^{-N}$ which is a uniform distribution and so cancels out in the MH accept-reject step.
-
-The end result of this is that we can generate a set of block membership samples $\{b^{(t)}\}_{t=1}^{T}$ with each $b^{(t)} \sim p(b | A, X)$.
-Each of these samples can then be used for the $\theta$-chain.
-
-\subsection{Sampling feature-to-block classifier parameters}
-
-The invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. We write this as follows:
-%
-\begin{align}
-	p(\theta | X, b) &\propto p(b | X, \theta) p(\theta) = \pi_\theta (\theta) \propto  \exp \left( - U(\theta) \right) \\
-	\therefore U(\theta) &= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}
-\end{align}
-%
-Where we have introduced $U(\theta)$ equal to the negative log posterior.
-This $U(\theta)$ term makes subsequent analysis more concise.
-Each of the constituent terms of $U(\theta)$ is easily computed (equation \ref{eqn:u-form}).
-To simplify notation, we define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
-%
-\begin{equation}
-	\log p(b | X, \theta) = \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
-	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta ||^2
-	\label{eqn:U-constituent-terms}
-\end{equation}
-%
-This means that if we ignore constant terms we can write $U(\theta)$ as in equation \ref{eqn:U-form}.
-We see that the prior effectively introduces a regularisation term.
-Note that $||\theta||^2 = \sum_{i} \theta_{i}^2 = \sum_{j=1}^{B} ||w_j||^2$ is the Euclidean norm of the vector of parameters $\theta$.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form}
-\end{equation}
-%
-$U(\theta)$ in equation \ref{eqn:U-form} appears a typical objective function to be minimised for neural network training.
-The first term is the cross-entropy between actual and predicted labels.
-The second term - introduced by the prior - brings a form of regularisation, preventing over-fitting.
-In traditional applications we only seek the value of $\theta$ that minimises the objective function $U(\theta)$, which in our case would yield the maximum a posteriori (MAP) estimate.
-This is often done through some kind of gradient descent as $\nabla U$ is easily computable (equation \ref{eqn:U-derivative}).
-
-However, our goal is not to find the MAP estimate but to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
-As discussed earlier, given the invariant $\pi_\theta(\cdot)$, it is sufficient to specify a proposal distribution and then apply a MH accept-reject step to ensure detailed balance of the Markov Chain.
-Nevertheless, we can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
-We therefore adopt the Metropolis Adjusted Langevin Algorithm (MALA) - first proposed by \citet{mala-tweedie} - which leverages just that.
-Given the current sample $\theta$, we propose a new sample $\theta'$ according to equation \ref{eqn:theta-update}
-%
-\begin{equation}
-	\theta' = \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi
-	\label{eqn:theta-update}
-\end{equation}
-%
-Where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter - which may vary with the sample index.
-Indeed without the injected noise term, this is equivalent to gradient descent.
-We require the noise term to fully explore the parameter space.
-As such the proposal distribution, is a simple multivariate Gaussian which can be easily evaluated.
-%
-\begin{equation}
-	q_\theta(\theta, \theta') = \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right)
-\end{equation}
-%
-The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}). By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as:
-%
-\begin{equation}
-	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
-	\label{eqn:U-derivative}
-\end{equation}
-%
-After a proposed move is generated, in typical Metropolis-Hastings fashion we accept the move with probability $\alpha_\theta$.
-%
-\begin{equation}
-	\alpha_\theta(\theta, \theta') = \min \left( 
-	\exp \left( U(\theta) - U(\theta')\right)
-	\frac{ 
-		q_\theta(\theta', \theta)
-	}{
-		q_\theta(\theta, \theta')
-	} 
-	, 1 \right)
-\end{equation}
-%
-This fully specifies, the sampling procedure to generate $\{\theta^{(t)}\}_{t=1}^T$. Bear in mind that each $\theta^{(t)}$ update step uses its corresponding $b^{(t)}$ block membership sample.
-
-\subsection{Sampling sequence}
-
-So far, each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample.
-This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance.
-This may lead to longer burn-in and autocorrelation times of the resulting Markov Chain.
-The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends on $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
-We would rather deal with the expectation of each $y_{ij}^{(t)}$:
-%
-\begin{equation}
-	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one(b_{i}^{(t)} = j) \right]
-	= p(b_i = j | A, X)
-\end{equation}
-%
-We obtain an unbiased estimate for this quantity as simply the empirical distribution of the block membership samples $\left\{ b^{(t)} \right\}_{t=1}^T$.
-%
-\begin{equation}
-	\hat{y}_{ij} \coloneqq \frac{1}{T} \sum_{t=1}^{T} y_{ij}^{(t)} = \frac{1}{T} \sum_{t=1}^{T} \one\{b_i^{(t)} = j\}
-\end{equation}
-%
-We therefore, choose to feed each $\theta^{(t)}$ update step the same $\hat{y}_{ij}$ for all $t$ rather than the corresponding $y^{(t)}_{ij}$.
-This means we no longer need to run the $b$ and $\theta$ Markov chains concurrently.
-Instead, we run the $b$-chain to completion and use it to generate $\hat{y}_{ij}$ for $i \in \{1 \dots N\}$ and $j \in \{1 \dots B\}$.
-This is an estimate of $p(b | A, X)$ that we use for every iteration of the $\theta$ Markov chain.
-This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$; wed refer to these as $T_b$ and $T_\theta$ henceforth.
-Furthermore, this changeover reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{Block membership sample generation} % give the algorithm a caption
-	\label{alg:b-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\FOR{$t \in \{0, 1 \dots T_b - 1\}$}
-			\STATE $b' \gets \sim q_b(b^{(t)}, b' | A)$
-			\STATE $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
-			\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-			\IF{$\log \eta < \log \alpha_b$}
-				\STATE $b^{(t+1)} \gets b'$
-			\ELSE
-				\STATE $b^{(t+1)} \gets b^{(t)}$
-			\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{LBGM parameter pseudo-marginal inference} % give the algorithm a caption
-	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\STATE $\mathcal{S} \gets \textrm{range}(\textrm{burn-in}, T_b, \textrm{thinning})$
-		\STATE $\hat{Y}_{ij} \gets \frac{1}{\mathcal{|S|}} \sum_{t \in \mathcal{S}} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
-
-		\item[]
-
-		\FOR{$t \in \{0, 1 \dots T_\theta - 1\}$}
-		\STATE $\xi \gets \sim \Gaussian(0, I)$
-		\STATE $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
-		\STATE $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
-		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-		\IF{$\log \eta < \log \alpha_\theta$}
-		\STATE $\theta^{(t+1)} \gets \theta'$
-		\ELSE
-		\STATE $\theta^{(t+1)} \gets \theta^{(t)}$
-		\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\subsection{Complexity Analysis}
-
-\section{Experiments}
-
-We apply the developed methods to a variety of datasets.
-
-\begin{itemize}
-	\item \textbf{College football teams} \cite{Evans_2010_football} - network of American college football teams and their interactions. Vertex labels are the division a team belongs to.
-	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.
-	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier} - network of the author's Facebook friends list.
-Each vertex has been manually labelled with a variety of features describing their relationship to the author.
-	\item \textbf{Twitch users} \cite{twitch} - a network of user-user friendships on the streaming service Twitch.
-Vertex labels are extracted according to video-games played, location and streaming habits.
-This dataset is also broken down into disjoint networks according to language.
-We only consider the English users with is a subnet with $N=7126$ vertices and $E=35324$ edges).
-\end{itemize}
-
-\section{Extensions}
-
-\section{Conclusion}
-
-\begin{ack}
-Use unnumbered first level headings for the acknowledgments. All acknowledgments
-go at the end of the paper before the list of references. Moreover, you are required to declare
-funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
-More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2021/PaperInformation/FundingDisclosure}.
-
-Do {\bf not} include this section in the anonymized submission, only in the final paper.
-You can use the \texttt{ack} environment provided in the style file to automatically hide this section in the anonymized submission.
-\end{ack}
-
-%references
-\bibliography{sources.bib}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Checklist}
-
-%%% BEGIN INSTRUCTIONS %%%
-The checklist follows the references.  Please
-read the checklist guidelines carefully for information on how to answer these
-questions.  For each question, change the default \answerTODO{} to \answerYes{},
-\answerNo{}, or \answerNA{}.  You are strongly encouraged to include a {\bf
-justification to your answer}, either by referencing the appropriate section of
-your paper or providing a brief inline description.  For example:
-\begin{itemize}
-  \item Did you include the license to the code and datasets? \answerYes{See Section gen-inst.}
-  \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
-  \item Did you include the license to the code and datasets? \answerNA{}
-\end{itemize}
-Please do not modify the questions and only use the provided macros for your
-answers.  Note that the Checklist section does not count towards the page
-limit.  In your paper, please delete this instructions block and only keep the
-Checklist section heading above along with the questions/answers below.
-%%% END INSTRUCTIONS %%%
-
-\begin{enumerate}
-
-\item For all authors...
-\begin{enumerate}
-  \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
-    \answerTODO{}
-  \item Did you describe the limitations of your work?
-    \answerTODO{}
-  \item Did you discuss any potential negative societal impacts of your work?
-    \answerTODO{}
-  \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you are including theoretical results...
-\begin{enumerate}
-  \item Did you state the full set of assumptions of all theoretical results?
-    \answerTODO{}
-	\item Did you include complete proofs of all theoretical results?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you ran experiments...
-\begin{enumerate}
-  \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
-    \answerTODO{}
-  \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
-    \answerTODO{}
-	\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
-    \answerTODO{}
-	\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
-\begin{enumerate}
-  \item If your work uses existing assets, did you cite the creators?
-    \answerTODO{}
-  \item Did you mention the license of the assets?
-    \answerTODO{}
-  \item Did you include any new assets either in the supplemental material or as a URL?
-    \answerTODO{}
-  \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
-    \answerTODO{}
-  \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you used crowdsourcing or conducted research with human subjects...
-\begin{enumerate}
-  \item Did you include the full text of instructions given to participants and screenshots, if applicable?
-    \answerTODO{}
-  \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
-    \answerTODO{}
-  \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
-    \answerTODO{}
-\end{enumerate}
-
-\end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\appendix
-
-\section{Appendix}
-
-\subsection{Derivation of conditional block distribution given feature matrix}
-\label{appdx:b|x}
-
-We wish to determine the form of $p(b| X)$. This can be done by integrating over the joint probability with respect to $\theta$.
-%
-\begin{align*}
-	p(b | X) &= \int p(b , \theta| X, \theta) d\theta = \int p(b | X, \theta) p(\theta | X) d\theta \\
-	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i=1}^{N} \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
-	&= \prod_{i=1}^{N} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j=1}^{B} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k=1}^{B} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}
-\end{align*}
-%
-We note that $b_i \in {1, 2, \dots B}$ and so the integral's value is unchanged with respect to $b_i$.
-The integrand has the same form no matter which value $b_i$ takes as the prior is the same for each $w_j$.
-As such the integral can only be a function of at most $\tilde{x}_i$ and $\sigma_\theta^2$ as it is symmetric with respect to $b_i$ and all the various $w_j$ are integrated out as they are dummy variables.
-Therefore, denoting the integral by the (unknown) function $f(\tilde{x}_i, \sigma_\theta^2)$, we write $p(b| X)$ as follows:
-%
-\begin{align*}
-	p(b | X) &= \prod_{i=1}^{N} f(\tilde{x}_i, \sigma_\theta^2) = \textrm{const w.r.t } b = c
-\end{align*}
-%
-As this is a constant with respect to $b$ we conclude that $p(b | X)$ must be a uniform distribution.
-$\nicefrac{1}{c}$ is simply the size of the set of values that $b$ can take.
-We know $b_i \in \mathcal{B} = \{1, 2, \dots B\}$.
-Therefore, $b \in \mathcal{B}^N$ and $|\mathcal{B}^N| = |\mathcal{B}|^N = B^N= \nicefrac{1}{c}$.
-Putting this all together we show that:
-%
-\begin{equation}
-	p(b | X) = B^{-N}
-\end{equation}
-
-\subsection{Derivation of U gradient with respect to feature parameters}
-\label{appdx:gradu}
-The goal is to determine $\nabla U(\theta)$, the gradient of the negative log posterior with respect to the parameters. We repeat the form of $U(\theta)$ in equation \ref{eqn:U-form-appdx}.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form-appdx}
-\end{equation}
-%
-Where $y_{ij}$ is independent of $\theta$ and $a_{ij}$ is the output from the softmax layer, with form as given in equation \ref{eqn:a-ij}.
-%
-\begin{equation}
-	a_{ij} \coloneqq \phi_{j} (x_i; \theta) = \frac{\exp(w_j^T \tilde{x}_i)}{\sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i)}
-	\label{eqn:a-ij} 
-\end{equation}
-%
-We note that $\theta = \{w_k\}_{k=1}^B$, and as such we can write this in vector form $\theta = \left[w_1^T, w_2^T \dots w_B^T  \right]^T$.
-Therefore, $\nabla U(\theta) = \left[\nicefrac{\partial U}{\partial w_1}^T,\nicefrac{\partial U}{\partial w_2}^T \dots \nicefrac{\partial U}{\partial w_B}^T  \right]^T$; to compute $\nabla U(\theta)$ it suffices to find the form of $\nicefrac{\partial U}{\partial w_k}$ with respect to a general $k$.
-
-To this end, we must first find partial derivatives of $a_{ij}$ and $||\theta||$ with respect to $w_k$. Starting with $a_{ij}$:
-%
-\begin{align}
-	\frac{\partial a_{ij}}{\partial w_k} &= \frac
-	{\tilde{x}_i \exp(w_j^T \tilde{x}_i) \delta_{jk} \cdot \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) 
-		- 
-	\exp(w_j^T \tilde{x}_i) \cdot \tilde{x}_i \exp(w_k^T \tilde{x}_i)}
-	{\left( \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) \right)^2} \nonumber \\
-	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right) 
-\end{align}
-%
-Where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$. Now moving onto the derivative of $||\theta||^2$:
-%
-\begin{equation}
-	\frac{ \partial}{\partial w_k} ||\theta||^2 = \frac{\partial}{\partial w_k} \left( \sum_{b=1}^B ||w_b||^2 \right) = 2w_k
-\end{equation}
-%
-We are ready to put this all together, to find the partial derivative of $U(\theta)$ with respect to each $w_k$:
-\begin{align}
-	\frac{\partial U}{\partial w_k} &= 
-	\sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} 
-	\left( \frac{-\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
-	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
-	&=  - \left( \sum_{i=1}^{N} \tilde{x}_i \left( y_{ik} - a_{ik} \sum_{j=1}^{B} y_{ij} \right)
-	- \frac{w_k}{\sigma_\theta^2} \right) \nonumber \\
-	&= - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
-\end{align}
-%
-This is the required result.
-This form can be computed efficiently through matrix operations.
-The only property of $y_{ij}$ we have used in the derivation is the sum-to-one constraint $\sum_{j=1}^{B} y_{ij} = 1$ for all $i$.
-
-\subsection{Choosing the MALA step-size}
-\label{appdx:step-size}
-
-For sampling from the $\theta$-chain of the block membership generator parameters, we employed the Metropolis Adjusted Langevin Algorithm (MALA). At iteration $t$, the proposed sample is generated by:
-%
-\begin{equation}
-	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi
-\end{equation}
-%
-There are two competing objectives when choosing the step-size $h_t$.
-On the one hand, we want the step-size to be large so that we arrive at a high density region quickly.
-However, too large a step-size will lead to a lower acceptance ratio and thus inefficient sampling.
-A solution to this problem would be to slowly decrease the step-size with $t$ - often called simulated annealing.
-Therefore, we still have a short burn-in time but will not bounce around the mode for large $t$.
-As well as the trivial constraint for $h_t$ to be strictly positive and we introduce two further constraints as outlined by \citet{Bayesian-SGLD}:
-%
-\begin{equation}
-	\sum_{t=1}^{\infty} h_t = \infty \qquad \textrm{and} \qquad
-	\sum_{t=1}^{\infty} h_t^2 < \infty
-	\label{eqn:h-constraints}
-\end{equation}
-%
-The first constraint ensures that we have cover sufficient distance to arrive at any arbitrary point in our domain, no matter the starting point.
-The second constraint ensures that once we converge to the mode rather than simply bouncing around it.
-\citet{Bayesian-SGLD} propose the following form for a polynomially decaying step-size which we adopt:
-%
-\begin{equation}
-	h_t = a(b + t)^{-\gamma}
-\end{equation}
-%
-Where $a, b, \gamma$ are hyper-parameters to be chosen.
-We require $a,b > 0$ and $\gamma \in (0.5, 1]$ to satisfy equation \ref{eqn:h-constraints}.
-We find empirically that $a=250, b=1000, \gamma=0.8$ yield good burn-in times for our MALA implementations.
-
-\end{document}

### 2021-05-16 nips/submission.tex
Split into multiple tex files

--- 2021-05-14 nips/submission.tex
@@
 
-\section{Introduction}
+\input{introduction}
 
-There is a wealth of networks in the world.
-Some examples of this graphical data are social networks, website hyperlinks and academic collaboration with more being produced each second.
-It is clear we need tools to analyse this increasingly omni-present form of data.
+\input{prelim}
 
-A somewhat surprising property of real-world graphs is that they exhibit strong community structure.
-In other words, each node will often belong to a cluster of densely connected nodes.
-This property is often exploited by graph compression algorithms and there is high interest in recovering the communities from the observed graph.
+\input{lbgm}
 
-A common subset of graphical data is the labelled network.
-This is a graph where we have information about the properties of each node.
-We shall refer to these node properties as features.
-One of the most common questions we can ask of this dataset is what features have the largest impact on the structure of the graph.
-For example, when analysing an academic collaboration graph, one may wish to ask what impact gender has on the structure of the network.
-Nevertheless, this analysis is often vulnerable to confounding variables.
-While gender very well may impact the structure of the graph, there is often a better explanatory variable for the structure.
+\input{inference}
 
-There is space to bridge the gap between these two approaches.
-Rather than determining whether a feature impacts the graphical structure directly, we can use the community concept as a stepping stone in our analysis.
-Therefore, we extract which features have the largest impact on overall graphical structure.
-This can be thought of as a form of dimensionality-reduction where we only keep the node features that are useful in predicting community memberships.
+\input{experiments}
 
-\section{Preliminaries}
+\input{extensions}
 
-This section defines some preliminary concepts required for the subsequent analysis.
-We first need a model for community-like structure in a graphical network.
-For this we adopt the stochastic block model (SBM) - widely used across academia.
-The premise is that each node in the graph belongs to a unique community called a block.
-The probability that two nodes are connected depends only on the block memberships of each.
-Graphs drawn from this ensemble exhibit community structure.
-Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}.
-A paraphrased definition is given below.
+\input{conclusion}
 
-\begin{definition}[Microcanonical NDC-SBM]
-	\label{defn:microcan-ndc-sbm}
-	Let $N \in \Integers^{+}$ denote the number of vertices in an undirected graph.
-The block memberships are encoded by a vector $b$ of length $N$ where each entry $b_i \in \{1, 2 \dots B\}$.
-$B \in \Integers^{+}$ is the number of non-empty blocks.
-Let $e$ be a $B \times B$ matrix of edge counts between blocks ($e_{rs}$ is number of edges from block $r$ onto block $s$ - or twice that number if $r=s$).
-For undirected graphs $e$ is symmetric.
-For a non-degree-corrected stochastic block model (NDC-SBM), we say that the graph $A$ is generated as follows:
-	%
-	\begin{equation}
-		A \sim \textrm{NDC-SBM}_{\textrm{MC}} (b, e)
-	\end{equation}
-	%
-	Where edges are placed uniformly at random but respecting the constraint imposed by $e$ and $b$.
-The additional parameters $N$ and $B$ are omitted as they are inferred from the shapes of $b$ and $e$.
-If we interpret $A$ as an adjacency matrix, then this constraint can be written formally as: $e_{rs} = \sum_{i,j} A_{ij} \one \{b_i = r\} \one \{b_j = s\}$.
-\end{definition}
-
-Nevertheless, this formulation does not accept high degree variability within blocks as is typical of real-world data.
-Indeed, the NDC-SBM favours a partition into high-degree and low-degree nodes rather than clusters of inter-connected nodes.
-We therefore introduce the degree-corrected SBM (DC-SBM) to circumvent these issues.
-
-\begin{definition}[Microcanonical DC-SBM]
-	\label{defn:microcan-dc-sbm}
-	This is much like the NDC-SBM but has an additional parameter $k$ which is an $N$-length vector encoding the degree sequence ($k_i$ is the degree of vertex $i$). Therefore, we write:
-	%
-	\begin{equation}
-		A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, e, k)
-	\end{equation}
-	%
-	Once again, edges are placed uniformly at random but respecting the constraints imposed by the parameters.
-The DC-SBM has the additional constraint that $k_i = \sum_{j} A_{ij}$.
-In what follows, we will always assume the degree-corrected model unless otherwise specified.
-
-\end{definition}
-
-\section{Latent block generative model}
-
-In this section we propose a novel generative model for modelling labelled networks.
-We restrict our analysis to labelled, undirected graphs with $N$ nodes.
-We define the vector $x_i \in \Xcal^D$ as the feature vector for the $i$'th vertex.
-Each vertex has $D$ total features and we assume all entries take values from the same set $\Xcal$.
-For the majority of datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$.
-The feature vectors $\{x_i\}_{i=1}^{N}$ are subsumed into the $N \times D$ matrix $X$.
-
-The proposed generative model (which we call the latent block generative model - LBGM) is given in figure \ref{fig:lbgm}.
-We start, with the feature matrix X and generate a vector of block memberships $b$.
-The parameters of this step are encapsulated by $\theta$.
-Each feature vector $x_i$ is treated independently and used to generate the block membership $b_i$.
-Each entry $b_i \in \{1, 2 \dots B\}$ where $B$ is the number of blocks and a constant in our model.
-We choose a single softmax layer to model $p(b_i | x_i, \theta)$.
-More complex models are possible but then deriving meaning from inferred parameter distributions is more difficult.
-Summarising, we write $p(b | X, \theta)$ as follows:
-%
-\begin{equation}
-	p(b| X, \theta) = \prod_{i=1}^{N} p(b_i | x_i, \theta) = \prod_{i=1}^{N} \phi_{b_i} (x_i; \theta)
-	= \prod_{i=1}^{N} \frac{\exp\left(w_{b_i}^T \tilde{x}_i\right)}{\sum_{k=1}^{B} \exp \left( w_k^T \tilde{x}_i\right)}
-\end{equation}
-%
-Where $\tilde{x} \coloneqq \left[x_1, x_2, \dots x_D, 1\right]^T$ is an augmented version of $x$ that allows for a bias term.
-The parameters for this stage are denoted $\theta$ and consist of the all $B$ weight vectors $\theta = \{w_k\}_{k=1}^{B}$.
-Each $w_k$ has dimension $D+1$.
-We could instead write the parameters $\theta$ as a $B \times (D+1)$ matrix of weights $W$; this form has use computationally as then $Z_i = W \tilde{x}_i$ is the input to the softmax activation function.
-
-\begin{figure}[!h]
-	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=7mm},
-		squarednode/.style={rectangle, draw=black, minimum size=7mm}
-		]
-		% nodes
-		\node[roundnode] (X) at (0, 0) {$X$};
-		\node[squarednode] (b) at (3, 0) {$b$};
-		\node[roundnode] (A) at (6, 0) {$A$};
-
-		% arrows
-		\draw[->] (X.east) -- node[above] {$\theta$} (b.west);
-		\draw[->] (b.east) -- node[above] {$\psi$}(A.west);
-	\end{tikzpicture}
-	\caption{Latent block generative model}
-	\label{fig:lbgm}
-\end{figure}
-
-With the block memberships $b$ generated, we then draw the graph $A$ from the microcanical DC-SBM (\ref{defn:microcan-dc-sbm}) with additional parameters $\psi = \{e, k\}$.
-In a slight abuse of notation we denote the inter-block edge count matrix with $e = \psi_e$ and the degree sequence $k = \psi_k$ to make explicit that these parameters are contained in $\psi$.
-%
-\begin{equation}
-	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k)
-\end{equation}
-
-\subsection{Prior selection}
-
-Before performing any inference, we must specify priors on $\theta$ and $\psi$.
-For $\theta$ it seems sensible to choose a Gaussian prior, with zero mean and variance matrix $\sigma^2_\theta I$ such that each element of $\theta$ is independent and distributed like $\sim \Gaussian(0, \sigma_\theta^2)$.
-In vector form, the prior for $\theta$ is therefore:
-%
-\begin{equation}
-	p(\theta) = \Gaussian \left( \theta ; 0, \sigma_\theta^2 I \right)
-	\label{eqn:theta-prior}
-\end{equation}
-%
-We will see that this form of prior is equivalent to a regularisation term in neural network training that penalises extreme weight magnitudes.
-For $\sigma_\theta^2 \rightarrow \infty$ this becomes an uninformative uniform prior.
-
-In our model, the block memberships vector $b$ is an intermediate latent variable and so we are not free to choose a prior for it.
-Nevertheless, as far as inference on the right-hand-side of figure \ref{fig:lbgm}, we regard $p(b | X)$ as a pseudo-prior on $b$.
-We can show (appendix \ref{appdx:b|x}) that our choice of prior for $p(\theta)$ in equation \ref{eqn:theta-prior} leads to a uniform $p(b | X)$ in equation \ref{eqn:b-pseudo-prior}.
-%
-\begin{equation}
-	p(b | X) = \int p(b | X, \theta) p(\theta) d\theta = B^{-N}
-	\label{eqn:b-pseudo-prior}
-\end{equation}
-%
-This is a uniform distribution that only depends on the number of blocks $B$.
-This simplifies the problem nicely as evaluating $p(b | X)$ does not require knowing the exact value $X$ takes.
-\citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for the additional microcanonical SBM parameters $\psi$ which we adopt.
-Peixoto's idea is to write the joint prior on $(b, e, k)$ as a product of conditionals $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$.
-For our purposes we must insert a conditioning on $X$, to form our pseudo-prior for $b$ and $\psi$, to give equation \ref{eqn:joint-pseudo-prior}.
-%
-\begin{equation}
-	p(b, \psi | X) = p(b | X) p(\psi | b, X) = p(b | X) p(\psi | b)
-	\label{eqn:joint-pseudo-prior}
-\end{equation}
-%
-Where the simplification is made apparent by noting $(\psi \indep X) | b$. We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference.
-%
-\begin{equation}
-	p(\psi | b) = p(e | b) p(k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
-	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right]
-\end{equation}
-%
-Where $\specialchoose{n}{m}$ is shorthand for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of distinct histograms with $n$ bins under the constraint they sum to $m$.
-$E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges in the graph.
-Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform with respect to $e$.
-The variable $\eta_j^r$ is introduced to denote the number of vertices in block $r$ that have degree $j$.
-Formally, $\eta_j^r \coloneqq \sum_{i} \one\left\{b_i = r \right\} \one \left\{k_i = j \right\}$.
-Furthermore, $q(m, n)$ is the number of different histograms with at most $n$ non-zero bins that sum to $m$.
-$q(m, n)$ is related to but different from $\specialchoose{n}{m}$.
-Lastly, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ is the number of vertices assigned to block $r$.
-Importantly, we have computable forms for $p(\theta)$ and $p(b, \psi | X)$ which will be useful for performing inference.
-
-\section{Inference}
-
-Now that we have defined the LBGM, we wish to leverage it to perform inference.
-Suppose we are presented with a vertex-labelled graph $(A, X)$; the goal is to draw samples for $\theta$ according to the posterior given the observed graph (equation \ref{eqn:theta-target}).
-%
-\begin{equation}
-	\label{eqn:theta-target}
-	\theta^{(t)} \sim p(\theta | A, X)
-\end{equation}
-%
-These samples allow us to approximate the posterior distribution for $\theta$ as well as compute a predictive distribution $p(b^* | x^*, A, X) = \int p(b^* | x^*, \theta) p(\theta | X, A) d\theta \approx \frac{1}{T} \sum_{t=1}^{T} p(b^* | x^*, \theta^{(t)})$.
-However, generating these samples is not easily done in practice.
-
-We instead propose an iterative approach.
-First drawing samples $b^{(t)}$ from the block membership posterior (equation \ref{eqn:b-samples}).
-We then use each $b^{(i)}$ to draw samples for $\theta$ as in equation \ref{eqn:theta-samples}.
-%
-\begin{align}
-	b^{(i)} &\sim p \Big( b | A, X \Big)  \label{eqn:b-samples}\\
-	\theta^{(i)} &\sim p\Big(\theta | X, b^{(i)} \Big) \label{eqn:theta-samples}
-\end{align}
-%
-Both of these sampling steps implemented with a Markov Chain through the Metropolis-Hastings algorithm \cite{hastings-alg}.
-We just need to define a proposal distribution $q(x, x')$ for proposing a move $x \rightarrow x'$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(\cdot)$, point-wise.
-The proposed move is then accepted with probability $\alpha$ (equation \ref{eqn:mh-accept}) else it is rejected and we stay at $x$.
-%
-\begin{equation}
-	\alpha = \min \left( \frac{\pi(x') q(x', x)}{\pi(x) q(x, x')} , 1 \right)
-	\label{eqn:mh-accept}
-\end{equation}
-%
-This accept-reject step ensures the resulting Markov Chain is in detailed balance with the target distribution $\pi(\cdot)$.
-What we propose in equations \ref{eqn:b-samples} and \ref{eqn:theta-samples} is therefore implemented through a 2-level Markov chain.
-The resulting samples for $\theta^{(i)}$ are unbiased in the sense that the expectation of their distribution is the posterior we are targeting in equation \ref{eqn:theta-target}.
-%
-\begin{align*}
-	\Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] &= \sum_{b \in \mathcal{B}^N} p(\theta | X, b) p(b | A, X) \\
-	&= \sum_{b \in \mathcal{B}^N} p(\theta, b | A, X) \\
-	&= p(\theta | A, X) \\
-\end{align*}
-%
-This is an example of a pseudo-marginal approach \cite{pseudo-marginal}.
-Indeed, the unbiased result is sufficient to prove that for sufficient samples, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{t})\right] = p(\theta
-\ A, X)$ which is exactly the distribution we were targetting (equation \ref{eqn:theta-target}).
-
-The reason we split the Markov chain into two stages is because the summation over all latent states $b \in \mathcal{B}^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in \Bcal^N} p(A | b) P(b | X, \theta)$ is intractable $O(B^N)$.
-%
-\begin{figure}[!h]
-	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=12mm},
-		squarednode/.style={rectangle, draw=black, minimum size=12mm}
-		]
-		% nodes
-		\node[roundnode] (b0) at (0, 2) {$b^{(t)}$};
-		\node[roundnode] (b1) at (4, 2) {$b'$};
-		\node[roundnode] (b2) at (8, 2) {$b^{(t+1)}$};
-		\node[roundnode] (t0) at (0, 0) {$\theta^{(t)}$};
-		\node[roundnode] (t1) at (4, 0) {$\theta'$};
-		\node[roundnode] (t2) at (8, 0) {$\theta^{(t+1)}$};
-
-		% arrows
-		\draw[->] (b0) to node[above] {$\sim q_b(b^{(t)}, b' | A)$} (b1);
-		\draw[->] (b1) to node[above] {$\alpha_b (b, b' | A)$} (b2);
-		\draw[->] (b0) [out=30, in=150] to node[above] {$1-\alpha_b$} (b2);
-
-		\draw[->] (t0) to node[above] {$\sim q_\theta(\theta, \theta' | X, b^{(t)})$} (t1);
-		\draw[->] (t1) to node[above] {$\alpha_\theta (\theta, \theta' | X, b^{(t)})$} (t2);
-		\draw[->] (t0) [out=-30, in=-150] to node[below] {$1-\alpha_\theta$} (t2);
-
-		\draw[dashed, ->] (b0) to (2, 0.8);
-		\draw[dashed, ->] (b0) to (5.5, 0.8);
-
-	\end{tikzpicture}
-	\caption{Sampling sequence}
-	\label{fig:samp-sequence}
-\end{figure}
-%
-Figure \ref{fig:samp-sequence} shows an overview of the proposed method.
-We have introduced subscripts and conditionings to make explicit what parameters each step utilises.
-In an important simplification, we note that $p(b| X) = B^{-N}$ which does not depend on the exact value of X.
-Therefore, we do not need to know the value of $X$ to perform the sampling on $b$.
-Conversely, for the $\theta^{(t)}$ samples, we use $b^{(t)}$ but not $A$ as $(\theta \indep A) | b$.
-
-\subsection{Sampling block memberships}
-
-\citet{Peixoto-MCMC} proposes a Monte Carlo method which we will base our approach on. It relies on writing the posterior in the following form:
-%
-\begin{equation}
-	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) = \pi_b(b)
-\end{equation}
-%
-Now $\pi_b(\cdot)$ is the un-normalised density we wish to sample from.
-In other words, we wish to construct a Markov chain that has $\pi_b(\cdot)$ as its invariant distribution.
-We can break $\pi_b$ down as follows:
-%
-\begin{align*}
-	\pi_b(b) &= p(b|X) \sum_{\psi} \nolimits p(A , \psi | b, X) \\
-	&= p(b|X) p(A, \psi^* | b, X) \\
-	&= p(A | b, \psi^*) \cdot p(\psi^* | b) \cdot p(b | X)
-\end{align*}
-%
-Since we are using the microcanonical SBM formulation, there is only one value of $\psi$ that is compatible with the given $(A, b)$ pair.
-We denote this value $\psi^* = \{k^*, e^*\}$.
-Specifically, $k^*_i = \sum_j A_{ij} $ and $e^*_{rs} = \sum_{i, j} A_{ij} \one \{b_i=r\} \one\{b_j=s\}$.
-Therefore, the summation over all $\psi$ reduces to just the single $\psi^*$ term.
-We also define the microcanonical entropy of the configuration as.
-%
-\begin{equation}
-	S(b) = - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big)
-\end{equation}
-%
-This entropy can be thought of as the description length of the graph because it is the sum of the information required to represent the graph given the parameters and the amount of information required to store the parameters (given the feature matrix $X$).
-The exact, from of the proposal distribution and accept-reject step is explored throughly by \citet{Peixoto-MCMC}.
-There is a widely available library, \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, implementing this algorithm.
-The only modification we make is in the block membership prior $p(b)$ which we replace with $p(b|X)=B^{-N}$ which is a uniform distribution and so cancels out in the MH accept-reject step.
-
-The end result of this is that we can generate a set of block membership samples $\{b^{(t)}\}_{t=1}^{T}$ with each $b^{(t)} \sim p(b | A, X)$.
-Each of these samples can then be used for the $\theta$-chain.
-
-\subsection{Sampling feature-to-block classifier parameters}
-
-The invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. We write this as follows:
-%
-\begin{align}
-	p(\theta | X, b) &\propto p(b | X, \theta) p(\theta) = \pi_\theta (\theta) \propto  \exp \left( - U(\theta) \right) \\
-	\therefore U(\theta) &= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}
-\end{align}
-%
-Where we have introduced $U(\theta)$ equal to the negative log posterior.
-This $U(\theta)$ term makes subsequent analysis more concise.
-Each of the constituent terms of $U(\theta)$ is easily computed (equation \ref{eqn:u-form}).
-To simplify notation, we define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
-%
-\begin{equation}
-	\log p(b | X, \theta) = \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
-	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta ||^2
-	\label{eqn:U-constituent-terms}
-\end{equation}
-%
-This means that if we ignore constant terms we can write $U(\theta)$ as in equation \ref{eqn:U-form}.
-We see that the prior effectively introduces a regularisation term.
-Note that $||\theta||^2 = \sum_{i} \theta_{i}^2 = \sum_{j=1}^{B} ||w_j||^2$ is the Euclidean norm of the vector of parameters $\theta$.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form}
-\end{equation}
-%
-$U(\theta)$ in equation \ref{eqn:U-form} appears a typical objective function to be minimised for neural network training.
-The first term is the cross-entropy between actual and predicted labels.
-The second term - introduced by the prior - brings a form of regularisation, preventing over-fitting.
-In traditional applications we only seek the value of $\theta$ that minimises the objective function $U(\theta)$, which in our case would yield the maximum a posteriori (MAP) estimate.
-This is often done through some kind of gradient descent as $\nabla U$ is easily computable (equation \ref{eqn:U-derivative}).
-
-However, our goal is not to find the MAP estimate but to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
-As discussed earlier, given the invariant $\pi_\theta(\cdot)$, it is sufficient to specify a proposal distribution and then apply a MH accept-reject step to ensure detailed balance of the Markov Chain.
-Nevertheless, we can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
-We therefore adopt the Metropolis Adjusted Langevin Algorithm (MALA) - first proposed by \citet{mala-tweedie} - which leverages just that.
-Given the current sample $\theta$, we propose a new sample $\theta'$ according to equation \ref{eqn:theta-update}
-%
-\begin{equation}
-	\theta' = \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi
-	\label{eqn:theta-update}
-\end{equation}
-%
-Where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter - which may vary with the sample index.
-Indeed without the injected noise term, this is equivalent to gradient descent.
-We require the noise term to fully explore the parameter space.
-As such the proposal distribution, is a simple multivariate Gaussian which can be easily evaluated.
-%
-\begin{equation}
-	q_\theta(\theta, \theta') = \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right)
-\end{equation}
-%
-The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}). By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as:
-%
-\begin{equation}
-	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
-	\label{eqn:U-derivative}
-\end{equation}
-%
-After a proposed move is generated, in typical Metropolis-Hastings fashion we accept the move with probability $\alpha_\theta$.
-%
-\begin{equation}
-	\alpha_\theta(\theta, \theta') = \min \left( 
-	\exp \left( U(\theta) - U(\theta')\right)
-	\frac{ 
-		q_\theta(\theta', \theta)
-	}{
-		q_\theta(\theta, \theta')
-	} 
-	, 1 \right)
-\end{equation}
-%
-This fully specifies, the sampling procedure to generate $\{\theta^{(t)}\}_{t=1}^T$. Bear in mind that each $\theta^{(t)}$ update step uses its corresponding $b^{(t)}$ block membership sample.
-
-\subsection{Sampling sequence}
-
-So far, each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample.
-This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance.
-This may lead to longer burn-in and autocorrelation times of the resulting Markov Chain.
-The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends on $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
-We would rather deal with the expectation of each $y_{ij}^{(t)}$:
-%
-\begin{equation}
-	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one(b_{i}^{(t)} = j) \right]
-	= p(b_i = j | A, X)
-\end{equation}
-%
-We obtain an unbiased estimate for this quantity as simply the empirical distribution of the block membership samples $\left\{ b^{(t)} \right\}_{t=1}^T$.
-%
-\begin{equation}
-	\hat{y}_{ij} \coloneqq \frac{1}{T} \sum_{t=1}^{T} y_{ij}^{(t)} = \frac{1}{T} \sum_{t=1}^{T} \one\{b_i^{(t)} = j\}
-\end{equation}
-%
-We therefore, choose to feed each $\theta^{(t)}$ update step the same $\hat{y}_{ij}$ for all $t$ rather than the corresponding $y^{(t)}_{ij}$.
-This means we no longer need to run the $b$ and $\theta$ Markov chains concurrently.
-Instead, we run the $b$-chain to completion and use it to generate $\hat{y}_{ij}$ for $i \in \{1 \dots N\}$ and $j \in \{1 \dots B\}$.
-This is an estimate of $p(b | A, X)$ that we use for every iteration of the $\theta$ Markov chain.
-This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$; wed refer to these as $T_b$ and $T_\theta$ henceforth.
-Furthermore, this changeover reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{Block membership sample generation} % give the algorithm a caption
-	\label{alg:b-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\FOR{$t \in \{0, 1 \dots T_b - 1\}$}
-			\STATE $b' \gets \sim q_b(b^{(t)}, b' | A)$
-			\STATE $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
-			\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-			\IF{$\log \eta < \log \alpha_b$}
-				\STATE $b^{(t+1)} \gets b'$
-			\ELSE
-				\STATE $b^{(t+1)} \gets b^{(t)}$
-			\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{LBGM parameter pseudo-marginal inference} % give the algorithm a caption
-	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\STATE $\mathcal{S} \gets \textrm{range}(\textrm{burn-in}, T_b, \textrm{thinning})$
-		\STATE $\hat{Y}_{ij} \gets \frac{1}{\mathcal{|S|}} \sum_{t \in \mathcal{S}} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
-
-		\item[]
-
-		\FOR{$t \in \{0, 1 \dots T_\theta - 1\}$}
-		\STATE $\xi \gets \sim \Gaussian(0, I)$
-		\STATE $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
-		\STATE $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
-		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-		\IF{$\log \eta < \log \alpha_\theta$}
-		\STATE $\theta^{(t+1)} \gets \theta'$
-		\ELSE
-		\STATE $\theta^{(t+1)} \gets \theta^{(t)}$
-		\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\subsection{Complexity Analysis}
-
-\section{Experiments}
-
-We apply the developed methods to a variety of datasets.
-
-\begin{itemize}
-	\item \textbf{College football teams} \cite{Evans_2010_football} - network of American college football teams and their interactions. Vertex labels are the division a team belongs to.
-	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.
-	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier} - network of the author's Facebook friends list.
-Each vertex has been manually labelled with a variety of features describing their relationship to the author.
-	\item \textbf{Twitch users} \cite{twitch} - a network of user-user friendships on the streaming service Twitch.
-Vertex labels are extracted according to video-games played, location and streaming habits.
-This dataset is also broken down into disjoint networks according to language.
-We only consider the English users with is a subnet with $N=7126$ vertices and $E=35324$ edges).
-\end{itemize}
-
-\section{Extensions}
-
-\section{Conclusion}
-
-\begin{ack}
-Use unnumbered first level headings for the acknowledgments. All acknowledgments
-go at the end of the paper before the list of references. Moreover, you are required to declare
-funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
-More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2021/PaperInformation/FundingDisclosure}.
-
-Do {\bf not} include this section in the anonymized submission, only in the final paper.
-You can use the \texttt{ack} environment provided in the style file to automatically hide this section in the anonymized submission.
-\end{ack}
+\input{ack}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Checklist}
-
-%%% BEGIN INSTRUCTIONS %%%
-The checklist follows the references.  Please
-read the checklist guidelines carefully for information on how to answer these
-questions.  For each question, change the default \answerTODO{} to \answerYes{},
-\answerNo{}, or \answerNA{}.  You are strongly encouraged to include a {\bf
-justification to your answer}, either by referencing the appropriate section of
-your paper or providing a brief inline description.  For example:
-\begin{itemize}
-  \item Did you include the license to the code and datasets? \answerYes{See Section gen-inst.}
-  \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
-  \item Did you include the license to the code and datasets? \answerNA{}
-\end{itemize}
-Please do not modify the questions and only use the provided macros for your
-answers.  Note that the Checklist section does not count towards the page
-limit.  In your paper, please delete this instructions block and only keep the
-Checklist section heading above along with the questions/answers below.
-%%% END INSTRUCTIONS %%%
-
-\begin{enumerate}
-
-\item For all authors...
-\begin{enumerate}
-  \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
-    \answerTODO{}
-  \item Did you describe the limitations of your work?
-    \answerTODO{}
-  \item Did you discuss any potential negative societal impacts of your work?
-    \answerTODO{}
-  \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you are including theoretical results...
-\begin{enumerate}
-  \item Did you state the full set of assumptions of all theoretical results?
-    \answerTODO{}
-	\item Did you include complete proofs of all theoretical results?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you ran experiments...
-\begin{enumerate}
-  \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
-    \answerTODO{}
-  \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
-    \answerTODO{}
-	\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
-    \answerTODO{}
-	\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
-\begin{enumerate}
-  \item If your work uses existing assets, did you cite the creators?
-    \answerTODO{}
-  \item Did you mention the license of the assets?
-    \answerTODO{}
-  \item Did you include any new assets either in the supplemental material or as a URL?
-    \answerTODO{}
-  \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
-    \answerTODO{}
-  \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
-    \answerTODO{}
-\end{enumerate}
-
-\item If you used crowdsourcing or conducted research with human subjects...
-\begin{enumerate}
-  \item Did you include the full text of instructions given to participants and screenshots, if applicable?
-    \answerTODO{}
-  \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
-    \answerTODO{}
-  \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
-    \answerTODO{}
-\end{enumerate}
-
-\end{enumerate}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\include{checklist}
 
@@
 
-\section{Appendix}
-
-\subsection{Derivation of conditional block distribution given feature matrix}
-\label{appdx:b|x}
-
-We wish to determine the form of $p(b| X)$. This can be done by integrating over the joint probability with respect to $\theta$.
-%
-\begin{align*}
-	p(b | X) &= \int p(b , \theta| X, \theta) d\theta = \int p(b | X, \theta) p(\theta | X) d\theta \\
-	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i=1}^{N} \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
-	&= \prod_{i=1}^{N} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j=1}^{B} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k=1}^{B} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}
-\end{align*}
-%
-We note that $b_i \in {1, 2, \dots B}$ and so the integral's value is unchanged with respect to $b_i$.
-The integrand has the same form no matter which value $b_i$ takes as the prior is the same for each $w_j$.
-As such the integral can only be a function of at most $\tilde{x}_i$ and $\sigma_\theta^2$ as it is symmetric with respect to $b_i$ and all the various $w_j$ are integrated out as they are dummy variables.
-Therefore, denoting the integral by the (unknown) function $f(\tilde{x}_i, \sigma_\theta^2)$, we write $p(b| X)$ as follows:
-%
-\begin{align*}
-	p(b | X) &= \prod_{i=1}^{N} f(\tilde{x}_i, \sigma_\theta^2) = \textrm{const w.r.t } b = c
-\end{align*}
-%
-As this is a constant with respect to $b$ we conclude that $p(b | X)$ must be a uniform distribution.
-$\nicefrac{1}{c}$ is simply the size of the set of values that $b$ can take.
-We know $b_i \in \mathcal{B} = \{1, 2, \dots B\}$.
-Therefore, $b \in \mathcal{B}^N$ and $|\mathcal{B}^N| = |\mathcal{B}|^N = B^N= \nicefrac{1}{c}$.
-Putting this all together we show that:
-%
-\begin{equation}
-	p(b | X) = B^{-N}
-\end{equation}
-
-\subsection{Derivation of U gradient with respect to feature parameters}
-\label{appdx:gradu}
-The goal is to determine $\nabla U(\theta)$, the gradient of the negative log posterior with respect to the parameters. We repeat the form of $U(\theta)$ in equation \ref{eqn:U-form-appdx}.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form-appdx}
-\end{equation}
-%
-Where $y_{ij}$ is independent of $\theta$ and $a_{ij}$ is the output from the softmax layer, with form as given in equation \ref{eqn:a-ij}.
-%
-\begin{equation}
-	a_{ij} \coloneqq \phi_{j} (x_i; \theta) = \frac{\exp(w_j^T \tilde{x}_i)}{\sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i)}
-	\label{eqn:a-ij} 
-\end{equation}
-%
-We note that $\theta = \{w_k\}_{k=1}^B$, and as such we can write this in vector form $\theta = \left[w_1^T, w_2^T \dots w_B^T  \right]^T$.
-Therefore, $\nabla U(\theta) = \left[\nicefrac{\partial U}{\partial w_1}^T,\nicefrac{\partial U}{\partial w_2}^T \dots \nicefrac{\partial U}{\partial w_B}^T  \right]^T$; to compute $\nabla U(\theta)$ it suffices to find the form of $\nicefrac{\partial U}{\partial w_k}$ with respect to a general $k$.
-
-To this end, we must first find partial derivatives of $a_{ij}$ and $||\theta||$ with respect to $w_k$. Starting with $a_{ij}$:
-%
-\begin{align}
-	\frac{\partial a_{ij}}{\partial w_k} &= \frac
-	{\tilde{x}_i \exp(w_j^T \tilde{x}_i) \delta_{jk} \cdot \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) 
-		- 
-	\exp(w_j^T \tilde{x}_i) \cdot \tilde{x}_i \exp(w_k^T \tilde{x}_i)}
-	{\left( \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) \right)^2} \nonumber \\
-	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right) 
-\end{align}
-%
-Where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$. Now moving onto the derivative of $||\theta||^2$:
-%
-\begin{equation}
-	\frac{ \partial}{\partial w_k} ||\theta||^2 = \frac{\partial}{\partial w_k} \left( \sum_{b=1}^B ||w_b||^2 \right) = 2w_k
-\end{equation}
-%
-We are ready to put this all together, to find the partial derivative of $U(\theta)$ with respect to each $w_k$:
-\begin{align}
-	\frac{\partial U}{\partial w_k} &= 
-	\sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} 
-	\left( \frac{-\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
-	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
-	&=  - \left( \sum_{i=1}^{N} \tilde{x}_i \left( y_{ik} - a_{ik} \sum_{j=1}^{B} y_{ij} \right)
-	- \frac{w_k}{\sigma_\theta^2} \right) \nonumber \\
-	&= - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
-\end{align}
-%
-This is the required result.
-This form can be computed efficiently through matrix operations.
-The only property of $y_{ij}$ we have used in the derivation is the sum-to-one constraint $\sum_{j=1}^{B} y_{ij} = 1$ for all $i$.
-
-\subsection{Choosing the MALA step-size}
-\label{appdx:step-size}
-
-For sampling from the $\theta$-chain of the block membership generator parameters, we employed the Metropolis Adjusted Langevin Algorithm (MALA). At iteration $t$, the proposed sample is generated by:
-%
-\begin{equation}
-	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi
-\end{equation}
-%
-There are two competing objectives when choosing the step-size $h_t$.
-On the one hand, we want the step-size to be large so that we arrive at a high density region quickly.
-However, too large a step-size will lead to a lower acceptance ratio and thus inefficient sampling.
-A solution to this problem would be to slowly decrease the step-size with $t$ - often called simulated annealing.
-Therefore, we still have a short burn-in time but will not bounce around the mode for large $t$.
-As well as the trivial constraint for $h_t$ to be strictly positive and we introduce two further constraints as outlined by \citet{Bayesian-SGLD}:
-%
-\begin{equation}
-	\sum_{t=1}^{\infty} h_t = \infty \qquad \textrm{and} \qquad
-	\sum_{t=1}^{\infty} h_t^2 < \infty
-	\label{eqn:h-constraints}
-\end{equation}
-%
-The first constraint ensures that we have cover sufficient distance to arrive at any arbitrary point in our domain, no matter the starting point.
-The second constraint ensures that once we converge to the mode rather than simply bouncing around it.
-\citet{Bayesian-SGLD} propose the following form for a polynomially decaying step-size which we adopt:
-%
-\begin{equation}
-	h_t = a(b + t)^{-\gamma}
-\end{equation}
-%
-Where $a, b, \gamma$ are hyper-parameters to be chosen.
-We require $a,b > 0$ and $\gamma \in (0.5, 1]$ to satisfy equation \ref{eqn:h-constraints}.
-We find empirically that $a=250, b=1000, \gamma=0.8$ yield good burn-in times for our MALA implementations.
+\include{appendices}
 

### 2021-05-16 nips/abstract.tex
Abstract

\begin{abstract}
	Labelled networks are an extremely common and important form of data. Often, we wish to determine how the vertex labels (referred to as features) affect graphical structure. The standard approach has been to partition the network into blocks grouped by distinct values of the feature of interest. We then use a block-based random graph model to test for evidence these extracted feature-based communities interact differently with one another.

	Nevertheless, these feature-based communities are often not a natural partition of the graph and thus the model is not a good fit. With this in mind, we present a novel generative model for better describing vertex-labelled undirected graphs. This is a two-stage model. First, we start with vertex features and probabilistically generate a block membership. Secondly, given these latent block memberships, we draw a graph according to the micro-canonical stochastic block model. We call this model the latent block generative model (LBGM).

	Armed with the newly defined LBGM, we present a method to efficiently sample the parameters of the feature to block generator. This allows us to automatically determine which features best explain the block-based graphical structure. Importantly, this analysis can be performed using the whole feature-set rather than considering each independently.

	We apply the developed methods to a variety of network data to show the derived results are meaningful and match our expectations. However, the vertex features available often do not readily explain the community structure in the resulting network. The approach is able to detect this and thus is protected against over-confidence. This is usually a failure of the data collection to capture all vertex features relevant to the resulting structure.

	Future work may benefit from extending the LBGM to multiple hierarchical levels. Currently, the single-level model only provides explanations for the widest separation between communities. Nevertheless, in real-world data each community may be composed of many sub-communities and so on. A hierarchical approach would be able to find explanations for network structure at each level of the hierarchy. 
\end{abstract}

### 2021-05-16 nips/submission.tex
Abstract

@@
 
-\begin{abstract}
-We present a novel generative model for describing vertex-labelled undirected graphs.
-This is a two-layered model.
-First a probabilistic mapping from vertex labels to latent block memberships and from these block memberships we generate a graph according to the micro-canonical stochastic block model.
-
-With this framework we are able to efficiently sample the parameters of the vertex label to block membership mapping.
-This allows us to automatically determine which features have the largest impact on graphical structure.
-\end{abstract}
+\input{abstract}
 
@@
 
-\include{checklist}
+\input{checklist}
 
@@
 
-\include{appendices}
+\input{appendices}
 

### 2021-05-16 nips/abstract.tex
Work on intro

@@
 Secondly, given these latent block memberships, we draw a graph according to the micro-canonical stochastic block model.
-We call this model the latent block generative model (LBGM).
+We call this model the feature-first block model (FFBM).
 
-	Armed with the newly defined LBGM, we present a method to efficiently sample the parameters of the feature to block generator.
+	Armed with the newly defined FFBM, we present a method to efficiently sample the parameters of the feature to block generator.
 This allows us to automatically determine which features best explain the block-based graphical structure.
@@
 
-	Future work may benefit from extending the LBGM to multiple hierarchical levels.
+	Future work may benefit from extending the FFBM to multiple hierarchical levels.
 Currently, the single-level model only provides explanations for the widest separation between communities.

### 2021-05-16 nips/inference.tex
Work on intro

@@
 
-Now that we have defined the LBGM, we wish to leverage it to perform inference.
+Now that we have defined the FFBM, we wish to leverage it to perform inference.
 Suppose we are presented with a vertex-labelled graph $(A, X)$; the goal is to draw samples for $\theta$ according to the posterior given the observed graph (equation \ref{eqn:theta-target}).
@@
 \begin{algorithm} % enter the algorithm environment
-	\caption{LBGM parameter pseudo-marginal inference} % give the algorithm a caption
+	\caption{FFBM parameter pseudo-marginal inference} % give the algorithm a caption
 	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
@@
 
-For the purposes of our model (the LBGM), the number of blocks $B$ is a constant which must be specified by the data scientist.
+For the purposes of our model (the FFBM), the number of blocks $B$ is a constant which must be specified by the data scientist.
 We could however, allow our choice of $B$ to be influenced by the observed data.

### 2021-05-16 nips/introduction.tex
Work on intro

@@
 
-There is a wealth of networks in the world.
-Some examples of this graphical data are social networks, website hyperlinks and academic collaboration with more being produced each second.
-It is clear we need tools to analyse this increasingly omni-present form of data.
+A somewhat surprising property of many real-world networks is that they exhibit strong community structure.
+In other words, each node will often belong to a cluster of densely connected nodes.
+There is high interest in recovering the latent communities from the observed graphs.
+The inferred communities can be exploited for compression algorithms \cite{cluster-compression} or used for link prediction in incomplete networks \cite{link-prediction} to name but a few applications.
 
-A somewhat surprising property of real-world graphs is that they exhibit strong community structure.
-In other words, each node will often belong to a cluster of densely connected nodes.
-This property is often exploited by graph compression algorithms and there is high interest in recovering the communities from the observed graph.
+Nevertheless, we restrict our analysis to labelled networks.
+These are graphs where we have information about the properties of each node.
+We shall refer to these node properties as features.
+One of the most common questions we can ask of labelled networks is whether a given vertex feature has an impact on structure of the graph.
+To answer this question from a Bayesian perspective we must use a random graph model; the most common form is called the stochastic block model (SBM) \cite{vanilla-sbm}.
+This is a latent variable model where each vertex belongs to a single block and the probability two nodes are connected depends only on the.
+There have been many variants to this model; the most popular are the mixed-membership stochastic block model (MMSBM) \cite{mixed-membership-sbm} and overlapping stochastic block model (OSBM) \cite{overlapping-sbm}.
+Effectively, these just extend the model to allow each vertex to belong to multiple blocks simultaneously.
 
-A common subset of graphical data is the labelled network.
-This is a graph where we have information about the properties of each node.
-We shall refer to these node properties as features.
-One of the most common questions we can ask of this dataset is what features have the largest impact on the structure of the graph.
-For example, when analysing an academic collaboration graph, one may wish to ask what impact gender has on the structure of the network.
-Nevertheless, this analysis is often vulnerable to confounding variables.
-While gender very well may impact the structure of the graph, there is often a better explanatory variable for the structure.
+However, a major drawback of these graphical models as applied to labelled networks is that they do not automatically include vertex features in the random graph generation process.
+Approaches based on graph neural networks \cite{sbm-gnn} that utilise vertex features have been developed but these lack the easy interpretability of the simpler models.
+To analyse a labelled network using one of the simple SBM variants, a typical inference procedure would be to partition the graph into blocks grouped by distinct values of the feature of interest.
+The associated model can then be used to test for evidence of heterogeneous connectivity between the feature-grouped blocks.
+Nevertheless, this approach is limited in that it can only consider one feature at a time.
+Furthermore, it is hard to rank order the magnitude of the impact of each feature if they are considered independently.
+We would instead prefer to partition the graph into its most natural blocks and then determine which feature best predict the resulting partition.
 
-There is space to bridge the gap between these two approaches.
-Rather than determining whether a feature impacts the graphical structure directly, we can use the community concept as a stepping stone in our analysis.
+With these desiderata in mind, we present a novel model for modelling labelled networks, which we call the feature-first block model (FFBM).
+This allows us to use the latent block membership as a stepping stone in our analysis.
 Therefore, we extract which features have the largest impact on overall graphical structure.

### 2021-05-16 nips/abstract.tex
abstract

@@
 	Labelled networks are an extremely common and important form of data.
-Often, we wish to determine how the vertex labels (referred to as features) affect graphical structure.
-The standard approach has been to partition the network into blocks grouped by distinct values of the feature of interest.
-We then use a block-based random graph model to test for evidence these extracted feature-based communities interact differently with one another.
+A typical inference goal is to determine how the vertex labels (called features) affect graphical structure.
+The standard approach to this problem is to partition the network into blocks grouped by distinct values of the feature of interest.
+We then use a block-based random graph model - typically a variant of the stochastic block model (SBM) - to test for evidence these extracted feature-based communities interact differently with one another.
 
 	Nevertheless, these feature-based communities are often not a natural partition of the graph and thus the model is not a good fit.
-With this in mind, we present a novel generative model for better describing vertex-labelled undirected graphs.
-This is a two-stage model.
-First, we start with vertex features and probabilistically generate a block membership.
-Secondly, given these latent block memberships, we draw a graph according to the micro-canonical stochastic block model.
-We call this model the feature-first block model (FFBM).
+With this in mind, we present a novel generative model, which we call the feature-first block model (FFBM), for better describing vertex-labelled undirected graphs.
+We present a method to efficiently sample the FFBM parameters.
+This allows us to automatically determine which features best explain the block-based graphical structure.
+Importantly, this analysis can be performed using the whole feature-set rather than considering each feature independently.
 
-	Armed with the newly defined FFBM, we present a method to efficiently sample the parameters of the feature to block generator.
-This allows us to automatically determine which features best explain the block-based graphical structure.
-Importantly, this analysis can be performed using the whole feature-set rather than considering each independently.
-
-	We apply the developed methods to a variety of network data to show the derived results are meaningful and match our expectations.
-However, the vertex features available often do not readily explain the community structure in the resulting network.
-The approach is able to detect this and thus is protected against over-confidence.
-This is usually a failure of the data collection to capture all vertex features relevant to the resulting structure.
-
+	We apply the developed method to a variety of network data to extract the most important features along which the vertices cluster.
+Those that do not impact the high-level structure can be discarded to reduce the problem dimension.
+In the case the vertex features available do not readily explain the community structure in the resulting network, the approach detects this and is protected against over-confidence.
 Future work may benefit from extending the FFBM to multiple hierarchical levels.
-Currently, the single-level model only provides explanations for the widest separation between communities.
-Nevertheless, in real-world data each community may be composed of many sub-communities and so on.
-A hierarchical approach would be able to find explanations for network structure at each level of the hierarchy.
+This allows the structure to be explained at each level of coarseness.
 \end{abstract}

### 2021-05-16 nips/abstract.tex
intro

@@
 A typical inference goal is to determine how the vertex labels (called features) affect graphical structure.
-The standard approach to this problem is to partition the network into blocks grouped by distinct values of the feature of interest.
-We then use a block-based random graph model - typically a variant of the stochastic block model (SBM) - to test for evidence these extracted feature-based communities interact differently with one another.
+The standard approach to this problem has been to partition the network into blocks grouped by distinct values of the feature of interest.
+A block-based random graph model - typically a variant of the stochastic block model (SBM) - is then used to test for evidence of asymmetric behaviour within these feature-based communities.
 
-	Nevertheless, these feature-based communities are often not a natural partition of the graph and thus the model is not a good fit.
+	Nevertheless, these feature-based communities are often not a natural partition of the graph and thus the models employed are rarely a good fit.
 With this in mind, we present a novel generative model, which we call the feature-first block model (FFBM), for better describing vertex-labelled undirected graphs.
-We present a method to efficiently sample the FFBM parameters.
-This allows us to automatically determine which features best explain the block-based graphical structure.
-Importantly, this analysis can be performed using the whole feature-set rather than considering each feature independently.
+This allows us to perform richer queries on labelled networks.
+We present a method to efficiently sample the FFBM parameters for inference.
+The FFBM's structure is kept deliberately simple to retain easy interpretability of the parameter values.
 
-	We apply the developed method to a variety of network data to extract the most important features along which the vertices cluster.
-Those that do not impact the high-level structure can be discarded to reduce the problem dimension.
+	We apply the developed methods to a variety of network data to extract the most important features along which the vertices divide themselves.
+The greatest advantage of the proposed approach is that the whole feature-space is used automatically and features can be rank-ordered implicitly according to impact.
+Any features that do not greatly impact the high-level structure can be discarded to reduce the problem dimension.
 In the case the vertex features available do not readily explain the community structure in the resulting network, the approach detects this and is protected against over-confidence.
 Future work may benefit from extending the FFBM to multiple hierarchical levels.
-This allows the structure to be explained at each level of coarseness.
+This would allow the structure to be explained at each level of coarseness rather than simply at the highest level.
 \end{abstract}

### 2021-05-16 nips/introduction.tex
intro

@@
 
-Nevertheless, we restrict our analysis to labelled networks.
-These are graphs where we have information about the properties of each node.
-We shall refer to these node properties as features.
-One of the most common questions we can ask of labelled networks is whether a given vertex feature has an impact on structure of the graph.
+We restrict our analysis to labelled networks.
+These are graphs where we additionally have information about the properties of each vertex.
+We shall refer to these vertex properties as features.
+One of the most common questions we can ask of labelled networks is whether a given vertex feature has an impact on the structure of the graph.
 To answer this question from a Bayesian perspective we must use a random graph model; the most common form is called the stochastic block model (SBM) \cite{vanilla-sbm}.
-This is a latent variable model where each vertex belongs to a single block and the probability two nodes are connected depends only on the.
-There have been many variants to this model; the most popular are the mixed-membership stochastic block model (MMSBM) \cite{mixed-membership-sbm} and overlapping stochastic block model (OSBM) \cite{overlapping-sbm}.
+This is a latent variable model where each vertex belongs to a single block and the probability two nodes are connected depends only on the block memberships of each.
+There have been many variants to this model; the most popular are the mixed-membership stochastic block model (MMSBM) \cite{mixed-membership-sbm} and the overlapping stochastic block model (OSBM) \cite{overlapping-sbm}.
 Effectively, these just extend the model to allow each vertex to belong to multiple blocks simultaneously.
@@
 Approaches based on graph neural networks \cite{sbm-gnn} that utilise vertex features have been developed but these lack the easy interpretability of the simpler models.
+
 To analyse a labelled network using one of the simple SBM variants, a typical inference procedure would be to partition the graph into blocks grouped by distinct values of the feature of interest.
 The associated model can then be used to test for evidence of heterogeneous connectivity between the feature-grouped blocks.
-Nevertheless, this approach is limited in that it can only consider one feature at a time.
-Furthermore, it is hard to rank order the magnitude of the impact of each feature if they are considered independently.
-We would instead prefer to partition the graph into its most natural blocks and then determine which feature best predict the resulting partition.
+Nevertheless, this approach is limited in that it can only consider one feature at a time; this means any conclusions drawn highly vulnerable to the presence of confounding variables.
+Furthermore, considering each feature separately makes it difficult to rank order the features by  magnitude of impact.
+Lastly, the feature-grouped blocks are often an unnatural partition of the graph, leading to a poor model fit.
+We would instead prefer to partition the graph into its most natural blocks and then determine which features best predict the resulting partition.
 
-With these desiderata in mind, we present a novel model for modelling labelled networks, which we call the feature-first block model (FFBM).
-This allows us to use the latent block membership as a stepping stone in our analysis.
-Therefore, we extract which features have the largest impact on overall graphical structure.
-This can be thought of as a form of dimensionality-reduction where we only keep the node features that are useful in predicting community memberships.
+With these desiderata in mind, we present a novel framework for modelling labelled networks, which we call the feature-first block model (FFBM).
+This can be thought of as an extension of the SBM to labelled networks.
+In the FFBM, we use the features first to generate the latent block membership for each vertex.
+The latent block membership is therefore a stepping stone rather than a starting point in our analysis.
+We go on to present an efficient algorithm for sampling from the parameters of the feature-to-block generator.
+We can interpret the sampled FFBM parameters to determine which features have the largest impact on overall graphical structure.
+Any features found to be irrelevant can be discarded as a form of dimensionality reduction.

### 2021-05-19 nips/experiments.tex
Improve legend
More work

@@
 	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.
-	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier} - network of the author's Facebook friends list.
+	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier}  ($N=349, E=2336, D=32$) - egonet of the author's Facebook friends list.
 Each vertex has been manually labelled with a variety of features describing their relationship to the author.
+For our purposes we remove all nodes of degree 1 (those that are only connected to the egonode) as these cannot be said to be part of any community present in the graph.
 	\item \textbf{Twitch users} \cite{twitch} - a network of user-user friendships on the streaming service Twitch.
@@
 
-\begin{table}[]
+\begin{table}[!h]
 	\centering
+	\caption{FFBM fit for various datasets}
+	\label{tab:results}
 	\begin{tabular}{c|ccc|c|cc}
@@
 		\includegraphics[width=\linewidth]{school-graph.png}
-		\caption{Softmax weighting samples for each block index}
+		\fbox{\includegraphics[width=0.5\linewidth]{school-legend.png}}
+		\caption{Graph coloured by inferred block memberships of each node}
 		\label{fig:school-graph}
@@
 		\includegraphics[width=\linewidth]{school-null.png}
-		\caption{}
+		\caption{Feature-to-block generator sampled weight parameters for each block index}
 		\label{fig:school-null}

### 2021-05-19 nips/experiments.tex
More work

@@
 We only consider the English users with is a subnet with $N=7126$ vertices and $E=35324$ edges).
+	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=1$) - network of Amazon book sales about U.S.
+politics, published close to the presidential election in 2004.
+Two books are connected if they were frequently co-purchased on Amazon.
+Vertex features encode the broad political affiliation of the author (liberal, conservative, neutral).
 \end{itemize}
 
-A table summarising the results is given in table \ref{tab:results}.
+We require metrics to assess the relative model evidence for the FFBM. . A table summarising the results is given in table \ref{tab:results}.
 
@@
 		Dataset & $N$ & $E$ & $D$ & $B$ & $\bar{S}_b /N$ & $\bar{U}_\theta /N$ \\ \hline
+		Political books & 105  & 441 & 1 & 3 & 11.7 & 0.707 \\
 		Primary school & 238 & 5539 & 13 &  18 & 43.0  & 2.75 \\
 		Maier FB egonet &  &     &     &   &           &             \\
-		&     &     &     &              &     &       
+
 	\end{tabular}

### 2021-05-19 nips/appendices.tex
All graphs loaded in

@@
 We find empirically that $a=250, b=1000, \gamma=0.8$ yield good burn-in times for our MALA implementations.
+
+\subsection{Algorithms}
+
+\begin{algorithm} % enter the algorithm environment
+	\caption{Block membership sample generation} % give the algorithm a caption
+	\label{alg:b-samples} % and a label for \ref{} commands later in the document
+	\begin{algorithmic} % enter the algorithmic environment
+
+		\FOR{$t \in \{0, 1 \dots T_b - 1\}$}
+		\STATE $b' \gets \sim q_b(b^{(t)}, b' | A)$
+		\STATE $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
+		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
+		\IF{$\log \eta < \log \alpha_b$}
+		\STATE $b^{(t+1)} \gets b'$
+		\ELSE
+		\STATE $b^{(t+1)} \gets b^{(t)}$
+		\ENDIF
+		\ENDFOR
+
+		\STATE \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$
+
+	\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm} % enter the algorithm environment
+	\caption{FFBM parameter pseudo-marginal inference} % give the algorithm a caption
+	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
+	\begin{algorithmic} % enter the algorithmic environment
+
+		\STATE $\hat{Y}_{ij} \gets \frac{1}{\mathcal{|S|}} \sum_{t \in \mathcal{S}} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
+
+		\item[]
+
+		\FOR{$t \in \{0, 1 \dots T_\theta - 1\}$}
+		\STATE $\xi \gets \sim \Gaussian(0, I)$
+		\STATE $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
+		\STATE $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
+		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
+		\IF{$\log \eta < \log \alpha_\theta$}
+		\STATE $\theta^{(t+1)} \gets \theta'$
+		\ELSE
+		\STATE $\theta^{(t+1)} \gets \theta^{(t)}$
+		\ENDIF
+		\ENDFOR
+
+		\STATE \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$
+
+	\end{algorithmic}
+\end{algorithm}

### 2021-05-19 nips/experiments.tex
All graphs loaded in

@@
 \begin{itemize}
+	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=1$) - network of Amazon book sales about U.S.
+politics, published close to the presidential election in 2004.
+Two books are connected if they were frequently co-purchased by the same customer.
+Vertex features encode the broad political affiliation of the author (liberal, conservative, neutral).
+
 	\item \textbf{Primary school dynamic contacts} \cite{schools} ($N=238, E=5539, D=13$) - network of face-to-face contacts amongst students and teachers at a primary school in Lyon, France.
@@
 We choose to analyse just the second day.
-	\item \textbf{College football teams} \cite{Evans_2010_football} - network of American college football teams and their interactions. Vertex labels are the division a team belongs to.
-	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.
+
+	\item \textbf{Facebook egonet} \cite{fb-snap} ($N=747, E=30025, D=480$) - an assortment of Facebook egonets.
+These are networks of a particular user's friends list and all the connections within that.
+Vertex features are extracted from each user's profile and is fully anonymized.
+We focus on the egonet with id 1912.
+
 	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier}  ($N=349, E=2336, D=32$) - egonet of the author's Facebook friends list.
@@
 For our purposes we remove all nodes of degree 1 (those that are only connected to the egonode) as these cannot be said to be part of any community present in the graph.
+
+	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.
+
 	\item \textbf{Twitch users} \cite{twitch} - a network of user-user friendships on the streaming service Twitch.
@@
 We only consider the English users with is a subnet with $N=7126$ vertices and $E=35324$ edges).
-	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=1$) - network of Amazon book sales about U.S.
-politics, published close to the presidential election in 2004.
-Two books are connected if they were frequently co-purchased on Amazon.
-Vertex features encode the broad political affiliation of the author (liberal, conservative, neutral).
+
 \end{itemize}
@@
 \end{table}
+
+\subsection{Political books}
+
+This dataset was collected by \citet{polbooks}.
+We wish to determine whether political affiliation is a good predictor of the overall network structure.
+We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral).
+The inferred block memberships are given in figure \ref{fig:books-graph}.
+
+\begin{figure}[!h]
+	\centering
+	\begin{subfigure}{0.3\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{polbooks-graph.png}
+		\fbox{\includegraphics[width=0.4\linewidth]{3-legend.png}}
+		\caption{Graph coloured by inferred block memberships of each node}
+		\label{fig:books-graph}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}{0.5\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{polbooks-null.png}
+		\caption{Feature-to-block generator sampled weight parameters for each block index}
+		\label{fig:book-null}
+	\end{subfigure}
+	\caption{Political books network}
+\end{figure}
 
@@
 	\end{subfigure}
-	\caption{Primary school dynamic contacts network \cite{schools}}
+	\caption{Primary school dynamic contacts network}
 \end{figure}
+
+\subsection{Facebook egonet}
+
+\begin{figure}[!h]
+	\centering
+	\begin{subfigure}{0.45\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{fb-graph.png}
+		\fbox{\includegraphics[width=0.3\linewidth]{10-legend.png}}
+		\caption{Graph coloured by inferred block memberships of each node}
+		\label{fig:fb-graph}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}{0.45\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{fb-null.png}
+		\caption{Feature-to-block generator sampled weight parameters for each block index}
+		\label{fig:fb-null}
+	\end{subfigure}
+	\caption{Primary school dynamic contacts network}
+\end{figure}

### 2021-05-19 nips/inference.tex
intro
All graphs loaded in

@@
 This is an estimate of $p(b | A, X)$ that we use for every iteration of the $\theta$ Markov chain.
-This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$; wed refer to these as $T_b$ and $T_\theta$ henceforth.
+This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$; we refer to these as $T_b$ and $T_\theta$ henceforth.
 Furthermore, this changeover reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{Block membership sample generation} % give the algorithm a caption
-	\label{alg:b-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\FOR{$t \in \{0, 1 \dots T_b - 1\}$}
-		\STATE $b' \gets \sim q_b(b^{(t)}, b' | A)$
-		\STATE $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
-		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-		\IF{$\log \eta < \log \alpha_b$}
-		\STATE $b^{(t+1)} \gets b'$
-		\ELSE
-		\STATE $b^{(t+1)} \gets b^{(t)}$
-		\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{FFBM parameter pseudo-marginal inference} % give the algorithm a caption
-	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\STATE $\mathcal{S} \gets \textrm{range}(\textrm{burn-in}, T_b, \textrm{thinning})$
-		\STATE $\hat{Y}_{ij} \gets \frac{1}{\mathcal{|S|}} \sum_{t \in \mathcal{S}} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
-
-		\item[]
-
-		\FOR{$t \in \{0, 1 \dots T_\theta - 1\}$}
-		\STATE $\xi \gets \sim \Gaussian(0, I)$
-		\STATE $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
-		\STATE $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
-		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-		\IF{$\log \eta < \log \alpha_\theta$}
-		\STATE $\theta^{(t+1)} \gets \theta'$
-		\ELSE
-		\STATE $\theta^{(t+1)} \gets \theta^{(t)}$
-		\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\subsection{Complexity Analysis}
 

### 2021-05-19 nips/conclusion.tex
Much work

\section{Conclusion}

The FFBM was developed to address the shortcomings of other graphical models when trying to test how vertex features affect graphical. The idea is to divide the graph into its most natural partition and only then bring in the vertex features to determine which ones best predict the partition. The inference algorithm proposed sidesteps an expensive summation over all latent block-states through the use of the microcanonical rather than the traditional SBM

The overall method is shown to be effective at extracting and describing the most natural communities in a labelled network. Nevertheless, the approach can only currently explain the structure on the macro-scale. Future work may benefit from extending the FFBM to be hierarchical in nature. That way, the structure of the network can be explained at each length-scale.

### 2021-05-19 nips/experiments.tex
Much work

@@
 We apply the developed methods to a variety of datasets, to show the power of the method across different scenarios.
-We choose a range of node counts $N$, edge counts $E$ and feature space dimension $D$.
+We choose datasets to span a range of node counts $N$, edge counts $E$ and feature space dimension $D$.
 
@@
 These are networks of a particular user's friends list and all the connections within that.
-Vertex features are extracted from each user's profile and is fully anonymized.
+Vertex features are extracted from each user's profile and are fully anonymized.
 We focus on the egonet with id 1912.
 
-	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier}  ($N=349, E=2336, D=32$) - egonet of the author's Facebook friends list.
+%	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier}  ($N=349, E=2336, D=32$) - egonet of the author's Facebook friends list.
 Each vertex has been manually labelled with a variety of features describing their relationship to the author.
 For our purposes we remove all nodes of degree 1 (those that are only connected to the egonode) as these cannot be said to be part of any community present in the graph.
-
-	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.
-
-	\item \textbf{Twitch users} \cite{twitch} - a network of user-user friendships on the streaming service Twitch.
+%		
+%	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.
+%	
+%	\item \textbf{Twitch users} \cite{twitch} - a network of user-user friendships on the streaming service Twitch.
 Vertex labels are extracted according to video-games played, location and streaming habits.
@@
 
-We require metrics to assess the relative model evidence for the FFBM. . A table summarising the results is given in table \ref{tab:results}.
+We require metrics to assess the relative model evidence for the FFBM.
+This can be split into two separate components: the microcanonical SBM fit (concerned with the $b$-samples) and the fit of the feature-to-block generator (concerned with the $\theta$-samples).
+For each of these we can evaluate the mean of the unnormalised log-target of each Markov chain.
+%
+\begin{equation}
+	\bar{S}_b \coloneqq \frac{1}{T_b} \sum_{t=1}^{T_b} S \left( b^{(t)} \right) \qquad \textrm{and} \qquad
+	\bar{U}_\theta \coloneqq \frac{1}{T_\theta} \sum_{t=1}^{T_\theta} U \left( \theta^{(t)} \right)
+\end{equation}
+%
+The lower these quantities, the better the fit of each stage of the model.
+To allow for rough comparison between datasets, we divide these quantities by the number of vertices in the graph $N$.
+Table \ref{tab:results} summarises the results for each experiment.
 
@@
 	\centering
-	\caption{FFBM fit for various datasets}
+	\caption{FFBM fit for the datasets}
 	\label{tab:results}
@@
 		Primary school & 238 & 5539 & 13 &  18 & 43.0  & 2.75 \\
-		Maier FB egonet &  &     &     &   &           &             \\
+		FB egonet & 747 & 30025 & 480 & 10  &           &             \\
 

### 2021-05-19 nips/inference.tex
Much work

@@
 Indeed, it is prudent to start our $b$-chain at this MDL solution as then the burn-in time is greatly reduced.
+
+\subsection{Dimensionality reduction}
+
+Once we have the samples $\left\{ \theta^{(t)} \right\}_{t=1}^{T_\theta} \sim p(\theta | A, X)$, we can compute the empirical mean and standard deviation of each component of $\theta$.
+Switching back to matrix notation we define $\theta = W$, such that $W_{ij}$ is the weight component for block $i$ and feature $j$, we can define:
+%
+\begin{equation}
+	\hat{\mu}_{ij} \coloneqq \frac{1}{T_\theta} \sum_{t=1}^{T_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
+	\hat{\sigma_{ij}} \coloneqq \frac{1}{T_\theta} \sum_{t=1}^{T_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2
+\end{equation}
+%
+A simple heuristic to discard the least important features requires specifying a cutoff $c > 0$ and a multiplier $k > 0$.
+We then discard all features $j$ wuch that equation \ref{eqn:discard} is satisfied.
+%
+\begin{equation}
+	(\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c) = \emptyset \quad \forall i \in \{1,2 \dots B\}
+	\label{eqn:discard}
+\end{equation}
+%
+Intuitively, this means discarding any feature for which $\hat{\mu}_{ij} \pm k\hat{\sigma_{ij}}$ lies within or spans the null region $(-c, c)$.
+If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \mu_{ij}, \sigma_{ij})$, then this is effectively a hypothesis test on the value of $W_{ij}$.
+%
+\begin{equation}
+	\begin{aligned}
+		H_0&: |W_{ij}| \leq c \\
+		H_1&: |W_{ij}| > c
+	\end{aligned}
+\end{equation}
+%
+With the multiplier $k$ specifying the degree of significance of the result (equivalent to the probability of a type-1 error).
+However, as the approximation is not exact we will only treat this dimensionality reduction method as a useful heuristic.

### 2021-05-19 nips/appendices.tex
Much work
First draft

@@
 \end{algorithm}
+
+\subsection{Inferring the number of blocks}
+
+For the purposes of our model (the FFBM), the number of blocks $B$ is a constant which must be specified by the data scientist.
+We could however, allow our choice of $B$ to be influenced by the observed data.
+This places us in the domain of empirical Bayes, which must be negotiated carefully.
+Prior beliefs must be determined a priori else they are not prior.
+However, as the number of blocks only specifies the coarseness of the analysis, it is fine to allow it to vary.
+Indeed, \citet{peixoto-determine-B} shows that for a fixed average degree the maximum number of detectable blocks scales as $O(\sqrt{N})$ where $N$ is the number of vertices.
+
+If we allow $B$ to vary in the $b$-chain (i.e.
+new blocks can be created and we permit empty blocks) then it can be run  until a minimum description length (MDL) solution is reached.
+We take the number of non-empty blocks at the MDL to be our fixed block number $B$ for subsequent analysis.
+Indeed, it is prudent to start our $b$-chain at this MDL solution as then the burn-in time is greatly reduced.
+
+\subsection{Implementation details}
+\label{sec:imp-details}
+
+All data analysis and visualisation was implemented in Python. Full source code is available at:
+
+\url{https://github.com/LozzaTray/Jormungandr}
+
+The scripts were run using a standard PC. Specs are:
+
+\begin{itemize}
+	\item \textbf{CPU}: Intel(R) Core(TM) i7-1065G7
+	\item \textbf{RAM}: 8GB
+	\item \textbf{GPU}: Intel(R) Iris(R) Plus Graphics
+\end{itemize}

### 2021-05-19 nips/conclusion.tex
First draft

@@
 
-The FFBM was developed to address the shortcomings of other graphical models when trying to test how vertex features affect graphical.
-The idea is to divide the graph into its most natural partition and only then bring in the vertex features to determine which ones best predict the partition.
-The inference algorithm proposed sidesteps an expensive summation over all latent block-states through the use of the microcanonical rather than the traditional SBM
+The FFBM was developed to address the shortcomings of other graphical models when testing how vertex features affect community structure.
+The idea is to divide the graph into its most natural partition and only then bring in the vertex features.
+We therefore hope to avoid a common pitfall in data science: correlation does not imply causation.
+It is very easy to find vertex features that are in some way correlated with the graphical structure.
+Nonetheless, only when we find the feature that best describes the most pronounced partition do we have a stronger case for causation.
+
+With the newly-defined FFBM, we go on to present an efficient inference algorithm to ascertain the parameters $\theta$ of the block generator.
+This is introduced as two concurrent Markov chains to sample the block memberships $b$ and block generator parameters $\theta$.
+The key trick to make the $b$-chain efficient is the the use of the microcanonical rather than the traditional SBM (as proposed by \citet{Peixoto-Bayesian-Microcanonical}) which allows us to bypass an expensive summation over the latent memberships $b$.
+Additionally, we use the empirical mean of the $b$-samples as the input to our $\theta$-chain, as this reduces the variance in our evaluation of the target distribution and thus shortens the burn-in and auto-correlation times of our $\theta$-chain.
 
 The overall method is shown to be effective at extracting and describing the most natural communities in a labelled network.
-Nevertheless, the approach can only currently explain the structure on the macro-scale.
-Future work may benefit from extending the FFBM to be hierarchical in nature.
-That way, the structure of the network can be explained at each length-scale.
+Nevertheless, the approach can only currently explain the structure at the macro-scale.
+We cannot explain structure within each community.
+Future work will benefit from extending the FFBM to be hierarchical in nature.
+That way, the structure of the network can be explained at all length-scales of interest.

### 2021-05-19 nips/experiments.tex
First draft

@@
 \begin{itemize}
-	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=1$) - network of Amazon book sales about U.S.
+	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=3$) - network of Amazon book sales about U.S.
 politics, published close to the presidential election in 2004.
 Two books are connected if they were frequently co-purchased by the same customer.
-Vertex features encode the broad political affiliation of the author (liberal, conservative, neutral).
+Vertex features encode the broad political affiliation of the author (liberal, conservative or neutral).
 
@@
 Two nodes are connected if the two parties shared a face-to-face interaction over the course of the day.
-Vertex features include class membership, gender and whether or not the student is a teacher or pupil.
+Vertex features include class membership, gender and whether or not the individual is a teacher or pupil.
 These data were collected on consecutive days in October 2009.
@@
 		Dataset & $N$ & $E$ & $D$ & $B$ & $\bar{S}_b /N$ & $\bar{U}_\theta /N$ \\ \hline
-		Political books & 105  & 441 & 1 & 3 & 11.7 & 0.707 \\
-		Primary school & 238 & 5539 & 13 &  18 & 43.0  & 2.75 \\
-		FB egonet & 747 & 30025 & 480 & 10  &           &             \\
-
+		Political books & 105  & 441 & 3 & 3 & 11.7 & 0.707 \\
+		Primary school & 238 & 5539 & 13 &  10 & 45.9  & 1.61 \\
+		FB egonet & 747 & 30025 & 480 & 10  & 66.9 & 7.47
 	\end{tabular}
@@
 
+\FloatBarrier
 \subsection{Political books}
@@
 The inferred block memberships are given in figure \ref{fig:books-graph}.
-
+We sample the block-generator parameters and plot the emprical mean and standard deviation of each sample on figure \ref{fig:book-null}.
+%
 \begin{figure}[!h]
@@
 		\fbox{\includegraphics[width=0.4\linewidth]{3-legend.png}}
-		\caption{Graph coloured by inferred block memberships of each node}
+		\caption{Sampled block memberships $\hat{Y}$}
 		\label{fig:books-graph}
@@
 		\includegraphics[width=\linewidth]{polbooks-null.png}
-		\caption{Feature-to-block generator sampled weight parameters for each block index}
+		\caption{Feature weights}
 		\label{fig:book-null}
 	\end{subfigure}
-	\caption{Political books network}
+	\caption{Political books network \cite{polbooks}}
 \end{figure}
+%
+Indeed, for all 3 blocks, each has a distinct political affiliation as its highest magnitude component.
+This is strong evidence that political affiliation is indeed the axis which best predicts the 3-way natural partition of the graph into blocks.
+Nevertheless, for block index 1, although \verb*|neutral| is the best predictor of the ones we have available, the value is not particularly extreme.
+Perhaps, there is an unobserved feature that places a book into this block.
+Indeed, block-0 is more of a centre-right block than true centre.
 
+\FloatBarrier
 \subsection{Primary school dynamic contacts}
 
-These data were originally collected by \citet{schools} to quantify the transmission opportunities for respiratory infections within a primary school context (ages 6-11 in France).
-However, we seek to ask a simpler question.
-What features best describe how people interact with one another in a primary school context.
+These data were originally collected by \citet{schools} to quantify the transmission opportunities for respiratory infections.
+However, we seek to ask which vertex features best describe how people interact with one another in a primary school context.
 The only vertex features we have available are school-class (one of 10 values - 2 per year group), gender and a distinction between teachers and pupils.
 
-We must first choose the number of blocks to consider $B$ to define the coarseness of our analysis.
-A total of 10 classes would suggest that $B=10$ is a natural starting point.
-With this value of $B$ we first sample from the block membership posterior $b^{(t)} \sim p(b | A, X)$.
-We visualise the inferred partition in figure \ref{fig:school-graph}
-
+We must first choose the number of blocks $B$ to define the coarseness of our analysis.
+A total of 10 school-classes would suggest that $B=10$ is a natural starting point.
+We visualise the inferred block memberships in figure \ref{fig:school-graph}.
+%
 \begin{figure}[!h]
@@
 		\includegraphics[width=\linewidth]{school-graph.png}
-		\fbox{\includegraphics[width=0.5\linewidth]{school-legend.png}}
-		\caption{Graph coloured by inferred block memberships of each node}
+		\fbox{\includegraphics[width=0.5\linewidth]{10-legend.png}}
+		\caption{Sampled block memberships $\hat{Y}$}
 		\label{fig:school-graph}
@@
 		\includegraphics[width=\linewidth]{school-null.png}
-		\caption{Feature-to-block generator sampled weight parameters for each block index}
+		\caption{Reduced dimension feature weights}
 		\label{fig:school-null}
 	\end{subfigure}
-	\caption{Primary school dynamic contacts network}
+	\caption{Primary school dynamic contacts network \cite{schools}}
 \end{figure}
+
+As before, we proceed to sample the block-generator parameters $\theta$ and employ the dimensionality reduction heuristic defined in section \ref{sec:dim-reduction} with null threshold $c=1$ and standard deviation multiplier $k=1$.
+We then plot the weights for the features that survive the discard step in figure \ref{sec:dim-reduction}.
+Immediately, we see that only the pupils' class memberships have survived (1A-5B); gender and teacher/student status have been discarded meaning that these are not good predictors of overall macro-structure.
+
+The vast majority of blocks are composed of a single class.
+However, some blocks have 2 comparably good classes as their predictor.
+For example, blocks 2 and 4 contain classes 4A and 4B as their 2 best predictors.
+This suggests that the social divide between classes is less pronounced for pupils in year 4 but there is nonetheless an unobserved variable that splits year 4 into 2 communities.
+The most surprising block is number 5 - which has comparable weightings for classes 5A and 1B.
+Perhaps there was a joint event between those two classes on the day the data were collected.
 
@@
 
+This dataset was chosen to showcase the power of the dimensionality reduction technique as the feature-space has high dimension ($D=480$).
+We sample the $b$-chain specifying $B=10$ total blocks and use this to construct the $\theta$-samples as before.
+
+We apply the dimensionality reduction heuristic outlined in section \ref{sec:dim-reduction}, choosing $k=1$ and $c=0.8$.
+This leads to a reduced dimension $D'=13$.
+These parameter values are plotted on figure \ref{fig:fb-null}.
+The features that remain are those that best explain the high-level community structure.
+More so than language, gender or surname - the school\footnote{This includes higher education institutes} you attend is the best explanatory variable for how the observed network's communities are partitioned at the highest level.
+%
 \begin{figure}[!h]
@@
 		\fbox{\includegraphics[width=0.3\linewidth]{10-legend.png}}
-		\caption{Graph coloured by inferred block memberships of each node}
+		\caption{Sampled block memberships $\hat{Y}$}
 		\label{fig:fb-graph}
@@
 		\includegraphics[width=\linewidth]{fb-null.png}
-		\caption{Feature-to-block generator sampled weight parameters for each block index}
+		\caption{Reduced dimension feature weights}
 		\label{fig:fb-null}
 	\end{subfigure}
-	\caption{Primary school dynamic contacts network}
+	\caption{Facebook egonet id 1912 \cite{fb-snap}}
 \end{figure}
+
+Nevertheless, this example also highlights a weakness of the method for high-dimensional feature spaces.
+When the feature dimension is very large, it becomes increasingly likely that a particular feature may uniquely identify a small set of nodes.
+If these nodes are all part of the same community then the classifier will overfit for that particular parameter.
+The regularisation term imposed by the prior goes some way to alleviating this problem but we see in figure \ref{fig:fb-null} that the feature \verb*|birthday-5| has a very high weight as it relates to block 0.
+It might be possible to shift the feature values such that they take values in $\{-1, 1\}$ rather than $\{0, 1\}$ but this approach falls down when input features are mutually correlated.
+Indeed, for features with discrete values such as the class memberships (1A, 1B, 2A \dots 5B) it is preferred to keep $x_i \in \{0, 1\}^D$ as then each block can be readily accept more than one mutually exclusive feature-group (i.e.
+two school-classes are part of the same block).

### 2021-05-19 nips/inference.tex
First draft

@@
 \section{Inference}
+\label{sec:inference}
 
@@
 
-\subsection{Choosing the number of blocks}
-
-For the purposes of our model (the FFBM), the number of blocks $B$ is a constant which must be specified by the data scientist.
-We could however, allow our choice of $B$ to be influenced by the observed data.
-This places us in the domain of empirical Bayes, which must be negotiated carefully.
-Prior beliefs must be determined a priori else they are not prior.
-However, as the number of blocks only specifies the coarseness of the analysis, it is fine to allow it to vary.
-Indeed, \citet{peixoto-determine-B} shows that for a fixed average degree the maximum number of detectable blocks scales as $O(\sqrt{N})$ where $N$ is the number of vertices.
-
-If we allow $B$ to vary in the $b$-chain (i.e.
-new blocks can be created and we permit empty blocks) then it can be run  until a minimum description length (MDL) solution is reached.
-We take the number of non-empty blocks at the MDL to be our fixed block number $B$ for subsequent analysis.
-Indeed, it is prudent to start our $b$-chain at this MDL solution as then the burn-in time is greatly reduced.
-
 \subsection{Dimensionality reduction}
+\label{sec:dim-reduction}
 
@@
 With the multiplier $k$ specifying the degree of significance of the result (equivalent to the probability of a type-1 error).
-However, as the approximation is not exact we will only treat this dimensionality reduction method as a useful heuristic.
+However, as the approximation is not exact we will only treat this dimensionality reduction method as a useful heuristic and not an exact method.

### 2021-05-19 nips/prelim.tex
First draft

@@
 The probability that two nodes are connected depends only on the block memberships of each.
-Graphs drawn from this ensemble exhibit community structure.
+Graphs drawn from the SBM ensemble exhibit community structure.
 Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}.
-A paraphrased definition is given below.
+A paraphrased definition is given below for the non-degree-corrected SBM (NDC-SBM).
 
@@
 Indeed, the NDC-SBM favours a partition into high-degree and low-degree nodes rather than clusters of inter-connected nodes.
-We therefore introduce the degree-corrected SBM (DC-SBM) to circumvent these issues.
+We therefore introduce the degree-corrected SBM (DC-SBM) \cite{Peixoto-Bayesian-Microcanonical} to circumvent these issues.
 

### 2021-05-20 nips/experiments.tex
More work on experiments

@@
 We choose datasets to span a range of node counts $N$, edge counts $E$ and feature space dimension $D$.
+We consider the following datasets:
 
@@
 This can be split into two separate components: the microcanonical SBM fit (concerned with the $b$-samples) and the fit of the feature-to-block generator (concerned with the $\theta$-samples).
-For each of these we can evaluate the mean of the unnormalised log-target of each Markov chain.
+The SBM has been extensively analysed in other works \cite{Peixoto-Bayesian-Microcanonical} so that will not be the emphasis of our analysis.
+We use a simple metric to gauge its fit: the mean description length per node of the $b$-samples (equation \ref{eqn:mean-dl}).
 %
 \begin{equation}
-	\bar{S}_b \coloneqq \frac{1}{T_b} \sum_{t=1}^{T_b} S \left( b^{(t)} \right) \qquad \textrm{and} \qquad
-	\bar{U}_\theta \coloneqq \frac{1}{T_\theta} \sum_{t=1}^{T_\theta} U \left( \theta^{(t)} \right)
+	\frac{1}{N} \bar{S}_b \coloneqq \frac{1}{N |\Tcal_b|} \sum_{t\in \Tcal_b} S \left( b^{(t)} \right)
+	\label{eqn:mean-dl}
 \end{equation}
 %
-The lower these quantities, the better the fit of each stage of the model.
-To allow for rough comparison between datasets, we divide these quantities by the number of vertices in the graph $N$.
+However, to assess the performance of the feature-to-block predictor, we must partition the vertex set $[N]$ into training and test sets -- $\Gcal_0$ and $\Gcal_1$ respectively\footnote{We choose to randomly create the partition on each experiment run such that 80\% of the available vertices form our training set $\Gcal$ and the remaining 20\% are held out to form our test set $\Gcal^C$}.
+The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal$ to train the $\theta$-chain.
+As $|\Gcal_0| \neq |\Gcal_1|$ sets, we cannot use the un-normalised log target $U$ (equation \ref{eqn:U-form}) for comparison between the two, as the prior term will be equal between the two but the total cross-entropy loss will be scaled by the size of each set.
+We therefore must use the average cross-entropy loss over each set (equation \ref{eqn:cross-entropy-loss}).
+%
+\begin{equation}
+	\bar{\Lcal}_a \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_a \left( \theta^{(t)} \right)
+	\qquad \textrm{where} \qquad
+	\Lcal_a \left( \theta^{(t)} \right) \coloneqq \frac{1}{|\Gcal_a|} \sum_{i \in \Gcal_a}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)}
+	\label{eqn:cross-entropy-loss}
+\end{equation}
+%
+Where $a \in \{0, 1\}$ has been introduced to toggle between training and test sets.
 Table \ref{tab:results} summarises the results for each experiment.
+We run each experiment 5 times with the same hyperparameter values\footnote{TODO hyperparam values} and report the mean plus or minus one standard deviation.
+
+However, we also wish to determine the power of the dimensionality reduction method.
+For this we discard all features that satisfy equation \ref{eqn:discard}, to reduce the dimension from $D$ to $D'$.
+We then retrain the feature-block predictor using just the retained features and report the loss over the training and test sets for the reduced classifier -- denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively.
+These are also
 
@@
 	\centering
-	\caption{FFBM fit for the datasets}
+	\caption{Table of results}
 	\label{tab:results}
-	\begin{tabular}{c|ccc|c|cc}
-		Dataset & $N$ & $E$ & $D$ & $B$ & $\bar{S}_b /N$ & $\bar{U}_\theta /N$ \\ \hline
-		Political books & 105  & 441 & 3 & 3 & 11.7 & 0.707 \\
-		Primary school & 238 & 5539 & 13 &  10 & 45.9  & 1.61 \\
-		FB egonet & 747 & 30025 & 480 & 10  & 66.9 & 7.47
+	\begin{tabular}{c|ccc|c|c|cc|c|cc}
+		Dataset & $N$ & $E$ & $D$ & $B$ & $\bar{S}_b /N$ & $\bar{\Lcal}_0$ & $\bar{\Lcal}_1$ & $D'$ & $\bar{\Lcal}_0'$ & $\bar{\Lcal}_1'$  \\ \hline
+		Polbooks & 105  & 441 & 3 & 3 & 11.7 & &&--&--&-- \\
+		School & 238 & 5539 & 13 &  10 & 45.9  & &&10&& \\
+		FB egonet & 747 & 30025 & 480 & 10  & 66.9 & &&&&
 	\end{tabular}

### 2021-05-20 nips/inference.tex
More work on experiments

@@
 	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
+	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2 = N \cdot \Lcal(\theta) + \frac{1}{2\sigma_\theta^2} ||\theta||^2
 	\label{eqn:U-form}
@@
 $U(\theta)$ in equation \ref{eqn:U-form} appears a typical objective function to be minimised for neural network training.
-The first term is the cross-entropy between actual and predicted labels.
-The second term - introduced by the prior - brings a form of regularisation, preventing over-fitting.
+The first term -- which we collect into $N \cdot \Lcal(\theta)$ -- is the cross-entropy between graph-predicted and feature-predicted block memberships.
+We denote the average cross-entropy $\Lcal(\theta)$ as it will become useful later on when performing experiments.
+The second term of equation \ref{eqn:U-form} -- introduced by the prior -- brings a form of regularisation, preventing over-fitting.
 In traditional applications we only seek the value of $\theta$ that minimises the objective function $U(\theta)$, which in our case would yield the maximum a posteriori (MAP) estimate.
@@
 Nevertheless, we can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
-We therefore adopt the Metropolis Adjusted Langevin Algorithm (MALA) - first proposed by \citet{mala-tweedie} - which leverages just that.
+We therefore adopt the Metropolis Adjusted Langevin Algorithm (MALA) -- first proposed by \citet{mala-tweedie} -- which leverages just that.
 Given the current sample $\theta$, we propose a new sample $\theta'$ according to equation \ref{eqn:theta-update}
@@
 
+TODO: burn-in and thinning
+
 \subsection{Dimensionality reduction}
@@
 \begin{equation}
-	\hat{\mu}_{ij} \coloneqq \frac{1}{T_\theta} \sum_{t=1}^{T_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
-	\hat{\sigma_{ij}} \coloneqq \frac{1}{T_\theta} \sum_{t=1}^{T_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2
+	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
+	\hat{\sigma}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2
 \end{equation}
@@
 A simple heuristic to discard the least important features requires specifying a cutoff $c > 0$ and a multiplier $k > 0$.
-We then discard all features $j$ wuch that equation \ref{eqn:discard} is satisfied.
+We then discard all features $j$ such that equation \ref{eqn:discard} is satisfied.
 %

### 2021-05-21 nips/experiments.tex
School experiment

@@
 The SBM has been extensively analysed in other works \cite{Peixoto-Bayesian-Microcanonical} so that will not be the emphasis of our analysis.
-We use a simple metric to gauge its fit: the mean description length per node of the $b$-samples (equation \ref{eqn:mean-dl}).
+$S(b)$ (equation \ref{eqn:dl-form}) can be interpreted as the description length of the partition imposed by $b$.
+In other words this is the amount of information needed to describe the graph and the associated SBM parameters.
+It is only natural to divide this quantity by the number of entities (nodes and edges) in our graph $N+E$ to allow for rough comparison between graphs.
+We use a simple metric to gauge the fit of the SBM: the description length per entity averaged over the $b$-samples (equation \ref{eqn:mean-dl}):
 %
 \begin{equation}
-	\frac{1}{N} \bar{S}_b \coloneqq \frac{1}{N |\Tcal_b|} \sum_{t\in \Tcal_b} S \left( b^{(t)} \right)
+	\bar{S}_e \coloneqq \frac{1}{(N+E) |\Tcal_b|} \sum_{t\in \Tcal_b} S \left( b^{(t)} \right)
 	\label{eqn:mean-dl}
@@
 We then retrain the feature-block predictor using just the retained features and report the loss over the training and test sets for the reduced classifier -- denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively.
-These are also
+These values are also included in the table except for the polbooks dataset as this is already very low.
 
@@
 	\centering
-	\caption{Table of results}
+	\caption{Results}
 	\label{tab:results}
-	\begin{tabular}{c|ccc|c|c|cc|c|cc}
-		Dataset & $N$ & $E$ & $D$ & $B$ & $\bar{S}_b /N$ & $\bar{\Lcal}_0$ & $\bar{\Lcal}_1$ & $D'$ & $\bar{\Lcal}_0'$ & $\bar{\Lcal}_1'$  \\ \hline
-		Polbooks & 105  & 441 & 3 & 3 & 11.7 & &&--&--&-- \\
-		School & 238 & 5539 & 13 &  10 & 45.9  & &&10&& \\
-		FB egonet & 747 & 30025 & 480 & 10  & 66.9 & &&&&
+	\resizebox{\textwidth}{!}{%
+	\begin{tabular}{c|c|c|c|cc|c|cc}
+		Dataset  & $B$ & $\bar{S}_e$ & D & $\bar{\Lcal}_0$ & $\bar{\Lcal}_1$ & $D'$ & $\bar{\Lcal}_0'$ & $\bar{\Lcal}_1'$  \\ \hline
+		Polbooks & 3 & $2.250 \pm 0.000$ & 3 & $0.550 \pm 0.023$ & $0.605 \pm 0.081$ &--&--&-- \\
+		School & 10 & $1.900 \pm 0.011$ & 13 & $0.778 \pm 0.066$ & $0.929 \pm 0.158$ &10 & $0.760 \pm 0.058$ & $0.847 \pm 0.148$ \\
+		FB egonet & 10  & 66.9 & 480 & & & & &
 	\end{tabular}
+	}
 \end{table}

### 2021-05-22 nips/appendices.tex
Much work

@@
 \begin{equation}
-	h_t = a(b + t)^{-\gamma}
+	h_t = \alpha(\beta + t)^{-\gamma}
 \end{equation}
 %
-Where $a, b, \gamma$ are hyper-parameters to be chosen.
-We require $a,b > 0$ and $\gamma \in (0.5, 1]$ to satisfy equation \ref{eqn:h-constraints}.
-We find empirically that $a=250, b=1000, \gamma=0.8$ yield good burn-in times for our MALA implementations.
+Where $\alpha, \beta, \gamma$ are hyper-parameters to be chosen.
+We require $\alpha,\beta > 0$ and $\gamma \in (0.5, 1]$ to satisfy equation \ref{eqn:h-constraints}.
+To reduce the number of hyperparameters we set these to have values given by equation \ref{eqn:step-size-params}.
+%
+\begin{equation}
+	\alpha = \frac{250 \cdot s}{N} \qquad \beta = 1000 \qquad \gamma = 0.8
+	\label{eqn:step-size-params}
+\end{equation}
+%
+Where $N$ is the number of data-points we are considering and now $s$ is the only free variable which we call the step-size scaling.
+
+\subsection{Burn-in and thinning}
+
+As with any MCMC method, we must deal with the issues presented by burn-in and thinning.
+We have introduced the notation $\Tcal_b$ and $\Tcal_\theta$ to denote the set of samples we keep from the $b$ and $\theta$ chains respectively.
+Note that we generate $T_b$ and $T_\theta$ samples total.
+The burn-in period refers to the time taken for the Markov Chain to converge to the stationary distribution.
+Sample thinning is necessary to ensure that neighbouring samples satisfy independence.
+However, as we do not leverage the independence property this is less important in our analysis.
+We can write the general set $\Tcal_\star$ as:
+%
+\begin{equation}
+	\Tcal_\star = \{T_\star \kappa_\star + i \lambda_\star :  
+	0 \leq i \leq \lfloor T_\star(1 - \kappa_\star) / \lambda_\star \rfloor \}
+\end{equation}
+%
+Where the parameter $\kappa_\star \in (0, 1)$ controls our burn-in and $\lambda_\star$ controls our thinning.
+$\kappa_\star$ can be determined by plotting the log-target (either $S(b^{(t)})$ or $U(\theta^{(t)})$ with respect to the epoch $t$.
+$\kappa_\star$ is then chosen to encompass the region where the log-target has roughly equilibrated.
+As we do not leverage sample independence $\lambda_\star$ can be chosen less rigorously.
+We often just use $\lambda_\star=5$
+
+\subsection{Hyperparameter values}
+
+\begin{table}[!h]
+	\centering
+	\caption{Hyperparameter values for each experiment}
+	\label{tab:hyperparams}
+	\resizebox{\textwidth}{!}{%
+		\begin{tabular}{c|ccc|ccc|cccc|cc|cccc}
+			Dataset & 
+			$B$ & $f$ & $\sigma_\theta$ & 
+			$T_b$ & $\kappa_b$ & $\lambda_b$ & 
+			$T_\theta$ & $\kappa_\theta$ & $\lambda_\theta$ & $s$ &
+			$k$ & $D'$ &
+			$T_\theta'$ & $\kappa_\theta'$ & $\lambda_\theta'$ & $s'$
+			\\ \hline
+			Polbooks &
+			3 & 0.8 & 1 &
+			1,000 & 0.2 & 5 &
+			10,000 & 0.2 & 10 & 0.2 &
+			-- & -- & 
+			-- & -- & -- & -- \\
+			School &
+			10 & 0.8 & 1 &
+			1,000 & 0.2 & 5 &
+			10,000 & 0.2 & 10 & 0.2 &
+			1 & 10 & 
+			10,000 & 0.2 & 10 & 0.2 \\
+			FB Egonet &
+			10 & 0.8 & 1 &
+			1,000 & 0.2 & 5 &
+			10,000 & 0.2 & 10 & 0.2 &
+			1 & 10 & 
+			10,000 & 0.2 & 10 & 0.2 \\
+		\end{tabular}
+	}
+\end{table}
 
@@
 
+\FloatBarrier
+
 \subsection{Inferring the number of blocks}
@@
 We take the number of non-empty blocks at the MDL to be our fixed block number $B$ for subsequent analysis.
-Indeed, it is prudent to start our $b$-chain at this MDL solution as then the burn-in time is greatly reduced.
+Indeed, it is prudent to start our $b$-chain at this MDL solution as then we can neglect the burn-in time.
 

### 2021-05-22 nips/experiments.tex
Much work

@@
 %
-However, to assess the performance of the feature-to-block predictor, we must partition the vertex set $[N]$ into training and test sets -- $\Gcal_0$ and $\Gcal_1$ respectively\footnote{We choose to randomly create the partition on each experiment run such that 80\% of the available vertices form our training set $\Gcal$ and the remaining 20\% are held out to form our test set $\Gcal^C$}.
+However, to assess the performance of the feature-to-block predictor, we must partition the vertex set $[N]$ into training and test sets -- $\Gcal_0$ and $\Gcal_1$ respectively
+\footnote{We choose to randomly create the partition on each experiment run such that a fraction $f$ of the available vertices form our training set $\Gcal$ and the remaining vertices are held out to form our test set $\Gcal^C$.
+For all the experiments $f=0.7$.}
+.
 The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal$ to train the $\theta$-chain.
@@
 	\resizebox{\textwidth}{!}{%
-	\begin{tabular}{c|c|c|c|cc|c|cc}
-		Dataset  & $B$ & $\bar{S}_e$ & D & $\bar{\Lcal}_0$ & $\bar{\Lcal}_1$ & $D'$ & $\bar{\Lcal}_0'$ & $\bar{\Lcal}_1'$  \\ \hline
-		Polbooks & 3 & $2.250 \pm 0.000$ & 3 & $0.550 \pm 0.023$ & $0.605 \pm 0.081$ &--&--&-- \\
-		School & 10 & $1.900 \pm 0.011$ & 13 & $0.778 \pm 0.066$ & $0.929 \pm 0.158$ &10 & $0.760 \pm 0.058$ & $0.847 \pm 0.148$ \\
-		FB egonet & 10  & 66.9 & 480 & & & & &
+	\begin{tabular}{c|ccc|c|cc|ccc}
+		Dataset  & $B$ & $D$ & $D'$ & $\bar{S}_e$ & $\bar{\Lcal}_0$ & $\bar{\Lcal}_1$ & $c^*$ & $\bar{\Lcal}_0'$ & $\bar{\Lcal}_1'$  \\ \hline
+		Polbooks & 3 & 3 & -- & $2.250 \pm 0.001$ & $0.584 \pm 0.033$ & $0.557 \pm 0.061$ & -- & -- & -- \\
+		School & 10 & 13 & 10 & $1.895 \pm 0.006$ & $0.789 \pm 0.109$ & $0.876 \pm 0.135$ & $1.114 \pm 0.224$ & $0.786 \pm 0.103$ & $0.833 \pm 0.126$ \\
+		FB egonet & 10  & 480 & 10 & & & & &
 	\end{tabular}

### 2021-05-22 nips/inference.tex
School experiment
Much work

@@
 	S(b) = - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big)
+	\label{eqn:dl-form}
 \end{equation}
@@
 
-Once we have the samples $\left\{ \theta^{(t)} \right\}_{t=1}^{T_\theta} \sim p(\theta | A, X)$, we can compute the empirical mean and standard deviation of each component of $\theta$.
+Once we have the samples $\left\{ \theta^{(t)} \right\} \sim p(\theta | A, X)$, we can compute the empirical mean and standard deviation of each component of $\theta$.
 Switching back to matrix notation we define $\theta = W$, such that $W_{ij}$ is the weight component for block $i$ and feature $j$, we can define:
@@
 A simple heuristic to discard the least important features requires specifying a cutoff $c > 0$ and a multiplier $k > 0$.
-We then discard all features $j$ such that equation \ref{eqn:discard} is satisfied.
+We define the function $\Fcal_i(j)$ as in \ref{eqn:fij} then only keep features in $\Dcal'$ given by equation \ref{eqn:kept-feature-set}.
 %
 \begin{equation}
-	(\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c) = \emptyset \quad \forall i \in \{1,2 \dots B\}
-	\label{eqn:discard}
+	\Fcal_i(j) \coloneqq (\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c)
+	\label{eqn:fij}
+\end{equation}
+\begin{equation}
+	\Dcal' = \left\{ j \in [D] : \exists i \in [B] \textrm{ s.t. }  \Fcal_i(j) \neq \emptyset \right\}
+	\label{eqn:kept-feature-set}
 \end{equation}
 %
-Intuitively, this means discarding any feature for which $\hat{\mu}_{ij} \pm k\hat{\sigma_{ij}}$ lies within or spans the null region $(-c, c)$.
+Intuitively, this means discarding any feature for which $\hat{\mu}_{ij} \pm k\hat{\sigma}_{ij}$ lies within or spans the null region $(-c, c)$ for all block indices.
 If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \mu_{ij}, \sigma_{ij})$, then this is effectively a hypothesis test on the value of $W_{ij}$.
+$\Dcal'$ then comprises all features for which $H_1$ is accepted.
 %
@@
 %
-With the multiplier $k$ specifying the degree of significance of the result (equivalent to the probability of a type-1 error).
+The multiplier $k$ specifies the degree of significance of the result (equivalent to the probability of a type-1 error).
 However, as the approximation is not exact we will only treat this dimensionality reduction method as a useful heuristic and not an exact method.
+
+Conversely, we could fix $k=k_0$ and the dimension of our reduced feature set $|\Dcal'|=D'$.
+We would then like to find the largest value of $c$ such that $|\Dcal'|=D'$ given $k=k_0$.
+This is summarised in equation \ref{eqn:c-star}.
+This approach may be preferred as it fixes the number of reduced dimensions.
+%
+\begin{equation}
+	c^* = \argmax_{c>0} (c : |\Dcal'| = D', k=k_0)
+	\label{eqn:c-star}
+\end{equation}
+%

### 2021-05-22 nips/experiments.tex
Results are in

@@
 Table \ref{tab:results} summarises the results for each experiment.
-We run each experiment 5 times with the same hyperparameter values\footnote{TODO hyperparam values} and report the mean plus or minus one standard deviation.
+We run each experiment $n=10$ times with the same hyper-parameter values\footnote{For a comprehensive list of the hyper-parameters used for each experiment please see appendix \ref{appdx:hyperparams}} and report the mean plus or minus one standard deviation.
 
@@
 		Polbooks & 3 & 3 & -- & $2.250 \pm 0.001$ & $0.584 \pm 0.033$ & $0.557 \pm 0.061$ & -- & -- & -- \\
-		School & 10 & 13 & 10 & $1.895 \pm 0.006$ & $0.789 \pm 0.109$ & $0.876 \pm 0.135$ & $1.114 \pm 0.224$ & $0.786 \pm 0.103$ & $0.833 \pm 0.126$ \\
-		FB egonet & 10  & 480 & 10 & & & & &
+		School & 10 & 13 & 10 & $1.894 \pm 0.006$ & $0.793 \pm 0.096$ & $0.914 \pm 0.112$ & $1.112 \pm 0.23$ & $0.791 \pm 0.096$ & $0.864 \pm 0.139$ \\
+		FB egonet & 10  & 480 & 10 & $1.626 \pm 0.003$ & $1.305 \pm 0.034$ & $1.539 \pm 0.087$ & $0.942 \pm 0.042$ & $1.496 \pm 0.093$ & $1.578 \pm 0.104$
 	\end{tabular}

### 2021-05-22 nips/inference.tex
Results are in

@@
 This may lead to longer burn-in and autocorrelation times of the resulting Markov Chain.
-The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends on $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
+The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends only on $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
 We would rather deal with the expectation of each $y_{ij}^{(t)}$:
@@
 \begin{equation}
-	\hat{y}_{ij} \coloneqq \frac{1}{T} \sum_{t=1}^{T} y_{ij}^{(t)} = \frac{1}{T} \sum_{t=1}^{T} \one\{b_i^{(t)} = j\}
+	\hat{y}_{ij} \coloneqq \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} y_{ij}^{(t)} = \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} \one\{b_i^{(t)} = j\}
 \end{equation}
@@
 
-TODO: burn-in and thinning
+A more fleshed-out discussion of how we choose to thin our samples is given in appendix \ref{appdx:burn-in-thinning}
 

### 2021-05-22 nips/ffbm.tex
Work on intro
Bit more work

--- 2021-05-16 nips/lbgm.tex
@@
-\section{Latent block generative model}
+\section{Feature-first block model}
 
@@
 
-The proposed generative model (which we call the latent block generative model - LBGM) is given in figure \ref{fig:lbgm}.
+The proposed generative model (which we call the feature-first block model - FFBM) is given in figure \ref{fig:ffbm}.
 We start, with the feature matrix X and generate a vector of block memberships $b$.
@@
 	\end{tikzpicture}
-	\caption{Latent block generative model}
-	\label{fig:lbgm}
+	\caption{Feature-first block model}
+	\label{fig:ffbm}
 \end{figure}
@@
 In our model, the block memberships vector $b$ is an intermediate latent variable and so we are not free to choose a prior for it.
-Nevertheless, as far as inference on the right-hand-side of figure \ref{fig:lbgm}, we regard $p(b | X)$ as a pseudo-prior on $b$.
+Nevertheless, as far as inference on the right-hand-side of figure \ref{fig:ffbm}, we regard $p(b | X)$ as a pseudo-prior on $b$.
 We can show (appendix \ref{appdx:b|x}) that our choice of prior for $p(\theta)$ in equation \ref{eqn:theta-prior} leads to a uniform $p(b | X)$ in equation \ref{eqn:b-pseudo-prior}.

### 2021-05-22 nips/experiments.tex
All plots

@@
 \FloatBarrier
+
+\begin{figure}[!h]
+	\centering
+	\begin{subfigure}[t]{0.28\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{polbooks-graph.png}
+		\caption{Polbooks}
+		\label{fig:polbooks-graph}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[t]{0.28\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{school-graph.png}
+		\caption{School}
+		\label{fig:school-graph}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[t]{0.28\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{fb-graph.png}
+		\caption{Facebook egonet}
+		\label{fig:fb-graph}
+	\end{subfigure}
+	\begin{subfigure}[t]{0.10\linewidth}
+		\centering
+		\includegraphics[width=0.8\linewidth]{10-vertical-legend.png}
+		\caption{Legend}
+		\label{fig:10-legend}
+	\end{subfigure}
+	\caption{Networks laid out and coloured according by inferred block memberships $\hat{y}$ for a given experiment iteration.
+Visualisation performed using \textit{graph-tool} \cite{peixoto_graph-tool_2014}}
+\end{figure}
+
+\begin{figure}[!h]
+	\centering
+	\begin{subfigure}[t]{0.32\linewidth}
+		\centering
+		\vskip 0pt
+		\includegraphics[width=\linewidth]{polbooks-null.png}
+		\caption{Polbooks}
+		\label{fig:polbooks-null}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[t]{0.32\linewidth}
+		\centering
+		\vskip 0pt
+		\includegraphics[width=\linewidth]{school-null.png}
+		\caption{School}
+		\label{fig:school-null}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[t]{0.32\linewidth}
+		\centering
+		\vskip 0pt
+		\includegraphics[width=\linewidth]{fb-null.png}
+		\caption{Facebook egonet}
+		\label{fig:fb-null}
+	\end{subfigure}
+	\caption{Reduced dimension feature-to-block generator weight samples}
+\end{figure}
+
 \subsection{Political books}
@@
 We sample the block-generator parameters and plot the emprical mean and standard deviation of each sample on figure \ref{fig:book-null}.
-%
-\begin{figure}[!h]
-	\centering
-	\begin{subfigure}{0.3\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{polbooks-graph.png}
-		\fbox{\includegraphics[width=0.4\linewidth]{3-legend.png}}
-		\caption{Sampled block memberships $\hat{Y}$}
-		\label{fig:books-graph}
-	\end{subfigure}
-	\hfill
-	\begin{subfigure}{0.5\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{polbooks-null.png}
-		\caption{Feature weights}
-		\label{fig:book-null}
-	\end{subfigure}
-	\caption{Political books network \cite{polbooks}}
-\end{figure}
-%
+
 Indeed, for all 3 blocks, each has a distinct political affiliation as its highest magnitude component.
@@
 We visualise the inferred block memberships in figure \ref{fig:school-graph}.
-%
-\begin{figure}[!h]
-	\centering
-	\begin{subfigure}{0.45\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{school-graph.png}
-		\fbox{\includegraphics[width=0.5\linewidth]{10-legend.png}}
-		\caption{Sampled block memberships $\hat{Y}$}
-		\label{fig:school-graph}
-	\end{subfigure}
-	\hfill
-	\begin{subfigure}{0.45\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{school-null.png}
-		\caption{Reduced dimension feature weights}
-		\label{fig:school-null}
-	\end{subfigure}
-	\caption{Primary school dynamic contacts network \cite{schools}}
-\end{figure}
 
@@
 More so than language, gender or surname - the school\footnote{This includes higher education institutes} you attend is the best explanatory variable for how the observed network's communities are partitioned at the highest level.
-%
-\begin{figure}[!h]
-	\centering
-	\begin{subfigure}{0.45\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{fb-graph.png}
-		\fbox{\includegraphics[width=0.3\linewidth]{10-legend.png}}
-		\caption{Sampled block memberships $\hat{Y}$}
-		\label{fig:fb-graph}
-	\end{subfigure}
-	\hfill
-	\begin{subfigure}{0.45\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{fb-null.png}
-		\caption{Reduced dimension feature weights}
-		\label{fig:fb-null}
-	\end{subfigure}
-	\caption{Facebook egonet id 1912 \cite{fb-snap}}
-\end{figure}
 

### 2021-05-23 nips/appendices.tex
Results are in
Finish writing

@@
 \subsection{Burn-in and thinning}
+\label{appdx:burn-in-thinning}
 
@@
 As we do not leverage sample independence $\lambda_\star$ can be chosen less rigorously.
-We often just use $\lambda_\star=5$
+We often just use $\lambda_b=5$ and $\lambda_\theta = 10$.
 
 \subsection{Hyperparameter values}
+\label{appdx:hyperparams}
 
@@
 	\centering
-	\caption{Hyperparameter values for each experiment}
+	\caption{Hyper-parameter values for each experiment}
 	\label{tab:hyperparams}
@@
 			Polbooks &
-			3 & 0.8 & 1 &
+			3 & 0.7 & 1 &
 			1,000 & 0.2 & 5 &
-			10,000 & 0.2 & 10 & 0.2 &
+			10,000 & 0.4 & 10 & 0.05 &
 			-- & -- & 
@@
 			School &
-			10 & 0.8 & 1 &
+			10 & 0.7 & 1 &
 			1,000 & 0.2 & 5 &
-			10,000 & 0.2 & 10 & 0.2 &
+			10,000 & 0.4 & 10 & 0.2 &
 			1 & 10 & 
-			10,000 & 0.2 & 10 & 0.2 \\
+			10,000 & 0.4 & 10 & 0.2 \\
 			FB Egonet &
-			10 & 0.8 & 1 &
+			10 & 0.7 & 1 &
 			1,000 & 0.2 & 5 &
-			10,000 & 0.2 & 10 & 0.2 &
+			10,000 & 0.4 & 10 & 0.017 &
 			1 & 10 & 
-			10,000 & 0.2 & 10 & 0.2 \\
+			10,000 & 0.4 & 10 & 0.5 \\
 		\end{tabular}
@@
 	\begin{algorithmic} % enter the algorithmic environment
-
-		\FOR{$t \in \{0, 1 \dots T_b - 1\}$}
+		\STATE $b^{(0)} \gets \argmax_b S(b|A)$
+		\FOR{$t \in [T_b]$}
 		\STATE $b' \gets \sim q_b(b^{(t)}, b' | A)$
@@
 
-		\STATE $\hat{Y}_{ij} \gets \frac{1}{\mathcal{|S|}} \sum_{t \in \mathcal{S}} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
+		\STATE $\hat{Y}_{ij} \gets \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
+		\STATE $\theta^{(0)} \gets \sim \Gaussian(0, \sigma_\theta I)$
 
@@
 
-		\FOR{$t \in \{0, 1 \dots T_\theta - 1\}$}
+		\FOR{$t \in [T_\theta]$}
 		\STATE $\xi \gets \sim \Gaussian(0, I)$
@@
 \subsection{Implementation details}
-\label{sec:imp-details}
+\label{appdx:imp-details}
 
@@
 \end{itemize}
+
+On this hardware each experiment iteration took the following amount of time to execute:
+
+\begin{table}[!h]
+	\centering
+	\caption{Compute-time for each experiment}
+	\label{tab:compute-time}
+	\begin{tabular}{c|ccc|c}
+		Dataset & $b$-chain & $\theta$-chain & Reduced $\theta$-chain & Overall compute time \\ \hline
+		Polbooks & $\sim$1s & $\sim$4s & -- & $\sim$5s \\
+		School & $\sim$10s & $\sim$10s & $\sim$10s & $\sim$30s \\
+		FB Egonet & $\sim$20s & $\sim$180s & $\sim$10s & $\sim$210s
+	\end{tabular}
+\end{table}

### 2021-05-23 nips/experiments.tex
Finish writing

@@
 \section{Experiments}
+\label{sec:experiments}
 
@@
 \begin{equation}
-	\bar{\Lcal}_a \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_a \left( \theta^{(t)} \right)
+	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star \left( \theta^{(t)} \right)
 	\qquad \textrm{where} \qquad
-	\Lcal_a \left( \theta^{(t)} \right) \coloneqq \frac{1}{|\Gcal_a|} \sum_{i \in \Gcal_a}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)}
+	\Lcal_\star \left( \theta^{(t)} \right) \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)}
 	\label{eqn:cross-entropy-loss}
@@
 %
-Where $a \in \{0, 1\}$ has been introduced to toggle between training and test sets.
+Where $\star \in \{0, 1\}$ has been introduced to toggle between training and test sets.
 Table \ref{tab:results} summarises the results for each experiment.
@@
 However, we also wish to determine the power of the dimensionality reduction method.
-For this we discard all features that satisfy equation \ref{eqn:discard}, to reduce the dimension from $D$ to $D'$.
-We then retrain the feature-block predictor using just the retained features and report the loss over the training and test sets for the reduced classifier -- denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively.
-These values are also included in the table except for the polbooks dataset as this is already very low.
+For this we apply equation \ref{eqn:c-star}, to reduce the dimension from $D$ to $D'$ with $k=1$.
+We then retrain the feature-block predictor using just the retained feature set $\Dcal'$ and report the loss over the training and test sets for the reduced classifier -- denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively.
+These values are also included in the table except for the polbooks dataset which is already very low dimension.
 
@@
 	\centering
-	\caption{Results}
+	\caption{Experimental results averaged over $n=10$ iterations (mean $\pm$ standard deviation)}
 	\label{tab:results}
@@
 		Polbooks & 3 & 3 & -- & $2.250 \pm 0.001$ & $0.584 \pm 0.033$ & $0.557 \pm 0.061$ & -- & -- & -- \\
-		School & 10 & 13 & 10 & $1.894 \pm 0.006$ & $0.793 \pm 0.096$ & $0.914 \pm 0.112$ & $1.112 \pm 0.23$ & $0.791 \pm 0.096$ & $0.864 \pm 0.139$ \\
+		School & 10 & 13 & 10 & $1.894 \pm 0.006$ & $0.793 \pm 0.096$ & $0.914 \pm 0.112$ & $1.112 \pm 0.230$ & $0.791 \pm 0.096$ & $0.864 \pm 0.139$ \\
 		FB egonet & 10  & 480 & 10 & $1.626 \pm 0.003$ & $1.305 \pm 0.034$ & $1.539 \pm 0.087$ & $0.942 \pm 0.042$ & $1.496 \pm 0.093$ & $1.578 \pm 0.104$
@@
 
-\FloatBarrier
+There are some general trends we can observe in these results.
+Firstly, the variance of the test loss $\bar{\Lcal}_1$ tends to be higher than the training loss $\bar{\Lcal}_0$.
+This is expected as our test set is smaller than the training set and so more susceptible to variability in its selection.
+Indeed, much of the variance in the evaluation of $\bar{\Lcal}_0$ and $\bar{\Lcal}_1$ comes from the random partitioning of the graph into training and test sets $\Gcal_0$ and $\Gcal_1$.
+Secondly, it can be seen that the dimensionality reduction procedure brings the training and test losses closer together.
+This implies that the features we keep are indeed correlated with the underlying graphical partition.
+
+The average description length of the graph per entity $\bar{S}_e$ has very low variance implying the detected communities can be found reliably (to within an arbitrary relabelling of blocks).
+For reference we plot an inferred partition for each of the graphs on figure \ref{fig:graphs-all}.
+Unsurprisingly, the polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do appear to partition the graph into densely connected clusters.
 
@@
 Visualisation performed using \textit{graph-tool} \cite{peixoto_graph-tool_2014}}
+	\label{fig:graphs-all}
 \end{figure}
+
+On figure \ref{fig:null-all} we give the sampled feature weights for the reduced feature-set $\Dcal'$ for a given experiment iteration (or just the original $\Dcal$ for the polbooks dataset).
+We go on to analyse each of these in turn.
 
@@
 	\caption{Reduced dimension feature-to-block generator weight samples}
+	\label{fig:null-all}
 \end{figure}
@@
 We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral).
-The inferred block memberships are given in figure \ref{fig:books-graph}.
-We sample the block-generator parameters and plot the emprical mean and standard deviation of each sample on figure \ref{fig:book-null}.
+The inferred block memberships are given in figure \ref{fig:polbooks-graph}.
+We sample the block-generator parameters and plot the emprical mean and standard deviation of each sample on figure \ref{fig:polbooks-null}.
 
@@
 This is strong evidence that political affiliation is indeed the axis which best predicts the 3-way natural partition of the graph into blocks.
-Nevertheless, for block index 1, although \verb*|neutral| is the best predictor of the ones we have available, the value is not particularly extreme.
+Nevertheless, for block index 0, although \verb*|neutral| is the best predictor of the ones we have available, the value is not particularly extreme.
 Perhaps, there is an unobserved feature that places a book into this block.
-Indeed, block-0 is more of a centre-right block than true centre.
+However, by inspecting figure \ref{fig:polbooks-graph} we note that there are very few edges between blocks 1 and 2 (the detected conservative and liberal blocks as per figure \ref{fig:school-null}) and indeed block 0 effectively acts as a separator for blocks 1 and 2.
+It is therefore, possible that the moderate value of the neutral weight for block 0 reflects the fact that this block contains a few peripheral conservative and liberal books at the corners of the block.
 
-\FloatBarrier
+From table \ref{tab:results} we see that the training and test losses are very similar (the test loss mean is even below the training loss) and both are quite low in magnitude.
+This is very strong evidence that the features we have available (the author's political affiliation) are indeed the drivers of the overall graphical structure.
+
 \subsection{Primary school dynamic contacts}
@@
 
-As before, we proceed to sample the block-generator parameters $\theta$ and employ the dimensionality reduction heuristic defined in section \ref{sec:dim-reduction} with null threshold $c=1$ and standard deviation multiplier $k=1$.
-We then plot the weights for the features that survive the discard step in figure \ref{sec:dim-reduction}.
+As before, we proceed to sample the block-generator parameters $\theta$ and employ the dimensionality reduction heuristic with standard deviation multiplier $k=1$ to pick out the top $D'=10$ features.
+We then plot the weights for the surviving features $d \in \Dcal'$ on figure \ref{fig:school-null}.
 Immediately, we see that only the pupils' class memberships have survived (1A-5B); gender and teacher/student status have been discarded meaning that these are not good predictors of overall macro-structure.
@@
 However, some blocks have 2 comparably good classes as their predictor.
-For example, blocks 2 and 4 contain classes 4A and 4B as their 2 best predictors.
-This suggests that the social divide between classes is less pronounced for pupils in year 4 but there is nonetheless an unobserved variable that splits year 4 into 2 communities.
+For example, blocks 1 and 3 contain classes 3A and 3B as their 2 best predictors.
+This suggests that the social divide between classes is less pronounced for pupils in year 3 but there is nonetheless an unobserved variable that splits year 3 into 2 communities.
 The most surprising block is number 5 - which has comparable weightings for classes 5A and 1B.
 Perhaps there was a joint event between those two classes on the day the data were collected.
+
+As for the training and test losses of the original and reduced classifiers.
+We see that the training loss stays roughly constant between the original and reduced classifiers $\bar{\Lcal}_0 \approx \bar{\Lcal}_0'$ but the test loss of the reduced classifier is ever-so-slightly decreased (within margin of error).
+This would imply that the features we have discarded (gender and teacher/student status) do not impact the high-level graphical structure.
+Nevertheless, the features we have available do not capture the full picture but they do quite well (low overall loss).
 
@@
 This dataset was chosen to showcase the power of the dimensionality reduction technique as the feature-space has high dimension ($D=480$).
-We sample the $b$-chain specifying $B=10$ total blocks and use this to construct the $\theta$-samples as before.
+We sample the $b$-chain, specifying $B=10$ total blocks and use this to construct the $\theta$-samples as before.
 
-We apply the dimensionality reduction heuristic outlined in section \ref{sec:dim-reduction}, choosing $k=1$ and $c=0.8$.
-This leads to a reduced dimension $D'=13$.
-These parameter values are plotted on figure \ref{fig:fb-null}.
-The features that remain are those that best explain the high-level community structure.
+We again apply the dimensionality reduction heuristic with $k=1$ to target $D'=10$ features in the reduced set.
+The parameter values for the surviving features are plotted on figure \ref{fig:fb-null}.
+The remaining features are those that best explain the high-level community structure.
+The majority of the surviving features are education related.
 More so than language, gender or surname - the school\footnote{This includes higher education institutes} you attend is the best explanatory variable for how the observed network's communities are partitioned at the highest level.
+Nevertheless, for $D'=10$ we do not have good explanations for the makeup of blocks 5-9 inclusive
 
@@
 It might be possible to shift the feature values such that they take values in $\{-1, 1\}$ rather than $\{0, 1\}$ but this approach falls down when input features are mutually correlated.
-Indeed, for features with discrete values such as the class memberships (1A, 1B, 2A \dots 5B) it is preferred to keep $x_i \in \{0, 1\}^D$ as then each block can be readily accept more than one mutually exclusive feature-group (i.e.
+Indeed, for features with discrete values such as the class memberships (1A-5B) of the school dataset it is preferred to keep $x_i \in \{0, 1\}^D$ as then each block can be readily accept more than one mutually exclusive feature-group (i.e.
 two school-classes are part of the same block).
+
+The training and test losses (table \ref{tab:results}) are the highest of the considered networks.
+This suggests that the features we have available do not readily explain the high-level structure.
+Nonetheless, after reducing the problem dimension our test loss is not greatly increased which suggests that the features we do keep are the best of the ones we have available at predicting the overall network structure.

### 2021-05-23 nips/prelim.tex
Succinct prelims

@@
 
-This section defines some preliminary concepts required for the subsequent analysis.
-We first need a model for community-like structure in a graphical network.
+We first need a model for community-like structure in a network.
 For this we adopt the stochastic block model (SBM) - widely used across academia.
@@
 Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}.
-A paraphrased definition is given below for the non-degree-corrected SBM (NDC-SBM).
+To allow for degree-variability between members of the same block, we must choose the degree-corrected formulation of the SBM (DC-SBM).
 
-\begin{definition}[Microcanonical NDC-SBM]
+\begin{definition}[Microcanonical DC-SBM]
 	\label{defn:microcan-ndc-sbm}
 	Let $N \in \Integers^{+}$ denote the number of vertices in an undirected graph.
-The block memberships are encoded by a vector $b$ of length $N$ where each entry $b_i \in \{1, 2 \dots B\}$.
-$B \in \Integers^{+}$ is the number of non-empty blocks.
-Let $e$ be a $B \times B$ matrix of edge counts between blocks ($e_{rs}$ is number of edges from block $r$ onto block $s$ - or twice that number if $r=s$).
+The block memberships are encoded by a vector $b \in [B]^N$,\footnote{We introduce the notation $[K] \coloneqq \{0, 1, 2 \dots K-1\}$ to compactly define a set of $K$ indices.
+Clearly, $[K]$ is only defined for $K \in \Integers^+$.}
+	where $B \in \Integers^{+}$ is the number of non-empty blocks.
+	Let $e \in (\Integers_0^+)^{N \times N}$ be the matrix of edge counts between blocks.
+$e_{rs}$ is then the number of edges from block $r$ onto block $s$ -- or twice that number if $r=s$.
 For undirected graphs $e$ is symmetric.
-For a non-degree-corrected stochastic block model (NDC-SBM), we say that the graph $A$ is generated as follows:
-	%
-	\begin{equation}
-		A \sim \textrm{NDC-SBM}_{\textrm{MC}} (b, e)
-	\end{equation}
-	%
-	Where edges are placed uniformly at random but respecting the constraint imposed by $e$ and $b$.
-The additional parameters $N$ and $B$ are omitted as they are inferred from the shapes of $b$ and $e$.
-If we interpret $A$ as an adjacency matrix, then this constraint can be written formally as: $e_{rs} = \sum_{i,j} A_{ij} \one \{b_i = r\} \one \{b_j = s\}$.
-\end{definition}
+	Let $k \in (\Integers^+_0)^N$ be a vector denoting the degree sequence of the graph. $k_i$ is then the degree of vertex $i$.
 
-Nevertheless, this formulation does not accept high degree variability within blocks as is typical of real-world data.
-Indeed, the NDC-SBM favours a partition into high-degree and low-degree nodes rather than clusters of inter-connected nodes.
-We therefore introduce the degree-corrected SBM (DC-SBM) \cite{Peixoto-Bayesian-Microcanonical} to circumvent these issues.
-
-\begin{definition}[Microcanonical DC-SBM]
-	\label{defn:microcan-dc-sbm}
-	This is much like the NDC-SBM but has an additional parameter $k$ which is an $N$-length vector encoding the degree sequence ($k_i$ is the degree of vertex $i$). Therefore, we write:
+	For the degree-corrected stochastic block model (DC-SBM), we say that the graph's adjacency matrix $A \in \{0,1\}^{N \times N}$ is generated as follows:
 	%
@@
 	%
-	Once again, edges are placed uniformly at random but respecting the constraints imposed by the parameters.
-The DC-SBM has the additional constraint that $k_i = \sum_{j} A_{ij}$.
-In what follows, we will always assume the degree-corrected model unless otherwise specified.
-
+	Where edges are placed uniformly at random but respecting the constraints imposed by $e$, $b$ and $k$ -- hence why this formulation is given the microcanonical moniker.
+Specifically, $A$ must satisfy the following:
+	%
+	\begin{equation}
+		e_{rs} = \sum_{i, j} A_{ij} \one \{b_i = r\} \one \{b_j = s\} \qquad \textrm{and} \qquad
+		k_i = \sum_j A_{ij}
+	\end{equation}
+	%
+	The additional parameters $N$ and $B$ are omitted as they can be inferred from the shapes of $b$ and $e$.
 \end{definition}

### 2021-05-23 nips/appendices.tex
Proof-read

@@
 	p(b | X) &= \int p(b , \theta| X, \theta) d\theta = \int p(b | X, \theta) p(\theta | X) d\theta \\
-	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i=1}^{N} \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
-	&= \prod_{i=1}^{N} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j=1}^{B} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k=1}^{B} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}
+	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i \in [N] } \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
+	&= \prod_{i \in [N]} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j \in [B]} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k \in [B]} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}
 \end{align*}
 %
-We note that $b_i \in {1, 2, \dots B}$ and so the integral's value is unchanged with respect to $b_i$.
+We note that $b_i \in [B]$ and so the integral's value is unchanged with respect to $b_i$.
 The integrand has the same form no matter which value $b_i$ takes as the prior is the same for each $w_j$.
@@
 $\nicefrac{1}{c}$ is simply the size of the set of values that $b$ can take.
-We know $b_i \in \mathcal{B} = \{1, 2, \dots B\}$.
-Therefore, $b \in \mathcal{B}^N$ and $|\mathcal{B}^N| = |\mathcal{B}|^N = B^N= \nicefrac{1}{c}$.
-Putting this all together we show that:
+We know $b_i \in [B]$.
+Therefore, $b \in [B]^N$ and $|[B]^N| = B^N = \nicefrac{1}{c}$.
+Putting this all together we conclude that:
 %
@@
 \begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
+	U(\theta) = \left( \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log \frac{1}{a_{ij}} \right)
 	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
@@
 \begin{equation}
-	a_{ij} \coloneqq \phi_{j} (x_i; \theta) = \frac{\exp(w_j^T \tilde{x}_i)}{\sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i)}
+	a_{ij} \coloneqq \phi_{j} (x_i; \theta) = \frac{\exp(w_j^T \tilde{x}_i)}{\sum_{r \in [B]} \exp(w_r^T \tilde{x}_i)}
 	\label{eqn:a-ij} 
@@
 	\frac{\partial a_{ij}}{\partial w_k} &= \frac
-	{\tilde{x}_i \exp(w_j^T \tilde{x}_i) \delta_{jk} \cdot \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) 
+	{\tilde{x}_i \exp(w_j^T \tilde{x}_i) \delta_{jk} \cdot \sum_{r \in [B]} \exp(w_r^T \tilde{x}_i) 
 		- 
 		\exp(w_j^T \tilde{x}_i) \cdot \tilde{x}_i \exp(w_k^T \tilde{x}_i)}
-	{\left( \sum_{b=1}^{B} \exp(w_b^T \tilde{x}_i) \right)^2} \nonumber \\
+	{\left( \sum_{r \in [B]} \exp(w_r^T \tilde{x}_i) \right)^2} \nonumber \\
 	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right) 
@@
 \begin{equation}
-	\frac{ \partial}{\partial w_k} ||\theta||^2 = \frac{\partial}{\partial w_k} \left( \sum_{b=1}^B ||w_b||^2 \right) = 2w_k
+	\frac{ \partial}{\partial w_k} ||\theta||^2 = \frac{\partial}{\partial w_k} \left( \sum_{r \in [B]} ||w_r||^2 \right) = 2w_k
 \end{equation}
@@
 Therefore, we still have a short burn-in time but will not bounce around the mode for large $t$.
-As well as the trivial constraint for $h_t$ to be strictly positive and we introduce two further constraints as outlined by \citet{Bayesian-SGLD}:
+As well as the trivial constraint for $h_t$ to be strictly positive, we introduce two further constraints as outlined by \citet{Bayesian-SGLD}:
 %
@@
 We require $\alpha,\beta > 0$ and $\gamma \in (0.5, 1]$ to satisfy equation \ref{eqn:h-constraints}.
-To reduce the number of hyperparameters we set these to have values given by equation \ref{eqn:step-size-params}.
+To reduce the number of hyperparameters we set these to have values given by the equations \ref{eqn:step-size-params}.
 %
@@
 Where $N$ is the number of data-points we are considering and now $s$ is the only free variable which we call the step-size scaling.
+For approximate methods, we can choose to bypass the MH accept-reject entirely to speed up computation.
+If this is done, the algorithm is instead called stochastic gradient Langevin diffusion (SGLD) \cite{Bayesian-SGLD}.
+This speeds up computation at the expense of exactness of the method.
 
@@
 
+\subsection{Initializing the b-chain}
+
+For the purposes of our model (the FFBM), the number of blocks $B$ is a constant which must be specified by the data scientist.
+We could however, allow our choice of $B$ to be influenced by the observed data.
+This places us in the domain of empirical Bayes, which must be negotiated carefully.
+Prior beliefs must be determined a priori else they are not prior.
+However, as the number of blocks only specifies the coarseness of the analysis, it is fine to allow it to vary.
+Indeed, \citet{peixoto-determine-B} shows that for a fixed average degree the maximum number of detectable blocks scales as $O(\sqrt{N})$ where $N$ is the number of vertices.
+
+If we allow $B$ to vary in the $b$-chain (i.e.
+new blocks can be created and we permit empty blocks) then it can be run  until a minimum description length (MDL) solution is reached.
+We take the number of non-empty blocks at the MDL to be our fixed block number $B$ for subsequent analysis.
+Indeed, it is prudent to start our $b$-chain at this MDL solution as then we can burn-in time is greatly reduced.
+
+\subsection{Algorithms}
+
+\begin{algorithm} % enter the algorithm environment
+	\caption{Block membership sample generation} % give the algorithm a caption
+	\label{alg:b-samples} % and a label for \ref{} commands later in the document
+	\begin{algorithmic} % enter the algorithmic environment
+		\State $b^{(0)} \gets \argmin_b S(b|A)$ \Comment{Implemented as greedy heuristic in \textit{graph-tool} library}
+		\For{$t \in \{0, 1 \dots T_b - 1\}$}
+		\State $b' \gets \sim q_b(b^{(t)}, b' | A)$
+		\State $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
+		\State $\eta \gets \sim \textrm{Unif}(0,1)$
+		\If{$\log \eta < \log \alpha_b$}
+		\State $b^{(t+1)} \gets b'$
+		\Else
+		\State $b^{(t+1)} \gets b^{(t)}$
+		\EndIf
+		\EndFor
+
+		\State \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$
+
+	\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm} % enter the algorithm environment
+	\caption{FFBM parameter pseudo-marginal inference} % give the algorithm a caption
+	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
+	\begin{algorithmic} % enter the algorithmic environment
+
+		\State $\hat{Y}_{ij} \gets \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
+		\State $\theta^{(0)} \gets \sim \Gaussian(0, \sigma_\theta I)$
+
+		\item[]
+
+		\For{$t \in \{0, 1 \dots T_\theta - 1\}$}
+		\State $\xi \gets \sim \Gaussian(0, I)$
+		\State $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
+		\State $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
+		\State $\eta \gets \sim \textrm{Unif}(0,1)$
+		\If{$\log \eta < \log \alpha_\theta$}
+		\State $\theta^{(t+1)} \gets \theta'$
+		\Else
+		\State $\theta^{(t+1)} \gets \theta^{(t)}$
+		\EndIf
+		\EndFor
+
+		\State \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$
+
+	\end{algorithmic}
+\end{algorithm}
+
+\clearpage
 \subsection{Hyperparameter values}
@@
 
-\subsection{Algorithms}
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{Block membership sample generation} % give the algorithm a caption
-	\label{alg:b-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-		\STATE $b^{(0)} \gets \argmax_b S(b|A)$
-		\FOR{$t \in [T_b]$}
-		\STATE $b' \gets \sim q_b(b^{(t)}, b' | A)$
-		\STATE $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
-		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-		\IF{$\log \eta < \log \alpha_b$}
-		\STATE $b^{(t+1)} \gets b'$
-		\ELSE
-		\STATE $b^{(t+1)} \gets b^{(t)}$
-		\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{FFBM parameter pseudo-marginal inference} % give the algorithm a caption
-	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\STATE $\hat{Y}_{ij} \gets \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
-		\STATE $\theta^{(0)} \gets \sim \Gaussian(0, \sigma_\theta I)$
-
-		\item[]
-
-		\FOR{$t \in [T_\theta]$}
-		\STATE $\xi \gets \sim \Gaussian(0, I)$
-		\STATE $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
-		\STATE $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
-		\STATE $\eta \gets \sim \textrm{Unif}(0,1)$
-		\IF{$\log \eta < \log \alpha_\theta$}
-		\STATE $\theta^{(t+1)} \gets \theta'$
-		\ELSE
-		\STATE $\theta^{(t+1)} \gets \theta^{(t)}$
-		\ENDIF
-		\ENDFOR
-
-		\STATE \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\FloatBarrier
-
-\subsection{Inferring the number of blocks}
-
-For the purposes of our model (the FFBM), the number of blocks $B$ is a constant which must be specified by the data scientist.
-We could however, allow our choice of $B$ to be influenced by the observed data.
-This places us in the domain of empirical Bayes, which must be negotiated carefully.
-Prior beliefs must be determined a priori else they are not prior.
-However, as the number of blocks only specifies the coarseness of the analysis, it is fine to allow it to vary.
-Indeed, \citet{peixoto-determine-B} shows that for a fixed average degree the maximum number of detectable blocks scales as $O(\sqrt{N})$ where $N$ is the number of vertices.
-
-If we allow $B$ to vary in the $b$-chain (i.e.
-new blocks can be created and we permit empty blocks) then it can be run  until a minimum description length (MDL) solution is reached.
-We take the number of non-empty blocks at the MDL to be our fixed block number $B$ for subsequent analysis.
-Indeed, it is prudent to start our $b$-chain at this MDL solution as then we can neglect the burn-in time.
-
 \subsection{Implementation details}
@@
 
-All data analysis and visualisation was implemented in Python. Full source code is available at:
-
-\url{https://github.com/LozzaTray/Jormungandr}
-
-The scripts were run using a standard PC. Specs are:
+All data analysis and visualisation was implemented in Python.
+Full source code is available in the supplementary material.
+The scripts were run using a standard PC using the Windows Subsystem for Linux (WSL) environment.
+Specs are:
 

### 2021-05-23 nips/conclusion.tex
Finish writing
Proof-read

@@
 \section{Conclusion}
+\label{sec:conclusion}
 
 The FFBM was developed to address the shortcomings of other graphical models when testing how vertex features affect community structure.
-The idea is to divide the graph into its most natural partition and only then bring in the vertex features.
+The idea is to divide the graph into its most natural partition and test whether the vertex features can accurately explain this partition.
 We therefore hope to avoid a common pitfall in data science: correlation does not imply causation.
@@
 
-With the newly-defined FFBM, we go on to present an efficient inference algorithm to ascertain the parameters $\theta$ of the block generator.
+With the newly-defined FFBM, we go on to present an efficient inference algorithm to sample the parameters $\theta$ of the feature-to-block generator.
 This is introduced as two concurrent Markov chains to sample the block memberships $b$ and block generator parameters $\theta$.
-The key trick to make the $b$-chain efficient is the the use of the microcanonical rather than the traditional SBM (as proposed by \citet{Peixoto-Bayesian-Microcanonical}) which allows us to bypass an expensive summation over the latent memberships $b$.
+The key trick to make the $b$-chain efficient is the use of the microcanonical rather than the traditional SBM (as proposed by \citet{Peixoto-Bayesian-Microcanonical}) which allows us to bypass an expensive summation over the latent memberships $b$.
 Additionally, we use the empirical mean of the $b$-samples as the input to our $\theta$-chain, as this reduces the variance in our evaluation of the target distribution and thus shortens the burn-in and auto-correlation times of our $\theta$-chain.
@@
 That way, the structure of the network can be explained at all length-scales of interest.
+This work may be built upon to develop better methods of analysing structure in networks and drawing evidence-backed conclusions.
+So long as data collection techniques remain ethical and care is taken to respect personal privacy, such empowered decision-making can only help humankind.

### 2021-05-23 nips/experiments.tex
Proof-read

@@
 
-We apply the developed methods to a variety of datasets, to show the power of the method across different scenarios.
-We choose datasets to span a range of node counts $N$, edge counts $E$ and feature space dimension $D$.
-We consider the following datasets:
+We apply the developed methods to a variety of datasets.
+Datasets are chosen to span a range of node counts $N$, edge counts $E$ and feature space dimension $D$.
+We consider the following datasets:\footnote{For information as to how these data were collected and how personally identifiable information was removed, please refer to the cited papers.}
 
 \begin{itemize}
-	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=3$) - network of Amazon book sales about U.S.
+	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=3$) -- network of Amazon book sales about U.S.
 politics, published close to the presidential election in 2004.
 Two books are connected if they were frequently co-purchased by the same customer.
-Vertex features encode the broad political affiliation of the author (liberal, conservative or neutral).
+Vertex features encode the political affiliation of the author (liberal, conservative or neutral).
 
-	\item \textbf{Primary school dynamic contacts} \cite{schools} ($N=238, E=5539, D=13$) - network of face-to-face contacts amongst students and teachers at a primary school in Lyon, France.
+	\item \textbf{Primary school dynamic contacts} \cite{schools} ($N=238, E=5539, D=13$) -- network of face-to-face contacts amongst students and teachers at a primary school in Lyon, France.
 Two nodes are connected if the two parties shared a face-to-face interaction over the course of the day.
@@
 
-	\item \textbf{Facebook egonet} \cite{fb-snap} ($N=747, E=30025, D=480$) - an assortment of Facebook egonets.
+	\item \textbf{Facebook egonet} \cite{fb-snap} ($N=747, E=30025, D=480$) -- an assortment of Facebook egonets.
 These are networks of a particular user's friends list and all the connections within that.
 Vertex features are extracted from each user's profile and are fully anonymized.
+They include information about education history, languages spoken, gender, home-town, date of birth, name to give a few examples.
 We focus on the egonet with id 1912.
-
-%	\item \textbf{Maier Facebook Egonet} \cite{FB-Maier}  ($N=349, E=2336, D=32$) - egonet of the author's Facebook friends list.
-Each vertex has been manually labelled with a variety of features describing their relationship to the author.
-For our purposes we remove all nodes of degree 1 (those that are only connected to the egonode) as these cannot be said to be part of any community present in the graph.
-%		
-%	\item \textbf{Law firm} \cite{LawFirm} - a network of relationships between members of a law firm. Each relationship is categorised according to type: coworkers, friends or advice.
-%	
-%	\item \textbf{Twitch users} \cite{twitch} - a network of user-user friendships on the streaming service Twitch.
-Vertex labels are extracted according to video-games played, location and streaming habits.
-This dataset is also broken down into disjoint networks according to language.
-We only consider the English users with is a subnet with $N=7126$ vertices and $E=35324$ edges).
 
@@
 This can be split into two separate components: the microcanonical SBM fit (concerned with the $b$-samples) and the fit of the feature-to-block generator (concerned with the $\theta$-samples).
-The SBM has been extensively analysed in other works \cite{Peixoto-Bayesian-Microcanonical} so that will not be the emphasis of our analysis.
-$S(b)$ (equation \ref{eqn:dl-form}) can be interpreted as the description length of the partition imposed by $b$.
+Starting with the SBM, $S(b)$ (equation \ref{eqn:dl-form}) can be interpreted as the description length of the partition imposed by $b$.
 In other words this is the amount of information needed to describe the graph and the associated SBM parameters.
 It is only natural to divide this quantity by the number of entities (nodes and edges) in our graph $N+E$ to allow for rough comparison between graphs.
-We use a simple metric to gauge the fit of the SBM: the description length per entity averaged over the $b$-samples (equation \ref{eqn:mean-dl}):
+This defines a simple metric to gauge the fit of the SBM: the description length per entity averaged over the $b$-samples (equation \ref{eqn:mean-dl}):
 %
@@
 However, to assess the performance of the feature-to-block predictor, we must partition the vertex set $[N]$ into training and test sets -- $\Gcal_0$ and $\Gcal_1$ respectively
-\footnote{We choose to randomly create the partition on each experiment run such that a fraction $f$ of the available vertices form our training set $\Gcal$ and the remaining vertices are held out to form our test set $\Gcal^C$.
+\footnote{We choose to randomly create the partition on each experiment run such that a constant fraction $f$ of the available vertices go to form our training set $\Gcal$ and the remaining vertices are held out to form our test set $\Gcal^C$.
 For all the experiments $f=0.7$.}
@@
 The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal$ to train the $\theta$-chain.
-As $|\Gcal_0| \neq |\Gcal_1|$ sets, we cannot use the un-normalised log target $U$ (equation \ref{eqn:U-form}) for comparison between the two, as the prior term will be equal between the two but the total cross-entropy loss will be scaled by the size of each set.
+As $|\Gcal_0| \neq |\Gcal_1|$ sets, we cannot use the un-normalised log target $U$ (equation \ref{eqn:U-form}) for comparison as the total cross-entropy loss is scaled by the size of each set -- even though the prior term stays constant.
 We therefore must use the average cross-entropy loss over each set (equation \ref{eqn:cross-entropy-loss}).
@@
 We then retrain the feature-block predictor using just the retained feature set $\Dcal'$ and report the loss over the training and test sets for the reduced classifier -- denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively.
-These values are also included in the table except for the polbooks dataset which is already very low dimension.
+These values are also included in table \ref{tab:results} except for the polbooks dataset which is already very low dimension.
 
@@
 
-There are some general trends we can observe in these results.
+Table \ref{tab:results} already highlights some general trends in the results.
 Firstly, the variance of the test loss $\bar{\Lcal}_1$ tends to be higher than the training loss $\bar{\Lcal}_0$.
-This is expected as our test set is smaller than the training set and so more susceptible to variability in its selection.
+This is expected as our test set is smaller than the training set and so more susceptible to variability in its construction.
 Indeed, much of the variance in the evaluation of $\bar{\Lcal}_0$ and $\bar{\Lcal}_1$ comes from the random partitioning of the graph into training and test sets $\Gcal_0$ and $\Gcal_1$.
@@
 
-The average description length of the graph per entity $\bar{S}_e$ has very low variance implying the detected communities can be found reliably (to within an arbitrary relabelling of blocks).
+The average description length per entity of the graph $\bar{S}_e$ has very low variance implying the detected communities can be found reliably (to within an arbitrary relabelling of blocks).
 For reference we plot an inferred partition for each of the graphs on figure \ref{fig:graphs-all}.
-Unsurprisingly, the polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do appear to partition the graph into densely connected clusters.
+The polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do succeed at partitioning the graph into densely connected clusters.
 
@@
 
-Indeed, for all 3 blocks, each has a distinct political affiliation as its highest magnitude component.
+Indeed, from figure \ref{fig:polbooks-null}, we see that all 3 blocks have a distinct political affiliation as their largest positive component.
 This is strong evidence that political affiliation is indeed the axis which best predicts the 3-way natural partition of the graph into blocks.
-Nevertheless, for block index 0, although \verb*|neutral| is the best predictor of the ones we have available, the value is not particularly extreme.
+Nevertheless, for block index 1, although \verb*|neutral| is the best predictor of the ones we have available, the value is not particularly extreme.
 Perhaps, there is an unobserved feature that places a book into this block.
-However, by inspecting figure \ref{fig:polbooks-graph} we note that there are very few edges between blocks 1 and 2 (the detected conservative and liberal blocks as per figure \ref{fig:school-null}) and indeed block 0 effectively acts as a separator for blocks 1 and 2.
-It is therefore, possible that the moderate value of the neutral weight for block 0 reflects the fact that this block contains a few peripheral conservative and liberal books at the corners of the block.
+However, by inspecting figure \ref{fig:polbooks-graph} we note that there are very few edges between blocks 2 and 3 (the detected conservative and liberal blocks as per figure \ref{fig:school-null}).
+Block 1 effectively acts as a separator between blocks 2 and 3; very few readers venture all the way across the political aisle for their next book.
+It is therefore, possible that the moderate value of the neutral weight for block 1 reflects the fact that this block contains a few peripheral conservative and liberal books at the corners of the block.
 
 From table \ref{tab:results} we see that the training and test losses are very similar (the test loss mean is even below the training loss) and both are quite low in magnitude.
-This is very strong evidence that the features we have available (the author's political affiliation) are indeed the drivers of the overall graphical structure.
+This is very strong evidence that the features we have available (the author's political affiliation) are indeed the drivers of the high-level graphical structure at $B=3$.
 
@@
 However, some blocks have 2 comparably good classes as their predictor.
-For example, blocks 1 and 3 contain classes 3A and 3B as their 2 best predictors.
-This suggests that the social divide between classes is less pronounced for pupils in year 3 but there is nonetheless an unobserved variable that splits year 3 into 2 communities.
+For example, block 2 contains classes 3A and 3B as its 2 best predictors.
+This suggests that the social divide between classes is less pronounced for pupils in year 3.
+Conversely, some classes are found to extend over two detected blocks (class 2B spans blocks 6 and 7) but we nonetheless do not have a feature which explains the division.
 The most surprising block is number 5 - which has comparable weightings for classes 5A and 1B.
@@
 We see that the training loss stays roughly constant between the original and reduced classifiers $\bar{\Lcal}_0 \approx \bar{\Lcal}_0'$ but the test loss of the reduced classifier is ever-so-slightly decreased (within margin of error).
-This would imply that the features we have discarded (gender and teacher/student status) do not impact the high-level graphical structure.
-Nevertheless, the features we have available do not capture the full picture but they do quite well (low overall loss).
+This improved generalisation implies that the features we have discarded (gender and teacher/student status) do not impact the high-level graphical structure and simply lead to over-fitting in the original classifier.
+The features we have available do not capture the entire picture but they do quite well (low overall loss).
 
@@
 
-Nevertheless, this example also highlights a weakness of the method for high-dimensional feature spaces.
+Nevertheless, this example highlights a weakness of the method for high-dimensional feature spaces.
 When the feature dimension is very large, it becomes increasingly likely that a particular feature may uniquely identify a small set of nodes.
 If these nodes are all part of the same community then the classifier will overfit for that particular parameter.
-The regularisation term imposed by the prior goes some way to alleviating this problem but we see in figure \ref{fig:fb-null} that the feature \verb*|birthday-5| has a very high weight as it relates to block 0.
+The regularisation term imposed by the prior goes some way to alleviating this problem but we see in figure \ref{fig:fb-null} that the feature \verb*|birthday-5| has a very high weight as it relates to block \textbf{XX}.
 It might be possible to shift the feature values such that they take values in $\{-1, 1\}$ rather than $\{0, 1\}$ but this approach falls down when input features are mutually correlated.
@@
 This suggests that the features we have available do not readily explain the high-level structure.
-Nonetheless, after reducing the problem dimension our test loss is not greatly increased which suggests that the features we do keep are the best of the ones we have available at predicting the overall network structure.
+Nonetheless, after reducing the problem dimension our test loss is not greatly increased which suggests that the features we do keep are the best from the ones we have available at predicting the overall network structure.

### 2021-05-23 nips/ffbm.tex
Proof-read

@@
 
-In this section we propose a novel generative model for modelling labelled networks.
-We restrict our analysis to labelled, undirected graphs with $N$ nodes.
+In this section we propose a novel generative model for labelled networks.
+As before, we let $N$ denote the number of nodes.
 We define the vector $x_i \in \Xcal^D$ as the feature vector for the $i$'th vertex.
 Each vertex has $D$ total features and we assume all entries take values from the same set $\Xcal$.
-For the majority of datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$.
-The feature vectors $\{x_i\}_{i=1}^{N}$ are subsumed into the $N \times D$ matrix $X$.
+For the datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$.
+The feature vectors $\{x_i\}_{i=1}^{N}$ may be compactly subsumed into the feature matrix $X \in \Xcal^{N \times D}$.
 
 The proposed generative model (which we call the feature-first block model - FFBM) is given in figure \ref{fig:ffbm}.
-We start, with the feature matrix X and generate a vector of block memberships $b$.
+First, we start with the feature matrix X and probabilistically generate a vector of block memberships $b$.
 The parameters of this step are encapsulated by $\theta$.
-Each feature vector $x_i$ is treated independently and used to generate the block membership $b_i$.
-Each entry $b_i \in \{1, 2 \dots B\}$ where $B$ is the number of blocks and a constant in our model.
+Each feature vector $x_i$ is treated independently and used to generate the corresponding block membership $b_i$.
+Each entry $b_i \in [B]$ where $B$ is the number of blocks and a constant in our model.
 We choose a single softmax layer to model $p(b_i | x_i, \theta)$.
-More complex models are possible but then deriving meaning from inferred parameter distributions is more difficult.
+More complex models are possible but then deriving meaning from the inferred parameter distributions is more difficult.
 Summarising, we write $p(b | X, \theta)$ as follows:
@@
 Where $\tilde{x} \coloneqq \left[x_1, x_2, \dots x_D, 1\right]^T$ is an augmented version of $x$ that allows for a bias term.
-The parameters for this stage are denoted $\theta$ and consist of the all $B$ weight vectors $\theta = \{w_k\}_{k=1}^{B}$.
+The parameter vector $\theta$ for this stage contains of the all the weight vectors $\theta = \{w_k\}_{k=1}^{B}$.
 Each $w_k$ has dimension $D+1$.
-We could instead write the parameters $\theta$ as a $B \times (D+1)$ matrix of weights $W$; this form has use computationally as then $Z_i = W \tilde{x}_i$ is the input to the softmax activation function.
+We could instead write the parameters $\theta$ as a $B \times (D+1)$ matrix of weights $W$; this form has computational benefits as then $z_i \coloneqq W \tilde{x}_i$ is the input to the softmax activation function.
 
@@
 
-With the block memberships $b$ generated, we then draw the graph $A$ from the microcanical DC-SBM (\ref{defn:microcan-dc-sbm}) with additional parameters $\psi = \{e, k\}$.
+Once the block memberships $b$ have been generated, we then draw the graph $A$ from the microcanonical DC-SBM (equation \ref{eqn:A-generation}) with additional parameters encapsulated by $\psi = \{e, k\}$.
 In a slight abuse of notation we denote the inter-block edge count matrix with $e = \psi_e$ and the degree sequence $k = \psi_k$ to make explicit that these parameters are contained in $\psi$.
@@
 	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k)
+	\label{eqn:A-generation}
 \end{equation}
@@
 This is a uniform distribution that only depends on the number of blocks $B$.
-This simplifies the problem nicely as evaluating $p(b | X)$ does not require knowing the exact value $X$ takes.
+This is an enormously important simplification as evaluating $p(b | X)$ does not require knowing the exact value $X$ takes and can be performed without an expensive Monte-Carlo integration over the $\theta$-domain.
 \citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for the additional microcanonical SBM parameters $\psi$ which we adopt.
@@
 %
-Where the simplification is made apparent by noting $(\psi \indep X) | b$. We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference.
+Where the simplification is made apparent by noting $(\psi \indep X) | b$. We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference:
 %

### 2021-05-23 nips/inference.tex
Proof-read

@@
 Now that we have defined the FFBM, we wish to leverage it to perform inference.
-Suppose we are presented with a vertex-labelled graph $(A, X)$; the goal is to draw samples for $\theta$ according to the posterior given the observed graph (equation \ref{eqn:theta-target}).
+Suppose we are presented with a vertex-labelled graph $(A, X)$; the goal is to draw samples for $\theta$ according to the posterior given the observed data:
 %
@@
 
-We instead propose an iterative approach.
-First drawing samples $b^{(t)}$ from the block membership posterior (equation \ref{eqn:b-samples}).
-We then use each $b^{(i)}$ to draw samples for $\theta$ as in equation \ref{eqn:theta-samples}.
+We propose an iterative approach.
+We first draw samples $b^{(t)}$ from the block membership posterior (equation \ref{eqn:b-samples}).
+We then use each $b^{(t)}$ to draw samples for $\theta$ as in equation \ref{eqn:theta-samples}.
 %
 \begin{align}
-	b^{(i)} &\sim p \Big( b | A, X \Big)  \label{eqn:b-samples}\\
-	\theta^{(i)} &\sim p\Big(\theta | X, b^{(i)} \Big) \label{eqn:theta-samples}
+	b^{(t)} &\sim p \Big( b | A, X \Big)  \label{eqn:b-samples}\\
+	\theta^{(t)} &\sim p\Big(\theta | X, b^{(t)} \Big) \label{eqn:theta-samples}
 \end{align}
 %
-Both of these sampling steps implemented with a Markov Chain through the Metropolis-Hastings algorithm \cite{hastings-alg}.
+Both of these sampling steps can be implemented with a Markov Chain through the Metropolis-Hastings algorithm \cite{hastings-alg}.
 We just need to define a proposal distribution $q(x, x')$ for proposing a move $x \rightarrow x'$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(\cdot)$, point-wise.
@@
 %
-\begin{align*}
-	\Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] &= \sum_{b \in \mathcal{B}^N} p(\theta | X, b) p(b | A, X) \\
-	&= \sum_{b \in \mathcal{B}^N} p(\theta, b | A, X) \\
-	&= p(\theta | A, X) \\
-\end{align*}
+\begin{equation}
+\Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] = \sum_{b \in [B]^N} p(\theta | X, b) p(b | A, X) = \sum_{b \in [B]^N} p(\theta, b | A, X) = p(\theta | A, X)
+\label{eqn:theta-unbiased}
+\end{equation}
 %
-This is an example of a pseudo-marginal approach \cite{pseudo-marginal}.
-Indeed, the unbiased result is sufficient to prove that for sufficient samples, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{t})\right] = p(\theta
-\ A, X)$ which is exactly the distribution we were targetting (equation \ref{eqn:theta-target}).
+This is an example of a pseudo-marginal approach.
+Indeed, \citet{pseudo-marginal} show that the unbiased result in equation \ref{eqn:theta-unbiased} is sufficient to prove that for sufficient samples, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{t})\right] = p(\theta
+| A, X)$ which is exactly the distribution we are targetting (equation \ref{eqn:theta-target}).
 
-The reason we split the Markov chain into two stages is because the summation over all latent states $b \in \mathcal{B}^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in \Bcal^N} p(A | b) P(b | X, \theta)$ is intractable $O(B^N)$.
+The reason we split the Markov chain into two stages is because the summation over all latent states $b \in [B]^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$ is intractable $O(B^N)$.
 %
@@
 We have introduced subscripts and conditionings to make explicit what parameters each step utilises.
-In an important simplification, we note that $p(b| X) = B^{-N}$ which does not depend on the exact value of X.
-Therefore, we do not need to know the value of $X$ to perform the sampling on $b$.
+We note the power of the simplification given by equation \ref{eqn:b-pseudo-prior}.
+As $p(b| X) = B^{-N}$ does not depend on the exact value of X, we do not need to know the value of $X$ to perform the sampling on $b$.
 Conversely, for the $\theta^{(t)}$ samples, we use $b^{(t)}$ but not $A$ as $(\theta \indep A) | b$.
@@
 %
-Now $\pi_b(\cdot)$ is the un-normalised density we wish to sample from.
+Now $\pi_b(\cdot)$ is the un-normalised density we wish to sample from for the $b$-chain.
 In other words, we wish to construct a Markov chain that has $\pi_b(\cdot)$ as its invariant distribution.
@@
 %
-Since we are using the microcanonical SBM formulation, there is only one value of $\psi$ that is compatible with the given $(A, b)$ pair.
-We denote this value $\psi^* = \{k^*, e^*\}$.
-Specifically, $k^*_i = \sum_j A_{ij} $ and $e^*_{rs} = \sum_{i, j} A_{ij} \one \{b_i=r\} \one\{b_j=s\}$.
-Therefore, the summation over all $\psi$ reduces to just the single $\psi^*$ term.
+Since we are using the microcanonical SBM formulation, there is only one value of $\psi$ that is compatible with the given $(A, b)$ pair (given in equation \ref{eqn:sbm-constraints}).
+We denote this value $\psi^* = \{\psi_k^*, \psi_e^*\}$.
+Therefore, the summation over all $\psi$ reduces to just the single $\psi^*$ term; this is the power of the microcanonical formulation.
 We also define the microcanonical entropy of the configuration as.
@@
 This entropy can be thought of as the description length of the graph because it is the sum of the information required to represent the graph given the parameters and the amount of information required to store the parameters (given the feature matrix $X$).
-The exact, from of the proposal distribution and accept-reject step is explored throughly by \citet{Peixoto-MCMC}.
+The exact form of the proposal distribution and accept-reject step is explored thoroughly by \citet{Peixoto-MCMC} and not repeated here.
 There is a widely used library for Python made available under LGPL called \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, which implements this algorithm.
@@
 %
-Where we have introduced $U(\theta)$ equal to the negative log posterior.
-This $U(\theta)$ term makes subsequent analysis more concise.
-Each of the constituent terms of $U(\theta)$ is easily computed (equation \ref{eqn:U-constituent-terms}).
-To simplify notation, we define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
+Where we have introduced $U(\theta)$ equal to the negative log posterior for conciseness.
+Each of the constituent terms of $U(\theta)$ are easily computed (equation \ref{eqn:U-constituent-terms}).
+Where we have defined $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
 %
 \begin{equation}
-	\log p(b | X, \theta) = \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
+	\log p(b | X, \theta) = \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
 	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta ||^2
@@
 %
-This means that if we ignore constant terms we can write $U(\theta)$ as in equation \ref{eqn:U-form}.
-We see that the prior effectively introduces a regularisation term.
+Discarding constant terms, we can write $U(\theta)$ as in equation \ref{eqn:U-form}.
 Note that $||\theta||^2 = \sum_{i} \theta_{i}^2 = \sum_{j=1}^{B} ||w_j||^2$ is the Euclidean norm of the vector of parameters $\theta$.
@@
 $U(\theta)$ in equation \ref{eqn:U-form} appears a typical objective function to be minimised for neural network training.
-The first term -- which we collect into $N \cdot \Lcal(\theta)$ -- is the cross-entropy between graph-predicted and feature-predicted block memberships.
-We denote the average cross-entropy $\Lcal(\theta)$ as it will become useful later on when performing experiments.
-The second term of equation \ref{eqn:U-form} -- introduced by the prior -- brings a form of regularisation, preventing over-fitting.
+The first term -- which we collect into $N \cdot \Lcal(\theta)$ -- is the cross-entropy between the graph-predicted and feature-predicted block memberships.
+We denote the average cross-entropy $\Lcal(\theta)$ as it will become useful later on for experimentation.
+The second term of equation \ref{eqn:U-form} -- introduced by the prior -- brings a form of regularisation, guarding against over-fitting.
 In traditional applications we only seek the value of $\theta$ that minimises the objective function $U(\theta)$, which in our case would yield the maximum a posteriori (MAP) estimate.
@@
 However, our goal is not to find the MAP estimate but to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
-As discussed earlier, given the invariant $\pi_\theta(\cdot)$, it is sufficient to specify a proposal distribution and then apply a MH accept-reject step to ensure detailed balance of the Markov Chain.
+As discussed earlier, given the invariant distribution $\pi_\theta(\cdot)$, it is sufficient to specify a proposal distribution and then apply a MH accept-reject step to ensure detailed balance of the Markov Chain.
 Nevertheless, we can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
-We therefore adopt the Metropolis Adjusted Langevin Algorithm (MALA) -- first proposed by \citet{mala-tweedie} -- which leverages just that.
+We therefore adopt the Metropolis Adjusted Langevin Algorithm (MALA) -- first proposed by \citet{mala-tweedie} -- which leverages the $\nabla U$ information.
 Given the current sample $\theta$, we propose a new sample $\theta'$ according to equation \ref{eqn:theta-update}
@@
 %
-Where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter - which may vary with the sample index.
-Indeed without the injected noise term, this is equivalent to gradient descent.
+Where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter -- which may vary with the sample index (appendix \ref{appdx:step-size} explores this more fully).
+Without the injected noise term, MALA is equivalent to gradient descent.
 We require the noise term to fully explore the parameter space.
@@
 %
-This fully specifies, the sampling procedure to generate $\{\theta^{(t)}\}_{t=1}^T$. Bear in mind that each $\theta^{(t)}$ update step uses its corresponding $b^{(t)}$ block membership sample.
+This fully specifies, the sampling procedure to generate $\{\theta^{(t)}\}_{t=1}^T$. Bear in mind that so far, each $\theta^{(t)}$ update step uses its corresponding $b^{(t)}$ block membership sample.
 
@@
 
-So far, each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample.
+Up to this point, each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample.
 This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance.
@@
 %
-We obtain an unbiased estimate for this quantity as simply the empirical distribution of the block membership samples $\left\{ b^{(t)} \right\}_{t=1}^T$.
+We can obtain an unbiased estimate for this quantity using the block membership samples $\left\{ b^{(t)} \right\}$.
+However, as with all MCMC methods, we must only uses samples after a burn-in and thinning has been applied.
+We introduce $\Tcal_b$ to denote the retained set of indices for the $b$-samples and $\Tcal_\theta$ similarly for the $\theta$-chain.
+An in-depth discussion of how these sets are chosen is given in appendix \ref{appdx:burn-in-thinning}.
+The unbiased estimate for $y_{ij}^{(t)}$ using the restricted sample set $\Tcal_b$ is denoted $\hat{y}_{ij}$ and has form:
 %
@@
 This means we no longer need to run the $b$ and $\theta$ Markov chains concurrently.
-Instead, we run the $b$-chain to completion and use it to generate $\hat{y}_{ij}$ for $i \in \{1 \dots N\}$ and $j \in \{1 \dots B\}$.
+Instead, we run the $b$-chain to completion and use it to generate $\hat{y}_{ij}$ for $i \in [N]$ and $j \in [B]$.
 This is an estimate of $p(b | A, X)$ that we use for every iteration of the $\theta$ Markov chain.
-This affords us the flexibility to vary the number of samples we draw for $b$ and $\theta$; we refer to these as $T_b$ and $T_\theta$ henceforth.
+
+This affords us the flexibility to vary the number of samples we draw for the $b$ and $\theta$-chains; we denote the total number of samples generated $T_b$ and $T_\theta$ henceforth.
+Note that due to burn-in and thinning, $|\Tcal_b| < T_b$ and $|\Tcal_\theta| < T_\theta$.
 Furthermore, this changeover reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
-
-A more fleshed-out discussion of how we choose to thin our samples is given in appendix \ref{appdx:burn-in-thinning}
 
@@
 Intuitively, this means discarding any feature for which $\hat{\mu}_{ij} \pm k\hat{\sigma}_{ij}$ lies within or spans the null region $(-c, c)$ for all block indices.
-If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \mu_{ij}, \sigma_{ij})$, then this is effectively a hypothesis test on the value of $W_{ij}$.
+If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \mu_{ij}, \sigma_{ij})$, then this is effectively a hypothesis test on the value of $W_{ij}$ (equation \ref{eqn:hyp-test-discard}).
 $\Dcal'$ then comprises all features for which $H_1$ is accepted.
@@
 	\end{aligned}
+	\label{eqn:hyp-test-discard}
 \end{equation}
 %
-The multiplier $k$ specifies the degree of significance of the result (equivalent to the probability of a type-1 error).
-However, as the approximation is not exact we will only treat this dimensionality reduction method as a useful heuristic and not an exact method.
+The multiplier $k$ determines the degree of significance of the result.
+However, as the Laplace approximation is not exact we will only treat this dimensionality reduction method as a useful heuristic and not an exact method.
 
@@
 This is summarised in equation \ref{eqn:c-star}.
-This approach may be preferred as it fixes the number of reduced dimensions.
+This approach is often preferred as it fixes the number of reduced dimensions.
 %

### 2021-05-23 nips/prelim.tex
Proof-read

@@
 \begin{definition}[Microcanonical DC-SBM]
-	\label{defn:microcan-ndc-sbm}
+	\label{defn:microcan-dc-sbm}
 	Let $N \in \Integers^{+}$ denote the number of vertices in an undirected graph.
-The block memberships are encoded by a vector $b \in [B]^N$,\footnote{We introduce the notation $[K] \coloneqq \{0, 1, 2 \dots K-1\}$ to compactly define a set of $K$ indices.
+The block memberships are encoded by a vector $b \in [B]^N$,\footnote{We introduce the notation $[K] \coloneqq \{1, 2 \dots K\}$ to compactly define a set of $K$ indices.
 Clearly, $[K]$ is only defined for $K \in \Integers^+$.}
@@
 		k_i = \sum_j A_{ij}
+		\label{eqn:sbm-constraints}
 	\end{equation}
@@
 \end{definition}
+
+In what follows, the SBM will always refer to the microcanonical DC-SBM variation.

### 2021-05-23 nips/appendices.tex
10 pages

@@
 \section{Appendix}
+
+\subsection{SBM prior choice explanation}
+\label{appdx:prior}
+
+We have borrowed the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$. These, are repeated here for reference:
+%
+\begin{equation}
+	p(\psi | b) = p(e | b) p(k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
+	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right]
+\end{equation}
+%
+Where $\specialchoose{n}{m}$ is shorthand for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of distinct histograms with $n$ bins under the constraint they sum to $m$.
+$E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges in the graph.
+Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform with respect to $e$.
+The variable $\eta_j^r$ is introduced to denote the number of vertices in block $r$ that have degree $j$.
+Formally, $\eta_j^r \coloneqq \sum_{i} \one\left\{b_i = r \right\} \one \left\{k_i = j \right\}$.
+Furthermore, $q(m, n)$ is the number of different histograms with at most $n$ non-zero bins that sum to $m$.
+$q(m, n)$ is related to but different from $\specialchoose{n}{m}$.
+Lastly, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ is the number of vertices assigned to block $r$.
+
+These were chosen carefully by \citet{Peixoto-Bayesian-Microcanonical} to more closely match the structure of empirical networks than simple uniform priors. We do not repeat his arguments here.
 
@@
 \subsection{Algorithms}
+\label{appdx:algorithms}
 
@@
 
-\clearpage
+\FloatBarrier
 \subsection{Hyperparameter values}

### 2021-05-23 nips/experiments.tex
10 pages

@@
 Two nodes are connected if the two parties shared a face-to-face interaction over the course of the day.
-Vertex features include class membership, gender and whether or not the individual is a teacher or pupil.
+Vertex features include class membership (one of 10 values 1A-5B), gender and whether or not the individual is a teacher or pupil.
 These data were collected on consecutive days in October 2009.
@@
 \end{itemize}
-
-We require metrics to assess the relative model evidence for the FFBM.
-This can be split into two separate components: the microcanonical SBM fit (concerned with the $b$-samples) and the fit of the feature-to-block generator (concerned with the $\theta$-samples).
-Starting with the SBM, $S(b)$ (equation \ref{eqn:dl-form}) can be interpreted as the description length of the partition imposed by $b$.
-In other words this is the amount of information needed to describe the graph and the associated SBM parameters.
-It is only natural to divide this quantity by the number of entities (nodes and edges) in our graph $N+E$ to allow for rough comparison between graphs.
-This defines a simple metric to gauge the fit of the SBM: the description length per entity averaged over the $b$-samples (equation \ref{eqn:mean-dl}):
-%
-\begin{equation}
-	\bar{S}_e \coloneqq \frac{1}{(N+E) |\Tcal_b|} \sum_{t\in \Tcal_b} S \left( b^{(t)} \right)
-	\label{eqn:mean-dl}
-\end{equation}
-%
-However, to assess the performance of the feature-to-block predictor, we must partition the vertex set $[N]$ into training and test sets -- $\Gcal_0$ and $\Gcal_1$ respectively
-\footnote{We choose to randomly create the partition on each experiment run such that a constant fraction $f$ of the available vertices go to form our training set $\Gcal$ and the remaining vertices are held out to form our test set $\Gcal^C$.
-For all the experiments $f=0.7$.}
-.
-The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal$ to train the $\theta$-chain.
-As $|\Gcal_0| \neq |\Gcal_1|$ sets, we cannot use the un-normalised log target $U$ (equation \ref{eqn:U-form}) for comparison as the total cross-entropy loss is scaled by the size of each set -- even though the prior term stays constant.
-We therefore must use the average cross-entropy loss over each set (equation \ref{eqn:cross-entropy-loss}).
-%
-\begin{equation}
-	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star \left( \theta^{(t)} \right)
-	\qquad \textrm{where} \qquad
-	\Lcal_\star \left( \theta^{(t)} \right) \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)}
-	\label{eqn:cross-entropy-loss}
-\end{equation}
-%
-Where $\star \in \{0, 1\}$ has been introduced to toggle between training and test sets.
-Table \ref{tab:results} summarises the results for each experiment.
-We run each experiment $n=10$ times with the same hyper-parameter values\footnote{For a comprehensive list of the hyper-parameters used for each experiment please see appendix \ref{appdx:hyperparams}} and report the mean plus or minus one standard deviation.
-
-However, we also wish to determine the power of the dimensionality reduction method.
-For this we apply equation \ref{eqn:c-star}, to reduce the dimension from $D$ to $D'$ with $k=1$.
-We then retrain the feature-block predictor using just the retained feature set $\Dcal'$ and report the loss over the training and test sets for the reduced classifier -- denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively.
-These values are also included in table \ref{tab:results} except for the polbooks dataset which is already very low dimension.
-
-\begin{table}[!h]
-	\centering
-	\caption{Experimental results averaged over $n=10$ iterations (mean $\pm$ standard deviation)}
-	\label{tab:results}
-	\resizebox{\textwidth}{!}{%
-	\begin{tabular}{c|ccc|c|cc|ccc}
-		Dataset  & $B$ & $D$ & $D'$ & $\bar{S}_e$ & $\bar{\Lcal}_0$ & $\bar{\Lcal}_1$ & $c^*$ & $\bar{\Lcal}_0'$ & $\bar{\Lcal}_1'$  \\ \hline
-		Polbooks & 3 & 3 & -- & $2.250 \pm 0.001$ & $0.584 \pm 0.033$ & $0.557 \pm 0.061$ & -- & -- & -- \\
-		School & 10 & 13 & 10 & $1.894 \pm 0.006$ & $0.793 \pm 0.096$ & $0.914 \pm 0.112$ & $1.112 \pm 0.230$ & $0.791 \pm 0.096$ & $0.864 \pm 0.139$ \\
-		FB egonet & 10  & 480 & 10 & $1.626 \pm 0.003$ & $1.305 \pm 0.034$ & $1.539 \pm 0.087$ & $0.942 \pm 0.042$ & $1.496 \pm 0.093$ & $1.578 \pm 0.104$
-	\end{tabular}
-	}
-\end{table}
-
-Table \ref{tab:results} already highlights some general trends in the results.
-Firstly, the variance of the test loss $\bar{\Lcal}_1$ tends to be higher than the training loss $\bar{\Lcal}_0$.
-This is expected as our test set is smaller than the training set and so more susceptible to variability in its construction.
-Indeed, much of the variance in the evaluation of $\bar{\Lcal}_0$ and $\bar{\Lcal}_1$ comes from the random partitioning of the graph into training and test sets $\Gcal_0$ and $\Gcal_1$.
-Secondly, it can be seen that the dimensionality reduction procedure brings the training and test losses closer together.
-This implies that the features we keep are indeed correlated with the underlying graphical partition.
-
-The average description length per entity of the graph $\bar{S}_e$ has very low variance implying the detected communities can be found reliably (to within an arbitrary relabelling of blocks).
-For reference we plot an inferred partition for each of the graphs on figure \ref{fig:graphs-all}.
-The polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do succeed at partitioning the graph into densely connected clusters.
 
@@
 \end{figure}
+
+We require metrics to assess performance.
+This can be split into two separate components: the microcanonical SBM fit (concerned with the $b$-samples) and the fit of the feature-to-block generator (concerned with the $\theta$-samples).
+Starting with the SBM, $S(b)$ (equation \ref{eqn:dl-form}) can be interpreted as the description length of the partition imposed by $b$.
+It is only natural to divide this quantity by the number of entities (nodes and edges) in our graph $N+E$ to allow for rough comparison between graphs.
+This defines a simple metric to gauge the fit of the SBM: the description length per entity averaged over the $b$-samples (equation \ref{eqn:mean-dl}):
+%
+\begin{equation}
+	\bar{S}_e \coloneqq \frac{1}{(N+E) |\Tcal_b|} \sum_{t\in \Tcal_b} S \left( b^{(t)} \right)
+	\label{eqn:mean-dl}
+\end{equation}
+%
+However, to assess the performance of the feature-to-block predictor, we must partition the vertex set $[N]$ into training and test sets: $\Gcal_0$ and $\Gcal_1$ respectively.\footnote{We choose to randomly create the partition on each experiment run such that a constant fraction $f$ of the available vertices go to form our training set $\Gcal_0$ and the remaining vertices are held out to form our test set $\Gcal_1$.
+For all the experiments $f=0.7$.}
+The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain.
+As $|\Gcal_0| \neq |\Gcal_1|$ sets, we cannot use the un-normalised log target $U$ (equation \ref{eqn:U-form}) for comparison; the total cross-entropy loss is scaled by the size of each set but the prior term stays constant.
+We therefore must use the average cross-entropy loss over each set (equation \ref{eqn:cross-entropy-loss}):
+%
+\begin{equation}
+	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star \left( \theta^{(t)} \right)
+	\qquad \textrm{where} \qquad
+	\Lcal_\star \left( \theta^{(t)} \right) \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)}
+	\label{eqn:cross-entropy-loss}
+\end{equation}
+%
+Where $\star \in \{0, 1\}$ has been introduced to toggle between training and test sets.
+Table \ref{tab:results} summarises the results for each experiment.\footnote{For a comprehensive list of the hyper-parameters used for each experiment please see appendix \ref{appdx:hyperparams}}
+
+We also apply the dimensionality reduction method on the two higher dimensional datasets (the school and FB egonet).
+For this we leverage equation \ref{eqn:c-star}, to reduce the dimension from $D$ to $D'$ with $k=1$ to yield $c^*$.
+We then retrain the feature-block predictor using just the retained feature set $\Dcal'$ and report the loss over the training and test sets for the reduced classifier -- denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively.
+These values are also included in table \ref{tab:results}.
+
+\begin{table}[!h]
+	\centering
+	\caption{Experimental results averaged over $n=10$ iterations (mean $\pm$ standard deviation)}
+	\label{tab:results}
+	\resizebox{\textwidth}{!}{%
+	\begin{tabular}{c|ccc|c|cc|ccc}
+		Dataset  & $B$ & $D$ & $D'$ & $\bar{S}_e$ & $\bar{\Lcal}_0$ & $\bar{\Lcal}_1$ & $c^*$ & $\bar{\Lcal}_0'$ & $\bar{\Lcal}_1'$  \\ \hline
+		Polbooks & 3 & 3 & -- & $2.250 \pm 0.001$ & $0.584 \pm 0.033$ & $0.557 \pm 0.061$ & -- & -- & -- \\
+		School & 10 & 13 & 10 & $1.894 \pm 0.006$ & $0.793 \pm 0.096$ & $0.914 \pm 0.112$ & $1.112 \pm 0.230$ & $0.791 \pm 0.096$ & $0.864 \pm 0.139$ \\
+		FB egonet & 10  & 480 & 10 & $1.626 \pm 0.003$ & $1.305 \pm 0.034$ & $1.539 \pm 0.087$ & $0.942 \pm 0.042$ & $1.496 \pm 0.093$ & $1.578 \pm 0.104$
+	\end{tabular}
+	}
+\end{table}
+
+Table \ref{tab:results} already highlights some general trends in the results.
+Firstly, the variance of the test loss $\bar{\Lcal}_1$ tends to be higher than the training loss $\bar{\Lcal}_0$.
+This is expected as our test set is smaller than the training set and so more susceptible to variability in its construction.
+Indeed, much of the variance in the evaluation of $\bar{\Lcal}_0$ and $\bar{\Lcal}_1$ comes from the random partitioning of the graph into training and test sets $\Gcal_0$ and $\Gcal_1$.
+Secondly, it can be seen that the dimensionality reduction procedure brings the training and test losses closer together.
+This implies that the features we keep are indeed correlated with the underlying graphical partition.
+
+The average description length per entity of the graph $\bar{S}_e$ has very low variance implying the detected communities can be found reliably (to within an arbitrary relabelling of blocks).
+For reference we plot an inferred partition for each of the graphs on figure \ref{fig:graphs-all}.
+The polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do succeed at partitioning the graph into densely connected clusters.
 
@@
 
-This dataset was collected by \citet{polbooks}.
 We wish to determine whether political affiliation is a good predictor of the overall network structure.
 We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral).
-The inferred block memberships are given in figure \ref{fig:polbooks-graph}.
-We sample the block-generator parameters and plot the emprical mean and standard deviation of each sample on figure \ref{fig:polbooks-null}.
 
-Indeed, from figure \ref{fig:polbooks-null}, we see that all 3 blocks have a distinct political affiliation as their largest positive component.
+From, figure \ref{fig:polbooks-null}, we see that all 3 blocks have a distinct political affiliation as their largest positive component.
 This is strong evidence that political affiliation is indeed the axis which best predicts the 3-way natural partition of the graph into blocks.
-Nevertheless, for block index 1, although \verb*|neutral| is the best predictor of the ones we have available, the value is not particularly extreme.
-Perhaps, there is an unobserved feature that places a book into this block.
-However, by inspecting figure \ref{fig:polbooks-graph} we note that there are very few edges between blocks 2 and 3 (the detected conservative and liberal blocks as per figure \ref{fig:school-null}).
+
+From figure \ref{fig:polbooks-graph}, we note that there are very few edges between blocks 2 and 3 (the detected conservative and liberal blocks as per figure \ref{fig:school-null}).
 Block 1 effectively acts as a separator between blocks 2 and 3; very few readers venture all the way across the political aisle for their next book.
-It is therefore, possible that the moderate value of the neutral weight for block 1 reflects the fact that this block contains a few peripheral conservative and liberal books at the corners of the block.
 
-From table \ref{tab:results} we see that the training and test losses are very similar (the test loss mean is even below the training loss) and both are quite low in magnitude.
+From table \ref{tab:results} we see that the training and test losses are very similar (the test loss mean is even below the training loss) and both are low in magnitude.
 This is very strong evidence that the features we have available (the author's political affiliation) are indeed the drivers of the high-level graphical structure at $B=3$.
@@
 \subsection{Primary school dynamic contacts}
-
-These data were originally collected by \citet{schools} to quantify the transmission opportunities for respiratory infections.
-However, we seek to ask which vertex features best describe how people interact with one another in a primary school context.
-The only vertex features we have available are school-class (one of 10 values - 2 per year group), gender and a distinction between teachers and pupils.
 
@@
 
-As before, we proceed to sample the block-generator parameters $\theta$ and employ the dimensionality reduction heuristic with standard deviation multiplier $k=1$ to pick out the top $D'=10$ features.
+As before, we sample the block-generator parameters $\theta$ and employ the dimensionality reduction technique with standard deviation multiplier $k=1$ to pick out the top $D'=10$ features.
 We then plot the weights for the surviving features $d \in \Dcal'$ on figure \ref{fig:school-null}.
@@
 This improved generalisation implies that the features we have discarded (gender and teacher/student status) do not impact the high-level graphical structure and simply lead to over-fitting in the original classifier.
-The features we have available do not capture the entire picture but they do quite well (low overall loss).
+The features we have available do not capture the entire picture but they do quite well (modestly low overall loss).
 
@@
 
-We again apply the dimensionality reduction heuristic with $k=1$ to target $D'=10$ features in the reduced set.
-The parameter values for the surviving features are plotted on figure \ref{fig:fb-null}.
-The remaining features are those that best explain the high-level community structure.
+We again apply the dimensionality reduction technique with $k=1$ to target $D'=10$ features in the reduced set.
+The remaining features (figure \ref{fig:fb-null}) are those that best explain the high-level community structure.
 The majority of the surviving features are education related.
-More so than language, gender or surname - the school\footnote{This includes higher education institutes} you attend is the best explanatory variable for how the observed network's communities are partitioned at the highest level.
-Nevertheless, for $D'=10$ we do not have good explanations for the makeup of blocks 5-9 inclusive
+Nevertheless, for $D'=10$ we only have good explanations for the makeup of some of the detected blocks.
 
-Nevertheless, this example highlights a weakness of the method for high-dimensional feature spaces.
 When the feature dimension is very large, it becomes increasingly likely that a particular feature may uniquely identify a small set of nodes.
 If these nodes are all part of the same community then the classifier will overfit for that particular parameter.
-The regularisation term imposed by the prior goes some way to alleviating this problem but we see in figure \ref{fig:fb-null} that the feature \verb*|birthday-5| has a very high weight as it relates to block \textbf{XX}.
-It might be possible to shift the feature values such that they take values in $\{-1, 1\}$ rather than $\{0, 1\}$ but this approach falls down when input features are mutually correlated.
-Indeed, for features with discrete values such as the class memberships (1A-5B) of the school dataset it is preferred to keep $x_i \in \{0, 1\}^D$ as then each block can be readily accept more than one mutually exclusive feature-group (i.e.
-two school-classes are part of the same block).
+The regularisation term imposed by the prior goes some way to alleviating this problem.
+Nevertheless, we see in figure \ref{fig:fb-null} that the feature \verb*|birthday-5| has a very high weight as it relates to block 1.
+It might be possible to shift the feature values such that they take values in $\{-1, 1\}$ rather than $\{0, 1\}$ but to accept more than one feature-group per block.
 
-The training and test losses (table \ref{tab:results}) are the highest of the considered networks.
-This suggests that the features we have available do not readily explain the high-level structure.
-Nonetheless, after reducing the problem dimension our test loss is not greatly increased which suggests that the features we do keep are the best from the ones we have available at predicting the overall network structure.
+The training and test losses (table \ref{tab:results}) are the highest of the considered networks but nonetheless still low.
+This suggests that the features we have available only partly explain the high-level structure.
+Promisingly, after reducing the problem dimension our test loss barely changes, suggesting that the features we do keep are the best from the ones we have available at explaining the overall network structure.

### 2021-05-23 nips/ffbm.tex
10 pages

@@
 %
-Where the simplification is made apparent by noting $(\psi \indep X) | b$. We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$, repeated here for reference:
-%
-\begin{equation}
-	p(\psi | b) = p(e | b) p(k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
-	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right]
-\end{equation}
-%
-Where $\specialchoose{n}{m}$ is shorthand for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of distinct histograms with $n$ bins under the constraint they sum to $m$.
-$E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges in the graph.
-Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform with respect to $e$.
-The variable $\eta_j^r$ is introduced to denote the number of vertices in block $r$ that have degree $j$.
-Formally, $\eta_j^r \coloneqq \sum_{i} \one\left\{b_i = r \right\} \one \left\{k_i = j \right\}$.
-Furthermore, $q(m, n)$ is the number of different histograms with at most $n$ non-zero bins that sum to $m$.
-$q(m, n)$ is related to but different from $\specialchoose{n}{m}$.
-Lastly, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ is the number of vertices assigned to block $r$.
-Importantly, we have computable forms for $p(\theta)$ and $p(b, \psi | X)$ which will be useful for performing inference.
+Where the simplification is made apparent by noting $(\psi \indep X) | b$.
+We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$ to complete our model.
+Please refer to appendix \ref{appdx:prior} for the exact form of $p(\psi | b)$.
+All that concerns us is that we have a form which is computable.

### 2021-05-23 nips/inference.tex
10 pages

@@
 | A, X)$ which is exactly the distribution we are targetting (equation \ref{eqn:theta-target}).
-
-The reason we split the Markov chain into two stages is because the summation over all latent states $b \in [B]^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$ is intractable $O(B^N)$.
 %
@@
 	\centering
+
 	\begin{tikzpicture}[
+		scale=0.7, every node/.style={transform shape},
 		roundnode/.style={circle, draw=black, minimum size=12mm},
@@
 		\node[roundnode] (b0) at (0, 2) {$b^{(t)}$};
-		\node[roundnode] (b1) at (4, 2) {$b'$};
-		\node[roundnode] (b2) at (8, 2) {$b^{(t+1)}$};
+		\node[roundnode] (b1) at (5, 2) {$b'$};
+		\node[roundnode] (b2) at (10, 2) {$b^{(t+1)}$};
 		\node[roundnode] (t0) at (0, 0) {$\theta^{(t)}$};
-		\node[roundnode] (t1) at (4, 0) {$\theta'$};
-		\node[roundnode] (t2) at (8, 0) {$\theta^{(t+1)}$};
+		\node[roundnode] (t1) at (5, 0) {$\theta'$};
+		\node[roundnode] (t2) at (10, 0) {$\theta^{(t+1)}$};
 
@@
 		\draw[dashed, ->] (b0) to (2, 0.8);
-		\draw[dashed, ->] (b0) to (5.5, 0.8);
+		\draw[dashed, ->] (b0) to (6.5, 0.8);
 
@@
 \end{figure}
-%
+
+The reason we split the Markov chain into two stages is because the summation over all latent states $b \in [B]^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$ is intractable $O(B^N)$.
+
 Figure \ref{fig:samp-sequence} shows an overview of the proposed method.
@@
 As $p(b| X) = B^{-N}$ does not depend on the exact value of X, we do not need to know the value of $X$ to perform the sampling on $b$.
-Conversely, for the $\theta^{(t)}$ samples, we use $b^{(t)}$ but not $A$ as $(\theta \indep A) | b$.
+Conversely, for the $\theta^{(t)}$ samples, we use only $b^{(t)}$ but not $A$ as $(\theta \indep A) | b$.
+
+\FloatBarrier
 
@@
 %
-\begin{align*}
-	\pi_b(b) &= p(b|X) \sum_{\psi} \nolimits p(A , \psi | b, X) \\
-	&= p(b|X) p(A, \psi^* | b, X) \\
-	&= p(A | b, \psi^*) \cdot p(\psi^* | b) \cdot p(b | X)
-\end{align*}
+\begin{equation}
+	\pi_b(b) = p(b|X) \sum_{\psi} \nolimits p(A , \psi | b, X) \\
+	= p(b|X) p(A, \psi^* | b, X) \\
+	= p(A | b, \psi^*) \cdot p(\psi^* | b) \cdot p(b | X)
+\end{equation}
 %
@@
 %
-This entropy can be thought of as the description length of the graph because it is the sum of the information required to represent the graph given the parameters and the amount of information required to store the parameters (given the feature matrix $X$).
+This entropy can equally be thought of as the description length of the graph -- the minimum amount of information required to store the graph and its parameters.
 The exact form of the proposal distribution and accept-reject step is explored thoroughly by \citet{Peixoto-MCMC} and not repeated here.
@@
 The only modification we make is in the block membership prior $p(b)$ which we replace with $p(b|X)=B^{-N}$ which is a uniform distribution and so cancels out in the MH accept-reject step.
-
-The end result of this is that we can generate a set of block membership samples $\{b^{(t)}\}_{t=1}^{T}$ with each $b^{(t)} \sim p(b | A, X)$.
-Each of these samples can then be used for the $\theta$-chain.
 
@@
 %
-Where we have introduced $U(\theta)$ equal to the negative log posterior for conciseness.
+Where we have introduced $U(\theta)$ equal to the negative log posterior.
 Each of the constituent terms of $U(\theta)$ are easily computed (equation \ref{eqn:U-constituent-terms}).
-Where we have defined $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
+Similarly, we define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
 %
@@
 %
-Discarding constant terms, we can write $U(\theta)$ as in equation \ref{eqn:U-form}.
+Discarding constant terms, we write $U(\theta)$ as in equation \ref{eqn:U-form}.
 Note that $||\theta||^2 = \sum_{i} \theta_{i}^2 = \sum_{j=1}^{B} ||w_j||^2$ is the Euclidean norm of the vector of parameters $\theta$.
@@
 %
-$U(\theta)$ in equation \ref{eqn:U-form} appears a typical objective function to be minimised for neural network training.
-The first term -- which we collect into $N \cdot \Lcal(\theta)$ -- is the cross-entropy between the graph-predicted and feature-predicted block memberships.
-We denote the average cross-entropy $\Lcal(\theta)$ as it will become useful later on for experimentation.
+$U(\theta)$ in equation \ref{eqn:U-form} appears a typical objective function for neural network training.
+The first term -- which we collect into $N \cdot \Lcal(\theta)$ -- is the cross-entropy between the graph-predicted and feature-predicted block memberships summed over all vertices.
 The second term of equation \ref{eqn:U-form} -- introduced by the prior -- brings a form of regularisation, guarding against over-fitting.
@@
 However, our goal is not to find the MAP estimate but to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
-As discussed earlier, given the invariant distribution $\pi_\theta(\cdot)$, it is sufficient to specify a proposal distribution and then apply a MH accept-reject step to ensure detailed balance of the Markov Chain.
-Nevertheless, we can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
+It is sufficient to specify a proposal distribution and then apply a MH accept-reject step to ensure detailed balance with $\pi_\theta(\cdot)$.
+However, we can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
 We therefore adopt the Metropolis Adjusted Langevin Algorithm (MALA) -- first proposed by \citet{mala-tweedie} -- which leverages the $\nabla U$ information.
-Given the current sample $\theta$, we propose a new sample $\theta'$ according to equation \ref{eqn:theta-update}
+Given the current sample $\theta$, we propose a new sample $\theta'$ according to:
 %
@@
 Without the injected noise term, MALA is equivalent to gradient descent.
-We require the noise term to fully explore the parameter space.
-As such the proposal distribution, is a simple multivariate Gaussian which can be easily evaluated.
+We require the noise term $\xi$ to fully explore the parameter space.
+We can write the proposal distribution simply as:
 %
@@
 %
-After a proposed move is generated, in typical Metropolis-Hastings fashion we accept the move with probability $\alpha_\theta$.
+After a proposed move is generated, in typical Metropolis-Hastings fashion we accept the move with probability $\alpha_\theta$:
 %
@@
 \end{equation}
-%
-This fully specifies, the sampling procedure to generate $\{\theta^{(t)}\}_{t=1}^T$. Bear in mind that so far, each $\theta^{(t)}$ update step uses its corresponding $b^{(t)}$ block membership sample.
 
@@
 This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance.
-This may lead to longer burn-in and autocorrelation times of the resulting Markov Chain.
-The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends only on $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
+This may lead to longer burn-in for the resulting Markov Chain.
+The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends only on the matrix $y^{(t)}$ with entries $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
 We would rather deal with the expectation of each $y_{ij}^{(t)}$:
@@
 We can obtain an unbiased estimate for this quantity using the block membership samples $\left\{ b^{(t)} \right\}$.
-However, as with all MCMC methods, we must only uses samples after a burn-in and thinning has been applied.
+However, as with all MCMC methods, we must only uses samples after burn-in and thinning have been applied.
 We introduce $\Tcal_b$ to denote the retained set of indices for the $b$-samples and $\Tcal_\theta$ similarly for the $\theta$-chain.
@@
 %
-We therefore, choose to feed each $\theta^{(t)}$ update step the same $\hat{y}_{ij}$ for all $t$ rather than the corresponding $y^{(t)}_{ij}$.
+We therefore, choose to feed each $\theta^{(t)}$ update step the same matrix $\hat{y}$ for all $t$ rather than the corresponding $y^{(t)}$.
 This means we no longer need to run the $b$ and $\theta$ Markov chains concurrently.
-Instead, we run the $b$-chain to completion and use it to generate $\hat{y}_{ij}$ for $i \in [N]$ and $j \in [B]$.
+Instead, we run the $b$-chain to completion and use it to generate $\hat{y}$.
 This is an estimate of $p(b | A, X)$ that we use for every iteration of the $\theta$ Markov chain.
@@
 Note that due to burn-in and thinning, $|\Tcal_b| < T_b$ and $|\Tcal_\theta| < T_\theta$.
-Furthermore, this changeover reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
+Furthermore, this changeover from $y^{(t)}$ to $\hat{y}$ reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
+A description of the overall algorithms is given in appendix \ref{appdx:algorithms}.
 
@@
 A simple heuristic to discard the least important features requires specifying a cutoff $c > 0$ and a multiplier $k > 0$.
-We define the function $\Fcal_i(j)$ as in \ref{eqn:fij} then only keep features in $\Dcal'$ given by equation \ref{eqn:kept-feature-set}.
+We define the function $\Fcal_i(j)$ as in \ref{eqn:fij} then only keep features with indices $d \in \Dcal'$, where $\Dcal'$ is constructed as in equation \ref{eqn:kept-feature-set}.
 %
-\begin{equation}
-	\Fcal_i(j) \coloneqq (\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c)
-	\label{eqn:fij}
-\end{equation}
-\begin{equation}
-	\Dcal' = \left\{ j \in [D] : \exists i \in [B] \textrm{ s.t. }  \Fcal_i(j) \neq \emptyset \right\}
+\begin{align}
+	\Fcal_i(j) &\coloneqq (\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c)
+	\label{eqn:fij} \\
+	\Dcal' &\coloneqq \left\{ j \in [D] : \exists i \in [B] \textrm{ s.t. }  \Fcal_i(j) \neq \emptyset \right\}
 	\label{eqn:kept-feature-set}
-\end{equation}
+\end{align}
 %
@@
 If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \mu_{ij}, \sigma_{ij})$, then this is effectively a hypothesis test on the value of $W_{ij}$ (equation \ref{eqn:hyp-test-discard}).
-$\Dcal'$ then comprises all features for which $H_1$ is accepted.
+$\Dcal'$ then comprises all features $i$ for which $H_1$ is accepted at least once for some $j \in [B]$.
 %
 \begin{equation}
-	\begin{aligned}
-		H_0&: |W_{ij}| \leq c \\
-		H_1&: |W_{ij}| > c
-	\end{aligned}
+	H_0: |\mu_{ij}| \leq c \qquad
+	H_1: |\mu_{ij}| > c
 	\label{eqn:hyp-test-discard}

### 2021-05-23 nips/introduction.tex
10 pages

@@
 One of the most common questions we can ask of labelled networks is whether a given vertex feature has an impact on the structure of the graph.
-To answer this question from a Bayesian perspective we must use a random graph model; the most common form is called the stochastic block model (SBM) \cite{vanilla-sbm}.
+To answer this question from a Bayesian perspective we must use a random graph model; the most common form is the stochastic block model (SBM) \cite{vanilla-sbm}.
 This is a latent variable model where each vertex belongs to a single block and the probability two nodes are connected depends only on the block memberships of each.
@@
 The associated model can then be used to test for evidence of heterogeneous connectivity between the feature-grouped blocks.
-Nevertheless, this approach is limited in that it can only consider one feature at a time; this means any conclusions drawn highly vulnerable to the presence of confounding variables.
+Nevertheless, this approach is limited in that it can only consider one feature at a time; this leaves any conclusions drawn highly vulnerable to the presence of confounding variables.
 Furthermore, considering each feature separately makes it difficult to rank order the features by  magnitude of impact.
 Lastly, the feature-grouped blocks are often an unnatural partition of the graph, leading to a poor model fit.
-We would instead prefer to partition the graph into its most natural blocks and then determine which features best predict the resulting partition.
+We would instead prefer to partition the graph into its most natural blocks and then find which of the available features -- if any -- best predict the resulting partition.
 
@@
 In the FFBM, we use the features first to generate the latent block membership for each vertex.
-The latent block membership is therefore a stepping stone rather than a starting point in our analysis.
+The latent block membership is therefore a stepping stone rather than a starting point in our model.
 We go on to present an efficient algorithm for sampling from the parameters of the feature-to-block generator.

### 2021-05-23 nips/conclusion.tex
Finish draft

@@
 The idea is to divide the graph into its most natural partition and test whether the vertex features can accurately explain this partition.
-We therefore hope to avoid a common pitfall in data science: correlation does not imply causation.
 It is very easy to find vertex features that are in some way correlated with the graphical structure.
@@
 This is introduced as two concurrent Markov chains to sample the block memberships $b$ and block generator parameters $\theta$.
-The key trick to make the $b$-chain efficient is the use of the microcanonical rather than the traditional SBM (as proposed by \citet{Peixoto-Bayesian-Microcanonical}) which allows us to bypass an expensive summation over the latent memberships $b$.
-Additionally, we use the empirical mean of the $b$-samples as the input to our $\theta$-chain, as this reduces the variance in our evaluation of the target distribution and thus shortens the burn-in and auto-correlation times of our $\theta$-chain.
+Nevertheless, we can serialise the chains and use the empirical mean of the $b$-samples as the input to our $\theta$-chain.
+This reduces the variance in our evaluation of the target distribution and thus shortens burn-in.
 
@@
 Nevertheless, the approach can only currently explain the structure at the macro-scale.
-We cannot explain structure within each community.
+We cannot explain structure within each block.
 Future work will benefit from extending the FFBM to be hierarchical in nature.
 That way, the structure of the network can be explained at all length-scales of interest.
-This work may be built upon to develop better methods of analysing structure in networks and drawing evidence-backed conclusions.
 So long as data collection techniques remain ethical and care is taken to respect personal privacy, such empowered decision-making can only help humankind.

### 2021-05-23 nips/experiments.tex
Finish draft

@@
 
-We apply the developed methods to a variety of datasets.
-Datasets are chosen to span a range of node counts $N$, edge counts $E$ and feature space dimension $D$.
-We consider the following datasets:\footnote{For information as to how these data were collected and how personally identifiable information was removed, please refer to the cited papers.}
+We apply the developed methods to a variety of datasets. These are chosen to span a range of node counts $N$, edge counts $E$ and feature space dimension $D$. We consider the following:
 
-\begin{itemize}
+\begin{itemize}[leftmargin=*]
 	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=3$) -- network of Amazon book sales about U.S.
 politics, published close to the presidential election in 2004.
-Two books are connected if they were frequently co-purchased by the same customer.
+Two books are connected if they were frequently co-purchased by customers.
 Vertex features encode the political affiliation of the author (liberal, conservative or neutral).
@@
 	\item \textbf{Primary school dynamic contacts} \cite{schools} ($N=238, E=5539, D=13$) -- network of face-to-face contacts amongst students and teachers at a primary school in Lyon, France.
-Two nodes are connected if the two parties shared a face-to-face interaction over the course of the day.
-Vertex features include class membership (one of 10 values 1A-5B), gender and whether or not the individual is a teacher or pupil.
-These data were collected on consecutive days in October 2009.
-We choose to analyse just the second day.
+Two nodes are connected if the two parties shared a face-to-face interaction over the school-day.
+Vertex features include class membership (one of 10 values: 1A-5B), gender and teacher/student distinction.
+No further identifiable information is retained.
+We choose to analyse just the second day of results.
 
-	\item \textbf{Facebook egonet} \cite{fb-snap} ($N=747, E=30025, D=480$) -- an assortment of Facebook egonets.
-These are networks of a particular user's friends list and all the connections within that.
-Vertex features are extracted from each user's profile and are fully anonymized.
-They include information about education history, languages spoken, gender, home-town, date of birth, name to give a few examples.
+	\item \textbf{Facebook egonet} \cite{fb-snap} ($N=747, E=30025, D=480$) -- an assortment of Facebook users' friends lists.
+Vertex features are extracted from each user's profile and are fully anonymised.
+They include information about education history, languages spoken, gender, home-town, birthday etc.
 We focus on the egonet with id 1912.
@@
 \end{figure}
-
-We require metrics to assess performance.
-This can be split into two separate components: the microcanonical SBM fit (concerned with the $b$-samples) and the fit of the feature-to-block generator (concerned with the $\theta$-samples).
-Starting with the SBM, $S(b)$ (equation \ref{eqn:dl-form}) can be interpreted as the description length of the partition imposed by $b$.
-It is only natural to divide this quantity by the number of entities (nodes and edges) in our graph $N+E$ to allow for rough comparison between graphs.
-This defines a simple metric to gauge the fit of the SBM: the description length per entity averaged over the $b$-samples (equation \ref{eqn:mean-dl}):
-%
-\begin{equation}
-	\bar{S}_e \coloneqq \frac{1}{(N+E) |\Tcal_b|} \sum_{t\in \Tcal_b} S \left( b^{(t)} \right)
-	\label{eqn:mean-dl}
-\end{equation}
-%
-However, to assess the performance of the feature-to-block predictor, we must partition the vertex set $[N]$ into training and test sets: $\Gcal_0$ and $\Gcal_1$ respectively.\footnote{We choose to randomly create the partition on each experiment run such that a constant fraction $f$ of the available vertices go to form our training set $\Gcal_0$ and the remaining vertices are held out to form our test set $\Gcal_1$.
-For all the experiments $f=0.7$.}
-The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain.
-As $|\Gcal_0| \neq |\Gcal_1|$ sets, we cannot use the un-normalised log target $U$ (equation \ref{eqn:U-form}) for comparison; the total cross-entropy loss is scaled by the size of each set but the prior term stays constant.
-We therefore must use the average cross-entropy loss over each set (equation \ref{eqn:cross-entropy-loss}):
-%
-\begin{equation}
-	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star \left( \theta^{(t)} \right)
-	\qquad \textrm{where} \qquad
-	\Lcal_\star \left( \theta^{(t)} \right) \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)}
-	\label{eqn:cross-entropy-loss}
-\end{equation}
-%
-Where $\star \in \{0, 1\}$ has been introduced to toggle between training and test sets.
-Table \ref{tab:results} summarises the results for each experiment.\footnote{For a comprehensive list of the hyper-parameters used for each experiment please see appendix \ref{appdx:hyperparams}}
-
-We also apply the dimensionality reduction method on the two higher dimensional datasets (the school and FB egonet).
-For this we leverage equation \ref{eqn:c-star}, to reduce the dimension from $D$ to $D'$ with $k=1$ to yield $c^*$.
-We then retrain the feature-block predictor using just the retained feature set $\Dcal'$ and report the loss over the training and test sets for the reduced classifier -- denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively.
-These values are also included in table \ref{tab:results}.
-
-\begin{table}[!h]
-	\centering
-	\caption{Experimental results averaged over $n=10$ iterations (mean $\pm$ standard deviation)}
-	\label{tab:results}
-	\resizebox{\textwidth}{!}{%
-	\begin{tabular}{c|ccc|c|cc|ccc}
-		Dataset  & $B$ & $D$ & $D'$ & $\bar{S}_e$ & $\bar{\Lcal}_0$ & $\bar{\Lcal}_1$ & $c^*$ & $\bar{\Lcal}_0'$ & $\bar{\Lcal}_1'$  \\ \hline
-		Polbooks & 3 & 3 & -- & $2.250 \pm 0.001$ & $0.584 \pm 0.033$ & $0.557 \pm 0.061$ & -- & -- & -- \\
-		School & 10 & 13 & 10 & $1.894 \pm 0.006$ & $0.793 \pm 0.096$ & $0.914 \pm 0.112$ & $1.112 \pm 0.230$ & $0.791 \pm 0.096$ & $0.864 \pm 0.139$ \\
-		FB egonet & 10  & 480 & 10 & $1.626 \pm 0.003$ & $1.305 \pm 0.034$ & $1.539 \pm 0.087$ & $0.942 \pm 0.042$ & $1.496 \pm 0.093$ & $1.578 \pm 0.104$
-	\end{tabular}
-	}
-\end{table}
-
-Table \ref{tab:results} already highlights some general trends in the results.
-Firstly, the variance of the test loss $\bar{\Lcal}_1$ tends to be higher than the training loss $\bar{\Lcal}_0$.
-This is expected as our test set is smaller than the training set and so more susceptible to variability in its construction.
-Indeed, much of the variance in the evaluation of $\bar{\Lcal}_0$ and $\bar{\Lcal}_1$ comes from the random partitioning of the graph into training and test sets $\Gcal_0$ and $\Gcal_1$.
-Secondly, it can be seen that the dimensionality reduction procedure brings the training and test losses closer together.
-This implies that the features we keep are indeed correlated with the underlying graphical partition.
-
-The average description length per entity of the graph $\bar{S}_e$ has very low variance implying the detected communities can be found reliably (to within an arbitrary relabelling of blocks).
-For reference we plot an inferred partition for each of the graphs on figure \ref{fig:graphs-all}.
-The polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do succeed at partitioning the graph into densely connected clusters.
-
-On figure \ref{fig:null-all} we give the sampled feature weights for the reduced feature-set $\Dcal'$ for a given experiment iteration (or just the original $\Dcal$ for the polbooks dataset).
-We go on to analyse each of these in turn.
-
 \begin{figure}[!h]
@@
 \end{figure}
+\begin{table}[!h]
+	\centering
+	\caption{Experimental results averaged over $n=10$ iterations (mean $\pm$ standard deviation)}
+	\label{tab:results}
+	\resizebox{\textwidth}{!}{%
+		\begin{tabular}{c|ccc|c|cc|ccc}
+			Dataset  & $B$ & $D$ & $D'$ & $\bar{S}_e$ & $\bar{\Lcal}_0$ & $\bar{\Lcal}_1$ & $c^*$ & $\bar{\Lcal}_0'$ & $\bar{\Lcal}_1'$  \\ \hline
+			Polbooks & 3 & 3 & -- & $2.250 \pm 0.001$ & $0.584 \pm 0.033$ & $0.557 \pm 0.061$ & -- & -- & -- \\
+			School & 10 & 13 & 10 & $1.894 \pm 0.006$ & $0.793 \pm 0.096$ & $0.914 \pm 0.112$ & $1.112 \pm 0.230$ & $0.791 \pm 0.096$ & $0.864 \pm 0.139$ \\
+			FB egonet & 10  & 480 & 10 & $1.626 \pm 0.003$ & $1.305 \pm 0.034$ & $1.539 \pm 0.087$ & $0.942 \pm 0.042$ & $1.496 \pm 0.093$ & $1.578 \pm 0.104$
+		\end{tabular}
+	}
+\end{table}
+%
+We require metrics to assess performance.
+This can be split into two separate components: the microcanonical SBM fit (concerned with the $b$-samples) and the fit of the feature-to-block generator (concerned with the $\theta$-samples).
+Starting with the SBM, $S(b)$ (equation \ref{eqn:dl-form}) can be interpreted as the description length of the partition imposed by $b$.
+It is only natural to divide this quantity by the number of entities (nodes and edges) in our graph $N+E$ to allow for rough comparison between graphs.
+This defines a simple metric to gauge the fit of the SBM: the description length per entity averaged over the $b$-samples (equation \ref{eqn:mean-dl}):
+%
+\begin{equation}
+	\bar{S}_e \coloneqq \frac{1}{(N+E) |\Tcal_b|} \sum_{t\in \Tcal_b} S \left( b^{(t)} \right)
+	\label{eqn:mean-dl}
+\end{equation}
+%
+However, to assess the performance of the feature-to-block predictor, we must partition the vertex set $[N]$ into training and test sets.
+We choose to randomly partition the vertices on each experiment run such that a constant fraction $f$ of the available vertices go to form our training set $\Gcal_0$ and the remainder are held out to form our test set $\Gcal_1$.
+The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain.
+As $|\Gcal_0| \neq |\Gcal_1|$ in general, we cannot use the un-normalised log target $U$ (equation \ref{eqn:U-form}) for comparison as the total cross-entropy loss is scaled by the size of each set but the prior term stays constant.
+We therefore must use the average cross-entropy loss over each set (equation \ref{eqn:cross-entropy-loss}):
+%
+\begin{equation}
+	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star \left( \theta^{(t)} \right)
+	\qquad \textrm{where} \qquad
+	\Lcal_\star \left( \theta^{(t)} \right) \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)}
+	\label{eqn:cross-entropy-loss}
+\end{equation}
+%
+Where $\star \in \{0, 1\}$ has been introduced to toggle between training and test sets.
+Table \ref{tab:results} summarises the results for each experiment.\footnote{For a comprehensive list of the hyper-parameters used for each experiment please see appendix \ref{appdx:hyperparams}} We also apply the dimensionality reduction method on the two higher dimensional datasets (the school and FB egonet).
+For this we leverage equation \ref{eqn:c-star}, to reduce the dimension from $D$ to $D'$ with $k=1$ to yield the maximal cutoff $c^*$.
+We then retrain the feature-block predictor using just the retained feature set $\Dcal'$ and report the loss over the training and test sets for the reduced classifier -- denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively.
+These values are also included in table \ref{tab:results}.
+
+Table \ref{tab:results} already highlights some general trends in the results.
+Firstly, the variance of the test loss $\bar{\Lcal}_1$ tends to be higher than the training loss $\bar{\Lcal}_0$.
+This is expected as our test set is smaller than the training set and so more susceptible to variability in its construction.
+Indeed, most of the variance in the evaluation of $\bar{\Lcal}_0$ and $\bar{\Lcal}_1$ comes from the random partitioning of the graph into training and test sets.
+Secondly, it can be seen that the dimensionality reduction procedure brings the training and test losses closer together.
+This implies that the features we keep are indeed correlated with the underlying graphical partition and that the approach generalises correctly.
+
+The average description length per entity of the graph $\bar{S}_e$ has very low variance implying the detected communities can be found reliably (to within an arbitrary relabelling of blocks).
+For reference we plot an inferred partition for each of the graphs on figure \ref{fig:graphs-all}.
+The polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do succeed at partitioning the graph into densely connected clusters.
 
@@
 
-We wish to determine whether political affiliation is a good predictor of the overall network structure.
+We wish to determine whether the author's political affiliation is a good predictor of the overall network structure.
 We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral).
-
 From, figure \ref{fig:polbooks-null}, we see that all 3 blocks have a distinct political affiliation as their largest positive component.
 This is strong evidence that political affiliation is indeed the axis which best predicts the 3-way natural partition of the graph into blocks.
-
-From figure \ref{fig:polbooks-graph}, we note that there are very few edges between blocks 2 and 3 (the detected conservative and liberal blocks as per figure \ref{fig:school-null}).
-Block 1 effectively acts as a separator between blocks 2 and 3; very few readers venture all the way across the political aisle for their next book.
-
-From table \ref{tab:results} we see that the training and test losses are very similar (the test loss mean is even below the training loss) and both are low in magnitude.
-This is very strong evidence that the features we have available (the author's political affiliation) are indeed the drivers of the high-level graphical structure at $B=3$.
+Furthermore, table \ref{tab:results} we see that the training and test losses are very similar (the test loss mean is even below the training loss) and both are low in magnitude.
+This provides further evidence to the claim that political affiliation is the best explanatory variable for the overall network structure.
 
@@
 
-We must first choose the number of blocks $B$ to define the coarseness of our analysis.
-A total of 10 school-classes would suggest that $B=10$ is a natural starting point.
-We visualise the inferred block memberships in figure \ref{fig:school-graph}.
-
+We choose $B=10$ in line with the total number of school-classes.
 As before, we sample the block-generator parameters $\theta$ and employ the dimensionality reduction technique with standard deviation multiplier $k=1$ to pick out the top $D'=10$ features.
@@
 
-As for the training and test losses of the original and reduced classifiers.
-We see that the training loss stays roughly constant between the original and reduced classifiers $\bar{\Lcal}_0 \approx \bar{\Lcal}_0'$ but the test loss of the reduced classifier is ever-so-slightly decreased (within margin of error).
-This improved generalisation implies that the features we have discarded (gender and teacher/student status) do not impact the high-level graphical structure and simply lead to over-fitting in the original classifier.
-The features we have available do not capture the entire picture but they do quite well (modestly low overall loss).
-
 \subsection{Facebook egonet}
 
-This dataset was chosen to showcase the power of the dimensionality reduction technique as the feature-space has high dimension ($D=480$).
-We sample the $b$-chain, specifying $B=10$ total blocks and use this to construct the $\theta$-samples as before.
-
-We again apply the dimensionality reduction technique with $k=1$ to target $D'=10$ features in the reduced set.
+We choose $B=10$ and $D'=10$ for this experiment.
 The remaining features (figure \ref{fig:fb-null}) are those that best explain the high-level community structure.
 The majority of the surviving features are education related.
-Nevertheless, for $D'=10$ we only have good explanations for the makeup of some of the detected blocks.
+Nevertheless, for $D'=10$ we only have good explanations for the makeup of some of the detected blocks; several blocks in figure \ref{fig:fb-null} do not have high-magnitude components for $D'=10$.
 
@@
 The regularisation term imposed by the prior goes some way to alleviating this problem.
-Nevertheless, we see in figure \ref{fig:fb-null} that the feature \verb*|birthday-5| has a very high weight as it relates to block 1.
-It might be possible to shift the feature values such that they take values in $\{-1, 1\}$ rather than $\{0, 1\}$ but to accept more than one feature-group per block.
-
-The training and test losses (table \ref{tab:results}) are the highest of the considered networks but nonetheless still low.
-This suggests that the features we have available only partly explain the high-level structure.
-Promisingly, after reducing the problem dimension our test loss barely changes, suggesting that the features we do keep are the best from the ones we have available at explaining the overall network structure.
+Nevertheless, we see in figure \ref{fig:fb-null} that the feature \verb*|birthday-5| has a very high weight as it relates to block 1 -- but it would be preposterous to conclude that birthdays determine graphical structure.
+The analyst must remain vigilant of such problems.

### 2021-05-23 nips/ffbm.tex
Finish draft

@@
 In this section we propose a novel generative model for labelled networks.
-As before, we let $N$ denote the number of nodes.
+We call this the feature-first block model (FFBM) and outline its structure in \ref{fig:ffbm} As before, we let $N$ denote the number of nodes and $B$ the number of blocks in our graph.
 We define the vector $x_i \in \Xcal^D$ as the feature vector for the $i$'th vertex.
-Each vertex has $D$ total features and we assume all entries take values from the same set $\Xcal$.
+$D$ is the number of features.
 For the datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$.
@@
 
-The proposed generative model (which we call the feature-first block model - FFBM) is given in figure \ref{fig:ffbm}.
-First, we start with the feature matrix X and probabilistically generate a vector of block memberships $b$.
+For the FFBM, we start with the feature matrix X and probabilistically generate a vector of block memberships $b \in [B]^N$.
 The parameters of this step are encapsulated by $\theta$.
-Each feature vector $x_i$ is treated independently and used to generate the corresponding block membership $b_i$.
-Each entry $b_i \in [B]$ where $B$ is the number of blocks and a constant in our model.
+Each feature vector $x_i$ is treated independently and used to generate the corresponding block membership $b_i \in [B]$.
 We choose a single softmax layer to model $p(b_i | x_i, \theta)$.
@@
 Where $\tilde{x} \coloneqq \left[x_1, x_2, \dots x_D, 1\right]^T$ is an augmented version of $x$ that allows for a bias term.
-The parameter vector $\theta$ for this stage contains of the all the weight vectors $\theta = \{w_k\}_{k=1}^{B}$.
+The parameter vector $\theta$ for this stage contains all the weight vectors $\theta = \{w_k\}_{k=1}^{B}$.
 Each $w_k$ has dimension $D+1$.
-We could instead write the parameters $\theta$ as a $B \times (D+1)$ matrix of weights $W$; this form has computational benefits as then $z_i \coloneqq W \tilde{x}_i$ is the input to the softmax activation function.
+We could instead write the parameters $\theta$ as a $B \times (D+1)$ matrix of weights $W$; this form has computational benefits as then $z_i \coloneqq W \tilde{x}_i$, which is the input to the softmax activation function.
 
@@
 	\end{tikzpicture}
-	\caption{Feature-first block model}
+	\caption{The feature-first block model (FFBM)}
 	\label{fig:ffbm}
@@
 
-Once the block memberships $b$ have been generated, we then draw the graph $A$ from the microcanonical DC-SBM (equation \ref{eqn:A-generation}) with additional parameters encapsulated by $\psi = \{e, k\}$.
-In a slight abuse of notation we denote the inter-block edge count matrix with $e = \psi_e$ and the degree sequence $k = \psi_k$ to make explicit that these parameters are contained in $\psi$.
+Once the block memberships $b$ have been generated, we then draw the graph $A$ from the microcanonical DC-SBM (equation \ref{eqn:A-generation}) with additional parameters encapsulated by $\psi = \{\psi_e, \psi_k\}$.
 %
@@
 %
-We will see that this form of prior is equivalent to a regularisation term in neural network training that penalises extreme weight magnitudes.
-For $\sigma_\theta^2 \rightarrow \infty$ this becomes an uninformative uniform prior.
-
 In our model, the block memberships vector $b$ is an intermediate latent variable and so we are not free to choose a prior for it.
@@
 %
-This is a uniform distribution that only depends on the number of blocks $B$.
-This is an enormously important simplification as evaluating $p(b | X)$ does not require knowing the exact value $X$ takes and can be performed without an expensive Monte-Carlo integration over the $\theta$-domain.
+This is an enormously important simplification as evaluating $p(b | X)$ does not require an expensive Monte-Carlo integration over the $\theta$-domain nor does it require the exact value of $X$.
 \citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for the additional microcanonical SBM parameters $\psi$ which we adopt.
@@
 %
-Where the simplification is made apparent by noting $(\psi \indep X) | b$.
+Where we leverage the fact $(\psi \indep X) | b$.
 We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$ to complete our model.
 Please refer to appendix \ref{appdx:prior} for the exact form of $p(\psi | b)$.
-All that concerns us is that we have a form which is computable.
+All that concerns the main argument is we have a computable form.

### 2021-05-24 cued/abstract.tex
Scaffold cued report

\begin{abstract}
	Labelled networks form a very common and important class of data,
	naturally appearing in numerous applications in science and engineering.
	A typical inference goal is to determine how the vertex labels
	(called features) affect the network's graph structure. A standard
	approach has been to partition the network into blocks grouped
	by distinct values of the feature of interest. A block-based random
	graph model -- typically a variant of the stochastic block model (SBM) --
	is then used to test for evidence of asymmetric behaviour within these
	feature-based communities. Nevertheless, the resulting communities
	often do not produce a natural partition of the graph. In this work
	we introduce a new generative model, the feature-first block model (FFBM),
	which is more effective at describing vertex-labelled undirected
	graphs and also facilitates the use of richer queries on labelled networks.
	We develop a Bayesian framework for inference with this model,
	and we present a method to efficiently sample from the posterior
	distribution of the FFBM parameters. The FFBM's structure is kept
	deliberately simple to retain easy interpretability of the parameter
	values. We apply the proposed methods to a variety of network data
	to extract the most important features along which the vertices
	are partitioned. The main advantages of the proposed approach are
	that the whole feature-space is used automatically, and features
	can be rank-ordered implicitly according to impact. Any features
	that do not significantly impact the high-level structure can be
	discarded to reduce the problem dimension. In cases where the vertex
	features available do not readily explain the community structure
	in the resulting network, the approach detects this and is protected
	against over-fitting. Results on several real-world datasets
	illustrate the performance of the proposed methods. 
\end{abstract}

### 2021-05-24 cued/appendices.tex
Scaffold cued report

--- 2021-05-23 nips/appendices.tex
@@
 \begin{equation}
-	p(\psi | b) = p(e | b) p(k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
+	p(\psi_e=e, \psi_k=k | b) = p(e | b) p(\psi_k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
 	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right]
@@
 \end{equation}
+
+\subsection{Derivation of U form}
+\label{appdx:form-U}
+
+The invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. We write this as follows:
+%
+\begin{align}
+	\pi_\theta(\theta) &\propto p(\theta | X, b) \propto p(b | X, \theta) p(\theta) \propto  \exp \left( - U(\theta) \right) \\
+	\therefore U(\theta) &= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}
+\end{align}
+%
+Where we have introduced $U(\theta)$ equal to the negative log posterior.
+Each of the constituent terms of $U(\theta)$ are easily computed (equation \ref{eqn:U-constituent-terms}) by defining $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
+%
+\begin{equation}
+	\log p(b | X, \theta) = \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
+	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta ||^2
+	\label{eqn:U-constituent-terms}
+\end{equation}
+%
+Discarding constant terms, we write $U(\theta)$ as in equation \ref{eqn:U-form-appdx-0}.
+Note that $||\theta||^2 = \sum_{i} \theta_{i}^2 = \sum_{j=1}^{B} ||w_j||^2$ is the Euclidean norm of the vector of parameters $\theta$.
+%
+\begin{equation}
+	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
+	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2 = N \cdot \Lcal(\theta) + \frac{1}{2\sigma_\theta^2} ||\theta||^2
+	\label{eqn:U-form-appdx-0}
+\end{equation}
+%
 

### 2021-05-24 cued/checklist.tex
Scaffold cued report

--- 2021-05-16 nips/checklist.tex
@@
 \section*{Checklist}
-
-%%% BEGIN INSTRUCTIONS %%%
-The checklist follows the references.  Please
-read the checklist guidelines carefully for information on how to answer these
-questions.  For each question, change the default \answerTODO{} to \answerYes{},
-\answerNo{}, or \answerNA{}.  You are strongly encouraged to include a {\bf
-	justification to your answer}, either by referencing the appropriate section of
-your paper or providing a brief inline description.  For example:
-\begin{itemize}
-	\item Did you include the license to the code and datasets? \answerYes{See Section gen-inst.}
-	\item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
-	\item Did you include the license to the code and datasets? \answerNA{}
-\end{itemize}
-Please do not modify the questions and only use the provided macros for your
-answers.  Note that the Checklist section does not count towards the page
-limit.  In your paper, please delete this instructions block and only keep the
-Checklist section heading above along with the questions/answers below.
-%%% END INSTRUCTIONS %%%
 
@@
 		\item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
-		\answerTODO{}
+		\answerYes{See experiments section \ref{sec:experiments}}
 		\item Did you describe the limitations of your work?
-		\answerTODO{}
+		\answerYes{See conclusion \ref{sec:conclusion} and experiments \ref{sec:experiments} sections}
 		\item Did you discuss any potential negative societal impacts of your work?
-		\answerTODO{}
+		\answerYes{See conclusion \ref{sec:conclusion}}
 		\item Have you read the ethics review guidelines and ensured that your paper conforms to them?
-		\answerTODO{}
+		\answerYes{}
 	\end{enumerate}
@@
 		\item Did you state the full set of assumptions of all theoretical results?
-		\answerTODO{}
+		\answerYes{See inference section \ref{sec:inference}}
 		\item Did you include complete proofs of all theoretical results?
-		\answerTODO{}
+		\answerYes{See inference section \ref{sec:inference}}
 	\end{enumerate}
@@
 		\item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
-		\answerTODO{}
+		\answerYes{See supplementary code}
 		\item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
-		\answerTODO{}
+		\answerYes{See appendix \ref{appdx:hyperparams}}
 		\item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
-		\answerTODO{}
+		\answerYes{See table \ref{tab:results}}
 		\item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
-		\answerTODO{}
+		\answerYes{See appendix \ref{appdx:imp-details}}
 	\end{enumerate}
@@
 		\item If your work uses existing assets, did you cite the creators?
-		\answerTODO{}
+		\answerYes{See section experiments \ref{sec:experiments}}
 		\item Did you mention the license of the assets?
-		\answerTODO{}
+		\answerYes{See section inference \ref{sec:inference}}
 		\item Did you include any new assets either in the supplemental material or as a URL?
-		\answerTODO{}
+		\answerYes{Supplementary material}
 		\item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
-		\answerTODO{}
+		\answerYes{Referred to original papers}
 		\item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
-		\answerTODO{}
+		\answerYes{Referred to original papers}
 	\end{enumerate}
@@
 		\item Did you include the full text of instructions given to participants and screenshots, if applicable?
-		\answerTODO{}
+		\answerNA{}
 		\item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
-		\answerTODO{}
+		\answerNA{}
 		\item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
-		\answerTODO{}
+		\answerNA{}
 	\end{enumerate}

### 2021-05-24 cued/inference.tex
Scaffold cued report

--- 2021-05-23 nips/inference.tex
@@
 %
-These samples allow us to approximate the posterior distribution for $\theta$ as well as compute a predictive distribution $p(b^* | x^*, A, X) = \int p(b^* | x^*, \theta) p(\theta | X, A) d\theta \approx \frac{1}{T} \sum_{t=1}^{T} p(b^* | x^*, \theta^{(t)})$.
 However, generating these samples is not easily done in practice.
-
-We propose an iterative approach.
-We first draw samples $b^{(t)}$ from the block membership posterior (equation \ref{eqn:b-samples}).
-We then use each $b^{(t)}$ to draw samples for $\theta$ as in equation \ref{eqn:theta-samples}.
+We therefore propose an iterative approach.
+We first draw samples $b^{(t)}$ from the block membership posterior (equation \ref{eqn:b-samples}) and then use each $b^{(t)}$ to draw samples for $\theta$ as in equation \ref{eqn:theta-samples}.
 %
@@
 \begin{equation}
-	\alpha = \min \left( \frac{\pi(x') q(x', x)}{\pi(x) q(x, x')} , 1 \right)
+	\alpha(x, x') = \min \left( \frac{\pi(x') q(x', x)}{\pi(x) q(x, x')} , 1 \right)
 	\label{eqn:mh-accept}
@@
 What we propose in equations \ref{eqn:b-samples} and \ref{eqn:theta-samples} is therefore implemented through a 2-level Markov chain.
-The resulting samples for $\theta^{(i)}$ are unbiased in the sense that the expectation of their distribution is the posterior we are targeting in equation \ref{eqn:theta-target}.
+The resulting samples for $\theta^{(t)}$ are unbiased in the sense that the expectation of their distribution is the posterior we are targetting:
 %
@@
 This is an example of a pseudo-marginal approach.
-Indeed, \citet{pseudo-marginal} show that the unbiased result in equation \ref{eqn:theta-unbiased} is sufficient to prove that for sufficient samples, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{t})\right] = p(\theta
+Indeed, \citet{pseudo-marginal} show that the unbiased result in equation \ref{eqn:theta-unbiased} is sufficient to prove that for large enough $t$, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{(t)})\right] = p(\theta
 | A, X)$ which is exactly the distribution we are targetting (equation \ref{eqn:theta-target}).
@@
 
-The reason we split the Markov chain into two stages is because the summation over all latent states $b \in [B]^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$ is intractable $O(B^N)$.
-
+The reason we split the Markov chain into two stages is because the summation over all latent states $b \in [B]^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$ is intractable -- $O(B^N)$.
 Figure \ref{fig:samp-sequence} shows an overview of the proposed method.
-We have introduced subscripts and conditionings to make explicit what parameters each step utilises.
+We have introduced subscripts and conditionings to make explicit what variables each step utilises.
 We note the power of the simplification given by equation \ref{eqn:b-pseudo-prior}.
-As $p(b| X) = B^{-N}$ does not depend on the exact value of X, we do not need to know the value of $X$ to perform the sampling on $b$.
+As $p(b| X)$ does not depend on the exact value of X, we do not need to know the value of $X$ to perform the sampling on $b$.
 Conversely, for the $\theta^{(t)}$ samples, we use only $b^{(t)}$ but not $A$ as $(\theta \indep A) | b$.
@@
 %
-This entropy can equally be thought of as the description length of the graph -- the minimum amount of information required to store the graph and its parameters.
-The exact form of the proposal distribution and accept-reject step is explored thoroughly by \citet{Peixoto-MCMC} and not repeated here.
+This entropy can equally be thought of as the description length of the graph.
+The exact form of the proposal $q_b$ is explored thoroughly by \citet{Peixoto-MCMC} and not repeated here.
 There is a widely used library for Python made available under LGPL called \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, which implements this algorithm.
-The only modification we make is in the block membership prior $p(b)$ which we replace with $p(b|X)=B^{-N}$ which is a uniform distribution and so cancels out in the MH accept-reject step.
+The only modification we make is in the block membership prior $p(b)$ which we replace with $p(b|X)=B^{-N}$, which cancels out in the MH accept-reject step as it is independent of $b$.
 
-\subsection{Sampling feature-to-block classifier parameters}
+\subsection{Sampling feature-to-block generator parameters}
 
@@
 %
-\begin{align}
-	p(\theta | X, b) &\propto p(b | X, \theta) p(\theta) = \pi_\theta (\theta) \propto  \exp \left( - U(\theta) \right) \\
-	\therefore U(\theta) &= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}
-\end{align}
+\begin{equation}
+	\pi_\theta(\theta) \propto p(\theta | X, b) \propto p(b | X, \theta) p(\theta) \propto \pi_\theta (\theta) \propto  \exp \left( - U(\theta) \right)
+\end{equation}
 %
 Where we have introduced $U(\theta)$ equal to the negative log posterior.
-Each of the constituent terms of $U(\theta)$ are easily computed (equation \ref{eqn:U-constituent-terms}).
-Similarly, we define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
-%
-\begin{equation}
-	\log p(b | X, \theta) = \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
-	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta ||^2
-	\label{eqn:U-constituent-terms}
-\end{equation}
-%
-Discarding constant terms, we write $U(\theta)$ as in equation \ref{eqn:U-form}.
-Note that $||\theta||^2 = \sum_{i} \theta_{i}^2 = \sum_{j=1}^{B} ||w_j||^2$ is the Euclidean norm of the vector of parameters $\theta$.
+We define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
+Discarding constant terms, we can write $U(\theta)$ as in equation \ref{eqn:U-form} (refer to appendix \ref{appdx:form-U} for the derivation).
 %
@@
 $U(\theta)$ in equation \ref{eqn:U-form} appears a typical objective function for neural network training.
-The first term -- which we collect into $N \cdot \Lcal(\theta)$ -- is the cross-entropy between the graph-predicted and feature-predicted block memberships summed over all vertices.
+The first term is introduced by the likelihood.
+We collect it into $N \cdot \Lcal(\theta)$, which is the cross-entropy between the graph-predicted and feature-predicted block memberships summed over all vertices.
 The second term of equation \ref{eqn:U-form} -- introduced by the prior -- brings a form of regularisation, guarding against over-fitting.
-In traditional applications we only seek the value of $\theta$ that minimises the objective function $U(\theta)$, which in our case would yield the maximum a posteriori (MAP) estimate.
-This is often done through some kind of gradient descent as $\nabla U$ is easily computable (equation \ref{eqn:U-derivative}).
-
-However, our goal is not to find the MAP estimate but to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
-It is sufficient to specify a proposal distribution and then apply a MH accept-reject step to ensure detailed balance with $\pi_\theta(\cdot)$.
-However, we can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
-We therefore adopt the Metropolis Adjusted Langevin Algorithm (MALA) -- first proposed by \citet{mala-tweedie} -- which leverages the $\nabla U$ information.
-Given the current sample $\theta$, we propose a new sample $\theta'$ according to:
+Different to traditional applications, our goal is not to find the minimiser of $U(\theta)$ but to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
+We can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
+We therefore adopt the Metropolis-adjusted Langevin algorithm (MALA) -- first proposed by \citet{mala-tweedie}.
+Given the current sample $\theta$, we generate a new sample $\theta'$ according to equation \ref{eqn:theta-update}.
 %
-\begin{equation}
-	\theta' = \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi
-	\label{eqn:theta-update}
-\end{equation}
+\begin{align}
+	\theta' &= \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi
+	\label{eqn:theta-update} \\
+	\therefore q_\theta(\theta, \theta') &= \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right)
+	\label{eqn:theta-proposal}
+\end{align}
 %
@@
 We require the noise term $\xi$ to fully explore the parameter space.
-We can write the proposal distribution simply as:
-%
-\begin{equation}
-	q_\theta(\theta, \theta') = \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right)
-\end{equation}
-%
-The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}). By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as:
+We can write the proposal distribution $q_\theta$ as in equation \ref{eqn:theta-proposal}.
+The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}).
+By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as:
 %
@@
 %
-After a proposed move is generated, in typical Metropolis-Hastings fashion we accept the move with probability $\alpha_\theta$:
-%
-\begin{equation}
-	\alpha_\theta(\theta, \theta') = \min \left( 
-	\exp \left( U(\theta) - U(\theta')\right)
-	\frac{ 
-		q_\theta(\theta', \theta)
-	}{
-		q_\theta(\theta, \theta')
-	} 
-	, 1 \right)
-\end{equation}
+After a proposed move is generated, in typical Metropolis-Hastings fashion we accept the move with probability $\alpha_\theta$, as in equation \ref{eqn:mh-accept}.
 
@@
 This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance.
-This may lead to longer burn-in for the resulting Markov Chain.
+This may lead to longer burn-in for the resulting Markov chain.
 The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends only on the matrix $y^{(t)}$ with entries $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
@@
 %
-We can obtain an unbiased estimate for this quantity using the block membership samples $\left\{ b^{(t)} \right\}$.
+We can obtain an unbiased estimate for this quantity using the set of $b$-samples.
 However, as with all MCMC methods, we must only uses samples after burn-in and thinning have been applied.
@@
 %
-We therefore, choose to feed each $\theta^{(t)}$ update step the same matrix $\hat{y}$ for all $t$ rather than the corresponding $y^{(t)}$.
+We choose to feed each $\theta^{(t)}$ update step the same matrix $\hat{y}$ for all $t$ rather than the corresponding $y^{(t)}$.
 This means we no longer need to run the $b$ and $\theta$ Markov chains concurrently.
 Instead, we run the $b$-chain to completion and use it to generate $\hat{y}$.
-This is an estimate of $p(b | A, X)$ that we use for every iteration of the $\theta$ Markov chain.
-
-This affords us the flexibility to vary the number of samples we draw for the $b$ and $\theta$-chains; we denote the total number of samples generated $T_b$ and $T_\theta$ henceforth.
-Note that due to burn-in and thinning, $|\Tcal_b| < T_b$ and $|\Tcal_\theta| < T_\theta$.
-Furthermore, this changeover from $y^{(t)}$ to $\hat{y}$ reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
+This affords us the flexibility to vary the lengths of the $b$ and $\theta$-chains.
+Furthermore, the changeover from $y^{(t)}$ to $\hat{y}$ reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
 A description of the overall algorithms is given in appendix \ref{appdx:algorithms}.
@@
 However, as the Laplace approximation is not exact we will only treat this dimensionality reduction method as a useful heuristic and not an exact method.
-
 Conversely, we could fix $k=k_0$ and the dimension of our reduced feature set $|\Dcal'|=D'$.
@@
 \end{equation}
-%

### 2021-05-24 cued/introduction.tex
Scaffold cued report

--- 2021-05-23 nips/introduction.tex
@@
 
-A somewhat surprising property of many real-world networks is that they exhibit strong community structure.
-In other words, each node will often belong to a cluster of densely connected nodes.
+A somewhat surprising property of many real-world networks is that they exhibit strong community structure; most nodes often belong to a densely connected cluster.
 There is high interest in recovering the latent communities from the observed graphs.
-The inferred communities can be exploited for compression algorithms \cite{cluster-compression} or used for link prediction in incomplete networks \cite{link-prediction} to name but a few applications.
+The inferred communities can be exploited for compression algorithms \cite{cluster-compression} or used for link prediction in incomplete networks \cite{link-prediction} to name a few applications.
 
-We restrict our analysis to labelled networks.
-These are graphs where we additionally have information about the properties of each vertex.
-We shall refer to these vertex properties as features.
-One of the most common questions we can ask of labelled networks is whether a given vertex feature has an impact on the structure of the graph.
-To answer this question from a Bayesian perspective we must use a random graph model; the most common form is the stochastic block model (SBM) \cite{vanilla-sbm}.
+We restrict our analysis to vertex-labelled networks.
+We shall refer to the vertex-labels as features.
+A common goal is to determine whether a given feature impacts graphical structure.
+To answer this from a Bayesian perspective we must use a random graph model; the standard is the stochastic block model (SBM) \cite{vanilla-sbm}.
 This is a latent variable model where each vertex belongs to a single block and the probability two nodes are connected depends only on the block memberships of each.
-There have been many variants to this model; the most popular are the mixed-membership stochastic block model (MMSBM) \cite{mixed-membership-sbm} and the overlapping stochastic block model (OSBM) \cite{overlapping-sbm}.
+There have been many variants to this model -- the most popular being the mixed-membership stochastic block model (MMSBM) \cite{mixed-membership-sbm} and the overlapping stochastic block model (OSBM) \cite{overlapping-sbm}.
 Effectively, these just extend the model to allow each vertex to belong to multiple blocks simultaneously.
-
 However, a major drawback of these graphical models as applied to labelled networks is that they do not automatically include vertex features in the random graph generation process.
@@
 The associated model can then be used to test for evidence of heterogeneous connectivity between the feature-grouped blocks.
-Nevertheless, this approach is limited in that it can only consider one feature at a time; this leaves any conclusions drawn highly vulnerable to the presence of confounding variables.
-Furthermore, considering each feature separately makes it difficult to rank order the features by  magnitude of impact.
+Nevertheless, this approach is limited in that it can only consider one feature at a time.
+This makes it difficult to rank order the features by  magnitude of impact.
 Lastly, the feature-grouped blocks are often an unnatural partition of the graph, leading to a poor model fit.
@@
 With these desiderata in mind, we present a novel framework for modelling labelled networks, which we call the feature-first block model (FFBM).
-This can be thought of as an extension of the SBM to labelled networks.
-In the FFBM, we use the features first to generate the latent block membership for each vertex.
-The latent block membership is therefore a stepping stone rather than a starting point in our model.
+This is an extension of the SBM to labelled networks.
 We go on to present an efficient algorithm for sampling from the parameters of the feature-to-block generator.
 We can interpret the sampled FFBM parameters to determine which features have the largest impact on overall graphical structure.
-Any features found to be irrelevant can be discarded as a form of dimensionality reduction.

### 2021-05-24 cued/prelim.tex
Scaffold cued report

--- 2021-05-23 nips/prelim.tex
@@
 For this we adopt the stochastic block model (SBM) - widely used across academia.
-The premise is that each node in the graph belongs to a unique community called a block.
+Each node in the graph belongs to a unique community called a block.
 The probability that two nodes are connected depends only on the block memberships of each.
-Graphs drawn from the SBM ensemble exhibit community structure.
 Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}.
-To allow for degree-variability between members of the same block, we must choose the degree-corrected formulation of the SBM (DC-SBM).
+To allow for degree-variability between members of the same block, we must choose the degree-corrected formulation (DC-SBM).
 
@@
 $e_{rs}$ is then the number of edges from block $r$ onto block $s$ -- or twice that number if $r=s$.
-For undirected graphs $e$ is symmetric.
-	Let $k \in (\Integers^+_0)^N$ be a vector denoting the degree sequence of the graph. $k_i$ is then the degree of vertex $i$.
-
-	For the degree-corrected stochastic block model (DC-SBM), we say that the graph's adjacency matrix $A \in \{0,1\}^{N \times N}$ is generated as follows:
+For undirected graphs, $e$ is symmetric.
+	Let $k \in (\Integers^+_0)^N$ be a vector denoting the degree sequence of the graph.
+$k_i$ is then the degree of vertex $i$.
+The graph's adjacency matrix $A \in \{0,1\}^{N \times N}$ is generated as follows:
 	%
@@
 	%
-	Where edges are placed uniformly at random but respecting the constraints imposed by $e$, $b$ and $k$ -- hence why this formulation is given the microcanonical moniker.
-Specifically, $A$ must satisfy the following:
+	Where edges are placed uniformly at random but respecting the constraints imposed by $e$, $b$ and $k$ -- hence the microcanonical moniker. Specifically, $A$ must satisfy the following:
 	%
 	\begin{equation}
-		e_{rs} = \sum_{i, j} A_{ij} \one \{b_i = r\} \one \{b_j = s\} \qquad \textrm{and} \qquad
-		k_i = \sum_j A_{ij}
+		e_{rs} = \sum_{i, j \in [N]} A_{ij} \one \{b_i = r\} \one \{b_j = s\} \quad \forall r,s \in [B]
+		\quad \textrm{and} \quad
+		k_i = \sum_{j \in [N]} A_{ij} \quad \forall i \in [N]
 		\label{eqn:sbm-constraints}
 	\end{equation}
-	%
-	The additional parameters $N$ and $B$ are omitted as they can be inferred from the shapes of $b$ and $e$.
 \end{definition}
-
-In what follows, the SBM will always refer to the microcanonical DC-SBM variation.

### 2021-05-24 cued/covid.tex
Font-size and title page

\section{Covid-19 disruption}

Please provide details of disrupted project work, including dates when the work was planned to take place and
reasons for the disruption (e.g. inaccessible equipment, illness).
If your work was disrupted, to what extent was it possible to redefine elements of your project so as to achieve
alternative goals? Please comment on the success, or otherwise, of the Covid-19 contingency plan that you
formulated last summer and documented on your Project Planning Form.
This appendix should not exceed one page in length.

### 2021-05-24 cued/abstract.tex
More work on CUED

@@
-\begin{abstract}
+\section*{Technical Abstract}
 	Labelled networks form a very common and important class of data,
@@
 	illustrate the performance of the proposed methods.
-\end{abstract}

### 2021-05-24 cued/covid.tex
Covid and risk retro

@@
 
-Please provide details of disrupted project work, including dates when the work was planned to take place and
-reasons for the disruption (e.g. inaccessible equipment, illness).
-If your work was disrupted, to what extent was it possible to redefine elements of your project so as to achieve
-alternative goals? Please comment on the success, or otherwise, of the Covid-19 contingency plan that you
-formulated last summer and documented on your Project Planning Form.
-This appendix should not exceed one page in length.
+Thankfully this project was not directly disrupted by Covid-19 as it was always going to be heavily theoretical and computational.
+The only disruption to speak of was the shift of supervisor meetings to online.
+Though shared whiteboards were alien at first, we grew to learn how to use them to effectively communicate ideas.
+
+I was ill with Covid-19 for 2 weeks in October but this did not majorly impact my ability to work.
+The Covid-19 contingency plan formulated last summer and documented on the Project Planning Form just involved switching supervisions to an online format.

### 2021-05-24 cued/risk.tex
Covid and risk retro

\section{Risk assessment retrospective}

As experimental work was entirely computer-based the risks encountered were not what one might typically associate with a more hands-on project: heavy machinery, dangerous chemicals etc. Most of the risks identified would be as a result of an improper working environment. These could be properly managed by ensuring the working environment felt comfortable. An external monitor went some way to alleviating the eye-strain of coding on a small screen for extended periods. Furthermore, an external mouse is far kinder on the wrist than a track-pad on laptops. There were no unexpected hazards encountered that were not identified at the start of the project.

### 2021-05-25 nips/additional.tex
Refactor appendices

--- 2021-05-23 nips/appendices.tex
@@
-\section{Appendix}
+\section{Appendix: Additional material}
 
-\subsection{SBM prior choice explanation}
-\label{appdx:prior}
+\subsection{SBM likelihood and prior}
+\label{appdx:sbm}
 
-We have borrowed the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$. These, are repeated here for reference:
+For reference, we provide the likelihoods and priors for the mircocanonical SBM proposed by \citet{Peixoto-Bayesian-Microcanonical}. We have that the graph $A$ is drawn from the SBM:
 %
 \begin{equation}
-	p(\psi | b) = p(e | b) p(k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
+	A \sim \textrm{DC-SBM}_{\textrm{MC}}(b, e, k)
+\end{equation}
+%
+With edges placed uniformly at random but respecting the constraints imposed by $b, e$ and $k$.
+The likelihood calculation for $p(A|k, e, b)$ then reduces to a case of counting configurations that yield the same adjacency matrix, $\Xi (A)$, and dividing by the total number of configurations possible, $\Xi(e)$.
+This is why this formulation is given the microcanonical moniker.
+If we consider the half-edges to be distinguishable for a moment, the total number of configurations that satisfy the $e$ constraint is:
+%
+\begin{equation}
+	\Omega(e) = \frac{\prod_{r} e_r !}{\prod_{r,s : r < s} e_{rs}! \cdot \prod_{r} e_{rr}!!}
+\end{equation}
+%
+Where $e_r \coloneqq \sum_{s} e_{rs}$ and $(2m)!!
+\coloneqq 2^m m!$.
+Nevertheless, a great number of these configurations yield the same graph $A$.
+We denote the number of configurations that yield the adjacency matrix $A$ with $\Xi(A)$, which can be computed as:
+%
+\begin{equation}
+	\Xi(A) \coloneqq \frac{\prod_i k_i !}{\prod i,j : i < j A_{ij} ! \prod_i A_{ii} !! }
+\end{equation}
+%
+Note the similarity between the forms of $\Omega(e)$ and $\Xi(A)$ as $e$ is effectively the adjacency matrix of the block-graph. With these defined we can write the overall likelihood:
+%
+\begin{equation}
+	p(A|k,e,b) = \frac{\Xi(A)}{\Omega(e)}
+\end{equation}
+%
+Obviously, this form is only defined if $A$ respects the constraints imposed by $(k,e,b)$ else the likelihood is 0.
+With the likelihood defined, we move on to the prior.
+As discussed in the main text, for the FFBM $b$ is an intermediate variable and not a parameter so we are not free to choose a prior for it.
+Nevertheless, we can borrow the conditional prior proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(e, k | b)$:
+%
+\begin{equation}
+	p(e, k | b) = p(e | b) p(k| e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
 	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right]
@@
 $q(m, n)$ is related to but different from $\specialchoose{n}{m}$.
-Lastly, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ is the number of vertices assigned to block $r$.
+Recall that $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ is the number of vertices assigned to block $r$.
 
-These were chosen carefully by \citet{Peixoto-Bayesian-Microcanonical} to more closely match the structure of empirical networks than simple uniform priors. We do not repeat his arguments here.
-
-\subsection{Derivation of conditional block distribution given feature matrix}
-\label{appdx:b|x}
-
-We wish to determine the form of $p(b| X)$. This can be done by integrating over the joint probability with respect to $\theta$.
-%
-\begin{align*}
-	p(b | X) &= \int p(b , \theta| X, \theta) d\theta = \int p(b | X, \theta) p(\theta | X) d\theta \\
-	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i \in [N] } \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
-	&= \prod_{i \in [N]} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j \in [B]} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k \in [B]} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}
-\end{align*}
-%
-We note that $b_i \in [B]$ and so the integral's value is unchanged with respect to $b_i$.
-The integrand has the same form no matter which value $b_i$ takes as the prior is the same for each $w_j$.
-As such the integral can only be a function of at most $\tilde{x}_i$ and $\sigma_\theta^2$ as it is symmetric with respect to $b_i$ and all the various $w_j$ are integrated out as they are dummy variables.
-Therefore, denoting the integral by the (unknown) function $f(\tilde{x}_i, \sigma_\theta^2)$, we write $p(b| X)$ as follows:
-%
-\begin{align*}
-	p(b | X) &= \prod_{i=1}^{N} f(\tilde{x}_i, \sigma_\theta^2) = \textrm{const w.r.t } b = c
-\end{align*}
-%
-As this is a constant with respect to $b$ we conclude that $p(b | X)$ must be a uniform distribution.
-$\nicefrac{1}{c}$ is simply the size of the set of values that $b$ can take.
-We know $b_i \in [B]$.
-Therefore, $b \in [B]^N$ and $|[B]^N| = B^N = \nicefrac{1}{c}$.
-Putting this all together we conclude that:
-%
-\begin{equation}
-	p(b | X) = B^{-N}
-\end{equation}
-
-\subsection{Derivation of U gradient with respect to feature parameters}
-\label{appdx:gradu}
-The goal is to determine $\nabla U(\theta)$, the gradient of the negative log posterior with respect to the parameters. We repeat the form of $U(\theta)$ in equation \ref{eqn:U-form-appdx}.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form-appdx}
-\end{equation}
-%
-Where $y_{ij}$ is independent of $\theta$ and $a_{ij}$ is the output from the softmax layer, with form as given in equation \ref{eqn:a-ij}.
-%
-\begin{equation}
-	a_{ij} \coloneqq \phi_{j} (x_i; \theta) = \frac{\exp(w_j^T \tilde{x}_i)}{\sum_{r \in [B]} \exp(w_r^T \tilde{x}_i)}
-	\label{eqn:a-ij} 
-\end{equation}
-%
-We note that $\theta = \{w_k\}_{k=1}^B$, and as such we can write this in vector form $\theta = \left[w_1^T, w_2^T \dots w_B^T  \right]^T$.
-Therefore, $\nabla U(\theta) = \left[\nicefrac{\partial U}{\partial w_1}^T,\nicefrac{\partial U}{\partial w_2}^T \dots \nicefrac{\partial U}{\partial w_B}^T  \right]^T$; to compute $\nabla U(\theta)$ it suffices to find the form of $\nicefrac{\partial U}{\partial w_k}$ with respect to a general $k$.
-
-To this end, we must first find partial derivatives of $a_{ij}$ and $||\theta||$ with respect to $w_k$. Starting with $a_{ij}$:
-%
-\begin{align}
-	\frac{\partial a_{ij}}{\partial w_k} &= \frac
-	{\tilde{x}_i \exp(w_j^T \tilde{x}_i) \delta_{jk} \cdot \sum_{r \in [B]} \exp(w_r^T \tilde{x}_i) 
-		- 
-		\exp(w_j^T \tilde{x}_i) \cdot \tilde{x}_i \exp(w_k^T \tilde{x}_i)}
-	{\left( \sum_{r \in [B]} \exp(w_r^T \tilde{x}_i) \right)^2} \nonumber \\
-	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right) 
-\end{align}
-%
-Where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$. Now moving onto the derivative of $||\theta||^2$:
-%
-\begin{equation}
-	\frac{ \partial}{\partial w_k} ||\theta||^2 = \frac{\partial}{\partial w_k} \left( \sum_{r \in [B]} ||w_r||^2 \right) = 2w_k
-\end{equation}
-%
-We are ready to put this all together, to find the partial derivative of $U(\theta)$ with respect to each $w_k$:
-\begin{align}
-	\frac{\partial U}{\partial w_k} &= 
-	\sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} 
-	\left( \frac{-\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
-	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
-	&=  - \left( \sum_{i=1}^{N} \tilde{x}_i \left( y_{ik} - a_{ik} \sum_{j=1}^{B} y_{ij} \right)
-	- \frac{w_k}{\sigma_\theta^2} \right) \nonumber \\
-	&= - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
-\end{align}
-%
-This is the required result.
-This form can be computed efficiently through matrix operations.
-The only property of $y_{ij}$ we have used in the derivation is the sum-to-one constraint $\sum_{j=1}^{B} y_{ij} = 1$ for all $i$.
+The form of these priors were chosen carefully by \citet{Peixoto-Bayesian-Microcanonical} to more closely match the structure of empirical networks than simple uniform priors.
+We do not repeat his arguments here.
 
@@
 Indeed, it is prudent to start our $b$-chain at this MDL solution as then we can burn-in time is greatly reduced.
-
-\subsection{Algorithms}
-\label{appdx:algorithms}
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{Block membership sample generation} % give the algorithm a caption
-	\label{alg:b-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-		\State $b^{(0)} \gets \argmin_b S(b|A)$ \Comment{Implemented as greedy heuristic in \textit{graph-tool} library}
-		\For{$t \in \{0, 1 \dots T_b - 1\}$}
-		\State $b' \gets \sim q_b(b^{(t)}, b' | A)$
-		\State $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
-		\State $\eta \gets \sim \textrm{Unif}(0,1)$
-		\If{$\log \eta < \log \alpha_b$}
-		\State $b^{(t+1)} \gets b'$
-		\Else
-		\State $b^{(t+1)} \gets b^{(t)}$
-		\EndIf
-		\EndFor
-
-		\State \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{FFBM parameter pseudo-marginal inference} % give the algorithm a caption
-	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\State $\hat{Y}_{ij} \gets \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
-		\State $\theta^{(0)} \gets \sim \Gaussian(0, \sigma_\theta I)$
-
-		\item[]
-
-		\For{$t \in \{0, 1 \dots T_\theta - 1\}$}
-		\State $\xi \gets \sim \Gaussian(0, I)$
-		\State $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
-		\State $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
-		\State $\eta \gets \sim \textrm{Unif}(0,1)$
-		\If{$\log \eta < \log \alpha_\theta$}
-		\State $\theta^{(t+1)} \gets \theta'$
-		\Else
-		\State $\theta^{(t+1)} \gets \theta^{(t)}$
-		\EndIf
-		\EndFor
-
-		\State \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\FloatBarrier
-\subsection{Hyperparameter values}
-\label{appdx:hyperparams}
-
-\begin{table}[!h]
-	\centering
-	\caption{Hyper-parameter values for each experiment}
-	\label{tab:hyperparams}
-	\resizebox{\textwidth}{!}{%
-		\begin{tabular}{c|ccc|ccc|cccc|cc|cccc}
-			Dataset & 
-			$B$ & $f$ & $\sigma_\theta$ & 
-			$T_b$ & $\kappa_b$ & $\lambda_b$ & 
-			$T_\theta$ & $\kappa_\theta$ & $\lambda_\theta$ & $s$ &
-			$k$ & $D'$ &
-			$T_\theta'$ & $\kappa_\theta'$ & $\lambda_\theta'$ & $s'$
-			\\ \hline
-			Polbooks &
-			3 & 0.7 & 1 &
-			1,000 & 0.2 & 5 &
-			10,000 & 0.4 & 10 & 0.05 &
-			-- & -- & 
-			-- & -- & -- & -- \\
-			School &
-			10 & 0.7 & 1 &
-			1,000 & 0.2 & 5 &
-			10,000 & 0.4 & 10 & 0.2 &
-			1 & 10 & 
-			10,000 & 0.4 & 10 & 0.2 \\
-			FB Egonet &
-			10 & 0.7 & 1 &
-			1,000 & 0.2 & 5 &
-			10,000 & 0.4 & 10 & 0.017 &
-			1 & 10 & 
-			10,000 & 0.4 & 10 & 0.5 \\
-		\end{tabular}
-	}
-\end{table}
-
-\subsection{Implementation details}
-\label{appdx:imp-details}
-
-All data analysis and visualisation was implemented in Python.
-Full source code is available in the supplementary material.
-The scripts were run using a standard PC using the Windows Subsystem for Linux (WSL) environment.
-Specs are:
-
-\begin{itemize}
-	\item \textbf{CPU}: Intel(R) Core(TM) i7-1065G7
-	\item \textbf{RAM}: 8GB
-	\item \textbf{GPU}: Intel(R) Iris(R) Plus Graphics
-\end{itemize}
-
-On this hardware each experiment iteration took the following amount of time to execute:
-
-\begin{table}[!h]
-	\centering
-	\caption{Compute-time for each experiment}
-	\label{tab:compute-time}
-	\begin{tabular}{c|ccc|c}
-		Dataset & $b$-chain & $\theta$-chain & Reduced $\theta$-chain & Overall compute time \\ \hline
-		Polbooks & $\sim$1s & $\sim$4s & -- & $\sim$5s \\
-		School & $\sim$10s & $\sim$10s & $\sim$10s & $\sim$30s \\
-		FB Egonet & $\sim$20s & $\sim$180s & $\sim$10s & $\sim$210s
-	\end{tabular}
-\end{table}

### 2021-05-25 nips/derivations.tex
Refactor appendices

--- 2021-05-23 nips/appendices.tex
@@
-\section{Appendix}
+\section{Appendix: Derivations}
 
@@
 The only property of $y_{ij}$ we have used in the derivation is the sum-to-one constraint $\sum_{j=1}^{B} y_{ij} = 1$ for all $i$.
-
-\subsection{Choosing the MALA step-size}
-\label{appdx:step-size}
-
-For sampling from the $\theta$-chain of the block membership generator parameters, we employed the Metropolis Adjusted Langevin Algorithm (MALA). At iteration $t$, the proposed sample is generated by:
-%
-\begin{equation}
-	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi
-\end{equation}
-%
-There are two competing objectives when choosing the step-size $h_t$.
-On the one hand, we want the step-size to be large so that we arrive at a high density region quickly.
-However, too large a step-size will lead to a lower acceptance ratio and thus inefficient sampling.
-A solution to this problem would be to slowly decrease the step-size with $t$ - often called simulated annealing.
-Therefore, we still have a short burn-in time but will not bounce around the mode for large $t$.
-As well as the trivial constraint for $h_t$ to be strictly positive, we introduce two further constraints as outlined by \citet{Bayesian-SGLD}:
-%
-\begin{equation}
-	\sum_{t=1}^{\infty} h_t = \infty \qquad \textrm{and} \qquad
-	\sum_{t=1}^{\infty} h_t^2 < \infty
-	\label{eqn:h-constraints}
-\end{equation}
-%
-The first constraint ensures that we have cover sufficient distance to arrive at any arbitrary point in our domain, no matter the starting point.
-The second constraint ensures that once we converge to the mode rather than simply bouncing around it.
-\citet{Bayesian-SGLD} propose the following form for a polynomially decaying step-size which we adopt:
-%
-\begin{equation}
-	h_t = \alpha(\beta + t)^{-\gamma}
-\end{equation}
-%
-Where $\alpha, \beta, \gamma$ are hyper-parameters to be chosen.
-We require $\alpha,\beta > 0$ and $\gamma \in (0.5, 1]$ to satisfy equation \ref{eqn:h-constraints}.
-To reduce the number of hyperparameters we set these to have values given by the equations \ref{eqn:step-size-params}.
-%
-\begin{equation}
-	\alpha = \frac{250 \cdot s}{N} \qquad \beta = 1000 \qquad \gamma = 0.8
-	\label{eqn:step-size-params}
-\end{equation}
-%
-Where $N$ is the number of data-points we are considering and now $s$ is the only free variable which we call the step-size scaling.
-For approximate methods, we can choose to bypass the MH accept-reject entirely to speed up computation.
-If this is done, the algorithm is instead called stochastic gradient Langevin diffusion (SGLD) \cite{Bayesian-SGLD}.
-This speeds up computation at the expense of exactness of the method.
-
-\subsection{Burn-in and thinning}
-\label{appdx:burn-in-thinning}
-
-As with any MCMC method, we must deal with the issues presented by burn-in and thinning.
-We have introduced the notation $\Tcal_b$ and $\Tcal_\theta$ to denote the set of samples we keep from the $b$ and $\theta$ chains respectively.
-Note that we generate $T_b$ and $T_\theta$ samples total.
-The burn-in period refers to the time taken for the Markov Chain to converge to the stationary distribution.
-Sample thinning is necessary to ensure that neighbouring samples satisfy independence.
-However, as we do not leverage the independence property this is less important in our analysis.
-We can write the general set $\Tcal_\star$ as:
-%
-\begin{equation}
-	\Tcal_\star = \{T_\star \kappa_\star + i \lambda_\star :  
-	0 \leq i \leq \lfloor T_\star(1 - \kappa_\star) / \lambda_\star \rfloor \}
-\end{equation}
-%
-Where the parameter $\kappa_\star \in (0, 1)$ controls our burn-in and $\lambda_\star$ controls our thinning.
-$\kappa_\star$ can be determined by plotting the log-target (either $S(b^{(t)})$ or $U(\theta^{(t)})$ with respect to the epoch $t$.
-$\kappa_\star$ is then chosen to encompass the region where the log-target has roughly equilibrated.
-As we do not leverage sample independence $\lambda_\star$ can be chosen less rigorously.
-We often just use $\lambda_b=5$ and $\lambda_\theta = 10$.
-
-\subsection{Initializing the b-chain}
-
-For the purposes of our model (the FFBM), the number of blocks $B$ is a constant which must be specified by the data scientist.
-We could however, allow our choice of $B$ to be influenced by the observed data.
-This places us in the domain of empirical Bayes, which must be negotiated carefully.
-Prior beliefs must be determined a priori else they are not prior.
-However, as the number of blocks only specifies the coarseness of the analysis, it is fine to allow it to vary.
-Indeed, \citet{peixoto-determine-B} shows that for a fixed average degree the maximum number of detectable blocks scales as $O(\sqrt{N})$ where $N$ is the number of vertices.
-
-If we allow $B$ to vary in the $b$-chain (i.e.
-new blocks can be created and we permit empty blocks) then it can be run  until a minimum description length (MDL) solution is reached.
-We take the number of non-empty blocks at the MDL to be our fixed block number $B$ for subsequent analysis.
-Indeed, it is prudent to start our $b$-chain at this MDL solution as then we can burn-in time is greatly reduced.
-
-\subsection{Algorithms}
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{Block membership sample generation} % give the algorithm a caption
-	\label{alg:b-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-		\State $b^{(0)} \gets \argmin_b S(b|A)$ \Comment{Implemented as greedy heuristic in \textit{graph-tool} library}
-		\For{$t \in \{0, 1 \dots T_b - 1\}$}
-		\State $b' \gets \sim q_b(b^{(t)}, b' | A)$
-		\State $\log \alpha_b \gets \log \alpha_b(b^{(t)}, b' | A)$
-		\State $\eta \gets \sim \textrm{Unif}(0,1)$
-		\If{$\log \eta < \log \alpha_b$}
-		\State $b^{(t+1)} \gets b'$
-		\Else
-		\State $b^{(t+1)} \gets b^{(t)}$
-		\EndIf
-		\EndFor
-
-		\State \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm} % enter the algorithm environment
-	\caption{FFBM parameter pseudo-marginal inference} % give the algorithm a caption
-	\label{alg:theta-samples} % and a label for \ref{} commands later in the document
-	\begin{algorithmic} % enter the algorithmic environment
-
-		\State $\hat{Y}_{ij} \gets \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
-		\State $\theta^{(0)} \gets \sim \Gaussian(0, \sigma_\theta I)$
-
-		\item[]
-
-		\For{$t \in \{0, 1 \dots T_\theta - 1\}$}
-		\State $\xi \gets \sim \Gaussian(0, I)$
-		\State $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
-		\State $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
-		\State $\eta \gets \sim \textrm{Unif}(0,1)$
-		\If{$\log \eta < \log \alpha_\theta$}
-		\State $\theta^{(t+1)} \gets \theta'$
-		\Else
-		\State $\theta^{(t+1)} \gets \theta^{(t)}$
-		\EndIf
-		\EndFor
-
-		\State \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$
-
-	\end{algorithmic}
-\end{algorithm}
-
-\clearpage
-\subsection{Hyperparameter values}
-\label{appdx:hyperparams}
-
-\begin{table}[!h]
-	\centering
-	\caption{Hyper-parameter values for each experiment}
-	\label{tab:hyperparams}
-	\resizebox{\textwidth}{!}{%
-		\begin{tabular}{c|ccc|ccc|cccc|cc|cccc}
-			Dataset & 
-			$B$ & $f$ & $\sigma_\theta$ & 
-			$T_b$ & $\kappa_b$ & $\lambda_b$ & 
-			$T_\theta$ & $\kappa_\theta$ & $\lambda_\theta$ & $s$ &
-			$k$ & $D'$ &
-			$T_\theta'$ & $\kappa_\theta'$ & $\lambda_\theta'$ & $s'$
-			\\ \hline
-			Polbooks &
-			3 & 0.7 & 1 &
-			1,000 & 0.2 & 5 &
-			10,000 & 0.4 & 10 & 0.05 &
-			-- & -- & 
-			-- & -- & -- & -- \\
-			School &
-			10 & 0.7 & 1 &
-			1,000 & 0.2 & 5 &
-			10,000 & 0.4 & 10 & 0.2 &
-			1 & 10 & 
-			10,000 & 0.4 & 10 & 0.2 \\
-			FB Egonet &
-			10 & 0.7 & 1 &
-			1,000 & 0.2 & 5 &
-			10,000 & 0.4 & 10 & 0.017 &
-			1 & 10 & 
-			10,000 & 0.4 & 10 & 0.5 \\
-		\end{tabular}
-	}
-\end{table}
-
-\subsection{Implementation details}
-\label{appdx:imp-details}
-
-All data analysis and visualisation was implemented in Python.
-Full source code is available in the supplementary material.
-The scripts were run using a standard PC using the Windows Subsystem for Linux (WSL) environment.
-Specs are:
-
-\begin{itemize}
-	\item \textbf{CPU}: Intel(R) Core(TM) i7-1065G7
-	\item \textbf{RAM}: 8GB
-	\item \textbf{GPU}: Intel(R) Iris(R) Plus Graphics
-\end{itemize}
-
-On this hardware each experiment iteration took the following amount of time to execute:
-
-\begin{table}[!h]
-	\centering
-	\caption{Compute-time for each experiment}
-	\label{tab:compute-time}
-	\begin{tabular}{c|ccc|c}
-		Dataset & $b$-chain & $\theta$-chain & Reduced $\theta$-chain & Overall compute time \\ \hline
-		Polbooks & $\sim$1s & $\sim$4s & -- & $\sim$5s \\
-		School & $\sim$10s & $\sim$10s & $\sim$10s & $\sim$30s \\
-		FB Egonet & $\sim$20s & $\sim$180s & $\sim$10s & $\sim$210s
-	\end{tabular}
-\end{table}

### 2021-05-25 nips/additional.tex
Remove bias

@@
 Indeed, it is prudent to start our $b$-chain at this MDL solution as then we can burn-in time is greatly reduced.
+
+\subsection{Dimensionality discussion}
+
+There is a challenge when dealing with vertex-features.
+Many vertex feature often take only one of several discrete values within a set.
+For example, with the primary school dataset \cite{schools}, a pupil's school-class can take one of 10 values from \{"1A", "1B", "2A", "2B" \dots "5A", "5B"\}.
+
+It is not immediately obvious how to encode this for the feature-to-block classifier.
+Although this is technically only a single dimension, we choose to expand the data into a set of 10 binary feature flags $\in \Xcal = \{0, 1\}$.
+Although only one flag will be set at a time it is the simplest method for representing discrete-valued data.
+The reason we choose $\Xcal=\{0, 1\}$ and not $\Xcal = \{-1, 1\}$ is that the former allows for simpler analysis of the resulting weight values.
+When a feature is switched off, it has no impact on the Softmax classifier.
+It is simpler to only model positive relationships.
+We prefer to say that this block is comprised of the vertices with feature $d$ on rather than this block contains the vertices with feature $d$ switched off.
+Nevertheless, this dimensionality expansion means that we cannot include a bias term in our classifier as then there is no unique MAP solution.

### 2021-05-25 nips/additional.tex
Remove bias sections in report

@@
 \subsection{Dimensionality discussion}
+\label{appdx:dimension}
 
@@
 The reason we choose $\Xcal=\{0, 1\}$ and not $\Xcal = \{-1, 1\}$ is that the former allows for simpler analysis of the resulting weight values.
-When a feature is switched off, it has no impact on the Softmax classifier.
+When a feature is switched off, it has no impact on the softmax classifier.
 It is simpler to only model positive relationships.
 We prefer to say that this block is comprised of the vertices with feature $d$ on rather than this block contains the vertices with feature $d$ switched off.
-Nevertheless, this dimensionality expansion means that we cannot include a bias term in our classifier as then there is no unique MAP solution.
+To make this explicit, say we have a feature encoding a pupil's gender: \{"male", "female"\}, we represent this as 2 binary feature flags "male": \{0, 1\} and "female": \{0, 1\}.
+This approach may seem inefficient but it is vital to ensure we can interpret the resulting parameter distributions.
+
+Nevertheless, this dimensionality expansion means that we cannot include a bias term in our classifier as then the MAP solution is less peaked and our $\theta$-samples will have higher variance.
+To illustrate this point we need only look at the form of the soft-max classifier with bias vector $\beta$, operating on a vector $x$ which only has feature $d$ turned on such that $x_i = \one\{i = d\}$.
+Component $j$ of the output is given by:
+%
+\begin{equation}
+	\phi_j(x) = \frac{\exp(w_j^T x + \beta_j)}{\sum_{k=1}^{B} \exp(w_k^T x + \beta_k)} = 
+	\frac{\exp(w_{jd} + \beta_j)}{\sum_{k=1}^{B} \exp({w_{kd} + \beta_k})}
+	\label{eqn:softmax-bias}
+\end{equation}
+%
+For the case that $x$ can only has one entry $x_d=1$ and the rest are 0, the term $w_{kd} + \beta_k$ is effectively the new weight for component $k$.
+The bias term $\beta_k$ becomes an unnecessary extra degree of freedom.
+To illustrate this point, we set $\beta_k=\beta_0 \forall j$.
+If this is the case then equation \ref{eqn:softmax-bias} can be simplified to:
+%
+\begin{equation}
+	\phi_j(x) = \frac{\exp(\beta_0) \cdot \exp(w_{jd})}{\sum_{k=1}^{B} \exp(\beta_0) \cdot \exp(w_{kd})}
+	= \frac{\exp(w_{jd})}{\sum_{k=1}^{B} \exp(w_{kd})}
+\end{equation}
+%
+This expression is independent of $\beta_0$ and so it is free to vary.
+This is not the whole picture as we also have the prior which introduces a regularisation term which favours weights closer to 0.
+Nevertheless, for mutually exclusive binary feature flags, the bias term does not add to the expressiveness of the model and only serves to complicate analysis.
+We therefore remove the bias term from the Softmax classifier.
+Even for data comprised of several such features (e.g.
+school class and data) we prefer to discard the bias term.
+The bias term only serves to decrease the loss function by leveraging information about the size of each detected block and not feature information.
+We do not wish to be overly confident in our model predictions when such prediction is influenced by block size and not feature information.
+Indeed, the approach was initially developed with a bias term but we found its removal yielded more reliable and reproducible results.

### 2021-05-25 nips/experiments.tex
Remove bias sections in report

@@
 Two nodes are connected if the two parties shared a face-to-face interaction over the school-day.
-Vertex features include class membership (one of 10 values: 1A-5B), gender and teacher/student distinction.
+Vertex features include class membership (one of 10 values: 1A-5B), gender (male, female) and teacher status encoded as an 11th school-class.
 No further identifiable information is retained.
@@
 			Dataset  & $B$ & $D$ & $D'$ & $\bar{S}_e$ & $\bar{\Lcal}_0$ & $\bar{\Lcal}_1$ & $c^*$ & $\bar{\Lcal}_0'$ & $\bar{\Lcal}_1'$  \\ \hline
-			Polbooks & 3 & 3 & -- & $2.250 \pm 0.001$ & $0.584 \pm 0.033$ & $0.557 \pm 0.061$ & -- & -- & -- \\
-			School & 10 & 13 & 10 & $1.894 \pm 0.006$ & $0.793 \pm 0.096$ & $0.914 \pm 0.112$ & $1.112 \pm 0.230$ & $0.791 \pm 0.096$ & $0.864 \pm 0.139$ \\
-			FB egonet & 10  & 480 & 10 & $1.626 \pm 0.003$ & $1.305 \pm 0.034$ & $1.539 \pm 0.087$ & $0.942 \pm 0.042$ & $1.496 \pm 0.093$ & $1.578 \pm 0.104$
+			Polbooks & 3 & 3 & -- & $2.250 \pm 0.000$ & $0.563 \pm 0.042$ & $0.595 \pm 0.089$ & -- & -- & -- \\
+			School & 10 & 13 & 10 & $1.894 \pm 0.004$ & $0.787 \pm 0.127$ & $0.885 \pm 0.129$ & $1.198 \pm 0.249$ & $0.793 \pm 0.132$ & $0.853 \pm 0.132$ \\
+			FB egonet & 10  & 480 & 10 & $1.626 \pm 0.003$ & $1.326 \pm 0.043$ & $1.538 \pm 0.069$ & $0.94 \pm 0.019$ & $1.580 \pm 0.150$ & $1.605 \pm 0.106$
 		\end{tabular}
@@
 This is strong evidence that political affiliation is indeed the axis which best predicts the 3-way natural partition of the graph into blocks.
-Furthermore, table \ref{tab:results} we see that the training and test losses are very similar (the test loss mean is even below the training loss) and both are low in magnitude.
+Furthermore, from table \ref{tab:results} we see that the training and test losses are very similar and both are low in magnitude.
 This provides further evidence to the claim that political affiliation is the best explanatory variable for the overall network structure.

### 2021-05-25 nips/ffbm.tex
Refactor appendices
Remove bias sections in report

@@
 	p(b| X, \theta) = \prod_{i=1}^{N} p(b_i | x_i, \theta) = \prod_{i=1}^{N} \phi_{b_i} (x_i; \theta)
-	= \prod_{i=1}^{N} \frac{\exp\left(w_{b_i}^T \tilde{x}_i\right)}{\sum_{k=1}^{B} \exp \left( w_k^T \tilde{x}_i\right)}
+	= \prod_{i=1}^{N} \frac{\exp\left(w_{b_i}^T x_i\right)}{\sum_{k=1}^{B} \exp \left( w_k^T x_i\right)}
 \end{equation}
 %
-Where $\tilde{x} \coloneqq \left[x_1, x_2, \dots x_D, 1\right]^T$ is an augmented version of $x$ that allows for a bias term.
+We deliberately exclude a bias term to ensure that the relationships we model are based on features and not information about the size of each detected block; a more complete discussion on this topic is given in \ref{appdx:dimension}.
 The parameter vector $\theta$ for this stage contains all the weight vectors $\theta = \{w_k\}_{k=1}^{B}$.
-Each $w_k$ has dimension $D+1$.
-We could instead write the parameters $\theta$ as a $B \times (D+1)$ matrix of weights $W$; this form has computational benefits as then $z_i \coloneqq W \tilde{x}_i$, which is the input to the softmax activation function.
+Each $w_k$ has dimension $D$.
+We could instead write the parameters $\theta$ as a $B \times D$ matrix of weights $W$; this form has computational benefits as then $z_i \coloneqq W x_i$, which is the input to the softmax activation function.
 
@@
 We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$ to complete our model.
-Please refer to appendix \ref{appdx:prior} for the exact form of $p(\psi | b)$.
+Please refer to appendix \ref{appdx:sbm} for the exact form of $p(\psi | b)$.
 All that concerns the main argument is we have a computable form.

### 2021-05-25 nips/additional.tex
Lines on graphs

@@
 
+By way of illustration, we can consider the primary school dataset \cite{schools}.
+We plot the normalised objective function $U\left( \theta^{(t)} \right) / N$ with respect to MALA iteration on figure \ref{fig:school-U-orginal}.
+We see that the chain converges to the modal neighbourhood quickly (within about 2000 iterations of a total 10,000).
+Nevertheless, we pick $\kappa_\theta=0.4 > 0.2$ just to be on the safe-side.
+$\lambda_\theta=10$ is chosen somewhat arbitrarily as we do not require neighbouring samples to be independent.
+Nevertheless, this thinning does speed up computation of quantities that require an average over all the retained samples -- as $|\Tcal_\theta| < T_\theta$.
+%
+\begin{figure}[!h]
+	\centering
+	\begin{subfigure}[t]{0.4\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{school-U-original.png}
+		\caption{Original samples $t \in [T_\theta]$}
+		\label{fig:school-U-orginal}
+	\end{subfigure}
+	\begin{subfigure}[t]{0.4\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{school-U-thinned.png}
+		\caption{Thinned samples $t \in \Tcal_\theta$ indexed by $i$}
+		\label{fig:school-U-thinned}
+	\end{subfigure}
+	\hspace{1cm}
+	\caption{Primary school dataset normalised objective function against sample index}
+\end{figure}
+
+We see on figure \ref{fig:school-U-thinned} that $U(\theta^{(i)})$ does bounce around a small amount.
+This is to be expected we are sampling from the posterior rather than converging to the MAP.
+The variations do appear to be rather long in time-scale but we can live with this as we do not require sample independence.
+\FloatBarrier
 \subsection{Initializing the b-chain}

### 2021-05-25 nips/inference.tex
Correct hyp test

@@
 \begin{equation}
-	H_0: |\mu_{ij}| \leq c \qquad
-	H_1: |\mu_{ij}| > c
+	H_0: |W_{ij}| \leq c \qquad
+	H_1: |W_{ij}| > c
 	\label{eqn:hyp-test-discard}

### 2021-05-26 nips/additional.tex
step-size appendix

@@
 
+Nevertheless, to choose the value of $s$ we must quantify the trade-off between acceptance ratio and burn-in time.
+We do this by way of an example with the primary school dataset \cite{schools}.
+We run the $b$-chain for $T_b = 1,000$ iterations and subsample such that $\Tcal_b$ is computed with $\kappa_b=0.2$ and $\lambda_b=5$.
+We then use this to run the $\theta$-chain for $T_\theta = 10,000$ iterations.
+The acceptance ratio -- which we denote $r_\alpha$ -- is easy enough to compute as simply the fraction of proposed moves which we accept.
+However, to quantify burn-in time we must use a different metric.
+We can plot the mean of the objective function averaged over all samples as a proxy for burn-in time:
+%
+\begin{equation}
+	\bar{U} \coloneqq \frac{1}{T_\theta} \sum_{t \in [T_\theta]} U\left( \theta^{(t)} \right)
+\end{equation}
+%
+As the chain will equilibrate in the vicinity of a minima in $U(\theta)$, the average over all samples is a rough indicator of the speed of burn-in.
+This assumes that the random starting point has high $U(\theta)$.
+The higher the value of $\bar{U}$, the longer the chain took to burn in and reach equilibrium.
+We plot the acceptance ratio $r_\alpha$ and average objective $\bar{U}$ for $10$ values of the step-size multiplier $s$ logarithmically spaced in the range $(10^{-2}, 10^1)$ on figure \ref{fig:step-size-tradeoff}.
+%
+\begin{figure}[!h]
+	\centering
+	\includegraphics[width=0.6\linewidth]{step-size-tradeoff}
+	\caption{Primary school dataset \cite{schools} illustration of step-size tradeoff}
+	\label{fig:step-size-tradeoff}
+\end{figure}
+
+We see immediately that acceptance ratio $r_\alpha$ steadily decreases with $s$.
+This is expected as it is unlikely we stay in a high density region of our target distribution if we are making very large steps each time.
+A low acceptance ratio leads to wasted samples and this inefficiency is obviously undesirable.
+Nevertheless, $\bar{U}$ decreases with $s$ indicating shorter burn-in times for larger step-sizes.
+Nevertheless, this effect plateaus for $s>0.2$ for this particular dataset.
+Therefore, we choose $s=0.2$ as our step-size multiplier for the primary school datasets as this yields the best trade-off between $r_\alpha$ and $\bar{U}$.
+Indeed, the acceptance ratio is still high for $r_\alpha \approx 0.8$ for $s=0.2$.
+The step-sizes for the other datasets were chosen following the same argument.
+
+\FloatBarrier
 \subsection{Burn-in and thinning}

### 2021-05-26 nips/derivations.tex
More dim reduction

@@
 The only property of $y_{ij}$ we have used in the derivation is the sum-to-one constraint $\sum_{j=1}^{B} y_{ij} = 1$ for all $i$.
+
+\subsection{Hypothesis test on feature weights}
+
+We are given samples $\{\theta^{(t)}\}_{t \in \Tcal_\theta} \sim p(\theta | A, X)$ and wish to determine the statistical significance of the weights.
+We adopt matrix notation for simplicity by representing $\theta$ with the matrix $B \times D$ matrix of feature weights $W$.
+The question is to determine the significance of a particular feature $d$ by examining the value of $W_{id}$ for all the values $i \in [B]$.
+We know that the posterior is proportional to the prior multiplied by the likelihood (assuming that the feature matrix $X$ is already given):
+%
+\begin{equation}
+	p(\theta|A, X) \propto p(\theta) \cdot p(A | \theta, X)
+\end{equation}
+%
+The prior term can be evaluated but the likelihood is intractable. The closest we can get is through a Monte-Carlo integration over $b$:
+%
+\begin{align}
+	p(A | \theta, X) &= \sum_{b \in [B]^N} p(A, b | \theta, X) \nonumber \\
+	&= \sum_{b \in [B]^N} p(A | b, \theta, X) \cdot p(b | \theta, X) \nonumber \\
+	&= \sum_{b \in [B]^N} p(A | b) \cdot p(b | \theta, X) \nonumber \\
+	&\approx \sum_{i} p\left( A | b^{(i)} \right) \quad \textrm{with} \quad b^{(i)} \sim p(b| \theta, X)
+	\label{eqn:mc-likelihood}
+\end{align}
+%
+This could be implemented for a single value of $\theta$ but such a form cannot be used to characterise the overall form of the posterior.
+Nevertheless, the form in equation \ref{eqn:mc-likelihood} still does highlight something interesting.
+The likelihood is peaked around areas of $\theta$ that generate a partition $b$ that is highly likely in the SBM sense -- high $p(A|b)$.
+This provides the motivation for using the Laplace approximation for modelling the posterior $p(\theta | A, X) \approx p(\theta; \mu, \Sigma)$.
+Indeed, the Laplace approximation is often used for modelling the posterior in logistic classification \cite{laplace}.
+As a simplification we assume that $\Sigma$ is diagonal, therefore each element of $\theta$ is independent.
+This assumption does not hold in general but it motivates the derivation of the dimensionality reduction we construct by analogy with hypothesis testing.
+Assuming, $\Sigma$ is diagonal then the posterior for each element of the weight matrix $W$ can be approximated by:
+%
+\begin{equation}
+	p(W_{ij}|A, X) \approx \Gaussian(W_{ij}|\hat{\mu}_{ij}, \hat{\sigma}_{ij}^2)
+\end{equation}
+%
+Where we have used the set of samples for $W$ drawn according to the exact posterior, to calculate unbiased estimates for the mean and standard deviation:
+%
+\begin{equation}
+	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
+	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2
+\end{equation}
+%
+The probability

### 2021-05-26 nips/inference.tex
Algorithms
More dim reduction

@@
 	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
-	\hat{\sigma}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2
+	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2
 \end{equation}
@@
 Intuitively, this means discarding any feature for which $\hat{\mu}_{ij} \pm k\hat{\sigma}_{ij}$ lies within or spans the null region $(-c, c)$ for all block indices.
-If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \mu_{ij}, \sigma_{ij})$, then this is effectively a hypothesis test on the value of $W_{ij}$ (equation \ref{eqn:hyp-test-discard}).
+If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \mu_{ij}, \sigma_{ij}^2)$, then this is effectively a hypothesis test on the value of $W_{ij}$ (equation \ref{eqn:hyp-test-discard}).
 $\Dcal'$ then comprises all features $i$ for which $H_1$ is accepted at least once for some $j \in [B]$.
@@
 \end{equation}
+%
+For an algorithmic implementation of this method refer to appendix \ref{appdx:algorithms}.

### 2021-05-26 nips/derivations.tex
More hyp testing

@@
 %
-The probability
+This approximation is not exact but can show it is accurate empirically.
+Indeed, if we run the primary school experiment with hyper-parameters given in appendix \ref{appdx:hyperparams}, we can then plot histograms of the collected $W$-samples and compare these to the Laplace approximation.
+The results are given on figure \ref{fig:school-histogram}.
+Even though the Laplace approximation is not exact, it is remarkably reliable.
+
+\begin{figure}[!h]
+	\centering
+	\begin{subfigure}[t]{0.3\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{school-sample-histogram-16.png}
+		\caption{Weight for block 1, class 3B}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[t]{0.3\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{school-sample-histogram-37.png}
+		\caption{Weight for block 3, class 4A}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[t]{0.3\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{school-sample-histogram-84.png}
+		\caption{Weight for block 8, class 2B}
+	\end{subfigure}
+
+	\caption{Histograms of sampled weights for the primary school experiment \cite{schools}. Dotted line is the applied Laplace approximation.}
+	\label{fig:school-histogram}
+\end{figure}
+
+Using the approximation, we can construct a test to determine whether a particular weight value has $|W_{ij}| > c$ with high probability. Assuming without loss of generality that
+%
+\begin{equation}
+	p(|W_{ij}| > c) = p(W_{ij})
+\end{equation}

### 2021-05-27 nips/derivations.tex
More hyp testing

@@
 \subsection{Hypothesis test on feature weights}
+\label{appdx:hyp-test}
 
@@
 
-Using the approximation, we can construct a test to determine whether a particular weight value has $|W_{ij}| > c$ with high probability. Assuming without loss of generality that
+Using the approximation, we wish to construct a test to determine whether a particular weight value has $|W_{ij}| > c$ with high probability, where $c>0$. More specifically, we wish to determine:
 %
 \begin{equation}
-	p(|W_{ij}| > c) = p(W_{ij})
+	\begin{aligned}
+		H_0&: |W_{ij}| \leq c \\
+		H_1&: |W_{ij}| > c
+	\end{aligned}
 \end{equation}
+%
+We can consider the probabilities $p(W_{ij} > c)$ and $p(W_{ij} < c)$ separately. Starting with $p(W_{ij} > c)$ we can write:
+%
+\begin{align*}
+	p(W_{ij} > c) &= p\left(Z > \frac{c - \hat{\mu}_{ij}}{\hat{\sigma}_{ij}}\right) \\
+	&= p\left(-Z < \frac{\hat{\mu}_{ij} - c}{\hat{\sigma}_{ij}}\right)\\
+	&= \Phi\left(\frac{\hat{\mu}_{ij} - c}{\hat{\sigma}_{ij}}\right)
+\end{align*}
+%
+Where $Z \sim \Gaussian(0,1)$ was introduced as a temporary variable and $\Phi(\cdot)$ is the standard Gaussian cdf.
+Now we introduce the variable $k$ to control the degree of significance of the result.
+By enforcing $k$ to be less than the argument of $\Phi$ we show that:
+%
+\begin{align*}
+	\frac{\hat{\mu}_{ij} - c}{\hat{\sigma}_{ij}} &> k \\
+	\hat{\mu}_{ij} - k \hat{\sigma}_{ij} &> c
+\end{align*}
+%
+As $\Phi(\cdot)$ is monotonically increasing, this result gives us that:
+%
+\begin{equation}
+	p(W_{ij} > c) > \Phi(k) \iff \hat{\mu}_{ij} - k \hat{\sigma}_{ij} > c
+	\label{eqn:hyp-test-positive}
+\end{equation}
+%
+A similar argument can be made to show that:
+%
+\begin{equation}
+	p(W_{ij} < -c) > \Phi(k) \iff \hat{\mu}_{ij} + k \hat{\sigma}_{ij} < -c
+	\label{eqn:hyp-test-negative}
+\end{equation}
+%
+Clearly equations \ref{eqn:hyp-test-positive} and \ref{eqn:hyp-test-negative} cannot both hold as $c>0$. Nevertheless, a necessary and sufficient condition for one of them holding is:
+%
+\begin{equation}
+	(\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c) = \emptyset
+	\label{eqn:hyp-test-empty}
+\end{equation}
+%
+We therefore choose to reject the null hypothesis $H_0$ if and only if equation \ref{eqn:hyp-test-empty}.
+This tells us with confidence greater than $\Phi(k)$ that either $W_{ij} < -c$ or $W_{ij} > c$ thus rejecting the null $H_0: |W_{ij}| \leq c$.
+Note that we choose to accept $H_0$ for the case where the combined probabilities of $p(W_{ij} < -c) + p(W_{ij} > c) > \Phi(k)$, if equation \ref{eqn:hyp-test-empty} is not satisfied.
+This is because we wish to apply this for dimensionality reduction purposes.
+We only want to keep weights that have a confident prediction outside the null-space not those that succeed by summing the two extremes of their distribution.
+Note that this case is very rare empirically.

### 2021-05-27 nips/inference.tex
More hyp testing

@@
 	\label{eqn:fij} \\
-	\Dcal' &\coloneqq \left\{ j \in [D] : \exists i \in [B] \textrm{ s.t. }  \Fcal_i(j) \neq \emptyset \right\}
+	\Dcal' &\coloneqq \left\{ j \in [D] : \exists i \in [B] \textrm{ s.t. }  \Fcal_i(j) = \emptyset \right\}
 	\label{eqn:kept-feature-set}
@@
 Intuitively, this means discarding any feature for which $\hat{\mu}_{ij} \pm k\hat{\sigma}_{ij}$ lies within or spans the null region $(-c, c)$ for all block indices.
-If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \mu_{ij}, \sigma_{ij}^2)$, then this is effectively a hypothesis test on the value of $W_{ij}$ (equation \ref{eqn:hyp-test-discard}).
+If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \mu_{ij}, \sigma_{ij}^2)$, then this is analogous to a hypothesis test on the value of $W_{ij}$ (equation \ref{eqn:hyp-test-discard}).
 $\Dcal'$ then comprises all features $i$ for which $H_1$ is accepted at least once for some $j \in [B]$.
@@
 %
-The multiplier $k$ determines the degree of significance of the result.
+The multiplier $k$ in equation \ref{eqn:fij} determines the degree of significance of the result.
+Appendix \ref{appdx:hyp-test} explores the derivation in more detail.
 However, as the Laplace approximation is not exact we will only treat this dimensionality reduction method as a useful heuristic and not an exact method.

### 2021-05-27 nips/additional.tex
More dim reduction
Proof-read appendices

@@
 
-For reference, we provide the likelihoods and priors for the mircocanonical SBM proposed by \citet{Peixoto-Bayesian-Microcanonical}. We have that the graph $A$ is drawn from the SBM:
+For reference, we provide the likelihood and priors for the microcanonical SBM proposed by \citet{Peixoto-Bayesian-Microcanonical}. We have that the graph $A$ is drawn from the SBM:
 %
@@
 With edges placed uniformly at random but respecting the constraints imposed by $b, e$ and $k$.
-The likelihood calculation for $p(A|k, e, b)$ then reduces to a case of counting configurations that yield the same adjacency matrix, $\Xi (A)$, and dividing by the total number of configurations possible, $\Xi(e)$.
+The likelihood calculation for $p(A|k, e, b)$ then reduces to a case of counting configurations that yield the same adjacency matrix, $\Xi (A)$, and dividing by the total number of configurations possible, $\Omega(e)$.
 This is why this formulation is given the microcanonical moniker.
@@
 \begin{equation}
-	\Xi(A) \coloneqq \frac{\prod_i k_i !}{\prod i,j : i < j A_{ij} ! \prod_i A_{ii} !! }
+	\Xi(A) \coloneqq \frac{\prod_i k_i !}{\prod_{i,j : i < j} A_{ij} ! \prod_i A_{ii} !! }
 \end{equation}
@@
 With the likelihood defined, we move on to the prior.
-As discussed in the main text, for the FFBM $b$ is an intermediate variable and not a parameter so we are not free to choose a prior for it.
-Nevertheless, we can borrow the conditional prior proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(e, k | b)$:
+As discussed in the main text, the block membership vector $b$ is an intermediate variable in the FFBM and not a parameter so we are not free to choose a prior for it.
+Nevertheless, we can borrow the conditional prior proposed by \citet{Peixoto-Bayesian-Microcanonical} for the other parameters to give:
 %
@@
 
-For sampling from the $\theta$-chain of the block membership generator parameters, we employed the Metropolis Adjusted Langevin Algorithm (MALA). At iteration $t$, the proposed sample is generated by:
+For sampling from the $\theta$-chain of the block membership generator parameters, we employ the Metropolis Adjusted Langevin Algorithm (MALA). At iteration $t$, the proposed sample is generated by:
 %
@@
 %
-The first constraint ensures that we have cover sufficient distance to arrive at any arbitrary point in our domain, no matter the starting point.
-The second constraint ensures that once we converge to the mode rather than simply bouncing around it.
+The first constraint ensures that we can cover sufficient distance to arrive at any arbitrary point in our domain, no matter the starting point.
+The second constraint ensures that once we converge to the mode we will stay there, rather than simply bouncing around it.
 \citet{Bayesian-SGLD} propose the following form for a polynomially decaying step-size which we adopt:
@@
 We require $\alpha,\beta > 0$ and $\gamma \in (0.5, 1]$ to satisfy equation \ref{eqn:h-constraints}.
-To reduce the number of hyperparameters we set these to have values given by the equations \ref{eqn:step-size-params}.
+To reduce the number of hyperparameters we set these to have values given by the equations in \ref{eqn:step-size-params}.
 %
@@
 If this is done, the algorithm is instead called stochastic gradient Langevin diffusion (SGLD) \cite{Bayesian-SGLD}.
-This speeds up computation at the expense of exactness of the method.
+This speeds up computation at the expense of the exactness of the method.
+Indeed, it relies on the fact that as the step-size is annealed to 0, the acceptance probability is close 1.
 
@@
 This assumes that the random starting point has high $U(\theta)$.
-The higher the value of $\bar{U}$, the longer the chain took to burn in and reach equilibrium.
+This assumption holds as the domain of $\theta$ is a high dimensional space so we are unlikely to get a low $U(\theta)$ simply from a lucky guess.
+The higher the value of $\bar{U}$, the longer we infer the chain took to burn in and reach equilibrium.
 We plot the acceptance ratio $r_\alpha$ and average objective $\bar{U}$ for $10$ values of the step-size multiplier $s$ logarithmically spaced in the range $(10^{-2}, 10^1)$ on figure \ref{fig:step-size-tradeoff}.
@@
 A low acceptance ratio leads to wasted samples and this inefficiency is obviously undesirable.
+Indeed, figure \ref{fig:step-size-tradeoff} suggests that a low step-size $s$ yields high acceptance probability thus justifying the approximation used in SGLD.
+
 Nevertheless, $\bar{U}$ decreases with $s$ indicating shorter burn-in times for larger step-sizes.
-Nevertheless, this effect plateaus for $s>0.2$ for this particular dataset.
+However, this effect plateaus for $s>0.2$ for this particular dataset.
 Therefore, we choose $s=0.2$ as our step-size multiplier for the primary school datasets as this yields the best trade-off between $r_\alpha$ and $\bar{U}$.
 Indeed, the acceptance ratio is still high for $r_\alpha \approx 0.8$ for $s=0.2$.
-The step-sizes for the other datasets were chosen following the same argument.
+The step-sizes for the other datasets were chosen following the same argument but we do not repeat the process here.
 
@@
 This is to be expected we are sampling from the posterior rather than converging to the MAP.
-The variations do appear to be rather long in time-scale but we can live with this as we do not require sample independence.
+There does appear to be variation over a long time-scale but we can live with this as we do not require sample independence.
 \FloatBarrier
@@
 new blocks can be created and we permit empty blocks) then it can be run  until a minimum description length (MDL) solution is reached.
-We take the number of non-empty blocks at the MDL to be our fixed block number $B$ for subsequent analysis.
+We take the number of non-empty blocks at the MDL solution to be our fixed block number $B$ for subsequent analysis.
 Indeed, it is prudent to start our $b$-chain at this MDL solution as then we can burn-in time is greatly reduced.
@@
 Although only one flag will be set at a time it is the simplest method for representing discrete-valued data.
-The reason we choose $\Xcal=\{0, 1\}$ and not $\Xcal = \{-1, 1\}$ is that the former allows for simpler analysis of the resulting weight values.
-When a feature is switched off, it has no impact on the softmax classifier.
-It is simpler to only model positive relationships.
-We prefer to say that this block is comprised of the vertices with feature $d$ on rather than this block contains the vertices with feature $d$ switched off.
+The reason we choose $\Xcal=\{0, 1\}$ and not $\Xcal = \{-1, 1\}$ is that the former allows for simpler analysis of the resulting weight values; when a feature $d$ is switched off then $x_{d}=0$ given $\Xcal = \{0, 1\}$, and the feature has no impact on the softmax layer.
+
+It is more intuitive to only model positive relationships.
+In other words, we prefer to say that this block is comprised of the vertices with feature $d$ turned on rather than this block contains the vertices with feature $d$ switched off.
 To make this explicit, say we have a feature encoding a pupil's gender: \{"male", "female"\}, we represent this as 2 binary feature flags "male": \{0, 1\} and "female": \{0, 1\}.
@@
 Nevertheless, this dimensionality expansion means that we cannot include a bias term in our classifier as then the MAP solution is less peaked and our $\theta$-samples will have higher variance.
-To illustrate this point we need only look at the form of the soft-max classifier with bias vector $\beta$, operating on a vector $x$ which only has feature $d$ turned on such that $x_i = \one\{i = d\}$.
+To illustrate this point we need only look at the form of the soft-max classifier with bias vector $\beta$, operating on a vector $x$ which only has one non-zero entry $x_d=1$.
 Component $j$ of the output is given by:
@@
 \begin{equation}
-	\phi_j(x) = \frac{\exp(w_j^T x + \beta_j)}{\sum_{k=1}^{B} \exp(w_k^T x + \beta_k)} = 
-	\frac{\exp(w_{jd} + \beta_j)}{\sum_{k=1}^{B} \exp({w_{kd} + \beta_k})}
+	\phi_j(x) = \frac{\exp(w_j^T x + \beta_j)}{\sum_{k \in [B]} \exp(w_k^T x + \beta_k)} = 
+	\frac{\exp(w_{jd} + \beta_j)}{\sum_{k \in [B]} \exp({w_{kd} + \beta_k})}
 	\label{eqn:softmax-bias}
@@
 %
-For the case that $x$ can only has one entry $x_d=1$ and the rest are 0, the term $w_{kd} + \beta_k$ is effectively the new weight for component $k$.
-The bias term $\beta_k$ becomes an unnecessary extra degree of freedom.
+The term $w_{kd} + \beta_k$ is effectively the new weight for component $k$ as this term always appears together.
+The bias term $\beta_k$ is an unnecessary extra degree of freedom.
 To illustrate this point, we set $\beta_k=\beta_0 \forall j$.
@@
 \begin{equation}
-	\phi_j(x) = \frac{\exp(\beta_0) \cdot \exp(w_{jd})}{\sum_{k=1}^{B} \exp(\beta_0) \cdot \exp(w_{kd})}
-	= \frac{\exp(w_{jd})}{\sum_{k=1}^{B} \exp(w_{kd})}
+	\phi_j(x) = \frac{\exp(\beta_0) \cdot \exp(w_{jd})}{\sum_{k \in [B]} \exp(\beta_0) \cdot \exp(w_{kd})}
+	= \frac{\exp(w_{jd})}{\sum_{k \in [B]} \exp(w_{kd})}
 \end{equation}
@@
 Even for data comprised of several such features (e.g.
-school class and data) we prefer to discard the bias term.
+school class and gender) we prefer to discard the bias term.
 The bias term only serves to decrease the loss function by leveraging information about the size of each detected block and not feature information.
-We do not wish to be overly confident in our model predictions when such prediction is influenced by block size and not feature information.
+We do not wish to be overly confident in our block predictions when such predictions are influenced by block size and not feature information.
 Indeed, the approach was initially developed with a bias term but we found its removal yielded more reliable and reproducible results.

### 2021-05-27 nips/abstract.tex
Yiannis version

@@
 \begin{abstract}
-	Labelled networks are an extremely common and important form of data.
-A typical inference goal is to determine how the vertex labels (called features) affect graphical structure.
-The standard approach to this problem has been to partition the network into blocks grouped by distinct values of the feature of interest.
-A block-based random graph model - typically a variant of the stochastic block model (SBM) - is then used to test for evidence of asymmetric behaviour within these feature-based communities.
-
-	Nevertheless, these feature-based communities are often not a natural partition of the graph and thus the models employed are rarely a good fit.
-With this in mind, we present a novel generative model, which we call the feature-first block model (FFBM), for better describing vertex-labelled undirected graphs.
-This allows us to perform richer queries on labelled networks.
-We present a method to efficiently sample the FFBM parameters for inference.
-The FFBM's structure is kept deliberately simple to retain easy interpretability of the parameter values.
-
-	We apply the developed methods to a variety of network data to extract the most important features along which the vertices divide themselves.
-The greatest advantage of the proposed approach is that the whole feature-space is used automatically and features can be rank-ordered implicitly according to impact.
-Any features that do not greatly impact the high-level structure can be discarded to reduce the problem dimension.
-In the case the vertex features available do not readily explain the community structure in the resulting network, the approach detects this and is protected against over-confidence.
-Future work may benefit from extending the FFBM to multiple hierarchical levels.
-This would allow the structure to be explained at each level of coarseness rather than simply at the highest level.
+	Labelled networks form a very common and important class of data,
+	naturally appearing in numerous applications in science and engineering.
+	A typical inference goal is to determine how the vertex labels
+	(or {\em features}) affect the network's graph structure. A standard
+	approach has been to partition the network into blocks grouped
+	by distinct values of the feature of interest. A block-based random
+	graph model -- typically a variant of the stochastic block model --
+	is then used to test for evidence of asymmetric behaviour within these
+	feature-based communities. Nevertheless, the resulting communities
+	often do not produce a natural partition of the graph. In this work
+	we introduce a new generative model, the feature-first block model (FFBM),
+	which is more effective at describing vertex-labelled undirected
+	graphs and also facilitates the use of richer queries on labelled networks.
+	We develop a Bayesian framework for inference with this model,
+	and we present a method to efficiently sample from the posterior
+	distribution of the FFBM parameters. The FFBM's structure is kept
+	deliberately simple to retain easy interpretability of the parameter
+	values. We apply the proposed methods to a variety of network data
+	to extract the most important features along which the vertices
+	are partitioned. The main advantages of the proposed approach are
+	that the whole feature-space is used automatically, and features
+	can be rank-ordered implicitly according to impact. Any features
+	that do not significantly impact the high-level structure can be
+	discarded to reduce the problem dimension. In cases where the vertex
+	features available do not readily explain the community structure
+	in the resulting network, the approach detects this and is protected
+	against over-fitting. Results on several real-world datasets
+	illustrate the performance of the proposed methods. 
 \end{abstract}

### 2021-05-27 nips/appendices.tex
Yiannis version

--- 2021-05-23 nips/appendices.tex
@@
 \section{Appendix}
+
+\subsection{SBM prior choice explanation}
+\label{appdx:prior}
+
+Here we recall for reference
+the priors $p(\psi | b)$ from \cite{Peixoto-Bayesian-Microcanonical}:
+%
+\begin{equation}
+	p(\psi_e=e, \psi_k=k | b) = p(e | b) p(\psi_k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
+	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right],
+\end{equation}
+%
+where $\specialchoose{n}{m}$ is shorthand 
+for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$,
+which can be thought of as the total number of distinct histograms 
+produced by $m$ samples in $n$ bins.
+The value
+$E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges in the graph. 
+Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform in $e$.
+The variable $\eta_j^r$ denotes the number of vertices in block $r$ 
+that have degree $j$; formally, $\eta_j^r \coloneqq \sum_{i} \one\left\{b_i = r \right\} \one \left\{k_i = j \right\}$. 
+The denominator $q(m, n)$ denotes the number of different histograms 
+produced by $m$ samples in 
+at most $n$ non-zero bins that sum to $m$. 
+Finally, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number 
+of half edges in block $r$ and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ 
+is the number of vertices assigned to block $r$.
+
+These were chosen carefully in \cite{Peixoto-Bayesian-Microcanonical} to 
+more closely match the structure of empirical networks than simple 
+uniform priors. We do not repeat these arguments here.
 
@@
 
-We wish to determine the form of $p(b| X)$. This can be done by integrating over the joint probability with respect to $\theta$.
+We determine the form of $p(b| X)$ by integrating
+out the parameters $\theta$. From the definitions we have:
 %
@@
 	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i \in [N] } \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
-	&= \prod_{i \in [N]} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j \in [B]} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k \in [B]} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}
+	&= \prod_{i \in [N]} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j \in [B]} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k \in [B]} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}.
 \end{align*}
 %
-We note that $b_i \in [B]$ and so the integral's value is unchanged with respect to $b_i$.
-The integrand has the same form no matter which value $b_i$ takes as the prior is the same for each $w_j$.
-As such the integral can only be a function of at most $\tilde{x}_i$ and $\sigma_\theta^2$ as it is symmetric with respect to $b_i$ and all the various $w_j$ are integrated out as they are dummy variables.
-Therefore, denoting the integral by the (unknown) function $f(\tilde{x}_i, \sigma_\theta^2)$, we write $p(b| X)$ as follows:
-%
-\begin{align*}
-	p(b | X) &= \prod_{i=1}^{N} f(\tilde{x}_i, \sigma_\theta^2) = \textrm{const w.r.t } b = c
-\end{align*}
-%
-As this is a constant with respect to $b$ we conclude that $p(b | X)$ must be a uniform distribution.
-$\nicefrac{1}{c}$ is simply the size of the set of values that $b$ can take.
-We know $b_i \in [B]$.
-Therefore, $b \in [B]^N$ and $|[B]^N| = B^N = \nicefrac{1}{c}$.
-Putting this all together we conclude that:
+The key observation here is that 
+the value of the integral is independent
+of the value of $b_i \in [B]$:
+The integrand has the same form regardless of $b_i$,
+because the prior is the same for each $w_j$. 
+Therefore, the integral is only a function of $\tilde{x}_i$ and $\sigma_\theta^2$,
+which means that, as function of $b$, $p(b|X)\propto 1$. And as
+$b$ takes values in $[B]^N$, we necessarily have:
 %
 \begin{equation}
-	p(b | X) = B^{-N}
+	p(b | X) = \frac{1}{\big|[B]^N\big|}=B^{-N}.
 \end{equation}
 
-\subsection{Derivation of U gradient with respect to feature parameters}
-\label{appdx:gradu}
-The goal is to determine $\nabla U(\theta)$, the gradient of the negative log posterior with respect to the parameters. We repeat the form of $U(\theta)$ in equation \ref{eqn:U-form-appdx}.
+\subsection{Derivation of {\boldmath $\;U(\theta)$} }
+\label{appdx:form-U}
+
+Recall from~(\ref{eq:U}) is Section~\ref{s:sfb} that,
+$$	\pi_\theta(\theta) \propto p(\theta | X, b) \propto p(b | X, \theta) p(\theta) \propto  \exp \left( - U(\theta) \right),
+$$ 
+so that $U$ can be expressed as,
+$$
+	U(\theta) 
+= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}.
+$$
+Writing,
+$y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and 
+$a_{ij} \coloneqq \phi_j(x_i; \theta)$, we have,
 %
 \begin{equation}
-	U(\theta) = \left( \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form-appdx}
+	\log p(b | X, \theta) = \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
+	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} 
+\|\theta \|^2,
+	\label{eqn:U-constituent-terms}
 \end{equation}
 %
-Where $y_{ij}$ is independent of $\theta$ and $a_{ij}$ is the output from the softmax layer, with form as given in equation \ref{eqn:a-ij}.
-%
-\begin{equation}
-	a_{ij} \coloneqq \phi_{j} (x_i; \theta) = \frac{\exp(w_j^T \tilde{x}_i)}{\sum_{r \in [B]} \exp(w_r^T \tilde{x}_i)}
-	\label{eqn:a-ij} 
-\end{equation}
-%
-We note that $\theta = \{w_k\}_{k=1}^B$, and as such we can write this in vector form $\theta = \left[w_1^T, w_2^T \dots w_B^T  \right]^T$.
-Therefore, $\nabla U(\theta) = \left[\nicefrac{\partial U}{\partial w_1}^T,\nicefrac{\partial U}{\partial w_2}^T \dots \nicefrac{\partial U}{\partial w_B}^T  \right]^T$; to compute $\nabla U(\theta)$ it suffices to find the form of $\nicefrac{\partial U}{\partial w_k}$ with respect to a general $k$.
+where
+$\|\theta\|^2 = \sum_{i} \theta_{i}^2 = \sum_{j=1}^{B} \|w_j\|^2$ 
+is the Euclidean norm of the vector of parameters $\theta$.
+Therefore, discarding constant terms, we 
+obtain exactly the representation~(\ref{eqn:U-form}), as claimed.
 
-To this end, we must first find partial derivatives of $a_{ij}$ and $||\theta||$ with respect to $w_k$. Starting with $a_{ij}$:
+\subsection{Derivation of {\boldmath $\;\nabla U(\theta)$}}
+\label{appdx:gradu}
+Here we show how the gradient 
+$\nabla U(\theta)$ can be computed explicitly.
+Recall the expression for $U(\theta)$ in~(\ref{eqn:U-form}).
+Writing $\theta$ as
+$\theta = \left[w_1^T, w_2^T \dots w_B^T  \right]^T$,
+in order to compute the gradient
+$\nabla U(\theta)$ we need to compute
+each of its components,
+$\nicefrac{\partial U}{\partial w_k}$,
+$1\leq k\leq B$.
+To that end, we first compute,
 %
@@
 	{\left( \sum_{r \in [B]} \exp(w_r^T \tilde{x}_i) \right)^2} \nonumber \\
-	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right) 
+	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right), 
+	\label{eq:dadw}
 \end{align}
 %
-Where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$. Now moving onto the derivative of $||\theta||^2$:
+where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$,
+and we also easily find,
 %
 \begin{equation}
-	\frac{ \partial}{\partial w_k} ||\theta||^2 = \frac{\partial}{\partial w_k} \left( \sum_{r \in [B]} ||w_r||^2 \right) = 2w_k
+	\frac{ \partial}{\partial w_k} \|\theta\|^2 = \frac{\partial}{\partial w_k} \left( \sum_{r \in [B]} \|w_r\|^2 \right) = 2w_k.
+\label{eq:dtsdw}
 \end{equation}
 %
-We are ready to put this all together, to find the partial derivative of $U(\theta)$ with respect to each $w_k$:
+Using~(\ref{eq:dadw}) and~(\ref{eq:dtsdw}), we obtain,
 \begin{align}
@@
 	\sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} 
-	\left( \frac{-\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
+	\left( -\frac{\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
 	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
@@
 	- \frac{w_k}{\sigma_\theta^2} \right) \nonumber \\
-	&= - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
+	&= - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right).
 \end{align}
 %
-This is the required result.
-This form can be computed efficiently through matrix operations.
-The only property of $y_{ij}$ we have used in the derivation is the sum-to-one constraint $\sum_{j=1}^{B} y_{ij} = 1$ for all $i$.
+This can be computed 
+efficiently through matrix operations. The only property of $y_{ij}$ 
+we have used in the derivation is the constraint $\sum_{j=1}^{B} y_{ij} = 1$,
+for all $i$.
 
@@
 
-For sampling from the $\theta$-chain of the block membership generator parameters, we employed the Metropolis Adjusted Langevin Algorithm (MALA). At iteration $t$, the proposed sample is generated by:
+Recall that in 
+Section~\ref{s:sfb} we used 
+the Metropolis Adjusted Langevin Algorithm (MALA)
+in order to 
+sample from the $\theta$-chain of the block membership 
+generator parameters.
+At iteration $t$, the proposed sample is generated by:
 %
 \begin{equation}
-	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi
+	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi.
 \end{equation}
@@
 There are two competing objectives when choosing the step-size $h_t$. 
-On the one hand, we want the step-size to be large so that we arrive at a high density region quickly.
-However, too large a step-size will lead to a lower acceptance ratio and thus inefficient sampling.
-A solution to this problem would be to slowly decrease the step-size with $t$ - often called simulated annealing.
-Therefore, we still have a short burn-in time but will not bounce around the mode for large $t$.
-As well as the trivial constraint for $h_t$ to be strictly positive, we introduce two further constraints as outlined by \citet{Bayesian-SGLD}:
+On the one hand, $h_t$ needs to be large so that the sampler
+arrives at a high density region quickly,
+while too large a step-size would lead to low acceptance rates and thus 
+inefficient sampling. An effective strategy is
+to use {\em simulated annealing}: allow $h_t$ to slowly decrease
+with $t$, as long as $h_t>0$ for all $t$ and also:
 %
 \begin{equation}
-	\sum_{t=1}^{\infty} h_t = \infty \qquad \textrm{and} \qquad
-	\sum_{t=1}^{\infty} h_t^2 < \infty
+	\sum_{t=1}^{\infty} h_t = \infty, \qquad \textrm{and} \qquad
+	\sum_{t=1}^{\infty} h_t^2 < \infty.
 	\label{eqn:h-constraints}
@@
 %
-The first constraint ensures that we have cover sufficient distance to arrive at any arbitrary point in our domain, no matter the starting point.
-The second constraint ensures that once we converge to the mode rather than simply bouncing around it.
-\citet{Bayesian-SGLD} propose the following form for a polynomially decaying step-size which we adopt:
+Following \citet{Bayesian-SGLD}, we adopt the 
+polynomially decaying step-sizes,
+%
+$h_t = \alpha(\beta + t)^{-\gamma}$,
+%
+where $\alpha>0$, $\beta>0$ and $\gamma\in(1/2,1]$ are hyper-parameters.
+We make the specific choices,
 %
 \begin{equation}
-	h_t = \alpha(\beta + t)^{-\gamma}
-\end{equation}
-%
-Where $\alpha, \beta, \gamma$ are hyper-parameters to be chosen.
-We require $\alpha,\beta > 0$ and $\gamma \in (0.5, 1]$ to satisfy equation \ref{eqn:h-constraints}.
-To reduce the number of hyperparameters we set these to have values given by the equations \ref{eqn:step-size-params}.
-%
-\begin{equation}
-	\alpha = \frac{250 \cdot s}{N} \qquad \beta = 1000 \qquad \gamma = 0.8
+	\alpha = \frac{250 \cdot s}{N}, \qquad \beta = 1000, \qquad \gamma = 0.8,
 	\label{eqn:step-size-params}
@@
 %
-Where $N$ is the number of data-points we are considering and now $s$ is the only free variable which we call the step-size scaling.
-For approximate methods, we can choose to bypass the MH accept-reject entirely to speed up computation.
+where $N$ is the number of data-points and $s$,
+the {\em step-size scaling}, is the only free parameter.
+
+% For approximate methods, we can choose to bypass the MH accept-reject entirely to speed up computation.
 If this is done, the algorithm is instead called stochastic gradient Langevin diffusion (SGLD) \cite{Bayesian-SGLD}.
@@
 
-As with any MCMC method, we must deal with the issues presented by burn-in and thinning.
-We have introduced the notation $\Tcal_b$ and $\Tcal_\theta$ to denote the set of samples we keep from the $b$ and $\theta$ chains respectively.
-Note that we generate $T_b$ and $T_\theta$ samples total.
-The burn-in period refers to the time taken for the Markov Chain to converge to the stationary distribution.
-Sample thinning is necessary to ensure that neighbouring samples satisfy independence.
-However, as we do not leverage the independence property this is less important in our analysis.
-We can write the general set $\Tcal_\star$ as:
+% As with any MCMC method, we must deal with the 
+% issues presented by burn-in and thinning.
+
+When sampling from the $b$- and $\theta$-chains described
+in Section~\ref{sec:inference}, we generate
+$T_b$ and $T_\theta$ samples total, respectively.
+We discard an initial proportion $\kappa_\star\in(0,1)$ of the samples 
+as corresponding to a ``burn-in'' period required for the distribution 
+of the chain to reach a distribution close to our target,
+and we also ``thin'' the remaining samples to 
+obain a less-dependent version. For $\star\in\{b,\theta\}$,
+the remaning sample sets are denoted $\Tcal_\star$
+in the notation of Section~\ref{s:ss},
 %
@@
 	\Tcal_\star = \{T_\star \kappa_\star + i \lambda_\star :  
-	0 \leq i \leq \lfloor T_\star(1 - \kappa_\star) / \lambda_\star \rfloor \}
+	0 \leq i \leq \lfloor T_\star(1 - \kappa_\star) / \lambda_\star \rfloor \},
 \end{equation}
 %
-Where the parameter $\kappa_\star \in (0, 1)$ controls our burn-in and $\lambda_\star$ controls our thinning.
-$\kappa_\star$ can be determined by plotting the log-target (either $S(b^{(t)})$ or $U(\theta^{(t)})$ with respect to the epoch $t$.
-$\kappa_\star$ is then chosen to encompass the region where the log-target has roughly equilibrated.
-As we do not leverage sample independence $\lambda_\star$ can be chosen less rigorously.
-We often just use $\lambda_b=5$ and $\lambda_\theta = 10$.
+where $\lambda_\star$ controls the thinning. The choice of
+$\kappa_\star$ can be determined by plotting the log-target (either $S(b^{(t)})$ 
+or $U(\theta^{(t)})$ as a function of $t$,
+and choosing $\kappa_\star$ to encompass the region where the log-target has 
+roughly reached equilibrium. As we do not leverage sample independence,
+$\lambda_\star$ can be chosen less rigorously; we often simply
+use $\lambda_b=5$ and $\lambda_\theta = 10$.
 
@@
 
-For the purposes of our model (the FFBM), the number of blocks $B$ is a constant which must be specified by the data scientist.
-We could however, allow our choice of $B$ to be influenced by the observed data.
-This places us in the domain of empirical Bayes, which must be negotiated carefully.
-Prior beliefs must be determined a priori else they are not prior.
-However, as the number of blocks only specifies the coarseness of the analysis, it is fine to allow it to vary.
-Indeed, \citet{peixoto-determine-B} shows that for a fixed average degree the maximum number of detectable blocks scales as $O(\sqrt{N})$ where $N$ is the number of vertices.
+For the purposes of the FFBM model, the number of blocks $B$ is a constant 
+which must be specified. If the choice of $B$ is influenced 
+by the observed data, then the analysis is no longer ``fully Bayesian''
+and belongs to the class of methods referred to as ``empirical Bayes.''
+However, as the number of blocks only specifies the coarseness of the 
+analysis, it is reasonable to allow it to vary. Indeed, 
+\citet{peixoto-determine-B} shows that for a fixed 
+average degree the maximum number of detectable blocks scales 
+as $O(\sqrt{N})$ where $N$ is the number of vertices.
 
-If we allow $B$ to vary in the $b$-chain (i.e.
-new blocks can be created and we permit empty blocks) then it can be run  until a minimum description length (MDL) solution is reached.
-We take the number of non-empty blocks at the MDL to be our fixed block number $B$ for subsequent analysis.
-Indeed, it is prudent to start our $b$-chain at this MDL solution as then we can burn-in time is greatly reduced.
+If $B$ is allowed to vary in the $b$-chain (i.e.,
+when new blocks can be created and empty blocks are allowed),
+then the chain can be run until a minimum description 
+length (MDL) solution is reached. We take the number of non-empty blocks 
+in the MDL solution to be our fixed block number $B$ for subsequent analysis. 
+Indeed, it is prudent to start the $b$-chain at this MDL solution as then 
+the necessary burn-in time can be greatly reduced.
 
 \subsection{Algorithms}
+\label{appdx:algorithms}
 
@@
 	\centering
-	\caption{Hyper-parameter values for each experiment}
+	\caption{Hyper-parameter values for each experiment.}
 	\label{tab:hyperparams}
@@
 	\centering
-	\caption{Compute-time for each experiment}
+	\caption{Compute-time for each experiment.}
 	\label{tab:compute-time}

### 2021-05-27 nips/conclusion.tex
Yiannis version

@@
 
-The FFBM was developed to address the shortcomings of other graphical models when testing how vertex features affect community structure.
-The idea is to divide the graph into its most natural partition and test whether the vertex features can accurately explain this partition.
-It is very easy to find vertex features that are in some way correlated with the graphical structure.
-Nonetheless, only when we find the feature that best describes the most pronounced partition do we have a stronger case for causation.
+The Feature-First Block Model (FFBM) introduced 
+in this paper is a new generative model for labelled networks,
+developed in order to address difficulties of other graphical models 
+when testing how vertex features affect community structure. The idea 
+is to divide the graph into its most natural partition and test whether 
+the vertex features can accurately explain this partition. 
+It is relatively easy to find vertex features that are in some way 
+correlated with the graphical structure. Nonetheless, only when 
+we find the feature that best describes the most pronounced partition 
+do we have a stronger case for causation.
 
-With the newly-defined FFBM, we go on to present an efficient inference algorithm to sample the parameters $\theta$ of the feature-to-block generator.
-This is introduced as two concurrent Markov chains to sample the block memberships $b$ and block generator parameters $\theta$.
-Nevertheless, we can serialise the chains and use the empirical mean of the $b$-samples as the input to our $\theta$-chain.
-This reduces the variance in our evaluation of the target distribution and thus shortens burn-in.
+Using this new model,
+we go on to describe an efficient inference algorithm to sample 
+the parameters $\theta$ of the feature-to-block generator. 
+This takes the form of two concurrent Markov chains,
+used to sample the block memberships $b$ and block generator 
+parameters $\theta$. These chains can either be executed
+in parallel or 
+the empirical mean of the $b$-samples can be used
+as the input to the $\theta$-chain. The latter option
+reduces the variance in our evaluation of the target distribution 
+and thus shortens burn-in.
 
-The overall method is shown to be effective at extracting and describing the most natural communities in a labelled network.
-Nevertheless, the approach can only currently explain the structure at the macro-scale.
-We cannot explain structure within each block.
-Future work will benefit from extending the FFBM to be hierarchical in nature.
-That way, the structure of the network can be explained at all length-scales of interest.
-So long as data collection techniques remain ethical and care is taken to respect personal privacy, such empowered decision-making can only help humankind.
+The overall method is shown to be effective at extracting and describing 
+the most natural communities in a labelled network. Nevertheless, the approach 
+can only currently explain the structure at the macro-scale. We cannot 
+explain structure within each block. Future work will benefit from extending 
+the FFBM to be hierarchical in nature. That way, the structure of the network 
+can be explained at all several scales of interest. So long as data 
+collection techniques remain ethical and care is taken to respect 
+personal privacy, such empowered decision-making can only help humankind.

### 2021-05-27 nips/experiments.tex
Yiannis version

@@
-\section{Experiments}
+\section{Experimental results}
 \label{sec:experiments}
@@
 Two books are connected if they were frequently co-purchased by customers.
-Vertex features encode the political affiliation of the author (liberal, conservative or neutral).
+Vertex features encode the political affiliation of the author (liberal, conservative, or neutral).
 
@@
 	\caption{Networks laid out and coloured according by inferred block memberships $\hat{y}$ for a given experiment iteration.
-Visualisation performed using \textit{graph-tool} \cite{peixoto_graph-tool_2014}}
+Visualisation performed using \textit{graph-tool} \cite{peixoto_graph-tool_2014}.}
 	\label{fig:graphs-all}
@@
 	\end{subfigure}
-	\caption{Reduced dimension feature-to-block generator weight samples}
+	\caption{Reduced dimension feature-to-block generator weight samples.}
 	\label{fig:null-all}
@@
 	\centering
-	\caption{Experimental results averaged over $n=10$ iterations (mean $\pm$ standard deviation)}
+	\caption{Experimental results averaged over $n=10$ iterations (mean $\pm$ standard deviation).}
 	\label{tab:results}
@@
 %
-We require metrics to assess performance.
-This can be split into two separate components: the microcanonical SBM fit (concerned with the $b$-samples) and the fit of the feature-to-block generator (concerned with the $\theta$-samples).
-Starting with the SBM, $S(b)$ (equation \ref{eqn:dl-form}) can be interpreted as the description length of the partition imposed by $b$.
-It is only natural to divide this quantity by the number of entities (nodes and edges) in our graph $N+E$ to allow for rough comparison between graphs.
-This defines a simple metric to gauge the fit of the SBM: the description length per entity averaged over the $b$-samples (equation \ref{eqn:mean-dl}):
+We require metrics to assess performance. This can be split into two separate 
+components: the microcanonical SBM fit (concerned with the $b$-samples) and 
+the fit of the feature-to-block generator (concerned with the $\theta$-samples).
+
+Starting with the SBM, recall that the quantity $S(b)$ in~(\ref{eqn:dl-form}) 
+can be interpreted as an ideal ``description length'' of the partition 
+imposed by $b$. We define a simple
+metric $\bar{S}_e$ to gauge the fit of the SBM,
+as the description length per entity 
+(i.e., divided by the total number $N+E$ of 
+nodes plus edges),
+averaged over the $b$-samples:
 %
 \begin{equation}
-	\bar{S}_e \coloneqq \frac{1}{(N+E) |\Tcal_b|} \sum_{t\in \Tcal_b} S \left( b^{(t)} \right)
+	\bar{S}_e \coloneqq \frac{1}{(N+E) |\Tcal_b|} \sum_{t\in \Tcal_b} S \left( b^{(t)} \right).
 	\label{eqn:mean-dl}
@@
 %
-However, to assess the performance of the feature-to-block predictor, we must partition the vertex set $[N]$ into training and test sets.
-We choose to randomly partition the vertices on each experiment run such that a constant fraction $f$ of the available vertices go to form our training set $\Gcal_0$ and the remainder are held out to form our test set $\Gcal_1$.
+
+Next, in order to assess the performance of the feature-to-block predictor, 
+we partition the vertex set $[N]$ into training and test sets. We choose to 
+randomly partition the vertices on each experiment run,
+so that a constant fraction $f$ of the available vertices go to form 
+the training set $\Gcal_0$ and the remainder are held out to form the
+test set $\Gcal_1$.
 The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain.
-As $|\Gcal_0| \neq |\Gcal_1|$ in general, we cannot use the un-normalised log target $U$ (equation \ref{eqn:U-form}) for comparison as the total cross-entropy loss is scaled by the size of each set but the prior term stays constant.
-We therefore must use the average cross-entropy loss over each set (equation \ref{eqn:cross-entropy-loss}):
+Because $|\Gcal_0| \neq |\Gcal_1|$ in general, we cannot use the un-normalised log-target $U$ from~(\ref{eqn:U-form}) for comparison,
+as the total cross-entropy loss scales with the size of each data set but 
+the prior term stays constant. We therefore use the average cross-entropy loss 
+over each set,
 %
 \begin{equation}
-	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star \left( \theta^{(t)} \right)
-	\qquad \textrm{where} \qquad
-	\Lcal_\star \left( \theta^{(t)} \right) \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)}
+	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star \left( \theta^{(t)} \right),
+	\quad \textrm{where} \quad
+	\Lcal_\star \left( \theta^{(t)} \right) \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)},
 	\label{eqn:cross-entropy-loss}
@@
 %
-Where $\star \in \{0, 1\}$ has been introduced to toggle between training and test sets.
-Table \ref{tab:results} summarises the results for each experiment.\footnote{For a comprehensive list of the hyper-parameters used for each experiment please see appendix \ref{appdx:hyperparams}} We also apply the dimensionality reduction method on the two higher dimensional datasets (the school and FB egonet).
-For this we leverage equation \ref{eqn:c-star}, to reduce the dimension from $D$ to $D'$ with $k=1$ to yield the maximal cutoff $c^*$.
-We then retrain the feature-block predictor using just the retained feature set $\Dcal'$ and report the loss over the training and test sets for the reduced classifier -- denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively.
-These values are also included in table \ref{tab:results}.
+and where $\star \in \{0, 1\}$ indicates whehter the training or test
+set is being considered.
 
-Table \ref{tab:results} already highlights some general trends in the results.
-Firstly, the variance of the test loss $\bar{\Lcal}_1$ tends to be higher than the training loss $\bar{\Lcal}_0$.
-This is expected as our test set is smaller than the training set and so more susceptible to variability in its construction.
-Indeed, most of the variance in the evaluation of $\bar{\Lcal}_0$ and $\bar{\Lcal}_1$ comes from the random partitioning of the graph into training and test sets.
-Secondly, it can be seen that the dimensionality reduction procedure brings the training and test losses closer together.
-This implies that the features we keep are indeed correlated with the underlying graphical partition and that the approach generalises correctly.
+Table~\ref{tab:results} summarises the results for each experiment;
+a complete list of all the hyperparameters used 
+is given in Appendix \ref{appdx:hyperparams}. We also apply the 
+dimensionality reduction method 
+of Section~\ref{sec:dim-reduction}
+to the two higher dimensional datasets (the school and FB egonet). 
+For this we use equation~(\ref{eqn:c-star}) with $k=1$,
+in order to reduce the dimension from 
+$D$ to a desired $D'$ with $k=1$ based on the maximal cutoff $c^*$. 
+We then retrain the feature-block predictor using only the retained 
+feature set $\Dcal'$, and report the log-loss over the training and 
+test sets for the reduced classifier -- 
+denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively. 
+These values are also shown in Table~\ref{tab:results}.
 
-The average description length per entity of the graph $\bar{S}_e$ has very low variance implying the detected communities can be found reliably (to within an arbitrary relabelling of blocks).
-For reference we plot an inferred partition for each of the graphs on figure \ref{fig:graphs-all}.
+Based on these results, some remarks are in order.
+Firstly, the variance of the test loss $\bar{\Lcal}_1$ tends to be higher 
+than the training loss $\bar{\Lcal}_0$. This is expected,
+as the test set is smaller than the training set and hence 
+more susceptible to variability in its construction. Indeed, 
+most of the variance in the evaluation of $\bar{\Lcal}_0$ and $\bar{\Lcal}_1$ 
+comes from the random partitioning of the graph into training and test sets. 
+Secondly, it can be seen that the dimensionality reduction procedure 
+brings the training and test losses closer together. This implies that 
+the features we keep are indeed correlated with the underlying graphical 
+partition and that the approach generalises correctly.
+
+The average description length per entity,
+$\bar{S}_e$, of the graph, 
+has very small variance, suggesting that
+the detected communities can be found reliably (to within an arbitrary 
+relabelling of blocks). For reference, we plot an inferred partition for each 
+of the graphs in Figure~\ref{fig:graphs-all}.
 The polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do succeed at partitioning the graph into densely connected clusters.
@@
 
-We wish to determine whether the author's political affiliation is a good predictor of the overall network structure.
-We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral).
-From, figure \ref{fig:polbooks-null}, we see that all 3 blocks have a distinct political affiliation as their largest positive component.
-This is strong evidence that political affiliation is indeed the axis which best predicts the 3-way natural partition of the graph into blocks.
-Furthermore, from table \ref{tab:results} we see that the training and test losses are very similar and both are low in magnitude.
-This provides further evidence to the claim that political affiliation is the best explanatory variable for the overall network structure.
+The goal here is to 
+determine whether the authors' political affiliations are a good predictor 
+of the overall network structure.
+We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal.
+or neutral).
+From Figure~\ref{fig:polbooks-null} we see that all 3 blocks have a distinct political affiliation as their largest positive component.
+This strongly indicates that political affiliation is indeed the axis 
+which best predicts the 3-way natural partition of the graph into blocks. 
+Furthermore, in Table~\ref{tab:results} we see that the training and test losses 
+are very similar and both are low in magnitude. This provides further evidence 
+to the claim that political affiliation is a very appropriate explanatory 
+variable for the overall network structure.
 
@@
 
-We choose $B=10$ in line with the total number of school-classes.
-As before, we sample the block-generator parameters $\theta$ and employ the dimensionality reduction technique with standard deviation multiplier $k=1$ to pick out the top $D'=10$ features.
-We then plot the weights for the surviving features $d \in \Dcal'$ on figure \ref{fig:school-null}.
-Immediately, we see that only the pupils' class memberships have survived (1A-5B); gender and teacher/student status have been discarded meaning that these are not good predictors of overall macro-structure.
+We choose the number of communities $B=10$, in line with the total number of 
+school classes. As before, we sample the block-generator parameters $\theta$ and 
+employ the dimensionality reduction technique 
+of Section~\ref{sec:dim-reduction},
+with standard deviation multiplier $k=1$ to pick out the top $D'=10$ features. 
+We then plot the weights for the resulting features $d \in \Dcal'$ 
+in Figure~\ref{fig:school-null}. It is rather obvious
+that only the pupils' class memberships (1A-5B) have remained
+significant;
+gender and teacher/student status have been discarded,
+meaning that these are not good predictors of overall macro-structure.
 
 The vast majority of blocks are composed of a single class. 
-However, some blocks have 2 comparably good classes as their predictor.
+However, some blocks have two comparably good classes as their predictor. 
 For example, block 2 contains classes 3A and 3B as its 2 best predictors. 
-This suggests that the social divide between classes is less pronounced for pupils in year 3.
-Conversely, some classes are found to extend over two detected blocks (class 2B spans blocks 6 and 7) but we nonetheless do not have a feature which explains the division.
-The most surprising block is number 5 - which has comparable weightings for classes 5A and 1B.
-Perhaps there was a joint event between those two classes on the day the data were collected.
+This suggests that the social divide between classes is less pronounced 
+for pupils in year 3. Conversely, some classes are found to extend over two 
+detected blocks (class 2B spans blocks 6 and 7) but we nonetheless do 
+not have a feature which explains the division. The most surprising block 
+is number 5, which has comparable weightings for classes 5A and 1B. 
+Perhaps there was a joint event between those two classes on the day 
+the data were collected.
 
@@
 
-We choose $B=10$ and $D'=10$ for this experiment.
-The remaining features (figure \ref{fig:fb-null}) are those that best explain the high-level community structure.
-The majority of the surviving features are education related.
-Nevertheless, for $D'=10$ we only have good explanations for the makeup of some of the detected blocks; several blocks in figure \ref{fig:fb-null} do not have high-magnitude components for $D'=10$.
+We choose $B=10$ and $D'=10$ for this experiment. The selected features 
+(Figure~\ref{fig:fb-null}) are those that best explain the high-level 
+community structure. The majority of them are education related. 
+Nevertheless, for $D'=10$ we only have good explanations for the makeup 
+of some of the detected blocks; several blocks in 
+Figure~\ref{fig:fb-null} do not have high-magnitude components for $D'=10$.
 
-When the feature dimension is very large, it becomes increasingly likely that a particular feature may uniquely identify a small set of nodes.
-If these nodes are all part of the same community then the classifier will overfit for that particular parameter.
-The regularisation term imposed by the prior goes some way to alleviating this problem.
-Nevertheless, we see in figure \ref{fig:fb-null} that the feature \verb*|birthday-5| has a very high weight as it relates to block 1 -- but it would be preposterous to conclude that birthdays determine graphical structure.
-The analyst must remain vigilant of such problems.
+When the feature dimension is very large, it becomes increasingly likely that 
+a particular feature may uniquely identify a small set of nodes. If these nodes 
+are all part of the same community, then the classifier will overfit for that 
+particular parameter. The regularisation term imposed by the prior goes some 
+way towards alleviating this problem. Nevertheless, we see in 
+Figure~\ref{fig:fb-null} that the feature \verb*|birthday-5| has a very 
+high weight as it relates to block 1 -- but it is highly unlikely
+that birthdays determine graphical structure. 
+% The analyst must remain vigilant of such problems.

### 2021-05-27 nips/ffbm.tex
Yiannis version

@@
 
-In this section we propose a novel generative model for labelled networks.
-We call this the feature-first block model (FFBM) and outline its structure in \ref{fig:ffbm} As before, we let $N$ denote the number of nodes and $B$ the number of blocks in our graph.
-We define the vector $x_i \in \Xcal^D$ as the feature vector for the $i$'th vertex.
-$D$ is the number of features.
-For the datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$.
-The feature vectors $\{x_i\}_{i=1}^{N}$ may be compactly subsumed into the feature matrix $X \in \Xcal^{N \times D}$.
+In this section we propose a novel generative model for labelled networks. We call this the feature-first block model (FFBM).
 
-For the FFBM, we start with the feature matrix X and probabilistically generate a vector of block memberships $b \in [B]^N$.
-The parameters of this step are encapsulated by $\theta$.
-Each feature vector $x_i$ is treated independently and used to generate the corresponding block membership $b_i \in [B]$.
-We choose a single softmax layer to model $p(b_i | x_i, \theta)$.
-More complex models are possible but then deriving meaning from the inferred parameter distributions is more difficult.
-Summarising, we write $p(b | X, \theta)$ as follows:
+Let $N$ denote the number of nodes, $B$ the number of blocks in the graph,
+and ${\cal X}$ the set of all possible features.
+We define the vector $x_i \in \Xcal^D$ as the feature vector of vertex $i$, 
+where $D$ is the number of features associated with each vertex.
+For the datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$. We write $X$ for the $N\times D$ {\em feature matrix} containing
+the feature vectors $\{x_i\}_{i=1}^{N}$ 
+as its rows.
+
+For the FFBM, we start with the feature matrix $X$ and generate a random
+vector of block memberships $b \in [B]^N$. For each node $i$, the
+block membership $b_i\in[B]$ is generated based on the feature
+vector $x_i$, independently between nodes. The conditional
+distribution of $b_i$ given $x_i$ also depends on a collection
+of weight vectors $\theta=\{w_k\}_{k=1}^B$, where each
+$w_k$ has dimension $D+1$. Specifically, 
+the distribution of $b$ given $X$ and $\theta$ is,
 %
@@
 %
-We deliberately exclude a bias term to ensure that the relationships we model are based on features and not information about the size of each detected block; a more complete discussion on this topic is given in \ref{appdx:dimension}.
-The parameter vector $\theta$ for this stage contains all the weight vectors $\theta = \{w_k\}_{k=1}^{B}$.
-Each $w_k$ has dimension $D$.
-We could instead write the parameters $\theta$ as a $B \times D$ matrix of weights $W$; this form has computational benefits as then $z_i \coloneqq W x_i$, which is the input to the softmax activation function.
+We deliberately exclude a bias term to ensure that the relationships we model are based on features and not information about the size of each detected block;
+a more complete discussion on this topic is given in \ref{appdx:dimension}. 
+We will later also find it convenient
+to write the parameters $\theta$ as a $B \times (D+1)$ matrix of weights $W$; this form has computational benefits as then $z_i \coloneqq W x_i$, which is the input to the softmax activation function.
+
+More complex models based on different choices for the distributions
+$\phi_{b_i}$ above are also possible, but then deriving meaning from 
+the inferred parameter distributions is more difficult.
 
@@
 
-Once the block memberships $b$ have been generated, we then draw the graph $A$ from the microcanonical DC-SBM (equation \ref{eqn:A-generation}) with additional parameters encapsulated by $\psi = \{\psi_e, \psi_k\}$.
+Once the block memberships $b$ have been generated, we then draw the 
+graph $A$ from the microcanonical DC-SBM with additional parameters 
+$\psi = \{\psi_e, \psi_k\}$:
 %
 \begin{equation}
-	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k)
+	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k).
 	\label{eqn:A-generation}
@@
 
-Before performing any inference, we must specify priors on $\theta$ and $\psi$.
-For $\theta$ it seems sensible to choose a Gaussian prior, with zero mean and variance matrix $\sigma^2_\theta I$ such that each element of $\theta$ is independent and distributed like $\sim \Gaussian(0, \sigma_\theta^2)$.
-In vector form, the prior for $\theta$ is therefore:
+To complete the description of our Bayesian framework,
+priors on $\theta$ and $\psi$ must also be specified. 
+We place a Gaussian prior on $\theta$ with zero mean
+and covariance matrix $\sigma^2_\theta I$, so that
+each element of $\theta$ has an independent ${\cal N}(0,\sigma_\theta^2)$
+prior, for a hyperparameter $\sigma_\theta^2$:
 %
 \begin{equation}
-	p(\theta) = \Gaussian \left( \theta ; 0, \sigma_\theta^2 I \right)
+	p(\theta) \sim \Gaussian \left( \theta ; 0, \sigma_\theta^2 I \right).
 	\label{eqn:theta-prior}
@@
 %
-In our model, the block memberships vector $b$ is an intermediate latent variable and so we are not free to choose a prior for it.
-Nevertheless, as far as inference on the right-hand-side of figure \ref{fig:ffbm}, we regard $p(b | X)$ as a pseudo-prior on $b$.
-We can show (appendix \ref{appdx:b|x}) that our choice of prior for $p(\theta)$ in equation \ref{eqn:theta-prior} leads to a uniform $p(b | X)$ in equation \ref{eqn:b-pseudo-prior}.
+Interestingly, this choice of prior leads to a particularly
+simple distribution on the block membership vector $b$.
+We show in Appendix~\ref{appdx:b|x} that,
+after integrating out $\theta$,
+$b$ is uniformly distributed:
 %
 \begin{equation}
-	p(b | X) = \int p(b | X, \theta) p(\theta) d\theta = B^{-N}
+	p(b | X) = \int p(b | X, \theta) p(\theta) d\theta = B^{-N}.
 	\label{eqn:b-pseudo-prior}
@@
 %
-This is an enormously important simplification as evaluating $p(b | X)$ does not require an expensive Monte-Carlo integration over the $\theta$-domain nor does it require the exact value of $X$.
-\citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for the additional microcanonical SBM parameters $\psi$ which we adopt.
-Peixoto's idea is to write the joint prior on $(b, e, k)$ as a product of conditionals $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$.
-For our purposes we must insert a conditioning on $X$, to form our pseudo-prior for $b$ and $\psi$, to give equation \ref{eqn:joint-pseudo-prior}.
+This is an important simplification as evaluating $p(b | X)$ does not require an expensive Monte Carlo integration over $\theta$ nor does it require the exact value of $X$. 
+\citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for 
+the additional microcanonical SBM parameters $\psi$, which we adopt. 
+The idea is to write the joint distribution on $(b, e, k)$ as a product of 
+conditionals, $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$. 
+In our case, conditioning on $X$ is also necessary, leading to,
 %
 \begin{equation}
-	p(b, \psi | X) = p(b | X) p(\psi | b, X) = p(b | X) p(\psi | b)
+	p(b, \psi | X) = p(b | X) p(\psi | b, X) = p(b | X) p(\psi | b),
 	\label{eqn:joint-pseudo-prior}
@@
 %
-Where we leverage the fact $(\psi \indep X) | b$.
-We then borrow the priors proposed by \citet{Peixoto-Bayesian-Microcanonical} for $p(\psi | b)$ to complete our model.
-Please refer to appendix \ref{appdx:sbm} for the exact form of $p(\psi | b)$.
-All that concerns the main argument is we have a computable form.
+where we used the fact $\psi$ and $X$ are conditionally 
+independent given $b$.  The exact form of the prior
+$p(\psi|b)$ is described in Appendix \ref{appdx:prior}.
+All that concerns the main argument is that it has
+an easily computable form.

### 2021-05-27 nips/inference.tex
Yiannis version

@@
 
-Now that we have defined the FFBM, we wish to leverage it to perform inference.
-Suppose we are presented with a vertex-labelled graph $(A, X)$; the goal is to draw samples for $\theta$ according to the posterior given the observed data:
+Having completed the definition of the FFBM, we wish to leverage it 
+to perform inference. Given a vertex-labelled graph $(A, X)$,
+the first step is to sample from the posterior distribution
+of $\theta$:
 %
@@
 	\label{eqn:theta-target}
-	\theta^{(t)} \sim p(\theta | A, X)
+	\theta \sim p(\theta | A, X).
 \end{equation}
 %
-However, generating these samples is not easily done in practice.
-We therefore propose an iterative approach.
-We first draw samples $b^{(t)}$ from the block membership posterior (equation \ref{eqn:b-samples}) and then use each $b^{(t)}$ to draw samples for $\theta$ as in equation \ref{eqn:theta-samples}.
+We propose an iterative approach to obtain samples
+$\theta^{(t)}$ from $p(\theta|A,X)$. We first draw samples $b^{(t)}$ 
+from the block membership posterior distribution,
+and then use each $b^{(t)}$ to obtain a corresponding
+sample $\theta^{(t)}$:
 %
 \begin{align}
-	b^{(t)} &\sim p \Big( b | A, X \Big)  \label{eqn:b-samples}\\
-	\theta^{(t)} &\sim p\Big(\theta | X, b^{(t)} \Big) \label{eqn:theta-samples}
+	b^{(t)} &\sim p ( b | A, X )  \label{eqn:b-samples},\\
+	\theta^{(t)} &\sim p(\theta | X, b^{(t)} ). \label{eqn:theta-samples}
 \end{align}
 %
-Both of these sampling steps can be implemented with a Markov Chain through the Metropolis-Hastings algorithm \cite{hastings-alg}.
-We just need to define a proposal distribution $q(x, x')$ for proposing a move $x \rightarrow x'$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(\cdot)$, point-wise.
+Both of these sampling steps can be implemented using
+Markov chain Monte Carlo via a Metropolis-Hastings 
+algorithm \cite{hastings-alg}, as a two-level Markov chain.
+
+% We just need to define a proposal distribution $q(x, x')$ for proposing a move $x \rightarrow x'$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(\cdot)$, point-wise.
 The proposed move is then accepted with probability $\alpha$ (equation \ref{eqn:mh-accept}) else it is rejected and we stay at $x$.
 %
-\begin{equation}
-	\alpha(x, x') = \min \left( \frac{\pi(x') q(x', x)}{\pi(x) q(x, x')} , 1 \right)
-	\label{eqn:mh-accept}
-\end{equation}
+% \begin{equation}
+	% \alpha(x, x') = \min \left( \frac{\pi(x') q(x', x)}{\pi(x) q(x, x')} , 1 \right)
+	% \label{eqn:mh-accept}
+% \end{equation}
 %
-This accept-reject step ensures the resulting Markov Chain is in detailed balance with the target distribution $\pi(\cdot)$.
-What we propose in equations \ref{eqn:b-samples} and \ref{eqn:theta-samples} is therefore implemented through a 2-level Markov chain.
-The resulting samples for $\theta^{(t)}$ are unbiased in the sense that the expectation of their distribution is the posterior we are targetting:
+% This accept-reject step ensures the resulting Markov Chain is in detailed balance with the target distribution $\pi(\cdot)$. 
+% What we propose in equations \ref{eqn:b-samples} and \ref{eqn:theta-samples} is therefore implemented through a 2-level Markov chain. 
+As we show in the following section, the resulting samples 
+for $\theta^{(t)}$ are unbiased in the sense that the expectation of 
+their distribution is the target posterior distribution:
 %
 \begin{equation}
-\Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] = \sum_{b \in [B]^N} p(\theta | X, b) p(b | A, X) = \sum_{b \in [B]^N} p(\theta, b | A, X) = p(\theta | A, X)
+\Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] = \sum_{b \in [B]^N} p(\theta | X, b) p(b | A, X) = \sum_{b \in [B]^N} p(\theta, b | A, X) = p(\theta | A, X).
 \label{eqn:theta-unbiased}
@@
 %
-This is an example of a pseudo-marginal approach.
-Indeed, \citet{pseudo-marginal} show that the unbiased result in equation \ref{eqn:theta-unbiased} is sufficient to prove that for large enough $t$, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{(t)})\right] = p(\theta
-| A, X)$ which is exactly the distribution we are targetting (equation \ref{eqn:theta-target}).
+This is an example of a pseudo-marginal approach. Indeed, \citet{pseudo-marginal} show that~(\ref{eqn:theta-unbiased})
+% the unbiased result in equation 
+is sufficient to prove that, for large enough $t$, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{(t)})\right] = p(\theta 
+| A, X)$, as required.
+%  which is exactly the distribution we are targetting (equation \ref{eqn:theta-target}).
 %
@@
 	\begin{tikzpicture}[
-		scale=0.7, every node/.style={transform shape},
+		scale=0.95, every node/.style={transform shape},
 		roundnode/.style={circle, draw=black, minimum size=12mm},
@@
 	\end{tikzpicture}
-	\caption{Sampling sequence}
+	\caption{Sampling sequence.}
 	\label{fig:samp-sequence}
@@
 
-The reason we split the Markov chain into two stages is because the summation over all latent states $b \in [B]^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$ is intractable -- $O(B^N)$.
+The reason for splitting the Markov chain into two levels is because the summation over all latent states $b \in [B]^N$ required to directly compute the likelihood, $p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$,
+is intractable, as the sum involves $B^N$ terms. 
 Figure \ref{fig:samp-sequence} shows an overview of the proposed method. 
-We have introduced subscripts and conditionings to make explicit what variables each step utilises.
-We note the power of the simplification given by equation \ref{eqn:b-pseudo-prior}.
-As $p(b| X)$ does not depend on the exact value of X, we do not need to know the value of $X$ to perform the sampling on $b$.
-Conversely, for the $\theta^{(t)}$ samples, we use only $b^{(t)}$ but not $A$ as $(\theta \indep A) | b$.
+The Metropolis-Hastings accept/reject probabilities 
+for the proposed $b$ and $\theta$ samples are denoted $\alpha_b$ and
+$\alpha_\theta$, respectively.
+% We have introduced subscripts and conditionings to make explicit what variables each step utilises.
+
+Note the importance of the simplification in~(\ref{eqn:b-pseudo-prior}). 
+As $p(b| X)$ in fact does not depend on $X$, 
+knowing $X$ is not necessary in order to obtain samples
+of the block membership vector $b$.
+And on the other level, in order to obtain 
+samples of the parameter vector $\theta$
+we use only $b$ but not $A$, as $\theta$ and $A$ are
+conditionally independnet given $b$.
 
@@
 
-\citet{Peixoto-MCMC} proposes a Monte Carlo method which we will base our approach on. It relies on writing the posterior in the following form:
+We adopt the Markov chain Monte Carlo procedure of
+\cite{Peixoto-MCMC},
+which relies on writing the posterior in the following form:
 %
 \begin{equation}
-	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) = \pi_b(b)
+	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) = \pi_b(b).
 \end{equation}
 %
-Now $\pi_b(\cdot)$ is the un-normalised density we wish to sample from for the $b$-chain.
-In other words, we wish to construct a Markov chain that has $\pi_b(\cdot)$ as its invariant distribution.
-We can break $\pi_b$ down as follows:
+Now $\pi_b(\cdot)$ is the un-normalised target density,
+which can be expressed as:
+%  we wish to sample from for the $b$-chain. In other words, we 
+% wish to construct a Markov chain that has $\pi_b(\cdot)$ 
+% as its invariant distribution. We can break $\pi_b$ down as follows:
 %
@@
 	= p(b|X) p(A, \psi^* | b, X) \\
-	= p(A | b, \psi^*) \cdot p(\psi^* | b) \cdot p(b | X)
+	= p(A | b, \psi^*) \cdot p(\psi^* | b) \cdot p(b | X).
 \end{equation}
 %
-Since we are using the microcanonical SBM formulation, there is only one value of $\psi$ that is compatible with the given $(A, b)$ pair (given in equation \ref{eqn:sbm-constraints}).
-We denote this value $\psi^* = \{\psi_k^*, \psi_e^*\}$.
-Therefore, the summation over all $\psi$ reduces to just the single $\psi^*$ term; this is the power of the microcanonical formulation.
-We also define the microcanonical entropy of the configuration as.
+Since we are using the microcanonical SBM formulation, there is only one 
+value of $\psi$ that is compatible with the given $(A, b)$ pair;
+recall the constraints in~(\ref{eqn:sbm-constraints}). 
+We denote this value $\psi^* = \{\psi_k^*, \psi_e^*\}$. Therefore, 
+the summation over all $\psi$ reduces to just the single $\psi^*$ term.
+% ; this is the power of the microcanonical formulation. 
+We also define the microcanonical entropy of the configuration as:
 %
 \begin{equation}
-	S(b) = - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big)
+	S(b) = - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big).
 	\label{eqn:dl-form}
@@
 %
-This entropy can equally be thought of as the description length of the graph.
-The exact form of the proposal $q_b$ is explored thoroughly by \citet{Peixoto-MCMC} and not repeated here.
-There is a widely used library for Python made available under LGPL called \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, which implements this algorithm.
-The only modification we make is in the block membership prior $p(b)$ which we replace with $p(b|X)=B^{-N}$, which cancels out in the MH accept-reject step as it is independent of $b$.
+This entropy can be thought of as the optimal
+``description length'' of the graph. 
+The exact form of the proposal $q_b$ is explored thoroughly in
+\cite{Peixoto-MCMC} and not repeated here. There is a widely used 
+library for Python made available under LGPL 
+called \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, 
+which implements this algorithm. The only modification we make is in 
+the block membership prior $p(b)$ that we replace with $p(b|X)=B^{-N}$, 
+which cancels out in the MH accept-reject step as it is independent of $b$.
 
 \subsection{Sampling feature-to-block generator parameters}
+\label{s:sfb}
 
-The invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. We write this as follows:
+The invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. 
+We write this as,
 %
 \begin{equation}
-	\pi_\theta(\theta) \propto p(\theta | X, b) \propto p(b | X, \theta) p(\theta) \propto \pi_\theta (\theta) \propto  \exp \left( - U(\theta) \right)
+	\pi_\theta(\theta) \propto p(\theta | X, b) \propto p(b | X, \theta) p(\theta) \propto \pi_\theta (\theta) \propto  \exp \left( - U(\theta) \right),
+	\label{eq:U}
 \end{equation}
 %
-Where we have introduced $U(\theta)$ equal to the negative log posterior.
-We define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
-Discarding constant terms, we can write $U(\theta)$ as in equation \ref{eqn:U-form} (refer to appendix \ref{appdx:form-U} for the derivation).
+where we write $U(\theta)$ for the negative log-posterior. Let $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$. 
+Discarding constant terms, we can write $U(\theta)$ as,
 %
@@
 	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2 = N \cdot \Lcal(\theta) + \frac{1}{2\sigma_\theta^2} ||\theta||^2
+	+ \frac{1}{2\sigma_\theta^2} \|\theta\|^2 = N \cdot \Lcal(\theta) + \frac{1}{2\sigma_\theta^2} \|\theta\|^2;
 	\label{eqn:U-form}
 \end{equation}
+see Appendix \ref{appdx:form-U} for the derivation.
 %
-$U(\theta)$ in equation \ref{eqn:U-form} appears a typical objective function for neural network training.
+The function $U(\theta)$ is a typical objective function for neural network training.
 The first term is introduced by the likelihood.
 We collect it into $N \cdot \Lcal(\theta)$, which is the cross-entropy between the graph-predicted and feature-predicted block memberships summed over all vertices.
-The second term of equation \ref{eqn:U-form} -- introduced by the prior -- brings a form of regularisation, guarding against over-fitting.
-Different to traditional applications, our goal is not to find the minimiser of $U(\theta)$ but to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
+The second, introduced by the prior, brings a form of regularisation, guarding against over-fitting. Unlike in many applications where the goal 
+is to find the minimiser of $U(\theta)$, our goal is to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
 We can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
-We therefore adopt the Metropolis-adjusted Langevin algorithm (MALA) -- first proposed by \citet{mala-tweedie}.
-Given the current sample $\theta$, we generate a new sample $\theta'$ according to equation \ref{eqn:theta-update}.
+We therefore adopt the Metropolis-adjusted Langevin algorithm (MALA) \cite{mala-tweedie}.
+Given the current sample $\theta$, we generate
+a new sample $\theta'$ as,
 %
-\begin{align}
-	\theta' &= \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi
-	\label{eqn:theta-update} \\
-	\therefore q_\theta(\theta, \theta') &= \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right)
-	\label{eqn:theta-proposal}
-\end{align}
-%
-Where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter -- which may vary with the sample index (appendix \ref{appdx:step-size} explores this more fully).
-Without the injected noise term, MALA is equivalent to gradient descent.
-We require the noise term $\xi$ to fully explore the parameter space.
-We can write the proposal distribution $q_\theta$ as in equation \ref{eqn:theta-proposal}.
-The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}).
-By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as:
+$$	\theta' = \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi,$$
+where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter 
+which may vary with the sample index (see Appendix \ref{appdx:step-size}).
+This leads to the proposal distribution,
+$$q_\theta(\theta, \theta') \sim \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right).$$
+Without the injected noise term, MALA is equivalent to gradient descent. We require the noise term $\xi$ to fully explore the parameter space. 
+The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}). By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as:
 %
 \begin{equation}
-	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
+	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right).
 	\label{eqn:U-derivative}
@@
 %
-After a proposed move is generated, in typical Metropolis-Hastings fashion we accept the move with probability $\alpha_\theta$, as in equation \ref{eqn:mh-accept}.
+After a proposed move is generated from $q_\theta$,
+it is either accepted or rejected in the standard
+Metropolis-Hastings accept/reject step.
 
 \subsection{Sampling sequence}
+\label{s:ss}
 
@@
 	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one(b_{i}^{(t)} = j) \right]
-	= p(b_i = j | A, X)
+	= p(b_i = j | A, X).
 \end{equation}
 %
-We can obtain an unbiased estimate for this quantity using the set of $b$-samples.
-However, as with all MCMC methods, we must only uses samples after burn-in and thinning have been applied.
-We introduce $\Tcal_b$ to denote the retained set of indices for the $b$-samples and $\Tcal_\theta$ similarly for the $\theta$-chain.
-An in-depth discussion of how these sets are chosen is given in appendix \ref{appdx:burn-in-thinning}.
-The unbiased estimate for $y_{ij}^{(t)}$ using the restricted sample set $\Tcal_b$ is denoted $\hat{y}_{ij}$ and has form:
+We can obtain an unbiased estimate for this quantity using 
+the thinned $b$-samples after burn-in; the length of the
+burn-in and the details of the thinning are described
+in Appendix \ref{appdx:burn-in-thinning}.
+Let $\Tcal_b$  denote the retained set of indices 
+for the $b$-samples and $\Tcal_\theta$ similarly for the $\theta$-chain. 
+The unbiased estimate for $y_{ij}^{(t)}$ using the 
+restricted sample set $\Tcal_b$ is denoted $\hat{y}_{ij}$:
 %
 \begin{equation}
-	\hat{y}_{ij} \coloneqq \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} y_{ij}^{(t)} = \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} \one\{b_i^{(t)} = j\}
+	\hat{y}_{ij} \coloneqq \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} y_{ij}^{(t)} = \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} \one\{b_i^{(t)} = j\}.
 \end{equation}
 %
-We choose to feed each $\theta^{(t)}$ update step the same matrix $\hat{y}$ for all $t$ rather than the corresponding $y^{(t)}$.
-This means we no longer need to run the $b$ and $\theta$ Markov chains concurrently.
-Instead, we run the $b$-chain to completion and use it to generate $\hat{y}$.
-This affords us the flexibility to vary the lengths of the $b$ and $\theta$-chains.
-Furthermore, the changeover from $y^{(t)}$ to $\hat{y}$ reduces the burn-in time for the $\theta$-chain by reducing the variance in our evaluation of $U$ and $\nabla U$.
-A description of the overall algorithms is given in appendix \ref{appdx:algorithms}.
+The same matrix $\hat{y}$ is used in the update step
+for each $\theta^{(t)}$.
+This way, it is not necessary to run the $b$ and $\theta$ Markov chains 
+concurrently. Instead, we run the $b$-chain to completion and use it 
+to generate $\hat{y}$. This affords us the flexibility to vary the 
+lengths of the $b$ and $\theta$-chains. Furthermore, the changeover 
+from $y^{(t)}$ to $\hat{y}$ reduces the burn-in time for 
+the $\theta$-chain by reducing the variance in the evaluation 
+of $U$ and $\nabla U$. A detailed description of the overall 
+algorithms is given in Appendix \ref{appdx:algorithms}.
 
@@
 
-Once we have the samples $\left\{ \theta^{(t)} \right\} \sim p(\theta | A, X)$, we can compute the empirical mean and standard deviation of each component of $\theta$.
+Once the samples $\left\{ \theta^{(t)} \right\} \sim p(\theta | A, X)$
+have been obtained, we can compute the empirical mean and standard deviation of each component of $\theta$.
 Switching back to matrix notation we define $\theta = W$, such that $W_{ij}$ is the weight component for block $i$ and feature $j$, we can define:
@@
 %
-A simple heuristic to discard the least important features requires specifying a cutoff $c > 0$ and a multiplier $k > 0$.
-We define the function $\Fcal_i(j)$ as in \ref{eqn:fij} then only keep features with indices $d \in \Dcal'$, where $\Dcal'$ is constructed as in equation \ref{eqn:kept-feature-set}.
+A simple heuristic to discard the least important features requires specifying a cutoff $c > 0$ and a multiplier $k > 0$. We define the function $\Fcal_i(j)$ 
+as in~(\ref{eqn:fij}) and only keep features with indices $d \in \Dcal'$, where $\Dcal'$ is given in~(\ref{eqn:kept-feature-set}).
 %
 \begin{align}
-	\Fcal_i(j) &\coloneqq (\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c)
+	\Fcal_i(j) &\coloneqq (\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c),
 	\label{eqn:fij} \\
@@
 %
-Intuitively, this means discarding any feature for which $\hat{\mu}_{ij} \pm k\hat{\sigma}_{ij}$ lies within or spans the null region $(-c, c)$ for all block indices.
-If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \mu_{ij}, \sigma_{ij}^2)$, then this is analogous to a hypothesis test on the value of $W_{ij}$ (equation \ref{eqn:hyp-test-discard}).
-$\Dcal'$ then comprises all features $i$ for which $H_1$ is accepted at least once for some $j \in [B]$.
+Intuitively, this means discarding any feature $j$ for which 
+$\hat{\mu}_{ij} \pm k\hat{\sigma}_{ij}$ lies outside the region
+$(-c, c)$ for all block indices $i$.
+If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \mu_{ij}, \sigma_{ij}^2)$, then this is analogous to a hypothesis test on the value of $W_{ij}$ as in~(\ref{eqn:hyp-test-discard});
+then $\Dcal'$ comprises all features $i$ for which $H_1$ is not rejected at least once for some $j \in [B]$
 %
@@
 \begin{equation}
-	c^* = \argmax_{c>0} (c : |\Dcal'| = D', k=k_0)
+	c^* = \argmax \{c>0\; : \;|\Dcal'| = D', k=k_0\}.
 	\label{eqn:c-star}
@@
 %
-For an algorithmic implementation of this method refer to appendix \ref{appdx:algorithms}.
+For an algorithmic implementation of this method refer to appendix \ref{appdx:algorithms}. This approach has the advantage that it fixes the number of reduced dimensions.

### 2021-05-27 nips/introduction.tex
Yiannis version

@@
 
-A somewhat surprising property of many real-world networks is that they exhibit strong community structure.
-In other words, each node will often belong to a cluster of densely connected nodes.
-There is high interest in recovering the latent communities from the observed graphs.
-The inferred communities can be exploited for compression algorithms \cite{cluster-compression} or used for link prediction in incomplete networks \cite{link-prediction} to name but a few applications.
+An important characteristic of many real-world networks is that they exhibit strong community structure, with most nodes often belonging to a densely connected cluster. 
+Finding ways to recover the latent communities from the observed graphs is an important
+task in many applications, includng
+graph/network compression \cite{cluster-compression} and link prediction in 
+incomplete networks \cite{link-prediction}.
 
-We restrict our analysis to labelled networks.
-These are graphs where we additionally have information about the properties of each vertex.
-We shall refer to these vertex properties as features.
-One of the most common questions we can ask of labelled networks is whether a given vertex feature has an impact on the structure of the graph.
-To answer this question from a Bayesian perspective we must use a random graph model; the most common form is the stochastic block model (SBM) \cite{vanilla-sbm}.
+In this work, we
+restrict our attention to vertex-labelled networks, and we
+refer to the vertex labels as {\em features}.
+A common goal is to determine whether a given feature impacts graphical structure.
+To answer this from a Bayesian perspective requires the use of a random graph model; the standard is the stochastic block model (SBM) \cite{vanilla-sbm}.
 This is a latent variable model where each vertex belongs to a single block and the probability two nodes are connected depends only on the block memberships of each.
-There have been many variants to this model; the most popular are the mixed-membership stochastic block model (MMSBM) \cite{mixed-membership-sbm} and the overlapping stochastic block model (OSBM) \cite{overlapping-sbm}.
-Effectively, these just extend the model to allow each vertex to belong to multiple blocks simultaneously.
-
-However, a major drawback of these graphical models as applied to labelled networks is that they do not automatically include vertex features in the random graph generation process.
+Numerous variants of this
+model have been considered -- the most popular ones being the mixed-membership stochastic block model (MMSBM) \cite{mixed-membership-sbm} and the overlapping stochastic block model (OSBM) \cite{overlapping-sbm}.
+Effectively, these extend the model to allow each vertex to belong to multiple blocks simultaneously.
+However, one drawback of these graphical models as applied to labelled networks is that they do not automatically include vertex features in the random graph generation process.
 Approaches based on graph neural networks \cite{sbm-gnn} that utilise vertex features have been developed but these lack the easy interpretability of the simpler models.
 
-To analyse a labelled network using one of the simple SBM variants, a typical inference procedure would be to partition the graph into blocks grouped by distinct values of the feature of interest.
-The associated model can then be used to test for evidence of heterogeneous connectivity between the feature-grouped blocks.
-Nevertheless, this approach is limited in that it can only consider one feature at a time; this leaves any conclusions drawn highly vulnerable to the presence of confounding variables.
-Furthermore, considering each feature separately makes it difficult to rank order the features by  magnitude of impact.
+To analyse a labelled network using one of the simple SBM variants, a typical inference procedure would be to first partition the graph into blocks grouped by distinct values of the feature of interest, and then use the associated model to test for evidence of heterogeneous connectivity between the feature-grouped blocks.
+Nevertheless, this approach is limited in that it can only consider one feature at a time.
+This makes it difficult to rank-order the features by  magnitude of impact.
 Lastly, the feature-grouped blocks are often an unnatural partition of the graph, leading to a poor model fit.
@@
 
-With these desiderata in mind, we present a novel framework for modelling labelled networks, which we call the feature-first block model (FFBM).
-This can be thought of as an extension of the SBM to labelled networks.
-In the FFBM, we use the features first to generate the latent block membership for each vertex.
-The latent block membership is therefore a stepping stone rather than a starting point in our model.
-We go on to present an efficient algorithm for sampling from the parameters of the feature-to-block generator.
-We can interpret the sampled FFBM parameters to determine which features have the largest impact on overall graphical structure.
-Any features found to be irrelevant can be discarded as a form of dimensionality reduction.
+Thus motivated, we present a novel framework for modelling labelled networks, which we call the feature-first block model (FFBM).
+This is an extension of the SBM to labelled networks.
+We go on to present an efficient algorithm for sampling from the parameters of the feature-to-block generator, and to describe how the
+sampled FFBM parameters can be interpreted in order to determine which features have the largest impact on overall graphical structure.

### 2021-05-27 nips/prelim.tex
10 pages
Yiannis version

--- 2021-05-23 nips/prelim.tex
@@
 
-We first need a model for community-like structure in a network.
-For this we adopt the stochastic block model (SBM) - widely used across academia.
-The premise is that each node in the graph belongs to a unique community called a block.
+We first need a model for community-like structure in a network. For this we adopt the widely-used stochastic block model (SBM).
+Each node in the graph belongs to a unique community called a block.
 The probability that two nodes are connected depends only on the block memberships of each.
-Graphs drawn from the SBM ensemble exhibit community structure.
 Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}.
-To allow for degree-variability between members of the same block, we must choose the degree-corrected formulation of the SBM (DC-SBM).
+To allow for degree-variability between members of the same block, we adopt the degree-corrected formulation (DC-SBM).
+
+For each integer $K\geq 1$, we use the notation $[K]:=\{1,2,\ldots,K\}$.
 
 \begin{definition}[Microcanonical DC-SBM]
-	\label{defn:microcan-ndc-sbm}
-	Let $N \in \Integers^{+}$ denote the number of vertices in an undirected graph.
-The block memberships are encoded by a vector $b \in [B]^N$,\footnote{We introduce the notation $[K] \coloneqq \{0, 1, 2 \dots K-1\}$ to compactly define a set of $K$ indices.
-Clearly, $[K]$ is only defined for $K \in \Integers^+$.}
-	where $B \in \Integers^{+}$ is the number of non-empty blocks.
-	Let $e \in (\Integers_0^+)^{N \times N}$ be the matrix of edge counts between blocks.
-$e_{rs}$ is then the number of edges from block $r$ onto block $s$ -- or twice that number if $r=s$.
-For undirected graphs $e$ is symmetric.
-	Let $k \in (\Integers^+_0)^N$ be a vector denoting the degree sequence of the graph. $k_i$ is then the degree of vertex $i$.
+	\label{defn:microcan-dc-sbm}
+	Let $N \geq 1$ denote the number of vertices in an undirected graph. The block memberships are encoded by a vector $b \in [B]^N$,
+where $B$ is the number of non-empty blocks.
+	Let $e=(e_{rs})$ be the $N\times N$ symmetric matrix of edge counts 
+between blocks, so that $e_{rs}$ is the number of edges from block $r$ to 
+block $s$ -- or twice that number if $r=s$. 
+	Let $k =(k_i)$ denote the vector of length $N$ containing the degree
+sequence of the graph, with $k_i$ being the degree of vertex $i$.
 
-	For the degree-corrected stochastic block model (DC-SBM), we say that the graph's adjacency matrix $A \in \{0,1\}^{N \times N}$ is generated as follows:
+The graph's adjacency matrix $A \in \{0,1\}^{N \times N}$ is generated 
+by placing edges uniformly at random, but respecting the constraints 
+imposed by $e$, $b$ and $k$ (hence the qualification `microcanonical').
+Specifically, $A$ must satisfy the following, for all $r,s\in[B]$
+and all $i\in[N]$:
 	%
 	\begin{equation}
-		A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, e, k)
+		e_{rs} = \sum_{i, j \in [N]} A_{ij} 
+	\one \{b_i = r\} \one \{b_j = s\} 
+		\qquad 
+		\textrm{and} \qquad
+		k_i = \sum_{j \in [N]} A_{ij}.
+		\label{eqn:sbm-constraints}
 	\end{equation}
-	%
-	Where edges are placed uniformly at random but respecting the constraints imposed by $e$, $b$ and $k$ -- hence why this formulation is given the microcanonical moniker.
-Specifically, $A$ must satisfy the following:
-	%
+We use the following notation to indicate this distribution:
 \begin{equation}
-		e_{rs} = \sum_{i, j} A_{ij} \one \{b_i = r\} \one \{b_j = s\} \qquad \textrm{and} \qquad
-		k_i = \sum_j A_{ij}
+		A \sim \mbox{\rm DC-SBM}_{\rm MC} (b, e, k).
 	\end{equation}
-	%
-	The additional parameters $N$ and $B$ are omitted as they can be inferred from the shapes of $b$ and $e$.
 \end{definition}

### 2021-05-27 nips/appendices.tex
More cleanup

@@
-\section{Appendix}
+\section*{Appendices}
 
+\section{Additional material}
 \subsection{SBM prior choice explanation}
@@
 
+\subsection{Choosing the MALA step-size}
+\label{appdx:step-size}
+
+Recall that in 
+Section~\ref{s:sfb} we used 
+the Metropolis Adjusted Langevin Algorithm (MALA)
+in order to 
+sample from the $\theta$-chain of the block membership 
+generator parameters.
+At iteration $t$, the proposed sample is generated by:
+%
+\begin{equation}
+	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi.
+\end{equation}
+%
+There are two competing objectives when choosing the step-size $h_t$. 
+On the one hand, $h_t$ needs to be large so that the sampler
+arrives at a high density region quickly,
+while too large a step-size would lead to low acceptance rates and thus 
+inefficient sampling. An effective strategy is
+to use {\em simulated annealing}: allow $h_t$ to slowly decrease
+with $t$, as long as $h_t>0$ for all $t$ and also:
+%
+\begin{equation}
+	\sum_{t=1}^{\infty} h_t = \infty, \qquad \textrm{and} \qquad
+	\sum_{t=1}^{\infty} h_t^2 < \infty.
+	\label{eqn:h-constraints}
+\end{equation}
+%
+Following \citet{Bayesian-SGLD}, we adopt the 
+polynomially decaying step-sizes,
+%
+$h_t = \alpha(\beta + t)^{-\gamma}$,
+%
+where $\alpha>0$, $\beta>0$ and $\gamma\in(1/2,1]$ are hyper-parameters.
+We make the specific choices,
+%
+\begin{equation}
+	\alpha = \frac{250 \cdot s}{N}, \qquad \beta = 1000, \qquad \gamma = 0.8,
+	\label{eqn:step-size-params}
+\end{equation}
+%
+where $N$ is the number of data-points and $s$,
+the {\em step-size scaling}, is the only free parameter.
+
+% For approximate methods, we can choose to bypass the MH accept-reject entirely to speed up computation.
+If this is done, the algorithm is instead called stochastic gradient Langevin diffusion (SGLD) \cite{Bayesian-SGLD}.
+This speeds up computation at the expense of exactness of the method.
+
+\subsection{Burn-in and thinning}
+\label{appdx:burn-in-thinning}
+
+% As with any MCMC method, we must deal with the 
+% issues presented by burn-in and thinning.
+
+When sampling from the $b$- and $\theta$-chains described
+in Section~\ref{sec:inference}, we generate
+$T_b$ and $T_\theta$ samples total, respectively.
+We discard an initial proportion $\kappa_\star\in(0,1)$ of the samples 
+as corresponding to a ``burn-in'' period required for the distribution 
+of the chain to reach a distribution close to our target,
+and we also ``thin'' the remaining samples to 
+obtain a less-dependent version. For $\star\in\{b,\theta\}$,
+the remaining sample sets are denoted $\Tcal_\star$
+in the notation of Section~\ref{s:ss},
+%
+\begin{equation}
+	\Tcal_\star = \{T_\star \kappa_\star + i \lambda_\star :  
+	0 \leq i \leq \lfloor T_\star(1 - \kappa_\star) / \lambda_\star \rfloor \},
+\end{equation}
+%
+where $\lambda_\star$ controls the thinning. The choice of
+$\kappa_\star$ can be determined by plotting the log-target (either $S(b^{(t)})$ 
+or $U(\theta^{(t)})$ as a function of $t$,
+and choosing $\kappa_\star$ to encompass the region where the log-target has 
+roughly reached equilibrium. As we do not leverage sample independence,
+$\lambda_\star$ can be chosen less rigorously; we often simply
+use $\lambda_b=5$ and $\lambda_\theta = 10$.
+
+\subsection{Initializing the b-chain}
+
+For the purposes of the FFBM model, the number of blocks $B$ is a constant 
+which must be specified. If the choice of $B$ is influenced 
+by the observed data, then the analysis is no longer ``fully Bayesian''
+and belongs to the class of methods referred to as ``empirical Bayes.''
+However, as the number of blocks only specifies the coarseness of the 
+analysis, it is reasonable to allow it to vary. Indeed, 
+\citet{peixoto-determine-B} shows that for a fixed 
+average degree the maximum number of detectable blocks scales 
+as $O(\sqrt{N})$ where $N$ is the number of vertices.
+
+If $B$ is allowed to vary in the $b$-chain (i.e.,
+when new blocks can be created and empty blocks are allowed),
+then the chain can be run until a minimum description 
+length (MDL) solution is reached. We take the number of non-empty blocks 
+in the MDL solution to be our fixed block number $B$ for subsequent analysis. 
+Indeed, it is prudent to start the $b$-chain at this MDL solution as then 
+the necessary burn-in time can be greatly reduced.
+
+\clearpage
+\section{Derivations}
 \subsection{Derivation of conditional block distribution given feature matrix}
@@
 
-Recall from~(\ref{eq:U}) is Section~\ref{s:sfb} that,
+Recall from~(\ref{eq:U}) in Section~\ref{s:sfb} that,
 $$	\pi_\theta(\theta) \propto p(\theta | X, b) \propto p(b | X, \theta) p(\theta) \propto  \exp \left( - U(\theta) \right),
@@
 
-\subsection{Choosing the MALA step-size}
-\label{appdx:step-size}
-
-Recall that in 
-Section~\ref{s:sfb} we used 
-the Metropolis Adjusted Langevin Algorithm (MALA)
-in order to 
-sample from the $\theta$-chain of the block membership 
-generator parameters.
-At iteration $t$, the proposed sample is generated by:
-%
-\begin{equation}
-	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi.
-\end{equation}
-%
-There are two competing objectives when choosing the step-size $h_t$. 
-On the one hand, $h_t$ needs to be large so that the sampler
-arrives at a high density region quickly,
-while too large a step-size would lead to low acceptance rates and thus 
-inefficient sampling. An effective strategy is
-to use {\em simulated annealing}: allow $h_t$ to slowly decrease
-with $t$, as long as $h_t>0$ for all $t$ and also:
-%
-\begin{equation}
-	\sum_{t=1}^{\infty} h_t = \infty, \qquad \textrm{and} \qquad
-	\sum_{t=1}^{\infty} h_t^2 < \infty.
-	\label{eqn:h-constraints}
-\end{equation}
-%
-Following \citet{Bayesian-SGLD}, we adopt the 
-polynomially decaying step-sizes,
-%
-$h_t = \alpha(\beta + t)^{-\gamma}$,
-%
-where $\alpha>0$, $\beta>0$ and $\gamma\in(1/2,1]$ are hyper-parameters.
-We make the specific choices,
-%
-\begin{equation}
-	\alpha = \frac{250 \cdot s}{N}, \qquad \beta = 1000, \qquad \gamma = 0.8,
-	\label{eqn:step-size-params}
-\end{equation}
-%
-where $N$ is the number of data-points and $s$,
-the {\em step-size scaling}, is the only free parameter.
-
-% For approximate methods, we can choose to bypass the MH accept-reject entirely to speed up computation.
-If this is done, the algorithm is instead called stochastic gradient Langevin diffusion (SGLD) \cite{Bayesian-SGLD}.
-This speeds up computation at the expense of exactness of the method.
-
-\subsection{Burn-in and thinning}
-\label{appdx:burn-in-thinning}
-
-% As with any MCMC method, we must deal with the 
-% issues presented by burn-in and thinning.
-
-When sampling from the $b$- and $\theta$-chains described
-in Section~\ref{sec:inference}, we generate
-$T_b$ and $T_\theta$ samples total, respectively.
-We discard an initial proportion $\kappa_\star\in(0,1)$ of the samples 
-as corresponding to a ``burn-in'' period required for the distribution 
-of the chain to reach a distribution close to our target,
-and we also ``thin'' the remaining samples to 
-obain a less-dependent version. For $\star\in\{b,\theta\}$,
-the remaning sample sets are denoted $\Tcal_\star$
-in the notation of Section~\ref{s:ss},
-%
-\begin{equation}
-	\Tcal_\star = \{T_\star \kappa_\star + i \lambda_\star :  
-	0 \leq i \leq \lfloor T_\star(1 - \kappa_\star) / \lambda_\star \rfloor \},
-\end{equation}
-%
-where $\lambda_\star$ controls the thinning. The choice of
-$\kappa_\star$ can be determined by plotting the log-target (either $S(b^{(t)})$ 
-or $U(\theta^{(t)})$ as a function of $t$,
-and choosing $\kappa_\star$ to encompass the region where the log-target has 
-roughly reached equilibrium. As we do not leverage sample independence,
-$\lambda_\star$ can be chosen less rigorously; we often simply
-use $\lambda_b=5$ and $\lambda_\theta = 10$.
-
-\subsection{Initializing the b-chain}
-
-For the purposes of the FFBM model, the number of blocks $B$ is a constant 
-which must be specified. If the choice of $B$ is influenced 
-by the observed data, then the analysis is no longer ``fully Bayesian''
-and belongs to the class of methods referred to as ``empirical Bayes.''
-However, as the number of blocks only specifies the coarseness of the 
-analysis, it is reasonable to allow it to vary. Indeed, 
-\citet{peixoto-determine-B} shows that for a fixed 
-average degree the maximum number of detectable blocks scales 
-as $O(\sqrt{N})$ where $N$ is the number of vertices.
-
-If $B$ is allowed to vary in the $b$-chain (i.e.,
-when new blocks can be created and empty blocks are allowed),
-then the chain can be run until a minimum description 
-length (MDL) solution is reached. We take the number of non-empty blocks 
-in the MDL solution to be our fixed block number $B$ for subsequent analysis. 
-Indeed, it is prudent to start the $b$-chain at this MDL solution as then 
-the necessary burn-in time can be greatly reduced.
+\clearpage
+\section{Computational details}
 

### 2021-05-27 nips/ffbm.tex
More cleanup

@@
 of weight vectors $\theta=\{w_k\}_{k=1}^B$, where each
-$w_k$ has dimension $D+1$. Specifically, 
+$w_k$ has dimension $D$. Specifically, 
 the distribution of $b$ given $X$ and $\theta$ is,
@@
 %
-We deliberately exclude a bias term to ensure that the relationships we model are based on features and not information about the size of each detected block;
-a more complete discussion on this topic is given in \ref{appdx:dimension}. 
+This is the form of a softmax; we deliberately exclude a bias term to ensure that the relationships we model are based on features and not information about the size of each detected block. 
 We will later also find it convenient
-to write the parameters $\theta$ as a $B \times (D+1)$ matrix of weights $W$; this form has computational benefits as then $z_i \coloneqq W x_i$, which is the input to the softmax activation function.
+to write the parameters $\theta$ as a $B \times D$ matrix of weights $W$.
 

### 2021-05-27 nips/appendices.tex
SUBMIT

@@
 the Metropolis Adjusted Langevin Algorithm (MALA)
-in order to 
+to 
 sample from the $\theta$-chain of the block membership 
@@
 
-When sampling from the $b$- and $\theta$-chains described
-in Section~\ref{sec:inference}, we generate
-$T_b$ and $T_\theta$ samples total, respectively.
+When sampling from the $b$ and $\theta$-chains described
+in Section~\ref{sec:inference}, we respectively generate
+$T_b$ and $T_\theta$ samples total.
 We discard an initial proportion $\kappa_\star\in(0,1)$ of the samples 
@@
 
-\clearpage
 \section{Derivations}
@@
 the value of the integral is independent
-of the value of $b_i \in [B]$:
-The integrand has the same form regardless of $b_i$,
+of the value of $b_i \in [B]$ as
+the integrand has the same form regardless of $b_i$. This is
 because the prior is the same for each $w_j$. 
 Therefore, the integral is only a function of $\tilde{x}_i$ and $\sigma_\theta^2$,
-which means that, as function of $b$, $p(b|X)\propto 1$. And as
+which means that, as a function of $b$, $p(b|X)\propto 1$. As
 $b$ takes values in $[B]^N$, we necessarily have:
@@
 $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and 
-$a_{ij} \coloneqq \phi_j(x_i; \theta)$, we have,
+$a_{ij} \coloneqq \phi_j(x_i; \theta)$, we have that,
 %
@@
 where
-$\|\theta\|^2 = \sum_{i} \theta_{i}^2 = \sum_{j=1}^{B} \|w_j\|^2$ 
+$\|\theta\|^2 = \sum_{i} \theta_{i}^2 = \sum_{j \in [B]} \|w_j\|^2$ 
 is the Euclidean norm of the vector of parameters $\theta$.
@@
 	\frac{\partial U}{\partial w_k} &= 
-	\sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} 
+	\sum_{i \in [N]} \sum_{j \in [B]} y_{ij} 
 	\left( -\frac{\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
 	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
-	&=  - \left( \sum_{i=1}^{N} \tilde{x}_i \left( y_{ik} - a_{ik} \sum_{j=1}^{B} y_{ij} \right)
+	&=  - \left( \sum_{i \in [N]} \tilde{x}_i \left( y_{ik} - a_{ik} \sum_{j \in [B]} y_{ij} \right)
 	- \frac{w_k}{\sigma_\theta^2} \right) \nonumber \\
-	&= - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right).
+	&= - \left( \sum_{i \in [N]} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right).
 \end{align}
@@
 efficiently through matrix operations. The only property of $y_{ij}$ 
-we have used in the derivation is the constraint $\sum_{j=1}^{B} y_{ij} = 1$,
+we have used in the derivation is the constraint $\sum_{j \in [B]} y_{ij} = 1$,
 for all $i$.
@@
 	\begin{algorithmic} % enter the algorithmic environment
+		\Procedure{SampleBlockMemberships}{$A, T_b$}
 		\State $b^{(0)} \gets \argmin_b S(b|A)$ \Comment{Implemented as greedy heuristic in \textit{graph-tool} library}
@@
 		\State \textbf{return} $\{b^{(t)}\}_{t=1}^{T_b}$
-
+		\EndProcedure
 	\end{algorithmic}
@@
 	\begin{algorithmic} % enter the algorithmic environment
-
+		\Procedure{SampleFeatureWeights}{$X, \{b^{(t)}\}, \Tcal_b, \sigma_\theta, s$}
 		\State $\hat{Y}_{ij} \gets \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} \one \{ b^{(t)}_i = j\} \quad \forall i,j$
@@
 		\State $\xi \gets \sim \Gaussian(0, I)$
-		\State $\theta' \gets \theta^{(t)} - h_t \nabla U(\theta^{(t)} | X, \hat{Y}) + \sqrt{2h_t} \cdot \xi$
+		\State $h_t \gets \frac{s}{N} \cdot 250(1000 + t)^{-0.8}$
+		\State $g_t \gets \nabla U(\theta^{(t)}| X, \hat{Y})$
+		\item[]
+		\State $\theta' \gets \theta^{(t)} - h_t \cdot g_t + \sqrt{2h_t} \cdot \xi$
 		\State $\log \alpha_\theta \gets \log \alpha_\theta(\theta^{(t)}, \theta' | A, \hat{Y})$
@@
 		\State \textbf{return} $\{\theta^{(t)}\}_{t=1}^{T_\theta}$
+		\EndProcedure
+	\end{algorithmic}
+\end{algorithm}
 
+\begin{algorithm} % enter the algorithm environment
+	\caption{Dimensionality reduction} % give the algorithm a caption
+	\label{alg:retained-set} % and a label for \ref{} commands later in the document
+	\begin{algorithmic} % enter the algorithmic environment
+		\Procedure{ReduceDimension}{$\{W^{(t)}\}, \Tcal_\theta, k, D'$}
+
+		\State $(B, D) \gets W^{(0)}$.shape
+		\State $\hat{\mu}_{ij} \gets \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} W_{ij}^{(t)} \quad \forall i \in [B], j \in [D]$
+		\State $\hat{\sigma}_{ij} \gets \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2 \quad \forall i \in [B], j \in [D]$
+
+		\item[]
+
+		\For{$d \in [D]$}
+		\For{$i \in [B]$}
+		\State $l_i \gets \hat{\mu}_{id} - k \cdot \hat{\sigma}_{id}$
+		\State $u_i \gets \hat{\mu}_{id} + k \cdot \hat{\sigma}_{id}$
+		\If{$l_i \leq 0$ and $u_i \geq 0$}
+		\State $l_i, u_i \gets 0$
+		\EndIf
+		\EndFor
+		\State $c_d \gets \max_i \min\left(|l_i|, |u_i| \right)$
+		\EndFor
+
+		\item[]
+
+		\State indexArray $\gets$ indexSort($c$, descending=True)[$0:D'$]
+		\State $d^* \gets$ indexArray[-1]
+		\State $\Dcal' \gets $ Set(indexArray)
+		\State $c^* \gets c_{d^*}$
+		\State \textbf{return} $\Dcal', c^*$
+		\EndProcedure
 	\end{algorithmic}
@@
 	\centering
-	\caption{Compute-time for each experiment.}
+	\caption{Compute-time for each experiment}
 	\label{tab:compute-time}
@@
 		Dataset & $b$-chain & $\theta$-chain & Reduced $\theta$-chain & Overall compute time \\ \hline
-		Polbooks & $\sim$1s & $\sim$4s & -- & $\sim$5s \\
-		School & $\sim$10s & $\sim$10s & $\sim$10s & $\sim$30s \\
-		FB Egonet & $\sim$20s & $\sim$180s & $\sim$10s & $\sim$210s
+		Polbooks & $\sim$1s & $\sim$10s & -- & $\sim$11s \\
+		School & $\sim$1s & $\sim$7s & $\sim$7s & $\sim$15s \\
+		FB Egonet & $\sim$2s & $\sim$50s & $\sim$8s & $\sim$60s
 	\end{tabular}

### 2021-05-27 nips/conclusion.tex
SUBMIT

@@
 the parameters $\theta$ of the feature-to-block generator. 
-This takes the form of two concurrent Markov chains,
+This takes the form of a two-level Markov chain,
 used to sample the block memberships $b$ and block generator 
@@
 the FFBM to be hierarchical in nature. That way, the structure of the network 
-can be explained at all several scales of interest. So long as data 
+can be explained at all length-scales of interest. So long as data 
 collection techniques remain ethical and care is taken to respect 

### 2021-05-27 nips/experiments.tex
SUBMIT

@@
 	\end{subfigure}
-	\caption{Reduced dimension feature-to-block generator weight samples.}
+	\caption{Reduced dimension feature-to-block generator weight samples. Dotted line is $\pm c^*$}
 	\label{fig:null-all}
@@
 
-Next, in order to assess the performance of the feature-to-block predictor, 
+Next, to assess the performance of the feature-to-block predictor, 
 we partition the vertex set $[N]$ into training and test sets. We choose to 
@@
 %
-and where $\star \in \{0, 1\}$ indicates whehter the training or test
+where $\star \in \{0, 1\}$ indicates whether the training or test
 set is being considered.
@@
 in order to reduce the dimension from 
-$D$ to a desired $D'$ with $k=1$ based on the maximal cutoff $c^*$. 
+$D$ to a desired $D'$. 
 We then retrain the feature-block predictor using only the retained 
@@
 of the graphs in Figure~\ref{fig:graphs-all}.
-The polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do succeed at partitioning the graph into densely connected clusters.
+The polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do succeed at dividing the graph into densely connected clusters.
 
@@
 of the overall network structure.
-We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal.
-or neutral).
+We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral).
 From Figure~\ref{fig:polbooks-null} we see that all 3 blocks have a distinct political affiliation as their largest positive component.

### 2021-05-28 nips/abstract.tex

@@
 	feature-based communities. Nevertheless, the resulting communities
-	often do not produce a natural partition of the graph. In this work
+	often do not produce a natural partition of the graph. In this work,
 	we introduce a new generative model, the feature-first block model (FFBM),

### 2021-05-28 nips/ffbm.tex

@@
 Let $N$ denote the number of nodes, $B$ the number of blocks in the graph,
-and ${\cal X}$ the set of all possible features.
+and ${\cal X}$ the set of values each feature can take.
 We define the vector $x_i \in \Xcal^D$ as the feature vector of vertex $i$, 
@@
 For the FFBM, we start with the feature matrix $X$ and generate a random
-vector of block memberships $b \in [B]^N$. For each node $i$, the
+vector of block memberships $b \in [B]^N$. For each vertex $i$, the
 block membership $b_i\in[B]$ is generated based on the feature
-vector $x_i$, independently between nodes. The conditional
+vector $x_i$, independently between vertices. The conditional
 distribution of $b_i$ given $x_i$ also depends on a collection
@@
 	p(b| X, \theta) = \prod_{i=1}^{N} p(b_i | x_i, \theta) = \prod_{i=1}^{N} \phi_{b_i} (x_i; \theta)
-	= \prod_{i=1}^{N} \frac{\exp\left(w_{b_i}^T x_i\right)}{\sum_{k=1}^{B} \exp \left( w_k^T x_i\right)}
+	= \prod_{i=1}^{N} \frac{\exp\left(w_{b_i}^T x_i\right)}{\sum_{k=1}^{B} \exp \left( w_k^T x_i\right)}.
 \end{equation}
 %
-This is the form of a softmax; we deliberately exclude a bias term to ensure that the relationships we model are based on features and not information about the size of each detected block. 
+$\phi_{b_i}$ is the form of a softmax activation function; we deliberately exclude a bias term to ensure that we only leverage feature information. 
 We will later also find it convenient
@@
 each element of $\theta$ has an independent ${\cal N}(0,\sigma_\theta^2)$
-prior, for a hyperparameter $\sigma_\theta^2$:
+prior, with hyperparameter $\sigma_\theta^2$:
 %
@@
 Interestingly, this choice of prior leads to a particularly
-simple distribution on the block membership vector $b$.
+simple distribution on the block membership vector $b$ given $X$.
 We show in Appendix~\ref{appdx:b|x} that,
 after integrating out $\theta$,
-$b$ is uniformly distributed:
+$b$ is uniformly distributed given $X$:
 %
@@
 \citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for 
-the additional microcanonical SBM parameters $\psi$, which we adopt. 
+the priors on the additional microcanonical SBM parameters $\psi$, which we adopt. 
 The idea is to write the joint distribution on $(b, e, k)$ as a product of 

### 2021-05-28 nips/prelim.tex

@@
 Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}.
-To allow for degree-variability between members of the same block, we adopt the degree-corrected formulation (DC-SBM).
+To allow for degree-variability between members of the same block, we adopt the degree-corrected formulation (DC-SBM), defined in~(\ref{defn:microcan-dc-sbm}).
 
@@
 	Let $e=(e_{rs})$ be the $N\times N$ symmetric matrix of edge counts 
-between blocks, so that $e_{rs}$ is the number of edges from block $r$ to 
+between blocks, such that $e_{rs}$ is the number of edges from block $r$ to 
 block $s$ -- or twice that number if $r=s$. 

### 2021-05-28 cued/derivations.tex
Repopulate cued folder

--- 2021-05-27 nips/derivations.tex
@@
 We note that $b_i \in [B]$ and so the integral's value is unchanged with respect to $b_i$.
-The integrand has the same form no matter which value $b_i$ takes as the prior is the same for each $w_j$.
+In other words, the integrand has the same form no matter which value $b_i$ takes as the prior is the same for each $w_j$.
 As such the integral can only be a function of at most $\tilde{x}_i$ and $\sigma_\theta^2$ as it is symmetric with respect to $b_i$ and all the various $w_j$ are integrated out as they are dummy variables.
@@
 \end{equation}
-
+%
+This is an extremely important result as it allows samples to be drawn from $p(b | A, X) \propto p(A|b, X) \cdot p(b | X) = p(A|b) \cdot p(b|X)$ without a complicated Monte-Carlo integration over $\theta$ to determine $p(b|X)$.
+Indeed, we do not even need to know the exact value of $X$ to compute $p(b|X)$.
 \subsection{Derivation of U form}
@@
 Therefore, $\nabla U(\theta) = \left[\nicefrac{\partial U}{\partial w_1}^T,\nicefrac{\partial U}{\partial w_2}^T \dots \nicefrac{\partial U}{\partial w_B}^T  \right]^T$; to compute $\nabla U(\theta)$ it suffices to find the form of $\nicefrac{\partial U}{\partial w_k}$ with respect to a general $k$.
-
-To this end, we must first find partial derivatives of $a_{ij}$ and $||\theta||$ with respect to $w_k$. Starting with $a_{ij}$:
+To this end, we must first find partial derivatives of $a_{ij}$ and $||\theta||$ with respect to $w_k$.
+Starting with $a_{ij}$:
 %
@@
 Indeed, the Laplace approximation is often used for modelling the posterior in logistic classification \cite{laplace}.
-As a simplification we assume that $\Sigma$ is diagonal, therefore each element of $\theta$ is independent.
-This assumption does not hold in general but it motivates the derivation of the dimensionality reduction we construct by analogy with hypothesis testing.
-Assuming, $\Sigma$ is diagonal then the posterior for each element of the weight matrix $W$ can be approximated by:
+This approximation is not exact but it motivates the derivation of the dimensionality reduction technique we construct by analogy with hypothesis testing.
+Marginalising over the other elements, the posterior for each element of the weight matrix $W$ can be approximated by:
 %
@@
 %
-This approximation is not exact but can show it is accurate empirically.
+This approximation is not exact but we can show it is accurate empirically.
 Indeed, if we run the primary school experiment with hyper-parameters given in appendix \ref{appdx:hyperparams}, we can then plot histograms of the collected $W$-samples and compare these to the Laplace approximation.
@@
 %
-Clearly equations \ref{eqn:hyp-test-positive} and \ref{eqn:hyp-test-negative} cannot both hold as $c>0$. Nevertheless, a necessary and sufficient condition for one of them holding is:
+Clearly equations \ref{eqn:hyp-test-positive} and \ref{eqn:hyp-test-negative} cannot both hold simultaneously as $c, \hat{\sigma}_{ij}, k>0$.
+Nevertheless, a necessary and sufficient condition for one of them to hold is:
 %
@@
 %
-We therefore choose to reject the null hypothesis $H_0$ if and only if equation \ref{eqn:hyp-test-empty}.
+We therefore choose to reject the null hypothesis $H_0$ if and only if equation \ref{eqn:hyp-test-empty} holds.
 This tells us with confidence greater than $\Phi(k)$ that either $W_{ij} < -c$ or $W_{ij} > c$ thus rejecting the null $H_0: |W_{ij}| \leq c$.

### 2021-05-28 cued/introduction.tex
Repopulate cued folder

@@
 
-A somewhat surprising property of many real-world networks is that they exhibit strong community structure; most nodes often belong to a densely connected cluster.
-There is high interest in recovering the latent communities from the observed graphs.
-The inferred communities can be exploited for compression algorithms \cite{cluster-compression} or used for link prediction in incomplete networks \cite{link-prediction} to name a few applications.
+An important characteristic of many real-world networks is that they exhibit strong community structure, with most nodes often belonging to a densely connected cluster. 
+Finding ways to recover the latent communities from the observed graph is an important
+task in many applications, including
+graph/network compression \cite{cluster-compression} and link prediction in 
+incomplete networks \cite{link-prediction}.
 
-We restrict our analysis to vertex-labelled networks.
-We shall refer to the vertex-labels as features.
+In this work, we
+restrict our attention to vertex-labelled networks, and we
+refer to the vertex labels as {\em features}.
 A common goal is to determine whether a given feature impacts graphical structure.
-To answer this from a Bayesian perspective we must use a random graph model; the standard is the stochastic block model (SBM) \cite{vanilla-sbm}.
+To answer this from a Bayesian perspective requires the use of a random graph model; the standard is the stochastic block model (SBM) \cite{vanilla-sbm}.
 This is a latent variable model where each vertex belongs to a single block and the probability two nodes are connected depends only on the block memberships of each.
-There have been many variants to this model -- the most popular being the mixed-membership stochastic block model (MMSBM) \cite{mixed-membership-sbm} and the overlapping stochastic block model (OSBM) \cite{overlapping-sbm}.
-Effectively, these just extend the model to allow each vertex to belong to multiple blocks simultaneously.
-However, a major drawback of these graphical models as applied to labelled networks is that they do not automatically include vertex features in the random graph generation process.
+Numerous variants of this
+model have been considered -- the most popular ones being the mixed-membership stochastic block model (MMSBM) \cite{mixed-membership-sbm} and the overlapping stochastic block model (OSBM) \cite{overlapping-sbm}.
+Effectively, these extend the model to allow each vertex to belong to multiple blocks simultaneously.
+However, one drawback of these graphical models as applied to labelled networks is that they do not automatically include vertex features in the random graph generation process.
 Approaches based on graph neural networks \cite{sbm-gnn} that utilise vertex features have been developed but these lack the easy interpretability of the simpler models.
 
-To analyse a labelled network using one of the simple SBM variants, a typical inference procedure would be to partition the graph into blocks grouped by distinct values of the feature of interest.
-The associated model can then be used to test for evidence of heterogeneous connectivity between the feature-grouped blocks.
+To analyse a labelled network using one of the simple SBM variants, a typical inference procedure would be to first partition the graph into blocks grouped by distinct values of the feature of interest, and then use the associated model to test for evidence of heterogeneous connectivity between the feature-grouped blocks.
 Nevertheless, this approach is limited in that it can only consider one feature at a time.
-This makes it difficult to rank order the features by  magnitude of impact.
+This makes it difficult to rank-order the features by  magnitude of impact.
 Lastly, the feature-grouped blocks are often an unnatural partition of the graph, leading to a poor model fit.
@@
 
-With these desiderata in mind, we present a novel framework for modelling labelled networks, which we call the feature-first block model (FFBM).
+Thus motivated, we present a novel framework for modelling labelled networks, which we call the feature-first block model (FFBM).
 This is an extension of the SBM to labelled networks.
-We go on to present an efficient algorithm for sampling from the parameters of the feature-to-block generator.
-We can interpret the sampled FFBM parameters to determine which features have the largest impact on overall graphical structure.
+We go on to present an efficient algorithm for sampling from the parameters of the feature-to-block generator, and to describe how the
+sampled FFBM parameters can be interpreted to determine which features have the largest impact on overall graphical structure.

### 2021-05-28 cued/tech-abstract.tex
Repopulate cued folder

--- 2021-05-28 nips/abstract.tex
@@
-\begin{abstract}
+\section*{Technical Abstract}
 Labelled networks form a very common and important class of data,
@@
 illustrate the performance of the proposed methods.
-\end{abstract}

### 2021-05-28 cued/additional.tex
Copy over relevant bits

--- 2021-05-27 nips/additional.tex
@@
 As discussed in the main text, the block membership vector $b$ is an intermediate variable in the FFBM and not a parameter so we are not free to choose a prior for it.
-Nevertheless, we can borrow the conditional prior proposed by \citet{Peixoto-Bayesian-Microcanonical} for the other parameters to give:
+
+Here we recall for reference
+the priors $p(\psi | b)$ from \cite{Peixoto-Bayesian-Microcanonical}:
 %
 \begin{equation}
-	p(e, k | b) = p(e | b) p(k| e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
-	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right]
+	p(\psi_e=e, \psi_k=k | b) = p(e | b) p(\psi_k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
+	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right],
 \end{equation}
 %
-Where $\specialchoose{n}{m}$ is shorthand for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$ which can be thought of as the total number of distinct histograms with $n$ bins under the constraint they sum to $m$.
+where $\specialchoose{n}{m}$ is shorthand 
+for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$,
+which can be thought of as the total number of distinct histograms 
+produced by $m$ samples in $n$ bins.
+The value
 $E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges in the graph. 
-Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform with respect to $e$.
-The variable $\eta_j^r$ is introduced to denote the number of vertices in block $r$ that have degree $j$.
-Formally, $\eta_j^r \coloneqq \sum_{i} \one\left\{b_i = r \right\} \one \left\{k_i = j \right\}$.
-Furthermore, $q(m, n)$ is the number of different histograms with at most $n$ non-zero bins that sum to $m$.
-$q(m, n)$ is related to but different from $\specialchoose{n}{m}$.
-Recall that $e_r \coloneqq \sum_{s} e_{rs}$ is the total number of half edges in block r and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ is the number of vertices assigned to block $r$.
+Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform in $e$.
+The variable $\eta_j^r$ denotes the number of vertices in block $r$ 
+that have degree $j$; formally, $\eta_j^r \coloneqq \sum_{i} \one\left\{b_i = r \right\} \one \left\{k_i = j \right\}$. 
+The denominator $q(m, n)$ denotes the number of different histograms 
+produced by $m$ samples in 
+at most $n$ non-zero bins that sum to $m$. 
+Finally, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number 
+of half edges in block $r$ and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ 
+is the number of vertices assigned to block $r$.
 
-The form of these priors were chosen carefully by \citet{Peixoto-Bayesian-Microcanonical} to more closely match the structure of empirical networks than simple uniform priors.
-We do not repeat his arguments here.
+These were chosen carefully in \cite{Peixoto-Bayesian-Microcanonical} to 
+more closely match the structure of empirical networks than simple 
+uniform priors. We do not repeat these arguments here.
 
@@
 
-For sampling from the $\theta$-chain of the block membership generator parameters, we employ the Metropolis Adjusted Langevin Algorithm (MALA). At iteration $t$, the proposed sample is generated by:
+Recall that in 
+Section~\ref{s:sfb} we used 
+the Metropolis Adjusted Langevin Algorithm (MALA)
+to 
+sample from the $\theta$-chain of the block membership 
+generator parameters.
+At iteration $t$, the proposed sample is generated by:
 %
 \begin{equation}
-	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi
+	\theta' = \theta^{(t)} - h_t \nabla U(\theta^{(t)}) + \sqrt{2h_t} \cdot \xi.
 \end{equation}
@@
 %
-The first constraint ensures that we can cover sufficient distance to arrive at any arbitrary point in our domain, no matter the starting point.
-The second constraint ensures that once we converge to the mode we will stay there, rather than simply bouncing around it.
-\citet{Bayesian-SGLD} propose the following form for a polynomially decaying step-size which we adopt:
+Following \citet{Bayesian-SGLD}, we adopt the 
+polynomially decaying step-sizes,
+%
+$h_t = \alpha(\beta + t)^{-\gamma}$,
+%
+where $\alpha>0$, $\beta>0$ and $\gamma\in(1/2,1]$ are hyper-parameters.
+We make the specific choices,
 %
 \begin{equation}
-	h_t = \alpha(\beta + t)^{-\gamma}
-\end{equation}
-%
-Where $\alpha, \beta, \gamma$ are hyper-parameters to be chosen.
-We require $\alpha,\beta > 0$ and $\gamma \in (0.5, 1]$ to satisfy equation \ref{eqn:h-constraints}.
-To reduce the number of hyperparameters we set these to have values given by the equations in \ref{eqn:step-size-params}.
-%
-\begin{equation}
-	\alpha = \frac{250 \cdot s}{N} \qquad \beta = 1000 \qquad \gamma = 0.8
+	\alpha = \frac{250 \cdot s}{N}, \qquad \beta = 1000, \qquad \gamma = 0.8,
 	\label{eqn:step-size-params}
@@
 %
-Where $N$ is the number of data-points we are considering and now $s$ is the only free variable which we call the step-size scaling.
-For approximate methods, we can choose to bypass the MH accept-reject entirely to speed up computation.
+where $N$ is the number of data-points and $s$,
+the {\em step-size scaling}, is the only free parameter.
+%
+%For approximate methods, we can choose to bypass the MH accept-reject entirely to speed up computation.
 If this is done, the algorithm is instead called stochastic gradient Langevin diffusion (SGLD) \cite{Bayesian-SGLD}.
@@
 Indeed, it relies on the fact that as the step-size is annealed to 0, the acceptance probability is close 1.
-
+%
 Nevertheless, to choose the value of $s$ we must quantify the trade-off between acceptance ratio and burn-in time.
@@
 
-As with any MCMC method, we must deal with the issues presented by burn-in and thinning.
-We have introduced the notation $\Tcal_b$ and $\Tcal_\theta$ to denote the set of samples we keep from the $b$ and $\theta$ chains respectively.
-Note that we generate $T_b$ and $T_\theta$ samples total.
-The burn-in period refers to the time taken for the Markov Chain to converge to the stationary distribution.
-Sample thinning is necessary to ensure that neighbouring samples satisfy independence.
-However, as we do not leverage the independence property this is less important in our analysis.
-We can write the general set $\Tcal_\star$ as:
+%As with any MCMC method, we must deal with the issues presented by burn-in and thinning.
+When sampling from the $b$ and $\theta$-chains described
+in Section~\ref{sec:inference}, we respectively generate
+$T_b$ and $T_\theta$ samples total.
+We discard an initial proportion $\kappa_\star\in(0,1)$ of the samples 
+as corresponding to a ``burn-in'' period required for the distribution 
+of the chain to reach a distribution close to our target,
+and we also ``thin'' the remaining samples to 
+obtain a less-dependent version. For $\star\in\{b,\theta\}$,
+the remaining sample sets are denoted $\Tcal_\star$
+in the notation of Section~\ref{s:ss},
 %
@@
 
-For the purposes of our model (the FFBM), the number of blocks $B$ is a constant which must be specified by the data scientist.
-We could however, allow our choice of $B$ to be influenced by the observed data.
-This places us in the domain of empirical Bayes, which must be negotiated carefully.
-Prior beliefs must be determined a priori else they are not prior.
-However, as the number of blocks only specifies the coarseness of the analysis, it is fine to allow it to vary.
-Indeed, \citet{peixoto-determine-B} shows that for a fixed average degree the maximum number of detectable blocks scales as $O(\sqrt{N})$ where $N$ is the number of vertices.
+For the purposes of the FFBM model, the number of blocks $B$ is a constant 
+which must be specified. If the choice of $B$ is influenced 
+by the observed data, then the analysis is no longer ``fully Bayesian''
+and belongs to the class of methods referred to as ``empirical Bayes.''
+However, as the number of blocks only specifies the coarseness of the 
+analysis, it is reasonable to allow it to vary. Indeed, 
+\citet{peixoto-determine-B} shows that for a fixed 
+average degree the maximum number of detectable blocks scales 
+as $O(\sqrt{N})$ where $N$ is the number of vertices.
 
-If we allow $B$ to vary in the $b$-chain (i.e.
-new blocks can be created and we permit empty blocks) then it can be run  until a minimum description length (MDL) solution is reached.
-We take the number of non-empty blocks at the MDL solution to be our fixed block number $B$ for subsequent analysis.
-Indeed, it is prudent to start our $b$-chain at this MDL solution as then we can burn-in time is greatly reduced.
+If $B$ is allowed to vary in the $b$-chain (i.e.,
+when new blocks can be created and empty blocks are allowed),
+then the chain can be run until a minimum description 
+length (MDL) solution is reached. We take the number of non-empty blocks 
+in the MDL solution to be our fixed block number $B$ for subsequent analysis. 
+Indeed, it is prudent to start the $b$-chain at this MDL solution as then 
+the necessary burn-in time can be greatly reduced.
 

### 2021-05-28 cued/derivations.tex
Copy over relevant bits

@@
 
-\subsection{Derivation of conditional block distribution given feature matrix}
+\subsection{Derivation of {\boldmath $p(b|X)$}}
 \label{appdx:b|x}
 
-We wish to determine the form of $p(b| X)$. This can be done by integrating over the joint probability with respect to $\theta$.
+We determine the form of $p(b| X)$ by integrating
+out the parameters $\theta$. From the definitions we have:
 %
@@
 	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i \in [N] } \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
-	&= \prod_{i \in [N]} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j \in [B]} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k \in [B]} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}
+	&= \prod_{i \in [N]} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j \in [B]} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k \in [B]} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}.
 \end{align*}
 %
-We note that $b_i \in [B]$ and so the integral's value is unchanged with respect to $b_i$.
-In other words, the integrand has the same form no matter which value $b_i$ takes as the prior is the same for each $w_j$.
-As such the integral can only be a function of at most $\tilde{x}_i$ and $\sigma_\theta^2$ as it is symmetric with respect to $b_i$ and all the various $w_j$ are integrated out as they are dummy variables.
-Therefore, denoting the integral by the (unknown) function $f(\tilde{x}_i, \sigma_\theta^2)$, we write $p(b| X)$ as follows:
-%
-\begin{align*}
-	p(b | X) &= \prod_{i=1}^{N} f(\tilde{x}_i, \sigma_\theta^2) = \textrm{const w.r.t } b = c
-\end{align*}
-%
-As this is a constant with respect to $b$ we conclude that $p(b | X)$ must be a uniform distribution.
-$\nicefrac{1}{c}$ is simply the size of the set of values that $b$ can take.
-We know $b_i \in [B]$.
-Therefore, $b \in [B]^N$ and $|[B]^N| = B^N = \nicefrac{1}{c}$.
-Putting this all together we conclude that:
+The key observation here is that 
+the value of the integral is independent
+of the value of $b_i \in [B]$ as
+the integrand has the same form regardless of $b_i$. This is
+because the prior is the same for each $w_j$. 
+Therefore, the integral is only a function of $\tilde{x}_i$ and $\sigma_\theta^2$,
+which means that, as a function of $b$, $p(b|X)\propto 1$. As
+$b$ takes values in $[B]^N$, we necessarily have:
 %
 \begin{equation}
-	p(b | X) = B^{-N}
+	p(b | X) = \frac{1}{\big|[B]^N\big|}=B^{-N}.
 \end{equation}
-%
-This is an extremely important result as it allows samples to be drawn from $p(b | A, X) \propto p(A|b, X) \cdot p(b | X) = p(A|b) \cdot p(b|X)$ without a complicated Monte-Carlo integration over $\theta$ to determine $p(b|X)$.
-Indeed, we do not even need to know the exact value of $X$ to compute $p(b|X)$.
-\subsection{Derivation of U form}
+
+\subsection{Derivation of {\boldmath $\;U(\theta)$}}
 \label{appdx:form-U}
 
-The invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. We write this as follows:
-%
-\begin{align}
-	\pi_\theta(\theta) &\propto p(\theta | X, b) \propto p(b | X, \theta) p(\theta) \propto  \exp \left( - U(\theta) \right) \\
-	\therefore U(\theta) &= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}
-\end{align}
-%
-Where we have introduced $U(\theta)$ equal to the negative log posterior.
-Each of the constituent terms of $U(\theta)$ are easily computed (equation \ref{eqn:U-constituent-terms}) by defining $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
+Recall from~(\ref{eq:U}) in Section~\ref{s:sfb} that,
+$$	\pi_\theta(\theta) \propto p(\theta | X, b) \propto p(b | X, \theta) p(\theta) \propto  \exp \left( - U(\theta) \right),
+$$ 
+so that $U$ can be expressed as,
+$$
+U(\theta) 
+= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}.
+$$
+Writing,
+$y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and 
+$a_{ij} \coloneqq \phi_j(x_i; \theta)$, we have that,
 %
@@
 	\log p(b | X, \theta) = \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
-	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} || \theta ||^2
+	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} 
+	\|\theta \|^2,
 	\label{eqn:U-constituent-terms}
@@
 %
-Discarding constant terms, we write $U(\theta)$ as in equation \ref{eqn:U-form-appdx-0}.
-Note that $||\theta||^2 = \sum_{i} \theta_{i}^2 = \sum_{j=1}^{B} ||w_j||^2$ is the Euclidean norm of the vector of parameters $\theta$.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2 = N \cdot \Lcal(\theta) + \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form-appdx-0}
-\end{equation}
-%
+where
+$\|\theta\|^2 = \sum_{i} \theta_{i}^2 = \sum_{j \in [B]} \|w_j\|^2$ 
+is the Euclidean norm of the vector of parameters $\theta$.
+Therefore, discarding constant terms, we 
+obtain exactly the representation~(\ref{eqn:U-form}), as claimed.
 
-\subsection{Derivation of U gradient with respect to feature parameters}
+\subsection{Derivation of {\boldmath $\;\nabla U(\theta)$}}
 \label{appdx:gradu}
-The goal is to determine $\nabla U(\theta)$, the gradient of the negative log posterior with respect to the parameters. We repeat the form of $U(\theta)$ in equation \ref{eqn:U-form-appdx}.
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2
-	\label{eqn:U-form-appdx}
-\end{equation}
-%
-Where $y_{ij}$ is independent of $\theta$ and $a_{ij}$ is the output from the softmax layer, with form as given in equation \ref{eqn:a-ij}.
-%
-\begin{equation}
-	a_{ij} \coloneqq \phi_{j} (x_i; \theta) = \frac{\exp(w_j^T \tilde{x}_i)}{\sum_{r \in [B]} \exp(w_r^T \tilde{x}_i)}
-	\label{eqn:a-ij} 
-\end{equation}
-%
-We note that $\theta = \{w_k\}_{k=1}^B$, and as such we can write this in vector form $\theta = \left[w_1^T, w_2^T \dots w_B^T  \right]^T$.
-Therefore, $\nabla U(\theta) = \left[\nicefrac{\partial U}{\partial w_1}^T,\nicefrac{\partial U}{\partial w_2}^T \dots \nicefrac{\partial U}{\partial w_B}^T  \right]^T$; to compute $\nabla U(\theta)$ it suffices to find the form of $\nicefrac{\partial U}{\partial w_k}$ with respect to a general $k$.
-To this end, we must first find partial derivatives of $a_{ij}$ and $||\theta||$ with respect to $w_k$.
-Starting with $a_{ij}$:
+Here we show how the gradient 
+$\nabla U(\theta)$ can be computed explicitly.
+Recall the expression for $U(\theta)$ in~(\ref{eqn:U-form}).
+Writing $\theta$ as
+$\theta = \left[w_1^T, w_2^T \dots w_B^T  \right]^T$,
+in order to compute the gradient
+$\nabla U(\theta)$ we need to compute
+each of its components,
+$\nicefrac{\partial U}{\partial w_k}$,
+$1\leq k\leq B$.
+To that end, we first compute,
 %
@@
 	{\left( \sum_{r \in [B]} \exp(w_r^T \tilde{x}_i) \right)^2} \nonumber \\
-	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right) 
+	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right), 
+	\label{eq:dadw}
 \end{align}
 %
-Where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$. Now moving onto the derivative of $||\theta||^2$:
+where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$,
+and we also easily find,
 %
 \begin{equation}
-	\frac{ \partial}{\partial w_k} ||\theta||^2 = \frac{\partial}{\partial w_k} \left( \sum_{r \in [B]} ||w_r||^2 \right) = 2w_k
+	\frac{ \partial}{\partial w_k} \|\theta\|^2 = \frac{\partial}{\partial w_k} \left( \sum_{r \in [B]} \|w_r\|^2 \right) = 2w_k.
+	\label{eq:dtsdw}
 \end{equation}
 %
-We are ready to put this all together, to find the partial derivative of $U(\theta)$ with respect to each $w_k$:
+Using~(\ref{eq:dadw}) and~(\ref{eq:dtsdw}), we obtain,
 \begin{align}
 	\frac{\partial U}{\partial w_k} &= 
-	\sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} 
-	\left( \frac{-\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
+	\sum_{i \in [N]} \sum_{j \in [B]} y_{ij} 
+	\left( -\frac{\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
 	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
-	&=  - \left( \sum_{i=1}^{N} \tilde{x}_i \left( y_{ik} - a_{ik} \sum_{j=1}^{B} y_{ij} \right)
+	&=  - \left( \sum_{i \in [N]} \tilde{x}_i \left( y_{ik} - a_{ik} \sum_{j \in [B]} y_{ij} \right)
 	- \frac{w_k}{\sigma_\theta^2} \right) \nonumber \\
-	&= - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
+	&= - \left( \sum_{i \in [N]} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right).
 \end{align}
 %
-This is the required result.
-This form can be computed efficiently through matrix operations.
-The only property of $y_{ij}$ we have used in the derivation is the sum-to-one constraint $\sum_{j=1}^{B} y_{ij} = 1$ for all $i$.
+This can be computed 
+efficiently through matrix operations. The only property of $y_{ij}$ 
+we have used in the derivation is the constraint $\sum_{j \in [B]} y_{ij} = 1$,
+for all $i$.
 

### 2021-05-28 cued/inference.tex

@@
 
-Now that we have defined the FFBM, we wish to leverage it to perform inference.
-Suppose we are presented with a vertex-labelled graph $(A, X)$; the goal is to draw samples for $\theta$ according to the posterior given the observed data:
+Having completed the definition of the FFBM, we wish to leverage it 
+to perform inference. Given a vertex-labelled graph $(A, X)$,
+the goal is to sample from the posterior distribution
+of $\theta$:
 %
@@
 %
-This is an example of a pseudo-marginal approach.
-Indeed, \citet{pseudo-marginal} show that the unbiased result in equation \ref{eqn:theta-unbiased} is sufficient to prove that for large enough $t$, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{(t)})\right] = p(\theta
-| A, X)$ which is exactly the distribution we are targetting (equation \ref{eqn:theta-target}).
-%
+This is an example of a pseudo-marginal approach. Indeed, \citet{pseudo-marginal} show that~(\ref{eqn:theta-unbiased})
+% the unbiased result in equation 
+is sufficient to prove that, for large enough $t$, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{(t)})\right] = p(\theta 
+| A, X)$, as required.
+%  which is exactly the distribution we are targetting (equation \ref{eqn:theta-target}).
+
 \begin{figure}[!h]
@@
 
-The reason we split the Markov chain into two stages is because the summation over all latent states $b \in [B]^N$ required to directly compute the likelihood $p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$ is intractable -- $O(B^N)$.
+The reason for splitting the Markov chain into two levels is that the summation over all latent states $b \in [B]^N$ required to directly compute the likelihood, $p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$,
+is intractable, as the sum involves $B^N$ terms. 
 Figure \ref{fig:samp-sequence} shows an overview of the proposed method. 
-We have introduced subscripts and conditionings to make explicit what variables each step utilises.
-We note the power of the simplification given by equation \ref{eqn:b-pseudo-prior}.
-As $p(b| X)$ does not depend on the exact value of X, we do not need to know the value of $X$ to perform the sampling on $b$.
-Conversely, for the $\theta^{(t)}$ samples, we use only $b^{(t)}$ but not $A$ as $(\theta \indep A) | b$.
+The Metropolis-Hastings accept/reject probabilities 
+for the proposed $b$ and $\theta$ samples are denoted $\alpha_b$ and
+$\alpha_\theta$, respectively.
+% We have introduced subscripts and conditionings to make explicit what variables each step utilises.
+
+Note the importance of the simplification in~(\ref{eqn:b-pseudo-prior}). 
+As $p(b| X)$ in fact does not depend on $X$, 
+knowing $X$ is not necessary in order to obtain samples
+of the block membership vector $b$.
+And on the other level, in order to obtain 
+samples of the parameter vector $\theta$
+we use only $b$ but not $A$, as $\theta$ and $A$ are
+conditionally independent given $b$.
 
 \FloatBarrier
-
+\clearpage
 \subsection{Sampling block memberships}
 
-\citet{Peixoto-MCMC} proposes a Monte Carlo method which we will base our approach on. It relies on writing the posterior in the following form:
+We adopt the Markov chain Monte Carlo procedure of
+\cite{Peixoto-MCMC},
+which relies on writing the posterior in the following form:
 %
 \begin{equation}
-	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) = \pi_b(b)
+	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) = \pi_b(b).
 \end{equation}
 %
-Now $\pi_b(\cdot)$ is the un-normalised density we wish to sample from for the $b$-chain.
-In other words, we wish to construct a Markov chain that has $\pi_b(\cdot)$ as its invariant distribution.
-We can break $\pi_b$ down as follows:
+Now $\pi_b(\cdot)$ is the un-normalised target density,
+which can be expressed as:
+%  we wish to sample from for the $b$-chain. In other words, we 
+% wish to construct a Markov chain that has $\pi_b(\cdot)$ 
+% as its invariant distribution. We can break $\pi_b$ down as follows:
 %
@@
 	= p(b|X) p(A, \psi^* | b, X) \\
-	= p(A | b, \psi^*) \cdot p(\psi^* | b) \cdot p(b | X)
+	= p(A | b, \psi^*) \cdot p(\psi^* | b) \cdot p(b | X).
 \end{equation}
 %
-Since we are using the microcanonical SBM formulation, there is only one value of $\psi$ that is compatible with the given $(A, b)$ pair (given in equation \ref{eqn:sbm-constraints}).
-We denote this value $\psi^* = \{\psi_k^*, \psi_e^*\}$.
-Therefore, the summation over all $\psi$ reduces to just the single $\psi^*$ term; this is the power of the microcanonical formulation.
-We also define the microcanonical entropy of the configuration as.
+Since we are using the microcanonical SBM formulation, there is only one 
+value of $\psi$ that is compatible with the given $(A, b)$ pair;
+recall the constraints in~(\ref{eqn:sbm-constraints}). 
+We denote this value $\psi^* = \{\psi_k^*, \psi_e^*\}$. Therefore, 
+the summation over all $\psi$ reduces to just the single $\psi^*$ term.
+% ; this is the power of the microcanonical formulation. 
+We also define the microcanonical entropy of the configuration as,
 %
 \begin{equation}
-	S(b) = - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big)
+	S(b) = - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big).
 	\label{eqn:dl-form}
@@
 %
-This entropy can equally be thought of as the description length of the graph.
-The exact form of the proposal $q_b$ is explored thoroughly by \citet{Peixoto-MCMC} and not repeated here.
-There is a widely used library for Python made available under LGPL called \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, which implements this algorithm.
-The only modification we make is in the block membership prior $p(b)$ which we replace with $p(b|X)=B^{-N}$, which cancels out in the MH accept-reject step as it is independent of $b$.
+This entropy can be thought of as the optimal
+``description length'' of the graph. 
+The exact form of the proposal $q_b$ is explored thoroughly in
+\cite{Peixoto-MCMC} and not repeated here. There is a widely used 
+library for Python made available under LGPL 
+called \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, 
+which implements this algorithm. The only modification we make is in 
+the block membership prior $p(b)$ that we replace with $p(b|X)=B^{-N}$, 
+which cancels out in the MH accept-reject step as it is independent of $b$.
 
 \subsection{Sampling feature-to-block generator parameters}
+\label{s:sfb}
 
-The invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. We write this as follows:
+The invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. 
+We write this as,
 %
 \begin{equation}
-	\pi_\theta(\theta) \propto p(\theta | X, b) \propto p(b | X, \theta) p(\theta) \propto \pi_\theta (\theta) \propto  \exp \left( - U(\theta) \right)
+	\pi_\theta(\theta) \propto p(\theta | X, b) \propto p(b | X, \theta) p(\theta) \propto  \exp \left( - U(\theta) \right),
+	\label{eq:U}
 \end{equation}
 %
-Where we have introduced $U(\theta)$ equal to the negative log posterior.
-We define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$.
-Discarding constant terms, we can write $U(\theta)$ as in equation \ref{eqn:U-form} (refer to appendix \ref{appdx:form-U} for the derivation).
+where we write $U(\theta)$ for the negative log-posterior. Let $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$. 
+Discarding constant terms, we can write $U(\theta)$ as,
 %
 \begin{equation}
-	U(\theta) = \left( \sum_{i=1}^{N} \sum_{j=1}^{B} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} ||\theta||^2 = N \cdot \Lcal(\theta) + \frac{1}{2\sigma_\theta^2} ||\theta||^2
+	U(\theta) = \left( \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log \frac{1}{a_{ij}} \right)
+	+ \frac{1}{2\sigma_\theta^2} \|\theta\|^2 = N \cdot \Lcal(\theta) + \frac{1}{2\sigma_\theta^2} \|\theta\|^2;
 	\label{eqn:U-form}
 \end{equation}
-%
-$U(\theta)$ in equation \ref{eqn:U-form} appears a typical objective function for neural network training.
+
+See Appendix \ref{appdx:form-U} for the derivation.
+The function $U(\theta)$ is a typical objective function for neural network training.
 The first term is introduced by the likelihood.
 We collect it into $N \cdot \Lcal(\theta)$, which is the cross-entropy between the graph-predicted and feature-predicted block memberships summed over all vertices.
-The second term of equation \ref{eqn:U-form} -- introduced by the prior -- brings a form of regularisation, guarding against over-fitting.
-Different to traditional applications, our goal is not to find the minimiser of $U(\theta)$ but to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
+The second, introduced by the prior, brings a form of regularisation, guarding against over-fitting. Unlike in many applications where the goal 
+is to find the minimiser of $U(\theta)$, our goal is to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
 We can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
-We therefore adopt the Metropolis-adjusted Langevin algorithm (MALA) -- first proposed by \citet{mala-tweedie}.
-Given the current sample $\theta$, we generate a new sample $\theta'$ according to equation \ref{eqn:theta-update}.
-%
-\begin{align}
-	\theta' &= \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi
-	\label{eqn:theta-update} \\
-	\therefore q_\theta(\theta, \theta') &= \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right)
-	\label{eqn:theta-proposal}
-\end{align}
-%
-Where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter -- which may vary with the sample index (appendix \ref{appdx:step-size} explores this more fully).
-Without the injected noise term, MALA is equivalent to gradient descent.
-We require the noise term $\xi$ to fully explore the parameter space.
-We can write the proposal distribution $q_\theta$ as in equation \ref{eqn:theta-proposal}.
-The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}).
-By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as:
+We therefore adopt the Metropolis-adjusted Langevin algorithm (MALA) \cite{mala-tweedie}.
+Given the current sample $\theta$, we generate
+a new sample $\theta'$ as,
+\begin{equation*}
+	\theta' = \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi,
+\end{equation*}
+
+where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter 
+which may vary with the sample index (see Appendix \ref{appdx:step-size}).
+This leads to the proposal distribution,
+\begin{equation*}
+	q_\theta(\theta, \theta') \sim \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right).
+\end{equation*}
+
+Without the injected noise term, MALA is equivalent to gradient descent. We require the noise term $\xi$ to fully explore the parameter space. 
+The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}). By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as,
 %
 \begin{equation}
-	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right)
+	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right).
 	\label{eqn:U-derivative}
@@
 	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
-	\hat{\sigma}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2
+	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2
 \end{equation}
 %
-A simple heuristic to discard the least important features requires specifying a cutoff $c > 0$ and a multiplier $k > 0$.
-We define the function $\Fcal_i(j)$ as in \ref{eqn:fij} then only keep features with indices $d \in \Dcal'$, where $\Dcal'$ is constructed as in equation \ref{eqn:kept-feature-set}.
+A simple heuristic to discard the least important features requires specifying a cutoff $c > 0$ and a multiplier $k > 0$. We define the function $\Fcal_i(j)$ 
+as in~(\ref{eqn:fij}) and only keep features with indices $d \in \Dcal'$, where $\Dcal'$ is given in~(\ref{eqn:kept-feature-set}).
 %
 \begin{align}
-	\Fcal_i(j) &\coloneqq (\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c)
+	\Fcal_i(j) &\coloneqq (\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c),
 	\label{eqn:fij} \\
-	\Dcal' &\coloneqq \left\{ j \in [D] : \exists i \in [B] \textrm{ s.t. }  \Fcal_i(j) \neq \emptyset \right\}
+	\Dcal' &\coloneqq \left\{ j \in [D] : \exists i \in [B] \textrm{ s.t. }  \Fcal_i(j) = \emptyset \right\}
 	\label{eqn:kept-feature-set}
@@
 %
-Intuitively, this means discarding any feature for which $\hat{\mu}_{ij} \pm k\hat{\sigma}_{ij}$ lies within or spans the null region $(-c, c)$ for all block indices.
-If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \mu_{ij}, \sigma_{ij})$, then this is effectively a hypothesis test on the value of $W_{ij}$ (equation \ref{eqn:hyp-test-discard}).
-$\Dcal'$ then comprises all features $i$ for which $H_1$ is accepted at least once for some $j \in [B]$.
+Intuitively, this means discarding any feature $j$ for which 
+$\hat{\mu}_{ij} \pm k\hat{\sigma}_{ij}$ lies outside the region
+$(-c, c)$ for all block indices $i$.
+If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \hat{\mu}_{ij}, \hat{\sigma}_{ij}^2)$, then this is analogous to a hypothesis test on the value of $W_{ij}$ as in~(\ref{eqn:hyp-test-discard});
+then $\Dcal'$ comprises all features $i$ for which $H_1$ is not rejected at least once for some $j \in [B]$
 %
 \begin{equation}
-	H_0: |\mu_{ij}| \leq c \qquad
-	H_1: |\mu_{ij}| > c
+	H_0: |W_{ij}| \leq c \qquad
+	H_1: |W_{ij}| > c
 	\label{eqn:hyp-test-discard}
@@
 %
-The multiplier $k$ determines the degree of significance of the result.
+The multiplier $k$ in equation \ref{eqn:fij} determines the degree of significance of the result.
 However, as the Laplace approximation is not exact we will only treat this dimensionality reduction method as a useful heuristic and not an exact method.
@@
 This is summarised in equation \ref{eqn:c-star}.
-This approach is often preferred as it fixes the number of reduced dimensions.
 %
 \begin{equation}
-	c^* = \argmax_{c>0} (c : |\Dcal'| = D', k=k_0)
+	c^* = \argmax \{c>0\; : \;|\Dcal'| = D', k=k_0\}.
 	\label{eqn:c-star}
 \end{equation}
+%
+For an algorithmic implementation of this method refer to appendix \ref{appdx:algorithms}. This approach has the advantage that it fixes the number of reduced dimensions.

### 2021-05-28 code/blogs.py
Remove code for submodule replacement

@@ -1,24 +1,0 @@
-from data.polblogs import PolBlog
-from model.graph_mcmc import Graph_MCMC
-
-
-def run():
-    graph = Graph_MCMC()
-    graph.read_from_file("polblogs.gml")
-    graph.filter_out_low_degree(2)
-    graph.remove_property("label")
-    graph.remove_property("source")
-    graph.partition(B_min=1, B_max=2)
-    graph.mcmc(100, verbose=True)
-
-    graph.draw("polblogs.png")
-    classifier = graph.sample_classifier_mala(10000, step_scaling=0.01, verbose=True)
-
-    classifier.thin_samples()
-    classifier.plot_sampled_weights(["Right-Wingness"])
-    classifier.plot_sample_histogram()
-    classifier.plot_sample_history()
-
-if __name__ == "__main__":
-    print("Analysing political blogs")
-    run()

### 2021-05-28 code/classifier.py

@@ -1,23 +1,0 @@
-from inference.softmax import SoftmaxNeuralNet
-import numpy as np
-
-
-def run():
-    B = 2
-    D = 4
-    classifier = SoftmaxNeuralNet([D, B], sigma=100)
-    X = np.array([[1, 0, 0, 0], [0, 1, 0, 0]])
-    Y = np.array([[0.8, 0.2], [0.2, 0.8]])
-
-    #classifier.fit(X, Y, verbose=True)
-    classifier.sgld_initialise()
-
-    for n in range(0, 1000):
-        cost = classifier.sgld_iterate(X, Y)
-        if n % 100 == 0:
-            print("cost:{}".format(cost))
-
-
-if __name__ == "__main__":
-    print("Testing softmax class")
-    run()

### 2021-05-28 code/data/aminer.py

@@ -1,113 +1,0 @@
-import csv
-import os
-from data.names import NameList
-from hypothesis.test_statistics import two_samples_mean_ll_ratio, students_z_test
-
-
-curr_dir = os.path.dirname(__file__)
-aminer_dir = os.path.join(curr_dir, "aminer")
-
-
-class AMiner:
-
-    def __init__(self, raw = True):
-        if raw:
-            self.gender_by_author_id = self._read_author_genders_raw()
-        else:
-            self.gender_by_author_id = self._read_genders_quick()
-
-        M = F = 0
-        for gender in self.gender_by_author_id.values():
-            if gender == "M":
-                M += 1
-            elif gender == "F":
-                F += 1
-
-        self.M = M
-        self.F = F
-
-
-    def _read_author_genders_raw(self):
-        name_list = NameList()
-        filename = "AMiner-Author.txt"
-        filepath = os.path.join(aminer_dir, filename)
-
-        outfilename = "AMiner-Author-Gender.txt"
-        outfilepath = os.path.join(aminer_dir, outfilename)
-        outfile = open(outfilepath, "w")
-
-        author_id = 0
-        gender_by_author_id = {}
-
-        with open(filepath) as fp:
-            for line in fp:
-                if line.startswith("#index"):
-                    author_id = int(line[7:])
-                elif line.startswith("#n"):
-                    full_name = line[3:]
-                    first_name = full_name.split(" ")[0]
-                    
-                    gender = "U"
-                    if "." not in first_name:
-                        gender = name_list.get_gender_of(first_name)
-
-
-                    gender_by_author_id[author_id] = gender
-                    outfile.write(str(author_id) + "," + gender + "\n")
-
-        outfile.close()
-        return gender_by_author_id
-
-
-    def _read_genders_quick(self):
-        filename = "AMiner-Author-Gender.txt"
-        filepath = os.path.join(aminer_dir, filename)
-
-        gender_by_author_id = {}
-        with open(filepath) as fp:
-            reader = csv.reader(fp, delimiter=",")
-            for row in reader:
-                author_id = int(row[0])
-                gender = row[1]
-                gender_by_author_id[author_id] = gender
-
-        return gender_by_author_id
-
-    def check_gender_disparity(self):
[…]

### 2021-05-28 code/data/facebook.py

@@ -1,405 +1,0 @@
-import os
-import csv
-from hypothesis.test_statistics import two_samples_mean_ll_ratio, students_z_test
-from distributions.sigmoid import sigmoid
-import numpy as np
-from tqdm import tqdm
-import matplotlib.pyplot as plt
-from utils.colors import color_between
-import pandas as pd
-from sklearn import linear_model
-from utils.plotting import sorted_bar_plot
-
-
-curr_dir = os.path.dirname(__file__)
-facebook_dir = os.path.join(curr_dir, "facebook")
-
-
-class FacebookGraph:
-    """
-    class for reading and manipulating facebook graphs
-    """
-    
-
-    def __init__(self, ego_id, significance_level=99):
-        self.ego_id = str(ego_id)
-        self.edges = self.read_edges()
-        self.feature_names = self.read_feature_names()
-        self.node_features = self.read_node_features()
-        self.significance_level = significance_level
-
-        # add edges from ego node
-        for node_id in self.node_features.keys():
-            self.edges.append((self.ego_id, node_id))
-
-        # add ego features
-        self.node_features[self.ego_id] = self.read_ego_features()
-
-        # constants
-        self.N = len(self.node_features)
-        self.M = len(self.edges)
-        self.D = len(self.feature_names)
-
-    
-    def read_graph_file(self, extension, delimiter, transform_function):
-        filepath = os.path.join(facebook_dir, self.ego_id + extension)
-        with open(filepath) as csvfile:
-            reader = csv.reader(csvfile, delimiter=delimiter)
-            rows = [transform_function(row) for row in reader]
-        return rows
-
-
-    def read_edges(self):
-        extension = ".edges"
-        delimiter = " "
-        edge_transform = lambda edge: (edge[0], edge[1])
-        return self.read_graph_file(extension, delimiter, edge_transform)
-
-
-    def read_feature_names(self):
-        extension = ".featnames"
-        delimiter = " "
-        feature_transform = lambda feature_row: (feature_row[0], self.convert_feature_name("_".join(feature_row[1:])))
-        features_arr = self.read_graph_file(extension, delimiter, feature_transform)
-        return {int(feature[0]): feature[1] for feature in features_arr}
-
-
-    def read_ego_features(self):
-        extension = ".egofeat"
-        delimiter = " "
-
-        boolean_arr_to_set = lambda boolean_arr: {idx for idx, feature_set in enumerate(boolean_arr) if feature_set == "1"}
[…]

### 2021-05-28 code/data/names.py

@@ -1,41 +1,0 @@
-import csv
-import os
-
-
-curr_dir = os.path.dirname(__file__)
-names_dir = os.path.join(curr_dir, "names")
-
-
-class NameList:
-
-    def __init__(self, year=2018):
-        filename = "yob" + str(year) + ".txt"
-        filepath = os.path.join(names_dir, filename)
-        
-        female_names = {}
-        male_names = {}
-
-        with open(filepath) as csvfile:
-            reader = csv.reader(csvfile, delimiter=",")
-            for row in reader:
-                [name, gender, count] = row
-                if gender == "F":
-                    female_names[name] = int(count)
-                elif gender == "M":
-                    male_names[name] = int(count)
-
-        self.female_names = female_names
-        self.male_names = male_names
-
-
-    def get_gender_of(self, name):
-        ratio = 10
-        female_count = self.female_names.get(name, 0)
-        male_count = self.male_names.get(name, 0)
-
-        if female_count > male_count * ratio:
-            return "F"
-        elif male_count > female_count * ratio:
-            return "M"
-        else:
-            return "U"

### 2021-05-28 code/data/polblogs.py

@@ -1,68 +1,0 @@
-import networkx as nx
-import matplotlib.pyplot as plt
-from sklearn.cluster import KMeans
-import os
-
-curr_dir = os.path.dirname(__file__)
-pol_dir = os.path.join(curr_dir, "polblogs")
-
-class PolBlog:
-
-    def __init__(self):
-        print("Reading in PolBlog data...")
-        filepath = os.path.join(pol_dir, "polblogs.gml")
-        self.graph = nx.read_gml(filepath)
-        self.graph = self.graph.to_undirected()
-        self.filter_out_low_degree()
-        self.pos = self.fruchterman_reingold()
-        print("Done")
-
-
-    def filter_out_low_degree(self, min_degree=2):
-        low_degree_nodes = [node for node, degree in dict(self.graph.degree).items() if degree < min_degree]
-        self.graph.remove_nodes_from(low_degree_nodes)
-
-
-    def political_partition(self):
-        politics_arr = []
-        for node_label, attributes in self.graph.nodes.items():
-            politics_arr.append(attributes["value"])
-
-        return politics_arr
-
-    
-    def fruchterman_reingold(self):
-        return nx.spring_layout(self.graph)
-
-
-    def k_means_centres(self, k=2):
-        print("Performing k-means...")
-        # nodes_by_degree_desc = sorted(self.graph.degree, key = lambda x: x[1], reverse=True)
-        node_pos_arr = list(self.pos.values())
-        kmeans = KMeans(n_clusters=k).fit(node_pos_arr) # can use flag random_state=0 to intialise seed
-        return kmeans.cluster_centers_
-
-    
-    def get_edge_list(self):
-        return self.graph.edges
-
-
-    def draw(self):
-        print("Drawing network...")
-        politics_arr = self.political_partition()
-        colors = ["blue" if value == 0 else "red" for value in politics_arr]
-        cluster_centres = self.k_means_centres()
-
-        nx.draw_networkx(self.graph, pos=self.pos, with_labels=False, node_size=5, width=0.1, node_color=colors)
-
-        plt.title("Political blogs (2004 US Election)")
-
-        # labels
-        plt.scatter([], [], c="blue", label="Democrat")
-        plt.scatter([], [], c="red", label="Republican")
-
-        for centre in cluster_centres:
-            plt.scatter(centre[0], centre[1], c="yellow", s=100, marker="x", zorder=100)
-
-        plt.legend()
-        plt.show()

### 2021-05-28 code/data/read.py

@@ -1,18 +1,0 @@
-import os
-import csv
-
-
-curr_dir = os.path.dirname(__file__)
-
-
-def read_file(filename):
-    if filename.endswith(".csv"):
-        return read_csv(filename)
-
-
-def read_csv(filename, transform_function=lambda x: x):
-    filepath = os.path.join(curr_dir, filename)
-    with open(filepath) as csvfile:
-        reader = csv.reader(csvfile)
-        rows = [transform_function(row) for row in reader]
-    return rows

### 2021-05-28 code/data/twitch.py

@@ -1,66 +1,0 @@
-import os
-import csv
-import json
-from model.graph_mcmc import Graph_MCMC
-
-
-curr_dir = os.path.dirname(__file__)
-twitch_dir = os.path.join(curr_dir, "twitch")
-
-
-class TwitchGraph:
-    """
-    class for reading and manipulating twitch graphs
-    """
-    
-
-    def __init__(self, country="ENGB"):
-        self.country = country
-        self.edge_filename = "musae_" + country + "_edges.csv"
-        self.feature_filename = "musae_" + country + "_features.json"
-
-        self.edges = self.read_edges()
-        self.node_features = self.read_node_features()
-
-    def read_edges(self):
-        delimiter = ","
-        edge_transform = lambda edge: (edge[0], edge[1])
-        filepath = os.path.join(twitch_dir, self.country, self.edge_filename)
-        
-        with open(filepath) as csvfile:
-            reader = csv.reader(csvfile, delimiter=delimiter)
-            reader.__next__() # discard header
-            rows = [edge_transform(row) for row in reader]
-        return rows
-
-    def read_node_features(self):
-        filepath = os.path.join(twitch_dir, self.country, self.feature_filename)
-
-        with open(filepath) as jsonfile:
-            features = json.load(jsonfile)
-
-        features = {key: val for key, val in features.items()}
-        return features            
-
-    
-    def generate_mcmc_graph(self):
-        graph = Graph_MCMC()
-        graph.read_from_edges(self.edges)
-
-        feature_set = set()
-
-        for features in self.node_features.values():
-            feature_set = feature_set.union(features)
-
-        vertices = graph.get_vertex_list()
-
-        for feat in feature_set:
-            readable_name = "feat-" + str(feat) 
-            value_arr = [feat in self.node_features[vertex] for vertex in vertices]
-            graph.add_property(readable_name, "bool", value_arr)
-
-        return graph
-        
-
-
-

### 2021-05-28 code/data/utils.py

@@ -1,9 +1,0 @@
-import os
-
-
-curr_dir = os.path.dirname(__file__)
-misc = os.path.join(curr_dir, "misc")
-
-
-def get_misc_path(filename):
-    return os.path.join(misc, filename)

### 2021-05-28 code/distributions/chi_squared.py

@@ -1,9 +1,0 @@
-from scipy.stats import chi2
-
-
-def pdf(x, dof=1):
-    return chi2.pdf(x, dof)
-
-
-def cdf(x, dof=1):
-    return chi2.cdf(x, dof)

### 2021-05-28 code/distributions/gaussian.py

@@ -1,12 +1,0 @@
-import math
-from scipy.stats import norm
-
-
-def pdf(x, mu=0, sigma=1):
-    """pdf for a gaussian"""
-    return norm.pdf(x, loc=mu, scale=sigma)
-
-
-def cdf(x, mu=0, sigma=1):
-    """cdf for a gaussian"""
-    return norm.cdf(x, loc=mu, scale=sigma)

### 2021-05-28 code/equality_of_means.py

@@ -1,40 +1,0 @@
-from data.read import read_csv
-from model.player import Player
-from typing import List
-from hypothesis.test_statistics import two_samples_mean_ll_ratio, students_z_test
-
-significance_level = 95
-
-def run():
-    t, p = two_samples_mean_ll_ratio(6, 9, 5, 3, debug=True)
-    print(t)
-    print(p)
-    z, p = students_z_test(6, 9, 5, 3, debug=True)
-    print
-
-
-    players: List[Player] = read_csv("training_attendance.csv", Player)
-    n = m = k = l = 0
-
-    for player in players:
-        if player.isFirstTeam():
-            k += player.attendances
-            n += player.attendances + player.absences
-        elif player.isSecondTeam():
-            l += player.attendances
-            m += player.attendances + player.absences
-
-    t, p = two_samples_mean_ll_ratio(n, m, k, l, debug=True)
-    z, p = students_z_test(n, m, k, l, debug=True)
-
-    if (100*p < (100 - significance_level)):
-        print("Null Hypothesis rejected at the {}% significance level".format(significance_level))
-        print("Means are different")
-    else:
-        print("Insufficient evidence to reject null")
-
-    
-
-if __name__ == "__main__":
-    print("Testing equality of means")
-    run()

### 2021-05-28 code/football.py

@@ -1,33 +1,0 @@
-from model.graph_mcmc import Graph_MCMC
-
-
-def run():
-    #dataset = "escorts"
-    dataset = "football_tsevans"
-
-    graph = Graph_MCMC()
-    #graph.read_from_gt("cond-mat")
-    graph.read_from_ns(dataset)
-    graph.remove_property("_graphml_vertex_id")
-    graph.remove_property("_pos")
-    graph.remove_property("label")
-    #graph.convert_props_to_flags()
-    graph.convert_to_flags("value", "conf-")
-
-    graph.partition(B_min=4, B_max=20)
-    graph.mcmc(1000)
-    graph.draw(dataset + ".png")
-
-    classifier = graph.sample_classifier_marginals(2500, step_scaling=0.01)
-
-    names = graph.get_feature_names()
-
-    classifier.thin_samples()
-    classifier.plot_sampled_weights(names, std_dev_multiplier=2)
-    classifier.plot_sample_histogram()
-    classifier.plot_sample_history()
-
-
-if __name__ == "__main__":
-    print("Analysing chosen dataset")
-    run()

### 2021-05-28 code/generated.py

@@ -1,26 +1,0 @@
-from model.graph_mcmc import Graph_MCMC
-
-def run():
-    graph = Graph_MCMC()
-    graph.read_from_file("generated.gml")
-
-    feature_names = graph.get_feature_names()
-    graph.partition(B_min=2, B_max=5)
-
-    # ml_classifier = graph.train_map_classifier()
-    # ml_classifier.plot_final_weights(feature_names)
-
-    # mcmc_classifier = graph.sample_classifier_mcmc(100, True)
-    # mcmc_classifier.plot_sampled_weights(feature_names)
-
-    _B_max = graph.mcmc(10, verbose=True)
-    marginal_classifier = graph.sample_classifier_marginals(2500, step_scaling=0.001, sigma=1, verbose=True)
-    marginal_classifier.plot_sampled_weights(feature_names)
-    marginal_classifier.plot_sample_histogram()
-    marginal_classifier.plot_sample_history()
-
-
-
-if __name__ == "__main__":
-    print("Analysing simulated data")
-    run()

### 2021-05-28 code/generation/sbm.py

@@ -1,102 +1,0 @@
-from graph_tool import Graph
-from graph_tool.generation import generate_maxent_sbm
-from graph_tool.all import graph_draw
-from graph_tool.inference import minimize_blockmodel_dl
-import numpy as np
-
-
-class SBM:
-
-    def __init__(self, n, B, prior, W):
-        """
-        Draw new SBM with:
-            n - vertices
-            prior - B length array denoting fraction in each community
-            W - BxB connectivity matrix
-        """
-
-        assert B == len(prior)
-        num_vertices_by_block = np.random.multinomial(n, prior)
-
-        vertex_block_labels = []
-        for idx, num_in_block in enumerate(num_vertices_by_block):
-            new_block_labels = [idx] * num_in_block # repeats num_in_block times [idx, idx .... idx]
-            vertex_block_labels.extend(new_block_labels)
-
-        out_fugacities = [1] * n
-
-        self.n = n
-        self.B = B
-        self.num_vertices_by_block = num_vertices_by_block
-        self.true_block_labels = vertex_block_labels
-        self.graph = generate_maxent_sbm(vertex_block_labels, W, out_fugacities)
-
-    
-    def add_property(self, name, prob_by_block):
-        """
-        Add property to node
-            name: string - property name
-            prob_by_block: array - B length array with probability feature turned on for each block
-        """
-
-        value_sequence = []
-
-        for b in range(0, self.B):
-            num_vertices = self.num_vertices_by_block[b]
-            prob = prob_by_block[b]
-            num_successes = np.random.binomial(num_vertices, prob)
-            value_sequence.extend([1] * num_successes)
-            value_sequence.extend([0] * (num_vertices - num_successes))
-
-        vertex_prop = self.graph.new_vertex_property("int", value_sequence)
-        self.graph.vertex_properties[name] = vertex_prop # add to graph
-
-    
-    def partition(self, B_min=None, B_max=None, degree_corrected=True):
-        """
-        Performs MCMC algorithm to minimise description length (DL)
-        returns partition array
-        """
-        print("Performing inference...")
-        self.state = minimize_blockmodel_dl(self.graph, B_min=B_min, B_max=B_max, deg_corr=degree_corrected, verbose=True)
-        print("Done")
-        return self.state.get_blocks()
-
-
-    def draw(self, output=None):
-        output = self.gen_output_path(output)
[…]

### 2021-05-28 code/hypothesis/test_statistics.py

@@ -1,55 +1,0 @@
-"""
-great website: https://machinelearningmastery.com/statistical-hypothesis-tests-in-python-cheat-sheet/ 
-"""
-import math
-from distributions import chi_squared, gaussian
-
-
-def bern_kl_divergence(p, q):
-    """
-    Returns the Kullback-Leibler Divergence between
-    2 Bernoulli distributions with parameters p and q
-    D(Bern(p) || Bern(q))
-    """
-    offset = 1E-10
-    p += offset
-    q += offset
-    return p * math.log(p / q) + (1-p) * math.log((1-p) / (1-q))
-
-
-def two_samples_mean_ll_ratio(n, m, k, l, debug=False):
-    p_hat = k / float(n)
-    q_hat = l / float(m)
-    r_hat = (k + l) / float(n + m)
-
-    t = n * bern_kl_divergence(p_hat, r_hat) + m * \
-        bern_kl_divergence(q_hat, r_hat)
-    p = 1 - chi_squared.cdf(2*t, 1) # 2*t ~ chi_1^2 
-
-    if debug:
-        print("Population sizes:\n n = {} , m = {}\n".format(n, m))
-        print("Sampled mean of each:\n p^ = {:.3f} , q^ = {:.3f}\n".format(p_hat, q_hat))
-        print("Joint mean:\n r^ = {:.3f}\n".format(r_hat))
-        print("t-statistic: t = {:.3f}".format(t))
-        print("p-value: p = {:.7f}\n".format(p))
-
-    return t, p
-
-
-def students_z_test(n, m, k, l, debug=False):
-    p_hat = k / float(n)
-    q_hat = l / float(m)
-    r_hat = (k + l) / float(n + m)
-
-    var = r_hat * (1 - r_hat) * (1 / float(n) + 1 / float(m))
-    z = (p_hat - q_hat) / math.sqrt(var)
-    p = 2 * (1 - gaussian.cdf(abs(z))) 
-
-    if debug:
-        print("Population sizes:\n n = {} , m = {}\n".format(n, m))
-        print("Sampled mean of each:\n p^ = {:.3f} , q^ = {:.3f}\n".format(p_hat, q_hat))
-        print("Joint mean:\n r^ = {:.3f}\n".format(r_hat))
-        print("z-statistic: z = {:.3f}".format(z))
-        print("p-value: p = {:.7f}\n".format(p))
-
-    return z, p

### 2021-05-28 code/inference/mnist_clothes.R

@@ -1,109 +1,0 @@
-# example from:
-# https://tensorflow.rstudio.com/tutorials/beginners/basic-ml/tutorial_basic_classification/
-
-library(keras)
-library(tfdatasets)
-library(dplyr)
-
-## % means binary operations
-
-print("Downloading dataset...")
-fashion_mnist <- dataset_fashion_mnist()
-
-c(train_images, train_labels) %<-% fashion_mnist$train
-c(test_images, test_labels) %<-% fashion_mnist$test
-
-class_names <- c(
-    "T-shirt/top",
-    "Trouser",
-    "Pullover",
-    "Dress",
-    "Coat",
-    "Sandal",
-    "Shirt",
-    "Sneaker",
-    "Bag",
-    "Ankle boot"
-)
-
-print("Dimensions: ")
-print(dim(train_images))
-
-library(tidyr)
-library(ggplot2)
-
-image_1 <- as.data.frame(train_images[1, , ])
-colnames(image_1) <- seq_len(ncol(image_1))
-image_1$y <- seq_len(nrow(image_1))
-image_1 <- gather(image_1, "x", "value", -y)
-image_1$x <- as.integer(image_1$x)
-
-# scale down from 0-255 to 0-1
-train_images <- train_images / 255
-test_images <- test_images / 255
-
-print("Plotting first 25 training images...")
-par(mfcol = c(5, 5))
-par(mar = c(0, 0, 1.5, 0), xaxs = "i", yaxs = "i")
-for (i in 1:25) {
-    img <- train_images[i, , ]
-    img <- t(apply(img, 2, rev))
-    image(1:28, 1:28, img,
-        col = gray((0:255) / 255), xaxt = "n", yaxt = "n",
-        main = paste(class_names[train_labels[i] + 1])
-    )
-}
-
-# %foo% is R syntax for custom binary operators
-# in dplyr package %>% is used as a pipe >> chains function calls
-
-# define model
-model <- keras_model_sequential()
-model %>%
-    layer_flatten(input_shape = c(28, 28)) %>%
-    layer_dense(units = 128, activation = "relu") %>%
-    layer_dense(units = 10, activation = "softmax")
-
-# compile model
-model %>% compile(
-    optimizer = "adam",
-    loss = "sparse_categorical_crossentropy",
-    metrics = c("accuracy")
-)
-
-# train
-model %>% fit(train_images, train_labels, epochs = 5, verbose = 2)
-
-# test metrics
[…]

### 2021-05-28 code/inference/softmax.py

@@ -1,743 +1,0 @@
-# adapted from http://www.adeveloperdiary.com/data-science/deep-learning/neural-network-with-softmax-in-python/
-
-import numpy as np
-import matplotlib.pyplot as plt
-from numpy.core.shape_base import block
-from sklearn.preprocessing import OneHotEncoder
-from scipy.stats import norm
-import math
-from inference.store import Store, compute_log_acceptance_prob
-from tqdm import tqdm
-from utils.colors import plt_color
-import matplotlib.mlab as mlab
-
-
-class SoftmaxNeuralNet:
-
-    def __init__(self, layers_size, sigma=1, a=250, b=1000, gamma=0.8):
-        """Initialise Neural Network"""
-        self.layers_size = layers_size
-        self.parameters = Store()
-        self.L = len(self.layers_size) - 1  # number of activation layers
-        self.n = 0
-        self.costs = []
-        self.sigma = 1
-        self.a = a
-        self.b = b
-        self.gamma = gamma
-        self._initialize_parameters()
-        self.store_history = []
-
-    def _sigmoid(self, Z):
-        return 1 / (1 + np.exp(-Z))
-
-    def _softmax(self, Z):
-        # keep numerically stable / avoid overflow
-        expZ = np.exp(Z - np.max(Z, axis=0, keepdims=True))
-        expZ = np.nan_to_num(expZ)
-        return expZ / expZ.sum(axis=0, keepdims=True)
-
-    def _initialize_parameters(self):
-        for l in range(1, len(self.layers_size)):
-            W = np.random.randn(
-                self.layers_size[l], self.layers_size[l - 1]) / np.sqrt(self.layers_size[l - 1])
-
-            self.parameters.set_W(W, l)
-
-    def _forward(self, X, store):
-        A = X.T
-        store.set_A(A, 0)
-        # hidden sigmoid layers
-        for l in range(1, self.L + 1):
-            Z = store.get_W(l).dot(A)
-
-            if l < self.L:
-                A = self._sigmoid(Z)
-            else:
-                A = self._softmax(Z)  # final softmax layer
-
-            store.set_A(A, l)
-            store.set_Z(Z, l)
-
-        return A
-
-    def _sigmoid_derivative(self, Z):
-        s = 1 / (1 + np.exp(-Z))
-        return s * (1 - s)
-
-    def _backward_cross_entropy_deriv(self, X, Y, store):
-        """performs backwards algorithm on store object in place
-
-            Parameters:
-                X: input matrix for training
-                Y: training target distribution
[…]

### 2021-05-28 code/inference/store.py

@@ -1,135 +1,0 @@
-import numpy as np
-from scipy.stats import norm
-
-
-class Store:
-
-    def __init__(self, W={}, dW={}, A={}, Z={}, U=None):
-        self.W = W
-        self.dW = dW
-        self.A = A
-        self.Z = Z
-        self.U = U
-
-    def shallow_copy(self):
-        """Creates a new store object with W, b and U copied through"""
-        new_store = Store(self.W.copy())
-        new_store.set_U(self.U)
-        return new_store
-
-    def full_copy(self):
-        """Creates a new store object with all props copied through"""
-        new_store = Store(
-            self.W.copy(), self.dW.copy(),
-            self.A.copy(), self.Z.copy(),
-            self.U
-        )
-        return new_store
-
-    # weight matrices
-
-    def get_W(self, l):
-        return self.W[l]
-
-    def set_W(self, W_instance, l):
-        self.W[l] = W_instance
-
-    # weight derivatives
-    def get_dW(self, l):
-        return self.dW[l]
-
-    def set_dW(self, dW_instance, l):
-        self.dW[l] = dW_instance
-
-    # activation output vector
-    def get_A(self, l):
-        return self.A[l]
-
-    def set_A(self, A_instance, l):
-        self.A[l] = A_instance
-
-    # activation input vector
-    def get_Z(self, l):
-        return self.Z[l]
-
-    def set_Z(self, Z_instance, l):
-        self.Z[l] = Z_instance
-
-    # -ve log-target
-    def get_U(self):
-        return self.U
-
-    def set_U(self, U):
-        self.U = U
-
-    # utility functions
-    def descend_gradient(self, step_size):
-        """
-        W and b params updated once in direction of decreasing gradient
-
-            Params:
-                step_size (float): step_size multiplier
-
-            Returns:
-                None: results stored in place
-        """
-
-        for l in self.W.keys():
-            self.W[l] = self.W[l] - step_size * self.dW[l]
-
-    def add_gaussian_noise(self, std_dev):
-        """
-        adds gaussian noise to each element of W and b
-
-            Params:
-                std_dev (float): standard deviation of noise
-
-            Returns:
-                None: results stored in place
-        """
-
-        for l in self.W.keys():
-            W = self.W[l]
-            self.W[l] = W + std_dev * np.random.randn(*W.shape)
-
-
-    def langevin_iterate(self, h):
-        """
[…]

### 2021-05-28 code/law.py

@@ -1,32 +1,0 @@
-from model.graph_mcmc import Graph_MCMC
-
-
-def run():
-    dataset = "law_firm"
-
-    graph = Graph_MCMC()
-    #graph.read_from_gt("cond-mat")
-    graph.read_from_ns(dataset)
-    graph.remove_property("_pos")
-    graph.remove_property("name")
-    graph.remove_property("nodeAge")
-    graph.convert_props_to_flags()
-    #graph.filter_edges("layer", 3) # 1: coworkers, 2: friendship, 3: advice
-
-    graph.partition(B_min=2, B_max=10)
-    graph.mcmc(1000)
-    graph.draw(dataset + ".png")
-
-    classifier = graph.sample_classifier_marginals(2500, step_scaling=0.01)
-
-    names = graph.get_feature_names()
-
-    classifier.thin_samples()
-    classifier.plot_sampled_weights(names, std_dev_multiplier=2)
-    classifier.plot_sample_histogram()
-    classifier.plot_sample_history()
-
-
-if __name__ == "__main__":
-    print("Analysing chosen dataset")
-    run()

### 2021-05-28 code/maier.py

@@ -1,39 +1,0 @@
-from model.graph_mcmc import Graph_MCMC
-
-
-def run():
-    graph = Graph_MCMC()
-    #graph.read_from_gt("cond-mat")
-    graph.read_from_ns("facebook_friends")
-    graph.remove_property("_graphml_vertex_id")
-    graph.remove_property("_pos")
-    graph.convert_props_to_flags()
-    graph.add_ego_node()
-
-    graph.partition(B_min=4, B_max=20)
-    graph.mcmc(1000, verbose=False)
-    graph.draw("maier-fb.png")
-
-    graph.plot_posterior_props()
-    names = graph.get_feature_names()
-
-    # map classifier
-    # map_classifier = graph.train_map_classifier()
-    # map_classifier.plot_final_weights(names)
-
-
-    # classifier = graph.sample_classifier_marginals(2500, step_scaling=0.001, verbose=True)
-    classifier = graph.sample_classifier_mala(2500, step_scaling=0.001, verbose=True)
-    classifier.thin_samples()
-
-
-    #classifier.plot_U()
-    #classifier.plot_final_weights(names)
-    classifier.plot_block_principal_dims(names.copy(), 1)
-    classifier.plot_sampled_weights(names.copy(), std_dev_multiplier=1, null_space=1)
-    #classifier.plot_sample_histogram()
-    #classifier.plot_sample_history()
-
-if __name__ == "__main__":
-    print("Analysing Maier facebook friends")
-    run()

### 2021-05-28 code/misc.py

@@ -1,29 +1,0 @@
-from model.graph_mcmc import Graph_MCMC
-
-
-def run():
-    dataset = "webkb/webkb_washington_cocite"
-
-    graph = Graph_MCMC()
-    graph.read_from_ns(dataset)
-    graph.remove_property("_pos")
-    graph.remove_property("name")
-    graph.convert_props_to_flags()
-
-    graph.partition(B_min=2, B_max=10)
-    graph.mcmc(1000)
-    graph.draw("webkb.png")
-
-    classifier = graph.sample_classifier_marginals(2500, step_scaling=0.01)
-
-    names = graph.get_feature_names()
-
-    classifier.thin_samples()
-    classifier.plot_sampled_weights(names, std_dev_multiplier=2)
-    classifier.plot_sample_histogram()
-    classifier.plot_sample_history()
-
-
-if __name__ == "__main__":
-    print("Analysing chosen dataset")
-    run()

### 2021-05-28 code/model/graph.py

@@ -1,214 +1,0 @@
-import numpy as np
-from tqdm import tqdm
-import math
-
-import networkx as nx
-import matplotlib.pyplot as plt
-from utils.colors import color_between
-
-
-class Graph:
-
-    def __init__(self, edges):
-
-        origin_vertices = set([edge[0] for edge in edges])
-        destin_vertices = set([edge[1] for edge in edges])
-
-        vertex_set = origin_vertices.union(destin_vertices)
-        
-        # constants
-        self.N = len(vertex_set)
-        self.M = len(edges)
-
-        # edges
-        self.edges_raw = edges
-
-        # relabel vertices with an index
-        self.index_to_vertex = {}
-        self.vertex_to_index = {}
-
-        for index, vertex in enumerate(vertex_set):
-            self.index_to_vertex[index] = vertex
-            self.vertex_to_index[vertex] = index
-
-        self.adjacency_list = [set() for v in range(0, self.N)]
-        self.adjacency_matrix = np.zeros((self.N, self.N))
-
-        for edge in edges:
-            a = self.vertex_to_index.get(edge[0])
-            b = self.vertex_to_index.get(edge[1])
-
-            self.adjacency_list[a].add(b)
-            self.adjacency_list[b].add(a)
-
-            self.adjacency_matrix[a, b] += 1
-            self.adjacency_matrix[b, a] += 1
-
-        self.assignments = None
-        print("Initialised graph with N={} nodes and M={} edges".format(self.N, self.M))
-
-
-    def abp(self, seed=None):
-        if seed is not None:
-            np.random.seed(seed)
-        # implement for r = 3
-        r = 3
-        T = math.floor(math.log2(self.N)) # num_iters - r
-
-        y = np.zeros((T, self.N, self.N)) # y[t, i, j] is message at time t from i -> j
-        z = np.zeros((T, self.N, self.N)) # z[t, i, j] is normalised message at t from i -> j
-        
-        # random initial setting
-        y[r-1, :, :] = np.random.normal(size=(self.N, self.N))
-
-        for t in tqdm(range(r, T)):
-            s = np.sum(y[t-1, :, :]) / (2 * self.M)
-            z[t-1, :, :] = y[t-1, :, :] - s
-
-            for v in range(0, self.N):
-                vd_set = self.adjacency_list[v]
-                
-                for vd in vd_set:
-                    y_vd_to_v = 0
-                    vdd_set = self.adjacency_list[vd].difference({v})
-
-                    for vdd in vdd_set:
[…]

### 2021-05-28 code/model/graph_mcmc.py

@@ -1,554 +1,0 @@
-from graph_tool.inference.blockmodel import BlockState
-import numpy as np
-# version focal seems to be winner
-from graph_tool import Graph as GT_Graph
-# X-server must be running else import will timeout
-from graph_tool.draw import graph_draw
-from graph_tool.inference import minimize_blockmodel_dl, mcmc_equilibrate, PartitionModeState, NestedBlockState
-from graph_tool.collection import data, ns
-import matplotlib.pyplot as plt
-from inference.softmax import SoftmaxNeuralNet, from_values_to_one_hot
-from data.utils import get_misc_path
-from tqdm import tqdm
-import os
-
-curr_dir = os.path.dirname(__file__)
-output_dir = os.path.join(curr_dir, "..", "output")
-
-
-def gen_output_path(filename):
-    ## valid extensions: .pdf, .png, .svg
-    if filename is not None:
-        return os.path.join(output_dir, filename)
-    return None
-
-
-class Graph_MCMC:
-
-
-    def __init__(self):
-        """Initialises empty graph"""
-        self.G = GT_Graph(directed=False)
-        
-        # initialise empty state
-        self.state = None
-        self.vertex_block_counts = None
-        self.B_max = None
-        self.relabelled_vertices = None
-
-        # treat prior on b as uniform
-        self.entropy_args = {"partition_dl": False}
-        self.mcmc_args = {"entropy_args": self.entropy_args}
-
-
-    # print helpers
-    def print_info(self):
-        N = self.G.num_vertices()
-        E = self.G.num_edges()
-        D = len(self.get_feature_names())
-        print("Graph with N={} nodes, E={} edges and D={} vertex features for training".format(N, E, D))
-
-    
-    def print_props(self):
-        print("All vertex props: " + str(set(self.G.vertex_properties.keys())))
-        print("Training vertex props: " + str(self.get_feature_names()))
-
-    
-    def list_props(self):
-        self.G.list_properties()
-
-
-    # read helpers
-    def read_from_edges(self, edges):
-        """Initialises graph based on edges"""
-        self.relabelled_vertices = self.G.add_edge_list(edges, hashed=True)
-
-    def read_from_file(self, filename):
-        filename = get_misc_path(filename)
-        self.G.load(filename)
-
-    
-    def read_from_gt(self, dataset_name):
-        self.G = data[dataset_name]
-        self.G.set_directed(False)
-
-    
-    def read_from_ns(self, dataset_name):
[…]

### 2021-05-28 code/model/player.py

@@ -1,17 +1,0 @@
-from typing import List
-
-
-class Player:
-    
-    
-    def __init__(self, csv_fields: List[str]):
-        self.team = int(csv_fields[0])
-        self.attendances = csv_fields.count("TRUE")
-        self.absences = csv_fields.count("FALSE")
-
-
-    def isFirstTeam(self):
-        return self.team == 1
-
-    def isSecondTeam(self):
-        return self.team == 2

### 2021-05-28 code/opus.py

@@ -1,242 +1,0 @@
-from model.graph_mcmc import Graph_MCMC
-from inference.softmax import SoftmaxNeuralNet
-from utils.subsampling import random_index_arr
-import numpy as np
-import matplotlib
-matplotlib.rcParams['mathtext.fontset'] = 'stix'
-matplotlib.rcParams['font.family'] = 'STIXGeneral'
-
-polbooks_args = {
-    "B": 3,
-    "f": 0.7,
-    "Tb": 1000,
-    "Tt": 10000,
-    "s": 0.05,
-    "burn-in": 0.4,
-    "thinning": 10,
-    "k": 1,
-    "D'": 3,
-    "burn-in-red": 0.4,
-    "thinning-red": 10,
-    "s-red": 0.05
-}
-
-school_args = {
-    "B": 10,
-    "f": 0.7,
-    "Tb": 1000,
-    "Tt": 10000,
-    "s": 0.2,
-    "burn-in": 0.4,
-    "thinning": 10,
-    "k": 1,
-    "D'": 10,
-    "burn-in-red": 0.4,
-    "thinning-red": 10,
-    "s-red": 0.2
-}
-
-fb_args = {
-    "B": 10,
-    "f": 0.7,
-    "Tb": 1000,
-    "Tt": 10000,
-    "s": 0.017,
-    "burn-in": 0.4,
-    "thinning": 10,
-    "k": 1,
-    "D'": 10,
-    "burn-in-red": 0.4,
-    "thinning-red": 10,
-    "s-red": 0.5
-}
-
-
-
-
-def create_polbooks_graph():
-    graph = Graph_MCMC()
-    graph.read_from_ns("polbooks")
-    graph.rename_property("label", "_label")
-    graph.convert_props_to_flags()
-    return graph, polbooks_args
-
-
-def create_school_graph():
-    graph = Graph_MCMC()
-    graph.read_from_ns("sp_primary_school/day_2")
-    graph.rename_property("label", "_label")
-    graph.rename_property("viz", "_viz")
-
-    graph.convert_props_to_flags()
-    graph.remove_property("Unknown")
-    return graph, school_args
-
-
-def create_maier_graph():
-    graph = Graph_MCMC()
-    graph.read_from_ns("facebook_friends")
-    graph.convert_props_to_flags()
-    graph.add_ego_node()
-    graph.filter_out_low_degree(min_degree=2)
-    args = {
-        "B": 18,
-        "f": 0.7,
[…]

### 2021-05-28 code/ret_script.R

@@ -1,58 +1,0 @@
-library("reticulate")
-
-print("Importing python module...")
-model <- import("model.graph_mcmc")
-
-print("Instantiating graph...")
-graph <- model$Graph_MCMC(vector())
-
-print("Reading from file...")
-graph$read_from_file("generated.gml")
-
-print("Detecting partition...")
-graph$partition(B_min = 2, B_max = 5)
-
-print("Sampling from posterior...")
-B <- graph$mcmc(100, verbose = TRUE)
-
-print("Generating features...")
-feature_names <- graph$get_feature_names()
-X <- graph$generate_feature_matrix()
-Y <- graph$generate_posterior()
-
-cat("X dimension: ", dim(X), "\n")
-cat("Y dimension: ", dim(Y), "\n")
-
-D <- dim(X)[2]
-
-
-library(keras)
-# define model
-print("Defining model...")
-model <- keras_model_sequential()
-model %>%
-    layer_dense(
-        units = B,
-        input_shape = c(D),
-        activation = "softmax",
-        name = "BlockProbs"
-    )
-
-
-# compile model
-model %>% compile(
-    optimizer = "adam",
-    loss = "categorical_crossentropy",
-    metrics = c("accuracy")
-)
-
-# train
-model %>% fit(X, Y, epochs = 10, verbose = 2)
-
-# test metrics
-score <- model %>% evaluate(X, Y, verbose = 0)
-
-cat("Training loss:", score$loss, "\n")
-
-# predict test set
-predictions <- model %>% predict(X)

### 2021-05-28 code/setup.py

@@ -1,16 +1,0 @@
-import setuptools
-
-setuptools.setup(
-    name="jormungandr",
-    version="0.0.1",
-    author="Lawrence Tray",
-    author_email="ltray99@gmail.com",
-    description="A small example package",
-    url="https://github.com/LozzaTray/Jormungandr",
-    packages=setuptools.find_packages(),
-    classifiers=[
-        "Programming Language :: Python :: 3",
-        "Operating System :: OS Independent",
-    ],
-    python_requires='>=3.6',
-)

### 2021-05-28 code/sgmcmc/interface.py

@@ -1,14 +1,0 @@
-import rpy2.robjects.packages as rpackages
-import matplotlib.pyplot as plt
-
-
-# import R's utility package
-utils = rpackages.importr('utils')
-tf = rpackages.importr('tensorflow')
-tf.install_tensorflow()
-
-try:
-    sgmcmc = rpackages.importr("sgmcmc")
-    print(sgmcmc)
-except e:
-    print(e)

### 2021-05-28 code/sgmcmc/sgmcmc.r

@@ -1,63 +1,0 @@
-library("sgmcmc")
-
-# Download and load covertype dataset
-covertype = getDataset("covertype")
-
-set.seed(13)
-test_observations = sample(nrow(covertype), 10^4)
-testSet = covertype[test_observations,]
-X = covertype[-c(test_observations),2:ncol(covertype)]
-y = covertype[-c(test_observations),1]
-dataset = list( "X" = X, "y" = y )
-
-# Get the dimension of X, needed to set shape of params$beta
-d = ncol(dataset$X)
-params = list( "bias" = 0, "beta" = matrix( rep( 0, d ), nrow = d ) )
-
-logLik = function(params, dataset) {
-    yEstimated = 1 / (1 + tf$exp( - tf$squeeze(params$bias + tf$matmul(dataset$X, params$beta))))
-    logLik = tf$reduce_sum(dataset$y * tf$log(yEstimated) + (1 - dataset$y) * tf$log(1 - yEstimated))
-    return(logLik)
-}
-
-logPrior = function(params) {
-    logPrior = - (tf$reduce_sum(tf$abs(params$beta)) + tf$reduce_sum(tf$abs(params$bias)))
-    return(logPrior)
-}
-
-stepsizesMCMC = list("beta" = 5e-6, "bias" = 5e-6)
-stepsizesOptimization = 5e-6
-
-print("hello")
-
-output = sgldcv(logLik, dataset, params, stepsizesMCMC, stepsizesOptimization, logPrior = logPrior, 
-        minibatchSize = 500, nIters = 100, verbose = TRUE, seed = 13 )
-
-print("again")
-
-yTest = testSet[,1]
-XTest = testSet[,2:ncol(testSet)]
-# Remove burn-in
-output$bias = output$bias[-c(1:1000)]
-output$beta = output$beta[-c(1:1000),,]
-iterations = seq(from = 1, to = 100, by = 10)
-logLoss = rep(0, length(iterations))
-# Calculate log loss every 10 iterations
-for ( iter in 1:length(iterations) ) {
-    j = iterations[iter]
-    # Get parameters at iteration j
-    beta0_j = output$bias[j]
-    beta_j = output$beta[j]
-    for ( i in 1:length(yTest) ) {
-        pihat_ij = 1 / (1 + exp(- beta0_j - sum(XTest[i,] * beta_j)))
-        y_i = yTest[i]
-        # Calculate log loss at current test set point
[…]

### 2021-05-28 code/social.py

@@ -1,66 +1,0 @@
-from data.facebook import FacebookGraph
-from typing import List
-from hypothesis.test_statistics import two_samples_mean_ll_ratio, students_z_test
-from model.graph import Graph
-from model.graph_mcmc import Graph_MCMC
-import numpy as np
-
-
-def standard():
-    ego_id = 0 # 0 or 107
-    gender_index = 77 # 77 or 264
-    fb = FacebookGraph(ego_id)
-    #fb.hypothesis_test_threeway(gender_index)
-    #fb.hypothesis_test_keyword("gender")
-
-    nodes_by_gender = fb.get_node_has_feature_dict(gender_index)
-    graph = Graph(fb.edges)
-    community_output = graph.abp(seed=1)
-    graph.proportions_in_each(nodes_by_gender)
-
-    #graph.draw_standard(title="Egonet id-{} post ABP".format(ego_id))
-    #graph.draw_partition(nodes_by_gender, "Egonet id-{} grouped by ABP hard partition".format(ego_id), "gender-77")
-    #graph.draw_standard(nodes_by_gender, "Egonet id-{}".format(ego_id), "gender-77")
-    #fb.gradient_ascent(community_output)
-    fb.linear_regression(community_output, ["gender", "language"])
-    #fb.linear_regression(community_output)
-
-    
-def mcmc():
-    ego_id = 0 # 0 or 107
-    fb = FacebookGraph(str(ego_id))
-    int_edges = [ (int(edge[0]), int(edge[1])) for edge in fb.edges]
-    int_edges = np.array(int_edges)
-
-    graph = Graph_MCMC()
-    graph.read_from_edges(int_edges)
-    vertices = graph.get_vertex_list()
-    
-    selected_feat_ids, selected_feat_names = fb.get_feat_ids_and_names(keywords=["gender", "language", "hometown", "last_name"])
-    #selected_feat_ids, selected_feat_names = fb.get_feat_ids_and_names()
-
-
-    for feat_id in selected_feat_ids:
-        feature_name = fb.feature_name(feat_id)
-        feature_flags = [fb.node_has_feature(str(vertex), feat_id) for vertex in vertices]
-        graph.add_property(feature_name, "bool", feature_flags)
-
-    graph.partition(B_min=2, B_max=10)
-    graph.mcmc(num_iter=100)
-    #graph.draw("partition.png")
-    #graph.plot_matrix()
-    #graph.plot_community_property_fractions()
-
[…]

### 2021-05-28 code/streaming.py

@@ -1,36 +1,0 @@
-from data.twitch import TwitchGraph
-from model.graph_mcmc import Graph_MCMC
-
-def save():
-    print("Loading data...")
-    twitch = TwitchGraph()
-    print("Converting to graph-tool...")
-    graph = twitch.generate_mcmc_graph()
-    print("Saving to gt file...")
-    graph.save_to_file("twitch")
-    print("Done")
-
-def analyse():
-    print("Loading from file...")
-    graph = Graph_MCMC()
-    graph.read_from_file("twitch.gt")
-    print("Done")
-
-    graph.partition(B_min=4, B_max=4)
-    graph.mcmc(100, verbose=True)
-    graph.draw("twitch.png")
-
-    #graph.plot_posterior_props()
-    names = graph.get_feature_names()
-
-    classifier = graph.sample_classifier_mala(100, step_scaling=0.001, verbose=True)
-    classifier.thin_samples()
-
-
-    #classifier.plot_U()
-    classifier.plot_block_principal_dims(names.copy(), 1)
-    classifier.plot_sampled_weights(names.copy(), std_dev_multiplier=1, null_space=1)
-
-
-if __name__ == "__main__":
-    analyse()

### 2021-05-28 code/utils/colors.py

@@ -1,19 +1,0 @@
-from colour import Color
-import math
-
-
-def color_between(val, start_col="blue", end_col="red", resolution=10):
-    """interpolates a value in range [0,1] to corresponding end colour"""
-    if val >= 1:
-        val = 0.99
-    elif val < 0:
-        val = 0
-        
-    start = Color(start_col)
-    end = Color(end_col)
-    index = math.floor(val * resolution)
-    return list(start.range_to(end, resolution))[index].hex
-
-
-def plt_color(index):
-    return "C" + str(index)

### 2021-05-28 code/utils/plotting.py

@@ -1,18 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-
-
-def sorted_bar_plot(values, labels, title, xlabel="", top=20):
-    sorted_indices = np.argsort(np.abs(values))
-    if len(sorted_indices) > top:
-        sorted_indices = sorted_indices[:top]
-
-    sorted_values = values[sorted_indices]
-    sorted_labels = labels[sorted_indices]
-
-    plt.figure()
-    plt.barh(sorted_labels, sorted_values)
-    plt.title(title)
-    plt.xlabel(xlabel)
-    plt.grid()
-    plt.show()

### 2021-05-28 code/utils/subsampling.py

@@ -1,20 +1,0 @@
-import numpy as np
-
-
-def random_index_arr(n, fraction):
-    """
-    Return two index arrays to form test and training set
-
-        Parameters:
-            n (int): length of array to index
-            fraction (float): fraction to use for training set
-
-        Returns:
-            train_indices (int[]): indices of training set
-            test_indices (int[]): indices of test set    
-    """
-    indices = np.random.permutation(n)
-    m = int(n * fraction)
-    train_indices = indices[:m]
-    test_indices = indices[m:]
-    return train_indices, test_indices

### 2021-05-28 cued/additional.tex
Additional appendix proof-read

@@
-\section{Appendix: Additional material}
+\section{Additional material}
 
@@
 
-For reference, we provide the likelihood and priors for the microcanonical SBM proposed by \citet{Peixoto-Bayesian-Microcanonical}. We have that the graph $A$ is drawn from the SBM:
+For reference, we provide the likelihood and priors for the microcanonical SBM proposed by \citet{Peixoto-Bayesian-Microcanonical}. We have that the graph $A$ is drawn from the SBM,
 %
 \begin{equation}
-	A \sim \textrm{DC-SBM}_{\textrm{MC}}(b, e, k)
+	A \sim \textrm{DC-SBM}_{\textrm{MC}}(b, e, k),
 \end{equation}
 %
-With edges placed uniformly at random but respecting the constraints imposed by $b, e$ and $k$.
-The likelihood calculation for $p(A|k, e, b)$ then reduces to a case of counting configurations that yield the same adjacency matrix, $\Xi (A)$, and dividing by the total number of configurations possible, $\Omega(e)$.
-This is why this formulation is given the microcanonical moniker.
-If we consider the half-edges to be distinguishable for a moment, the total number of configurations that satisfy the $e$ constraint is:
+with edges placed uniformly at random but respecting the constraints imposed by $b, e$ and $k$.
+The computation of the likelihood, $p(A|k, e, b)$, is simply a case of counting the number of configurations that yield the same adjacency matrix and dividing by the total number of configurations possible.
+This is what gives this formulation the `microcanonical' moniker.
+If we consider the half-edges to be distinguishable for a moment, the total number of configurations that satisfy the $e$ constraint is,
 %
 \begin{equation}
-	\Omega(e) = \frac{\prod_{r} e_r !}{\prod_{r,s : r < s} e_{rs}! \cdot \prod_{r} e_{rr}!!}
+	\Omega(e) = \frac{\prod_{r} e_r !}{\prod_{r,s : r < s} e_{rs}! \cdot \prod_{r} e_{rr}!!},
 \end{equation}
 %
-Where $e_r \coloneqq \sum_{s} e_{rs}$ and $(2m)!!
+where $e_r \coloneqq \sum_{s} e_{rs}$ and $(2m)!!
 \coloneqq 2^m m!$.
-Nevertheless, a great number of these configurations yield the same graph $A$.
-We denote the number of configurations that yield the adjacency matrix $A$ with $\Xi(A)$, which can be computed as:
+Nevertheless, a great number of these $\Omega(e)$ total configurations yield the same graph $A$.
+We denote the number of configurations that yield the adjacency matrix $A$ with,
 %
 \begin{equation}
-	\Xi(A) \coloneqq \frac{\prod_i k_i !}{\prod_{i,j : i < j} A_{ij} ! \prod_i A_{ii} !! }
+	\Xi(A) \coloneqq \frac{\prod_i k_i !}{\prod_{i,j : i < j} A_{ij} ! \prod_i A_{ii} !! }.
 \end{equation}
 %
-Note the similarity between the forms of $\Omega(e)$ and $\Xi(A)$ as $e$ is effectively the adjacency matrix of the block-graph. With these defined we can write the overall likelihood:
+Note the similarity between the forms of $\Omega(e)$ and $\Xi(A)$ as $e$ is effectively the adjacency matrix of the block-graph.
+With these defined we can write the overall likelihood as the ratio of the two:
 %
 \begin{equation}
-	p(A|k,e,b) = \frac{\Xi(A)}{\Omega(e)}
+	p(A|k,e,b) = \frac{\Xi(A)}{\Omega(e)}.
 \end{equation}
@@
 Obviously, this form is only defined if $A$ respects the constraints imposed by $(k,e,b)$ else the likelihood is 0.
-With the likelihood defined, we move on to the prior.
-As discussed in the main text, the block membership vector $b$ is an intermediate variable in the FFBM and not a parameter so we are not free to choose a prior for it.
 
-Here we recall for reference
-the priors $p(\psi | b)$ from \cite{Peixoto-Bayesian-Microcanonical}:
+With the likelihood defined, we turn our attention to the prior, necessary for Bayesian analysis.
+As discussed in the main text, the block membership vector $b$ is an intermediate variable in the FFBM so we are not free to choose a prior for it.
+Nevertheless, we can define the prior on $(e,k)$ as a conditional on $b$.
+We switch back to the FFBM notation of $\psi = \{\psi_e, \psi_k\}$ to encapsulate $e$ and $k$.
+We recall for reference
+the prior $p(\psi | b)$ from \cite{Peixoto-Bayesian-Microcanonical}:
 %
 \begin{equation}
-	p(\psi_e=e, \psi_k=k | b) = p(e | b) p(\psi_k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
+	p(\psi_e=e, \psi_k=k | b) = p(e | b) p(k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
 	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right],
@@
 
-These were chosen carefully in \cite{Peixoto-Bayesian-Microcanonical} to 
+These priors were chosen carefully in \cite{Peixoto-Bayesian-Microcanonical} to 
 more closely match the structure of empirical networks than simple 
@@
 Section~\ref{s:sfb} we used 
-the Metropolis Adjusted Langevin Algorithm (MALA)
+the Metropolis-adjusted Langevin algorithm (MALA)
 to 
@@
 the {\em step-size scaling}, is the only free parameter.
-%
-%For approximate methods, we can choose to bypass the MH accept-reject entirely to speed up computation.
-If this is done, the algorithm is instead called stochastic gradient Langevin diffusion (SGLD) \cite{Bayesian-SGLD}.
+
+As an aside, we note that an interesting modification to MALA can be made if we skip out the Metropolis-Hastings accept/reject step entirely.
+If this is done, the algorithm is instead called the stochastic gradient Langevin dynamics (SGLD) \cite{Bayesian-SGLD}.
 This speeds up computation at the expense of the exactness of the method.
-Indeed, it relies on the fact that as the step-size is annealed to 0, the acceptance probability is close 1.
-%
-Nevertheless, to choose the value of $s$ we must quantify the trade-off between acceptance ratio and burn-in time.
-We do this by way of an example with the primary school dataset \cite{schools}.
+Indeed, it relies on the fact that as the step-size is annealed to 0, the acceptance probability gets close to 1.
+Figure \ref{fig:step-size-tradeoff} gives empirical backing to this approximate rule-of-thumb.
+Nevertheless, as SGLD is not exact, we only use MALA for experimentation.
+
+In order to choose the exact value of the {\em step-size scaling} $s$ for each experiment, we must quantify the trade-off between acceptance ratio and burn-in time.
+We illustrate the adopted method using the primary school dataset \cite{schools} as an example.
 We run the $b$-chain for $T_b = 1,000$ iterations and subsample such that $\Tcal_b$ is computed with $\kappa_b=0.2$ and $\lambda_b=5$.
 We then use this to run the $\theta$-chain for $T_\theta = 10,000$ iterations.
-The acceptance ratio -- which we denote $r_\alpha$ -- is easy enough to compute as simply the fraction of proposed moves which we accept.
-However, to quantify burn-in time we must use a different metric.
-We can plot the mean of the objective function averaged over all samples as a proxy for burn-in time:
+The acceptance ratio -- which we denote $r_\alpha$ -- is simply the fraction of proposed moves which we accept.
+However, to measure burn-in we define a new metric: the mean of the objective function averaged over all samples,
 %
 \begin{equation}
-	\bar{U} \coloneqq \frac{1}{T_\theta} \sum_{t \in [T_\theta]} U\left( \theta^{(t)} \right)
+	\bar{U} \coloneqq \frac{1}{T_\theta} \sum_{t \in [T_\theta]} U\left( \theta^{(t)} \right).
 \end{equation}
 %
-As the chain will equilibrate in the vicinity of a minima in $U(\theta)$, the average over all samples is a rough indicator of the speed of burn-in.
-This assumes that the random starting point has high $U(\theta)$.
-This assumption holds as the domain of $\theta$ is a high dimensional space so we are unlikely to get a low $U(\theta)$ simply from a lucky guess.
+Since the chain will equilibrate in the vicinity of a minima in $U(\theta)$, the average over all samples is a rough indicator of the speed of burn-in.
 The higher the value of $\bar{U}$, the longer we infer the chain took to burn in and reach equilibrium.
-We plot the acceptance ratio $r_\alpha$ and average objective $\bar{U}$ for $10$ values of the step-size multiplier $s$ logarithmically spaced in the range $(10^{-2}, 10^1)$ on figure \ref{fig:step-size-tradeoff}.
+This method relies on the assumption that the random starting point $\theta^{(0)}$ has a high value for the objective function $U(\theta^{(0)})$; this assumption holds as the domain of $\theta$ is a high dimensional space so we are unlikely to get a low $U(\theta)$ from a lucky guess.
+We plot the acceptance ratio $r_\alpha$ and average objective $\bar{U}$ for $10$ values of the step-size multiplier $s$ logarithmically spaced in the range $(10^{-2}, 10^1)$ on Figure \ref{fig:step-size-tradeoff}.
 %
@@
 
-We see immediately that acceptance ratio $r_\alpha$ steadily decreases with $s$.
-This is expected as it is unlikely we stay in a high density region of our target distribution if we are making very large steps each time.
-A low acceptance ratio leads to wasted samples and this inefficiency is obviously undesirable.
-Indeed, figure \ref{fig:step-size-tradeoff} suggests that a low step-size $s$ yields high acceptance probability thus justifying the approximation used in SGLD.
+We see immediately that the acceptance ratio $r_\alpha$ steadily decreases with $s$.
+This is expected as large jumps for each proposal make it unlikely we stay in a high density region of our target distribution.
+A low acceptance ratio leads to wasted samples and is obviously undesirable.
 
@@
 However, this effect plateaus for $s>0.2$ for this particular dataset.
-Therefore, we choose $s=0.2$ as our step-size multiplier for the primary school datasets as this yields the best trade-off between $r_\alpha$ and $\bar{U}$.
-Indeed, the acceptance ratio is still high for $r_\alpha \approx 0.8$ for $s=0.2$.
+Therefore, we choose $s=0.2$ as our step-size multiplier for the primary school dataset as this yields the best trade-off between $r_\alpha$ and $\bar{U}$.
 The step-sizes for the other datasets were chosen following the same argument but we do not repeat the process here.
@@
 By way of illustration, we can consider the primary school dataset \cite{schools}.
-We plot the normalised objective function $U\left( \theta^{(t)} \right) / N$ with respect to MALA iteration on figure \ref{fig:school-U-orginal}.
+We plot the normalised objective function $U\left( \theta^{(t)} \right) / N$ with respect to MALA iteration on Figure \ref{fig:school-U-orginal} -- using the hyperparameters given in Appendix \ref{appdx:hyperparams}.
 We see that the chain converges to the modal neighbourhood quickly (within about 2000 iterations of a total 10,000).
-Nevertheless, we pick $\kappa_\theta=0.4 > 0.2$ just to be on the safe-side.
+Nevertheless, we pick $\kappa_\theta=0.4 > 0.2$ to be on the safe-side.
 $\lambda_\theta=10$ is chosen somewhat arbitrarily as we do not require neighbouring samples to be independent.
-Nevertheless, this thinning does speed up computation of quantities that require an average over all the retained samples -- as $|\Tcal_\theta| < T_\theta$.
+However, this thinning also has the advantage of speeding up computation of quantities that require an average over all the retained samples -- as $|\Tcal_\theta| < T_\theta$.
 %
@@
 
-We see on figure \ref{fig:school-U-thinned} that $U(\theta^{(i)})$ does bounce around a small amount.
-This is to be expected we are sampling from the posterior rather than converging to the MAP.
-There does appear to be variation over a long time-scale but we can live with this as we do not require sample independence.
+We plot the retained samples on Figure \ref{fig:school-U-thinned}.
+$U(\theta^{(i)})/N$ does bounce around a small amount but this is expected as we are sampling from the posterior rather than converging to the MAP solution.
+There does appear to be variation over a long time-scale but this is acceptable as we do not strictly require sample independence.
 \FloatBarrier
@@
 
-There is a challenge when dealing with vertex-features.
 Many vertex feature often take only one of several discrete values within a set.
-For example, with the primary school dataset \cite{schools}, a pupil's school-class can take one of 10 values from \{"1A", "1B", "2A", "2B" \dots "5A", "5B"\}.
+For example, with the primary school dataset \cite{schools}, a pupil's school-class can take one of 10 values from \{``1A", ``1B", ``2A", ``2B" \dots ``5A", "5B"\}.
 
 It is not immediately obvious how to encode this for the feature-to-block classifier.
-Although this is technically only a single dimension, we choose to expand the data into a set of 10 binary feature flags $\in \Xcal = \{0, 1\}$.
+Although this is technically only a single dimension, we choose to expand the data into a set of 10 binary feature flags all taking values in the set $\Xcal = \{0, 1\}$.
 Although only one flag will be set at a time it is the simplest method for representing discrete-valued data.
-The reason we choose $\Xcal=\{0, 1\}$ and not $\Xcal = \{-1, 1\}$ is that the former allows for simpler analysis of the resulting weight values; when a feature $d$ is switched off then $x_{d}=0$ given $\Xcal = \{0, 1\}$, and the feature has no impact on the softmax layer.
+The reason we choose $\Xcal=\{0, 1\}$ and not $\Xcal = \{-1, 1\}$ is that the former allows for simpler analysis of the resulting weight values; when a feature $d$ is switched off then $x_{d}=0$ given $\Xcal = \{0, 1\}$, and the feature has no impact on the softmax output.
+This would not be the case for $\Xcal = \{-1, 1\}$.
 
-It is more intuitive to only model positive relationships.
+Furthermore, it is more intuitive to only model positive relationships.
 In other words, we prefer to say that this block is comprised of the vertices with feature $d$ turned on rather than this block contains the vertices with feature $d$ switched off.
-To make this explicit, say we have a feature encoding a pupil's gender: \{"male", "female"\}, we represent this as 2 binary feature flags "male": \{0, 1\} and "female": \{0, 1\}.
+To summarise the approach explicitly, say we have a feature encoding a pupil's gender: \{"male", "female"\}, we represent this as 2 binary feature flags "male": \{0, 1\} and "female": \{0, 1\}.
 This approach may seem inefficient but it is vital to ensure we can interpret the resulting parameter distributions.
 
-Nevertheless, this dimensionality expansion means that we cannot include a bias term in our classifier as then the MAP solution is less peaked and our $\theta$-samples will have higher variance.
-To illustrate this point we need only look at the form of the soft-max classifier with bias vector $\beta$, operating on a vector $x$ which only has one non-zero entry $x_d=1$.
+Nevertheless, this dimensionality expansion means that we should not include a bias term in our softmax as then the MAP solution is less peaked and our $\theta$-samples will have higher variance.
+To illustrate this point we need only look at the form of the softmax activation function with bias vector $\beta$, operating on a vector $x$ which only has one non-zero entry $x_d=1$.
 Component $j$ of the output is given by:
@@
 	\phi_j(x) = \frac{\exp(w_j^T x + \beta_j)}{\sum_{k \in [B]} \exp(w_k^T x + \beta_k)} = 
-	\frac{\exp(w_{jd} + \beta_j)}{\sum_{k \in [B]} \exp({w_{kd} + \beta_k})}
+	\frac{\exp(w_{jd} + \beta_j)}{\sum_{k \in [B]} \exp({w_{kd} + \beta_k})}.
 	\label{eqn:softmax-bias}
@@
 %
-The term $w_{kd} + \beta_k$ is effectively the new weight for component $k$ as this term always appears together.
-The bias term $\beta_k$ is an unnecessary extra degree of freedom.
-To illustrate this point, we set $\beta_k=\beta_0 \forall j$.
+The term $w_{kd} + \beta_k$ is effectively the new weight for component $k$ as these two terms will always appear together.
+The bias term $\beta_k$ is therefore an unnecessary extra degree of freedom.
+To illustrate this point, we set $\beta_k=\beta_0$ for all $j$.
 If this is the case then equation \ref{eqn:softmax-bias} can be simplified to:
@@
 This is not the whole picture as we also have the prior which introduces a regularisation term which favours weights closer to 0.
+This ensures the MAP solution is unique.
 Nevertheless, for mutually exclusive binary feature flags, the bias term does not add to the expressiveness of the model and only serves to complicate analysis.
-We therefore remove the bias term from the Softmax classifier.
-Even for data comprised of several such features (e.g.
+We therefore remove the bias term from the softmax.
+
+Even for data comprised of several feature-classes each expanded into flags (e.g.
 school class and gender) we prefer to discard the bias term.
-The bias term only serves to decrease the loss function by leveraging information about the size of each detected block and not feature information.
+The bias term only serves to decrease the objective function by leveraging information about the size of each detected block and not feature information.
 We do not wish to be overly confident in our block predictions when such predictions are influenced by block size and not feature information.

### 2021-05-28 cued/additional.tex
Edit hyp-test appendix

@@
 Indeed, the approach was initially developed with a bias term but we found its removal yielded more reliable and reproducible results.
+
+\subsection{Hypothesis test on feature weights}
+\label{appdx:hyp-test}
+
+We are given samples $\{\theta^{(t)}\}_{t \in \Tcal_\theta} \sim p(\theta | A, X)$ and wish to determine the statistical significance of the weights.
+We adopt matrix notation for simplicity by representing $\theta$ with the matrix $B \times D$ matrix of feature weights $W$.
+The question is to determine the significance of a particular feature $d$ by examining the value of $W_{id}$ for all the values $i \in [B]$.
+We know that the posterior is proportional to the prior multiplied by the likelihood,
+%
+\begin{equation}
+	p(\theta|A, X) \propto p(\theta) \cdot p(A | \theta, X),
+\end{equation}
+%
+assuming that the feature matrix $X$ is already given. The prior term can be evaluated but the likelihood is intractable. The closest we can get is through a Monte-Carlo integration over $b$,
+%
+\begin{align}
+	p(A | \theta, X) &= \sum_{b \in [B]^N} p(A, b | \theta, X) \nonumber \\
+	&= \sum_{b \in [B]^N} p(A | b, \theta, X) \cdot p(b | \theta, X) \nonumber \\
+	&= \sum_{b \in [B]^N} p(A | b) \cdot p(b | \theta, X) \nonumber \\
+	&\approx \sum_{i} p\left( A | b^{(i)} \right) \quad \textrm{with} \quad b^{(i)} \sim p(b| \theta, X).
+	\label{eqn:mc-likelihood}
+\end{align}
+%
+This could be implemented for a single value of $\theta$ but such a form cannot be used to characterise the overall form of the posterior.
+Nevertheless, the form in~(\ref{eqn:mc-likelihood}) highlights something interesting.
+The likelihood is peaked around areas of $\theta$ that generate a partition $b$ that is highly likely in the SBM sense -- high $p(A|b)$.
+This provides the motivation for using the Laplace approximation for modelling the posterior $p(\theta | A, X) \approx p(\theta; \hat{\mu}, \hat{\Sigma})$; we use a multivariate Gaussian to approximate the posterior.
+Indeed, the Laplace approximation is often used for modelling the posterior in logistic classification \cite{laplace}.
+This approximation is not exact but it motivates the derivation of the dimensionality reduction technique we construct by analogy with hypothesis testing.
+Marginalising over the other elements, the posterior for each element of the weight matrix $W$ can be approximated by,
+%
+\begin{equation}
+	p(W_{ij}|A, X) \approx \Gaussian(W_{ij}|\hat{\mu}_{ij}, \hat{\sigma}_{ij}^2).
+\end{equation}
+%
+Where we have used the set of samples for $W$ drawn according to the exact posterior, to calculate unbiased estimates for the mean and standard deviation:
+%
+\begin{equation}
+	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
+	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2.
+\end{equation}
+%
+This approximation is not exact but we can show it is accurate empirically.
+Indeed, if we run the primary school experiment with hyper-parameters given in Appendix \ref{appdx:hyperparams}, we can then plot histograms of the collected $W$-samples and compare these to the Laplace approximation.
+The results are given on Figure \ref{fig:school-histogram}.
+Even though the Laplace approximation is not exact, it is remarkably reliable.
+
+\begin{figure}[!h]
+	\centering
+	\begin{subfigure}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{school-sample-histogram-16.png}
+		\caption{Weight for block 1, class 3B}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{school-sample-histogram-37.png}
+		\caption{Weight for block 3, class 4A}
+	\end{subfigure}
+	\hfill
+	\begin{subfigure}[t]{0.32\linewidth}
+		\centering
+		\includegraphics[width=\linewidth]{school-sample-histogram-84.png}
+		\caption{Weight for block 8, class 2B}
+	\end{subfigure}
+
+	\caption{Histograms of sampled weights for the primary school experiment \cite{schools}. Dotted line is the applied Laplace approximation.}
+	\label{fig:school-histogram}
+\end{figure}
+
+Using the approximation, we wish to construct a test to determine whether a particular weight value has $|W_{ij}| > c$ with high probability, where $c>0$. More specifically, we wish to decide between:
+%
+\begin{equation}
+	\begin{aligned}
+		H_0&: |W_{ij}| \leq c \\
+		H_1&: |W_{ij}| > c
+	\end{aligned}
+\end{equation}
+%
+We can consider the probabilities $p(W_{ij} > c)$ and $p(W_{ij} < c)$ separately. Starting with $p(W_{ij} > c)$, we write this as,
+%
+\begin{align*}
+	p(W_{ij} > c) &= p\left(Z > \frac{c - \hat{\mu}_{ij}}{\hat{\sigma}_{ij}}\right) \\
+	&= p\left(-Z < \frac{\hat{\mu}_{ij} - c}{\hat{\sigma}_{ij}}\right)\\
+	&= \Phi\left(\frac{\hat{\mu}_{ij} - c}{\hat{\sigma}_{ij}}\right),
+\end{align*}
+%
+where $Z \sim \Gaussian(0,1)$ and $\Phi(\cdot)$ is the standard Gaussian cumulative density function.
+We introduce the variable $k$ to control the degree of significance of the result.
+Enforcing $k$ to be less than the argument of $\Phi$ gives us that:
+%
+\begin{equation*}
+	\hat{\mu}_{ij} - k \hat{\sigma}_{ij} > c.
+\end{equation*}
+%
+As $\Phi(\cdot)$ is monotonically increasing, this yields the following:
+%
+\begin{equation}
+	p(W_{ij} > c) > \Phi(k) \iff \hat{\mu}_{ij} - k \hat{\sigma}_{ij} > c.
+	\label{eqn:hyp-test-positive}
+\end{equation}
+%
+A similar argument can be made to show that:
+%
+\begin{equation}
+	p(W_{ij} < -c) > \Phi(k) \iff \hat{\mu}_{ij} + k \hat{\sigma}_{ij} < -c.
+	\label{eqn:hyp-test-negative}
+\end{equation}
+%
+Clearly equations \ref{eqn:hyp-test-positive} and \ref{eqn:hyp-test-negative} cannot both hold simultaneously as $c, \hat{\sigma}_{ij}, k>0$.
+Nevertheless, a necessary and sufficient condition for one of them to hold is:
+%
+\begin{equation}
+	(\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c) = \emptyset
+	\label{eqn:hyp-test-empty}
+\end{equation}
+%
+This is equivalent to saying that $\hat{\mu}_{ij}$ is at least $k$ standard deviations away from the range $(-c, c)$.
+We choose to reject the null hypothesis $H_0$ if and only if equation \ref{eqn:hyp-test-empty} holds.
+We can say with probability greater than or equal to $\Phi(k)$ that either $W_{ij} < -c$ or $W_{ij} > c$ thus rejecting the null $H_0: |W_{ij}| \leq c$.
+
+Note that we choose to accept $H_0$ even in the case that the combined probabilities of $p(W_{ij} < -c) + p(W_{ij} > c) > \Phi(k)$, if equation \ref{eqn:hyp-test-empty} is not satisfied.
+This is because we wish to apply this for dimensionality reduction purposes.
+We only want to keep weights for which we have a confident prediction outside the range $(-c, c)$ not those that succeed by summing the two extremes of their distribution.
+Note that this case is very rare empirically but it nonetheless must be specified.

### 2021-05-28 cued/derivations.tex
Additional appendix proof-read
Edit hyp-test appendix

@@
-\section{Appendix: Derivations}
+\section{Derivations}
 
@@
 We determine the form of $p(b| X)$ by integrating
-out the parameters $\theta$. From the definitions we have:
+out the parameters $\theta$. From the definitions, we have:
 %
@@
 	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i \in [N] } \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
-	&= \prod_{i \in [N]} \int \frac{\exp(w_{b_i}^T \tilde{x}_i) \prod_{j \in [B]} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k \in [B]} \exp(w_{k}^T \tilde{x}_i)} dw_{1:B}.
+	&= \prod_{i \in [N]} \int \frac{\exp(w_{b_i}^T x_i) \prod_{j \in [B]} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k \in [B]} \exp(w_{k}^T x_i)} dw_{1:B}.
 \end{align*}
@@
 because the prior is the same for each $w_j$. 
-Therefore, the integral is only a function of $\tilde{x}_i$ and $\sigma_\theta^2$,
+Therefore, the integral is only a function of $x_i$ and $\sigma_\theta^2$,
 which means that, as a function of $b$, $p(b|X)\propto 1$. As
@@
 Therefore, discarding constant terms, we 
-obtain exactly the representation~(\ref{eqn:U-form}), as claimed.
+obtain:
+%
+\begin{equation}
+	U(\theta) = \left( \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log \frac{1}{a_{ij}} \right)
+	+ \frac{1}{2\sigma_\theta^2} \|\theta\|^2.
+\end{equation}
+%
+This is exactly the representation~(\ref{eqn:U-form}), as claimed.
 
@@
 for all $i$.
-
-\subsection{Hypothesis test on feature weights}
-\label{appdx:hyp-test}
-
-We are given samples $\{\theta^{(t)}\}_{t \in \Tcal_\theta} \sim p(\theta | A, X)$ and wish to determine the statistical significance of the weights.
-We adopt matrix notation for simplicity by representing $\theta$ with the matrix $B \times D$ matrix of feature weights $W$.
-The question is to determine the significance of a particular feature $d$ by examining the value of $W_{id}$ for all the values $i \in [B]$.
-We know that the posterior is proportional to the prior multiplied by the likelihood (assuming that the feature matrix $X$ is already given):
-%
-\begin{equation}
-	p(\theta|A, X) \propto p(\theta) \cdot p(A | \theta, X)
-\end{equation}
-%
-The prior term can be evaluated but the likelihood is intractable. The closest we can get is through a Monte-Carlo integration over $b$:
-%
-\begin{align}
-	p(A | \theta, X) &= \sum_{b \in [B]^N} p(A, b | \theta, X) \nonumber \\
-	&= \sum_{b \in [B]^N} p(A | b, \theta, X) \cdot p(b | \theta, X) \nonumber \\
-	&= \sum_{b \in [B]^N} p(A | b) \cdot p(b | \theta, X) \nonumber \\
-	&\approx \sum_{i} p\left( A | b^{(i)} \right) \quad \textrm{with} \quad b^{(i)} \sim p(b| \theta, X)
-	\label{eqn:mc-likelihood}
-\end{align}
-%
-This could be implemented for a single value of $\theta$ but such a form cannot be used to characterise the overall form of the posterior.
-Nevertheless, the form in equation \ref{eqn:mc-likelihood} still does highlight something interesting.
-The likelihood is peaked around areas of $\theta$ that generate a partition $b$ that is highly likely in the SBM sense -- high $p(A|b)$.
-This provides the motivation for using the Laplace approximation for modelling the posterior $p(\theta | A, X) \approx p(\theta; \mu, \Sigma)$.
-Indeed, the Laplace approximation is often used for modelling the posterior in logistic classification \cite{laplace}.
-This approximation is not exact but it motivates the derivation of the dimensionality reduction technique we construct by analogy with hypothesis testing.
-Marginalising over the other elements, the posterior for each element of the weight matrix $W$ can be approximated by:
-%
-\begin{equation}
-	p(W_{ij}|A, X) \approx \Gaussian(W_{ij}|\hat{\mu}_{ij}, \hat{\sigma}_{ij}^2)
-\end{equation}
-%
-Where we have used the set of samples for $W$ drawn according to the exact posterior, to calculate unbiased estimates for the mean and standard deviation:
-%
-\begin{equation}
-	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
-	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2
-\end{equation}
-%
-This approximation is not exact but we can show it is accurate empirically.
-Indeed, if we run the primary school experiment with hyper-parameters given in appendix \ref{appdx:hyperparams}, we can then plot histograms of the collected $W$-samples and compare these to the Laplace approximation.
-The results are given on figure \ref{fig:school-histogram}.
-Even though the Laplace approximation is not exact, it is remarkably reliable.
-
-\begin{figure}[!h]
-	\centering
-	\begin{subfigure}[t]{0.3\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{school-sample-histogram-16.png}
-		\caption{Weight for block 1, class 3B}
-	\end{subfigure}
-	\hfill
-	\begin{subfigure}[t]{0.3\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{school-sample-histogram-37.png}
-		\caption{Weight for block 3, class 4A}
-	\end{subfigure}
-	\hfill
-	\begin{subfigure}[t]{0.3\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{school-sample-histogram-84.png}
-		\caption{Weight for block 8, class 2B}
-	\end{subfigure}
-
-	\caption{Histograms of sampled weights for the primary school experiment \cite{schools}. Dotted line is the applied Laplace approximation.}
-	\label{fig:school-histogram}
-\end{figure}
-
-Using the approximation, we wish to construct a test to determine whether a particular weight value has $|W_{ij}| > c$ with high probability, where $c>0$. More specifically, we wish to determine:
-%
-\begin{equation}
-	\begin{aligned}
-		H_0&: |W_{ij}| \leq c \\
-		H_1&: |W_{ij}| > c
-	\end{aligned}
-\end{equation}
-%
-We can consider the probabilities $p(W_{ij} > c)$ and $p(W_{ij} < c)$ separately. Starting with $p(W_{ij} > c)$ we can write:
-%
-\begin{align*}
-	p(W_{ij} > c) &= p\left(Z > \frac{c - \hat{\mu}_{ij}}{\hat{\sigma}_{ij}}\right) \\
-	&= p\left(-Z < \frac{\hat{\mu}_{ij} - c}{\hat{\sigma}_{ij}}\right)\\
-	&= \Phi\left(\frac{\hat{\mu}_{ij} - c}{\hat{\sigma}_{ij}}\right)
-\end{align*}
-%
-Where $Z \sim \Gaussian(0,1)$ was introduced as a temporary variable and $\Phi(\cdot)$ is the standard Gaussian cdf.
-Now we introduce the variable $k$ to control the degree of significance of the result.
-By enforcing $k$ to be less than the argument of $\Phi$ we show that:
-%
-\begin{align*}
-	\frac{\hat{\mu}_{ij} - c}{\hat{\sigma}_{ij}} &> k \\
-	\hat{\mu}_{ij} - k \hat{\sigma}_{ij} &> c
-\end{align*}
-%
-As $\Phi(\cdot)$ is monotonically increasing, this result gives us that:
-%
-\begin{equation}
-	p(W_{ij} > c) > \Phi(k) \iff \hat{\mu}_{ij} - k \hat{\sigma}_{ij} > c
-	\label{eqn:hyp-test-positive}
-\end{equation}
-%
-A similar argument can be made to show that:
-%
-\begin{equation}
-	p(W_{ij} < -c) > \Phi(k) \iff \hat{\mu}_{ij} + k \hat{\sigma}_{ij} < -c
-	\label{eqn:hyp-test-negative}
-\end{equation}
-%
-Clearly equations \ref{eqn:hyp-test-positive} and \ref{eqn:hyp-test-negative} cannot both hold simultaneously as $c, \hat{\sigma}_{ij}, k>0$.
-Nevertheless, a necessary and sufficient condition for one of them to hold is:
-%
-\begin{equation}
-	(\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c) = \emptyset
-	\label{eqn:hyp-test-empty}
-\end{equation}
-%
-We therefore choose to reject the null hypothesis $H_0$ if and only if equation \ref{eqn:hyp-test-empty} holds.
-This tells us with confidence greater than $\Phi(k)$ that either $W_{ij} < -c$ or $W_{ij} > c$ thus rejecting the null $H_0: |W_{ij}| \leq c$.
-Note that we choose to accept $H_0$ for the case where the combined probabilities of $p(W_{ij} < -c) + p(W_{ij} > c) > \Phi(k)$, if equation \ref{eqn:hyp-test-empty} is not satisfied.
-This is because we wish to apply this for dimensionality reduction purposes.
-We only want to keep weights that have a confident prediction outside the null-space not those that succeed by summing the two extremes of their distribution.
-Note that this case is very rare empirically.

### 2021-05-29 cued/additional.tex
More appendix

@@
 \subsection{SBM likelihood and prior}
-\label{appdx:sbm}
+\label{appdx:sbm-likelihood}
 
@@
 
-With the likelihood defined, we turn our attention to the prior, necessary for Bayesian analysis.
+\subsection{SBM prior}
+\label{appdx:sbm-prior}
+
+We turn our attention to the prior on the SBM parameters, necessary for Bayesian analysis.
 As discussed in the main text, the block membership vector $b$ is an intermediate variable in the FFBM so we are not free to choose a prior for it.
@@
 uniform priors. We do not repeat these arguments here.
+
+\subsection{Metropolis-Hastings}
+\label{appdx:metropolis-hastings}
+
+Recall from Section \ref{sec:inference}, we state that we can implement sampling from the $b$ and $\theta$-chains through the Metropolis-Hastings algorithm \cite{hastings-alg}.
+For reference, we define the prerequisites for implementing the Metropolis-Hastings algorithm.
+
+Suppose we wish to draw samples for some general $x$ from the normalised density $\pi^*(x)$.
+We just need to define a proposal distribution $q(x, x')$ for proposing a move $x \rightarrow x'$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(x) \propto \pi*(x)$, point-wise.
+Each proposed move $x \rightarrow x'$ is then accepted with probability $\alpha$~(\ref{eqn:mh-accept}) else it is rejected and we stay at $x$.
+
+\begin{equation}
+\alpha(x, x') = \min \left( \frac{\pi(x') q(x', x)}{\pi(x) q(x, x')} , 1 \right).
+\label{eqn:mh-accept}
+\end{equation}
+
+This accept-reject step ensures the resulting Markov Chain is in detailed balance with the target distribution $\pi(\cdot)$. Therefore, the set of samples $\{x^{(t)}\} \sim \pi^*(x)$
 

### 2021-05-29 cued/prelim.tex
More appendix

@@
 
-We first need a model for community-like structure in a network.
-For this we adopt the stochastic block model (SBM) - widely used across academia.
+We first need a model for community-like structure in a network. For this we adopt the widely-used stochastic block model (SBM).
 Each node in the graph belongs to a unique community called a block.
@@
 Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}.
-To allow for degree-variability between members of the same block, we must choose the degree-corrected formulation (DC-SBM).
+To allow for degree-variability between members of the same block, we adopt the degree-corrected formulation (DC-SBM), defined in~(\ref{defn:microcan-dc-sbm}).
+
+For each integer $K\geq 1$, we use the notation $[K]:=\{1,2,\ldots,K\}$.
 
@@
 	\label{defn:microcan-dc-sbm}
-	Let $N \in \Integers^{+}$ denote the number of vertices in an undirected graph.
-The block memberships are encoded by a vector $b \in [B]^N$,\footnote{We introduce the notation $[K] \coloneqq \{1, 2 \dots K\}$ to compactly define a set of $K$ indices.
-Clearly, $[K]$ is only defined for $K \in \Integers^+$.}
-	where $B \in \Integers^{+}$ is the number of non-empty blocks.
-	Let $e \in (\Integers_0^+)^{N \times N}$ be the matrix of edge counts between blocks.
-$e_{rs}$ is then the number of edges from block $r$ onto block $s$ -- or twice that number if $r=s$.
-For undirected graphs, $e$ is symmetric.
-	Let $k \in (\Integers^+_0)^N$ be a vector denoting the degree sequence of the graph.
-$k_i$ is then the degree of vertex $i$.
-The graph's adjacency matrix $A \in \{0,1\}^{N \times N}$ is generated as follows:
+	Let $N \geq 1$ denote the number of vertices in an undirected graph. The block memberships are encoded by a vector $b \in [B]^N$,
+where $B$ is the number of non-empty blocks.
+	Let $e=(e_{rs})$ be the $N\times N$ symmetric matrix of edge counts 
+between blocks, such that $e_{rs}$ is the number of edges from block $r$ to 
+block $s$ -- or twice that number if $r=s$. 
+	Let $k =(k_i)$ denote the vector of length $N$ containing the degree
+sequence of the graph, with $k_i$ being the degree of vertex $i$.
+
+The graph's adjacency matrix $A \in \{0,1\}^{N \times N}$ is generated 
+by placing edges uniformly at random, but respecting the constraints 
+imposed by $e$, $b$ and $k$ (hence the qualification `microcanonical').
+Specifically, $A$ must satisfy the following, for all $r,s\in[B]$
+and all $i\in[N]$:
 	%
 	\begin{equation}
-		A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, e, k)
-	\end{equation}
-	%
-	Where edges are placed uniformly at random but respecting the constraints imposed by $e$, $b$ and $k$ -- hence the microcanonical moniker. Specifically, $A$ must satisfy the following:
-	%
-	\begin{equation}
-		e_{rs} = \sum_{i, j \in [N]} A_{ij} \one \{b_i = r\} \one \{b_j = s\} \quad \forall r,s \in [B]
-		\quad \textrm{and} \quad
-		k_i = \sum_{j \in [N]} A_{ij} \quad \forall i \in [N]
+		e_{rs} = \sum_{i, j \in [N]} A_{ij} 
+	\one \{b_i = r\} \one \{b_j = s\} 
+		\qquad 
+		\textrm{and} \qquad
+		k_i = \sum_{j \in [N]} A_{ij}.
 		\label{eqn:sbm-constraints}
 	\end{equation}
+We use the following notation to indicate this distribution:
+\begin{equation}
+		A \sim \mbox{\rm DC-SBM}_{\rm MC} (b, e, k).
+	\end{equation}
 \end{definition}
+
+Appendix \ref{appdx:sbm} describes how to compute the likelihood, $p(A|b,e,k)$, for graphs generated from the microcanonical DC-SBM.

### 2021-05-29 cued/tech-abstract.tex
More appendix

@@
 A typical inference goal is to determine how the vertex labels
-(or {\em features}) affect the network's graph structure. A standard
-approach has been to partition the network into blocks grouped
+(or {\em features}) affect the network's graph structure.
+
+A standard approach has been to partition the network into blocks grouped
 by distinct values of the feature of interest. A block-based random
@@
 feature-based communities. Nevertheless, the resulting communities
-often do not produce a natural partition of the graph. In this work,
-we introduce a new generative model, the feature-first block model (FFBM),
+often do not produce a natural partition of the graph.
+
+In this work, we introduce a new generative model, the feature-first block model (FFBM),
 which is more effective at describing vertex-labelled undirected
@@
 deliberately simple to retain easy interpretability of the parameter
-values. We apply the proposed methods to a variety of network data
+values.
+
+We apply the proposed methods to a variety of network data
 to extract the most important features along which the vertices
 are partitioned. The main advantages of the proposed approach are
-that the whole feature-space is used automatically, and features
+that the whole feature-space is used automatically and that features
 can be rank-ordered implicitly according to impact. Any features

### 2021-05-30 cued/experiments.tex
Add accuracy discussion

--- 2021-05-27 nips/experiments.tex
@@
 
-\begin{itemize}[leftmargin=*]
+\begin{itemize}
 	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=3$) -- network of Amazon book sales about U.S.
@@
 	\label{fig:graphs-all}
-\end{figure}
-\begin{figure}[!h]
-	\centering
-	\begin{subfigure}[t]{0.32\linewidth}
-		\centering
-		\vskip 0pt
-		\includegraphics[width=\linewidth]{polbooks-null.png}
-		\caption{Polbooks}
-		\label{fig:polbooks-null}
-	\end{subfigure}
-	\hfill
-	\begin{subfigure}[t]{0.32\linewidth}
-		\centering
-		\vskip 0pt
-		\includegraphics[width=\linewidth]{school-null.png}
-		\caption{School}
-		\label{fig:school-null}
-	\end{subfigure}
-	\hfill
-	\begin{subfigure}[t]{0.32\linewidth}
-		\centering
-		\vskip 0pt
-		\includegraphics[width=\linewidth]{fb-null.png}
-		\caption{Facebook egonet}
-		\label{fig:fb-null}
-	\end{subfigure}
-	\caption{Reduced dimension feature-to-block generator weight samples. Dotted line is $\pm c^*$}
-	\label{fig:null-all}
 \end{figure}
@@
 
+Nevertheless, the cross-entropy loss over the whole training set may be too coarse a measure of model fit.
+It is often the case that we have good feature-based explanations for some but not all of the detected blocks.
+We wish to define a new measure of fit specific to each detected block.
+We could use a binary cross-entropy function but this would necessarily use the whole test/training set and so few vertices will belong to the block of interest.
+We instead prefer to compute quantities over the set of all vertices associated with block $j$, defined as,
+%
+\begin{equation}
+	\Bcal_\star(j) \coloneqq \{i \in \Gcal_\star : \hat{b}_i = j\} \quad \textrm{where} 
+	\quad \hat{b}_i \coloneqq \argmax_j \hat{y}_{ij}.
+\end{equation}
+%
+Recall that $\hat{y}_{ij}$ is our estimate for the block membership posterior~(\ref{eqn:y-hat}) using only information from the adjacency matrix $A$.
+Again, $\star \in \{0, 1\}$ toggles between training and test sets.
+Now $\Bcal_\star(j)$ is the set of all vertices within the training or test set that have maximum a posteriori probability of belonging to set $j$.
+We choose to resolve ties consistently by picking the lower index.
+We now define the accuracy for block $j$ as,
+%
+\begin{equation}
+	\eta_\star(j) \coloneqq \frac{1}{|\Bcal_\star (j)| \cdot 
+	|\Tcal_\theta| } 
+	\sum_{i \in \Bcal_\star (j)}  \sum_{t \in \Tcal_\theta}
+	\one \left\{\hat{b}_i = \argmax_j \phi_j \left( x_i; \theta^{(t)} \right) \right\}.
+	\label{eqn:accuracy}
+\end{equation}
+%
+This is effectively testing whether the feature-to-block predictions and the graph-based predictions agree in their largest component.
+We call this metric, $\eta_\star(j)$, the {\em block-accuracy} for block for $j$.
+
+For each of the experiments, we plot the collected $\theta$-samples for the features that survive the dimensionality reduction procedure.
+We also plot the per-block accuracy for the original-dimension classifiers.
+
 \subsection{Political books}
@@
 determine whether the authors' political affiliations are a good predictor 
-of the overall network structure.
-We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral).
+of the overall network structure. We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral).
+%
+\begin{figure}[!h]
+	\centering
+	\begin{subfigure}[t]{0.45\linewidth}
+		\centering
+		\vskip 0pt
+		\includegraphics[width=\linewidth]{polbooks-null}
+		\caption{$\theta$-samples}
+		\label{fig:polbooks-null}
+	\end{subfigure}
+	\begin{subfigure}[t]{0.45\linewidth}
+		\centering
+		\vskip 0pt
+		\includegraphics[width=\linewidth]{polbooks-null}
+		\caption{Per-block accuracy $\eta(j)$}
+		\label{fig:polbooks-accuracy}
+	\end{subfigure}
+	\caption{Political books dataset}
+	\label{fig:polbooks}
+\end{figure}
+%
 From Figure~\ref{fig:polbooks-null} we see that all 3 blocks have a distinct political affiliation as their largest positive component. 
@@
 meaning that these are not good predictors of overall macro-structure.
-
+%
+\begin{figure}[!h]
+	\centering
+	\begin{subfigure}[t]{0.45\linewidth}
+		\centering
+		\vskip 0pt
+		\includegraphics[width=\linewidth]{school-null}
+		\caption{$\theta$-samples}
+		\label{fig:school-null}
+	\end{subfigure}
+	\begin{subfigure}[t]{0.45\linewidth}
+		\centering
+		\vskip 0pt
+		\includegraphics[width=\linewidth]{school-null}
+		\caption{Per-block accuracy $\eta(j)$}
+		\label{fig:school-accuracy}
+	\end{subfigure}
+	\caption{Primary school dynamic contacts dataset}
+	\label{fig:school}
+\end{figure}
+%
 The vast majority of blocks are composed of a single class. 
@@
 Figure~\ref{fig:fb-null} do not have high-magnitude components for $D'=10$.
-
+%
+\begin{figure}[!h]
+	\centering
+	\begin{subfigure}[t]{0.45\linewidth}
+		\centering
+		\vskip 0pt
+		\includegraphics[width=\linewidth]{fb-null}
+		\caption{$\theta$-samples}
+		\label{fig:fb-null}
+	\end{subfigure}
+	\begin{subfigure}[t]{0.45\linewidth}
+		\centering
+		\vskip 0pt
+		\includegraphics[width=\linewidth]{fb-null}
+		\caption{Per-block accuracy $\eta(j)$}
+		\label{fig:fb-accuracy}
+	\end{subfigure}
+	\caption{Facebook egonet dataset}
+	\label{fig:fb}
+\end{figure}
+%
 When the feature dimension is very large, it becomes increasingly likely that 

### 2021-05-30 cued/experiments.tex
Add accuracy

@@
 		\includegraphics[width=\linewidth]{polbooks-graph.png}
-		\caption{Polbooks}
+		\caption{Polbooks.}
 		\label{fig:polbooks-graph}
@@
 		\includegraphics[width=\linewidth]{school-graph.png}
-		\caption{School}
+		\caption{School.}
 		\label{fig:school-graph}
@@
 		\includegraphics[width=\linewidth]{fb-graph.png}
-		\caption{Facebook egonet}
+		\caption{Facebook egonet.}
 		\label{fig:fb-graph}
@@
 		\includegraphics[width=0.8\linewidth]{10-vertical-legend.png}
-		\caption{Legend}
+		\caption{Legend.}
 		\label{fig:10-legend}
@@
 	\centering
-	\begin{subfigure}[t]{0.45\linewidth}
+	\begin{subfigure}[t]{0.452\linewidth}
 		\centering
 		\vskip 0pt
-		\includegraphics[width=\linewidth]{polbooks-null}
-		\caption{$\theta$-samples}
+		\includegraphics[width=\linewidth]{polbooks-null-1}
+		\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
 		\label{fig:polbooks-null}
@@
 		\vskip 0pt
-		\includegraphics[width=\linewidth]{polbooks-null}
-		\caption{Per-block accuracy $\eta(j)$}
+		\includegraphics[width=\linewidth]{polbooks-accuracy-1}
+		\caption{Per-block accuracy $\eta(j)$.}
 		\label{fig:polbooks-accuracy}
 	\end{subfigure}
-	\caption{Political books dataset}
+	\caption{Political books dataset.}
 	\label{fig:polbooks}
 \end{figure}
-%
+
 From Figure~\ref{fig:polbooks-null} we see that all 3 blocks have a distinct political affiliation as their largest positive component. 
@@
 variable for the overall network structure.
+
+Nevertheless, from Figure~\ref{fig:polbooks-accuracy} we see that block 1 has low accuracy.
+This suggests that detected block 1 is not solely composed of ``neutral" books but also contains some ``liberal" and ``conservative" authors.
+Indeed, by examining Figure~\ref{fig:polbooks-graph} we see that block 1 is effectively a buffer between blocks 2 and 3.
+It is therefore not surprising that some books from either side leak into block 1.
+Perhaps, the three distinct categories (``conservative", ``liberal", ``neutral") are too coarse a measure of political affiliation; it may be the case that the nominally ``conservative" books found in block 1 are in fact more centre-right.
+It would be interesting to rerun the experiment using soft labels for political affiliation now encoded as a continuous variable $(-1,1)$.
+Nevertheless, political affiliation encoded as 3 distinct labels remains a fantastic predictor of network structure.
 
@@
 		\centering
-		\vskip 0pt
-		\includegraphics[width=\linewidth]{school-null}
-		\caption{$\theta$-samples}
+		\imagebox{6.2cm}{\includegraphics[width=\linewidth]{school-null-1}}
+		\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
 		\label{fig:school-null}
@@
 		\centering
-		\vskip 0pt
-		\includegraphics[width=\linewidth]{school-null}
-		\caption{Per-block accuracy $\eta(j)$}
+		\imagebox{6.2cm}{\includegraphics[width=\linewidth]{school-accuracy-1}}
+		\caption{Per-block accuracy $\eta(j)$.}
 		\label{fig:school-accuracy}
 	\end{subfigure}
-	\caption{Primary school dynamic contacts dataset}
+	\caption{Primary school dynamic contacts dataset.}
 	\label{fig:school}
@@
 		\centering
-		\vskip 0pt
-		\includegraphics[width=\linewidth]{fb-null}
-		\caption{$\theta$-samples}
+		\imagebox{7cm}{\includegraphics[width=\linewidth]{fb-null-1}}
+		\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
 		\label{fig:fb-null}
@@
 		\centering
-		\vskip 0pt
-		\includegraphics[width=\linewidth]{fb-null}
-		\caption{Per-block accuracy $\eta(j)$}
+		\imagebox{7cm}{\includegraphics[width=\linewidth]{fb-accuracy-1}}
+		\caption{Per-block accuracy $\eta(j)$.}
 		\label{fig:fb-accuracy}
 	\end{subfigure}
-	\caption{Facebook egonet dataset}
+	\caption{Facebook egonet dataset.}
 	\label{fig:fb}

### 2021-05-31 cued/experiments.tex
Finish

@@
 		\includegraphics[width=\linewidth]{polbooks-graph.png}
-		\caption{Polbooks.}
+		\caption{Polbooks}
 		\label{fig:polbooks-graph}
@@
 		\includegraphics[width=\linewidth]{school-graph.png}
-		\caption{School.}
+		\caption{School}
 		\label{fig:school-graph}
@@
 		\includegraphics[width=\linewidth]{fb-graph.png}
-		\caption{Facebook egonet.}
+		\caption{Facebook egonet}
 		\label{fig:fb-graph}
 	\end{subfigure}
-	\begin{subfigure}[t]{0.10\linewidth}
+	\begin{subfigure}[t]{0.11\linewidth}
 		\centering
 		\includegraphics[width=0.8\linewidth]{10-vertical-legend.png}
-		\caption{Legend.}
+		\caption{Legend}
 		\label{fig:10-legend}
@@
 We wish to define a new measure of fit specific to each detected block.
-We could use a binary cross-entropy function but this would necessarily use the whole test/training set and so few vertices will belong to the block of interest.
-We instead prefer to compute quantities over the set of all vertices associated with block $j$, defined as,
+This requires defining the set of all vertices associated with block $j$, as,
 %
@@
 This is effectively testing whether the feature-to-block predictions and the graph-based predictions agree in their largest component.
-We call this metric, $\eta_\star(j)$, the {\em block-accuracy} for block for $j$.
+We call this metric, $\eta_\star(j)$, the {\em block-accuracy} for block $j$.
+It is clearly bounded $0 \leq \eta_\star(j) \leq 1$, with an accuracy of 1 meaning perfect agreement for the vertices in detected block $j$.
 
@@
 We also plot the per-block accuracy for the original-dimension classifiers.
+We discuss the results specific to each dataset in turn.
 
@@
 
-Nevertheless, from Figure~\ref{fig:polbooks-accuracy} we see that block 1 has low accuracy.
+However, from Figure~\ref{fig:polbooks-accuracy} we see that block 1 has low accuracy.
 This suggests that detected block 1 is not solely composed of ``neutral" books but also contains some ``liberal" and ``conservative" authors.
-Indeed, by examining Figure~\ref{fig:polbooks-graph} we see that block 1 is effectively a buffer between blocks 2 and 3.
+Indeed, by examining Figure~\ref{fig:polbooks-graph} we see that block 1 is effectively a buffer between blocks 2 and 3 -- there are very few edges between blocks 2 and 3.
 It is therefore not surprising that some books from either side leak into block 1.
 Perhaps, the three distinct categories (``conservative", ``liberal", ``neutral") are too coarse a measure of political affiliation; it may be the case that the nominally ``conservative" books found in block 1 are in fact more centre-right.
-It would be interesting to rerun the experiment using soft labels for political affiliation now encoded as a continuous variable $(-1,1)$.
+In the absence of more granular labels we cannot test this hypothesis.
 Nevertheless, political affiliation encoded as 3 distinct labels remains a fantastic predictor of network structure.
@@
 \end{figure}
-%
+
 The vast majority of blocks are composed of a single class. 
 However, some blocks have two comparably good classes as their predictor. 
-For example, block 2 contains classes 3A and 3B as its 2 best predictors. 
+For example, both blocks 2 and 5 contain classes 3A and 3B as their 2 best predictors. 
 This suggests that the social divide between classes is less pronounced 
 for pupils in year 3. Conversely, some classes are found to extend over two 
-detected blocks (class 2B spans blocks 6 and 7) but we nonetheless do 
+detected blocks (class 2B spans blocks 8 and 9) but we do 
 not have a feature which explains the division. The most surprising block 
-is number 5, which has comparable weightings for classes 5A and 1B. 
+is number 7, which has comparable weightings for classes 5A and 1B. 
 Perhaps there was a joint event between those two classes on the day 
-the data were collected.
+the data were collected. Nevertheless, a hierarchical approach would help further subdivide these blocks.
+
+Figure~\ref{fig:school-accuracy} shows excellent accuracy for the majority of blocks.
+In fact the only blocks with low accuracy are those that have a school-class span two blocks such that we cannot reliably distinguish between the two.
+This is much more pronounced when we apply hard classification rather than the soft cross-entropy loss.
+Indeed block 1 has low accuracy because the class 1A has a higher weight component for block 3 than it does for block 1.
+As we use hard classification to compute accuracy, this means vertices belonging to class 1A are assigned to block 3.
+A similar effect is seen in block 5 which often loses out to block 2.
+Depending on the sample index $t$, class 2B is assigned to block 8 or 9 with maximal probability.
+This explains the rough 50\% accuracy for blocks 8 and 9.
 
@@
 Figure~\ref{fig:fb-null} do not have high-magnitude components for $D'=10$.
+This is further emphasised by the disparate accuracies in Figure~\ref{fig:fb-accuracy}.
+Nevertheless, observe that the accuracy is only high for blocks that contain high magnitude weights.
+The only exception to this rule is block 9 which nonetheless, although this may have high magnitude weights just below the cut-off $c^*$.
+Recall that the accuracy is computed as ana average over all samples and so is likely to be lower than the MAP solution.
 %
@@
 \end{figure}
-%
+
 When the feature dimension is very large, it becomes increasingly likely that 
@@
 that birthdays determine graphical structure.
-% The analyst must remain vigilant of such problems.
+
+However, we see the advantage of having $\Xcal=\{0,1\}$ rather than $\Xcal={-1, 1}$ as argued in Appendix \ref{appdx:dimension}.
+Block 1 in Figure~\ref{fig:fb-null} is allowed to contain 3 distinct feature categories.
+Nevertheless, one would have to apply a hierarchical approach to subdivide block 1 into smaller blocks to test whether these features determine structure at the lower-level.

### 2021-05-31 cued/experiments.tex
Modifications

@@
 Perhaps, the three distinct categories (``conservative", ``liberal", ``neutral") are too coarse a measure of political affiliation; it may be the case that the nominally ``conservative" books found in block 1 are in fact more centre-right.
-In the absence of more granular labels we cannot test this hypothesis.
+In the absence of more granular labels, we cannot test this hypothesis.
 Nevertheless, political affiliation encoded as 3 distinct labels remains a fantastic predictor of network structure.
@@
 The vast majority of blocks are composed of a single class. 
-However, some blocks have two comparably good classes as their predictor. 
-For example, both blocks 2 and 5 contain classes 3A and 3B as their 2 best predictors. 
+However, some blocks have two comparably strong classes as their predictors. 
+For example, blocks 2 and 5 both contain classes 3A and 3B as their 2 best predictors. 
 This suggests that the social divide between classes is less pronounced 
@@
 Perhaps there was a joint event between those two classes on the day 
-the data were collected. Nevertheless, a hierarchical approach would help further subdivide these blocks.
+the data were collected.
 
@@
 This is much more pronounced when we apply hard classification rather than the soft cross-entropy loss.
-Indeed block 1 has low accuracy because the class 1A has a higher weight component for block 3 than it does for block 1.
-As we use hard classification to compute accuracy, this means vertices belonging to class 1A are assigned to block 3.
+Indeed block 1 has low accuracy because class 1A has a higher weight component for block 3 than it does for block 1.
+As we use hard classification to compute accuracy, vertices belonging to class 1A are predicted to belong to block 3.
 A similar effect is seen in block 5 which often loses out to block 2.
-Depending on the sample index $t$, class 2B is assigned to block 8 or 9 with maximal probability.
-This explains the rough 50\% accuracy for blocks 8 and 9.
+The same can be said of blocks 8 and 9.
+However, in this case the weights for class 2B are very similar -- thus explaining the roughly evenly split accuracy for blocks 8 and 9.
 
@@
 Nevertheless, observe that the accuracy is only high for blocks that contain high magnitude weights.
-The only exception to this rule is block 9 which nonetheless, although this may have high magnitude weights just below the cut-off $c^*$.
-Recall that the accuracy is computed as ana average over all samples and so is likely to be lower than the MAP solution.
+The only exception to this rule is block 9 which nonetheless may have high magnitude weights just below the cut-off $c^*$.
 %
@@
 a particular feature may uniquely identify a small set of nodes. If these nodes 
-are all part of the same community, then the classifier will overfit for that 
+are all part of the same community, then the classifier may overfit for that 
 particular parameter. The regularisation term imposed by the prior goes some 
@@
 that birthdays determine graphical structure.
-
-However, we see the advantage of having $\Xcal=\{0,1\}$ rather than $\Xcal={-1, 1}$ as argued in Appendix \ref{appdx:dimension}.
-Block 1 in Figure~\ref{fig:fb-null} is allowed to contain 3 distinct feature categories.
-Nevertheless, one would have to apply a hierarchical approach to subdivide block 1 into smaller blocks to test whether these features determine structure at the lower-level.
+This issue could have been alleviated by choosing $\Xcal=\{-1, +1\}$ such that having a feature turned off supports excluding a vertex from the block.
+However, Figure~\ref{fig:fb-null} demonstrates why we made the deliberate choice $\Xcal=\{0,1\}$, as argued in Appendix \ref{appdx:dimension}.
+Because of the choice $\Xcal=\{0,1\}$, block 1 is allowed to contain 3 distinct feature categories rather than just one.
+We accept this tradeoff: risking spurious high magnitude components (e.g.
+\verb*|birthday-5|) for the benefit of allowing a top-level block span multiple feature-groups.
+Nevertheless, one would have to apply a hierarchical approach to subdivide each top-level block into sub-blocks to test whether the constituent feature-groups determine structure at the lower-level.

### 2021-05-31 cued/tech-abstract.tex
Modifications

@@
 in the resulting network, the approach detects this and is protected
-against over-fitting. Results on several real-world datasets
-illustrate the performance of the proposed methods.
+against over-fitting.
+
+Future work on this topic could benefit from extending the FFBM to be hierarchical in nature.
+Such a modification would be relatively natural and would increase the descriptive power of the model.
+The current FFBM formulation is limited in that it can only explain structure at the macro-scale.
+Nevertheless, a hierarchical formulation would allow explanations of structure at any arbitrary length-scale.

### 2021-06-01 cued/conclusion.tex
Proof-read main report

--- 2021-05-27 nips/conclusion.tex
@@
 in this paper is a new generative model for labelled networks,
-developed in order to address difficulties of other graphical models 
+developed to address difficulties of other graphical models 
 when testing how vertex features affect community structure. The idea 
-is to divide the graph into its most natural partition and test whether 
+is to divide the graph into its most natural partition and determine whether 
 the vertex features can accurately explain this partition. 
@@
 we go on to describe an efficient inference algorithm to sample 
-the parameters $\theta$ of the feature-to-block generator. 
+the parameters of the FFBM. 
 This takes the form of a two-level Markov chain,

### 2021-06-01 cued/experiments.tex
Proof-read main report

@@
 denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively. 
-These values are also shown in Table~\ref{tab:results}.
+These values are also given in Table~\ref{tab:results}.
 
@@
 However, from Figure~\ref{fig:polbooks-accuracy} we see that block 1 has low accuracy. 
-This suggests that detected block 1 is not solely composed of ``neutral" books but also contains some ``liberal" and ``conservative" authors.
-Indeed, by examining Figure~\ref{fig:polbooks-graph} we see that block 1 is effectively a buffer between blocks 2 and 3 -- there are very few edges between blocks 2 and 3.
-It is therefore not surprising that some books from either side leak into block 1.
-Perhaps, the three distinct categories (``conservative", ``liberal", ``neutral") are too coarse a measure of political affiliation; it may be the case that the nominally ``conservative" books found in block 1 are in fact more centre-right.
+This suggests that detected block 1 is not solely composed of ``neutral" books but also 
+contains some ``liberal" and ``conservative" authors. Indeed, by examining 
+Figure~\ref{fig:polbooks-graph} we see that block 1 is effectively a buffer between 
+blocks 2 and 3; there are very few edges between blocks 2 and 3. It is therefore not 
+surprising that some books from either side leak into block 1.
+Perhaps, the three distinct categories (``conservative", ``liberal", ``neutral") are 
+too coarse a measure of political affiliation; it may be the case that the nominally 
+``conservative" books found in block 1 are in fact more centre-right. 
 In the absence of more granular labels, we cannot test this hypothesis. 
-Nevertheless, political affiliation encoded as 3 distinct labels remains a fantastic predictor of network structure.
+Nevertheless, political affiliation encoded as 3 distinct labels 
+remains a fantastic predictor of network structure.
 
@@
 high weight as it relates to block 1 -- but it is highly unlikely
-that birthdays determine graphical structure.
-This issue could have been alleviated by choosing $\Xcal=\{-1, +1\}$ such that having a feature turned off supports excluding a vertex from the block.
-However, Figure~\ref{fig:fb-null} demonstrates why we made the deliberate choice $\Xcal=\{0,1\}$, as argued in Appendix \ref{appdx:dimension}.
-Because of the choice $\Xcal=\{0,1\}$, block 1 is allowed to contain 3 distinct feature categories rather than just one.
-We accept this tradeoff: risking spurious high magnitude components (e.g.
-\verb*|birthday-5|) for the benefit of allowing a top-level block span multiple feature-groups.
-Nevertheless, one would have to apply a hierarchical approach to subdivide each top-level block into sub-blocks to test whether the constituent feature-groups determine structure at the lower-level.
+that birthdays determine graphical structure. This issue could have been alleviated by 
+choosing $\Xcal=\{-1, +1\}$ such that having a feature turned off supports excluding 
+a vertex from the block. However, Figure~\ref{fig:fb-null} demonstrates why we made 
+the deliberate choice $\Xcal=\{0,1\}$, as argued in Appendix \ref{appdx:dimension}. 
+Because of the choice $\Xcal=\{0,1\}$, block 1 is allowed to contain 3 distinct feature 
+categories rather than just one. We accept this trade-off: risking spurious high magnitude 
+components (e.g. \verb*|birthday-5|) for the benefit of allowing a top-level block to span 
+multiple feature-groups. Nevertheless, one would have to apply a hierarchical approach 
+to subdivide each top-level block into sub-blocks to test whether the constituent 
+feature-groups determine structure at the lower-level.

### 2021-06-01 cued/ffbm.tex
More appendix
Proof-read main report

--- 2021-05-28 nips/ffbm.tex
@@
 
-Let $N$ denote the number of nodes, $B$ the number of blocks in the graph,
+Let $N$ denote the number of vertices, $B$ the number of blocks in the graph,
 and ${\cal X}$ the set of values each feature can take.
@@
 \begin{equation}
-	p(b| X, \theta) = \prod_{i=1}^{N} p(b_i | x_i, \theta) = \prod_{i=1}^{N} \phi_{b_i} (x_i; \theta)
-	= \prod_{i=1}^{N} \frac{\exp\left(w_{b_i}^T x_i\right)}{\sum_{k=1}^{B} \exp \left( w_k^T x_i\right)}.
+	p(b| X, \theta) = \prod_{i \in [N]} p(b_i | x_i, \theta) = \prod_{i \in [N]} \phi_{b_i} (x_i; \theta)
+	= \prod_{i \in [N]} \frac{\exp\left(w_{b_i}^T x_i\right)}{\sum_{k \in [B]} \exp \left( w_k^T x_i\right)}.
 \end{equation}
 %
-$\phi_{b_i}$ is the form of a softmax activation function; we deliberately exclude a bias term to ensure that we only leverage feature information. 
+Note that $\phi_{b_i}$ has the form of a softmax activation function;
+we deliberately exclude a bias term to ensure that we only leverage feature information.
+This decision is justified in greater detail in Appendix~\ref{appdx:dimension}.
 We will later also find it convenient
@@
 independent given $b$.  The exact form of the prior
-$p(\psi|b)$ is described in Appendix \ref{appdx:prior}.
+$p(\psi|b)$ is described in Appendix \ref{appdx:sbm-prior}.
 All that concerns the main argument is that it has

### 2021-06-01 cued/inference.tex
More appendix
Add accuracy discussion
Fix misc errors
Proof-read main report

@@
 %
-Both of these sampling steps can be implemented using
-Markov chain Monte Carlo via a Metropolis-Hastings 
-algorithm \cite{hastings-alg}, as a two-level Markov chain.
+Both of these sampling steps can be implemented through a
+Markov chain Monte-Carlo approach via a Metropolis-Hastings 
+algorithm \cite{hastings-alg}, as a two-level Markov chain. See Appendix \ref{appdx:metropolis-hastings} for more details on the Metropolis-Hastings algorithm.
 
-% We just need to define a proposal distribution $q(x, x')$ for proposing a move $x \rightarrow x'$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(\cdot)$, point-wise.
-The proposed move is then accepted with probability $\alpha$ (equation \ref{eqn:mh-accept}) else it is rejected and we stay at $x$.
-%
-% \begin{equation}
-	% \alpha(x, x') = \min \left( \frac{\pi(x') q(x', x)}{\pi(x) q(x, x')} , 1 \right)
-	% \label{eqn:mh-accept}
-% \end{equation}
-%
-% This accept-reject step ensures the resulting Markov Chain is in detailed balance with the target distribution $\pi(\cdot)$. 
-% What we propose in equations \ref{eqn:b-samples} and \ref{eqn:theta-samples} is therefore implemented through a 2-level Markov chain. 
 As we show in the following section, the resulting samples 
@@
 
-		\draw[->] (t0) to node[above] {$\sim q_\theta(\theta, \theta' | X, b^{(t)})$} (t1);
+		\draw[->] (t0) to node[above] {$\sim q_\theta(\theta^{(t)}, \theta' | X, b^{(t)})$} (t1);
 		\draw[->] (t1) to node[above] {$\alpha_\theta (\theta, \theta' | X, b^{(t)})$} (t2);
@@
 
-The reason for splitting the Markov chain into two levels is that the summation over all latent states $b \in [B]^N$ required to directly compute the likelihood, $p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$,
-is intractable, as the sum involves $B^N$ terms. 
+The reason for splitting the Markov chain into two levels is that the summation over all latent states $b \in [B]^N$ required to directly compute the likelihood, $p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$, is intractable, as the sum involves $B^N$ terms.
 Figure \ref{fig:samp-sequence} shows an overview of the proposed method. 
@@
 \FloatBarrier
-\clearpage
 \subsection{Sampling block memberships}
@@
 %
-where we write $U(\theta)$ for the negative log-posterior. Let $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$. 
-Discarding constant terms, we can write $U(\theta)$ as,
+where we write $U(\theta)$ for the negative log-posterior. We define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$. 
+Discarding constant terms, we can then write $U(\theta)$ as,
 %
@@
 The function $U(\theta)$ is a typical objective function for neural network training.
-The first term is introduced by the likelihood.
-We collect it into $N \cdot \Lcal(\theta)$, which is the cross-entropy between the graph-predicted and feature-predicted block memberships summed over all vertices.
+The first term is introduced by the likelihood; we collect it into $N \cdot \Lcal(\theta)$, which is the cross-entropy between the graph-predicted and feature-predicted block memberships summed over all vertices.
 The second, introduced by the prior, brings a form of regularisation, guarding against over-fitting. Unlike in many applications where the goal 
@@
 a new sample $\theta'$ as,
+%
 \begin{equation*}
@@
 \end{equation*}
-
+%
 where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter 
-which may vary with the sample index (see Appendix \ref{appdx:step-size}).
+which may vary with the sample index (see Appendix \ref{appdx:step-size} for a discussion of step-size).
 This leads to the proposal distribution,
@@
 \end{equation*}
-
+%
 Without the injected noise term, MALA is equivalent to gradient descent. We require the noise term $\xi$ to fully explore the parameter space. 
@@
 \begin{equation}
-	\frac{\partial U}{\partial w_k} = - \left( \sum_{i=1}^{N} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right).
+	\frac{\partial U}{\partial w_k} = - \left( \sum_{i \in [N]} \Big\{ x_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right).
 	\label{eqn:U-derivative}
@@
 	\hat{y}_{ij} \coloneqq \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} y_{ij}^{(t)} = \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} \one\{b_i^{(t)} = j\}.
+	\label{eqn:y-hat}
 \end{equation}
@@
 have been obtained, we can compute the empirical mean and standard deviation of each component of $\theta$.
-Switching back to matrix notation we define $\theta = W$, such that $W_{ij}$ is the weight component for block $i$ and feature $j$, we can define:
+Switching back to matrix notation we define $\theta = W$, such that $W_{ij}$ is the weight component for block $i$ and feature $j$.
+We can then define:
 %
@@
 	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
-	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2
+	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2.
 \end{equation}
@@
 Intuitively, this means discarding any feature $j$ for which 
-$\hat{\mu}_{ij} \pm k\hat{\sigma}_{ij}$ lies outside the region
+$(\hat{\mu}_{ij} - k\hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \sigma)$ overlaps with
 $(-c, c)$ for all block indices $i$.
 If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \hat{\mu}_{ij}, \hat{\sigma}_{ij}^2)$, then this is analogous to a hypothesis test on the value of $W_{ij}$ as in~(\ref{eqn:hyp-test-discard});
-then $\Dcal'$ comprises all features $i$ for which $H_1$ is not rejected at least once for some $j \in [B]$
+then $\Dcal'$ comprises all features $j$ for which $H_1$ is accepted at least once for some $i \in [B]$.
 %
@@
 %
-The multiplier $k$ in equation \ref{eqn:fij} determines the degree of significance of the result.
+The multiplier $k$ in~(\ref{eqn:fij}) determines the degree of significance of the result.
 However, as the Laplace approximation is not exact we will only treat this dimensionality reduction method as a useful heuristic and not an exact method.
+Appendix~\ref{appdx:hyp-test} explores this subject at greater depth.
+
 Conversely, we could fix $k=k_0$ and the dimension of our reduced feature set $|\Dcal'|=D'$.

### 2021-06-01 cued/tech-abstract.tex
Proof-read main report

@@
 graph model -- typically a variant of the stochastic block model --
-is then used to test for evidence of asymmetric behaviour within these
+is then used to test for evidence of asymmetric behaviour between these
 feature-based communities. Nevertheless, the resulting communities

### 2021-06-01 cued/additional.tex
Fix misc errors
Proof-read

@@
 
-\subsection{SBM likelihood and prior}
+\subsection{SBM likelihood}
 \label{appdx:sbm-likelihood}
@@
 \begin{equation}
-	\Omega(e) = \frac{\prod_{r} e_r !}{\prod_{r,s : r < s} e_{rs}! \cdot \prod_{r} e_{rr}!!},
+	\Omega(e) = \frac{\prod_{r} e_r !}{\prod_{r,s : r < s} e_{rs}! \prod_{r} e_{rr}!!},
 \end{equation}
@@
 $E = \frac{1}{2} \sum_{r,s} e_{rs}$ is the total number of edges in the graph. 
-Importantly, $E$ is not allowed to vary and so $p(e|b)$ is uniform in $e$.
+Importantly, $E$ is fixed by $A$ and so $p(e|b)$ is uniform in $e$.
 The variable $\eta_j^r$ denotes the number of vertices in block $r$ 
@@
 is the number of vertices assigned to block $r$.
-
 These priors were chosen carefully in \cite{Peixoto-Bayesian-Microcanonical} to 
@@
 Suppose we wish to draw samples for some general $x$ from the normalised density $\pi^*(x)$.
-We just need to define a proposal distribution $q(x, x')$ for proposing a move $x \rightarrow x'$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(x) \propto \pi*(x)$, point-wise.
+We just need to define a proposal distribution $q(x, x')$ and be able to evaluate an un-normalised form of the target distribution, denoted $\pi(x) \propto \pi^*(x)$, point-wise.
 Each proposed move $x \rightarrow x'$ is then accepted with probability $\alpha$~(\ref{eqn:mh-accept}) else it is rejected and we stay at $x$.
-
+%
 \begin{equation}
@@
 \end{equation}
-
-This accept-reject step ensures the resulting Markov Chain is in detailed balance with the target distribution $\pi(\cdot)$. Therefore, the set of samples $\{x^{(t)}\} \sim \pi^*(x)$
+%
+This accept-reject step ensures the resulting Markov Chain is in detailed balance with the target distribution $\pi(\cdot)$. Therefore, the set of samples produced $\{x^{(t)}\} \sim \pi^*(x)$.
 
@@
 	\includegraphics[width=0.6\linewidth]{step-size-tradeoff}
-	\caption{Primary school dataset \cite{schools} illustration of step-size tradeoff}
+	\caption{Primary school dataset illustration of step-size tradeoff.}
 	\label{fig:step-size-tradeoff}
@@
 
-%As with any MCMC method, we must deal with the issues presented by burn-in and thinning.
 When sampling from the $b$ and $\theta$-chains described
@@
 We discard an initial proportion $\kappa_\star\in(0,1)$ of the samples 
-as corresponding to a ``burn-in'' period required for the distribution 
-of the chain to reach a distribution close to our target,
+as corresponding to a ``burn-in'' period required for the chain to equilibrate,
 and we also ``thin'' the remaining samples to 
@@
 where $\lambda_\star$ controls the thinning. The choice of
-$\kappa_\star$ can be determined by plotting the log-target (either $S(b^{(t)})$ 
-or $U(\theta^{(t)})$ as a function of $t$,
+$\kappa_\star$ can be determined by plotting the log-target -- either $S(b^{(t)})$ 
+or $U(\theta^{(t)})$ -- as a function of $t$,
 and choosing $\kappa_\star$ to encompass the region where the log-target has 
@@
 By way of illustration, we can consider the primary school dataset \cite{schools}.
-We plot the normalised objective function $U\left( \theta^{(t)} \right) / N$ with respect to MALA iteration on Figure \ref{fig:school-U-orginal} -- using the hyperparameters given in Appendix \ref{appdx:hyperparams}.
+We plot the normalised objective function with respect to MALA iteration on Figure \ref{fig:school-U-orginal} -- using the hyperparameters given in Appendix \ref{appdx:hyperparams}.
 We see that the chain converges to the modal neighbourhood quickly (within about 2000 iterations of a total 10,000).
@@
 	\centering
-	\begin{subfigure}[t]{0.4\linewidth}
+	\begin{subfigure}[t]{0.42\linewidth}
 		\centering
 		\includegraphics[width=\linewidth]{school-U-original.png}
-		\caption{Original samples $t \in [T_\theta]$}
+		\caption{Original samples $t \in [T_\theta]$.}
 		\label{fig:school-U-orginal}
 	\end{subfigure}
-	\begin{subfigure}[t]{0.4\linewidth}
+	\begin{subfigure}[t]{0.42\linewidth}
 		\centering
 		\includegraphics[width=\linewidth]{school-U-thinned.png}
-		\caption{Thinned samples $t \in \Tcal_\theta$ indexed by $i$}
+		\caption{Thinned samples $t \in \Tcal_\theta$ indexed by $i$.}
 		\label{fig:school-U-thinned}
@@
 	\hspace{1cm}
-	\caption{Primary school dataset normalised objective function against sample index}
+	\caption{Primary school dataset normalised objective function against sample index.}
 \end{figure}
@@
 in the MDL solution to be our fixed block number $B$ for subsequent analysis. 
-Indeed, it is prudent to start the $b$-chain at this MDL solution as then 
-the necessary burn-in time can be greatly reduced.
+We can then restart the $b$-chain at this MDL solution 
+but this time fixing the number of blocks $B=B^{\textrm{MDL}}$ so that all our $b$-samples have the same dimension.
+Another benefit of starting the chain at the MDL solution is that our burn-in time is reduced.
 
@@
 In other words, we prefer to say that this block is comprised of the vertices with feature $d$ turned on rather than this block contains the vertices with feature $d$ switched off.
-To summarise the approach explicitly, say we have a feature encoding a pupil's gender: \{"male", "female"\}, we represent this as 2 binary feature flags "male": \{0, 1\} and "female": \{0, 1\}.
+To summarise the approach explicitly, say we have a feature encoding a pupil's gender: \{``male", ``female"\}, we represent this as 2 binary feature flags ``male": \{0, 1\} and ``female": \{0, 1\}.
 This approach may seem inefficient but it is vital to ensure we can interpret the resulting parameter distributions.
@@
 	\phi_j(x) = \frac{\exp(\beta_0) \cdot \exp(w_{jd})}{\sum_{k \in [B]} \exp(\beta_0) \cdot \exp(w_{kd})}
-	= \frac{\exp(w_{jd})}{\sum_{k \in [B]} \exp(w_{kd})}
+	= \frac{\exp(w_{jd})}{\sum_{k \in [B]} \exp(w_{kd})}.
 \end{equation}
@@
 This expression is independent of $\beta_0$ and so it is free to vary.
-This is not the whole picture as we also have the prior which introduces a regularisation term which favours weights closer to 0.
+This is not the whole picture as we also have the prior which introduces a regularisation term, favouring weights closer to 0.
 This ensures the MAP solution is unique.
 Nevertheless, for mutually exclusive binary feature flags, the bias term does not add to the expressiveness of the model and only serves to complicate analysis.
-We therefore remove the bias term from the softmax.
+We therefore choose to remove the bias term from the softmax.
 
@@
 Indeed, the approach was initially developed with a bias term but we found its removal yielded more reliable and reproducible results.
+\clearpage
 
@@
 We are given samples $\{\theta^{(t)}\}_{t \in \Tcal_\theta} \sim p(\theta | A, X)$ and wish to determine the statistical significance of the weights.
-We adopt matrix notation for simplicity by representing $\theta$ with the matrix $B \times D$ matrix of feature weights $W$.
-The question is to determine the significance of a particular feature $d$ by examining the value of $W_{id}$ for all the values $i \in [B]$.
+We adopt matrix notation for simplicity by representing $\theta$ with the $B \times D$ matrix of feature weights $W$.
+The question is to determine the significance of a particular feature $d$ from the sampled values of $W_{id}$ for all the values $i \in [B]$.
 We know that the posterior is proportional to the prior multiplied by the likelihood,
@@
 The likelihood is peaked around areas of $\theta$ that generate a partition $b$ that is highly likely in the SBM sense -- high $p(A|b)$.
-This provides the motivation for using the Laplace approximation for modelling the posterior $p(\theta | A, X) \approx p(\theta; \hat{\mu}, \hat{\Sigma})$; we use a multivariate Gaussian to approximate the posterior.
+This provides the motivation for using the Laplace approximation for the posterior $p(\theta | A, X) \approx p(\theta; \hat{\mu}, \hat{\Sigma})$.
 Indeed, the Laplace approximation is often used for modelling the posterior in logistic classification \cite{laplace}.
-This approximation is not exact but it motivates the derivation of the dimensionality reduction technique we construct by analogy with hypothesis testing.
 Marginalising over the other elements, the posterior for each element of the weight matrix $W$ can be approximated by,
@@
 %
-Where we have used the set of samples for $W$ drawn according to the exact posterior, to calculate unbiased estimates for the mean and standard deviation:
+We use the set of samples for $W$ drawn according to the exact posterior, to calculate unbiased estimates for the mean and standard deviation:
 %
@@
 This approximation is not exact but we can show it is accurate empirically.
-Indeed, if we run the primary school experiment with hyper-parameters given in Appendix \ref{appdx:hyperparams}, we can then plot histograms of the collected $W$-samples and compare these to the Laplace approximation.
-The results are given on Figure \ref{fig:school-histogram}.
-Even though the Laplace approximation is not exact, it is remarkably reliable.
+Indeed, if we run the primary school experiment with hyper-parameters given in Appendix \ref{appdx:hyperparams}, we can then plot histograms of the collected $W^{(t)}$-samples (for $t \in \Tcal_\theta$) and compare these to the Laplace approximation.
+The results are given on Figure \ref{fig:school-histogram}, which show the Laplace approximation is remarkably reliable.
 

### 2021-06-01 cued/additional.tex
Proof-read night before

@@
 
-For reference, we provide the likelihood and priors for the microcanonical SBM proposed by \citet{Peixoto-Bayesian-Microcanonical}. We have that the graph $A$ is drawn from the SBM,
+For reference, we provide the likelihood for the microcanonical SBM proposed by \citet{Peixoto-Bayesian-Microcanonical}. We have that the graph $A$ is drawn from the SBM,
 %
@@
 %
-Obviously, this form is only defined if $A$ respects the constraints imposed by $(k,e,b)$ else the likelihood is 0.
+Obviously, this form is only applicable if $A$ respects the constraints imposed by $(k,e,b)$ else the likelihood is 0.
 
@@
 
-Nevertheless, $\bar{U}$ decreases with $s$ indicating shorter burn-in times for larger step-sizes.
+Nevertheless, $\bar{U}$ decreases with $s$, indicating shorter burn-in times for larger step-sizes.
 However, this effect plateaus for $s>0.2$ for this particular dataset.
@@
 by the observed data, then the analysis is no longer ``fully Bayesian''
-and belongs to the class of methods referred to as ``empirical Bayes.''
+and belongs to the class of methods referred to as ``empirical Bayes''.
 However, as the number of blocks only specifies the coarseness of the 
@@
 
-Many vertex feature often take only one of several discrete values within a set.
+Many vertex features often take only one of several discrete values within a set.
 For example, with the primary school dataset \cite{schools}, a pupil's school-class can take one of 10 values from \{``1A", ``1B", ``2A", ``2B" \dots ``5A", "5B"\}.
@@
 Nevertheless, this dimensionality expansion means that we should not include a bias term in our softmax as then the MAP solution is less peaked and our $\theta$-samples will have higher variance.
-To illustrate this point we need only look at the form of the softmax activation function with bias vector $\beta$, operating on a vector $x$ which only has one non-zero entry $x_d=1$.
+To illustrate this point, we need only look at the form of the softmax activation function with bias vector $\beta$, operating on a vector $x$ which only has one non-zero entry $x_d=1$.
 Component $j$ of the output is given by:
@@
 To illustrate this point, we set $\beta_k=\beta_0$ for all $j$.
-If this is the case then equation \ref{eqn:softmax-bias} can be simplified to:
+If this is the case,~(\ref{eqn:softmax-bias}) can be simplified to:
 %
@@
 We are given samples $\{\theta^{(t)}\}_{t \in \Tcal_\theta} \sim p(\theta | A, X)$ and wish to determine the statistical significance of the weights.
-We adopt matrix notation for simplicity by representing $\theta$ with the $B \times D$ matrix of feature weights $W$.
-The question is to determine the significance of a particular feature $d$ from the sampled values of $W_{id}$ for all the values $i \in [B]$.
+We adopt matrix notation for simplicity, representing $\theta$ with the $B \times D$ matrix of feature weights $W$.
+The question is to determine the significance of a particular feature $d$ from the sampled values of $W_{id}$ for $i \in [B]$.
 We know that the posterior is proportional to the prior multiplied by the likelihood,
@@
 We choose to reject the null hypothesis $H_0$ if and only if equation \ref{eqn:hyp-test-empty} holds.
-We can say with probability greater than or equal to $\Phi(k)$ that either $W_{ij} < -c$ or $W_{ij} > c$ thus rejecting the null $H_0: |W_{ij}| \leq c$.
+This allows us to state with confidence greater than $\Phi(k)$ that either $W_{ij} < -c$ or $W_{ij} > c$ thus rejecting the null $H_0: |W_{ij}| \leq c$.
 
@@
 We only want to keep weights for which we have a confident prediction outside the range $(-c, c)$ not those that succeed by summing the two extremes of their distribution.
-Note that this case is very rare empirically but it nonetheless must be specified.
+This case is very rare empirically but it nonetheless must be specified.

### 2021-06-01 cued/derivations.tex
Proof-read
Proof-read night before

@@
 because the prior is the same for each $w_j$. 
-Therefore, the integral is only a function of $x_i$ and $\sigma_\theta^2$,
+Therefore, the integral can only be a function of at most $x_i$ and $\sigma_\theta^2$,
 which means that, as a function of $b$, $p(b|X)\propto 1$. As
@@
 	\frac{\partial a_{ij}}{\partial w_k} &= \frac
-	{\tilde{x}_i \exp(w_j^T \tilde{x}_i) \delta_{jk} \cdot \sum_{r \in [B]} \exp(w_r^T \tilde{x}_i) 
+	{x_i \exp(w_j^T x_i) \delta_{jk} \cdot \sum_{r \in [B]} \exp(w_r^T x_i) 
 		- 
-		\exp(w_j^T \tilde{x}_i) \cdot \tilde{x}_i \exp(w_k^T \tilde{x}_i)}
-	{\left( \sum_{r \in [B]} \exp(w_r^T \tilde{x}_i) \right)^2} \nonumber \\
-	&= \tilde{x}_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right), 
+		\exp(w_j^T x_i) \cdot x_i \exp(w_k^T x_i)}
+	{\left( \sum_{r \in [B]} \exp(w_r^T x_i) \right)^2} \nonumber \\
+	&= x_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right), 
 	\label{eq:dadw}
@@
 	\sum_{i \in [N]} \sum_{j \in [B]} y_{ij} 
-	\left( -\frac{\tilde{x}_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
+	\left( -\frac{x_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
 	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
-	&=  - \left( \sum_{i \in [N]} \tilde{x}_i \left( y_{ik} - a_{ik} \sum_{j \in [B]} y_{ij} \right)
+	&=  - \left( \sum_{i \in [N]} x_i \left( y_{ik} - a_{ik} \sum_{j \in [B]} y_{ij} \right)
 	- \frac{w_k}{\sigma_\theta^2} \right) \nonumber \\
-	&= - \left( \sum_{i \in [N]} \Big\{ \tilde{x}_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right).
+	&= - \left( \sum_{i \in [N]} \Big\{ x_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right).
 \end{align}

### 2021-06-01 cued/inference.tex
Proof-read night before

@@
 This entropy can be thought of as the optimal
-``description length'' of the graph. 
+``description length'' of the graph; if we 
+constructed an optimal code for the SBM ensemble,
+this would be the minimum amount of information 
+required to store the graph and its parameters. 
 The exact form of the proposal $q_b$ is explored thoroughly in
@@
 where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter 
-which may vary with the sample index (see Appendix \ref{appdx:step-size} for a discussion of step-size).
+which may vary with the sample index (see Appendix \ref{appdx:step-size} for a discussion of step-size variation).
 This leads to the proposal distribution,
@@
 Intuitively, this means discarding any feature $j$ for which 
-$(\hat{\mu}_{ij} - k\hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \sigma)$ overlaps with
+$(\hat{\mu}_{ij} - k\hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij})$ overlaps with
 $(-c, c)$ for all block indices $i$.
@@
 The multiplier $k$ in~(\ref{eqn:fij}) determines the degree of significance of the result.
-However, as the Laplace approximation is not exact we will only treat this dimensionality reduction method as a useful heuristic and not an exact method.
+However, as the Laplace approximation is not exact, we only treat this dimensionality reduction method as a useful heuristic and not an exact method.
 Appendix~\ref{appdx:hyp-test} explores this subject at greater depth.

### 2021-06-01 cued/risk.tex
Proof-read night before

@@
 
-As experimental work was entirely computer-based the risks encountered were not what one might typically associate with a more hands-on project: heavy machinery, dangerous chemicals etc.
+As experimental work was entirely computer-based, the risks encountered were not what one might typically associate with a more hands-on project: heavy machinery, dangerous chemicals etc.
 Most of the risks identified would be as a result of an improper working environment.

### 2021-06-02 cued/additional.tex
SUBMIT

@@
 %
-This accept-reject step ensures the resulting Markov Chain is in detailed balance with the target distribution $\pi(\cdot)$. Therefore, the set of samples produced $\{x^{(t)}\} \sim \pi^*(x)$.
+This accept-reject step ensures the resulting Markov chain is in detailed balance with the target distribution $\pi(\cdot)$. Therefore, the set of samples produced $\{x^{(t)}\} \sim \pi^*(x)$.
 
@@
 This method relies on the assumption that the random starting point $\theta^{(0)}$ has a high value for the objective function $U(\theta^{(0)})$; this assumption holds as the domain of $\theta$ is a high dimensional space so we are unlikely to get a low $U(\theta)$ from a lucky guess.
-We plot the acceptance ratio $r_\alpha$ and average objective $\bar{U}$ for $10$ values of the step-size multiplier $s$ logarithmically spaced in the range $(10^{-2}, 10^1)$ on Figure \ref{fig:step-size-tradeoff}.
+We plot the acceptance ratio $r_\alpha$ and average objective $\bar{U}$ for $10$ values of the step-size multiplier $s$, logarithmically spaced in the range $(10^{-2}, 10^1)$ on Figure \ref{fig:step-size-tradeoff}.
 %
@@
 Many vertex features often take only one of several discrete values within a set.
-For example, with the primary school dataset \cite{schools}, a pupil's school-class can take one of 10 values from \{``1A", ``1B", ``2A", ``2B" \dots ``5A", "5B"\}.
+For example, with the primary school dataset \cite{schools}, a pupil's school-class can take one of 10 values from \{``1A", ``1B", ``2A", ``2B" \dots ``5A", ``5B"\}.
 

### 2021-06-02 cued/inference.tex
SUBMIT

@@
 
-We adopt the Markov chain Monte Carlo procedure of
+We adopt the Markov chain Monte-Carlo procedure of
 \cite{Peixoto-MCMC},
@@
 have been obtained, we can compute the empirical mean and standard deviation of each component of $\theta$.
-Switching back to matrix notation we define $\theta = W$, such that $W_{ij}$ is the weight component for block $i$ and feature $j$.
+Switching back to matrix notation, we define $\theta = W$, such that $W_{ij}$ is the weight component for block $i$ and feature $j$.
 We can then define:

### 2021-10-13 publication/abstract.tex
scaffold publication

--- 2021-06-01 cued/tech-abstract.tex
@@
-\section*{Technical Abstract}
+\begin{abstract}
 Labelled networks form a very common and important class of data,
@@
 A typical inference goal is to determine how the vertex labels
-(or {\em features}) affect the network's graph structure.
+(or {\em features}) affect the network's structure.
 
@@
 Nevertheless, a hierarchical formulation would allow explanations of structure at any arbitrary length-scale.
+\end{abstract}

### 2021-10-13 publication/extensions.tex
scaffold publication

\section{Extensions}

The greatest limitation of the current FFBM formulation is that it can only explain structure at the macro-scale. In other words, we cannot explain structure within each detected block. Future work will benefit from extending the FFBM to be hierarchical in nature. This would be a relatively natural extension. Indeed, the SBM has already been extended to a hierarchical form, often called the nested SBM \cite{SBM-hierarchical}. The idea is to divide each block into sub-blocks and so on recursively until a specified depth is reached. The full block membership for a particular vertex now encodes the memberships at all levels of the hierarchy.

The necessary modification of the feature-to-block generator is also rather natural. Given the nested SBM, we would have a hierarchy of generators, each generating a block membership at a particular level of the hierarchy. Nevertheless, this does pose some practical issues for scalability; supposing we have $L$ levels in our hierarchy and each divides the parent block into $B$ sub-blocks, then the number of distinct generators necessary scales as $O(B^L)$. To avoid exponential growth in the number of model parameters, we could apply some form of dimensionality reduction as we descend the layers so that each generator is only given relevant features as input.

### 2021-10-13 publication/ffbm.tex
scaffold publication

--- 2021-06-01 cued/ffbm.tex
@@
 and ${\cal X}$ the set of values each feature can take.
-We define the vector $x_i \in \Xcal^D$ as the feature vector of vertex $i$, 
+We define the vector $x_i \in \Xcal^D$ as the feature vector for vertex $i$, 
 where $D$ is the number of features associated with each vertex.
@@
 \end{equation}
+%
+Figure~\ref{fig:ffbm} illustrates the proposed model.
 

### 2021-10-13 publication/introduction.tex
scaffold publication

--- 2021-05-28 cued/introduction.tex
@@
 To analyse a labelled network using one of the simple SBM variants, a typical inference procedure would be to first partition the graph into blocks grouped by distinct values of the feature of interest, and then use the associated model to test for evidence of heterogeneous connectivity between the feature-grouped blocks.
-Nevertheless, this approach is limited in that it can only consider one feature at a time.
+Nevertheless, this approach is limited as it can only consider one feature at a time.
 This makes it difficult to rank-order the features by  magnitude of impact.
@@
 This is an extension of the SBM to labelled networks.
-We go on to present an efficient algorithm for sampling from the parameters of the feature-to-block generator, and to describe how the
-sampled FFBM parameters can be interpreted to determine which features have the largest impact on overall graphical structure.
+We go on to present an efficient algorithm for sampling from the FFBM parameters, and to describe how the sampled parameters can be interpreted to determine which features have the largest impact on overall graphical structure.

### 2021-10-19 publication/prelim.tex
address comments

--- 2021-05-29 cued/prelim.tex
@@
 where $B$ is the number of non-empty blocks.
-	Let $e=(e_{rs})$ be the $N\times N$ symmetric matrix of edge counts 
+	Let $e=(e_{rs})$ be the $B \times B$ symmetric matrix of edge counts 
 between blocks, such that $e_{rs}$ is the number of edges from block $r$ to 
@@
 
-Appendix \ref{appdx:sbm} describes how to compute the likelihood, $p(A|b,e,k)$, for graphs generated from the microcanonical DC-SBM.
+Appendix \ref{appdx:sbm-likelihood} describes how to compute the likelihood, $p(A|b,e,k)$, for graphs generated from the microcanonical DC-SBM.

### 2021-10-19 publication/abstract.tex
More scaffolding

@@
 \begin{abstract}
-Labelled networks form a very common and important class of data,
-naturally appearing in numerous applications in science and engineering.
+Labelled networks are a very common and important class of data.
 A typical inference goal is to determine how the vertex labels
@@
 In this work, we introduce a new generative model, the feature-first block model (FFBM),
-which is more effective at describing vertex-labelled undirected
-graphs and also facilitates the use of richer queries on labelled networks.
+which facilitates the use of richer queries on labelled network structure.
 We develop a Bayesian framework for inference with this model,
 and we present a method to efficiently sample from the posterior
-distribution of the FFBM parameters. The FFBM's structure is kept
-deliberately simple to retain easy interpretability of the parameter
-values.
+distribution of the FFBM parameters.
 
@@
 that the whole feature-space is used automatically and that features
-can be rank-ordered implicitly according to impact. Any features
-that do not significantly impact the high-level structure can be
-discarded to reduce the problem dimension. In cases where the vertex
-features available do not readily explain the community structure
-in the resulting network, the approach detects this and is protected
-against over-fitting.
+can be rank-ordered implicitly according to impact.
 
-Future work on this topic could benefit from extending the FFBM to be hierarchical in nature. 
-Such a modification would be relatively natural and would increase the descriptive power of the model. 
-The current FFBM formulation is limited in that it can only explain structure at the macro-scale. 
-Nevertheless, a hierarchical formulation would allow explanations of structure at any arbitrary length-scale.
+\keywords{Stochastic Block Models \and Bayesian Inference.}
+
 \end{abstract}

### 2021-10-26 publication/experiments.tex
cosmetic changes

--- 2021-06-01 cued/experiments.tex
@@
 Vertex features include class membership (one of 10 values: 1A-5B), gender (male, female) and teacher status encoded as an 11th school-class.
-No further identifiable information is retained.
 We choose to analyse just the second day of results.
@@
 		\includegraphics[width=0.8\linewidth]{10-vertical-legend.png}
-		\caption{Legend}
+		\caption{Block Legend}
 		\label{fig:10-legend}
 	\end{subfigure}
-	\caption{Networks laid out and coloured according by inferred block memberships $\hat{y}$ for a given experiment iteration.
+	\caption{Networks laid out and coloured according to inferred block memberships $\hat{y}$ for a given experiment iteration.
 Visualisation performed using \textit{graph-tool} \cite{peixoto_graph-tool_2014}.}
@@
 	\centering
-	\caption{Experimental results averaged over $n=10$ iterations (mean $\pm$ standard deviation).}
+	\caption{Experimental results averaged over $n=10$ iterations (mean $\pm$ std. dev.).}
 	\label{tab:results}
@@
 
-Table~\ref{tab:results} summarises the results for each experiment;
-a complete list of all the hyperparameters used 
-is given in Appendix \ref{appdx:hyperparams}. We also apply the 
+Table~\ref{tab:results} summarises the results for each experiment. We also apply the 
 dimensionality reduction method 
@@
 		\centering
-		\imagebox{6.2cm}{\includegraphics[width=\linewidth]{school-null-1}}
+		\imagebox{0.8\linewidth}{\includegraphics[width=\linewidth]{school-null-1}}
 		\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
@@
 		\centering
-		\imagebox{6.2cm}{\includegraphics[width=\linewidth]{school-accuracy-1}}
+		\imagebox{0.8\linewidth}{\includegraphics[width=\linewidth]{school-accuracy-1}}
 		\caption{Per-block accuracy $\eta(j)$.}
@@
 		\centering
-		\imagebox{7cm}{\includegraphics[width=\linewidth]{fb-null-1}}
+		\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{fb-null-1}}
 		\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
@@
 		\centering
-		\imagebox{7cm}{\includegraphics[width=\linewidth]{fb-accuracy-1}}
+		\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{fb-accuracy-1}}
 		\caption{Per-block accuracy $\eta(j)$.}
@@
 a vertex from the block. However, Figure~\ref{fig:fb-null} demonstrates why we made 
-the deliberate choice $\Xcal=\{0,1\}$, as argued in Appendix \ref{appdx:dimension}. 
+the deliberate choice $\Xcal=\{0,1\}$. 
 Because of the choice $\Xcal=\{0,1\}$, block 1 is allowed to contain 3 distinct feature 

### 2021-10-26 publication/inference.tex
cosmetic changes

--- 2021-06-02 cued/inference.tex
@@
 Markov chain Monte-Carlo approach via a Metropolis-Hastings 
-algorithm \cite{hastings-alg}, as a two-level Markov chain. See Appendix \ref{appdx:metropolis-hastings} for more details on the Metropolis-Hastings algorithm.
+algorithm \cite{hastings-alg}, as a two-level Markov chain.
 
@@
 	\begin{tikzpicture}[
-		scale=0.95, every node/.style={transform shape},
+		scale=0.8, every node/.style={transform shape},
 		roundnode/.style={circle, draw=black, minimum size=12mm},
@@
 where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter 
-which may vary with the sample index (see Appendix \ref{appdx:step-size} for a discussion of step-size variation).
+which may vary with the sample index.
 This leads to the proposal distribution,
@@
 We can obtain an unbiased estimate for this quantity using 
-the thinned $b$-samples after burn-in; the length of the
-burn-in and the details of the thinning are described
-in Appendix \ref{appdx:burn-in-thinning}.
+the thinned $b$-samples after burn-in.
 Let $\Tcal_b$  denote the retained set of indices 
@@
 the $\theta$-chain by reducing the variance in the evaluation 
-of $U$ and $\nabla U$. A detailed description of the overall 
-algorithms is given in Appendix \ref{appdx:algorithms}.
+of $U$ and $\nabla U$.
 
@@
 However, as the Laplace approximation is not exact, we only treat this dimensionality reduction method as a useful heuristic and not an exact method.
-Appendix~\ref{appdx:hyp-test} explores this subject at greater depth.
 
@@
 %
-For an algorithmic implementation of this method refer to appendix \ref{appdx:algorithms}. This approach has the advantage that it fixes the number of reduced dimensions.
+This approach has the advantage that it fixes the number of reduced dimensions.

### 2021-10-26 publication/detailed-experiments.tex
Editing

--- 2021-10-26 publication/experiments.tex
@@
-\section{Experimental results}
-\label{sec:experiments}
-
-We apply the developed methods to a variety of datasets. These are chosen to span a range of node counts $N$, edge counts $E$ and feature space dimension $D$. We consider the following:
-
-\begin{itemize}
-	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=3$) -- network of Amazon book sales about U.S.
-politics, published close to the presidential election in 2004.
-Two books are connected if they were frequently co-purchased by customers.
-Vertex features encode the political affiliation of the author (liberal, conservative, or neutral).
-
-	\item \textbf{Primary school dynamic contacts} \cite{schools} ($N=238, E=5539, D=13$) -- network of face-to-face contacts amongst students and teachers at a primary school in Lyon, France.
-Two nodes are connected if the two parties shared a face-to-face interaction over the school-day.
-Vertex features include class membership (one of 10 values: 1A-5B), gender (male, female) and teacher status encoded as an 11th school-class.
-We choose to analyse just the second day of results.
-
-	\item \textbf{Facebook egonet} \cite{fb-snap} ($N=747, E=30025, D=480$) -- an assortment of Facebook users' friends lists.
-Vertex features are extracted from each user's profile and are fully anonymised.
-They include information about education history, languages spoken, gender, home-town, birthday etc.
-We focus on the egonet with id 1912.
-
-\end{itemize}
-
-\begin{figure}[!h]
-	\centering
-	\begin{subfigure}[t]{0.28\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{polbooks-graph.png}
-		\caption{Polbooks}
-		\label{fig:polbooks-graph}
-	\end{subfigure}
-	\hfill
-	\begin{subfigure}[t]{0.28\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{school-graph.png}
-		\caption{School}
-		\label{fig:school-graph}
-	\end{subfigure}
-	\hfill
-	\begin{subfigure}[t]{0.28\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{fb-graph.png}
-		\caption{Facebook egonet}
-		\label{fig:fb-graph}
-	\end{subfigure}
-	\begin{subfigure}[t]{0.11\linewidth}
-		\centering
-		\includegraphics[width=0.8\linewidth]{10-vertical-legend.png}
-		\caption{Block Legend}
-		\label{fig:10-legend}
-	\end{subfigure}
-	\caption{Networks laid out and coloured according to inferred block memberships $\hat{y}$ for a given experiment iteration.
-Visualisation performed using \textit{graph-tool} \cite{peixoto_graph-tool_2014}.}
-	\label{fig:graphs-all}
-\end{figure}
-\begin{table}[!h]
-	\centering
-	\caption{Experimental results averaged over $n=10$ iterations (mean $\pm$ std. dev.).}
-	\label{tab:results}
-	\resizebox{\textwidth}{!}{%
-		\begin{tabular}{c|ccc|c|cc|ccc}
-			Dataset  & $B$ & $D$ & $D'$ & $\bar{S}_e$ & $\bar{\Lcal}_0$ & $\bar{\Lcal}_1$ & $c^*$ & $\bar{\Lcal}_0'$ & $\bar{\Lcal}_1'$  \\ \hline
-			Polbooks & 3 & 3 & -- & $2.250 \pm 0.000$ & $0.563 \pm 0.042$ & $0.595 \pm 0.089$ & -- & -- & -- \\
-			School & 10 & 13 & 10 & $1.894 \pm 0.004$ & $0.787 \pm 0.127$ & $0.885 \pm 0.129$ & $1.198 \pm 0.249$ & $0.793 \pm 0.132$ & $0.853 \pm 0.132$ \\
-			FB egonet & 10  & 480 & 10 & $1.626 \pm 0.003$ & $1.326 \pm 0.043$ & $1.538 \pm 0.069$ & $0.94 \pm 0.019$ & $1.580 \pm 0.150$ & $1.605 \pm 0.106$
-		\end{tabular}
-	}
-\end{table}
-%
-We require metrics to assess performance. This can be split into two separate 
-components: the microcanonical SBM fit (concerned with the $b$-samples) and 
-the fit of the feature-to-block generator (concerned with the $\theta$-samples).
-
-Starting with the SBM, recall that the quantity $S(b)$ in~(\ref{eqn:dl-form}) 
-can be interpreted as an ideal ``description length'' of the partition 
-imposed by $b$. We define a simple
-metric $\bar{S}_e$ to gauge the fit of the SBM,
-as the description length per entity 
-(i.e., divided by the total number $N+E$ of 
-nodes plus edges),
-averaged over the $b$-samples:
-%
-\begin{equation}
-	\bar{S}_e \coloneqq \frac{1}{(N+E) |\Tcal_b|} \sum_{t\in \Tcal_b} S \left( b^{(t)} \right).
-	\label{eqn:mean-dl}
-\end{equation}
-%
-
-Next, to assess the performance of the feature-to-block predictor, 
-we partition the vertex set $[N]$ into training and test sets. We choose to 
-randomly partition the vertices on each experiment run,
-so that a constant fraction $f$ of the available vertices go to form 
-the training set $\Gcal_0$ and the remainder are held out to form the
-test set $\Gcal_1$.
-The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain.
-Because $|\Gcal_0| \neq |\Gcal_1|$ in general, we cannot use the un-normalised log-target $U$ from~(\ref{eqn:U-form}) for comparison,
-as the total cross-entropy loss scales with the size of each data set but 
-the prior term stays constant. We therefore use the average cross-entropy loss 
-over each set,
-%
-\begin{equation}
-	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star \left( \theta^{(t)} \right),
-	\quad \textrm{where} \quad
-	\Lcal_\star \left( \theta^{(t)} \right) \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)},
-	\label{eqn:cross-entropy-loss}
-\end{equation}
-%
-where $\star \in \{0, 1\}$ indicates whether the training or test
-set is being considered.
-
-Table~\ref{tab:results} summarises the results for each experiment. We also apply the 
-dimensionality reduction method 
-of Section~\ref{sec:dim-reduction}
-to the two higher dimensional datasets (the school and FB egonet). 
-For this we use equation~(\ref{eqn:c-star}) with $k=1$,
-in order to reduce the dimension from 
-$D$ to a desired $D'$. 
-We then retrain the feature-block predictor using only the retained 
-feature set $\Dcal'$, and report the log-loss over the training and 
-test sets for the reduced classifier -- 
-denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively. 
-These values are also given in Table~\ref{tab:results}.
-
-Based on these results, some remarks are in order.
-Firstly, the variance of the test loss $\bar{\Lcal}_1$ tends to be higher 
-than the training loss $\bar{\Lcal}_0$. This is expected,
-as the test set is smaller than the training set and hence 
-more susceptible to variability in its construction. Indeed, 
-most of the variance in the evaluation of $\bar{\Lcal}_0$ and $\bar{\Lcal}_1$ 
-comes from the random partitioning of the graph into training and test sets. 
-Secondly, it can be seen that the dimensionality reduction procedure 
-brings the training and test losses closer together. This implies that 
-the features we keep are indeed correlated with the underlying graphical 
-partition and that the approach generalises correctly.
-
-The average description length per entity,
-$\bar{S}_e$, of the graph, 
-has very small variance, suggesting that
-the detected communities can be found reliably (to within an arbitrary 
-relabelling of blocks). For reference, we plot an inferred partition for each 
-of the graphs in Figure~\ref{fig:graphs-all}.
-The polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do succeed at dividing the graph into densely connected clusters.
-
-Nevertheless, the cross-entropy loss over the whole training set may be too coarse a measure of model fit.
-It is often the case that we have good feature-based explanations for some but not all of the detected blocks.
-We wish to define a new measure of fit specific to each detected block.
-This requires defining the set of all vertices associated with block $j$, as,
-%
-\begin{equation}
-	\Bcal_\star(j) \coloneqq \{i \in \Gcal_\star : \hat{b}_i = j\} \quad \textrm{where} 
-	\quad \hat{b}_i \coloneqq \argmax_j \hat{y}_{ij}.
-\end{equation}
-%
-Recall that $\hat{y}_{ij}$ is our estimate for the block membership posterior~(\ref{eqn:y-hat}) using only information from the adjacency matrix $A$.
-Again, $\star \in \{0, 1\}$ toggles between training and test sets.
-Now $\Bcal_\star(j)$ is the set of all vertices within the training or test set that have maximum a posteriori probability of belonging to set $j$.
-We choose to resolve ties consistently by picking the lower index.
-We now define the accuracy for block $j$ as,
-%
-\begin{equation}
-	\eta_\star(j) \coloneqq \frac{1}{|\Bcal_\star (j)| \cdot 
-	|\Tcal_\theta| } 
-	\sum_{i \in \Bcal_\star (j)}  \sum_{t \in \Tcal_\theta}
-	\one \left\{\hat{b}_i = \argmax_j \phi_j \left( x_i; \theta^{(t)} \right) \right\}.
-	\label{eqn:accuracy}
-\end{equation}
-%
-This is effectively testing whether the feature-to-block predictions and the graph-based predictions agree in their largest component.
-We call this metric, $\eta_\star(j)$, the {\em block-accuracy} for block $j$.
-It is clearly bounded $0 \leq \eta_\star(j) \leq 1$, with an accuracy of 1 meaning perfect agreement for the vertices in detected block $j$.
-
-For each of the experiments, we plot the collected $\theta$-samples for the features that survive the dimensionality reduction procedure.
-We also plot the per-block accuracy for the original-dimension classifiers.
-We discuss the results specific to each dataset in turn.
-
 \subsection{Political books}

### 2021-10-26 publication/experiments.tex
Editing

@@
 \begin{itemize}
-	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=3$) -- network of Amazon book sales about U.S.
-politics, published close to the presidential election in 2004.
-Two books are connected if they were frequently co-purchased by customers.
+	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=3$) -- network of Amazon political book sales, published close to the 2004 presidential election.
+Two books are connected if they were frequently co-purchased.
 Vertex features encode the political affiliation of the author (liberal, conservative, or neutral).
@@
 	\item \textbf{Primary school dynamic contacts} \cite{schools} ($N=238, E=5539, D=13$) -- network of face-to-face contacts amongst students and teachers at a primary school in Lyon, France.
-Two nodes are connected if the two parties shared a face-to-face interaction over the school-day.
 Vertex features include class membership (one of 10 values: 1A-5B), gender (male, female) and teacher status encoded as an 11th school-class.
@@
 	\item \textbf{Facebook egonet} \cite{fb-snap} ($N=747, E=30025, D=480$) -- an assortment of Facebook users' friends lists.
-Vertex features are extracted from each user's profile and are fully anonymised.
-They include information about education history, languages spoken, gender, home-town, birthday etc.
+Vertex features are fully anonymised and encode information about each user's education history, languages spoken, gender, home-town, birthday etc.
 We focus on the egonet with id 1912.
@@
 \begin{equation}
-	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star \left( \theta^{(t)} \right),
+	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star^{(t)},
 	\quad \textrm{where} \quad
-	\Lcal_\star \left( \theta^{(t)} \right) \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)},
+	\Lcal_\star^{(t)} \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)},
 	\label{eqn:cross-entropy-loss}
@@
 as the test set is smaller than the training set and hence 
-more susceptible to variability in its construction. Indeed, 
-most of the variance in the evaluation of $\bar{\Lcal}_0$ and $\bar{\Lcal}_1$ 
-comes from the random partitioning of the graph into training and test sets. 
+more susceptible to variability in its construction. 
 Secondly, it can be seen that the dimensionality reduction procedure 
@@
 We wish to define a new measure of fit specific to each detected block.
-This requires defining the set of all vertices associated with block $j$, as,
-%
-\begin{equation}
-	\Bcal_\star(j) \coloneqq \{i \in \Gcal_\star : \hat{b}_i = j\} \quad \textrm{where} 
+This requires defining the set of all vertices associated with block $j$ as,
+$
+	\Bcal_\star(j) \coloneqq \{i \in \Gcal_\star : \hat{b}_i = j\}
+$
+where
+$ 
 	\quad \hat{b}_i \coloneqq \argmax_j \hat{y}_{ij}.
-\end{equation}
-%
+$
 Recall that $\hat{y}_{ij}$ is our estimate for the block membership posterior~(\ref{eqn:y-hat}) using only information from the adjacency matrix $A$.
 Again, $\star \in \{0, 1\}$ toggles between training and test sets.
-Now $\Bcal_\star(j)$ is the set of all vertices within the training or test set that have maximum a posteriori probability of belonging to set $j$.
+Now $\Bcal_\star(j)$ is the set of vertices that have maximum a posteriori probability of belonging to block $j$.
 We choose to resolve ties consistently by picking the lower index.
@@
 %
-This is effectively testing whether the feature-to-block predictions and the graph-based predictions agree in their largest component.
+This is effectively testing whether the feature-to-block and the graph-to=block predictions agree in their largest component.
 We call this metric, $\eta_\star(j)$, the {\em block-accuracy} for block $j$.
@@
 
-\subsection{Political books}
-
-The goal here is to 
-determine whether the authors' political affiliations are a good predictor 
-of the overall network structure. We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral).
-%
-\begin{figure}[!h]
-	\centering
-	\begin{subfigure}[t]{0.452\linewidth}
-		\centering
-		\vskip 0pt
-		\includegraphics[width=\linewidth]{polbooks-null-1}
-		\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
-		\label{fig:polbooks-null}
-	\end{subfigure}
-	\begin{subfigure}[t]{0.45\linewidth}
-		\centering
-		\vskip 0pt
-		\includegraphics[width=\linewidth]{polbooks-accuracy-1}
-		\caption{Per-block accuracy $\eta(j)$.}
-		\label{fig:polbooks-accuracy}
-	\end{subfigure}
-	\caption{Political books dataset.}
-	\label{fig:polbooks}
-\end{figure}
-
-From Figure~\ref{fig:polbooks-null} we see that all 3 blocks have a distinct political affiliation as their largest positive component. 
-This strongly indicates that political affiliation is indeed the axis 
-which best predicts the 3-way natural partition of the graph into blocks. 
-Furthermore, in Table~\ref{tab:results} we see that the training and test losses 
-are very similar and both are low in magnitude. This provides further evidence 
-to the claim that political affiliation is a very appropriate explanatory 
-variable for the overall network structure.
-
-However, from Figure~\ref{fig:polbooks-accuracy} we see that block 1 has low accuracy. 
-This suggests that detected block 1 is not solely composed of ``neutral" books but also 
-contains some ``liberal" and ``conservative" authors. Indeed, by examining 
-Figure~\ref{fig:polbooks-graph} we see that block 1 is effectively a buffer between 
-blocks 2 and 3; there are very few edges between blocks 2 and 3. It is therefore not 
-surprising that some books from either side leak into block 1.
-Perhaps, the three distinct categories (``conservative", ``liberal", ``neutral") are 
-too coarse a measure of political affiliation; it may be the case that the nominally 
-``conservative" books found in block 1 are in fact more centre-right. 
-In the absence of more granular labels, we cannot test this hypothesis. 
-Nevertheless, political affiliation encoded as 3 distinct labels 
-remains a fantastic predictor of network structure.
-
-\subsection{Primary school dynamic contacts}
-
-We choose the number of communities $B=10$, in line with the total number of 
-school classes. As before, we sample the block-generator parameters $\theta$ and 
-employ the dimensionality reduction technique 
-of Section~\ref{sec:dim-reduction},
-with standard deviation multiplier $k=1$ to pick out the top $D'=10$ features. 
-We then plot the weights for the resulting features $d \in \Dcal'$ 
-in Figure~\ref{fig:school-null}. It is rather obvious
-that only the pupils' class memberships (1A-5B) have remained
-significant;
-gender and teacher/student status have been discarded,
-meaning that these are not good predictors of overall macro-structure.
-%
-\begin{figure}[!h]
-	\centering
-	\begin{subfigure}[t]{0.45\linewidth}
-		\centering
-		\imagebox{0.8\linewidth}{\includegraphics[width=\linewidth]{school-null-1}}
-		\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
-		\label{fig:school-null}
-	\end{subfigure}
-	\begin{subfigure}[t]{0.45\linewidth}
-		\centering
-		\imagebox{0.8\linewidth}{\includegraphics[width=\linewidth]{school-accuracy-1}}
-		\caption{Per-block accuracy $\eta(j)$.}
-		\label{fig:school-accuracy}
-	\end{subfigure}
-	\caption{Primary school dynamic contacts dataset.}
-	\label{fig:school}
-\end{figure}
-
-The vast majority of blocks are composed of a single class. 
-However, some blocks have two comparably strong classes as their predictors. 
-For example, blocks 2 and 5 both contain classes 3A and 3B as their 2 best predictors. 
-This suggests that the social divide between classes is less pronounced 
-for pupils in year 3. Conversely, some classes are found to extend over two 
-detected blocks (class 2B spans blocks 8 and 9) but we do 
-not have a feature which explains the division. The most surprising block 
-is number 7, which has comparable weightings for classes 5A and 1B. 
-Perhaps there was a joint event between those two classes on the day 
-the data were collected.
-
-Figure~\ref{fig:school-accuracy} shows excellent accuracy for the majority of blocks.
-In fact the only blocks with low accuracy are those that have a school-class span two blocks such that we cannot reliably distinguish between the two.
-This is much more pronounced when we apply hard classification rather than the soft cross-entropy loss.
-Indeed block 1 has low accuracy because class 1A has a higher weight component for block 3 than it does for block 1.
-As we use hard classification to compute accuracy, vertices belonging to class 1A are predicted to belong to block 3.
-A similar effect is seen in block 5 which often loses out to block 2.
-The same can be said of blocks 8 and 9.
-However, in this case the weights for class 2B are very similar -- thus explaining the roughly evenly split accuracy for blocks 8 and 9.
-
-\subsection{Facebook egonet}
-
-We choose $B=10$ and $D'=10$ for this experiment. The selected features 
-(Figure~\ref{fig:fb-null}) are those that best explain the high-level 
-community structure. The majority of them are education related. 
-Nevertheless, for $D'=10$ we only have good explanations for the makeup 
-of some of the detected blocks; several blocks in 
-Figure~\ref{fig:fb-null} do not have high-magnitude components for $D'=10$.
-This is further emphasised by the disparate accuracies in Figure~\ref{fig:fb-accuracy}.
-Nevertheless, observe that the accuracy is only high for blocks that contain high magnitude weights.
-The only exception to this rule is block 9 which nonetheless may have high magnitude weights just below the cut-off $c^*$.
-%
-\begin{figure}[!h]
-	\centering
-	\begin{subfigure}[t]{0.45\linewidth}
-		\centering
-		\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{fb-null-1}}
-		\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
-		\label{fig:fb-null}
-	\end{subfigure}
-	\begin{subfigure}[t]{0.45\linewidth}
-		\centering
-		\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{fb-accuracy-1}}
-		\caption{Per-block accuracy $\eta(j)$.}
-		\label{fig:fb-accuracy}
-	\end{subfigure}
-	\caption{Facebook egonet dataset.}
-	\label{fig:fb}
-\end{figure}
-
-When the feature dimension is very large, it becomes increasingly likely that 
-a particular feature may uniquely identify a small set of nodes. If these nodes 
-are all part of the same community, then the classifier may overfit for that 
-particular parameter. The regularisation term imposed by the prior goes some 
-way towards alleviating this problem. Nevertheless, we see in 
-Figure~\ref{fig:fb-null} that the feature \verb*|birthday-5| has a very 
-high weight as it relates to block 1 -- but it is highly unlikely
-that birthdays determine graphical structure. This issue could have been alleviated by 
-choosing $\Xcal=\{-1, +1\}$ such that having a feature turned off supports excluding 
-a vertex from the block. However, Figure~\ref{fig:fb-null} demonstrates why we made 
-the deliberate choice $\Xcal=\{0,1\}$. 
-Because of the choice $\Xcal=\{0,1\}$, block 1 is allowed to contain 3 distinct feature 
-categories rather than just one. We accept this trade-off: risking spurious high magnitude 
-components (e.g. \verb*|birthday-5|) for the benefit of allowing a top-level block to span 
-multiple feature-groups. Nevertheless, one would have to apply a hierarchical approach 
-to subdivide each top-level block into sub-blocks to test whether the constituent 
-feature-groups determine structure at the lower-level.
+\include{detailed-experiments}

### 2021-10-26 publication/detailed-experiments.tex
More edits

@@
+%\begin{figure}[!h]
+%	\centering
+%	\begin{subfigure}[t]{0.32\linewidth}
+%		\centering
+%		\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{polbooks-null-1}}
+%		\caption{Political Books.}
+%		\label{fig:polbooks-null}
+%	\end{subfigure}
+%	\begin{subfigure}[t]{0.32\linewidth}
+%		\centering
+%		\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{school-null-1}}
+%		\caption{Primary School.}
+%		\label{fig:school-null}
+%	\end{subfigure}
+%	\begin{subfigure}[t]{0.32\linewidth}
+%		\centering
+%		\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{fb-null-1}}
+%		\caption{Facebook Egonet.}
+%		\label{fig:fb-null}
+%	\end{subfigure}
+%	\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
+%\end{figure}
+%%
+%\begin{figure}[!h]
+%	\centering
+%	\begin{subfigure}[t]{0.32\linewidth}
+%		\centering
+%		\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{polbooks-accuracy-1}}
+%		\caption{Political Books.}
+%		\label{fig:polbooks-null}
+%	\end{subfigure}
+%	\begin{subfigure}[t]{0.32\linewidth}
+%		\centering
+%		\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{school-accuracy-1}}
+%		\caption{Primary School.}
+%		\label{fig:school-null}
+%	\end{subfigure}
+%	\begin{subfigure}[t]{0.32\linewidth}
+%		\centering
+%		\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{fb-accuracy-1}}
+%		\caption{Facebook Egonet.}
+%		\label{fig:fb-null}
+%	\end{subfigure}
+%	\caption{Per-block accuracy $\eta(j)$.}
+%\end{figure}
+
 \subsection{Political books}
@@
 	\centering
-	\begin{subfigure}[t]{0.452\linewidth}
+	\begin{subfigure}[t]{0.45\linewidth}
 		\centering
@@
 From Figure~\ref{fig:polbooks-null} we see that all 3 blocks have a distinct political affiliation as their largest positive component.  
-This strongly indicates that political affiliation is indeed the axis 
-which best predicts the 3-way natural partition of the graph into blocks. 
 Furthermore, in Table~\ref{tab:results} we see that the training and test losses 
-are very similar and both are low in magnitude. This provides further evidence 
-to the claim that political affiliation is a very appropriate explanatory 
+are very similar and both are low in magnitude. This is strong evidence 
+that political affiliation is a very appropriate explanatory 
 variable for the overall network structure.
@@
 surprising that some books from either side leak into block 1.
-Perhaps, the three distinct categories (``conservative", ``liberal", ``neutral") are 
-too coarse a measure of political affiliation; it may be the case that the nominally 
-``conservative" books found in block 1 are in fact more centre-right. 
+The three distinct categories may be 
+too coarse a measure of political affiliation; perhaps the nominally 
+``conservative" books found in block 1 are closer to the centre. 
 In the absence of more granular labels, we cannot test this hypothesis. 
@@
 We choose the number of communities $B=10$, in line with the total number of 
-school classes. As before, we sample the block-generator parameters $\theta$ and 
-employ the dimensionality reduction technique 
-of Section~\ref{sec:dim-reduction},
-with standard deviation multiplier $k=1$ to pick out the top $D'=10$ features. 
+school classes. We 
+employ the dimensionality reduction technique to pick out the top $D'=10$ features. 
 We then plot the weights for the resulting features $d \in \Dcal'$ 

### 2021-10-30 publication/ffbm.tex
cosmetic changes
Edit first few sections

@@
 
-In this section we propose a novel generative model for labelled networks. We call this the feature-first block model (FFBM).
+In this section we propose a novel generative model for labelled networks. We call this the feature-first block model (FFBM -- illustrated in figure~\ref{fig:ffbm}).
 
-Let $N$ denote the number of vertices, $B$ the number of blocks in the graph,
+Let $N$ denote the number of vertices, $B$ the number of blocks
 and ${\cal X}$ the set of values each feature can take.
@@
 we deliberately exclude a bias term to ensure that we only leverage feature information.
-This decision is justified in greater detail in Appendix~\ref{appdx:dimension}.
 We will later also find it convenient
 to write the parameters $\theta$ as a $B \times D$ matrix of weights $W$.
-
 More complex models based on different choices for the distributions
@@
 the inferred parameter distributions is more difficult. 
-
-\begin{figure}[!h]
+%
+\begin{figure}
 	\centering
@@
 \end{equation}
-%
-Figure~\ref{fig:ffbm} illustrates the proposed model.
 
@@
 %
-Interestingly, this choice of prior leads to a particularly
-simple distribution on the block membership vector $b$ given $X$.
-We show in Appendix~\ref{appdx:b|x} that,
-after integrating out $\theta$,
-$b$ is uniformly distributed given $X$:
+This choice of prior gives a very
+simple form for the conditional distribution of the block membership vector $b$ given $X$. 
+It is a uniform distribution (\ref{eqn:b-pseudo-prior})
+ -- proof in Appendix~\ref{appdx:b|x}. This an important simplification as evaluating $p(b | X)$ does not require an expensive  integration over $\theta$ nor does it require the exact value of $X$.
 %
@@
 %
-This is an important simplification as evaluating $p(b | X)$ does not require an expensive Monte Carlo integration over $\theta$ nor does it require the exact value of $X$. 
+
 \citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for 
@@
 where we used the fact $\psi$ and $X$ are conditionally 
-independent given $b$.  The exact form of the prior
-$p(\psi|b)$ is described in Appendix \ref{appdx:sbm-prior}.
+independent given $b$.
 All that concerns the main argument is that it has

### 2021-10-30 publication/inference.tex
Edit first few sections

@@
 	b^{(t)} &\sim p ( b | A, X )  \label{eqn:b-samples},\\
-	\theta^{(t)} &\sim p(\theta | X, b^{(t)} ). \label{eqn:theta-samples}
+	\theta^{(t)} &\sim p\left(\theta | X, b^{(t)} \right). \label{eqn:theta-samples}
 \end{align}
@@
 algorithm \cite{hastings-alg}, as a two-level Markov chain.
-
-As we show in the following section, the resulting samples 
+The resulting samples 
 for $\theta^{(t)}$ are unbiased in the sense that the expectation of 
@@
 \begin{equation}
-\Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] = \sum_{b \in [B]^N} p(\theta | X, b) p(b | A, X) = \sum_{b \in [B]^N} p(\theta, b | A, X) = p(\theta | A, X).
+\Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] = \sum_{b \in [B]^N} p(\theta | X, b) p(b | A, X) = p(\theta | A, X).
 \label{eqn:theta-unbiased}
@@
 % the unbiased result in equation 
-is sufficient to prove that, for large enough $t$, $\theta^{(t)} \sim \Expect_{b^{(i)}} \left[ p(\theta | X, b^{(t)})\right] = p(\theta 
-| A, X)$, as required.
-%  which is exactly the distribution we are targetting (equation \ref{eqn:theta-target}).
+is sufficient to prove that, for large enough $t$, $\theta^{(t)} \sim p(\theta | A, X)$, as required.
 
@@
 Figure \ref{fig:samp-sequence} shows an overview of the proposed method. 
-The Metropolis-Hastings accept/reject probabilities 
-for the proposed $b$ and $\theta$ samples are denoted $\alpha_b$ and
-$\alpha_\theta$, respectively.
-% We have introduced subscripts and conditionings to make explicit what variables each step utilises.
+The Metropolis-Hastings proposal distribution is denoted $q$ and the acceptance probability denoted $\alpha$.
 

### 2021-10-30 publication/introduction.tex
Edit first few sections

@@
 
-An important characteristic of many real-world networks is that they exhibit strong community structure, with most nodes often belonging to a densely connected cluster. 
+Many real-world networks exhibit strong community structure, with most nodes often belonging to a densely connected cluster. 
 Finding ways to recover the latent communities from the observed graph is an important
 task in many applications, including
-graph/network compression \cite{cluster-compression} and link prediction in 
+network compression \cite{cluster-compression} and link prediction in 
 incomplete networks \cite{link-prediction}.
@@
 In this work, we
-restrict our attention to vertex-labelled networks, and we
-refer to the vertex labels as {\em features}.
-A common goal is to determine whether a given feature impacts graphical structure.
+restrict our attention to vertex-labelled networks,referring to the vertex labels as {\em features}.
+A typical goal is to determine whether a given feature impacts graphical structure.
 To answer this from a Bayesian perspective requires the use of a random graph model; the standard is the stochastic block model (SBM) \cite{vanilla-sbm}.
@@
 Effectively, these extend the model to allow each vertex to belong to multiple blocks simultaneously.
-However, one drawback of these graphical models as applied to labelled networks is that they do not automatically include vertex features in the random graph generation process.
+However, these models do not automatically include vertex features in the random graph generation process.
 Approaches based on graph neural networks \cite{sbm-gnn} that utilise vertex features have been developed but these lack the easy interpretability of the simpler models.
 
-To analyse a labelled network using one of the simple SBM variants, a typical inference procedure would be to first partition the graph into blocks grouped by distinct values of the feature of interest, and then use the associated model to test for evidence of heterogeneous connectivity between the feature-grouped blocks.
-Nevertheless, this approach is limited as it can only consider one feature at a time.
-This makes it difficult to rank-order the features by  magnitude of impact.
-Lastly, the feature-grouped blocks are often an unnatural partition of the graph, leading to a poor model fit.
+To analyse a labelled network using one of the simple SBM variants, a typical procedure would be to first partition the graph into blocks grouped by distinct values of the feature of interest.
+The associated model can then be used to test for evidence of heterogeneous connectivity between the feature-grouped blocks.
+Nevertheless, this approach can only consider one feature set at a time, making rank-ordering the features by importance difficult.
+Furthermore, the feature-grouped blocks are often an unnatural partition of the graph, leading to a poor model fit.
+
 We would instead prefer to partition the graph into its most natural blocks and then find which of the available features -- if any -- best predict the resulting partition.
-
 Thus motivated, we present a novel framework for modelling labelled networks, which we call the feature-first block model (FFBM).

### 2021-10-30 publication/prelim.tex
cosmetic changes
Edit first few sections

--- 2021-05-29 cued/prelim.tex
@@
 We first need a model for community-like structure in a network. For this we adopt the widely-used stochastic block model (SBM).
-Each node in the graph belongs to a unique community called a block.
-The probability that two nodes are connected depends only on the block memberships of each.
 Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}.
 To allow for degree-variability between members of the same block, we adopt the degree-corrected formulation (DC-SBM), defined in~(\ref{defn:microcan-dc-sbm}).
-
-For each integer $K\geq 1$, we use the notation $[K]:=\{1,2,\ldots,K\}$.
 
@@
 	Let $N \geq 1$ denote the number of vertices in an undirected graph. The block memberships are encoded by a vector $b \in [B]^N$,
-where $B$ is the number of non-empty blocks.
-	Let $e=(e_{rs})$ be the $N\times N$ symmetric matrix of edge counts 
+where $B$ is the number of non-empty blocks.\footnote{For each integer $K\geq 1$, we use the notation $[K]:=\{1,2,\ldots,K\}$.}
+	Let $e=(e_{rs})$ be the $B \times B$ symmetric matrix of edge counts 
 between blocks, such that $e_{rs}$ is the number of edges from block $r$ to 
-block $s$ -- or twice that number if $r=s$. 
-	Let $k =(k_i)$ denote the vector of length $N$ containing the degree
-sequence of the graph, with $k_i$ being the degree of vertex $i$.
+block $s$. 
+	Let $k =(k_i)$ denote a vector of length $N$, with $k_i$ being the degree of vertex $i$.
 
@@
 \end{definition}
-
-Appendix \ref{appdx:sbm} describes how to compute the likelihood, $p(A|b,e,k)$, for graphs generated from the microcanonical DC-SBM.

### 2021-10-31 publication/experiments.tex
Reduce to two figures

--- 2021-10-26 publication/experiments.tex
@@
 
+\begin{figure}[!h]
+	\centering
+	\begin{subfigure}[t]{0.32\linewidth}
+			\centering
+			\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{polbooks-null-1}}
+			\caption{Political Books.}
+			\label{fig:polbooks-null}
+		\end{subfigure}
+	\begin{subfigure}[t]{0.32\linewidth}
+			\centering
+			\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{school-null-1}}
+			\caption{Primary School.}
+			\label{fig:school-null}
+		\end{subfigure}
+	\begin{subfigure}[t]{0.32\linewidth}
+			\centering
+			\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{fb-null-1}}
+			\caption{Facebook Egonet.}
+			\label{fig:fb-null}
+		\end{subfigure}
+	\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
+\end{figure}
+%
+\begin{figure}[!h]
+	\centering
+	\begin{subfigure}[t]{0.32\linewidth}
+			\centering
+			\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{polbooks-accuracy-2}}
+			\caption{Political Books.}
+			\label{fig:polbooks-null}
+		\end{subfigure}
+	\begin{subfigure}[t]{0.32\linewidth}
+			\centering
+			\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{school-accuracy-1}}
+			\caption{Primary School.}
+			\label{fig:school-null}
+		\end{subfigure}
+	\begin{subfigure}[t]{0.32\linewidth}
+			\centering
+			\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{fb-accuracy-1}}
+			\caption{Facebook Egonet.}
+			\label{fig:fb-null}
+		\end{subfigure}
+	\caption{Per-block accuracy $\eta(j)$.}
+\end{figure}
+
 \subsection{Political books}
@@
 of the overall network structure. We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral).
-%
-\begin{figure}[!h]
-	\centering
-	\begin{subfigure}[t]{0.452\linewidth}
-		\centering
-		\vskip 0pt
-		\includegraphics[width=\linewidth]{polbooks-null-1}
-		\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
-		\label{fig:polbooks-null}
-	\end{subfigure}
-	\begin{subfigure}[t]{0.45\linewidth}
-		\centering
-		\vskip 0pt
-		\includegraphics[width=\linewidth]{polbooks-accuracy-1}
-		\caption{Per-block accuracy $\eta(j)$.}
-		\label{fig:polbooks-accuracy}
-	\end{subfigure}
-	\caption{Political books dataset.}
-	\label{fig:polbooks}
-\end{figure}
 
 From Figure~\ref{fig:polbooks-null} we see that all 3 blocks have a distinct political affiliation as their largest positive component.  
-This strongly indicates that political affiliation is indeed the axis 
-which best predicts the 3-way natural partition of the graph into blocks. 
 Furthermore, in Table~\ref{tab:results} we see that the training and test losses 
-are very similar and both are low in magnitude. This provides further evidence 
-to the claim that political affiliation is a very appropriate explanatory 
+are very similar and both are low in magnitude. This is strong evidence 
+that political affiliation is a very appropriate explanatory 
 variable for the overall network structure.
@@
 meaning that these are not good predictors of overall macro-structure.
-%
-\begin{figure}[!h]
-	\centering
-	\begin{subfigure}[t]{0.45\linewidth}
-		\centering
-		\imagebox{0.8\linewidth}{\includegraphics[width=\linewidth]{school-null-1}}
-		\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
-		\label{fig:school-null}
-	\end{subfigure}
-	\begin{subfigure}[t]{0.45\linewidth}
-		\centering
-		\imagebox{0.8\linewidth}{\includegraphics[width=\linewidth]{school-accuracy-1}}
-		\caption{Per-block accuracy $\eta(j)$.}
-		\label{fig:school-accuracy}
-	\end{subfigure}
-	\caption{Primary school dynamic contacts dataset.}
-	\label{fig:school}
-\end{figure}
 
@@
 The only exception to this rule is block 9 which nonetheless may have high magnitude weights just below the cut-off $c^*$.
-%
-\begin{figure}[!h]
-	\centering
-	\begin{subfigure}[t]{0.45\linewidth}
-		\centering
-		\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{fb-null-1}}
-		\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
-		\label{fig:fb-null}
-	\end{subfigure}
-	\begin{subfigure}[t]{0.45\linewidth}
-		\centering
-		\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{fb-accuracy-1}}
-		\caption{Per-block accuracy $\eta(j)$.}
-		\label{fig:fb-accuracy}
-	\end{subfigure}
-	\caption{Facebook egonet dataset.}
-	\label{fig:fb}
-\end{figure}
 

### 2021-10-31 publication/inference.tex
Reduce to two figures

@@
 		% arrows
-		\draw[->] (b0) to node[above] {$\sim q_b(b^{(t)}, b' | A)$} (b1);
-		\draw[->] (b1) to node[above] {$\alpha_b (b, b' | A)$} (b2);
-		\draw[->] (b0) [out=30, in=150] to node[above] {$1-\alpha_b$} (b2);
+		\draw[->] (b0) to node[above] {$\sim q_b ( b^{(t)}, b' | A )$} (b1);
+		\draw[->] (b1) to node[above] {$\alpha_b (b, b' | A )$} (b2);
+		\draw[->] (b0) [out=25, in=155] to node[above] {$1-\alpha_b$} (b2);
 
@@
 		\draw[->] (t1) to node[above] {$\alpha_\theta (\theta, \theta' | X, b^{(t)})$} (t2);
-		\draw[->] (t0) [out=-30, in=-150] to node[below] {$1-\alpha_\theta$} (t2);
+		\draw[->] (t0) [out=-25, in=-155] to node[below] {$1-\alpha_\theta$} (t2);
 
@@
 %
-Now $\pi_b(\cdot)$ is the un-normalised target density,
-which can be expressed as:
-%  we wish to sample from for the $b$-chain. In other words, we 
-% wish to construct a Markov chain that has $\pi_b(\cdot)$ 
-% as its invariant distribution. We can break $\pi_b$ down as follows:
-%
-\begin{equation}
-	\pi_b(b) = p(b|X) \sum_{\psi} \nolimits p(A , \psi | b, X) \\
-	= p(b|X) p(A, \psi^* | b, X) \\
-	= p(A | b, \psi^*) \cdot p(\psi^* | b) \cdot p(b | X).
-\end{equation}
-%
+Now $\pi_b(\cdot)$ is the un-normalised target density.
 Since we are using the microcanonical SBM formulation, there is only one 
@@
 We denote this value $\psi^* = \{\psi_k^*, \psi_e^*\}$. Therefore, 
-the summation over all $\psi$ reduces to just the single $\psi^*$ term.
-% ; this is the power of the microcanonical formulation. 
+the summation over all $\psi$ needed to evaluate $p(A | b, X)$ reduces to just the single $\psi^*$ term:
+$p(A | b, X) = \sum_{\psi} \nolimits p(A , \psi | b, X) = p(A, \psi^* | b, X)$.
 We also define the microcanonical entropy of the configuration as,
@@
 \begin{equation}
-	S(b) = - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big).
+	S(b) \coloneqq - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big).
 	\label{eqn:dl-form}
@@
 This entropy can be thought of as the optimal
-``description length'' of the graph; if we 
-constructed an optimal code for the SBM ensemble,
-this would be the minimum amount of information 
-required to store the graph and its parameters. 
+``description length'' of the graph. 
 The exact form of the proposal $q_b$ is explored thoroughly in
-\cite{Peixoto-MCMC} and not repeated here. There is a widely used 
-library for Python made available under LGPL 
-called \verb*|graph-tool| \cite{peixoto_graph-tool_2014}, 
-which implements this algorithm. The only modification we make is in 
+\cite{Peixoto-MCMC} and not repeated here. The \verb*|graph-tool| \cite{peixoto_graph-tool_2014}
+library for Python available under LGPL implements this algorithm.
+The only modification we make is in 
 the block membership prior $p(b)$ that we replace with $p(b|X)=B^{-N}$, 
@@
 This leads to the proposal distribution,
-\begin{equation*}
-	q_\theta(\theta, \theta') \sim \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right).
-\end{equation*}
-%
-Without the injected noise term, MALA is equivalent to gradient descent. We require the noise term $\xi$ to fully explore the parameter space. 
-The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}). By noting that $\theta = \{w_k\}_{k=1}^{B}$, we write the derivative with respect to each $w_k$ as,
-%
-\begin{equation}
-	\frac{\partial U}{\partial w_k} = - \left( \sum_{i \in [N]} \Big\{ x_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right).
-	\label{eqn:U-derivative}
-\end{equation}
-%
-After a proposed move is generated from $q_\theta$,
-it is either accepted or rejected in the standard
-Metropolis-Hastings accept/reject step.
+$q_\theta(\theta, \theta') \sim \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right).$
+Without the injected noise term $\xi$, MALA is equivalent to gradient descent. We require $\xi$ to fully explore the parameter space. 
+The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}).
 
@@
 
-Up to this point, each $\theta^{(t)}$ update uses its corresponding $b^{(t)}$ sample.
-This means that the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance.
-This may lead to longer burn-in for the resulting Markov chain.
+So far, each $\theta^{(t)}$ update has used its corresponding $b^{(t)}$ sample.
+This means the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance, leading to longer burn-in.
 The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends only on the matrix $y^{(t)}$ with entries $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
@@
 \begin{equation}
-	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one(b_{i}^{(t)} = j) \right]
+	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one \left( b_{i}^{(t)} = j \right) \right]
 	= p(b_i = j | A, X).
@@
 concurrently. Instead, we run the $b$-chain to completion and use it 
-to generate $\hat{y}$. This affords us the flexibility to vary the 
-lengths of the $b$ and $\theta$-chains. Furthermore, the changeover 
-from $y^{(t)}$ to $\hat{y}$ reduces the burn-in time for 
-the $\theta$-chain by reducing the variance in the evaluation 
-of $U$ and $\nabla U$.
+to generate $\hat{y}$ also allowing us to vary the lengths of each. This change shortens the burn-in of the $\theta$-chain by reducing the variance of our proposal distribution.
 
@@
 The multiplier $k$ in~(\ref{eqn:fij}) determines the degree of significance of the result.
-However, as the Laplace approximation is not exact, we only treat this dimensionality reduction method as a useful heuristic and not an exact method.
-
+However, as the Laplace approximation is not exact, we regard this as a useful heuristic and not an exact method.
 Conversely, we could fix $k=k_0$ and the dimension of our reduced feature set $|\Dcal'|=D'$.
-We would then like to find the largest value of $c$ such that $|\Dcal'|=D'$ given $k=k_0$.
-This is summarised in equation \ref{eqn:c-star}.
+We would then like to find the largest value of $c$ such that $|\Dcal'|=D'$ given $k=k_0$:
 %
@@
 %
-This approach has the advantage that it fixes the number of reduced dimensions.
+This has the advantage of fixing the number of reduced dimensions.

### 2021-10-31 publication/experiments.tex
Banter
Brevify polbooks

@@
 			\caption{Political Books.}
-			\label{fig:polbooks-null}
+			\label{fig:polbooks-accuracy}
 		\end{subfigure}
@@
 			\caption{Primary School.}
-			\label{fig:school-null}
+			\label{fig:school-accuracy}
 		\end{subfigure}
@@
 			\caption{Facebook Egonet.}
-			\label{fig:fb-null}
+			\label{fig:fb-accuracy}
 		\end{subfigure}
@@
 
-The goal here is to 
-determine whether the authors' political affiliations are a good predictor 
-of the overall network structure. We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral).
-
+We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral).
 From Figure~\ref{fig:polbooks-null} we see that all 3 blocks have a distinct political affiliation as their largest positive component.  
@@
 This suggests that detected block 1 is not solely composed of ``neutral" books but also 
-contains some ``liberal" and ``conservative" authors. Indeed, by examining 
-Figure~\ref{fig:polbooks-graph} we see that block 1 is effectively a buffer between 
-blocks 2 and 3; there are very few edges between blocks 2 and 3. It is therefore not 
-surprising that some books from either side leak into block 1.
-The three distinct categories may be 
-too coarse a measure of political affiliation; perhaps the nominally 
-``conservative" books found in block 1 are closer to the centre. 
-In the absence of more granular labels, we cannot test this hypothesis. 
-Nevertheless, political affiliation encoded as 3 distinct labels 
-remains a fantastic predictor of network structure.
+contains some ``liberal" and ``conservative" authors. Examining 
+Figure~\ref{fig:polbooks-graph}, we see the majority of paths between blocks 2 and 3 go through block 1.
+Block 1 is in effect a bridge between the ``conservative'' and ``liberal'' blocks so it is unsurprising that some books from either side leak into block 1.
 

### 2021-11-02 publication/experiments.tex
Brevify experiments

@@
 
-\begin{figure}[!h]
+\begin{figure}
 	\centering
-	\begin{subfigure}[t]{0.32\linewidth}
+	\begin{minipage}{0.42\linewidth}
 		\centering
-			\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{polbooks-null-1}}
+		\begin{subfigure}{\linewidth}
+			\centering
+			\imagebox{0.8\linewidth}{\includegraphics[width=\linewidth]{polbooks-null-1}}
 			\caption{Political Books.}
@@
 		\end{subfigure}
-	\begin{subfigure}[t]{0.32\linewidth}
+
+		\begin{subfigure}{\linewidth}
 			\centering
-			\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{school-null-1}}
+			\imagebox{0.85\linewidth}{\includegraphics[width=\linewidth]{school-null-1}}
 			\caption{Primary School.}
@@
 		\end{subfigure}
-	\begin{subfigure}[t]{0.32\linewidth}
+
+		\begin{subfigure}{\linewidth}
 			\centering
@@
 		\end{subfigure}
-	\caption{$\theta$-samples. Dotted line is $\pm c^*$.}
-\end{figure}
+		\caption{$\theta$-samples. Dotted line $\pm c^*$.}
+	\end{minipage}
 	%
-\begin{figure}[!h]
+	\begin{minipage}{0.42\linewidth}
 		\centering
-	\begin{subfigure}[t]{0.32\linewidth}
+		\begin{subfigure}[t]{\linewidth}
 			\centering
-			\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{polbooks-accuracy-2}}
+			\imagebox{0.8\linewidth}{\includegraphics[width=\linewidth]{polbooks-accuracy-2}}
 			\caption{Political Books.}
@@
 		\end{subfigure}
-	\begin{subfigure}[t]{0.32\linewidth}
+
+		\begin{subfigure}{\linewidth}
 			\centering
-			\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{school-accuracy-1}}
+			\imagebox{0.85\linewidth}{\includegraphics[width=\linewidth]{school-accuracy-1}}
 			\caption{Primary School.}
@@
 		\caption{Per-block accuracy $\eta(j)$.}
+	\end{minipage}
 \end{figure}
 
-\subsection{Political books}
+\subsubsection{Political books.}
 
@@
 
-\subsection{Primary school dynamic contacts}
+\subsubsection{Primary school dynamic contacts.}
 
 We choose the number of communities $B=10$, in line with the total number of 
-school classes. We 
-employ the dimensionality reduction technique to pick out the top $D'=10$ features. 
-We then plot the weights for the resulting features $d \in \Dcal'$ 
-in Figure~\ref{fig:school-null}. It is rather obvious
-that only the pupils' class memberships (1A-5B) have remained
-significant;
+school classes. From Figure~\ref{fig:school-null}, we see only the pupils' class memberships (1A-5B) have survived
+the dimensionality-reduction process for $D'=10$;
 gender and teacher/student status have been discarded,
-meaning that these are not good predictors of overall macro-structure.
+meaning these are poor predictors of overall macro-structure.
 
@@
 detected blocks (class 2B spans blocks 8 and 9) but we do 
-not have a feature which explains the division. The most surprising block 
-is number 7, which has comparable weightings for classes 5A and 1B. 
-Perhaps there was a joint event between those two classes on the day 
-the data were collected.
+not have a feature which explains the division.
 
@@
 In fact the only blocks with low accuracy are those that have a school-class span two blocks such that we cannot reliably distinguish between the two.
-This is much more pronounced when we apply hard classification rather than the soft cross-entropy loss.
+This is more pronounced when we apply hard classification rather than the soft cross-entropy loss.
 Indeed block 1 has low accuracy because class 1A has a higher weight component for block 3 than it does for block 1.
@@
 
-\subsection{Facebook egonet}
+\subsubsection{Facebook egonet.}
 
@@
 community structure. The majority of them are education related. 
-Nevertheless, for $D'=10$ we only have good explanations for the makeup 
-of some of the detected blocks; several blocks in 
+Nevertheless, for $D'=10$ we only have good explanations for some of the detected blocks; several blocks in 
 Figure~\ref{fig:fb-null} do not have high-magnitude components for $D'=10$.
@@
 high weight as it relates to block 1 -- but it is highly unlikely
-that birthdays determine graphical structure. This issue could have been alleviated by 
-choosing $\Xcal=\{-1, +1\}$ such that having a feature turned off supports excluding 
-a vertex from the block. However, Figure~\ref{fig:fb-null} demonstrates why we made 
-the deliberate choice $\Xcal=\{0,1\}$. 
-Because of the choice $\Xcal=\{0,1\}$, block 1 is allowed to contain 3 distinct feature 
-categories rather than just one. We accept this trade-off: risking spurious high magnitude 
-components (e.g. \verb*|birthday-5|) for the benefit of allowing a top-level block to span 
-multiple feature-groups. Nevertheless, one would have to apply a hierarchical approach 
-to subdivide each top-level block into sub-blocks to test whether the constituent 
-feature-groups determine structure at the lower-level.
+that birthdays determine graphical structure.

### 2021-11-03 publication/conclusion.tex
Editing
Fix citet

--- 2021-06-01 cued/conclusion.tex
@@
 The Feature-First Block Model (FFBM) introduced 
-in this paper is a new generative model for labelled networks,
-developed to address difficulties of other graphical models 
-when testing how vertex features affect community structure. The idea 
-is to divide the graph into its most natural partition and determine whether 
+in this paper is a new generative model for labelled networks well-suited for describing how features
+affect structure. The idea is to divide the graph into its most natural partition and determine whether 
 the vertex features can accurately explain this partition. 
-It is relatively easy to find vertex features that are in some way 
-correlated with the graphical structure. Nonetheless, only when 
-we find the feature that best describes the most pronounced partition 
+It is easy to find vertex features that are in some way 
+correlated with the graphical structure; only when 
+we find the feature that best describes the most pronounced partition,
 do we have a stronger case for causation.
@@
 the parameters of the FFBM. 
-This takes the form of a two-level Markov chain,
-used to sample the block memberships $b$ and block generator 
-parameters $\theta$. These chains can either be executed
-in parallel or 
-the empirical mean of the $b$-samples can be used
-as the input to the $\theta$-chain. The latter option
-reduces the variance in our evaluation of the target distribution 
-and thus shortens burn-in.
-
-The overall method is shown to be effective at extracting and describing 
+This is shown to be effective at extracting and describing 
 the most natural communities in a labelled network. Nevertheless, the approach 
-can only currently explain the structure at the macro-scale. We cannot 
-explain structure within each block. Future work will benefit from extending 
+can only currently explain the structure at the macro-scale. Future work will benefit from extending 
 the FFBM to be hierarchical in nature. That way, the structure of the network 
-can be explained at all length-scales of interest. So long as data 
-collection techniques remain ethical and care is taken to respect 
-personal privacy, such empowered decision-making can only help humankind.
+can be explained at all length-scales of interest.

### 2021-11-03 publication/extensions.tex
Fix citet

@@
 
-The greatest limitation of the current FFBM formulation is that it can only explain structure at the macro-scale.
-In other words, we cannot explain structure within each detected block.
+The current FFBM formulation can only explain macro-structure and not structure within each detected block.
 Future work will benefit from extending the FFBM to be hierarchical in nature.
-This would be a relatively natural extension.
 Indeed, the SBM has already been extended to a hierarchical form, often called the nested SBM \cite{SBM-hierarchical}.
-The idea is to divide each block into sub-blocks and so on recursively until a specified depth is reached.
-The full block membership for a particular vertex now encodes the memberships at all levels of the hierarchy.
 
@@
 Given the nested SBM, we would have a hierarchy of generators, each generating a block membership at a particular level of the hierarchy.
-Nevertheless, this does pose some practical issues for scalability; supposing we have $L$ levels in our hierarchy and each divides the parent block into $B$ sub-blocks, then the number of distinct generators necessary scales as $O(B^L)$.
 To avoid exponential growth in the number of model parameters, we could apply some form of dimensionality reduction as we descend the layers so that each generator is only given relevant features as input.

### 2021-11-03 publication/ffbm.tex
Brevify polbooks
Fix citet

@@
 
-In this section we propose a novel generative model for labelled networks. We call this the feature-first block model (FFBM -- illustrated in figure~\ref{fig:ffbm}).
+In this section we propose a novel generative model for labelled networks. We call this the feature-first block model (FFBM -- illustrated in Figure~\ref{fig:ffbm}).
 
@@
 
-\citet{Peixoto-Bayesian-Microcanonical} proposes careful choices for 
+Peixoto \cite{Peixoto-Bayesian-Microcanonical} proposes careful choices for 
 the priors on the additional microcanonical SBM parameters $\psi$, which we adopt. 

### 2021-11-03 publication/abstract.tex
cosmetic changes
1-page intro

@@
 (or {\em features}) affect the network's structure.
-
-A standard approach has been to partition the network into blocks grouped
-by distinct values of the feature of interest. A block-based random
-graph model -- typically a variant of the stochastic block model --
-is then used to test for evidence of asymmetric behaviour between these
-feature-based communities. Nevertheless, the resulting communities
-often do not produce a natural partition of the graph.
 
@@
 
-\keywords{Stochastic Block Models \and Bayesian Inference.}
+\keywords{Labelled Networks \and Bayesian Inference.}
 

### 2021-11-03 publication/introduction.tex
1-page intro

@@
 Finding ways to recover the latent communities from the observed graph is an important
-task in many applications, including
-network compression \cite{cluster-compression} and link prediction in 
-incomplete networks \cite{link-prediction}.
+task in many applications. In this work, we restrict our attention to vertex-labelled networks, 
+referring to the labels as {\em features}.
+A typical goal is to determine whether a given feature impacts graphical structure.
+Answering this requires a random graph model; the standard is the stochastic block model (SBM) \cite{vanilla-sbm}.
+Numerous variants of the SBM  have been proposed -- such as the MMSBM \cite{mixed-membership-sbm} and OSBM \cite{overlapping-sbm}.
+However, these do not automatically include vertex features in the graph generation process.
 
-In this work, we
-restrict our attention to vertex-labelled networks,referring to the vertex labels as {\em features}.
-A typical goal is to determine whether a given feature impacts graphical structure.
-To answer this from a Bayesian perspective requires the use of a random graph model; the standard is the stochastic block model (SBM) \cite{vanilla-sbm}.
-This is a latent variable model where each vertex belongs to a single block and the probability two nodes are connected depends only on the block memberships of each.
-Numerous variants of this
-model have been considered -- the most popular ones being the mixed-membership stochastic block model (MMSBM) \cite{mixed-membership-sbm} and the overlapping stochastic block model (OSBM) \cite{overlapping-sbm}.
-Effectively, these extend the model to allow each vertex to belong to multiple blocks simultaneously.
-However, these models do not automatically include vertex features in the random graph generation process.
-Approaches based on graph neural networks \cite{sbm-gnn} that utilise vertex features have been developed but these lack the easy interpretability of the simpler models.
-
-To analyse a labelled network using one of the simple SBM variants, a typical procedure would be to first partition the graph into blocks grouped by distinct values of the feature of interest.
+To analyse a labelled network using one of the simple SBM variants, a typical procedure would be to partition the graph into blocks grouped by distinct values of the feature of interest.
 The associated model can then be used to test for evidence of heterogeneous connectivity between the feature-grouped blocks.
-Nevertheless, this approach can only consider one feature set at a time, making rank-ordering the features by importance difficult.
-Furthermore, the feature-grouped blocks are often an unnatural partition of the graph, leading to a poor model fit.
+Nevertheless, this approach can only consider one feature set at a time and the feature-grouped blocks are often an unnatural partition of the graph.
 
@@
 This is an extension of the SBM to labelled networks.
-We go on to present an efficient algorithm for sampling from the FFBM parameters, and to describe how the sampled parameters can be interpreted to determine which features have the largest impact on overall graphical structure.

### 2021-11-03 publication/prelim.tex
Fix citet
1-page intro

@@
 
-We first need a model for community-like structure in a network. For this we adopt the widely-used stochastic block model (SBM).
-Specifically, we will use the microcanonical variant of the SBM, proposed by \citet{Peixoto-Bayesian-Microcanonical}.
+We first need a model for community-like structure in a network.
+For this we adopt the widely-used stochastic block model (SBM).
+This is a latent variable model where each vertex belongs to a single block and the probability two nodes are connected depends only on the block memberships of each.
+Specifically, we will use the microcanonical variant of the SBM, proposed by Peixoto \cite{Peixoto-Bayesian-Microcanonical}.
 To allow for degree-variability between members of the same block, we adopt the degree-corrected formulation (DC-SBM), defined in~(\ref{defn:microcan-dc-sbm}).
@@
 by placing edges uniformly at random, but respecting the constraints 
-imposed by $e$, $b$ and $k$ (hence the qualification `microcanonical').
-Specifically, $A$ must satisfy the following, for all $r,s\in[B]$
+imposed by $e$, $b$ and $k$.
+Specifically, if $A \sim \mbox{\rm DC-SBM}_{\rm MC} (b, e, k)$ then it must satisfy the following, for all $r,s\in[B]$
 and all $i\in[N]$:
@@
 	\end{equation}
-We use the following notation to indicate this distribution:
-\begin{equation}
-		A \sim \mbox{\rm DC-SBM}_{\rm MC} (b, e, k).
-	\end{equation}
 \end{definition}

### 2021-11-03 publication/inference.tex
Fix citet
dimensionality reduction

@@
 %
-This is an example of a pseudo-marginal approach. Indeed, \citet{pseudo-marginal} show that~(\ref{eqn:theta-unbiased})
+This is an example of a pseudo-marginal approach. Indeed, Andrieu and Roberts \cite{pseudo-marginal} show that~(\ref{eqn:theta-unbiased})
 % the unbiased result in equation 
@@
 Once the samples $\left\{ \theta^{(t)} \right\} \sim p(\theta | A, X)$
-have been obtained, we can compute the empirical mean and standard deviation of each component of $\theta$.
-Switching back to matrix notation, we define $\theta = W$, such that $W_{ij}$ is the weight component for block $i$ and feature $j$.
-We can then define:
+have been obtained, we can compute the empirical mean and standard deviation of each component of $\theta = \{ w_i \}_{i=1}^{B}$:
 %
 \begin{equation}
-	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
-	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2.
+	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} w_{ij}^{(t)} \qquad \textrm{and} \qquad
+	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( w_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2.
 \end{equation}
@@
 $(-c, c)$ for all block indices $i$.
-If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \hat{\mu}_{ij}, \hat{\sigma}_{ij}^2)$, then this is analogous to a hypothesis test on the value of $W_{ij}$ as in~(\ref{eqn:hyp-test-discard});
-then $\Dcal'$ comprises all features $j$ for which $H_1$ is accepted at least once for some $i \in [B]$.
-%
-\begin{equation}
-	H_0: |W_{ij}| \leq c \qquad
-	H_1: |W_{ij}| > c
-	\label{eqn:hyp-test-discard}
-\end{equation}
-%
-The multiplier $k$ in~(\ref{eqn:fij}) determines the degree of significance of the result.
+If we were to use the Laplace approximation for the posterior $p(w_{ij} | A, X) \approx \Gaussian(w_{ij}; \hat{\mu}_{ij}, \hat{\sigma}_{ij}^2)$, then this would be analogous to a hypothesis test on the magnitude of $w_{ij}$ compared to $c$ with multiplier $k$ in~(\ref{eqn:fij}) determines the degree of significance of the result.
 However, as the Laplace approximation is not exact, we regard this as a useful heuristic and not an exact method.
-Conversely, we could fix $k=k_0$ and the dimension of our reduced feature set $|\Dcal'|=D'$.
-We would then like to find the largest value of $c$ such that $|\Dcal'|=D'$ given $k=k_0$:
+Conversely, if we want to fix the number of dimensions in our reduced feature set $|\Dcal'|=D'$, the problem then becomes finding the largest value of $c$ such that $|\Dcal'|=D'$ given $k=k_0$:
 %
@@
 \end{equation}
-%
-This has the advantage of fixing the number of reduced dimensions.

### 2021-11-03 publication/ffbm.tex
1-page intro
dimensionality reduction
Misc

@@
-\section{Feature-first block model}
+\section{Feature-First Block Model}
 
@@
 	\end{tikzpicture}
-	\caption{The feature-first block model (FFBM)}
+	\caption{The Feature-First Block Model (FFBM)}
 	\label{fig:ffbm}
@@
 priors on $\theta$ and $\psi$ must also be specified. 
-We place a Gaussian prior on $\theta$ with zero mean
-and covariance matrix $\sigma^2_\theta I$, so that
+We place a Gaussian prior on $\theta$ such that
 each element of $\theta$ has an independent ${\cal N}(0,\sigma_\theta^2)$
@@
 This choice of prior gives a very
-simple form for the conditional distribution of the block membership vector $b$ given $X$. 
-It is a uniform distribution (\ref{eqn:b-pseudo-prior})
- -- proof in Appendix~\ref{appdx:b|x}. This an important simplification as evaluating $p(b | X)$ does not require an expensive  integration over $\theta$ nor does it require the exact value of $X$.
+simple form for the conditional distribution of the block membership vector $b$ given $X$; it is a uniform distribution:
 %
@@
 %
-
+The proof is given in Appendix~\ref{appdx:b|x}.
+This an important simplification as evaluating $p(b | X)$ does not require an expensive  integration over $\theta$ nor does it require the exact value of $X$.
 Peixoto \cite{Peixoto-Bayesian-Microcanonical} proposes careful choices for 

### 2021-11-03 publication/inference.tex
Misc

@@
 Once the samples $\left\{ \theta^{(t)} \right\} \sim p(\theta | A, X)$
-have been obtained, we can compute the empirical mean and standard deviation of each component of $\theta = \{ w_i \}_{i=1}^{B}$:
+have been obtained, we can compute the empirical mean and standard deviation of each component of $\theta$. Switching to matrix notation $\theta = W$ we define:
 %
 \begin{equation}
-	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} w_{ij}^{(t)} \qquad \textrm{and} \qquad
-	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( w_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2.
+	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
+	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2.
 \end{equation}
@@
 $(-c, c)$ for all block indices $i$.
-If we were to use the Laplace approximation for the posterior $p(w_{ij} | A, X) \approx \Gaussian(w_{ij}; \hat{\mu}_{ij}, \hat{\sigma}_{ij}^2)$, then this would be analogous to a hypothesis test on the magnitude of $w_{ij}$ compared to $c$ with multiplier $k$ in~(\ref{eqn:fij}) determines the degree of significance of the result.
+If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \hat{\mu}_{ij}, \hat{\sigma}_{ij}^2)$, then this would be analogous to a hypothesis test on the magnitude of $W_{ij}$ compared to $c$ with multiplier $k$ in~(\ref{eqn:fij}) determines the degree of significance of the result.
 However, as the Laplace approximation is not exact, we regard this as a useful heuristic and not an exact method.

### 2021-11-03 publication/inference.tex
MH sampling on one page

@@
 We propose an iterative approach to obtain samples
-$\theta^{(t)}$ from $p(\theta|A,X)$. We first draw samples $b^{(t)}$ 
-from the block membership posterior distribution,
-and then use each $b^{(t)}$ to obtain a corresponding
+$\theta^{(t)}$ from $p(\theta|A,X)$. We first draw a sample $b^{(t)}$ 
+from the block membership posterior,
+and then use $b^{(t)}$ to obtain a corresponding
 sample $\theta^{(t)}$:
 %
-\begin{align}
-	b^{(t)} &\sim p ( b | A, X )  \label{eqn:b-samples},\\
-	\theta^{(t)} &\sim p\left(\theta | X, b^{(t)} \right). \label{eqn:theta-samples}
-\end{align}
+\begin{equation}
+	b^{(t)} \sim p \Big( b | A, X \Big) \quad \textrm{then} \quad
+	\theta^{(t)} \sim p\left(\theta | X, b^{(t)} \right).
+\end{equation}
 %
-Both of these sampling steps can be implemented through a
-Markov chain Monte-Carlo approach via a Metropolis-Hastings 
-algorithm \cite{hastings-alg}, as a two-level Markov chain.
-The resulting samples 
-for $\theta^{(t)}$ are unbiased in the sense that the expectation of 
-their distribution is the target posterior distribution:
+This can be achieved through a two-tiered
+Markov chain Monte-Carlo approach via the Metropolis-Hastings (MH) 
+algorithm \cite{hastings-alg}.
+The resulting $\theta^{(t)}$ samples are unbiased in that the expectation of 
+their distribution is the true posterior:
 %
@@
 is sufficient to prove that, for large enough $t$, $\theta^{(t)} \sim p(\theta | A, X)$, as required.
-
+%
 \begin{figure}[!h]
@@
 
-The reason for splitting the Markov chain into two levels is that the summation over all latent states $b \in [B]^N$ required to directly compute the likelihood, $p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$, is intractable, as the sum involves $B^N$ terms.
-Figure \ref{fig:samp-sequence} shows an overview of the proposed method. 
-The Metropolis-Hastings proposal distribution is denoted $q$ and the acceptance probability denoted $\alpha$.
+We have split the Markov chain into two levels as the summation over all latent $b \in [B]^N$ required to directly compute the likelihood, $p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$, is intractable.
+Figure \ref{fig:samp-sequence} shows an overview of the proposed method, with $q$ and $\alpha$ denoting the MH proposal distribution and acceptance probability respectively.
 
 Note the importance of the simplification in~(\ref{eqn:b-pseudo-prior}). 
-As $p(b| X)$ in fact does not depend on $X$, 
-knowing $X$ is not necessary in order to obtain samples
-of the block membership vector $b$.
+As evaluating $p(b| X)$ does not depend on $X$, 
+we do not need $X$ to sample $b$.
 And on the other level, in order to obtain 
-samples of the parameter vector $\theta$
-we use only $b$ but not $A$, as $\theta$ and $A$ are
-conditionally independent given $b$.
+samples for $\theta$
+we use only $b$ but not $A$, as $(\theta \indep A )| b$.
 
@@
 
-We adopt the Markov chain Monte-Carlo procedure of
+To generate our $b$-samples, we adopt the Markov chain Monte-Carlo procedure of
 \cite{Peixoto-MCMC},
@@
 %
-This entropy can be thought of as the optimal
+This can be thought of as the optimal
 ``description length'' of the graph. 
 The exact form of the proposal $q_b$ is explored thoroughly in
-\cite{Peixoto-MCMC} and not repeated here. The \verb*|graph-tool| \cite{peixoto_graph-tool_2014}
-library for Python available under LGPL implements this algorithm.
-The only modification we make is in 
-the block membership prior $p(b)$ that we replace with $p(b|X)=B^{-N}$, 
+\cite{Peixoto-MCMC} and not repeated here. We use the \verb*|graph-tool| \cite{peixoto_graph-tool_2014}
+library for Python, which implements this algorithm.
+The only modification is in 
+the prior $p(b)$ that we replace with $p(b|X)=B^{-N}$, 
 which cancels out in the MH accept-reject step as it is independent of $b$.
@@
 
-The invariant distribution we wish to target for the $\theta$ samples is the posterior of $\theta$ given the values of the pair $(X, b)$. 
+The invariant distribution we wish to target for the $\theta$-samples is the posterior of $\theta$ given the values of the pair $(X, b)$. 
 We write this as,
@@
 	U(\theta) = \left( \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} \|\theta\|^2 = N \cdot \Lcal(\theta) + \frac{1}{2\sigma_\theta^2} \|\theta\|^2;
+	+ \frac{1}{2\sigma_\theta^2} \|\theta\|^2 = N \cdot \Lcal(\theta) + \frac{1}{2\sigma_\theta^2} \|\theta\|^2.
 	\label{eqn:U-form}
 \end{equation}
-
+%
 See Appendix \ref{appdx:form-U} for the derivation.
 The function $U(\theta)$ is a typical objective function for neural network training.
-The first term is introduced by the likelihood; we collect it into $N \cdot \Lcal(\theta)$, which is the cross-entropy between the graph-predicted and feature-predicted block memberships summed over all vertices.
-The second, introduced by the prior, brings a form of regularisation, guarding against over-fitting. Unlike in many applications where the goal 
-is to find the minimiser of $U(\theta)$, our goal is to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
-We can use $\nabla U$ as a useful heuristic to bias our proposal towards regions of higher target density.
-We therefore adopt the Metropolis-adjusted Langevin algorithm (MALA) \cite{mala-tweedie}.
+The first term $N \cdot \Lcal(\theta)$ is introduced by the likelihood and represents the cross-entropy between the graph-predicted and feature-predicted block memberships.
+The second term, introduced by the prior, brings a form of regularisation, guarding against over-fitting.
+Our goal is to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
+We choose to adopt the Metropolis-adjusted Langevin algorithm (MALA) \cite{mala-tweedie}, which uses $\nabla U$ to bias the proposal towards regions of higher density.
 Given the current sample $\theta$, we generate
@@
 \begin{equation*}
-	\theta' = \theta - h \nabla U(\theta) + \sqrt{2h} \cdot \xi,
+	\theta' \sim q_\theta\big(\theta, \theta'\big) 
+	= \Gaussian \big( \theta' ; \theta - h \nabla U(\theta), 2h I \big),
 \end{equation*}
@@
 which may vary with the sample index.
-This leads to the proposal distribution,
-$q_\theta(\theta, \theta') \sim \Gaussian \left( \theta' ; \theta - h \nabla U(\theta), 2h I \right).$
 Without the injected noise term $\xi$, MALA is equivalent to gradient descent. We require $\xi$ to fully explore the parameter space. 

### 2021-11-03 publication/inference.tex
6 pages up until experiments

@@
 So far, each $\theta^{(t)}$ update has used its corresponding $b^{(t)}$ sample.
-This means the evaluation of $U(\theta)$ and $\nabla U(\theta)$ has high variance, leading to longer burn-in.
-The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U(\theta)$ and $\nabla U(\theta)$ which depends only on the matrix $y^{(t)}$ with entries $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
+This means the evaluation of $U^{(t)}$ and $\nabla U^{(t)}$ has high variance, leading to longer burn-in.
+The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U^{(t)}$ and $\nabla U^{(t)}$ which depends only on the matrix $y^{(t)}$ with entries $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
 We would rather deal with the expectation of each $y_{ij}^{(t)}$:
@@
 for the $b$-samples and $\Tcal_\theta$ similarly for the $\theta$-chain. 
-The unbiased estimate for $y_{ij}^{(t)}$ using the 
-restricted sample set $\Tcal_b$ is denoted $\hat{y}_{ij}$:
+The unbiased estimate for $y_{ij}^{(t)}$ is then:
 %

### 2021-11-06 publication/derivations.tex
cut down appendices

--- 2021-06-01 cued/derivations.tex
@@
 \begin{align*}
-	p(b | X) &= \int p(b , \theta| X, \theta) d\theta = \int p(b | X, \theta) p(\theta | X) d\theta \\
-	&=\int p(b | X, \theta) p(\theta) d\theta = \int \prod_{i \in [N] } \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
+	p(b | X) 
+	&= \int p(b , \theta| X) d\theta 
+	= \int p(b | X, \theta) p(\theta | X) d\theta
+	= \int \prod_{i \in [N] } \phi_{b_i}(x_i; \theta) p(\theta) d\theta \\
 	&= \prod_{i \in [N]} \int \frac{\exp(w_{b_i}^T x_i) \prod_{j \in [B]} \Gaussian(w_j; 0, \sigma_\theta^2 I)}{\sum_{k \in [B]} \exp(w_{k}^T x_i)} dw_{1:B}.
@@
 	\log p(b | X, \theta) = \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
-	\log p(\theta) = -\frac{(D+1)(B)}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} 
+	\log p(\theta) = -\frac{D B}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} 
 	\|\theta \|^2,
@@
 Therefore, discarding constant terms, we 
-obtain:
+obtain,
 %
@@
 \end{equation}
-%
-This is exactly the representation~(\ref{eqn:U-form}), as claimed.
 
@@
 $\nabla U(\theta)$ can be computed explicitly.
-Recall the expression for $U(\theta)$ in~(\ref{eqn:U-form}).
-Writing $\theta$ as
-$\theta = \left[w_1^T, w_2^T \dots w_B^T  \right]^T$,
-in order to compute the gradient
+As
+$\theta = \{ w_k \}_{k=1}^{B}$,
+to compute the gradient
 $\nabla U(\theta)$ we need to compute
@@
 %
-Using~(\ref{eq:dadw}) and~(\ref{eq:dtsdw}), we obtain,
+Recall the expression for $U(\theta)$ in~(\ref{eqn:U-form}). Using~(\ref{eq:dadw}) and~(\ref{eq:dtsdw}), we obtain,
 \begin{align}

### 2021-11-06 publication/experiments-detailed.tex
Refactor experiments section

--- 2021-11-02 publication/experiments.tex
@@
-\section{Experimental results}
-\label{sec:experiments}
+\subsubsection{Political books.}
 
-We apply the developed methods to a variety of datasets. These are chosen to span a range of node counts $N$, edge counts $E$ and feature space dimension $D$. We consider the following:
-
-\begin{itemize}
-	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=3$) -- network of Amazon political book sales, published close to the 2004 presidential election.
+\cite{polbooks} ($N=105, E=441, D=3$) -- network of Amazon political book sales, published close to the 2004 presidential election.
 Two books are connected if they were frequently co-purchased.
 Vertex features encode the political affiliation of the author (liberal, conservative, or neutral).
-
-	\item \textbf{Primary school dynamic contacts} \cite{schools} ($N=238, E=5539, D=13$) -- network of face-to-face contacts amongst students and teachers at a primary school in Lyon, France.
-Vertex features include class membership (one of 10 values: 1A-5B), gender (male, female) and teacher status encoded as an 11th school-class.
-We choose to analyse just the second day of results.
-
-	\item \textbf{Facebook egonet} \cite{fb-snap} ($N=747, E=30025, D=480$) -- an assortment of Facebook users' friends lists.
-Vertex features are fully anonymised and encode information about each user's education history, languages spoken, gender, home-town, birthday etc.
-We focus on the egonet with id 1912.
-
-\end{itemize}
-
-\begin{figure}[!h]
-	\centering
-	\begin{subfigure}[t]{0.28\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{polbooks-graph.png}
-		\caption{Polbooks}
-		\label{fig:polbooks-graph}
-	\end{subfigure}
-	\hfill
-	\begin{subfigure}[t]{0.28\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{school-graph.png}
-		\caption{School}
-		\label{fig:school-graph}
-	\end{subfigure}
-	\hfill
-	\begin{subfigure}[t]{0.28\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{fb-graph.png}
-		\caption{Facebook egonet}
-		\label{fig:fb-graph}
-	\end{subfigure}
-	\begin{subfigure}[t]{0.11\linewidth}
-		\centering
-		\includegraphics[width=0.8\linewidth]{10-vertical-legend.png}
-		\caption{Block Legend}
-		\label{fig:10-legend}
-	\end{subfigure}
-	\caption{Networks laid out and coloured according to inferred block memberships $\hat{y}$ for a given experiment iteration.
-Visualisation performed using \textit{graph-tool} \cite{peixoto_graph-tool_2014}.}
-	\label{fig:graphs-all}
-\end{figure}
-\begin{table}[!h]
-	\centering
-	\caption{Experimental results averaged over $n=10$ iterations (mean $\pm$ std. dev.).}
-	\label{tab:results}
-	\resizebox{\textwidth}{!}{%
-		\begin{tabular}{c|ccc|c|cc|ccc}
-			Dataset  & $B$ & $D$ & $D'$ & $\bar{S}_e$ & $\bar{\Lcal}_0$ & $\bar{\Lcal}_1$ & $c^*$ & $\bar{\Lcal}_0'$ & $\bar{\Lcal}_1'$  \\ \hline
-			Polbooks & 3 & 3 & -- & $2.250 \pm 0.000$ & $0.563 \pm 0.042$ & $0.595 \pm 0.089$ & -- & -- & -- \\
-			School & 10 & 13 & 10 & $1.894 \pm 0.004$ & $0.787 \pm 0.127$ & $0.885 \pm 0.129$ & $1.198 \pm 0.249$ & $0.793 \pm 0.132$ & $0.853 \pm 0.132$ \\
-			FB egonet & 10  & 480 & 10 & $1.626 \pm 0.003$ & $1.326 \pm 0.043$ & $1.538 \pm 0.069$ & $0.94 \pm 0.019$ & $1.580 \pm 0.150$ & $1.605 \pm 0.106$
-		\end{tabular}
-	}
-\end{table}
-%
-We require metrics to assess performance. This can be split into two separate 
-components: the microcanonical SBM fit (concerned with the $b$-samples) and 
-the fit of the feature-to-block generator (concerned with the $\theta$-samples).
-
-Starting with the SBM, recall that the quantity $S(b)$ in~(\ref{eqn:dl-form}) 
-can be interpreted as an ideal ``description length'' of the partition 
-imposed by $b$. We define a simple
-metric $\bar{S}_e$ to gauge the fit of the SBM,
-as the description length per entity 
-(i.e., divided by the total number $N+E$ of 
-nodes plus edges),
-averaged over the $b$-samples:
-%
-\begin{equation}
-	\bar{S}_e \coloneqq \frac{1}{(N+E) |\Tcal_b|} \sum_{t\in \Tcal_b} S \left( b^{(t)} \right).
-	\label{eqn:mean-dl}
-\end{equation}
-%
-
-Next, to assess the performance of the feature-to-block predictor, 
-we partition the vertex set $[N]$ into training and test sets. We choose to 
-randomly partition the vertices on each experiment run,
-so that a constant fraction $f$ of the available vertices go to form 
-the training set $\Gcal_0$ and the remainder are held out to form the
-test set $\Gcal_1$.
-The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain.
-Because $|\Gcal_0| \neq |\Gcal_1|$ in general, we cannot use the un-normalised log-target $U$ from~(\ref{eqn:U-form}) for comparison,
-as the total cross-entropy loss scales with the size of each data set but 
-the prior term stays constant. We therefore use the average cross-entropy loss 
-over each set,
-%
-\begin{equation}
-	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star^{(t)},
-	\quad \textrm{where} \quad
-	\Lcal_\star^{(t)} \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)},
-	\label{eqn:cross-entropy-loss}
-\end{equation}
-%
-where $\star \in \{0, 1\}$ indicates whether the training or test
-set is being considered.
-
-Table~\ref{tab:results} summarises the results for each experiment. We also apply the 
-dimensionality reduction method 
-of Section~\ref{sec:dim-reduction}
-to the two higher dimensional datasets (the school and FB egonet). 
-For this we use equation~(\ref{eqn:c-star}) with $k=1$,
-in order to reduce the dimension from 
-$D$ to a desired $D'$. 
-We then retrain the feature-block predictor using only the retained 
-feature set $\Dcal'$, and report the log-loss over the training and 
-test sets for the reduced classifier -- 
-denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively. 
-These values are also given in Table~\ref{tab:results}.
-
-Based on these results, some remarks are in order.
-Firstly, the variance of the test loss $\bar{\Lcal}_1$ tends to be higher 
-than the training loss $\bar{\Lcal}_0$. This is expected,
-as the test set is smaller than the training set and hence 
-more susceptible to variability in its construction. 
-Secondly, it can be seen that the dimensionality reduction procedure 
-brings the training and test losses closer together. This implies that 
-the features we keep are indeed correlated with the underlying graphical 
-partition and that the approach generalises correctly.
-
-The average description length per entity,
-$\bar{S}_e$, of the graph, 
-has very small variance, suggesting that
-the detected communities can be found reliably (to within an arbitrary 
-relabelling of blocks). For reference, we plot an inferred partition for each 
-of the graphs in Figure~\ref{fig:graphs-all}.
-The polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do succeed at dividing the graph into densely connected clusters.
-
-Nevertheless, the cross-entropy loss over the whole training set may be too coarse a measure of model fit.
-It is often the case that we have good feature-based explanations for some but not all of the detected blocks.
-We wish to define a new measure of fit specific to each detected block.
-This requires defining the set of all vertices associated with block $j$ as,
-$
-	\Bcal_\star(j) \coloneqq \{i \in \Gcal_\star : \hat{b}_i = j\}
-$
-where
-$ 
-	\quad \hat{b}_i \coloneqq \argmax_j \hat{y}_{ij}.
-$
-Recall that $\hat{y}_{ij}$ is our estimate for the block membership posterior~(\ref{eqn:y-hat}) using only information from the adjacency matrix $A$.
-Again, $\star \in \{0, 1\}$ toggles between training and test sets.
-Now $\Bcal_\star(j)$ is the set of vertices that have maximum a posteriori probability of belonging to block $j$.
-We choose to resolve ties consistently by picking the lower index.
-We now define the accuracy for block $j$ as,
-%
-\begin{equation}
-	\eta_\star(j) \coloneqq \frac{1}{|\Bcal_\star (j)| \cdot 
-	|\Tcal_\theta| } 
-	\sum_{i \in \Bcal_\star (j)}  \sum_{t \in \Tcal_\theta}
-	\one \left\{\hat{b}_i = \argmax_j \phi_j \left( x_i; \theta^{(t)} \right) \right\}.
-	\label{eqn:accuracy}
-\end{equation}
-%
-This is effectively testing whether the feature-to-block and the graph-to=block predictions agree in their largest component.
-We call this metric, $\eta_\star(j)$, the {\em block-accuracy} for block $j$.
-It is clearly bounded $0 \leq \eta_\star(j) \leq 1$, with an accuracy of 1 meaning perfect agreement for the vertices in detected block $j$.
-
-For each of the experiments, we plot the collected $\theta$-samples for the features that survive the dimensionality reduction procedure.
-We also plot the per-block accuracy for the original-dimension classifiers.
-We discuss the results specific to each dataset in turn.
-
-\begin{figure}
-	\centering
-	\begin{minipage}{0.42\linewidth}
-		\centering
-		\begin{subfigure}{\linewidth}
-			\centering
-			\imagebox{0.8\linewidth}{\includegraphics[width=\linewidth]{polbooks-null-1}}
-			\caption{Political Books.}
-			\label{fig:polbooks-null}
-		\end{subfigure}
-
-		\begin{subfigure}{\linewidth}
-			\centering
-			\imagebox{0.85\linewidth}{\includegraphics[width=\linewidth]{school-null-1}}
-			\caption{Primary School.}
-			\label{fig:school-null}
-		\end{subfigure}
-
-		\begin{subfigure}{\linewidth}
-			\centering
-			\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{fb-null-1}}
-			\caption{Facebook Egonet.}
-			\label{fig:fb-null}
-		\end{subfigure}
-		\caption{$\theta$-samples. Dotted line $\pm c^*$.}
-	\end{minipage}
-	%
-	\begin{minipage}{0.42\linewidth}
-		\centering
-		\begin{subfigure}[t]{\linewidth}
-			\centering
-			\imagebox{0.8\linewidth}{\includegraphics[width=\linewidth]{polbooks-accuracy-2}}
-			\caption{Political Books.}
-			\label{fig:polbooks-accuracy}
-		\end{subfigure}
-
-		\begin{subfigure}{\linewidth}
-			\centering
-			\imagebox{0.85\linewidth}{\includegraphics[width=\linewidth]{school-accuracy-1}}
-			\caption{Primary School.}
-			\label{fig:school-accuracy}
-		\end{subfigure}
-
-		\begin{subfigure}{\linewidth}
-			\centering
-			\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{fb-accuracy-1}}
-			\caption{Facebook Egonet.}
-			\label{fig:fb-accuracy}
-		\end{subfigure}
-		\caption{Per-block accuracy $\eta(j)$.}
-	\end{minipage}
-\end{figure}
-
-\subsubsection{Political books.}
 
@@
 \subsubsection{Primary school dynamic contacts.}
+
+\cite{schools} ($N=238, E=5539, D=13$) -- network of face-to-face contacts amongst students and teachers at a primary school in Lyon, France.
+Vertex features include class membership (one of 10 values: 1A-5B), gender (male, female) and teacher status encoded as an 11th school-class.
+We choose to analyse just the second day of results.
 
@@
 
+\cite{fb-snap} ($N=747, E=30025, D=480$) -- an assortment of Facebook users' friends lists.
+Vertex features are fully anonymised and encode information about each user's education history, languages spoken, gender, home-town, birthday etc.
+We focus on the egonet with id 1912.
+
 We choose $B=10$ and $D'=10$ for this experiment. The selected features 

### 2021-11-06 publication/experiments.tex
Misc
Refactor experiments section

@@
 
-We apply the developed methods to a variety of datasets. These are chosen to span a range of node counts $N$, edge counts $E$ and feature space dimension $D$. We consider the following:
-
-\begin{itemize}
-	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=3$) -- network of Amazon political book sales, published close to the 2004 presidential election.
-Two books are connected if they were frequently co-purchased.
-Vertex features encode the political affiliation of the author (liberal, conservative, or neutral).
-
-	\item \textbf{Primary school dynamic contacts} \cite{schools} ($N=238, E=5539, D=13$) -- network of face-to-face contacts amongst students and teachers at a primary school in Lyon, France.
-Vertex features include class membership (one of 10 values: 1A-5B), gender (male, female) and teacher status encoded as an 11th school-class.
-We choose to analyse just the second day of results.
-
-	\item \textbf{Facebook egonet} \cite{fb-snap} ($N=747, E=30025, D=480$) -- an assortment of Facebook users' friends lists.
-Vertex features are fully anonymised and encode information about each user's education history, languages spoken, gender, home-town, birthday etc.
-We focus on the egonet with id 1912.
-
-\end{itemize}
-
-\begin{figure}[!h]
-	\centering
-	\begin{subfigure}[t]{0.28\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{polbooks-graph.png}
-		\caption{Polbooks}
-		\label{fig:polbooks-graph}
-	\end{subfigure}
-	\hfill
-	\begin{subfigure}[t]{0.28\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{school-graph.png}
-		\caption{School}
-		\label{fig:school-graph}
-	\end{subfigure}
-	\hfill
-	\begin{subfigure}[t]{0.28\linewidth}
-		\centering
-		\includegraphics[width=\linewidth]{fb-graph.png}
-		\caption{Facebook egonet}
-		\label{fig:fb-graph}
-	\end{subfigure}
-	\begin{subfigure}[t]{0.11\linewidth}
-		\centering
-		\includegraphics[width=0.8\linewidth]{10-vertical-legend.png}
-		\caption{Block Legend}
-		\label{fig:10-legend}
-	\end{subfigure}
-	\caption{Networks laid out and coloured according to inferred block memberships $\hat{y}$ for a given experiment iteration.
-Visualisation performed using \textit{graph-tool} \cite{peixoto_graph-tool_2014}.}
-	\label{fig:graphs-all}
-\end{figure}
-\begin{table}[!h]
-	\centering
-	\caption{Experimental results averaged over $n=10$ iterations (mean $\pm$ std. dev.).}
-	\label{tab:results}
-	\resizebox{\textwidth}{!}{%
-		\begin{tabular}{c|ccc|c|cc|ccc}
-			Dataset  & $B$ & $D$ & $D'$ & $\bar{S}_e$ & $\bar{\Lcal}_0$ & $\bar{\Lcal}_1$ & $c^*$ & $\bar{\Lcal}_0'$ & $\bar{\Lcal}_1'$  \\ \hline
-			Polbooks & 3 & 3 & -- & $2.250 \pm 0.000$ & $0.563 \pm 0.042$ & $0.595 \pm 0.089$ & -- & -- & -- \\
-			School & 10 & 13 & 10 & $1.894 \pm 0.004$ & $0.787 \pm 0.127$ & $0.885 \pm 0.129$ & $1.198 \pm 0.249$ & $0.793 \pm 0.132$ & $0.853 \pm 0.132$ \\
-			FB egonet & 10  & 480 & 10 & $1.626 \pm 0.003$ & $1.326 \pm 0.043$ & $1.538 \pm 0.069$ & $0.94 \pm 0.019$ & $1.580 \pm 0.150$ & $1.605 \pm 0.106$
-		\end{tabular}
-	}
-\end{table}
+We apply the developed methods to a variety of datasets. These are chosen to span a range of node counts $N$, edge counts $E$ and feature space dimension $D$.
 %
@@
 
-\begin{figure}
-	\centering
-	\begin{minipage}{0.42\linewidth}
-		\centering
-		\begin{subfigure}{\linewidth}
-			\centering
-			\imagebox{0.8\linewidth}{\includegraphics[width=\linewidth]{polbooks-null-1}}
-			\caption{Political Books.}
-			\label{fig:polbooks-null}
-		\end{subfigure}
-
-		\begin{subfigure}{\linewidth}
-			\centering
-			\imagebox{0.85\linewidth}{\includegraphics[width=\linewidth]{school-null-1}}
-			\caption{Primary School.}
-			\label{fig:school-null}
-		\end{subfigure}
-
-		\begin{subfigure}{\linewidth}
-			\centering
-			\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{fb-null-1}}
-			\caption{Facebook Egonet.}
-			\label{fig:fb-null}
-		\end{subfigure}
-		\caption{$\theta$-samples. Dotted line $\pm c^*$.}
-	\end{minipage}
-	%
-	\begin{minipage}{0.42\linewidth}
-		\centering
-		\begin{subfigure}[t]{\linewidth}
-			\centering
-			\imagebox{0.8\linewidth}{\includegraphics[width=\linewidth]{polbooks-accuracy-2}}
-			\caption{Political Books.}
-			\label{fig:polbooks-accuracy}
-		\end{subfigure}
-
-		\begin{subfigure}{\linewidth}
-			\centering
-			\imagebox{0.85\linewidth}{\includegraphics[width=\linewidth]{school-accuracy-1}}
-			\caption{Primary School.}
-			\label{fig:school-accuracy}
-		\end{subfigure}
-
-		\begin{subfigure}{\linewidth}
-			\centering
-			\imagebox{0.9\linewidth}{\includegraphics[width=\linewidth]{fb-accuracy-1}}
-			\caption{Facebook Egonet.}
-			\label{fig:fb-accuracy}
-		\end{subfigure}
-		\caption{Per-block accuracy $\eta(j)$.}
-	\end{minipage}
-\end{figure}
-
-\subsubsection{Political books.}
-
-We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral).
-From Figure~\ref{fig:polbooks-null} we see that all 3 blocks have a distinct political affiliation as their largest positive component.  
-Furthermore, in Table~\ref{tab:results} we see that the training and test losses 
-are very similar and both are low in magnitude. This is strong evidence 
-that political affiliation is a very appropriate explanatory 
-variable for the overall network structure.
-
-However, from Figure~\ref{fig:polbooks-accuracy} we see that block 1 has low accuracy. 
-This suggests that detected block 1 is not solely composed of ``neutral" books but also 
-contains some ``liberal" and ``conservative" authors. Examining 
-Figure~\ref{fig:polbooks-graph}, we see the majority of paths between blocks 2 and 3 go through block 1.
-Block 1 is in effect a bridge between the ``conservative'' and ``liberal'' blocks so it is unsurprising that some books from either side leak into block 1.
-
-\subsubsection{Primary school dynamic contacts.}
-
-We choose the number of communities $B=10$, in line with the total number of 
-school classes. From Figure~\ref{fig:school-null}, we see only the pupils' class memberships (1A-5B) have survived
-the dimensionality-reduction process for $D'=10$;
-gender and teacher/student status have been discarded,
-meaning these are poor predictors of overall macro-structure.
-
-The vast majority of blocks are composed of a single class. 
-However, some blocks have two comparably strong classes as their predictors. 
-For example, blocks 2 and 5 both contain classes 3A and 3B as their 2 best predictors. 
-This suggests that the social divide between classes is less pronounced 
-for pupils in year 3. Conversely, some classes are found to extend over two 
-detected blocks (class 2B spans blocks 8 and 9) but we do 
-not have a feature which explains the division.
-
-Figure~\ref{fig:school-accuracy} shows excellent accuracy for the majority of blocks.
-In fact the only blocks with low accuracy are those that have a school-class span two blocks such that we cannot reliably distinguish between the two.
-This is more pronounced when we apply hard classification rather than the soft cross-entropy loss.
-Indeed block 1 has low accuracy because class 1A has a higher weight component for block 3 than it does for block 1.
-As we use hard classification to compute accuracy, vertices belonging to class 1A are predicted to belong to block 3.
-A similar effect is seen in block 5 which often loses out to block 2.
-The same can be said of blocks 8 and 9.
-However, in this case the weights for class 2B are very similar -- thus explaining the roughly evenly split accuracy for blocks 8 and 9.
-
-\subsubsection{Facebook egonet.}
-
-We choose $B=10$ and $D'=10$ for this experiment. The selected features 
-(Figure~\ref{fig:fb-null}) are those that best explain the high-level 
-community structure. The majority of them are education related. 
-Nevertheless, for $D'=10$ we only have good explanations for some of the detected blocks; several blocks in 
-Figure~\ref{fig:fb-null} do not have high-magnitude components for $D'=10$.
-This is further emphasised by the disparate accuracies in Figure~\ref{fig:fb-accuracy}.
-Nevertheless, observe that the accuracy is only high for blocks that contain high magnitude weights.
-The only exception to this rule is block 9 which nonetheless may have high magnitude weights just below the cut-off $c^*$.
-
-When the feature dimension is very large, it becomes increasingly likely that 
-a particular feature may uniquely identify a small set of nodes. If these nodes 
-are all part of the same community, then the classifier may overfit for that 
-particular parameter. The regularisation term imposed by the prior goes some 
-way towards alleviating this problem. Nevertheless, we see in 
-Figure~\ref{fig:fb-null} that the feature \verb*|birthday-5| has a very 
-high weight as it relates to block 1 -- but it is highly unlikely
-that birthdays determine graphical structure.
+\input{experiments-detailed}
+\input{experiments-figures}

### 2021-11-06 publication/experiments-detailed.tex
Refactor

@@
-\subsubsection{Political books.}
+Table~\ref{tab:results} summarises the results for each experiment. We also apply the 
+dimensionality reduction method 
+of Section~\ref{sec:dim-reduction}
+to the two higher dimensional datasets (the school and FB egonet). 
+For this we use equation~(\ref{eqn:c-star}) with $k=1$,
+in order to reduce the dimension from 
+$D$ to a desired $D'$. 
+We then retrain the feature-block predictor using only the retained 
+feature set $\Dcal'$, and report the log-loss over the training and 
+test sets for the reduced classifier -- 
+denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively. 
+These values are also given in Table~\ref{tab:results}.
+
+Based on these results, some remarks are in order.
+Firstly, the variance of the test loss $\bar{\Lcal}_1$ tends to be higher 
+than the training loss $\bar{\Lcal}_0$. This is expected,
+as the test set is smaller than the training set and hence 
+more susceptible to variability in its construction. 
+Secondly, it can be seen that the dimensionality reduction procedure 
+brings the training and test losses closer together. This implies that 
+the features we keep are indeed correlated with the underlying graphical 
+partition and that the approach generalises correctly.
+
+The average description length per entity,
+$\bar{S}_e$, of the graph, 
+has very small variance, suggesting that
+the detected communities can be found reliably (to within an arbitrary 
+relabelling of blocks). For reference, we plot an inferred partition for each 
+of the graphs in Figure~\ref{fig:graphs-all}.
+The polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do succeed at dividing the graph into densely connected clusters.
+
+\subsubsection{Political books}
 
@@
 
-\subsubsection{Primary school dynamic contacts.}
+\subsubsection{Primary school dynamic contacts}
 
@@
 
-\subsubsection{Facebook egonet.}
+\subsubsection{Facebook egonet}
 

### 2021-11-06 publication/experiments.tex
Refactor

@@
-\section{Experimental results}
+\section{Experiments}
 \label{sec:experiments}
 
-We apply the developed methods to a variety of datasets. These are chosen to span a range of node counts $N$, edge counts $E$ and feature space dimension $D$.
-%
-We require metrics to assess performance. This can be split into two separate 
+We apply the developed methods to a vareity of datasets chosen to span a range of edge counts $E$ and feature-space dimension $D$. 
+However, we first require metrics to assess model performance. This can be split into two separate 
 components: the microcanonical SBM fit (concerned with the $b$-samples) and 
@@
 can be interpreted as an ideal ``description length'' of the partition 
-imposed by $b$. We define a simple
-metric $\bar{S}_e$ to gauge the fit of the SBM,
-as the description length per entity 
-(i.e., divided by the total number $N+E$ of 
-nodes plus edges),
-averaged over the $b$-samples:
+imposed by $b$. We define the average
+description length per entity (nodes and edges) $\bar{S}_e$ to gauge the SBM fit:
 %
@@
 \end{equation}
-%
 
 Next, to assess the performance of the feature-to-block predictor, 
-we partition the vertex set $[N]$ into training and test sets. We choose to 
-randomly partition the vertices on each experiment run,
-so that a constant fraction $f$ of the available vertices go to form 
+we partition the vertex set $[N]$ into training and test sets. On each experiment run,
+a constant fraction $f$ of the available vertices go to form 
 the training set $\Gcal_0$ and the remainder are held out to form the
-test set $\Gcal_1$.
+test set $\Gcal_1$. These sets are chosen randomly.
 The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain.
-Because $|\Gcal_0| \neq |\Gcal_1|$ in general, we cannot use the un-normalised log-target $U$ from~(\ref{eqn:U-form}) for comparison,
+Because $|\Gcal_0| \neq |\Gcal_1|$ in general, we cannot use the un-normalised log-target $U$ from~(\ref{eqn:U-form}) for benchmarking,
 as the total cross-entropy loss scales with the size of each data set but 
@@
 set is being considered.
-
-Table~\ref{tab:results} summarises the results for each experiment. We also apply the 
-dimensionality reduction method 
-of Section~\ref{sec:dim-reduction}
-to the two higher dimensional datasets (the school and FB egonet). 
-For this we use equation~(\ref{eqn:c-star}) with $k=1$,
-in order to reduce the dimension from 
-$D$ to a desired $D'$. 
-We then retrain the feature-block predictor using only the retained 
-feature set $\Dcal'$, and report the log-loss over the training and 
-test sets for the reduced classifier -- 
-denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively. 
-These values are also given in Table~\ref{tab:results}.
-
-Based on these results, some remarks are in order.
-Firstly, the variance of the test loss $\bar{\Lcal}_1$ tends to be higher 
-than the training loss $\bar{\Lcal}_0$. This is expected,
-as the test set is smaller than the training set and hence 
-more susceptible to variability in its construction. 
-Secondly, it can be seen that the dimensionality reduction procedure 
-brings the training and test losses closer together. This implies that 
-the features we keep are indeed correlated with the underlying graphical 
-partition and that the approach generalises correctly.
-
-The average description length per entity,
-$\bar{S}_e$, of the graph, 
-has very small variance, suggesting that
-the detected communities can be found reliably (to within an arbitrary 
-relabelling of blocks). For reference, we plot an inferred partition for each 
-of the graphs in Figure~\ref{fig:graphs-all}.
-The polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do succeed at dividing the graph into densely connected clusters.
 

### 2021-11-06 publication/derivations.tex
13 pages

@@
-\section{Derivations}
+\section{Appendix: Derivations}
 
@@
 	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
-	&=  - \left( \sum_{i \in [N]} x_i \left( y_{ik} - a_{ik} \sum_{j \in [B]} y_{ij} \right)
-	- \frac{w_k}{\sigma_\theta^2} \right) \nonumber \\
 	&= - \left( \sum_{i \in [N]} \Big\{ x_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right).

### 2021-11-06 publication/experiments-detailed.tex
1-page experiment metrics defn

@@
 
-\cite{polbooks} ($N=105, E=441, D=3$) -- network of Amazon political book sales, published close to the 2004 presidential election.
-Two books are connected if they were frequently co-purchased.
-Vertex features encode the political affiliation of the author (liberal, conservative, or neutral).
-
 We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral).
@@
 \subsubsection{Primary school dynamic contacts}
-
-\cite{schools} ($N=238, E=5539, D=13$) -- network of face-to-face contacts amongst students and teachers at a primary school in Lyon, France.
-Vertex features include class membership (one of 10 values: 1A-5B), gender (male, female) and teacher status encoded as an 11th school-class.
-We choose to analyse just the second day of results.
 
@@
 
-\cite{fb-snap} ($N=747, E=30025, D=480$) -- an assortment of Facebook users' friends lists.
-Vertex features are fully anonymised and encode information about each user's education history, languages spoken, gender, home-town, birthday etc.
-We focus on the egonet with id 1912.
-
 We choose $B=10$ and $D'=10$ for this experiment. The selected features 

### 2021-11-06 publication/experiments.tex
13 pages
1-page experiment metrics defn

@@
 
-We apply the developed methods to a vareity of datasets chosen to span a range of edge counts $E$ and feature-space dimension $D$. 
-However, we first require metrics to assess model performance. This can be split into two separate 
-components: the microcanonical SBM fit (concerned with the $b$-samples) and 
-the fit of the feature-to-block generator (concerned with the $\theta$-samples).
+We apply the outlined methods to a variety of datasets:
 
-Starting with the SBM, recall that the quantity $S(b)$ in~(\ref{eqn:dl-form}) 
-can be interpreted as an ideal ``description length'' of the partition 
-imposed by $b$. We define the average
-description length per entity (nodes and edges) $\bar{S}_e$ to gauge the SBM fit:
+\begin{itemize}
+	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=3$) -- network of Amazon political book sales, published close to the 2004 presidential election.
+Two books are connected if they were frequently co-purchased.
+Vertex features encode the political affiliation of the author (liberal, conservative, or neutral).
+	\item \textbf{Primary school dynamic contacts} \cite{schools} ($N=238, E=5539, D=13$) -- network of face-to-face contacts amongst students and teachers at a primary school in Lyon, France.
+Vertex features include class membership (one of 10 values: 1A-5B), gender (male, female) and teacher status encoded as an 11th school-class.
+We choose to analyse just the second day of results.
+	\item \textbf{Facebook egonet} \cite{fb-snap} ($N=747, E=30025, D=480$) -- an assortment of Facebook users' friends lists.
+Vertex features are fully anonymised and encode information about each user's education history, languages spoken, gender, home-town, birthday etc.
+We focus on the egonet with id 1912.
+\end{itemize}
+%
+We first require metrics to assess model performance. We define the average
+description length per entity (nodes and edges) $\bar{S}_e$ as a suitable metric to gauge the SBM fit:
 %
@@
 Next, to assess the performance of the feature-to-block predictor, 
-we partition the vertex set $[N]$ into training and test sets. On each experiment run,
-a constant fraction $f$ of the available vertices go to form 
-the training set $\Gcal_0$ and the remainder are held out to form the
-test set $\Gcal_1$. These sets are chosen randomly.
-The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain.
-Because $|\Gcal_0| \neq |\Gcal_1|$ in general, we cannot use the un-normalised log-target $U$ from~(\ref{eqn:U-form}) for benchmarking,
-as the total cross-entropy loss scales with the size of each data set but 
-the prior term stays constant. We therefore use the average cross-entropy loss 
-over each set,
+we randomly partition the vertex set $[N]$ into training and test sets: $\Gcal_0$ and $\Gcal_1$.
+The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain. As $|\Gcal_0| \neq |\Gcal_1|$ in general, we use the average cross-entropy loss 
+over each set to gauge fit,
 %
@@
 %
-where $\star \in \{0, 1\}$ indicates whether the training or test
-set is being considered.
-
-Nevertheless, the cross-entropy loss over the whole training set may be too coarse a measure of model fit.
-It is often the case that we have good feature-based explanations for some but not all of the detected blocks.
-We wish to define a new measure of fit specific to each detected block.
-This requires defining the set of all vertices associated with block $j$ as,
+where $\star \in \{0, 1\}$ toggles between the training and test sets.
+%
+Nevertheless, the cross-entropy loss is a coarse measure of fit. We wish to define a new measure of fit specific to each detected block. Let us define
 $
@@
 $ 
-	\quad \hat{b}_i \coloneqq \argmax_j \hat{y}_{ij}.
+	\hat{b}_i \coloneqq \argmax_j \hat{y}_{ij}.
 $
-Recall that $\hat{y}_{ij}$ is our estimate for the block membership posterior~(\ref{eqn:y-hat}) using only information from the adjacency matrix $A$.
-Again, $\star \in \{0, 1\}$ toggles between training and test sets.
-Now $\Bcal_\star(j)$ is the set of vertices that have maximum a posteriori probability of belonging to block $j$.
-We choose to resolve ties consistently by picking the lower index.
-We now define the accuracy for block $j$ as,
+Now $\Bcal_\star(j)$ is the set of vertices that have maximum a posteriori probability of belonging to block $j$. We now define the accuracy for block $j$ as,
 %
@@
 \end{equation}
-%
-This is effectively testing whether the feature-to-block and the graph-to=block predictions agree in their largest component.
+
+This effectively tests whether the feature-to-block and the graph-to-block predictions agree in their largest component.
 We call this metric, $\eta_\star(j)$, the {\em block-accuracy} for block $j$.
@@
 
-For each of the experiments, we plot the collected $\theta$-samples for the features that survive the dimensionality reduction procedure.
-We also plot the per-block accuracy for the original-dimension classifiers.
-We discuss the results specific to each dataset in turn.
-
 \input{experiments-detailed}

### 2021-11-06 publication/experiments-detailed.tex
12-pages

@@
-Table~\ref{tab:results} summarises the results for each experiment. We also apply the 
+For the higher-dimensional datasets, we also apply the 
 dimensionality reduction method 
-of Section~\ref{sec:dim-reduction}
-to the two higher dimensional datasets (the school and FB egonet). 
-For this we use equation~(\ref{eqn:c-star}) with $k=1$,
-in order to reduce the dimension from 
-$D$ to a desired $D'$. 
+of Section~\ref{sec:dim-reduction}.  
 We then retrain the feature-block predictor using only the retained 
@@
 denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively.
-These values are also given in Table~\ref{tab:results}.
 
-Based on these results, some remarks are in order.
-Firstly, the variance of the test loss $\bar{\Lcal}_1$ tends to be higher 
-than the training loss $\bar{\Lcal}_0$. This is expected,
-as the test set is smaller than the training set and hence 
-more susceptible to variability in its construction. 
-Secondly, it can be seen that the dimensionality reduction procedure 
+Table~\ref{tab:results} summarises the results for each experiment.
+We see that the dimensionality reduction procedure 
 brings the training and test losses closer together. This implies that 
@@
 partition and that the approach generalises correctly.
-
+%
 The average description length per entity,
@@
 the detected communities can be found reliably (to within an arbitrary 
-relabelling of blocks). For reference, we plot an inferred partition for each 
-of the graphs in Figure~\ref{fig:graphs-all}.
-The polbooks graph yields the cleanest separation between blocks but nonetheless the inferred partitions for the other datasets do succeed at dividing the graph into densely connected clusters.
+relabelling of blocks).
 
-\subsubsection{Political books}
+\paragraph{\textbf{Political books.}}
 
@@
 From Figure~\ref{fig:polbooks-null} we see that all 3 blocks have a distinct political affiliation as their largest positive component.  
-Furthermore, in Table~\ref{tab:results} we see that the training and test losses 
+Furthermore, the training and test losses from Table~\ref{tab:results}  
 are very similar and both are low in magnitude. This is strong evidence 
@@
 variable for the overall network structure.
-
+%
 However, from Figure~\ref{fig:polbooks-accuracy} we see that block 1 has low accuracy. 
@@
 
-\subsubsection{Primary school dynamic contacts}
+\paragraph{\textbf{Primary school.}}
 
 We choose the number of communities $B=10$, in line with the total number of 
-school classes. From Figure~\ref{fig:school-null}, we see only the pupils' class memberships (1A-5B) have survived
-the dimensionality-reduction process for $D'=10$;
+school classes. Only the pupils' class memberships (1A-5B) survive
+the dimensionality-reduction process (Figur~\ref{fig:school-null}e);
 gender and teacher/student status have been discarded,
 meaning these are poor predictors of overall macro-structure.
-
+%
 The vast majority of blocks are composed of a single class. 
-However, some blocks have two comparably strong classes as their predictors. 
-For example, blocks 2 and 5 both contain classes 3A and 3B as their 2 best predictors. 
-This suggests that the social divide between classes is less pronounced 
-for pupils in year 3. Conversely, some classes are found to extend over two 
+However, some blocks have two comparably strong classes as their predictors (e.g. blocks 2 and 5). 
+Conversely, some classes are found to extend over two 
 detected blocks (class 2B spans blocks 8 and 9) but we do 
 not have a feature which explains the division.
-
+%
 Figure~\ref{fig:school-accuracy} shows excellent accuracy for the majority of blocks.
@@
 This is more pronounced when we apply hard classification rather than the soft cross-entropy loss.
-Indeed block 1 has low accuracy because class 1A has a higher weight component for block 3 than it does for block 1.
-As we use hard classification to compute accuracy, vertices belonging to class 1A are predicted to belong to block 3.
-A similar effect is seen in block 5 which often loses out to block 2.
-The same can be said of blocks 8 and 9.
-However, in this case the weights for class 2B are very similar -- thus explaining the roughly evenly split accuracy for blocks 8 and 9.
+Perhapsthere are unobserved features which explain this divide.
 
-\subsubsection{Facebook egonet}
+\paragraph{\textbf{Facebook egonet.}}
 
@@
 Nevertheless, for $D'=10$ we only have good explanations for some of the detected blocks; several blocks in 
-Figure~\ref{fig:fb-null} do not have high-magnitude components for $D'=10$.
-This is further emphasised by the disparate accuracies in Figure~\ref{fig:fb-accuracy}.
-Nevertheless, observe that the accuracy is only high for blocks that contain high magnitude weights.
-The only exception to this rule is block 9 which nonetheless may have high magnitude weights just below the cut-off $c^*$.
-
-When the feature dimension is very large, it becomes increasingly likely that 
-a particular feature may uniquely identify a small set of nodes. If these nodes 
-are all part of the same community, then the classifier may overfit for that 
-particular parameter. The regularisation term imposed by the prior goes some 
-way towards alleviating this problem. Nevertheless, we see in 
-Figure~\ref{fig:fb-null} that the feature \verb*|birthday-5| has a very 
-high weight as it relates to block 1 -- but it is highly unlikely
-that birthdays determine graphical structure.
+Figure~\ref{fig:fb-null} do not have high-magnitude components for $D'=10$. This is further emphasised by the disparate accuracies in Figure~\ref{fig:fb-accuracy}.

### 2021-11-06 publication/experiments.tex
12-pages

@@
 
-We apply the outlined methods to a variety of datasets:
+We apply the outlined methods to a variety of labelled networks. For reference, the inferred partitions for all of these are potted on Figure~\ref{fig:graphs-all}.
 
@@
 we randomly partition the vertex set $[N]$ into training and test sets: $\Gcal_0$ and $\Gcal_1$.
+\footnote{We choose $f=0.7$ as the fraction of vertices that go to form our test set. Please refer to \url{https://github.com/LozzaTray/Jormungandr-code} for full source code.}
 The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain. As $|\Gcal_0| \neq |\Gcal_1|$ in general, we use the average cross-entropy loss 
@@
 $
-Now $\Bcal_\star(j)$ is the set of vertices that have maximum a posteriori probability of belonging to block $j$. We now define the accuracy for block $j$ as,
+Now $\Bcal_\star(j)$ is the set of vertices that have maximum a posteriori probability of belonging to block $j$.
+We now define the {\em block-accuracy} for block $j$ as in \ref{eqn:accuracy}.
+This effectively tests whether the feature-to-block and the graph-to-block predictions agree in their largest component.
 %
@@
 
-This effectively tests whether the feature-to-block and the graph-to-block predictions agree in their largest component.
-We call this metric, $\eta_\star(j)$, the {\em block-accuracy} for block $j$.
-It is clearly bounded $0 \leq \eta_\star(j) \leq 1$, with an accuracy of 1 meaning perfect agreement for the vertices in detected block $j$.
-
+\input{experiments-figures}
 \input{experiments-detailed}
-\input{experiments-figures}

### 2021-11-07 publication/derivations.tex
neaten up appendix
Add impl details

@@
-\section{Appendix: Derivations}
+\section{Appendix}
 
@@
 
-\subsection{Derivation of {\boldmath $\;U(\theta)$}}
+\subsection{Derivation of {\boldmath $U(\theta)$} and {\boldmath $\nabla U(\theta)$}}
 \label{appdx:form-U}
@@
 %
-\begin{equation}
+\begin{equation*}
 	\log p(b | X, \theta) = \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
@@
 	\label{eqn:U-constituent-terms}
-\end{equation}
+\end{equation*}
 %
@@
 
-\subsection{Derivation of {\boldmath $\;\nabla U(\theta)$}}
-\label{appdx:gradu}
-Here we show how the gradient 
-$\nabla U(\theta)$ can be computed explicitly.
 As
@@
 for all $i$.
+
+\subsection{Implementation details}
+\label{appdx:hyperparams}
+
+All data analysis and visualisation was implemented in Python. Full source code is available at:
+\url{https://github.com/LozzaTray/Jormungandr-code}. Table~\ref{tab:hyperparams} contains the hyper-parameters for each experiment run. The thinned samples are generated as,
+\begin{equation}
+	\Tcal_\star = \{T_\star \kappa_\star + i \lambda_\star :  
+	0 \leq i \leq \lfloor T_\star(1 - \kappa_\star) / \lambda_\star \rfloor \}.
+\end{equation}
+
+\begin{table}[!h]
+	\centering
+	\caption{Hyper-parameter values for each experiment}
+	\label{tab:hyperparams}
+	\resizebox{\textwidth}{!}{%
+		\begin{tabular}{c|ccc|ccc|cccc|cc|cccc}
+			Dataset & 
+			$B$ & $f$ & $\sigma_\theta$ & 
+			$T_b$ & $\kappa_b$ & $\lambda_b$ & 
+			$T_\theta$ & $\kappa_\theta$ & $\lambda_\theta$ & $s$ &
+			$k$ & $D'$ &
+			$T_\theta'$ & $\kappa_\theta'$ & $\lambda_\theta'$ & $s'$
+			\\ \hline
+			Polbooks &
+			3 & 0.7 & 1 &
+			1,000 & 0.2 & 5 &
+			10,000 & 0.4 & 10 & 0.05 &
+			-- & -- & 
+			-- & -- & -- & -- \\
+			School &
+			10 & 0.7 & 1 &
+			1,000 & 0.2 & 5 &
+			10,000 & 0.4 & 10 & 0.2 &
+			1 & 10 & 
+			10,000 & 0.4 & 10 & 0.2 \\
+			FB Egonet &
+			10 & 0.7 & 1 &
+			1,000 & 0.2 & 5 &
+			10,000 & 0.4 & 10 & 0.017 &
+			1 & 10 & 
+			10,000 & 0.4 & 10 & 0.5 \\
+		\end{tabular}
+	}
+\end{table}

### 2021-11-07 publication/experiments.tex
Add impl details

@@
 Next, to assess the performance of the feature-to-block predictor, 
-we randomly partition the vertex set $[N]$ into training and test sets: $\Gcal_0$ and $\Gcal_1$.
-\footnote{We choose $f=0.7$ as the fraction of vertices that go to form our test set. Please refer to \url{https://github.com/LozzaTray/Jormungandr-code} for full source code.}
+we randomly partition the vertex set $[N]$ such that a constant fraction $f$ forms our training set $\Gcal_0$ and the remainder form our test set $\Gcal_1$.
 The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain. As $|\Gcal_0| \neq |\Gcal_1|$ in general, we use the average cross-entropy loss 

### 2021-11-07 publication/abstract.tex
Flesh out inference

@@
 In this work, we introduce a new generative model, the feature-first block model (FFBM),
-which facilitates the use of richer queries on labelled network structure.
-We develop a Bayesian framework for inference with this model,
-and we present a method to efficiently sample from the posterior
-distribution of the FFBM parameters.
+to facilitate the use of richer queries on labelled network structure.
+We develop a Bayesian framework and devise a two-level Markov-Chain Monte-Carlo approach to efficiently sample the FFBM parameters.
+This allows us to infer if and how the observed vertex-features affect macro-structure.
 

### 2021-11-07 publication/derivations.tex
Flesh out inference

@@
 All data analysis and visualisation was implemented in Python. Full source code is available at:
-\url{https://github.com/LozzaTray/Jormungandr-code}. Table~\ref{tab:hyperparams} contains the hyper-parameters for each experiment run. The thinned samples are generated as,
-\begin{equation}
+\url{https://github.com/LozzaTray/Jormungandr-code}. Table~\ref{tab:hyperparams} contains the hyper-parameters for each experiment run. The set of retained samples are generated as,
+$
 	\Tcal_\star = \{T_\star \kappa_\star + i \lambda_\star :  
-	0 \leq i \leq \lfloor T_\star(1 - \kappa_\star) / \lambda_\star \rfloor \}.
-\end{equation}
+	0 \leq i \leq \lfloor T_\star(1 - \kappa_\star) / \lambda_\star \rfloor \},
+$
+with $\kappa_star$ controlling burn-in and $\lambda_star$ controlling thinning.
 
@@
 	\resizebox{\textwidth}{!}{%
-		\begin{tabular}{c|ccc|ccc|cccc|cc|cccc}
+		\begin{tabular}{c|ccc|ccc|ccc|cc|ccc}
 			Dataset & 
@@
 			$T_b$ & $\kappa_b$ & $\lambda_b$ & 
-			$T_\theta$ & $\kappa_\theta$ & $\lambda_\theta$ & $s$ &
+			$T_\theta$ & $\kappa_\theta$ & $\lambda_\theta$ & %$s$ &
 			$k$ & $D'$ &
-			$T_\theta'$ & $\kappa_\theta'$ & $\lambda_\theta'$ & $s'$
+			$T_\theta'$ & $\kappa_\theta'$ & $\lambda_\theta'$ %& $s'$
 			\\ \hline
@@
 			1,000 & 0.2 & 5 &
-			10,000 & 0.4 & 10 & 0.05 &
+			10,000 & 0.4 & 10 & %0.05 &
 			-- & -- & 
-			-- & -- & -- & -- \\
+			-- & -- & -- %& -- 
+			\\
 			School &
@@
 			1,000 & 0.2 & 5 &
-			10,000 & 0.4 & 10 & 0.2 &
+			10,000 & 0.4 & 10 & %0.2 &
 			1 & 10 & 
-			10,000 & 0.4 & 10 & 0.2 \\
+			10,000 & 0.4 & 10 %& 0.2 
+			\\
 			FB Egonet &
@@
 			1,000 & 0.2 & 5 &
-			10,000 & 0.4 & 10 & 0.017 &
+			10,000 & 0.4 & 10 & %0.017 &
 			1 & 10 & 
-			10,000 & 0.4 & 10 & 0.5 \\
+			10,000 & 0.4 & 10 %& 0.5 
+			\\
 		\end{tabular}

### 2021-11-07 publication/ffbm.tex
Flesh out inference

@@
 of weight vectors $\theta=\{w_k\}_{k=1}^B$, where each
-$w_k$ has dimension $D$. Specifically, 
+$w_k$ has dimension $D$. We will later find it convenient
+to write $\theta$ as a $B \times D$ matrix of weights $W$. Specifically, 
 the distribution of $b$ given $X$ and $\theta$ is,
@@
 \end{equation}
-%
-Note that $\phi_{b_i}$ has the form of a softmax activation function;
-we deliberately exclude a bias term to ensure that we only leverage feature information.
-We will later also find it convenient
-to write the parameters $\theta$ as a $B \times D$ matrix of weights $W$.
+
+Note that $\phi_{b_i}$ has the form of a softmax activation function.
 More complex models based on different choices for the distributions
@@
 \end{equation}
-%
+
 This choice of prior gives a very
@@
 \end{equation}
-%
+
 The proof is given in Appendix~\ref{appdx:b|x}.
@@
 In our case, conditioning on $X$ is also necessary, leading to,
-%
-\begin{equation}
+$
 	p(b, \psi | X) = p(b | X) p(\psi | b, X) = p(b | X) p(\psi | b),
-	\label{eqn:joint-pseudo-prior}
-\end{equation}
-%
+$
 where we used the fact $\psi$ and $X$ are conditionally 

### 2021-11-07 publication/inference.tex
Add impl details
Flesh out inference

@@
 Having completed the definition of the FFBM, we wish to leverage it 
-to perform inference. Given a vertex-labelled graph $(A, X)$,
-the goal is to sample from the posterior distribution
-of $\theta$:
+to perform inference. Specifically, given a labelled network $(A, X)$, we wish to infer if and how the observed features $(X)$ impact graphical structure $(A)$. Formally,
+this means characterising the posterior distribution:
+$
+p(\theta|A, X) \propto p(\theta) \cdot p(A | X, \theta).
+$
+Although the prior is easily computable, computing the likelihood would require a summation over all latent block-states, $p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$, and is therefore intractable.
+This approach is made doubly intractable as we would also need to compute the normalising constant $p(A|X) = \int p(\theta) \cdot p(A | X, \theta) d\theta$.
+Clearly we cannot directly compute the form of the posterior, so instead we aim to draw samples from it,
 %
@@
 	\label{eqn:theta-target}
-	\theta \sim p(\theta | A, X).
+	\theta^{(t)} \sim p(\theta | A, X).
 \end{equation}
-%
-We propose an iterative approach to obtain samples
-$\theta^{(t)}$ from $p(\theta|A,X)$. We first draw a sample $b^{(t)}$ 
+
+We propose an iterative approach to obtain these samples
+$\{\theta^{(t)}\}_t$. We first draw a sample $b^{(t)}$ 
 from the block membership posterior,
@@
 \end{equation}
-%
-This can be achieved through a two-tiered
-Markov chain Monte-Carlo approach via the Metropolis-Hastings (MH) 
+
+This can be implemented through a two-level
+Markov-Chain via the Metropolis-Hastings (MH) 
 algorithm \cite{hastings-alg}.
+The splitting of the Markov-Chain into two levels allows us to side-step the summation over
+all latent $b \in [B]^N$ required to directly compute the likelihood, $p(A| X, \theta)$.
 The resulting $\theta^{(t)}$ samples are unbiased in that the expectation of 
@@
 \end{equation}
-%
+
 This is an example of a pseudo-marginal approach. Indeed, Andrieu and Roberts \cite{pseudo-marginal} show that~(\ref{eqn:theta-unbiased})
-% the unbiased result in equation 
 is sufficient to prove that, for large enough $t$, $\theta^{(t)} \sim p(\theta | A, X)$, as required.
@@
 	\end{tikzpicture}
-	\caption{Sampling sequence.}
+	\caption{$\theta$-sample generation.}
 	\label{fig:samp-sequence}
@@
 
-We have split the Markov chain into two levels as the summation over all latent $b \in [B]^N$ required to directly compute the likelihood, $p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$, is intractable.
 Figure \ref{fig:samp-sequence} shows an overview of the proposed method, with $q$ and $\alpha$ denoting the MH proposal distribution and acceptance probability respectively.
-
 Note the importance of the simplification in~(\ref{eqn:b-pseudo-prior}). 
@@
 Without the injected noise term $\xi$, MALA is equivalent to gradient descent. We require $\xi$ to fully explore the parameter space. 
-The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:gradu}).
+The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:form-U}).
 

### 2021-11-07 publication/conclusion.tex
IDA-draft

@@
 
-The Feature-First Block Model (FFBM) introduced 
-in this paper is a new generative model for labelled networks well-suited for describing how features
-affect structure. The idea is to divide the graph into its most natural partition and determine whether 
-the vertex features can accurately explain this partition. 
+The proposed Feature-First Block Model (FFBM)
+is a new generative model for labelled networks well-suited for describing how features
+affect structure.
 It is easy to find vertex features that are in some way 
@@
 the parameters of the FFBM.
+The idea is to divide the graph into its most natural partition and determine whether 
+the vertex features can accurately explain this partition. 
 This is shown to be effective at extracting and describing 

### 2021-11-07 publication/derivations.tex
IDA-draft

@@
 	+ \frac{1}{2\sigma_\theta^2} \|\theta\|^2.
+	\label{eqn:U-form-appdx}
 \end{equation}
-
-As
-$\theta = \{ w_k \}_{k=1}^{B}$,
-to compute the gradient
-$\nabla U(\theta)$ we need to compute
+%
+Now to find $\nabla U(\theta)$, we need to compute
 each of its components,
-$\nicefrac{\partial U}{\partial w_k}$,
-$1\leq k\leq B$.
+$\nicefrac{\partial U}{\partial w_k}$ for $k \in [B]$.
 To that end, we first compute,
@@
 %
-Recall the expression for $U(\theta)$ in~(\ref{eqn:U-form}). Using~(\ref{eq:dadw}) and~(\ref{eq:dtsdw}), we obtain,
+Combining the expression for $U(\theta)$ in~(\ref{eqn:U-form-appdx}) and the expressions of~(\ref{eq:dadw}) and~(\ref{eq:dtsdw}), we obtain,
 \begin{align}
@@
 
-All data analysis and visualisation was implemented in Python. Full source code is available at:
+Full source code is available at:
 \url{https://github.com/LozzaTray/Jormungandr-code}. Table~\ref{tab:hyperparams} contains the hyper-parameters for each experiment run. The set of retained samples are generated as,
@@
 	\Tcal_\star = \{T_\star \kappa_\star + i \lambda_\star :  
-	0 \leq i \leq \lfloor T_\star(1 - \kappa_\star) / \lambda_\star \rfloor \},
+	0 \leq i \leq \lfloor T_\star(1 - \kappa_\star) / \lambda_\star \rfloor \}.
 $
-with $\kappa_star$ controlling burn-in and $\lambda_star$ controlling thinning.
-
+%
 \begin{table}[!h]
@@
 	\label{tab:hyperparams}
-	\resizebox{\textwidth}{!}{%
 	\begin{tabular}{c|ccc|ccc|ccc|cc|ccc}
@@
 	\end{tabular}
-	}
 \end{table}

### 2021-11-07 publication/experiments-detailed.tex
IDA-draft

@@
-For the higher-dimensional datasets, we also apply the 
-dimensionality reduction method 
-of Section~\ref{sec:dim-reduction}.  
-We then retrain the feature-block predictor using only the retained 
-feature set $\Dcal'$, and report the log-loss over the training and 
-test sets for the reduced classifier -- 
-denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively.
-
 Table~\ref{tab:results} summarises the results for each experiment. 
@@
 partition and that the approach generalises correctly.
+The test loss variance is higher than the training loss variance as the test set is smaller and so more susceptible to variability in its construction.
 %
@@
 
-We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation (conservative, liberal or neutral).
+We choose to partition the network into $B=3$ communities as we only have this many distinct values for political affiliation.
 From Figure~\ref{fig:polbooks-null} we see that all 3 blocks have a distinct political affiliation as their largest positive component.  
@@
 Figure~\ref{fig:polbooks-graph}, we see the majority of paths between blocks 2 and 3 go through block 1.
-Block 1 is in effect a bridge between the ``conservative'' and ``liberal'' blocks so it is unsurprising that some books from either side leak into block 1.
+Block 1 is in effect a bridge between the ``conservative'' and ``liberal'' blocks so it is unsurprising that some of these leak into block 1.
 
@@
 school classes. Only the pupils' class memberships (1A-5B) survive
-the dimensionality-reduction process (Figur~\ref{fig:school-null}e);
+the dimensionality-reduction process (Figure~\ref{fig:school-null});
 gender and teacher/student status have been discarded,
@@
 This is more pronounced when we apply hard classification rather than the soft cross-entropy loss.
-Perhapsthere are unobserved features which explain this divide.
+Perhaps there are unobserved features which explain this divide.
 
@@
 
-We choose $B=10$ and $D'=10$ for this experiment. The selected features 
+The retained features 
 (Figure~\ref{fig:fb-null}) are those that best explain the high-level 
-community structure. The majority of them are education related. 
+community structure. The majority of these are education related. 
 Nevertheless, for $D'=10$ we only have good explanations for some of the detected blocks; several blocks in 
-Figure~\ref{fig:fb-null} do not have high-magnitude components for $D'=10$. This is further emphasised by the disparate accuracies in Figure~\ref{fig:fb-accuracy}.
+Figure~\ref{fig:fb-null} do not have high-magnitude components. This is further emphasised by the disparate accuracies in Figure~\ref{fig:fb-accuracy}.
+%
+For a high-dimensional feature-space, it is likely that a particular
+feature may uniquely identify a small set of vertices; if these are all in the same block, then the classifier may overfit despite the penalty imposed by the prior. Indeed, we see in
+Figure~\ref{fig:fb-null} that the feature birthday-5 has a very high weight as it relates to block 1 – but it is unlikely that birthdays determine graphical structure.

### 2021-11-07 publication/experiments.tex
IDA-draft

@@
 
-We apply the outlined methods to a variety of labelled networks. For reference, the inferred partitions for all of these are potted on Figure~\ref{fig:graphs-all}.
+We apply the outlined methods to a variety of labelled networks:
 
@@
 %
+For reference, the inferred partitions for all of these are given on Figure~\ref{fig:graphs-all}.
 We first require metrics to assess model performance. We define the average
@@
 \end{equation}
-
+%
 Next, to assess the performance of the feature-to-block predictor, 
@@
 \end{equation}
-
+%
+For the higher-dimensional datasets, we also apply the 
+dimensionality reduction method 
+of Section~\ref{sec:dim-reduction}.  
+We then retrain the feature-block predictor using only the retained 
+feature set $\Dcal'$, and report the log-loss over the training and 
+test sets for the reduced classifier -- 
+denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively. 
 \input{experiments-figures}

### 2021-11-07 publication/inference.tex
IDA-draft

@@
 Although the prior is easily computable, computing the likelihood would require a summation over all latent block-states, $p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$, and is therefore intractable.
-This approach is made doubly intractable as we would also need to compute the normalising constant $p(A|X) = \int p(\theta) \cdot p(A | X, \theta) d\theta$.
+This approach is doubly intractable as we would also need to compute the normalising constant $p(A|X)$.
 Clearly we cannot directly compute the form of the posterior, so instead we aim to draw samples from it,
@@
 \end{equation}
-
+%
 We propose an iterative approach to obtain these samples
-$\{\theta^{(t)}\}_t$. We first draw a sample $b^{(t)}$ 
+$\{\theta^{(t)}\}$. We first draw a sample $b^{(t)}$ 
 from the block membership posterior,
@@
 \end{equation}
-
+%
 This can be implemented through a two-level
@@
 \end{equation}
-
+%
 This is an example of a pseudo-marginal approach. Indeed, Andrieu and Roberts \cite{pseudo-marginal} show that~(\ref{eqn:theta-unbiased})
@@
 concurrently. Instead, we run the $b$-chain to completion and use it 
-to generate $\hat{y}$ also allowing us to vary the lengths of each. This change shortens the burn-in of the $\theta$-chain by reducing the variance of our proposal distribution.
+to generate $\hat{y}$ also allowing us to vary the lengths of each.
 
@@
 
-Once the samples $\left\{ \theta^{(t)} \right\} \sim p(\theta | A, X)$
-have been obtained, we can compute the empirical mean and standard deviation of each component of $\theta$. Switching to matrix notation $\theta = W$ we define:
+Evaluating $U$ and $\nabla U$ is linear in the dimension of our feature-space $D$ so there is computational incentive to reduce $D$.
+Given the samples $\left\{ \theta^{(t)} \right\} \sim p(\theta | A, X)$, we can compute the empirical mean and standard deviation of each component of $\theta$.
+Switching to matrix notation $\theta = W$ we define:
 %
@@
 $(-c, c)$ for all block indices $i$.
-If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \hat{\mu}_{ij}, \hat{\sigma}_{ij}^2)$, then this would be analogous to a hypothesis test on the magnitude of $W_{ij}$ compared to $c$ with multiplier $k$ in~(\ref{eqn:fij}) determines the degree of significance of the result.
-However, as the Laplace approximation is not exact, we regard this as a useful heuristic and not an exact method.
+If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \hat{\mu}_{ij}, \hat{\sigma}_{ij}^2)$, then this would be analogous to a hypothesis test on the magnitude of $W_{ij}$ compared to $c$ with multiplier $k$ in~(\ref{eqn:fij}) determining the degree of significance of the result.
 Conversely, if we want to fix the number of dimensions in our reduced feature set $|\Dcal'|=D'$, the problem then becomes finding the largest value of $c$ such that $|\Dcal'|=D'$ given $k=k_0$:

### 2021-11-16 publication/abstract.tex
IDA-draft
IK Edits

@@
 \begin{abstract}
-Labelled networks are a very common and important class of data.
+Labelled networks are an important class of data,
+naturally appearing
+in numerous applications in science and engineering.
 A typical inference goal is to determine how the vertex labels
 (or {\em features}) affect the network's structure.
-
 In this work, we introduce a new generative model, the feature-first block model (FFBM),
-to facilitate the use of richer queries on labelled network structure.
-We develop a Bayesian framework and devise a two-level Markov-Chain Monte-Carlo approach to efficiently sample the FFBM parameters.
-This allows us to infer if and how the observed vertex-features affect macro-structure.
-
+that facilitates the use of rich queries on labelled networks.
+We develop a Bayesian framework and devise a two-level Markov chain Monte 
+Carlo approach to efficiently sample from the
+relevant posterior distribution of the FFBM parameters. This allows us to infer if and how the observed vertex-features affect macro-structure.
 We apply the proposed methods to a variety of network data
@@
 
-\keywords{Labelled Networks \and Bayesian Inference.}
+\keywords{Stochastic Block Model \and Labelled Networks \and Inference.}
 

### 2021-11-16 publication/conclusion.tex
IK Edits

@@
 The proposed Feature-First Block Model (FFBM)
-is a new generative model for labelled networks well-suited for describing how features
-affect structure.
-It is easy to find vertex features that are in some way 
-correlated with the graphical structure; only when 
-we find the feature that best describes the most pronounced partition,
-do we have a stronger case for causation.
+is a new generative model for labelled networks.
+It is a hierachical Bayesian model, 
+well-suited for describing how features
+affect network structure.
+The Bayesian inference tools developed in this 
+work facilitate the identification of 
+vertex features that are in some way 
+correlated with the network's graphical structure.
+Consequently, 
+finding the features that best describe the most 
+pronounced partition, makes it possible in practice
+to examine the existence of -- and to make a case for --
+causal relationships.
 
-Using this new model,
-we go on to describe an efficient inference algorithm to sample 
-the parameters of the FFBM.
-The idea is to divide the graph into its most natural partition and determine whether 
-the vertex features can accurately explain this partition. 
-This is shown to be effective at extracting and describing 
-the most natural communities in a labelled network. Nevertheless, the approach 
-can only currently explain the structure at the macro-scale. Future work will benefit from extending 
-the FFBM to be hierarchical in nature. That way, the structure of the network 
-can be explained at all length-scales of interest.
+An an efficient MCMC algorithm 
+is developed for sampling 
+from the posterior distribution of the
+the relevanat parameters in the FFBM:
+The main idea is to divide up the graph into 
+the most natural partition under the associated
+parameter values, and then to determine whether 
+the vertex features can accurately explain the partition. 
+Through several applications on empirical
+network data, this approach 
+is shown to be effective at extracting and describing 
+the most natural communities in a labelled network. 
+Nevertheless, it
+can only currently explain the structure at the macroscopic
+scale. Future work will benefit from extending 
+the FFBM to a further hierarchical model,
+so that
+the structure of the network 
+can be explained at all scales of interest.

### 2021-11-16 publication/derivations.tex
IK Edits

@@
 because the prior is the same for each $w_j$. 
-Therefore, the integral can only be a function of at most $x_i$ and $\sigma_\theta^2$,
+Therefore, the integral can only be a function of $x_i$ and $\sigma_\theta^2$,
 which means that, as a function of $b$, $p(b|X)\propto 1$. As
 $b$ takes values in $[B]^N$, we necessarily have:
-%
-\begin{equation}
-	p(b | X) = \frac{1}{\big|[B]^N\big|}=B^{-N}.
-\end{equation}
+$p(b | X) =1/\big|[B]^N\big|=B^{-N}$.
 
@@
 $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and 
-$a_{ij} \coloneqq \phi_j(x_i; \theta)$, we have that,
+$a_{ij} \coloneqq \phi_j(x_i; \theta)$ as before, we have that,
 %
@@
 each of its components,
-$\nicefrac{\partial U}{\partial w_k}$ for $k \in [B]$.
+$\nicefrac{\partial U}{\partial w_k}$, for $k \in [B]$.
 To that end, we first compute,
@@
 
-Full source code is available at:
-\url{https://github.com/LozzaTray/Jormungandr-code}. Table~\ref{tab:hyperparams} contains the hyper-parameters for each experiment run. The set of retained samples are generated as,
+Full source code is available at:\\
+\centerline{\url{https://github.com/LozzaTray/Jormungandr-code}}
+Table~\ref{tab:hyperparams} contains all 
+hyper-parameter values used in our experiments. The set of retained 
+samples are generated as,
 $
@@
 %
-\begin{table}[!h]
+\begin{table}[!ht]
 	\centering
-	\caption{Hyper-parameter values for each experiment}
+	\caption{Hyper-parameter values for each experiment.}
 	\label{tab:hyperparams}
@@
 		\\
-		FB Egonet &
+		FB egonet &
 		10 & 0.7 & 1 &

### 2021-11-16 publication/experiments-detailed.tex
IK Edits

@@
 We see that the dimensionality reduction procedure 
-brings the training and test losses closer together. This implies that 
-the features we keep are indeed correlated with the underlying graphical 
+brings the training and test losses closer. This indicates that 
+the retained
+are indeed well correlated with the underlying graphical 
 partition and that the approach generalises correctly.
@@
 %
-The average description length per entity,
-$\bar{S}_e$, of the graph, 
+The average description length per entity of the graph,
+$\bar{S}_e$,
 has very small variance, suggesting that
@@
 %
-The vast majority of blocks are composed of a single class. 
+Almost all blocks are composed of a single class. 
 However, some blocks have two comparably strong classes as their predictors (e.g. blocks 2 and 5). 
@@
 %
-Figure~\ref{fig:school-accuracy} shows excellent accuracy for the majority of blocks.
-In fact the only blocks with low accuracy are those that have a school-class span two blocks such that we cannot reliably distinguish between the two.
-This is more pronounced when we apply hard classification rather than the soft cross-entropy loss.
-Perhaps there are unobserved features which explain this divide.
+Figure~\ref{fig:school-accuracy} shows excellent accuracy for most blocks. In fact the only blocks with low accuracy are those with a `school-class'
+feature that spans two blocks, such that we cannot reliably distinguish between the two.
+This is more pronounced when we apply hard classification rather than cross-entropy loss.
+It is possible that there are unobserved features here
+which explain this divide.
 
@@
 feature may uniquely identify a small set of vertices; if these are all in the same block, then the classifier may overfit despite the penalty imposed by the prior. Indeed, we see in
-Figure~\ref{fig:fb-null} that the feature birthday-5 has a very high weight as it relates to block 1 – but it is unlikely that birthdays determine graphical structure.
+Figure~\ref{fig:fb-null} that the feature `birthday-5' has a very high weight as it relates to block 1 – but it is unlikely that birthdays determine graphical structure.

### 2021-11-16 publication/experiments.tex
IK Edits

@@
-\section{Experiments}
+\section{Experimental Results}
 \label{sec:experiments}
 
-We apply the outlined methods to a variety of labelled networks:
+We apply our proposed methods to a variety of labelled networks:
 
 \begin{itemize}
-	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=3$) -- network of Amazon political book sales, published close to the 2004 presidential election.
+	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=3$) -- network of Amazon political books published close to the 2004 presidential election.
 Two books are connected if they were frequently co-purchased.
 Vertex features encode the political affiliation of the author (liberal, conservative, or neutral).
-	\item \textbf{Primary school dynamic contacts} \cite{schools} ($N=238, E=5539, D=13$) -- network of face-to-face contacts amongst students and teachers at a primary school in Lyon, France.
-Vertex features include class membership (one of 10 values: 1A-5B), gender (male, female) and teacher status encoded as an 11th school-class.
-We choose to analyse just the second day of results.
-	\item \textbf{Facebook egonet} \cite{fb-snap} ($N=747, E=30025, D=480$) -- an assortment of Facebook users' friends lists.
-Vertex features are fully anonymised and encode information about each user's education history, languages spoken, gender, home-town, birthday etc.
-We focus on the egonet with id 1912.
+	\item \textbf{Primary school dynamic contacts} \cite{schools} ($N=238, E=5539, D=13$) -- network of 238 individuals (students and teachers),
+with edges denoting face-to-face contacts at a primary school in Lyon, France. 
+The vertex features are class membership (one of 10 values: 1A-5B), gender (male, female), and status (teacher, student). We choose to analyse just the second day of results.
+	\item \textbf{Facebook egonet} \cite{fb-snap} ($N=747, E=30025, D=480$) -- network of Facebook users with edges denoting ``friends''.
+Vertex features are fully anonymised and encode information about each user's education history, languages spoken, gender, home-town, birthday etc. We focus on the egonet with id 1912.
 \end{itemize}
@@
 For reference, the inferred partitions for all of these are given on Figure~\ref{fig:graphs-all}.
-We first require metrics to assess model performance. We define the average
-description length per entity (nodes and edges) $\bar{S}_e$ as a suitable metric to gauge the SBM fit:
+We employ the following metrics to assess model performance. 
+First, the average
+description length per entity (nodes and edges) 
+$\bar{S}_e$ 
+used to gauge the SBM fit is defined as:
 %
@@
 Next, to assess the performance of the feature-to-block predictor, 
-we randomly partition the vertex set $[N]$ such that a constant fraction $f$ forms our training set $\Gcal_0$ and the remainder form our test set $\Gcal_1$.
-The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain. As $|\Gcal_0| \neq |\Gcal_1|$ in general, we use the average cross-entropy loss 
-over each set to gauge fit,
+the vertex set $[N]$ 
+is partitioned at random so that 
+a constant fraction $f$ of vertices form the training set $\Gcal_0$ and 
+the remainder form the test set $\Gcal_1$.
+The $b$-chain is run using the whole network but we $\in \Gcal_0$ is 
+used to train the $\theta$-chain. 
+% As $|\Gcal_0| \neq |\Gcal_1|$, in general, 
+Then the  the average cross-entropy loss 
+over each set is used to gauge the quality of the fit,
 %
@@
 where $\star \in \{0, 1\}$ toggles between the training and test sets.
+and $\hat{y}_{ij}$ is defined in~(\ref{eqn:y-hat}).
 %
-Nevertheless, the cross-entropy loss is a coarse measure of fit. We wish to define a new measure of fit specific to each detected block. Let us define
+Nevertheless, the cross-entropy loss is a coarse measure of fit. 
+A new measure, specific to each detected block,
+can be defined as follows. Let
+$\Bcal_\star(j)$ 
+be the set of vertices with maximum a posteriori probability of belonging 
+to block~$j$,
 $
-	\Bcal_\star(j) \coloneqq \{i \in \Gcal_\star : \hat{b}_i = j\}
+	\Bcal_\star(j) \coloneqq \{i \in \Gcal_\star : \hat{b}_i = j\},
 $
@@
 $ 
-	\hat{b}_i \coloneqq \argmax_j \hat{y}_{ij}.
+	\hat{b}_i \coloneqq \argmax_j \hat{y}_{ij},
 $
-Now $\Bcal_\star(j)$ is the set of vertices that have maximum a posteriori probability of belonging to block $j$.
-We now define the {\em block-accuracy} for block $j$ as in \ref{eqn:accuracy}.
-This effectively tests whether the feature-to-block and the graph-to-block predictions agree in their largest component.
+and
+define the {\em block-accuracy} for block $j$ as,
 %
@@
 %
+This effectively tests whether the feature-to-block and 
+graph-to-block predictions agree in their largest component.
 For the higher-dimensional datasets, we also apply the 

### 2021-11-16 publication/ffbm.tex
IDA-draft
IK Edits

@@
 
-In this section we propose a novel generative model for labelled networks. We call this the feature-first block model (FFBM -- illustrated in Figure~\ref{fig:ffbm}).
+In this section we propose a novel generative model for labelled networks. We call this the feature-first block model (FFBM),
+illustrated in Figure~\ref{fig:ffbm}).
 
@@
 where $D$ is the number of features associated with each vertex.
-For the datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$. We write $X$ for the $N\times D$ {\em feature matrix} containing
+For example, in the datasets we analyse we deal with binary feature flags
+(denoting the presence/absence of each feature),
+so $\Xcal = \{0, 1\}$. We write $X$ for the $N\times D$ {\em feature matrix} containing
 the feature vectors $\{x_i\}_{i=1}^{N}$ 
@@
 \end{equation}
-
+%
 Note that $\phi_{b_i}$ has the form of a softmax activation function.
@@
 %
-\begin{figure}
+\begin{figure}[!ht]
 	\centering
@@
 \end{equation}
-
+%
 This choice of prior gives a very
@@
 \end{equation}
-
-The proof is given in Appendix~\ref{appdx:b|x}.
-This an important simplification as evaluating $p(b | X)$ does not require an expensive  integration over $\theta$ nor does it require the exact value of $X$.
+%
+The proof is given in Appendix~\ref{appdx:b|x}. This an important simplification as evaluating $p(b | X)$ does not require an expensive  integration over $\theta$ nor does it depend on $X$.
 Peixoto \cite{Peixoto-Bayesian-Microcanonical} proposes careful choices for 
-the priors on the additional microcanonical SBM parameters $\psi$, which we adopt. 
+the priors on the additional microcanonical SBM parameters $\psi$, which we adopt without repeating their exact form here. 
 The idea is to write the joint distribution on $(b, e, k)$ as a product of 
@@
 independent given $b$.
-All that concerns the main argument is that it has
+All that concerns the main argument is that $p(\psi|b)$ has
 an easily computable form.

### 2021-11-16 publication/inference.tex
IK Edits

@@
 Having completed the definition of the FFBM, we wish to leverage it 
-to perform inference. Specifically, given a labelled network $(A, X)$, we wish to infer if and how the observed features $(X)$ impact graphical structure $(A)$. Formally,
+to perform inference. Specifically, given a labelled network $(A, X)$, we wish to infer if and how the observed features $X$ impact the graphical structure $A$. Formally,
 this means characterising the posterior distribution:
@@
 $
-Although the prior is easily computable, computing the likelihood would require a summation over all latent block-states, $p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$, and is therefore intractable.
-This approach is doubly intractable as we would also need to compute the normalising constant $p(A|X)$.
-Clearly we cannot directly compute the form of the posterior, so instead we aim to draw samples from it,
+Although the prior is easily computable, 
+computing the likelihood 
+requires summing over all latent block-states, 
+$p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$, which is 
+clearly impractical. In fact, this
+approach is doubly intractable as we would also 
+need to compute the normalising constant $p(A|X)$.
+Therefore, following standard Bayesian practice,
+instead we aim to draw samples from the posterior,
 %
@@
 %
-We propose an iterative approach to obtain these samples
+We propose an iterative Markov chain Monte Carlo
+(MCMC) approach to obtain these samples
 $\{\theta^{(t)}\}$. We first draw a sample $b^{(t)}$ 
@@
 \begin{equation}
-	b^{(t)} \sim p \Big( b | A, X \Big) \quad \textrm{then} \quad
-	\theta^{(t)} \sim p\left(\theta | X, b^{(t)} \right).
+	b^{(t)} \stackrel{\rm distr}{\approx} p \big( b | A, X \big) 
+	\quad \textrm{then} \quad
+	\theta^{(t)} \stackrel{\rm distr}{\approx} 
+	p\big(\theta | X, b^{(t)} \big),
 \end{equation}
 %
-This can be implemented through a two-level
-Markov-Chain via the Metropolis-Hastings (MH) 
+where these approximations become exact as
+the number of MCMC iterations $t\to\infty$.
+As described in the following subsections,
+this can be implemented through a two-level
+Markov chain via the Metropolis-Hastings (MH) 
 algorithm \cite{hastings-alg}.
-The splitting of the Markov-Chain into two levels allows us to side-step the summation over
+The splitting of the Markov chain into two levels allows us to side-step the summation over
 all latent $b \in [B]^N$ required to directly compute the likelihood, $p(A| X, \theta)$.
-The resulting $\theta^{(t)}$ samples are unbiased in that the expectation of 
-their distribution is the true posterior:
+The resulting $\theta^{(t)}$ samples are asymptotically
+unbiased in that the expectation of 
+their distribution converges to the true posterior:
 %
 \begin{equation}
+\lim_{t\to\infty}
 \Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] = \sum_{b \in [B]^N} p(\theta | X, b) p(b | A, X) = p(\theta | A, X).
@@
 %
-This is an example of a pseudo-marginal approach. Indeed, Andrieu and Roberts \cite{pseudo-marginal} show that~(\ref{eqn:theta-unbiased})
-is sufficient to prove that, for large enough $t$, $\theta^{(t)} \sim p(\theta | A, X)$, as required.
+This is an example of a pseudo-marginal approach;
+see, e.g., Andrieu and Roberts~\cite{pseudo-marginal} 
+for a detailed rigorous derivation based on~(\ref{eqn:theta-unbiased}).
 %
-\begin{figure}[!h]
+\begin{figure}[!ht]
 	\centering
@@
 
-To generate our $b$-samples, we adopt the Markov chain Monte-Carlo procedure of
+To generate the required $b$-samples, we adopt the MCMC
+procedure of
 \cite{Peixoto-MCMC},
-which relies on writing the posterior in the following form:
+which relies on writing the posterior in the following form,
 %
 \begin{equation}
-	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) = \pi_b(b).
+	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) = \pi_b(b),
 \end{equation}
 %
-Now $\pi_b(\cdot)$ is the un-normalised target density.
+where $\pi_b(\cdot)$ denotes the un-normalised target density.
 Since we are using the microcanonical SBM formulation, there is only one 
@@
 $p(A | b, X) = \sum_{\psi} \nolimits p(A , \psi | b, X) = p(A, \psi^* | b, X)$.
-We also define the microcanonical entropy of the configuration as,
+In
+this context, the microcanonical entropy of the configuration $b$
+is,
 %
 \begin{equation}
-	S(b) \coloneqq - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big).
+	S(b) \coloneqq - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big),
 	\label{eqn:dl-form}
@@
 %
-This can be thought of as the optimal
+which can be thought of as the optimal
 ``description length'' of the graph. 
+This expression will later be employed 
+to help evaluate experimental results. 
 The exact form of the proposal $q_b$ is explored thoroughly in
@@
 
-The invariant distribution we wish to target for the $\theta$-samples is the posterior of $\theta$ given the values of the pair $(X, b)$. 
+The target distribution for the required $\theta$-samples 
+is the posterior of $\theta$ given the values of the pair $(X, b)$. 
 We write this as,
@@
 %
-where we write $U(\theta)$ for the negative log-posterior. We define $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$. 
-Discarding constant terms, we can then write $U(\theta)$ as,
+where $U(\theta)$ denotes the negative log-posterior. Let $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$. 
+Discarding constant terms, $U(\theta)$ can be expressed as,
 %
@@
 	U(\theta) = \left( \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} \|\theta\|^2 = N \cdot \Lcal(\theta) + \frac{1}{2\sigma_\theta^2} \|\theta\|^2.
+	+ \frac{1}{2\sigma_\theta^2} \|\theta\|^2 = N \cdot \Lcal(\theta) + \frac{1}{2\sigma_\theta^2} \|\theta\|^2;
 	\label{eqn:U-form}
@@
 %
-See Appendix \ref{appdx:form-U} for the derivation.
+see Appendix \ref{appdx:form-U}.
 The function $U(\theta)$ is a typical objective function for neural network training.
 The first term $N \cdot \Lcal(\theta)$ is introduced by the likelihood and represents the cross-entropy between the graph-predicted and feature-predicted block memberships.
-The second term, introduced by the prior, brings a form of regularisation, guarding against over-fitting.
-Our goal is to draw samples from the posterior $\pi_\theta(\cdot) \propto \exp(-U(\cdot))$.
-We choose to adopt the Metropolis-adjusted Langevin algorithm (MALA) \cite{mala-tweedie}, which uses $\nabla U$ to bias the proposal towards regions of higher density.
-Given the current sample $\theta$, we generate
-a new sample $\theta'$ as,
+The second term, introduced by the prior, brings a form of regularisation, guarding against over-fitting. In order to draw samples from the posterior 
+$\pi_\theta \propto \exp(-U)$ we adopt the Metropolis-adjusted Langevin 
+algorithm (MALA) \cite{mala-tweedie}, which uses $\nabla U$ to bias the 
+proposal towards regions of higher density. Given the current 
+sample $\theta$, a proposed 
+new sample $\theta'$ is generated from,
 %
@@
 So far, each $\theta^{(t)}$ update has used its corresponding $b^{(t)}$ sample.
-This means the evaluation of $U^{(t)}$ and $\nabla U^{(t)}$ has high variance, leading to longer burn-in.
+This means the evaluation of $U^{(t)}$ and $\nabla U^{(t)}$ has high variance, leading to longer burn-in and possibly slower MCMC convergence.
 The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U^{(t)}$ and $\nabla U^{(t)}$ which depends only on the matrix $y^{(t)}$ with entries $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
@@
 %
-We can obtain an unbiased estimate for this quantity using 
+An unbiased estimate for this can be obtained using 
 the thinned $b$-samples after burn-in.
@@
 %
-The same matrix $\hat{y}$ is used in the update step
-for each $\theta^{(t)}$.
+The same matrix $\hat{y}$ is used in each $\theta^{(t)}$ update step.
 This way, it is not necessary to run the $b$ and $\theta$ Markov chains 
@@
 
-Evaluating $U$ and $\nabla U$ is linear in the dimension of our feature-space $D$ so there is computational incentive to reduce $D$.
-Given the samples $\left\{ \theta^{(t)} \right\} \sim p(\theta | A, X)$, we can compute the empirical mean and standard deviation of each component of $\theta$.
-Switching to matrix notation $\theta = W$ we define:
+The complexity of evaluating $U$ and $\nabla U$ is linear in 
+the dimension of the feature space $D$,
+so there is computational incentive to reduce $D$.
+Given the samples $\left\{ \theta^{(t)} \right\}$, we can compute the empirical mean and standard deviation of each component of $\theta$. 
+Switching to the matrix notation $W$ for $\theta$,
+let:
 %
@@
 	\label{eqn:fij} \\
-	\Dcal' &\coloneqq \left\{ j \in [D] : \exists i \in [B] \textrm{ s.t. }  \Fcal_i(j) = \emptyset \right\}
+	\Dcal' &\coloneqq \left\{ j \in [D] : \exists i \in [B] \textrm{ s.t. }  \Fcal_i(j) = \emptyset \right\}.
 	\label{eqn:kept-feature-set}
@@
 $(\hat{\mu}_{ij} - k\hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij})$ overlaps with
-$(-c, c)$ for all block indices $i$.
+$(-c, c)$ for all $i$.
 If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \hat{\mu}_{ij}, \hat{\sigma}_{ij}^2)$, then this would be analogous to a hypothesis test on the magnitude of $W_{ij}$ compared to $c$ with multiplier $k$ in~(\ref{eqn:fij}) determining the degree of significance of the result.

### 2021-11-16 publication/introduction.tex
Flesh out inference
IK Edits

@@
 
-Many real-world networks exhibit strong community structure, with most nodes often belonging to a densely connected cluster. 
+Many real-world networks exhibit strong community structure, with most nodes belonging to densely connected clusters. 
 Finding ways to recover the latent communities from the observed graph is an important
-task in many applications. In this work, we restrict our attention to vertex-labelled networks, 
+task in many applications,
+including
+compression \cite{cluster-compression} and link prediction 
+\cite{link-prediction}.
+In this work, we examine vertex-labelled networks, 
 referring to the labels as {\em features}.
@@
 Answering this requires a random graph model; the standard is the stochastic block model (SBM) \cite{vanilla-sbm}.
-Numerous variants of the SBM  have been proposed -- such as the MMSBM \cite{mixed-membership-sbm} and OSBM \cite{overlapping-sbm}.
-However, these do not automatically include vertex features in the graph generation process.
+Numerous variants of the SBM  have been proposed, e.g., the MMSBM \cite{mixed-membership-sbm} and OSBM \cite{overlapping-sbm}, but these do not include features in the graph generation process.
 
@@
 The associated model can then be used to test for evidence of heterogeneous connectivity between the feature-grouped blocks.
-Nevertheless, this approach can only consider one feature set at a time and the feature-grouped blocks are often an unnatural partition of the graph.
+Nevertheless, this approach can only consider disjoint feature sets and the feature-grouped blocks are often an unnatural partition of the graph.
 

### 2021-11-16 publication/prelim.tex
IK Edits

@@
 Specifically, we will use the microcanonical variant of the SBM, proposed by Peixoto \cite{Peixoto-Bayesian-Microcanonical}.
-To allow for degree-variability between members of the same block, we adopt the degree-corrected formulation (DC-SBM), defined in~(\ref{defn:microcan-dc-sbm}).
+To allow for degree-variability between members of the same block, we adopt the following degree-corrected
+formulation (DC-SBM):
+% , defined in~(\ref{defn:microcan-dc-sbm}).
 
@@
 	\label{defn:microcan-dc-sbm}
-	Let $N \geq 1$ denote the number of vertices in an undirected graph. The block memberships are encoded by a vector $b \in [B]^N$,
+	Let $N \geq 1$ denote the number of vertices in an undirected graph
+with $E$ edges. The block memberships are encoded by a vector $b \in [B]^N$,
 where $B$ is the number of non-empty blocks.\footnote{For each integer $K\geq 1$, we use the notation $[K]:=\{1,2,\ldots,K\}$.}
@@
 The graph's adjacency matrix $A \in \{0,1\}^{N \times N}$ is generated 
-by placing edges uniformly at random, but respecting the constraints 
-imposed by $e$, $b$ and $k$.
-Specifically, if $A \sim \mbox{\rm DC-SBM}_{\rm MC} (b, e, k)$ then it must satisfy the following, for all $r,s\in[B]$
+by placing edges uniformly at random, conditional
+on the constraints imposed by $b$, $e$ and $k$ being satisfied.
+Specifically, if $A \sim \mbox{\rm DC-SBM}_{\rm MC} (b, e, k)$,
+then with probability~1 it satisfies,
+for all $r,s\in[B]$
 and all $i\in[N]$:

### 2021-11-16 publication/experiments.tex
Final IDA draft

@@
-\section{Experimental Results}
+\section{Experimental results}
 \label{sec:experiments}
@@
 the remainder form the test set $\Gcal_1$.
-The $b$-chain is run using the whole network but we $\in \Gcal_0$ is 
-used to train the $\theta$-chain. 
-% As $|\Gcal_0| \neq |\Gcal_1|$, in general, 
+The $b$-chain is run using the whole network but only vertices $v \in \Gcal_0$ are 
+used for the $\theta$-chain. 
 Then the  the average cross-entropy loss 
@@
 %
-where $\star \in \{0, 1\}$ toggles between the training and test sets.
+where $\star \in \{0, 1\}$ toggles between the training and test sets
 and $\hat{y}_{ij}$ is defined in~(\ref{eqn:y-hat}).

### 2021-11-16 publication/ffbm.tex
Final IDA draft

@@
 In this section we propose a novel generative model for labelled networks. We call this the feature-first block model (FFBM),
-illustrated in Figure~\ref{fig:ffbm}).
+illustrated in Figure~\ref{fig:ffbm}.
 
@@
 where $D$ is the number of features associated with each vertex.
-For example, in the datasets we analyse we deal with binary feature flags
+For example, in the datasets we analyse, we deal with binary feature flags
 (denoting the presence/absence of each feature),
@@
 %
-The proof is given in Appendix~\ref{appdx:b|x}. This an important simplification as evaluating $p(b | X)$ does not require an expensive  integration over $\theta$ nor does it depend on $X$.
+The proof is given in Appendix~\ref{appdx:b|x}. This is an important simplification as evaluating $p(b | X)$ does not require an expensive  integration over $\theta$ nor does it depend on $X$.
 Peixoto \cite{Peixoto-Bayesian-Microcanonical} proposes careful choices for 

### 2021-11-16 publication/prelim.tex
Final IDA draft

@@
 For this we adopt the widely-used stochastic block model (SBM).
-This is a latent variable model where each vertex belongs to a single block and the probability two nodes are connected depends only on the block memberships of each.
+This is a latent variable model where each vertex belongs to a single block and the probability two vertices are connected depends only on the block memberships of each.
 Specifically, we will use the microcanonical variant of the SBM, proposed by Peixoto \cite{Peixoto-Bayesian-Microcanonical}.

### 2022-02-20 iwsm/abstract.tex
IWSM full scaffold

--- 2021-11-16 publication/abstract.tex
@@
-\begin{abstract}
+\abstract{
 Labelled networks are an important class of data,
@@
 can be rank-ordered implicitly according to impact.
-
-\keywords{Stochastic Block Model \and Labelled Networks \and Inference.}
-
-\end{abstract}
+}

### 2022-02-20 iwsm/additional.tex
IWSM full scaffold

--- 2021-06-02 cued/additional.tex
@@
 %
-where $e_r \coloneqq \sum_{s} e_{rs}$ and $(2m)!!
-\coloneqq 2^m m!$.
+where $e_r \triangleq \sum_{s} e_{rs}$ and $(2m)!!
+\triangleq 2^m m!$.
 Nevertheless, a great number of these $\Omega(e)$ total configurations yield the same graph $A$.
@@
 \begin{equation}
-	\Xi(A) \coloneqq \frac{\prod_i k_i !}{\prod_{i,j : i < j} A_{ij} ! \prod_i A_{ii} !! }.
+	\Xi(A) \triangleq \frac{\prod_i k_i !}{\prod_{i,j : i < j} A_{ij} ! \prod_i A_{ii} !! }.
 \end{equation}
@@
 \begin{equation}
-	p(\psi_e=e, \psi_k=k | b) = p(e | b) p(k | e, b) = \left[ \specialchoose{ \specialchoose{B}{2} }{ E} \right]^{-1} 
+	p(\psi_e=e, \psi_k=k | b) = p(e | b) p(k | e, b) = \left[ \genfrac{\{}{\}}{0pt}{}{ \genfrac{\{}{\}}{0pt}{}{B}{2} }{ E} \right]^{-1} 
 	\cdot \left[ \prod_r \frac{\prod_j \eta_j^r !}{n_r! q(e_r, n_r)} \right],
@@
 %
-where $\specialchoose{n}{m}$ is shorthand 
+where $\genfrac{\{}{\}}{0pt}{}{n}{m}$ is shorthand 
 for $\binom{n+m-1}{m} = \frac{(n+m-1)!}{(n-1)!(m)!}$,
@@
 The variable $\eta_j^r$ denotes the number of vertices in block $r$ 
-that have degree $j$; formally, $\eta_j^r \coloneqq \sum_{i} \one\left\{b_i = r \right\} \one \left\{k_i = j \right\}$. 
+that have degree $j$; formally, $\eta_j^r \triangleq \sum_{i} \boldsymbol{1}\left\{b_i = r \right\} \boldsymbol{1} \left\{k_i = j \right\}$. 
 The denominator $q(m, n)$ denotes the number of different histograms 
@@
 at most $n$ non-zero bins that sum to $m$. 
-Finally, $e_r \coloneqq \sum_{s} e_{rs}$ is the total number 
-of half edges in block $r$ and $n_r \coloneqq \sum_{i} \one\{b_i = r\}$ 
+Finally, $e_r \triangleq \sum_{s} e_{rs}$ is the total number 
+of half edges in block $r$ and $n_r \triangleq \sum_{i} \boldsymbol{1}\{b_i = r\}$ 
 is the number of vertices assigned to block $r$.
@@
 We illustrate the adopted method using the primary school dataset \cite{schools} as an example.
-We run the $b$-chain for $T_b = 1,000$ iterations and subsample such that $\Tcal_b$ is computed with $\kappa_b=0.2$ and $\lambda_b=5$.
+We run the $b$-chain for $T_b = 1,000$ iterations and subsample such that $\mathcal{T}_b$ is computed with $\kappa_b=0.2$ and $\lambda_b=5$.
 We then use this to run the $\theta$-chain for $T_\theta = 10,000$ iterations.
@@
 \begin{equation}
-	\bar{U} \coloneqq \frac{1}{T_\theta} \sum_{t \in [T_\theta]} U\left( \theta^{(t)} \right).
+	\bar{U} \triangleq \frac{1}{T_\theta} \sum_{t \in [T_\theta]} U\left( \theta^{(t)} \right).
 \end{equation}
@@
 obtain a less-dependent version. For $\star\in\{b,\theta\}$,
-the remaining sample sets are denoted $\Tcal_\star$
+the remaining sample sets are denoted $\mathcal{T}_\star$
 in the notation of Section~\ref{s:ss},
@@
 \begin{equation}
-	\Tcal_\star = \{T_\star \kappa_\star + i \lambda_\star :  
+	\mathcal{T}_\star = \{T_\star \kappa_\star + i \lambda_\star :  
 	0 \leq i \leq \lfloor T_\star(1 - \kappa_\star) / \lambda_\star \rfloor \},
@@
 $\lambda_\theta=10$ is chosen somewhat arbitrarily as we do not require neighbouring samples to be independent.
-However, this thinning also has the advantage of speeding up computation of quantities that require an average over all the retained samples -- as $|\Tcal_\theta| < T_\theta$.
+However, this thinning also has the advantage of speeding up computation of quantities that require an average over all the retained samples -- as $|\mathcal{T}_\theta| < T_\theta$.
 %
@@
 		\includegraphics[width=\linewidth]{school-U-thinned.png}
-		\caption{Thinned samples $t \in \Tcal_\theta$ indexed by $i$.}
+		\caption{Thinned samples $t \in \mathcal{T}_\theta$ indexed by $i$.}
 		\label{fig:school-U-thinned}
@@
 It is not immediately obvious how to encode this for the feature-to-block classifier.
-Although this is technically only a single dimension, we choose to expand the data into a set of 10 binary feature flags all taking values in the set $\Xcal = \{0, 1\}$.
+Although this is technically only a single dimension, we choose to expand the data into a set of 10 binary feature flags all taking values in the set $\mathcal{X} = \{0, 1\}$.
 Although only one flag will be set at a time it is the simplest method for representing discrete-valued data.
-The reason we choose $\Xcal=\{0, 1\}$ and not $\Xcal = \{-1, 1\}$ is that the former allows for simpler analysis of the resulting weight values; when a feature $d$ is switched off then $x_{d}=0$ given $\Xcal = \{0, 1\}$, and the feature has no impact on the softmax output.
-This would not be the case for $\Xcal = \{-1, 1\}$.
+The reason we choose $\mathcal{X}=\{0, 1\}$ and not $\mathcal{X} = \{-1, 1\}$ is that the former allows for simpler analysis of the resulting weight values; when a feature $d$ is switched off then $x_{d}=0$ given $\mathcal{X} = \{0, 1\}$, and the feature has no impact on the softmax output.
+This would not be the case for $\mathcal{X} = \{-1, 1\}$.
 
@@
 
-We are given samples $\{\theta^{(t)}\}_{t \in \Tcal_\theta} \sim p(\theta | A, X)$ and wish to determine the statistical significance of the weights.
+We are given samples $\{\theta^{(t)}\}_{t \in \mathcal{T}_\theta} \sim p(\theta | A, X)$ and wish to determine the statistical significance of the weights.
 We adopt matrix notation for simplicity, representing $\theta$ with the $B \times D$ matrix of feature weights $W$.
@@
 \begin{equation}
-	p(W_{ij}|A, X) \approx \Gaussian(W_{ij}|\hat{\mu}_{ij}, \hat{\sigma}_{ij}^2).
+	p(W_{ij}|A, X) \approx \mathcal{N}(W_{ij}|\hat{\mu}_{ij}, \hat{\sigma}_{ij}^2).
 \end{equation}
@@
 \begin{equation}
-	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
-	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2.
+	\hat{\mu}_{ij} \triangleq \frac{1}{|\mathcal{T}_\theta|} \sum_{t \in \mathcal{T}_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
+	\hat{\sigma}_{ij}^2 \triangleq \frac{1}{|\mathcal{T}_\theta|} \sum_{t \in \mathcal{T}_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2.
 \end{equation}
@@
 This approximation is not exact but we can show it is accurate empirically.
-Indeed, if we run the primary school experiment with hyper-parameters given in Appendix \ref{appdx:hyperparams}, we can then plot histograms of the collected $W^{(t)}$-samples (for $t \in \Tcal_\theta$) and compare these to the Laplace approximation.
+Indeed, if we run the primary school experiment with hyper-parameters given in Appendix \ref{appdx:hyperparams}, we can then plot histograms of the collected $W^{(t)}$-samples (for $t \in \mathcal{T}_\theta$) and compare these to the Laplace approximation.
 The results are given on Figure \ref{fig:school-histogram}, which show the Laplace approximation is remarkably reliable.
@@
 %
-where $Z \sim \Gaussian(0,1)$ and $\Phi(\cdot)$ is the standard Gaussian cumulative density function.
+where $Z \sim \mathcal{N}(0,1)$ and $\Phi(\cdot)$ is the standard Gaussian cumulative density function.
 We introduce the variable $k$ to control the degree of significance of the result.

### 2022-02-20 iwsm/conclusion.tex
IWSM full scaffold

--- 2021-11-16 publication/conclusion.tex
@@
 is a new generative model for labelled networks.
-It is a hierachical Bayesian model, 
+It is a hierarchical Bayesian model, 
 well-suited for describing how features
@@
 
-An an efficient MCMC algorithm 
+An efficient MCMC algorithm 
 is developed for sampling 
-from the posterior distribution of the
-the relevanat parameters in the FFBM:
-The main idea is to divide up the graph into 
-the most natural partition under the associated
+from the posterior distribution of
+the relevant parameters in the FFBM;
+the main idea is to divide up the graph into 
+its most natural partition under the associated
 parameter values, and then to determine whether 

### 2022-02-20 iwsm/experiments.tex
IWSM full scaffold

--- 2021-11-16 publication/experiments.tex
@@
 \begin{equation}
-	\bar{S}_e \coloneqq \frac{1}{(N+E) |\Tcal_b|} \sum_{t\in \Tcal_b} S \left( b^{(t)} \right).
+	\bar{S}_e \triangleq \frac{1}{(N+E) |\mathcal{T}_b|} \sum_{t\in \mathcal{T}_b} S \left( b^{(t)} \right).
 	\label{eqn:mean-dl}
@@
 is partitioned at random so that 
-a constant fraction $f$ of vertices form the training set $\Gcal_0$ and 
-the remainder form the test set $\Gcal_1$.
-The $b$-chain is run using the whole network but only vertices $v \in \Gcal_0$ are 
+a constant fraction $f$ of vertices form the training set $\mathcal{G}_0$ and 
+the remainder form the test set $\mathcal{G}_1$.
+The $b$-chain is run using the whole network but only vertices $v \in \mathcal{G}_0$ are 
 used for the $\theta$-chain. 
@@
 \begin{equation}
-	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star^{(t)},
+	\bar{\mathcal{L}}_\star \triangleq \frac{1}{|\mathcal{T}_\theta|} \sum_{t \in \mathcal{T}_\theta} \mathcal{L}_\star^{(t)},
 	\quad \textrm{where} \quad
-	\Lcal_\star^{(t)} \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)},
+	\mathcal{L}_\star^{(t)} \triangleq \frac{1}{|\mathcal{G}_\star|} \sum_{i \in \mathcal{G}_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)},
 	\label{eqn:cross-entropy-loss}
@@
 can be defined as follows. Let
-$\Bcal_\star(j)$ 
+$\mathcal{B}_\star(j)$ 
 be the set of vertices with maximum a posteriori probability of belonging 
@@
 $
-	\Bcal_\star(j) \coloneqq \{i \in \Gcal_\star : \hat{b}_i = j\},
+	\mathcal{B}_\star(j) \triangleq \{i \in \mathcal{G}_\star : \hat{b}_i = j\},
 $
@@
 $ 
-	\hat{b}_i \coloneqq \argmax_j \hat{y}_{ij},
+	\hat{b}_i \triangleq \underset{j}{\operatorname{argmax}}\hat{y}_{ij},
 $
@@
 \begin{equation}
-	\eta_\star(j) \coloneqq \frac{1}{|\Bcal_\star (j)| \cdot 
-	|\Tcal_\theta| } 
-	\sum_{i \in \Bcal_\star (j)}  \sum_{t \in \Tcal_\theta}
-	\one \left\{\hat{b}_i = \argmax_j \phi_j \left( x_i; \theta^{(t)} \right) \right\}.
+	\eta_\star(j) \triangleq \frac{1}{|\mathcal{B}_\star (j)| \cdot 
+	|\mathcal{T}_\theta| } 
+	\sum_{i \in \mathcal{B}_\star (j)}  \sum_{t \in \mathcal{T}_\theta}
+	\boldsymbol{1} \left\{\hat{b}_i = \underset{j}{\operatorname{argmax}}\phi_j \left( x_i; \theta^{(t)} \right) \right\}.
 	\label{eqn:accuracy}
@@
 We then retrain the feature-block predictor using only the retained 
-feature set $\Dcal'$, and report the log-loss over the training and 
+feature set $\mathcal{D}'$, and report the log-loss over the training and 
 test sets for the reduced classifier -- 
-denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively. 
-\input{experiments-figures}
-\input{experiments-detailed}
+denoted $\bar{\mathcal{L}}_0'$ and $\bar{\mathcal{L}}_1'$ respectively.

### 2022-02-20 iwsm/ffbm-detailed.tex
IWSM full scaffold

--- 2021-11-16 publication/ffbm.tex
@@
-\section{Feature-First Block Model}
-
-In this section we propose a novel generative model for labelled networks. We call this the feature-first block model (FFBM),
-illustrated in Figure~\ref{fig:ffbm}.
-
-Let $N$ denote the number of vertices, $B$ the number of blocks
-and ${\cal X}$ the set of values each feature can take.
-We define the vector $x_i \in \Xcal^D$ as the feature vector for vertex $i$, 
-where $D$ is the number of features associated with each vertex.
-For example, in the datasets we analyse, we deal with binary feature flags
-(denoting the presence/absence of each feature),
-so $\Xcal = \{0, 1\}$. We write $X$ for the $N\times D$ {\em feature matrix} containing
-the feature vectors $\{x_i\}_{i=1}^{N}$ 
-as its rows.
-
-For the FFBM, we start with the feature matrix $X$ and generate a random
-vector of block memberships $b \in [B]^N$. For each vertex $i$, the
-block membership $b_i\in[B]$ is generated based on the feature
-vector $x_i$, independently between vertices. The conditional
-distribution of $b_i$ given $x_i$ also depends on a collection
-of weight vectors $\theta=\{w_k\}_{k=1}^B$, where each
-$w_k$ has dimension $D$. We will later find it convenient
-to write $\theta$ as a $B \times D$ matrix of weights $W$. Specifically, 
-the distribution of $b$ given $X$ and $\theta$ is,
-%
-\begin{equation}
-	p(b| X, \theta) = \prod_{i \in [N]} p(b_i | x_i, \theta) = \prod_{i \in [N]} \phi_{b_i} (x_i; \theta)
-	= \prod_{i \in [N]} \frac{\exp\left(w_{b_i}^T x_i\right)}{\sum_{k \in [B]} \exp \left( w_k^T x_i\right)}.
-\end{equation}
-%
-Note that $\phi_{b_i}$ has the form of a softmax activation function.
-More complex models based on different choices for the distributions
-$\phi_{b_i}$ above are also possible, but then deriving meaning from 
-the inferred parameter distributions is more difficult. 
-%
-\begin{figure}[!ht]
-	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=7mm},
-		squarednode/.style={rectangle, draw=black, minimum size=7mm}
-		]
-		% nodes
-		\node[roundnode] (X) at (0, 0) {$X$};
-		\node[squarednode] (b) at (3, 0) {$b$};
-		\node[roundnode] (A) at (6, 0) {$A$};
-
-		% arrows
-		\draw[->] (X.east) -- node[above] {$\theta$} (b.west);
-		\draw[->] (b.east) -- node[above] {$\psi$}(A.west);
-	\end{tikzpicture}
-	\caption{The Feature-First Block Model (FFBM)}
-	\label{fig:ffbm}
-\end{figure}
-
-Once the block memberships $b$ have been generated, we then draw the 
-graph $A$ from the microcanonical DC-SBM with additional parameters 
-$\psi = \{\psi_e, \psi_k\}$:
-%
-\begin{equation}
-	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k).
-	\label{eqn:A-generation}
-\end{equation}
-
 \subsection{Prior selection}
@@
 \begin{equation}
-	p(\theta) \sim \Gaussian \left( \theta ; 0, \sigma_\theta^2 I \right).
+	p(\theta) \sim \mathcal{N} \left( \theta ; 0, \sigma_\theta^2 I \right).
 	\label{eqn:theta-prior}

### 2022-02-20 iwsm/ffbm.tex
IWSM full scaffold

--- 2021-11-16 publication/ffbm.tex
@@
 and ${\cal X}$ the set of values each feature can take.
-We define the vector $x_i \in \Xcal^D$ as the feature vector for vertex $i$, 
+We define the vector $x_i \in \mathcal{X}^D$ as the feature vector for vertex $i$, 
 where $D$ is the number of features associated with each vertex.
@@
 (denoting the presence/absence of each feature),
-so $\Xcal = \{0, 1\}$. We write $X$ for the $N\times D$ {\em feature matrix} containing
+so $\mathcal{X} = \{0, 1\}$. We write $X$ for the $N\times D$ {\em feature matrix} containing
 the feature vectors $\{x_i\}_{i=1}^{N}$ 
@@
 	\centering
-	\begin{tikzpicture}[
-		roundnode/.style={circle, draw=black, minimum size=7mm},
-		squarednode/.style={rectangle, draw=black, minimum size=7mm}
-		]
-		% nodes
-		\node[roundnode] (X) at (0, 0) {$X$};
-		\node[squarednode] (b) at (3, 0) {$b$};
-		\node[roundnode] (A) at (6, 0) {$A$};
-
-		% arrows
-		\draw[->] (X.east) -- node[above] {$\theta$} (b.west);
-		\draw[->] (b.east) -- node[above] {$\psi$}(A.west);
-	\end{tikzpicture}
+%	\begin{tikzpicture}[
+%		roundnode/.style={circle, draw=black, minimum size=7mm},
+%		squarednode/.style={rectangle, draw=black, minimum size=7mm}
+%		]
+%		% nodes
+%		\node[roundnode] (X) at (0, 0) {$X$};
+%		\node[squarednode] (b) at (3, 0) {$b$};
+%		\node[roundnode] (A) at (6, 0) {$A$};
+%		
+%		% arrows
+%		\draw[->] (X.east) -- node[above] {$\theta$} (b.west);
+%		\draw[->] (b.east) -- node[above] {$\psi$}(A.west);
+%	\end{tikzpicture}
+	\includegraphics[width=0.8\textwidth]{img/ffbm.png}
 	\caption{The Feature-First Block Model (FFBM)}
@@
 \end{equation}
-
-\subsection{Prior selection}
-
-To complete the description of our Bayesian framework,
-priors on $\theta$ and $\psi$ must also be specified. 
-We place a Gaussian prior on $\theta$ such that
-each element of $\theta$ has an independent ${\cal N}(0,\sigma_\theta^2)$
-prior, with hyperparameter $\sigma_\theta^2$:
-%
-\begin{equation}
-	p(\theta) \sim \Gaussian \left( \theta ; 0, \sigma_\theta^2 I \right).
-	\label{eqn:theta-prior}
-\end{equation}
-%
-This choice of prior gives a very
-simple form for the conditional distribution of the block membership vector $b$ given $X$; it is a uniform distribution:
-%
-\begin{equation}
-	p(b | X) = \int p(b | X, \theta) p(\theta) d\theta = B^{-N}.
-	\label{eqn:b-pseudo-prior}
-\end{equation}
-%
-The proof is given in Appendix~\ref{appdx:b|x}. This is an important simplification as evaluating $p(b | X)$ does not require an expensive  integration over $\theta$ nor does it depend on $X$.
-Peixoto \cite{Peixoto-Bayesian-Microcanonical} proposes careful choices for 
-the priors on the additional microcanonical SBM parameters $\psi$, which we adopt without repeating their exact form here. 
-The idea is to write the joint distribution on $(b, e, k)$ as a product of 
-conditionals, $p(b, e, k) = p(b) p(e | b) p(k | e, b)= p(b) p(\psi | b)$. 
-In our case, conditioning on $X$ is also necessary, leading to,
-$
-	p(b, \psi | X) = p(b | X) p(\psi | b, X) = p(b | X) p(\psi | b),
-$
-where we used the fact $\psi$ and $X$ are conditionally 
-independent given $b$.
-All that concerns the main argument is that $p(\psi|b)$ has
-an easily computable form.

### 2022-02-20 iwsm/inference-detailed.tex
IWSM full scaffold

--- 2021-11-16 publication/inference.tex
@@
-\section{Inference}
-\label{sec:inference}
-
-Having completed the definition of the FFBM, we wish to leverage it 
-to perform inference. Specifically, given a labelled network $(A, X)$, we wish to infer if and how the observed features $X$ impact the graphical structure $A$. Formally,
-this means characterising the posterior distribution:
-$
-p(\theta|A, X) \propto p(\theta) \cdot p(A | X, \theta).
-$
-Although the prior is easily computable, 
-computing the likelihood 
-requires summing over all latent block-states, 
-$p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$, which is 
-clearly impractical. In fact, this
-approach is doubly intractable as we would also 
-need to compute the normalising constant $p(A|X)$.
-Therefore, following standard Bayesian practice,
-instead we aim to draw samples from the posterior,
-%
-\begin{equation}
-	\label{eqn:theta-target}
-	\theta^{(t)} \sim p(\theta | A, X).
-\end{equation}
-%
-We propose an iterative Markov chain Monte Carlo
-(MCMC) approach to obtain these samples
-$\{\theta^{(t)}\}$. We first draw a sample $b^{(t)}$ 
-from the block membership posterior,
-and then use $b^{(t)}$ to obtain a corresponding
-sample $\theta^{(t)}$:
-%
-\begin{equation}
-	b^{(t)} \stackrel{\rm distr}{\approx} p \big( b | A, X \big) 
-	\quad \textrm{then} \quad
-	\theta^{(t)} \stackrel{\rm distr}{\approx} 
-	p\big(\theta | X, b^{(t)} \big),
-\end{equation}
-%
-where these approximations become exact as
-the number of MCMC iterations $t\to\infty$.
-As described in the following subsections,
-this can be implemented through a two-level
-Markov chain via the Metropolis-Hastings (MH) 
-algorithm \cite{hastings-alg}.
-The splitting of the Markov chain into two levels allows us to side-step the summation over
-all latent $b \in [B]^N$ required to directly compute the likelihood, $p(A| X, \theta)$.
-The resulting $\theta^{(t)}$ samples are asymptotically
-unbiased in that the expectation of 
-their distribution converges to the true posterior:
-%
-\begin{equation}
-\lim_{t\to\infty}
-\Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] = \sum_{b \in [B]^N} p(\theta | X, b) p(b | A, X) = p(\theta | A, X).
-\label{eqn:theta-unbiased}
-\end{equation}
-%
-This is an example of a pseudo-marginal approach;
-see, e.g., Andrieu and Roberts~\cite{pseudo-marginal} 
-for a detailed rigorous derivation based on~(\ref{eqn:theta-unbiased}).
-%
-\begin{figure}[!ht]
-	\centering
-
-	\begin{tikzpicture}[
-		scale=0.8, every node/.style={transform shape},
-		roundnode/.style={circle, draw=black, minimum size=12mm},
-		squarednode/.style={rectangle, draw=black, minimum size=12mm}
-		]
-		% nodes
-		\node[roundnode] (b0) at (0, 2) {$b^{(t)}$};
-		\node[roundnode] (b1) at (5, 2) {$b'$};
-		\node[roundnode] (b2) at (10, 2) {$b^{(t+1)}$};
-		\node[roundnode] (t0) at (0, 0) {$\theta^{(t)}$};
-		\node[roundnode] (t1) at (5, 0) {$\theta'$};
-		\node[roundnode] (t2) at (10, 0) {$\theta^{(t+1)}$};
-
-		% arrows
-		\draw[->] (b0) to node[above] {$\sim q_b ( b^{(t)}, b' | A )$} (b1);
-		\draw[->] (b1) to node[above] {$\alpha_b (b, b' | A )$} (b2);
-		\draw[->] (b0) [out=25, in=155] to node[above] {$1-\alpha_b$} (b2);
-
-		\draw[->] (t0) to node[above] {$\sim q_\theta(\theta^{(t)}, \theta' | X, b^{(t)})$} (t1);
-		\draw[->] (t1) to node[above] {$\alpha_\theta (\theta, \theta' | X, b^{(t)})$} (t2);
-		\draw[->] (t0) [out=-25, in=-155] to node[below] {$1-\alpha_\theta$} (t2);
-
-		\draw[dashed, ->] (b0) to (2, 0.8);
-		\draw[dashed, ->] (b0) to (6.5, 0.8);
-
-	\end{tikzpicture}
-	\caption{$\theta$-sample generation.}
-	\label{fig:samp-sequence}
-\end{figure}
-
-Figure \ref{fig:samp-sequence} shows an overview of the proposed method, with $q$ and $\alpha$ denoting the MH proposal distribution and acceptance probability respectively.
-Note the importance of the simplification in~(\ref{eqn:b-pseudo-prior}). 
-As evaluating $p(b| X)$ does not depend on $X$, 
-we do not need $X$ to sample $b$.
-And on the other level, in order to obtain 
-samples for $\theta$
-we use only $b$ but not $A$, as $(\theta \indep A )| b$.
-
-\FloatBarrier
 \subsection{Sampling block memberships}
@@
 \begin{equation}
-	S(b) \coloneqq - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big),
+	S(b) \triangleq - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big),
 	\label{eqn:dl-form}
@@
 %
-where $U(\theta)$ denotes the negative log-posterior. Let $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$. 
+where $U(\theta)$ denotes the negative log-posterior. Let $y_{ij} \triangleq \boldsymbol{1} \left\{ b_i = j \right\}$ and $a_{ij} \triangleq \phi_j(x_i; \theta)$. 
 Discarding constant terms, $U(\theta)$ can be expressed as,
@@
 	U(\theta) = \left( \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} \|\theta\|^2 = N \cdot \Lcal(\theta) + \frac{1}{2\sigma_\theta^2} \|\theta\|^2;
+	+ \frac{1}{2\sigma_\theta^2} \|\theta\|^2 = N \cdot \mathcal{L}(\theta) + \frac{1}{2\sigma_\theta^2} \|\theta\|^2;
 	\label{eqn:U-form}
@@
 The function $U(\theta)$ is a typical objective function for neural network training.
-The first term $N \cdot \Lcal(\theta)$ is introduced by the likelihood and represents the cross-entropy between the graph-predicted and feature-predicted block memberships.
+The first term $N \cdot \mathcal{L}(\theta)$ is introduced by the likelihood and represents the cross-entropy between the graph-predicted and feature-predicted block memberships.
 The second term, introduced by the prior, brings a form of regularisation, guarding against over-fitting. In order to draw samples from the posterior 
@@
 	\theta' \sim q_\theta\big(\theta, \theta'\big) 
-	= \Gaussian \big( \theta' ; \theta - h \nabla U(\theta), 2h I \big),
+	= \mathcal{N} \big( \theta' ; \theta - h \nabla U(\theta), 2h I \big),
 \end{equation*}
 %
-where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter 
+where $\xi \sim \mathcal{N}(0, I)$ and $h$ is a step-size parameter 
 which may vary with the sample index.
@@
 This means the evaluation of $U^{(t)}$ and $\nabla U^{(t)}$ has high variance, leading to longer burn-in and possibly slower MCMC convergence.
-The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U^{(t)}$ and $\nabla U^{(t)}$ which depends only on the matrix $y^{(t)}$ with entries $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
+The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U^{(t)}$ and $\nabla U^{(t)}$ which depends only on the matrix $y^{(t)}$ with entries $y_{ij}^{(t)} \triangleq \boldsymbol{1}\{b_i^{(t)} = j\}$.
 We would rather deal with the expectation of each $y_{ij}^{(t)}$:
@@
 \begin{equation}
-	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one \left( b_{i}^{(t)} = j \right) \right]
+	\mathbb{E} \left[ y_{ij}^{(t)} \right] = \mathbb{E}_{b^{(t)}} \left[ \boldsymbol{1} \left( b_{i}^{(t)} = j \right) \right]
 	= p(b_i = j | A, X).
@@
 the thinned $b$-samples after burn-in.
-Let $\Tcal_b$  denote the retained set of indices 
-for the $b$-samples and $\Tcal_\theta$ similarly for the $\theta$-chain. 
+Let $\mathcal{T}_b$  denote the retained set of indices 
+for the $b$-samples and $\mathcal{T}_\theta$ similarly for the $\theta$-chain. 
 The unbiased estimate for $y_{ij}^{(t)}$ is then:
@@
 \begin{equation}
-	\hat{y}_{ij} \coloneqq \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} y_{ij}^{(t)} = \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} \one\{b_i^{(t)} = j\}.
+	\hat{y}_{ij} \triangleq \frac{1}{|\mathcal{T}_b|} \sum_{t \in \mathcal{T}_b} y_{ij}^{(t)} = \frac{1}{|\mathcal{T}_b|} \sum_{t \in \mathcal{T}_b} \boldsymbol{1}\{b_i^{(t)} = j\}.
 	\label{eqn:y-hat}
@@
 \begin{equation}
-	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
-	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2.
+	\hat{\mu}_{ij} \triangleq \frac{1}{|\mathcal{T}_\theta|} \sum_{t \in \mathcal{T}_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
+	\hat{\sigma}_{ij}^2 \triangleq \frac{1}{|\mathcal{T}_\theta|} \sum_{t \in \mathcal{T}_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2.
 \end{equation}
 %
-A simple heuristic to discard the least important features requires specifying a cutoff $c > 0$ and a multiplier $k > 0$. We define the function $\Fcal_i(j)$ 
-as in~(\ref{eqn:fij}) and only keep features with indices $d \in \Dcal'$, where $\Dcal'$ is given in~(\ref{eqn:kept-feature-set}).
+A simple heuristic to discard the least important features requires specifying a cutoff $c > 0$ and a multiplier $k > 0$. We define the function $\mathcal{F}_i(j)$ 
+as in~(\ref{eqn:fij}) and only keep features with indices $d \in \mathcal{D}'$, where $\mathcal{D}'$ is given in~(\ref{eqn:kept-feature-set}).
 %
 \begin{align}
-	\Fcal_i(j) &\coloneqq (\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c),
+	\mathcal{F}_i(j) &\triangleq (\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c),
 	\label{eqn:fij} \\
-	\Dcal' &\coloneqq \left\{ j \in [D] : \exists i \in [B] \textrm{ s.t. }  \Fcal_i(j) = \emptyset \right\}.
+	\mathcal{D}' &\triangleq \left\{ j \in [D] : \exists i \in [B] \textrm{ s.t. }  \mathcal{F}_i(j) = \emptyset \right\}.
 	\label{eqn:kept-feature-set}
@@
 $(-c, c)$ for all $i$.
-If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \hat{\mu}_{ij}, \hat{\sigma}_{ij}^2)$, then this would be analogous to a hypothesis test on the magnitude of $W_{ij}$ compared to $c$ with multiplier $k$ in~(\ref{eqn:fij}) determining the degree of significance of the result.
-Conversely, if we want to fix the number of dimensions in our reduced feature set $|\Dcal'|=D'$, the problem then becomes finding the largest value of $c$ such that $|\Dcal'|=D'$ given $k=k_0$:
+If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \mathcal{N}(W_{ij}; \hat{\mu}_{ij}, \hat{\sigma}_{ij}^2)$, then this would be analogous to a hypothesis test on the magnitude of $W_{ij}$ compared to $c$ with multiplier $k$ in~(\ref{eqn:fij}) determining the degree of significance of the result.
+Conversely, if we want to fix the number of dimensions in our reduced feature set $|\mathcal{D}'|=D'$, the problem then becomes finding the largest value of $c$ such that $|\mathcal{D}'|=D'$ given $k=k_0$:
 %
 \begin{equation}
-	c^* = \argmax \{c>0\; : \;|\Dcal'| = D', k=k_0\}.
+	c^* = {\operatorname{argmax}} \{c>0\; : \;|\mathcal{D}'| = D', k=k_0\}.
 	\label{eqn:c-star}

### 2022-02-20 iwsm/inference.tex
IWSM full scaffold

--- 2021-11-16 publication/inference.tex
@@
 \lim_{t\to\infty}
-\Expect_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] = \sum_{b \in [B]^N} p(\theta | X, b) p(b | A, X) = p(\theta | A, X).
+\mathbb{E}_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] = \sum_{b \in [B]^N} p(\theta | X, b) p(b | A, X) = p(\theta | A, X).
 \label{eqn:theta-unbiased}
@@
 
-	\begin{tikzpicture}[
-		scale=0.8, every node/.style={transform shape},
-		roundnode/.style={circle, draw=black, minimum size=12mm},
-		squarednode/.style={rectangle, draw=black, minimum size=12mm}
-		]
-		% nodes
-		\node[roundnode] (b0) at (0, 2) {$b^{(t)}$};
-		\node[roundnode] (b1) at (5, 2) {$b'$};
-		\node[roundnode] (b2) at (10, 2) {$b^{(t+1)}$};
-		\node[roundnode] (t0) at (0, 0) {$\theta^{(t)}$};
-		\node[roundnode] (t1) at (5, 0) {$\theta'$};
-		\node[roundnode] (t2) at (10, 0) {$\theta^{(t+1)}$};
-
-		% arrows
-		\draw[->] (b0) to node[above] {$\sim q_b ( b^{(t)}, b' | A )$} (b1);
-		\draw[->] (b1) to node[above] {$\alpha_b (b, b' | A )$} (b2);
-		\draw[->] (b0) [out=25, in=155] to node[above] {$1-\alpha_b$} (b2);
-
-		\draw[->] (t0) to node[above] {$\sim q_\theta(\theta^{(t)}, \theta' | X, b^{(t)})$} (t1);
-		\draw[->] (t1) to node[above] {$\alpha_\theta (\theta, \theta' | X, b^{(t)})$} (t2);
-		\draw[->] (t0) [out=-25, in=-155] to node[below] {$1-\alpha_\theta$} (t2);
-
-		\draw[dashed, ->] (b0) to (2, 0.8);
-		\draw[dashed, ->] (b0) to (6.5, 0.8);
-
-	\end{tikzpicture}
+%	\begin{tikzpicture}[
+%		scale=0.8, every node/.style={transform shape},
+%		roundnode/.style={circle, draw=black, minimum size=12mm},
+%		squarednode/.style={rectangle, draw=black, minimum size=12mm}
+%		]
+%		% nodes
+%		\node[roundnode] (b0) at (0, 2) {$b^{(t)}$};
+%		\node[roundnode] (b1) at (5, 2) {$b'$};
+%		\node[roundnode] (b2) at (10, 2) {$b^{(t+1)}$};
+%		\node[roundnode] (t0) at (0, 0) {$\theta^{(t)}$};
+%		\node[roundnode] (t1) at (5, 0) {$\theta'$};
+%		\node[roundnode] (t2) at (10, 0) {$\theta^{(t+1)}$};
+%		
+%		% arrows
+%		\draw[->] (b0) to node[above] {$\sim q_b ( b^{(t)}, b' | A )$} (b1);
+%		\draw[->] (b1) to node[above] {$\alpha_b (b, b' | A )$} (b2);
+%		\draw[->] (b0) [out=25, in=155] to node[above] {$1-\alpha_b$} (b2);
+%		
+%		\draw[->] (t0) to node[above] {$\sim q_\theta(\theta^{(t)}, \theta' | X, b^{(t)})$} (t1);
+%		\draw[->] (t1) to node[above] {$\alpha_\theta (\theta, \theta' | X, b^{(t)})$} (t2);
+%		\draw[->] (t0) [out=-25, in=-155] to node[below] {$1-\alpha_\theta$} (t2);
+%		
+%		\draw[dashed, ->] (b0) to (2, 0.8);
+%		\draw[dashed, ->] (b0) to (6.5, 0.8);
+%		
+%	\end{tikzpicture}
+	\includegraphics[width=0.8\textwidth]{img/sampling-sequence}
 	\caption{$\theta$-sample generation.}
@@
 samples for $\theta$
-we use only $b$ but not $A$, as $(\theta \indep A )| b$.
-
-\FloatBarrier
-\subsection{Sampling block memberships}
-
-To generate the required $b$-samples, we adopt the MCMC
-procedure of
-\cite{Peixoto-MCMC},
-which relies on writing the posterior in the following form,
-%
-\begin{equation}
-	p(b | A, X) \propto p(A | b, X) \cdot p(b | X) = \pi_b(b),
-\end{equation}
-%
-where $\pi_b(\cdot)$ denotes the un-normalised target density.
-Since we are using the microcanonical SBM formulation, there is only one 
-value of $\psi$ that is compatible with the given $(A, b)$ pair;
-recall the constraints in~(\ref{eqn:sbm-constraints}).
-We denote this value $\psi^* = \{\psi_k^*, \psi_e^*\}$. Therefore, 
-the summation over all $\psi$ needed to evaluate $p(A | b, X)$ reduces to just the single $\psi^*$ term:
-$p(A | b, X) = \sum_{\psi} \nolimits p(A , \psi | b, X) = p(A, \psi^* | b, X)$.
-In
-this context, the microcanonical entropy of the configuration $b$
-is,
-%
-\begin{equation}
-	S(b) \coloneqq - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big),
-	\label{eqn:dl-form}
-\end{equation}
-%
-which can be thought of as the optimal
-``description length'' of the graph. 
-This expression will later be employed 
-to help evaluate experimental results. 
-The exact form of the proposal $q_b$ is explored thoroughly in
-\cite{Peixoto-MCMC} and not repeated here. We use the \verb*|graph-tool| \cite{peixoto_graph-tool_2014}
-library for Python, which implements this algorithm.
-The only modification is in 
-the prior $p(b)$ that we replace with $p(b|X)=B^{-N}$, 
-which cancels out in the MH accept-reject step as it is independent of $b$.
-
-\subsection{Sampling feature-to-block generator parameters}
-\label{s:sfb}
-
-The target distribution for the required $\theta$-samples 
-is the posterior of $\theta$ given the values of the pair $(X, b)$. 
-We write this as,
-%
-\begin{equation}
-	\pi_\theta(\theta) \propto p(\theta | X, b) \propto p(b | X, \theta) p(\theta) \propto  \exp \left( - U(\theta) \right),
-	\label{eq:U}
-\end{equation}
-%
-where $U(\theta)$ denotes the negative log-posterior. Let $y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and $a_{ij} \coloneqq \phi_j(x_i; \theta)$. 
-Discarding constant terms, $U(\theta)$ can be expressed as,
-%
-\begin{equation}
-	U(\theta) = \left( \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} \|\theta\|^2 = N \cdot \Lcal(\theta) + \frac{1}{2\sigma_\theta^2} \|\theta\|^2;
-	\label{eqn:U-form}
-\end{equation}
-%
-see Appendix \ref{appdx:form-U}.
-The function $U(\theta)$ is a typical objective function for neural network training.
-The first term $N \cdot \Lcal(\theta)$ is introduced by the likelihood and represents the cross-entropy between the graph-predicted and feature-predicted block memberships.
-The second term, introduced by the prior, brings a form of regularisation, guarding against over-fitting. In order to draw samples from the posterior 
-$\pi_\theta \propto \exp(-U)$ we adopt the Metropolis-adjusted Langevin 
-algorithm (MALA) \cite{mala-tweedie}, which uses $\nabla U$ to bias the 
-proposal towards regions of higher density. Given the current 
-sample $\theta$, a proposed 
-new sample $\theta'$ is generated from,
-%
-\begin{equation*}
-	\theta' \sim q_\theta\big(\theta, \theta'\big) 
-	= \Gaussian \big( \theta' ; \theta - h \nabla U(\theta), 2h I \big),
-\end{equation*}
-%
-where $\xi \sim \Gaussian(0, I)$ and $h$ is a step-size parameter 
-which may vary with the sample index.
-Without the injected noise term $\xi$, MALA is equivalent to gradient descent. We require $\xi$ to fully explore the parameter space. 
-The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:form-U}).
-
-\subsection{Sampling sequence}
-\label{s:ss}
-
-So far, each $\theta^{(t)}$ update has used its corresponding $b^{(t)}$ sample.
-This means the evaluation of $U^{(t)}$ and $\nabla U^{(t)}$ has high variance, leading to longer burn-in and possibly slower MCMC convergence.
-The only link between $b^{(t)}$ and $\theta^{(t)}$ is in the evaluation of $U^{(t)}$ and $\nabla U^{(t)}$ which depends only on the matrix $y^{(t)}$ with entries $y_{ij}^{(t)} \coloneqq \one\{b_i^{(t)} = j\}$.
-We would rather deal with the expectation of each $y_{ij}^{(t)}$:
-%
-\begin{equation}
-	\Expect \left[ y_{ij}^{(t)} \right] = \Expect_{b^{(t)}} \left[ \one \left( b_{i}^{(t)} = j \right) \right]
-	= p(b_i = j | A, X).
-\end{equation}
-%
-An unbiased estimate for this can be obtained using 
-the thinned $b$-samples after burn-in.
-Let $\Tcal_b$  denote the retained set of indices 
-for the $b$-samples and $\Tcal_\theta$ similarly for the $\theta$-chain. 
-The unbiased estimate for $y_{ij}^{(t)}$ is then:
-%
-\begin{equation}
-	\hat{y}_{ij} \coloneqq \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} y_{ij}^{(t)} = \frac{1}{|\Tcal_b|} \sum_{t \in \Tcal_b} \one\{b_i^{(t)} = j\}.
-	\label{eqn:y-hat}
-\end{equation}
-%
-The same matrix $\hat{y}$ is used in each $\theta^{(t)}$ update step.
-This way, it is not necessary to run the $b$ and $\theta$ Markov chains 
-concurrently. Instead, we run the $b$-chain to completion and use it 
-to generate $\hat{y}$ also allowing us to vary the lengths of each.
-
-\subsection{Dimensionality reduction}
-\label{sec:dim-reduction}
-
-The complexity of evaluating $U$ and $\nabla U$ is linear in 
-the dimension of the feature space $D$,
-so there is computational incentive to reduce $D$.
-Given the samples $\left\{ \theta^{(t)} \right\}$, we can compute the empirical mean and standard deviation of each component of $\theta$. 
-Switching to the matrix notation $W$ for $\theta$,
-let:
-%
-\begin{equation}
-	\hat{\mu}_{ij} \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} W_{ij}^{(t)} \qquad \textrm{and} \qquad
-	\hat{\sigma}_{ij}^2 \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \left( W_{ij}^{(t)} - \hat{\mu}_{ij} \right)^2.
-\end{equation}
-%
-A simple heuristic to discard the least important features requires specifying a cutoff $c > 0$ and a multiplier $k > 0$. We define the function $\Fcal_i(j)$ 
-as in~(\ref{eqn:fij}) and only keep features with indices $d \in \Dcal'$, where $\Dcal'$ is given in~(\ref{eqn:kept-feature-set}).
-%
-\begin{align}
-	\Fcal_i(j) &\coloneqq (\hat{\mu}_{ij} - k \hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij}) \cap (-c, +c),
-	\label{eqn:fij} \\
-	\Dcal' &\coloneqq \left\{ j \in [D] : \exists i \in [B] \textrm{ s.t. }  \Fcal_i(j) = \emptyset \right\}.
-	\label{eqn:kept-feature-set}
-\end{align}
-%
-Intuitively, this means discarding any feature $j$ for which 
-$(\hat{\mu}_{ij} - k\hat{\sigma}_{ij}, \hat{\mu}_{ij} + k \hat{\sigma}_{ij})$ overlaps with
-$(-c, c)$ for all $i$.
-If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \Gaussian(W_{ij}; \hat{\mu}_{ij}, \hat{\sigma}_{ij}^2)$, then this would be analogous to a hypothesis test on the magnitude of $W_{ij}$ compared to $c$ with multiplier $k$ in~(\ref{eqn:fij}) determining the degree of significance of the result.
-Conversely, if we want to fix the number of dimensions in our reduced feature set $|\Dcal'|=D'$, the problem then becomes finding the largest value of $c$ such that $|\Dcal'|=D'$ given $k=k_0$:
-%
-\begin{equation}
-	c^* = \argmax \{c>0\; : \;|\Dcal'| = D', k=k_0\}.
-	\label{eqn:c-star}
-\end{equation}
+we use only $b$ but not $A$, as $(\theta \perp \!\!\! \perp A )| b$.

### 2022-02-20 iwsm/references.tex
IWSM full scaffold

% References should be placed in the text in author (year) form.
% The list of references should be placed below IN ALPHABETICAL ORDER.
% (Please follow the format of the examples very tightly).

\references
\begin{description}

	\item[Abbe, E.] (2016) Graph compression: The effect of clusters.
	In: {\it 54th Annual Allerton Conference on Communication, Control, and Computing.} pp. 1–-8.

	\item[Airoldi, E.M., Blei, D., Fienberg, S., Xing, E.] (2009).
	Mixed membership stochastic blockmodels. 
	In: {\it Advances in Neural Information Processing Systems} vol. 21.

	\item[Andrieu, C., Roberts, G.O.] (2009). 
	The pseudo-marginal approach for efficient Monte Carlo computations.
	The Annals of Statistics 37(2), 697–-725.

	\item[Gaucher, S., Klopp, O., Robin, G.] (2021).
	Outliers detection in networks with missing links.
	Computational Statistics \& Data Analysis 164, 107308.

	\item[Hastings, W.K.] (1970).
	Monte Carlo sampling methods using Markov chains and their applications.
	Biometrika 57(1), 97-–109.

	\item[Leskovec, J., Mcauley, J.] (2012).
	Learning to discover social circles in ego networks.
	In: {\it Advances in Neural Information Processing Systems} vol. 25

	\item[Nowicki, K., Snijders, T.A.B.] (2001).
	Estimation and prediction for stochastic blockstructures. 
	Journal of the American Statistical Association 96(455), 1077–-1087.

	\item[Pasternak, B., Ivask, I.] (1970).
	Four unpublished letters. Books Abroad 44(2), 196–-200.

	\item[Peixoto, T.P.] (2014).
	Efficient Monte Carlo and greedy heuristic for the inference of stochastic block models.
	Physical Review E 89(1).

	\item[Peixoto, T.P.] (2017).
	Nonparametric Bayesian inference of the microcanonical
	stochastic block model. Physical Review E 95(1).

	\item[Peixoto, T.] (2014).
	The graph-tool Python library. figshare (2014), figshare. com/articles/graph\_tool/1164194

	\item[Roberts, G.O., Tweedie, R.L.] (1996).
	Exponential convergence of Langevin distributions and their discrete approximations.
	Bernoulli 2(4), 341–-363.

	\item[Stehle, J. et al] (2011).
	High resolution measurements of face-to-face contact patterns in a primary school.
	PLoS ONE 6(8), 1–-13.

	\item[Zhu, J., Song, J., Chen, B.] (2016).
	Max-margin nonparametric latent feature models for link prediction.
	arXiv preprint cs.LG:1602.07428.
\end{description}

### 2022-03-01 iwsm/experiments.tex
Edit some more

--- 2021-11-16 publication/experiments.tex
@@
 
-We apply our proposed methods to a variety of labelled networks:
-
-\begin{itemize}
-	\item \textbf{Political books} \cite{polbooks} ($N=105, E=441, D=3$) -- network of Amazon political books published close to the 2004 presidential election.
-Two books are connected if they were frequently co-purchased.
-Vertex features encode the political affiliation of the author (liberal, conservative, or neutral).
-	\item \textbf{Primary school dynamic contacts} \cite{schools} ($N=238, E=5539, D=13$) -- network of 238 individuals (students and teachers),
-with edges denoting face-to-face contacts at a primary school in Lyon, France. 
-The vertex features are class membership (one of 10 values: 1A-5B), gender (male, female), and status (teacher, student). We choose to analyse just the second day of results.
-	\item \textbf{Facebook egonet} \cite{fb-snap} ($N=747, E=30025, D=480$) -- network of Facebook users with edges denoting ``friends''.
-Vertex features are fully anonymised and encode information about each user's education history, languages spoken, gender, home-town, birthday etc. We focus on the egonet with id 1912.
-\end{itemize}
-%
+We apply our proposed methods to a variety of real-world datasets.
 For reference, the inferred partitions for all of these are given on Figure~\ref{fig:graphs-all}.
-We employ the following metrics to assess model performance. 
-First, the average
-description length per entity (nodes and edges) 
-$\bar{S}_e$ 
-used to gauge the SBM fit is defined as:
-%
-\begin{equation}
-	\bar{S}_e \coloneqq \frac{1}{(N+E) |\Tcal_b|} \sum_{t\in \Tcal_b} S \left( b^{(t)} \right).
-	\label{eqn:mean-dl}
-\end{equation}
-%
-Next, to assess the performance of the feature-to-block predictor, 
-the vertex set $[N]$ 
-is partitioned at random so that 
-a constant fraction $f$ of vertices form the training set $\Gcal_0$ and 
-the remainder form the test set $\Gcal_1$.
-The $b$-chain is run using the whole network but only vertices $v \in \Gcal_0$ are 
-used for the $\theta$-chain. 
-Then the  the average cross-entropy loss 
-over each set is used to gauge the quality of the fit,
-%
-\begin{equation}
-	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star^{(t)},
-	\quad \textrm{where} \quad
-	\Lcal_\star^{(t)} \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)},
-	\label{eqn:cross-entropy-loss}
-\end{equation}
-%
-where $\star \in \{0, 1\}$ toggles between the training and test sets
-and $\hat{y}_{ij}$ is defined in~(\ref{eqn:y-hat}).
-%
-Nevertheless, the cross-entropy loss is a coarse measure of fit. 
-A new measure, specific to each detected block,
-can be defined as follows. Let
-$\Bcal_\star(j)$ 
-be the set of vertices with maximum a posteriori probability of belonging 
-to block~$j$,
-$
-	\Bcal_\star(j) \coloneqq \{i \in \Gcal_\star : \hat{b}_i = j\},
-$
-where
-$ 
-	\hat{b}_i \coloneqq \argmax_j \hat{y}_{ij},
-$
-and
-define the {\em block-accuracy} for block $j$ as,
-%
-\begin{equation}
-	\eta_\star(j) \coloneqq \frac{1}{|\Bcal_\star (j)| \cdot 
-	|\Tcal_\theta| } 
-	\sum_{i \in \Bcal_\star (j)}  \sum_{t \in \Tcal_\theta}
-	\one \left\{\hat{b}_i = \argmax_j \phi_j \left( x_i; \theta^{(t)} \right) \right\}.
-	\label{eqn:accuracy}
-\end{equation}
-%
-This effectively tests whether the feature-to-block and 
-graph-to-block predictions agree in their largest component.
-For the higher-dimensional datasets, we also apply the 
-dimensionality reduction method 
-of Section~\ref{sec:dim-reduction}.  
-We then retrain the feature-block predictor using only the retained 
-feature set $\Dcal'$, and report the log-loss over the training and 
-test sets for the reduced classifier -- 
-denoted $\bar{\Lcal}_0'$ and $\bar{\Lcal}_1'$ respectively. 
-\input{experiments-figures}
-\input{experiments-detailed}
+%We employ the following metrics to assess model performance. 
+%First, the average
+%description length per entity (nodes and edges) 
+%$\bar{S}_e$ 
+%used to gauge the SBM fit is defined as:
+%%
+%\begin{equation}
+%	\bar{S}_e \triangleq \frac{1}{(N+E) |\mathcal{T}_b|} \sum_{t\in \mathcal{T}_b} S \left( b^{(t)} \right).
+%	\label{eqn:mean-dl}
+%\end{equation}
+%%
+%Next, to assess the performance of the feature-to-block predictor, 
+%the vertex set $[N]$ 
+%is partitioned at random so that 
+%a constant fraction $f$ of vertices form the training set $\mathcal{G}_0$ and 
+%the remainder form the test set $\mathcal{G}_1$.
+%The $b$-chain is run using the whole network but only vertices $v \in \mathcal{G}_0$ are 
+%used for the $\theta$-chain. 
+%Then the  the average cross-entropy loss 
+%over each set is used to gauge the quality of the fit,
+%%
+%\begin{equation}
+%	\bar{\mathcal{L}}_\star \triangleq \frac{1}{|\mathcal{T}_\theta|} \sum_{t \in \mathcal{T}_\theta} \mathcal{L}_\star^{(t)},
+%	\quad \textrm{where} \quad
+%	\mathcal{L}_\star^{(t)} \triangleq \frac{1}{|\mathcal{G}_\star|} \sum_{i \in \mathcal{G}_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)},
+%	\label{eqn:cross-entropy-loss}
+%\end{equation}
+%%
+%where $\star \in \{0, 1\}$ toggles between the training and test sets
+%and $\hat{y}_{ij}$ is defined in~(\ref{eqn:y-hat}).
+%%
+%Nevertheless, the cross-entropy loss is a coarse measure of fit. 
+%A new measure, specific to each detected block,
+%can be defined as follows. Let
+%$\mathcal{B}_\star(j)$ 
+%be the set of vertices with maximum a posteriori probability of belonging 
+%to block~$j$,
+%$
+%	\mathcal{B}_\star(j) \triangleq \{i \in \mathcal{G}_\star : \hat{b}_i = j\},
+%$
+%where
+%$ 
+%	\hat{b}_i \triangleq \underset{j}{\operatorname{argmax}}\hat{y}_{ij},
+%$
+%and
+%define the {\em block-accuracy} for block $j$ as,
+%%
+%\begin{equation}
+%	\eta_\star(j) \triangleq \frac{1}{|\mathcal{B}_\star (j)| \cdot 
+%	|\mathcal{T}_\theta| } 
+%	\sum_{i \in \mathcal{B}_\star (j)}  \sum_{t \in \mathcal{T}_\theta}
+%	\boldsymbol{1} \left\{\hat{b}_i = \underset{j}{\operatorname{argmax}}\phi_j \left( x_i; \theta^{(t)} \right) \right\}.
+%	\label{eqn:accuracy}
+%\end{equation}
+%%
+%This effectively tests whether the feature-to-block and 
+%graph-to-block predictions agree in their largest component.
+%For the higher-dimensional datasets, we also apply the 
+%dimensionality reduction method 
+%of Section~\ref{sec:dim-reduction}.  
+%We then retrain the feature-block predictor using only the retained 
+%feature set $\mathcal{D}'$, and report the log-loss over the training and 
+%test sets for the reduced classifier -- 
+%denoted $\bar{\mathcal{L}}_0'$ and $\bar{\mathcal{L}}_1'$ respectively.

### 2022-03-01 iwsm/inference.tex
Edit some more

@@
 Although the prior is easily computable, 
-computing the likelihood 
-requires summing over all latent block-states, 
-$p(A| X, \theta) = \sum_{b \in [B]^N} p(A | b) P(b | X, \theta)$, which is 
+computing the likelihood $p(A| X. \theta)$
+requires summing over all $B^N$ latent block-states which is 
 clearly impractical. In fact, this
@@
 Markov chain via the Metropolis-Hastings (MH) 
-algorithm \cite{hastings-alg}.
+algorithm [Hastings (1970)].
 The splitting of the Markov chain into two levels allows us to side-step the summation over
@@
 This is an example of a pseudo-marginal approach;
-see, e.g., Andrieu and Roberts~\cite{pseudo-marginal} 
+see, e.g., Andrieu and Roberts (2009) 
 for a detailed rigorous derivation based on~(\ref{eqn:theta-unbiased}).
@@
 Figure \ref{fig:samp-sequence} shows an overview of the proposed method, with $q$ and $\alpha$ denoting the MH proposal distribution and acceptance probability respectively.
-Note the importance of the simplification in~(\ref{eqn:b-pseudo-prior}). 
-As evaluating $p(b| X)$ does not depend on $X$, 
+Due to our formulation, evaluating $p(b| X)$ does not depend on $X$ so
 we do not need $X$ to sample $b$.

### 2022-03-01 iwsm/introduction.tex
Edit some more

--- 2021-11-16 publication/introduction.tex
@@
 including
-compression \cite{cluster-compression} and link prediction 
-\cite{link-prediction}.
+compression [Abbe (2016)] and link prediction 
+[Zhu et al. (2016)].
 In this work, we examine vertex-labelled networks, 
-referring to the labels as {\em features}.
-A typical goal is to determine whether a given feature impacts graphical structure.
-Answering this requires a random graph model; the standard is the stochastic block model (SBM) \cite{vanilla-sbm}.
-Numerous variants of the SBM  have been proposed, e.g., the MMSBM \cite{mixed-membership-sbm} and OSBM \cite{overlapping-sbm}, but these do not include features in the graph generation process.
+referring to the labels as {\em features}. A typical goal is to determine whether a given feature impacts graphical structure. Answering this requires a random graph model;
+the standard is the stochastic block model (SBM) [Nowicki and Snijders (2001)].
+Numerous variants of the SBM  have been proposed, e.g., the MMSBM [Airoldi et al.
+(2009)] and OSBM [Zhu et al.
+(2016)], but these do not include features in the graph generation process.
 

### 2022-03-01 iwsm/prelim.tex
IWSM full scaffold
Edit some more

--- 2021-11-16 publication/prelim.tex
@@
 This is a latent variable model where each vertex belongs to a single block and the probability two vertices are connected depends only on the block memberships of each.
-Specifically, we will use the microcanonical variant of the SBM, proposed by Peixoto \cite{Peixoto-Bayesian-Microcanonical}.
-To allow for degree-variability between members of the same block, we adopt the following degree-corrected
+Specifically, we will use the microcanonical variant of the SBM, proposed by Peixoto (2016). To allow for degree-variability between members of the same block, we adopt the following degree-corrected 
 formulation (DC-SBM):
@@
 
-\begin{definition}[Microcanonical DC-SBM]
-	\label{defn:microcan-dc-sbm}
+\textbf{Microcanonical DC-SBM} --
 Let $N \geq 1$ denote the number of vertices in an undirected graph
@@
 	e_{rs} = \sum_{i, j \in [N]} A_{ij} 
-	\one \{b_i = r\} \one \{b_j = s\} 
+\boldsymbol{1} \{b_i = r\} \boldsymbol{1} \{b_j = s\} 
 	\qquad 
@@
 \end{equation}
-\end{definition}

### 2022-03-05 iwsm/ffbm.tex
Edit some more
4 Pages

@@
 
-In this section we propose a novel generative model for labelled networks. We call this the feature-first block model (FFBM),
+In this section, we propose a novel generative model for labelled networks. We call this the feature-first block model (FFBM),
 illustrated in Figure~\ref{fig:ffbm}.
-
+%
 Let $N$ denote the number of vertices, $B$ the number of blocks
 and ${\cal X}$ the set of values each feature can take.
-We define the vector $x_i \in \mathcal{X}^D$ as the feature vector for vertex $i$, 
-where $D$ is the number of features associated with each vertex.
-For example, in the datasets we analyse, we deal with binary feature flags
-(denoting the presence/absence of each feature),
-so $\mathcal{X} = \{0, 1\}$. We write $X$ for the $N\times D$ {\em feature matrix} containing
+%
+%We define the vector $x_i \in \mathcal{X}^D$ as the feature vector for vertex $i$, 
+%where $D$ is the number of features associated with each vertex.
+%For example, in the datasets we analyse, we deal with binary feature flags
+%(denoting the presence/absence of each feature),
+%so $\mathcal{X} = \{0, 1\}$.
+%
+We write $X$ for the $N\times D$ {\em feature matrix} containing
 the feature vectors $\{x_i\}_{i=1}^{N}$ 
@@
 block membership $b_i\in[B]$ is generated based on the feature
-vector $x_i$, independently between vertices. The conditional
-distribution of $b_i$ given $x_i$ also depends on a collection
-of weight vectors $\theta=\{w_k\}_{k=1}^B$, where each
-$w_k$ has dimension $D$. We will later find it convenient
-to write $\theta$ as a $B \times D$ matrix of weights $W$. Specifically, 
-the distribution of $b$ given $X$ and $\theta$ is,
+vector $x_i$, independently between vertices, 
+$p(b| X, \theta) = \prod_{i \in [N]} \phi_{b_i} (x_i; \theta)$.
+%The conditional
+%distribution of $b_i$ given $x_i$ also depends on a collection
+%of weight vectors $\theta=\{w_k\}_{k=1}^B$, where each
+%$w_k$ has dimension $D$.
 %
-\begin{equation}
-	p(b| X, \theta) = \prod_{i \in [N]} p(b_i | x_i, \theta) = \prod_{i \in [N]} \phi_{b_i} (x_i; \theta)
-	= \prod_{i \in [N]} \frac{\exp\left(w_{b_i}^T x_i\right)}{\sum_{k \in [B]} \exp \left( w_k^T x_i\right)}.
-\end{equation}
+%We will later find it convenient
+%to write $\theta$ as a $B \times D$ matrix of weights $W$. Specifically, 
+%the distribution of $b$ given $X$ and $\theta$ is,
 %
-Note that $\phi_{b_i}$ has the form of a softmax activation function.
-More complex models based on different choices for the distributions
-$\phi_{b_i}$ above are also possible, but then deriving meaning from 
-the inferred parameter distributions is more difficult. 
+%\begin{equation}
+%	p(b| X, \theta) = \prod_{i \in [N]} \phi_{b_i} (x_i; \theta)
+%	= \prod_{i \in [N]} \frac{\exp\left(w_{b_i}^T x_i\right)}{\sum_{k \in [B]} \exp \left( w_k^T x_i\right)}.
+%\end{equation}
+%
+%Note that $\phi_{b_i}$ has the form of a softmax activation function.
+%More complex models based on different choices for the distributions
+%$\phi_{b_i}$ above are also possible, but then deriving meaning from 
+%the inferred parameter distributions is more difficult. 
 %
@@
 Once the block memberships $b$ have been generated, we then draw the 
-graph $A$ from the microcanonical DC-SBM with additional parameters 
-$\psi = \{\psi_e, \psi_k\}$:
+adjacency matrix of the graph, $A$, from the microcanonical DC-SBM [Peixoto (2017)] with additional parameters 
+$\psi$,
 %
 \begin{equation}
-	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k).
+	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi).
 	\label{eqn:A-generation}

### 2022-03-05 iwsm/inference.tex
4 Pages

@@
 $
-Although the prior is easily computable, 
-computing the likelihood $p(A| X. \theta)$
-requires summing over all $B^N$ latent block-states which is 
-clearly impractical. In fact, this
-approach is doubly intractable as we would also 
-need to compute the normalising constant $p(A|X)$.
+%Although the prior is easily computable, 
+%computing the likelihood $p(A| X. \theta)$
+%requires summing over all $B^N$ latent block-states which is 
+%clearly impractical.
+%In fact, this
+%approach is doubly intractable as we would also 
+%need to compute the normalising constant $p(A|X)$.
 Therefore, following standard Bayesian practice,
@@
 %
-where these approximations become exact as
-the number of MCMC iterations $t\to\infty$.
-As described in the following subsections,
-this can be implemented through a two-level
-Markov chain via the Metropolis-Hastings (MH) 
-algorithm [Hastings (1970)].
-The splitting of the Markov chain into two levels allows us to side-step the summation over
+%where these approximations become exact as
+%the number of MCMC iterations $t\to\infty$.
+Splitting of the Markov chain into two levels side-steps the intractable summation over
 all latent $b \in [B]^N$ required to directly compute the likelihood, $p(A| X, \theta)$.
@@
 unbiased in that the expectation of 
-their distribution converges to the true posterior:
-%
-\begin{equation}
-\lim_{t\to\infty}
-\mathbb{E}_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] = \sum_{b \in [B]^N} p(\theta | X, b) p(b | A, X) = p(\theta | A, X).
-\label{eqn:theta-unbiased}
-\end{equation}
-%
-This is an example of a pseudo-marginal approach;
-see, e.g., Andrieu and Roberts (2009) 
-for a detailed rigorous derivation based on~(\ref{eqn:theta-unbiased}).
+their distribution converges to the true posterior.
+%%
+%\begin{equation}
+%\lim_{t\to\infty}
+%\mathbb{E}_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] = \sum_{b \in [B]^N} p(\theta | X, b) p(b | A, X) = p(\theta | A, X).
+%\label{eqn:theta-unbiased}
+%\end{equation}
+%%
+%This is an example of a pseudo-marginal approach;
+%see, e.g., Andrieu and Roberts (2009) 
+%for a detailed rigorous derivation based on~(\ref{eqn:theta-unbiased}).
 %
@@
 %	\end{tikzpicture}
-	\includegraphics[width=0.8\textwidth]{img/sampling-sequence}
+	\includegraphics[width=0.6\textwidth]{img/sampling-sequence}
 	\caption{$\theta$-sample generation.}
@@
 
-Figure \ref{fig:samp-sequence} shows an overview of the proposed method, with $q$ and $\alpha$ denoting the MH proposal distribution and acceptance probability respectively.
-Due to our formulation, evaluating $p(b| X)$ does not depend on $X$ so
+Figure \ref{fig:samp-sequence} shows an overview of the proposed method, with $q$ and $\alpha$ denoting the Metropolis-Hastings proposal distribution and acceptance probability respectively.
+Due to the FFBM's formulation, evaluating $p(b| X)$ does not depend on $X$ so
 we do not need $X$ to sample $b$.

### 2022-03-05 iwsm/introduction.tex
4 Pages

@@
 Many real-world networks exhibit strong community structure, with most nodes belonging to densely connected clusters. 
-Finding ways to recover the latent communities from the observed graph is an important
-task in many applications,
-including
-compression [Abbe (2016)] and link prediction 
-[Zhu et al. (2016)].
+%Finding ways to recover the latent communities from the observed graph is an important
+%task in many applications,
+%including
+%compression [Abbe (2016)] and link prediction 
+%[Zhu et al. (2016)].
 In this work, we examine vertex-labelled networks, 
@@
 the standard is the stochastic block model (SBM) [Nowicki and Snijders (2001)].
-Numerous variants of the SBM  have been proposed, e.g., the MMSBM [Airoldi et al.
-(2009)] and OSBM [Zhu et al.
-(2016)], but these do not include features in the graph generation process.
+%Numerous variants of the SBM  have been proposed, e.g., the MMSBM [Airoldi et al. (2009)]
+%and OSBM [Zhu et al. (2016)], but these do not include features in the graph generation
+%process.
+
+INSERT REBUTTAL OF CURRENT METHODS
 

### 2022-03-05 iwsm/references.tex
Edit some more
4 Pages

@@
 	\item[Abbe, E.] (2016) Graph compression: The effect of clusters.
-	In: {\it 54th Annual Allerton Conference on Communication, Control, and Computing.} pp. 1–-8.
+	In: {\it 54th Annual Allerton Conference on Communication, Control, and Computing.}.
 
-	\item[Airoldi, E.M., Blei, D., Fienberg, S., Xing, E.] (2009).
-	Mixed membership stochastic blockmodels. 
-	In: {\it Advances in Neural Information Processing Systems} vol. 21.
-
-	\item[Andrieu, C., Roberts, G.O.] (2009). 
-	The pseudo-marginal approach for efficient Monte Carlo computations.
-	The Annals of Statistics 37(2), 697–-725.
+%	\item[Airoldi, E.M., Blei, D., Fienberg, S., Xing, E.] (2009).
+%	Mixed membership stochastic blockmodels. 
+%	In: {\it Advances in Neural Information Processing Systems} vol. 21.
 
@@
 	Outliers detection in networks with missing links.
-	Computational Statistics \& Data Analysis 164, 107308.
-
-	\item[Hastings, W.K.] (1970).
-	Monte Carlo sampling methods using Markov chains and their applications.
-	Biometrika 57(1), 97-–109.
+	Computational Statistics \& Data Analysis 164.
 
@@
 
-	\item[Peixoto, T.P.] (2014).
-	Efficient Monte Carlo and greedy heuristic for the inference of stochastic block models.
-	Physical Review E 89(1).
+%	\item[Peixoto, T.P.] (2013).
+%	The graph-tool Python library. figshare (2014), figshare. com/articles/graph\_tool/1164194
+
+%	\item[Peixoto, T.P.] (2014).
+%	Efficient Monte Carlo and greedy heuristic for the inference of stochastic block models.
+%	Physical Review E 89(1).
 
@@
 
-	\item[Peixoto, T.] (2014).
-	The graph-tool Python library. figshare (2014), figshare. com/articles/graph\_tool/1164194
-
-	\item[Roberts, G.O., Tweedie, R.L.] (1996).
-	Exponential convergence of Langevin distributions and their discrete approximations.
-	Bernoulli 2(4), 341–-363.
+%	\item[Roberts, G.O., Tweedie, R.L.] (1996).
+%	Exponential convergence of Langevin distributions and their discrete approximations.
+%	Bernoulli 2(4), 341–-363.
 
@@
 
-	\item[Zhu, J., Song, J., Chen, B.] (2016).
-	Max-margin nonparametric latent feature models for link prediction.
-	arXiv preprint cs.LG:1602.07428.
+%	\item[Zhu, J., Song, J., Chen, B.] (2016).
+%	Max-margin nonparametric latent feature models for link prediction.
+%	arXiv preprint cs.LG:1602.07428.
 \end{description}

### 2022-03-05 iwsm/inference.tex
Sort out references

@@
 %the number of MCMC iterations $t\to\infty$.
-Splitting of the Markov chain into two levels side-steps the intractable summation over
+Splitting the Markov chain into two levels side-steps the intractable summation over
 all latent $b \in [B]^N$ required to directly compute the likelihood, $p(A| X, \theta)$.

### 2022-03-05 iwsm/references.tex
Sort out references

@@
 
-	\item[Abbe, E.] (2016) Graph compression: The effect of clusters.
-	In: {\it 54th Annual Allerton Conference on Communication, Control, and Computing.}.
+%	\item[Abbe, E.] (2016) Graph compression: The effect of clusters.
+%	In: {\it 54th Annual Allerton Conference on Communication, Control, and Computing.}.
 
@@
 
-	\item[Gaucher, S., Klopp, O., Robin, G.] (2021).
-	Outliers detection in networks with missing links.
-	Computational Statistics \& Data Analysis 164.
+%	\item[Gaucher, S., Klopp, O., Robin, G.] (2021).
+%	Outliers detection in networks with missing links.
+%	Computational Statistics \& Data Analysis 164.
+
+	\item[Krebs, V.] (2004) . The political books network,
+	http://www.orgnet.com/.
 
@@
 	Learning to discover social circles in ego networks.
-	In: {\it Advances in Neural Information Processing Systems} vol. 25
+	{\it Advances in Neural Information Processing Systems} vol. 25
 
@@
 	Journal of the American Statistical Association 96(455), 1077–-1087.
-
-	\item[Pasternak, B., Ivask, I.] (1970).
-	Four unpublished letters. Books Abroad 44(2), 196–-200.
 

### 2022-03-05 iwsm/abstract.tex
4 pages

@@
 relevant posterior distribution of the FFBM parameters. This allows us to infer if and how the observed vertex-features affect macro-structure.
-We apply the proposed methods to a variety of network data
+We apply the proposed methods to several real-world networks
 to extract the most important features along which the vertices
-are partitioned. The main advantages of the proposed approach are
-that the whole feature-space is used automatically and that features
-can be rank-ordered implicitly according to impact.
+are partitioned. The approach stands out from its peers in
+that the whole feature-space is used automatically and features
+can be rank-ordered implicitly by importance.
 }

### 2022-03-05 iwsm/conclusion.tex
4 pages

@@
 
-The proposed Feature-First Block Model (FFBM)
-is a new generative model for labelled networks.
-It is a hierarchical Bayesian model, 
-well-suited for describing how features
-affect network structure.
-The Bayesian inference tools developed in this 
-work facilitate the identification of 
-vertex features that are in some way 
-correlated with the network's graphical structure.
-Consequently, 
-finding the features that best describe the most 
-pronounced partition, makes it possible in practice
-to examine the existence of -- and to make a case for --
-causal relationships.
+%The proposed Feature-First Block Model (FFBM)
+%is a new generative model for labelled networks.
+%It is a hierarchical Bayesian model, 
+%well-suited for describing how features
+%affect network structure.
+%The Bayesian inference tools developed in this 
+%work facilitate the identification of 
+%vertex features that are in some way 
+%correlated with the network's graphical structure.
+%Consequently, 
+%finding the features that best describe the most 
+%pronounced partition, makes it possible in practice
+%to examine the existence of -- and to make a case for --
+%causal relationships.
 

### 2022-03-05 iwsm/experiments-detailed.tex
IWSM full scaffold
4 pages

--- 2021-11-16 publication/experiments-detailed.tex
@@
-Table~\ref{tab:results} summarises the results for each experiment. 
-We see that the dimensionality reduction procedure 
-brings the training and test losses closer. This indicates that 
-the retained
-are indeed well correlated with the underlying graphical 
-partition and that the approach generalises correctly.
-The test loss variance is higher than the training loss variance as the test set is smaller and so more susceptible to variability in its construction.
-%
-The average description length per entity of the graph,
-$\bar{S}_e$,
-has very small variance, suggesting that
-the detected communities can be found reliably (to within an arbitrary 
-relabelling of blocks).
-
 \paragraph{\textbf{Political books.}}

### 2022-03-05 iwsm/experiments.tex
4 pages

@@
 We apply our proposed methods to a variety of real-world datasets.
-For reference, the inferred partitions for all of these are given on Figure~\ref{fig:graphs-all}.
-%We employ the following metrics to assess model performance. 
-%First, the average
-%description length per entity (nodes and edges) 
-%$\bar{S}_e$ 
-%used to gauge the SBM fit is defined as:
+For reference, the inferred partitions $b$ for all of these are given on Figure~\ref{fig:graphs-all}.
+We require metrics to assess model performance. 
+First, the average
+description length per entity (nodes and edges) 
+$\bar{S}_e$ 
+is used to gauge the SBM fit.
 %%
@@
 %%
-%Next, to assess the performance of the feature-to-block predictor, 
-%the vertex set $[N]$ 
-%is partitioned at random so that 
-%a constant fraction $f$ of vertices form the training set $\mathcal{G}_0$ and 
-%the remainder form the test set $\mathcal{G}_1$.
-%The $b$-chain is run using the whole network but only vertices $v \in \mathcal{G}_0$ are 
-%used for the $\theta$-chain. 
-%Then the  the average cross-entropy loss 
-%over each set is used to gauge the quality of the fit,
+Second, to assess the performance of the feature-to-block predictor, 
+the vertex set $[N]$ 
+is partitioned at random into training and test sets, $\mathcal{G}_0$ and $\mathcal{G}_1$. 
+Then the average cross-entropy loss 
+over each set is used to gauge the quality of the fit, $\bar{\mathcal{L}}_\star$,
 %%
@@
 %%
-%where $\star \in \{0, 1\}$ toggles between the training and test sets
+%where $\star \in \{0, 1\}$ toggles between the training and test sets.
+%
 %and $\hat{y}_{ij}$ is defined in~(\ref{eqn:y-hat}).
-%%
+%
 %Nevertheless, the cross-entropy loss is a coarse measure of fit. 
@@
 %graph-to-block predictions agree in their largest component.
-%For the higher-dimensional datasets, we also apply the 
-%dimensionality reduction method 
-%of Section~\ref{sec:dim-reduction}.  
-%We then retrain the feature-block predictor using only the retained 
-%feature set $\mathcal{D}'$, and report the log-loss over the training and 
-%test sets for the reduced classifier -- 
-%denoted $\bar{\mathcal{L}}_0'$ and $\bar{\mathcal{L}}_1'$ respectively.
+%
+For the higher-dimensional datasets, we also develop and apply a novel 
+dimensionality reduction method to select only the top $D'$ features.  
+We then retrain the feature-block predictor using only the retained 
+feature set, and report the cross-entropy loss over the training and 
+test sets for the reduced classifier -- 
+denoted $\bar{\mathcal{L}}_\star'$.
+
+Table~\ref{tab:results} summarises the high-level results for each experiment. 
+We see that the dimensionality reduction procedure 
+brings the training and test losses closer. This indicates that
+the retained features
+are indeed well correlated with the underlying graphical 
+partition and that the approach generalises correctly. 
+%The test loss variance is higher than the training loss variance as the test set is smaller and so more susceptible to variability in its construction.
+%%
+%The average description length per entity of the graph,
+%$\bar{S}_e$,
+%has very small variance, suggesting that
+%the detected communities can be found reliably (to within an arbitrary 
+%relabelling of blocks).

### 2022-03-05 iwsm/ffbm.tex
4 pages

@@
 
-In this section, we propose a novel generative model for labelled networks. We call this the feature-first block model (FFBM),
+We propose a novel generative model for labelled networks -- which we call the feature-first block model (FFBM),
 illustrated in Figure~\ref{fig:ffbm}.
@@
 Once the block memberships $b$ have been generated, we then draw the 
-adjacency matrix of the graph, $A$, from the microcanonical DC-SBM [Peixoto (2017)] with additional parameters 
+adjacency matrix of the graph, $A$, from the microcanonical DC-SBM, Peixoto (2017), with additional parameters 
 $\psi$,
-%
-\begin{equation}
+$
 	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi).
 	\label{eqn:A-generation}
-\end{equation}
+$

### 2022-03-05 iwsm/inference.tex
4 pages

@@
 
-Having completed the definition of the FFBM, we wish to leverage it 
-to perform inference. Specifically, given a labelled network $(A, X)$, we wish to infer if and how the observed features $X$ impact the graphical structure $A$. Formally,
+We wish to leverage the FFBM for inference. Specifically, given a labelled network $(A, X)$, we wish to infer if and how the observed features $X$ impact the graphical structure $A$. Formally,
 this means characterising the posterior distribution:
@@
 instead we aim to draw samples from the posterior,
-%
-\begin{equation}
+$
 	\label{eqn:theta-target}
 	\theta^{(t)} \sim p(\theta | A, X).
-\end{equation}
-%
+$
 We propose an iterative Markov chain Monte Carlo
@@
 
-Figure \ref{fig:samp-sequence} shows an overview of the proposed method, with $q$ and $\alpha$ denoting the Metropolis-Hastings proposal distribution and acceptance probability respectively.
-Due to the FFBM's formulation, evaluating $p(b| X)$ does not depend on $X$ so
+Figure \ref{fig:samp-sequence} shows an overview of the proposed method, with $q$ and $\alpha$ denoting the Metropolis-Hastings proposal distribution and acceptance probability.
+Due to the formulation of the FFBM, evaluating $p(b| X)$ does not depend on $X$ so
 we do not need $X$ to sample $b$.

### 2022-03-05 iwsm/introduction.tex
4 pages

@@
 referring to the labels as {\em features}. A typical goal is to determine whether a given feature impacts graphical structure. Answering this requires a random graph model;
-the standard is the stochastic block model (SBM) [Nowicki and Snijders (2001)].
-%Numerous variants of the SBM  have been proposed, e.g., the MMSBM [Airoldi et al. (2009)]
-%and OSBM [Zhu et al. (2016)], but these do not include features in the graph generation
+the standard is the stochastic block model (SBM) -- see Peixoto (2017).
+%Numerous variants of the SBM  have been proposed, e.g., the MMSBM -- Airoldi et al. (2009) --
+%and OSBM -- Zhu et al. (2016) -- but these do not include features in the graph generation
 %process.
 
-INSERT REBUTTAL OF CURRENT METHODS
-
-To analyse a labelled network using one of the simple SBM variants, a typical procedure would be to partition the graph into blocks grouped by distinct values of the feature of interest.
+Analysing a labelled network with a simple SBM variant, requires partitioning the graph into blocks grouped by distinct values of the feature of interest.
 The associated model can then be used to test for evidence of heterogeneous connectivity between the feature-grouped blocks.
@@
 We would instead prefer to partition the graph into its most natural blocks and then find which of the available features -- if any -- best predict the resulting partition.
-Thus motivated, we present a novel framework for modelling labelled networks, which we call the feature-first block model (FFBM).
-This is an extension of the SBM to labelled networks.
+Thus motivated, we present a novel framework for modelling labelled networks.
+This would not the first extension of the SBM to labelled networks, e.g.
+Stanley et al (2019).
+However, most of the current approaches are focused on leveraging feature information to partition the graph more reliably in the presence of noise.
+We seek instead to develop a model well suited for inferring how vertex features impact graphical structure and to report our confidence in those conclusions.

### 2022-03-05 iwsm/conclusion.tex
Refactor

--- 2022-02-20 iwsm/conclusion.tex
@@
 \label{sec:conclusion}
-
-The proposed Feature-First Block Model (FFBM)
-is a new generative model for labelled networks.
-It is a hierarchical Bayesian model, 
-well-suited for describing how features
-affect network structure.
-The Bayesian inference tools developed in this 
-work facilitate the identification of 
-vertex features that are in some way 
-correlated with the network's graphical structure.
-Consequently, 
-finding the features that best describe the most 
-pronounced partition, makes it possible in practice
-to examine the existence of -- and to make a case for --
-causal relationships.
 

### 2022-03-05 iwsm/ffbm.tex
Refactor
Proofread

@@
 illustrated in Figure~\ref{fig:ffbm}.
-%
 Let $N$ denote the number of vertices, $B$ the number of blocks
-and ${\cal X}$ the set of values each feature can take.
-%
-%We define the vector $x_i \in \mathcal{X}^D$ as the feature vector for vertex $i$, 
-%where $D$ is the number of features associated with each vertex.
-%For example, in the datasets we analyse, we deal with binary feature flags
-%(denoting the presence/absence of each feature),
-%so $\mathcal{X} = \{0, 1\}$.
-%
+and $D$ the number of features associated with each vertex.
 We write $X$ for the $N\times D$ {\em feature matrix} containing
@@
 as its rows.
-
+%
 For the FFBM, we start with the feature matrix $X$ and generate a random
-vector of block memberships $b \in [B]^N$. For each vertex $i$, the
+vector of block memberships $b \in [B]^N$ -- noting the notation $[K] = \{1, 2 \dots K\}$. For each vertex $i$, the
 block membership $b_i\in[B]$ is generated based on the feature
@@
 $p(b| X, \theta) = \prod_{i \in [N]} \phi_{b_i} (x_i; \theta)$.
-%The conditional
-%distribution of $b_i$ given $x_i$ also depends on a collection
-%of weight vectors $\theta=\{w_k\}_{k=1}^B$, where each
-%$w_k$ has dimension $D$.
-%
-%We will later find it convenient
-%to write $\theta$ as a $B \times D$ matrix of weights $W$. Specifically, 
-%the distribution of $b$ given $X$ and $\theta$ is,
-%
-%\begin{equation}
-%	p(b| X, \theta) = \prod_{i \in [N]} \phi_{b_i} (x_i; \theta)
-%	= \prod_{i \in [N]} \frac{\exp\left(w_{b_i}^T x_i\right)}{\sum_{k \in [B]} \exp \left( w_k^T x_i\right)}.
-%\end{equation}
-%
-%Note that $\phi_{b_i}$ has the form of a softmax activation function.
-%More complex models based on different choices for the distributions
-%$\phi_{b_i}$ above are also possible, but then deriving meaning from 
-%the inferred parameter distributions is more difficult. 
 %
@@
 	\centering
+%   IWSM Template does not support TIKZ
 %	\begin{tikzpicture}[
@@
 $
+Priors on the parameters $\theta$ and $\psi$ are chosen to complete the Bayesian framework.

### 2022-03-05 iwsm/inference.tex
Refactor
Proofread

@@
 
-We wish to leverage the FFBM for inference. Specifically, given a labelled network $(A, X)$, we wish to infer if and how the observed features $X$ impact the graphical structure $A$. Formally,
-this means characterising the posterior distribution:
+Given a labelled network $(A, X)$, we wish to infer if and how the observed features $X$ impact the graphical structure $A$. Formally,
+this means characterising the posterior distribution for $\theta$,
 $
@@
 $
-%Although the prior is easily computable, 
-%computing the likelihood $p(A| X. \theta)$
-%requires summing over all $B^N$ latent block-states which is 
-%clearly impractical.
-%In fact, this
-%approach is doubly intractable as we would also 
-%need to compute the normalising constant $p(A|X)$.
 Therefore, following standard Bayesian practice,
-instead we aim to draw samples from the posterior,
+we aim to draw samples from this posterior,
 $
@@
 %
-%where these approximations become exact as
-%the number of MCMC iterations $t\to\infty$.
+where these approximations become exact as
+the number of MCMC iterations $t\to\infty$.
 Splitting the Markov chain into two levels side-steps the intractable summation over
@@
 their distribution converges to the true posterior.
-%%
-%\begin{equation}
-%\lim_{t\to\infty}
-%\mathbb{E}_{b^{(t)}} \left[p \left( \theta | X, b^{(t)} \right) \right] = \sum_{b \in [B]^N} p(\theta | X, b) p(b | A, X) = p(\theta | A, X).
-%\label{eqn:theta-unbiased}
-%\end{equation}
-%%
-%This is an example of a pseudo-marginal approach;
-%see, e.g., Andrieu and Roberts (2009) 
-%for a detailed rigorous derivation based on~(\ref{eqn:theta-unbiased}).
 %
@@
 	\centering
-
+%   IWSM Template does not support tikz
 %	\begin{tikzpicture}[

### 2022-03-20 iwsm/abstract.tex
Proofread
Integrate review comments

@@
 naturally appearing
-in numerous applications in science and engineering.
+in applications in science and engineering.
 A typical inference goal is to determine how the vertex labels
 (or {\em features}) affect the network's structure.
-In this work, we introduce a new generative model, the feature-first block model (FFBM),
+We introduce a new generative model, the feature-first block model (FFBM),
 that facilitates the use of rich queries on labelled networks.
@@
 Carlo approach to efficiently sample from the
-relevant posterior distribution of the FFBM parameters. This allows us to infer if and how the observed vertex-features affect macro-structure.
+posterior distribution of the FFBM parameters. This allows us to infer if and how the observed vertex-features affect macro-structure.
 We apply the proposed methods to several real-world networks
 to extract the most important features along which the vertices
-are partitioned. The approach stands out from its peers in
-that the whole feature-space is used automatically and features
+are partitioned. Importantly, the whole feature-space is used automatically and features
 can be rank-ordered implicitly by importance.
+[The full version of this paper is available on arXiv as [cs.LG] 2105.13762.]
 }

### 2022-03-20 iwsm/experiments.tex
Refactor
Integrate review comments

--- 2022-03-01 iwsm/experiments.tex
@@
 We apply our proposed methods to a variety of real-world datasets.
-For reference, the inferred partitions for all of these are given on Figure~\ref{fig:graphs-all}.
-%We employ the following metrics to assess model performance. 
-%First, the average
-%description length per entity (nodes and edges) 
-%$\bar{S}_e$ 
-%used to gauge the SBM fit is defined as:
-%%
-%\begin{equation}
-%	\bar{S}_e \triangleq \frac{1}{(N+E) |\mathcal{T}_b|} \sum_{t\in \mathcal{T}_b} S \left( b^{(t)} \right).
-%	\label{eqn:mean-dl}
-%\end{equation}
-%%
-%Next, to assess the performance of the feature-to-block predictor, 
-%the vertex set $[N]$ 
-%is partitioned at random so that 
-%a constant fraction $f$ of vertices form the training set $\mathcal{G}_0$ and 
-%the remainder form the test set $\mathcal{G}_1$.
-%The $b$-chain is run using the whole network but only vertices $v \in \mathcal{G}_0$ are 
-%used for the $\theta$-chain. 
-%Then the  the average cross-entropy loss 
-%over each set is used to gauge the quality of the fit,
-%%
-%\begin{equation}
-%	\bar{\mathcal{L}}_\star \triangleq \frac{1}{|\mathcal{T}_\theta|} \sum_{t \in \mathcal{T}_\theta} \mathcal{L}_\star^{(t)},
-%	\quad \textrm{where} \quad
-%	\mathcal{L}_\star^{(t)} \triangleq \frac{1}{|\mathcal{G}_\star|} \sum_{i \in \mathcal{G}_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)},
-%	\label{eqn:cross-entropy-loss}
-%\end{equation}
-%%
-%where $\star \in \{0, 1\}$ toggles between the training and test sets
-%and $\hat{y}_{ij}$ is defined in~(\ref{eqn:y-hat}).
-%%
-%Nevertheless, the cross-entropy loss is a coarse measure of fit. 
-%A new measure, specific to each detected block,
-%can be defined as follows. Let
-%$\mathcal{B}_\star(j)$ 
-%be the set of vertices with maximum a posteriori probability of belonging 
-%to block~$j$,
-%$
-%	\mathcal{B}_\star(j) \triangleq \{i \in \mathcal{G}_\star : \hat{b}_i = j\},
-%$
-%where
-%$ 
-%	\hat{b}_i \triangleq \underset{j}{\operatorname{argmax}}\hat{y}_{ij},
-%$
-%and
-%define the {\em block-accuracy} for block $j$ as,
-%%
-%\begin{equation}
-%	\eta_\star(j) \triangleq \frac{1}{|\mathcal{B}_\star (j)| \cdot 
-%	|\mathcal{T}_\theta| } 
-%	\sum_{i \in \mathcal{B}_\star (j)}  \sum_{t \in \mathcal{T}_\theta}
-%	\boldsymbol{1} \left\{\hat{b}_i = \underset{j}{\operatorname{argmax}}\phi_j \left( x_i; \theta^{(t)} \right) \right\}.
-%	\label{eqn:accuracy}
-%\end{equation}
-%%
-%This effectively tests whether the feature-to-block and 
-%graph-to-block predictions agree in their largest component.
-%For the higher-dimensional datasets, we also apply the 
-%dimensionality reduction method 
-%of Section~\ref{sec:dim-reduction}.  
-%We then retrain the feature-block predictor using only the retained 
-%feature set $\mathcal{D}'$, and report the log-loss over the training and 
-%test sets for the reduced classifier -- 
-%denoted $\bar{\mathcal{L}}_0'$ and $\bar{\mathcal{L}}_1'$ respectively.
+The inferred partitions $b$ for all of these are given on Figure~\ref{fig:graphs-all}.
+To assess model performance,
+%
+the average
+description length per entity (nodes and edges) 
+$\bar{S}_e$ 
+is used to gauge the SBM fit,
+and
+the vertex set $[N]$ is partitioned at random into training and test sets, 
+$\mathcal{G}_0$ and $\mathcal{G}_1$,
+%
+to assess the performance of the feature-to-block predictor.
+The average cross-entropy loss 
+over each set,
+denoted $\bar{\mathcal{L}}_\star$,
+is used to gauge the quality of the fit.
+
+For higher-dimensional datasets, we develop a novel 
+dimensionality reduction method to select only the top $D'$ features.  
+We then retrain the feature-block predictor using only the retained 
+feature set, and report the cross-entropy loss 
+$\bar{\mathcal{L}}_\star'$
+over the training and 
+test sets for the reduced classifier.
+
+Table~\ref{tab:results} summarises the results for each experiment. 
+We see that the dimensionality reduction procedure 
+brings the training and test losses closer, indicating that
+the retained features
+are indeed well correlated with the underlying graphical 
+partition and that the approach generalises correctly.

### 2022-03-20 iwsm/ffbm.tex
Integrate review comments

@@
 
-We propose a novel generative model for labelled networks -- which we call the feature-first block model (FFBM),
+We propose a novel generative model for labelled networks, which we call the feature-first block model (FFBM),
 illustrated in Figure~\ref{fig:ffbm}.
@@
 For the FFBM, we start with the feature matrix $X$ and generate a random
-vector of block memberships $b \in [B]^N$ -- noting the notation $[K] = \{1, 2 \dots K\}$. For each vertex $i$, the
+vector of block memberships $b \in [B]^N$, where we write $[K] = \{1, 2 \dots K\}$. For each vertex~$i$, the
 block membership $b_i\in[B]$ is generated based on the feature
-vector $x_i$, independently between vertices, 
+vector~$x_i$, independently between vertices, so
 $p(b| X, \theta) = \prod_{i \in [N]} \phi_{b_i} (x_i; \theta)$.
@@
 %	\end{tikzpicture}
-	\includegraphics[width=0.8\textwidth]{img/ffbm.png}
+	\includegraphics[width=0.6\textwidth]{img/ffbm.png}
 	\caption{The Feature-First Block Model (FFBM)}
@@
 Once the block memberships $b$ have been generated, we then draw the 
-adjacency matrix of the graph, $A$, from the microcanonical DC-SBM, Peixoto (2017), with additional parameters 
-$\psi$,
-$
-	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi).
-	\label{eqn:A-generation}
-$
-Priors on the parameters $\theta$ and $\psi$ are chosen to complete the Bayesian framework.
+adjacency matrix of the graph 
+	$A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi)$
+from the microcanonical DC-SBM, Peixoto (2017), with additional parameters 
+$\psi$.
+Appropriate priors are placed on the parameters $\theta$ and $\psi$ to 
+complete the Bayesian framework.

### 2022-03-20 iwsm/inference.tex
Integrate review comments

@@
 $
-Therefore, following standard Bayesian practice,
-we aim to draw samples from this posterior,
-$
-	\label{eqn:theta-target}
-	\theta^{(t)} \sim p(\theta | A, X).
-$
-We propose an iterative Markov chain Monte Carlo
-(MCMC) approach to obtain these samples
-$\{\theta^{(t)}\}$. We first draw a sample $b^{(t)}$ 
+Following standard Bayesian practice,
+we propose an iterative Markov chain Monte Carlo
+(MCMC) approach to obtain 
+samples $\theta^{(t)}$ from this posterior.
+First we sample $b^{(t)}$ 
 from the block membership posterior,
 and then use $b^{(t)}$ to obtain a corresponding
-sample $\theta^{(t)}$:
+sample $\theta^{(t)}$,
 %
 \begin{equation}
-	b^{(t)} \stackrel{\rm distr}{\approx} p \big( b | A, X \big) 
+	b^{(t)} \sim p \big( b | A, X \big) 
 	\quad \textrm{then} \quad
-	\theta^{(t)} \stackrel{\rm distr}{\approx} 
-	p\big(\theta | X, b^{(t)} \big),
+	\theta^{(t)}\sim
+	p\big(\theta | X, b^{(t)} \big).
 \end{equation}
 %
-where these approximations become exact as
-the number of MCMC iterations $t\to\infty$.
+% where these approximations become exact as
+% the number of MCMC iterations $t\to\infty$.
 Splitting the Markov chain into two levels side-steps the intractable summation over
@@
 	\includegraphics[width=0.6\textwidth]{img/sampling-sequence}
+	\vspace*{-0.1in}
 	\caption{$\theta$-sample generation.}

### 2022-03-20 iwsm/introduction.tex
Refactor
Proofread
Integrate review comments

@@
 Many real-world networks exhibit strong community structure, with most nodes belonging to densely connected clusters. 
-%Finding ways to recover the latent communities from the observed graph is an important
-%task in many applications,
-%including
-%compression [Abbe (2016)] and link prediction 
-%[Zhu et al. (2016)].
 In this work, we examine vertex-labelled networks, 
 referring to the labels as {\em features}. A typical goal is to determine whether a given feature impacts graphical structure. Answering this requires a random graph model;
-the standard is the stochastic block model (SBM) -- see Peixoto (2017).
-%Numerous variants of the SBM  have been proposed, e.g., the MMSBM -- Airoldi et al. (2009) --
-%and OSBM -- Zhu et al. (2016) -- but these do not include features in the graph generation
-%process.
+the standard is the stochastic block model (SBM), see Peixoto (2017).
 
-Analysing a labelled network with a simple SBM variant, requires partitioning the graph into blocks grouped by distinct values of the feature of interest.
+Analysing a labelled network with one of the standard SBM variants requires partitioning the graph into blocks grouped by distinct values of the feature of interest.
 The associated model can then be used to test for evidence of heterogeneous connectivity between the feature-grouped blocks.
-Nevertheless, this approach can only consider disjoint feature sets and the feature-grouped blocks are often an unnatural partition of the graph.
+But this approach can only consider disjoint feature sets and the feature-grouped blocks often provide an unnatural partition.
 
@@
 Thus motivated, we present a novel framework for modelling labelled networks.
-This would not the first extension of the SBM to labelled networks, e.g.
+This is not the first extension of the SBM to labelled networks, e.g.
 Stanley et al (2019).

### 2022-05-22 abstract_IWSM2022_Tray_final/conclusion.tex
Expand draft

--- 2022-03-05 iwsm/conclusion.tex
@@
 
+The feature-first block model
+(FFBM) is introduced, as a 
+new generative model for labelled 
+networks with communities.
 An efficient MCMC algorithm  
@@
 the vertex features can accurately explain the partition. 
-Through several applications on empirical
+Through applications on empirical
 network data, this approach 
-is shown to be effective at extracting and describing 
+is demonstrated to be effective at extracting and describing 
 the most natural communities in a labelled network. 

### 2022-05-22 abstract_IWSM2022_Tray_final/inference-detailed.tex
Expand draft

--- 2022-02-20 iwsm/inference-detailed.tex
@@
 procedure of
-\cite{Peixoto-MCMC},
+Peixoto (2017),
 which relies on writing the posterior in the following form,
@@
 where $\pi_b(\cdot)$ denotes the un-normalised target density.
-Since we are using the microcanonical SBM formulation, there is only one 
-value of $\psi$ that is compatible with the given $(A, b)$ pair;
-recall the constraints in~(\ref{eqn:sbm-constraints}).
-We denote this value $\psi^* = \{\psi_k^*, \psi_e^*\}$. Therefore, 
-the summation over all $\psi$ needed to evaluate $p(A | b, X)$ reduces to just the single $\psi^*$ term:
+As we are using the microcanonical SBM formulation, there is only one 
+value of $\psi$ - denoted $\psi^*$ - that is compatible with the given $(A, b)$ pair;
+see Peixoto (2017).
+Therefore, the summation needed to evaluate $p(A | b, X)$ reduces to just one term:
 $p(A | b, X) = \sum_{\psi} \nolimits p(A , \psi | b, X) = p(A, \psi^* | b, X)$.
@@
 is,
-%
-\begin{equation}
-	S(b) \triangleq - \log \pi_b(b) = - \Big( \log p(A | b, \psi^*) + \log p(\psi^*, b | X) \Big),
-	\label{eqn:dl-form}
-\end{equation}
-%
+$S(b) \triangleq - \log \pi_b(b),$
 which can be thought of as the optimal
-``description length'' of the graph. 
-This expression will later be employed 
-to help evaluate experimental results. 
-The exact form of the proposal $q_b$ is explored thoroughly in
-\cite{Peixoto-MCMC} and not repeated here. We use the \verb*|graph-tool| \cite{peixoto_graph-tool_2014}
-library for Python, which implements this algorithm.
-The only modification is in 
-the prior $p(b)$ that we replace with $p(b|X)=B^{-N}$, 
-which cancels out in the MH accept-reject step as it is independent of $b$.
+``description length'' of the graph; 
+this will later be employed 
+to evaluate experimental results.
 
@@
 	U(\theta) = \left( \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log \frac{1}{a_{ij}} \right)
-	+ \frac{1}{2\sigma_\theta^2} \|\theta\|^2 = N \cdot \mathcal{L}(\theta) + \frac{1}{2\sigma_\theta^2} \|\theta\|^2;
+	+ \frac{1}{2\sigma_\theta^2} \|\theta\|^2 = N \cdot \mathcal{L}(\theta) + \frac{1}{2\sigma_\theta^2} \|\theta\|^2.
 	\label{eqn:U-form}
@@
 %
-see Appendix \ref{appdx:form-U}.
 The function $U(\theta)$ is a typical objective function for neural network training. 
-The first term $N \cdot \mathcal{L}(\theta)$ is introduced by the likelihood and represents the cross-entropy between the graph-predicted and feature-predicted block memberships.
-The second term, introduced by the prior, brings a form of regularisation, guarding against over-fitting. In order to draw samples from the posterior 
+The first term $N \cdot \mathcal{L}(\theta)$ comes from the likelihood and is the cross-entropy between the graph-predicted and feature-predicted block memberships. 
+The second term, introduced by the prior, brings a form of regularisation, guarding against over-fitting. 
+To draw samples from the posterior 
 $\pi_\theta \propto \exp(-U)$ we adopt the Metropolis-adjusted Langevin
-algorithm (MALA) \cite{mala-tweedie}, which uses $\nabla U$ to bias the 
-proposal towards regions of higher density. Given the current 
-sample $\theta$, a proposed 
-new sample $\theta'$ is generated from,
-%
-\begin{equation*}
-	\theta' \sim q_\theta\big(\theta, \theta'\big) 
-	= \mathcal{N} \big( \theta' ; \theta - h \nabla U(\theta), 2h I \big),
-\end{equation*}
-%
-where $\xi \sim \mathcal{N}(0, I)$ and $h$ is a step-size parameter 
-which may vary with the sample index.
-Without the injected noise term $\xi$, MALA is equivalent to gradient descent. We require $\xi$ to fully explore the parameter space. 
-The term $\nabla U$ has an easy to compute analytic form (derived in Appendix \ref{appdx:form-U}).
+algorithm, MALA - Roberts and Tweedie (1996), which uses $\nabla U$ to bias the 
+proposal towards regions of higher density.
 
@@
 We would rather deal with the expectation of each $y_{ij}^{(t)}$:
-%
-\begin{equation}
+$
 	\mathbb{E} \left[ y_{ij}^{(t)} \right] = \mathbb{E}_{b^{(t)}} \left[ \boldsymbol{1} \left( b_{i}^{(t)} = j \right) \right]
 	= p(b_i = j | A, X).
-\end{equation}
-%
+$
 An unbiased estimate for this can be obtained using 
@@
 
-The complexity of evaluating $U$ and $\nabla U$ is linear in 
+Evaluating $U$ and $\nabla U$ is linear in 
 the dimension of the feature space $D$,
 so there is computational incentive to reduce $D$.
-Given the samples $\left\{ \theta^{(t)} \right\}$, we can compute the empirical mean and standard deviation of each component of $\theta$. 
+Given the samples $\left\{ \theta^{(t)} \right\}$, we compute the empirical mean and standard deviation of each component. 
 Switching to the matrix notation $W$ for $\theta$,
@@
 $(-c, c)$ for all $i$. 
-If we were to use the Laplace approximation for the posterior $p(W_{ij} | A, X) \approx \mathcal{N}(W_{ij}; \hat{\mu}_{ij}, \hat{\sigma}_{ij}^2)$, then this would be analogous to a hypothesis test on the magnitude of $W_{ij}$ compared to $c$ with multiplier $k$ in~(\ref{eqn:fij}) determining the degree of significance of the result.
 Conversely, if we want to fix the number of dimensions in our reduced feature set $|\mathcal{D}'|=D'$, the problem then becomes finding the largest value of $c$ such that $|\mathcal{D}'|=D'$ given $k=k_0$:
